### 2019-11-09 Mathematical Statatistics/ch_02.rmd
ch2 work
ch2 work

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Exploratory Data Analysis}
   \rfoot{\color{headergrey}Chapter 2}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 2

### 2.1

Compute the mean $\bar{x}$ and median, $m$ ,of the six numbers: 3, 5, 8, 15, 20, 21, 24, then apply the natural log to the data.

```{r, c2_1, echo = T}

x <- c(3, 5, 8, 15, 20, 21, 24)
xt <- log(x)

```

Does $\bar{x}$ = $\tilde{x}$ ?

```{r, c2_1a, echo = T}

```

### 2.3

Import the flights data.

```{r, c2_3, echo = T}

flights <- read.csv(paste0(data.dir, "FlightDelays.csv"),
                 header = T)

```

### 2019-11-09 Mathematical Statatistics/ch_02.rmd
ch2 work
ch2 work

@@
 
-```{r, c2_1a, echo = T}
+```{r echo = T}
+
+log(mean(x)) == mean(xt)
 
@@
 
-### 2.3
+Does $m$ = $\tilde{m}$ ?
+
+```{r echo = T}
+
+log(median(x)) == median(xt)
+
+```
+
+### 2.2
+
+Compute the median $\bar{x}$ and median of the eight numbers: 1, 2, 4, 5, 6, 8, 11, 15.
+
+Let $f(x) = \sqrt{x}$
+
+Apply the transformation, then compute the mean, $tilde{x}$ and median, m, of the transformed data.
+
+```{r echo = T}
+
+x <- c(1, 2, 4, 5, 6, 8, 11, 15)
+xt <- sqrt(x)
+
+```
+
++ Is $f(\bar{x}) = \tilde{x}$?
+
+```{r echo = T}
+
+sqrt(mean(x)) == mean(xt)
+
+```
+
++ Is $f(m) = \tilde{m}$?
+
+```{r echo = T}
+
+sqrt(median(x)) == median(xt)
+
+```
+
+### 2.4
 
@@
 
-```{r, c2_3, echo = T}
+```{r echo = T}
 
-flights <- read.csv(paste0(data.dir, "FlightDelays.csv"),
-                 header = T)
+flights <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
+                 header = T))
 
 ```
+
+a.) Create a table and bar chart of the departure times _(DepartTimes)_
+
+```{r echo = T}
+
+table(flights$DepartTime)
+
+```
+
+```{r echo = T, fig.height=3.5}
+
+ggplot(flights, aes(DepartTime)) +
+   geom_bar(aes(fill = ..count..)) +
+   labs(title = "Depart Times")
+
+```
+
+\newpage
+
+b.) Create a contingency table of the variables _Day_ and _Delay30_.
+
+```{r echo = T}
+
+delay <- table(flights$Day, flights$Delayed30)
+
+pretty_kable(delay, "Flight Delays")
+
+```
+
+Show the proportions of delayed flights, by day:
+
+```{r echo = T}
+
+pretty_kable(round(prop.table(delay), 4) * 100, "Flight Delays Proportions")
+
+```
+
+\newpage
+
+c.) Create side-by-side boxplots of the lengths of flight times, grouped by wether or not the flight was delayed at least 30 minutes:
+
+```{r echo = T, fig.height=3.5}
+
+ggplot(flights) +
+   geom_boxplot(aes(Delayed30, FlightLength, fill = Delayed30)) +
+   labs(title = "Flight Length by Delayed30")
+
+```
+
+d.) Do you think there is a relationship between the length of the flight and whether or not the departure time is delayed by at least 30 minutes?
+
+The average flight time is the same, however, the flights that are delayed 30 minutes or more seem to be shorter overall.

### 2019-11-10 Mathematical Statatistics/ch_02.rmd
mathematical statistics, ch2
mathematical statistics, ch2

@@
 The average flight time is the same, however, the flights that are delayed 30 minutes or more seem to be shorter overall.
+
+### 2.5
+
+Import the General Social Survey data.
+
+```{r echo = T}
+
+gss <- data.table(read.csv(paste0(data.dir, "GSS2002.csv"),
+                 header = T))
+
+```
+
+a.) Create a table and a bar chart of the response to the question about the death penalty.
+
+```{r echo = T}
+
+table(gss$DeathPenalty)
+
+```
+
+```{r echo = T, fig.height=3.5}
+
+ggplot(gss, aes(DeathPenalty)) +
+   geom_bar(aes(fill = ..count..)) +
+   labs(title = "Death Penalty Survey Results")
+
+```
+
+b.) Use the _table_ command and the summary command in R on the gun ownership variable. What additional information does the summary command give that the table does not?
+
+```{r echo = T}
+
+table(gss$OwnGun)
+
+```
+
+```{r echo = T}
+
+summary(gss$OwnGun)
+
+```
+
+The summary tells us how many people didn't respond at all to the question.
+
+c.) Create a contingency table displaying the relationship between opinions about the death penalty to that about gun ownership.
+
+```{r echo = T}
+
+with(gss, {
+   table(OwnGun, DeathPenalty)
+})
+
+```
+
+d.) What proportion of gun owners favor the death penalty? Does it appear to be different from the proportion among those who do not own guns?
+
+```{r echo = T}
+
+round(prop.table(with(gss, {
+   table(OwnGun, DeathPenalty)
+})) * 100, 2)
+
+```
+
+It does seem that gun owners are overwhelmingly in favor of the death penalty compared to those who do not own a gun.
+
+### 2.6
+
+Import the data from the recidivism case study in section 1.4.
+
+Import the General Social Survey data.
+
+```{r echo = T}
+
+recid <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
+                 header = T))
+
+```
+
+a.) Create a table and bar chart  of the __Recid__ variable.
+
+```{r echo = T}
+table(recid$Recid)
+```
+
+```{r echo = T}
+ggplot(recid, aes(Recid)) +
+   geom_bar(aes(fill = ..count..)) +
+   labs(title = "Recidivism")
+```
+
+b.) Create a contingency table summarizing the relationship between recidivism (__Recid__) by age (__Age25__).
+
+```{r echo = T}
+with(recid, {
+   table(Recid, Age25)
+})
+```
+
+Of those over 25 years of age, what proportion were sent back to prision?
+
+```{r echo = T}
+round(prop.table(with(recid, {
+   table(Recid, Age25)
+})) * 100, 2)
+```
+
+c.) Create side-by-side boxplots of the number of days to recidivism grouped by type of violation, and give three comparative statements about the distributions.
+
+```{r echo = T, fig.height=3.5}
+ggplot(recid[!is.na(Days)], aes(Type, Days)) +
+   geom_boxplot(aes(fill = Type)) +
+   labs(title = "Days Until Recidivism by Crime Type")
+
+```
+
++ 1.) Technical violations seem to happen quicker after initial release than new crimes.
+
++ 2.) There are more outliers present in the technical violations, with a cluster of them occuring after 900 days.
+
++ 3.) The variance is days until recidivism is larger for new crimes. The chances of a technical violation occuring after 450 days drop dramatically.
+
+```{r echo = T}
+recid[!is.na(Days), .(Variance = comma(var(Days))), by = Type]
+```
+
+d.) Use the quantile command to obtain the quartiles of the number of days to recidivism.
+Since there are missing values (__NA__) for those released offeneders who had not recidivated, you will need to add the argument __na.rm = T__ to the __quartile__ command to exclude those observations.
+
+```{r echo = T}
+quantile(recid$Days, na.rm = T)
+```
+e.) Create ecdf's of days to recidivism for those under 25 years of age and those 25 years of age or older.
+
+```{r echo = T}
+
+with(recid, {
+   recid$U25 <- Age25 == "Under 25"
+
+   plot.ecdf(recid[U25 == T]$Days)
+   plot.ecdf(recid[U25 == F]$Days, col="cornflowerblue", add = T)
+})
+
+```
+
+Approximately what proportion in each age group were sent back to prision 400 days after release?
+
+Of those who were sent back to prision,
+
+```{r echo = T}
+
+recid[!is.na(Days), 
+      .(Days, O400 = Days >= 400), by = Age25][,
+                                               .(Proportion = sum(O400) / .N), by = Age25]
+
+```
+
+went back after being release 400 days or more.

### 2019-11-10 Mathematical Statatistics/ch_02.R

@@ -1,7 +1,4 @@
 
-
-D:\Projects\Statistical-Computing
-
 qnorm(.5)
 
 
@@ -14,11 +11,15 @@ plot(x, dnorm(x))
 
 ?pnorm
 
-f <- function(x, mu = 0, sgma = 1) (1 / sgma*sqrt(2*pi)) ^ exp( (-1/2) * ( (x - mu)/sgma)**2 )
+
+x <- seq(-3, 3, 0.01)
 
 par(mfrow=c(1,1))
-plot(x, f(x))
+plot(x, f(x), lty=3)
 segments(a, 0, a, f(a), col = "blue")
+
+f_norm <- function(x, mu = 0, sgma = 1) (1 / sgma*sqrt(2*pi)) ^ exp( (-1/2) * ( (x - mu)/sgma)**2 )
+curve(f_norm, from = -3, to = 3)
 
 qnorm(.455)
 pnorm(-.11)

### 2019-11-11 Computational Finance/cf_ch_01.rmd
computational finance work
computational finance work

--- 2019-11-09 Mathematical Statatistics/ch_02.rmd
@@
 library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
+library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
+setwd("D:/Projects/Statistical-Computing/RDS")
+
 ```
@@
 
-#### Chapter 2
+#### Chapter 1
 
-### 2.1
+### Ex 1.1
 
-Compute the mean $\bar{x}$ and median, $m$ ,of the six numbers: 3, 5, 8, 15, 20, 21, 24, then apply the natural log to the data.
+We shall use the __getSymbols__ function in the quantmod package to retrieve financial data for General Electric (__GE__).
 
-```{r, c2_1, echo = T}
+```{r echo = T}
 
-x <- c(3, 5, 8, 15, 20, 21, 24)
-xt <- log(x)
+getSymbols("GE", src = "yahoo", from = "2000-01-01", to = "2009-12-30")
+
+names(GE)
+
+GE["2000-01-01/2000-01-20"]
+
+geAdj = GE$GE.Adjusted["2000-01-01-/2000-01-20"]; geAdj
+
+max(geAdj); min(geAdj); mean(geAdj)
 
@@
 
-Does $\bar{x}$ = $\tilde{x}$ ?
-
-```{r, c2_1a, echo = T}
-
+```{r echo = T, fig.height=3.5}
+chartSeries(GE)
 ```
 
-### 2.3
+```{r echo = T, fig.height=3.5}
+chartSeries(GE, TA=NULL, subset='2001-01::2001-02')
+```
 
-Import the flights data.
-
-```{r, c2_3, echo = T}
-
-flights <- read.csv(paste0(data.dir, "FlightDelays.csv"),
-                 header = T)
-
+```{r echo=T}
+saveRDS(GE, file = "GE.rds")
 ```

### 2019-11-11 Mathematical Statatistics/ms_ch_02.rmd
computational finance work
computational finance work

--- 2019-11-10 Mathematical Statatistics/ch_02.rmd
@@
 
-Apply the transformation, then compute the mean, $tilde{x}$ and median, m, of the transformed data.
+Apply the transformation, then compute the mean, $\tilde{x}$ and median, m, of the transformed data.
 
@@
 ```
+
 e.) Create ecdf's of days to recidivism for those under 25 years of age and those 25 years of age or older.
@@
 went back after being release 400 days or more.
+
+### 2.7
+
+Import the data from the black spruce case study in 1.10.
+
+```{r echo=T}
+spruce <- data.table(read.csv(paste0(data.dir, "Spruce.csv"),
+                 header = T))
+```
+
+a.) Compute the numeric summaries for the hight changes of the seedlings.
+
+```{r echo=T}
+summary(spruce$Ht.change)
+```
+
+b.) Create a histogram and normal quantile plot for the height changes of the seedlings.
+
+Is the distribution approximately normal?
+
+```{r echo=T, fig.height=3}
+ggplot(spruce) +
+   geom_histogram(aes(x = Ht.change, y = ..density.., fill = ..count..), bins = 20) +
+   geom_density(aes(x = Ht.change, y = ..density.., col = "darkred"), lwd = 1) +
+   guides(col = "none") +
+   labs(title = "Height-Change Densities")
+```
+```{r echo=T, fig.height=3}
+ggplot(spruce) +
+   geom_qq(aes(sample = Ht.change)) +
+   geom_qq_line(aes(sample = Ht.change)) +
+   labs(title = "Height Change QQ-Plot")
+```
+
+For the height change variable, we see an uneven density plot, qq-plot with outliers in the tails. The data appers to be "approximately" normal.
+
+c.) Create a boxplot to compare the distribution of the change in diameters of the seedlings (__Di.change__), grouped by wheather or not they were in fertilized plots.
+
+```{r echo=T, fig.height = 3}
+ggplot(spruce) +
+   geom_boxplot(aes( y = Di.change, fill = Fertilizer ))
+```
+
+d.) Use the __tapply__ command to find the numeric summaries of the diameter changes for the two levels of fertilization.
+
+```{r echo=T}
+tapply(spruce$Di.change, spruce$Fertilizer, summary)
+```
+
+e.) Create a scatter plot of the hight change against the diameter changes, and describe the relationship.
+
+```{r echo=T, fig.height=3.5}
+ggplot(spruce, aes(Di.change, Ht.change)) +
+   geom_point() +
+   geom_smooth(method = "lm") +
+   labs(title = "Height Change ~ Diameter Change")
+```
+
+\newpage
+
+### 2.8
+
+Import the mobile ads data from section 1.12.
+
+```{r echo=T}
+mobile <- data.table(read.csv(paste0(data.dir, "MobileAds.csv"),
+                 header = T))
+```

### 2019-11-12 Mathematical Statatistics/ms_ch_02.rmd
mathematical statistics work

@@
 ```
+
+a.) Create histograms of the variables __m.cpc_pre__ and __m.cpc_post__, and describe their distributions.
+
+```{r echo=T, fig.height=3.5}
+
+fit <- fitdistr(mobile$m.cpc_pre, "lognormal")
+
+ggplot(mobile, aes(m.cpc_pre)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(y = ..density.., col = "darkred"), lwd = 1) +
+   stat_function(fun = dlnorm, size = 1, color = 'gray',
+                    args = list(mean = fit$estimate[1], sd = fit$estimate[2])) +
+   guides(col = "none") +
+   labs(title = "CPC Pre vs lognormal")
+
+```
+
+```{r echo=T, fig.height=3.5}
+
+pct_zero <- sum(mobile$m.cpc_post == 0) / nrow(mobile) # 7% of values zero
+
+# we will replace the zeros with half the min value here to fit a lognormal.
+
+t_cpc_post <- mobile$d.cpc_post
+rep_val <- min( t_cpc_post[t_cpc_post > 0] ) / 2
+t_cpc_post[t_cpc_post == 0] <- rep_val
+
+fit <- fitdistr(t_cpc_post, "lognormal")
+
+ggplot(mobile, aes(t_cpc_post)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(y = ..density.., col = "darkred"), lwd = 1) +
+   stat_function(fun = dlnorm, size = 1.2, color = "darkgrey",
+                    args = list(mean = fit$estimate[1], sd = fit$estimate[2])) +
+   guides(col = "none") +
+   labs(title = "CPC Post vs lognormal") +
+   theme(legend.position = "none")
+```
+
+The distribution of cpc pre and post seem to follow an approximately lognormal distribution.
+
+b.) Compute the difference between these two variables, create a histogram, and describe the distribution.
+
+```{r echo = T}
+
+```

### 2019-11-12 Mathematical Statatistics/ms_ch_02.rmd
mathematical statistics work
mathematical statistics work

@@
 
-```{r echo = T}
+```{r echo = T, fig.height=3.5}
+
+diff <- data.table(Diff = with(mobile, { m.cpc_pre - m.cpc_post}))[, Index := .I ]
+
+suppressWarnings({
+   fit <- fitdistr(diff$Diff, "t")
+})
+
+ggplot(diff, aes(Diff)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..))+
+   geom_density(aes(y = ..density..), col = "cornflowerblue", lwd = 1.2) +
+   stat_function(fun = dt, size = 1.2, color = "darkred",
+                    args = list(df = fit$estimate[3])) +
+   stat_function(fun = dstd, size = 1.2, color = "darkgreen",
+                    args = list(mean = fit$estimate[1], sd = fit$estimate[2])) +   
+   guides(col = "none") +
+   labs(title = "CPC Pre-Post vs t-dist") +
+   theme(legend.position = "none")
 
 ```
+
+We fit the delta data to a t-distribution, estimating the parameters with __fitdist__.
+The distribution is way too tail heavy for a normal, and the standard t-distribution has tails that are too light, while the normal t-distribution has tails that are too heavy.
+
+We also note that the true tails in the distribution are polynomial, so getting a close fit would be a bit challenging.
+
+d.) Create a normal quantile plot of the difference. Does it appear to be normally distributed?
+
+```{r echo = T, fig.height=3.5}
+
+ggplot(diff, aes(sample = Diff)) +
+   geom_qq() +
+   geom_qq_line() +
+   labs(title = "Normal QQ-Plot of Pre-Post")
+
+```
+
+This data definitely does not fit a normal distribution.
+As we saw with the density plots, there are numerious problems with fitting the tails of the data.
+A simple normal estimate is a terrible fit, while even parametric approaches have problems due to the polynomial tails.
+
+### 2.9
+
+Let $x_1 < x_2 < \ldots < x_n$ and $y_1 < y_2 < \ldots < y_n$ be two sets of data with means $\bar{x}, \bar{y}$ and means $m_x, m_y$, respectively.
+
+Let $w_i = x_i + y_i$ for i = 1, 2, ..., n.
+
+a.) Prove or give a counterexample: $\bar{x} + \bar{y}$ is the mean of $w_1, w_2, ..., w_n$.
+
+b.) Prove or give a counterexample: $m_x + m_y$ is the median of $w_1, w_2, ..., w_n$.

### 2019-11-14 Mathematical Statatistics/ms_ch_02.rmd
mathematical statistics ch2 work
mathematical statistics ch2 work
mathematical statistics, ch2 work
mathematical statistics, ch2 work

@@
 
-Let $w_i = x_i + y_i$ for i = 1, 2, ..., n.
+Let $w_i = x_i + y_i$ for $i = 1, 2, \ldots, n$.
 
@@
 
+\setlength{\leftskip}{2cm}
+
+$\bar{w} = \bar{x} + \bar{y} = \frac{1}{n}\sum_{i=1}^{n}x_n + \frac{1}{n}\sum_{i=1}^{n}y_n$
+
+$\ldots = \frac{1}{n}\sum_{i=1}^{n}[x_n + y_n]$
+
+\setlength{\leftskip}{0pt}
+
 b.) Prove or give a counterexample: $m_x + m_y$ is the median of $w_1, w_2, ..., w_n$.
+
+\setlength{\leftskip}{2cm}
+
+$m_x = \frac{x_{[(\#x+1)\div2]} + x_{[(\#x+1)\div2]}}{2}$
+
+$m_y = \frac{y_{[(\#y+1)\div2]} + y_{[(\#y+1)\div2]}}{2}$
+
+$m_x + m_y = \frac{x_{[(\#x+1)\div2]} + x_{[(\#x+1)\div2]}}{2} + \frac{y_{[(\#y+1)\div2]} + y_{[(\#y+1)\div2]}}{2}$
+
+$\ldots = \frac{1}{2}(x_{[(\#x+1)\div2]} + x_{[(\#x+1)\div2]} + y_{[(\#y+1)\div2]} + y_{[(\#y+1)\div2]})$
+
+\setlength{\leftskip}{0pt}
+
+\newpage
+
+### 2.10
+
+Find the median __m__ and first and third quartiles for the random variable $X$ having:
+
+a.) The exponential distribution with $pdf f(x) = \lambda e^{-\lambda x}$
+
+\setlength{\leftskip}{2cm}
+
+__m__ = $\frac{ln(2)}{\lambda}$ =
+
+$q_p = -\frac{ln(1 - p)}{\lambda}$, where q in .25, .5, .75.
+
+\setlength{\leftskip}{0cm}
+
+b.) The Pareto distribution with parameter $\alpha > 0$ with $pdf f(x) = \frac{\alpha}{X^{\alpha+1}}$
+
+### 2.11
+
+Let the random variable X have a Cauchy distribution with $pdf f(x) = \frac{1}{\pi(1 + x(x - \theta)^2))}$ for $-\infty < x < \infty$.
+
+a.) Show that the mean of X does not exist.
+
+b.) More generally, will $\mathbb{E}[X^k]$ exist? (k = 1, 2, 3, ...).
+
+c.) Show that $\theta$ is the median of the distribution.

### 2019-11-23 template.rmd
sln refactor
sln refactor

--- 2019-11-11 Computational Finance/cf_ch_01.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Exploratory Data Analysis}
-   \rfoot{\color{headergrey}Chapter 2}
+   \fancyhead[C]{\Large\bfseries\color{headergrey} - Title}
+   \rfoot{\color{headergrey} -Section}
    \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
+   \fancyfoot[C]{\rmfamily\color{headergrey} -LF}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 
-#### Chapter 1
+#### Section
 
-### Ex 1.1
-
-We shall use the __getSymbols__ function in the quantmod package to retrieve financial data for General Electric (__GE__).
-
-```{r echo = T}
-
-getSymbols("GE", src = "yahoo", from = "2000-01-01", to = "2009-12-30")
-
-names(GE)
-
-GE["2000-01-01/2000-01-20"]
-
-geAdj = GE$GE.Adjusted["2000-01-01-/2000-01-20"]; geAdj
-
-max(geAdj); min(geAdj); mean(geAdj)
-
-```
-
-```{r echo = T, fig.height=3.5}
-chartSeries(GE)
-```
-
-```{r echo = T, fig.height=3.5}
-chartSeries(GE, TA=NULL, subset='2001-01::2001-02')
-```
-
-```{r echo=T}
-saveRDS(GE, file = "GE.rds")
-```
+### Subsection

### 2019-11-23 Driver/Driver.cpp

@@ -1,20 +1,0 @@
-// Driver.cpp : This file contains the 'main' function. Program execution begins and ends there.
-//
-
-#include <iostream>
-
-int main()
-{
-    std::cout << "Hello World!\n";
-}
-
-// Run program: Ctrl + F5 or Debug > Start Without Debugging menu
-// Debug program: F5 or Debug > Start Debugging menu
-
-// Tips for Getting Started: 
-//   1. Use the Solution Explorer window to add/manage files
-//   2. Use the Team Explorer window to connect to source control
-//   3. Use the Output window to see build output and other messages
-//   4. Use the Error List window to view errors
-//   5. Go to Project > Add New Item to create new code files, or Project > Add Existing Item to add existing code files to the project
-//   6. In the future, to open this project again, go to File > Open > Project and select the .sln file

### 2019-11-23 EntryPoint/main.cpp

@@ -1,22 +1,0 @@
-#include <iostream>
-#include <string>
-
-#include "storageSize.h"
-#include "printBinary.h"
-
-using namespace std;
-
-int main()
-{
-    string value;
-
-    cout << "Enter a number to convert to binary: ";
-    cin >> value;
-
-    printBinary(stoi(value));
-
-    string in;
-    cin >> in;
-
-    return 0;
-}

### 2019-11-23 EntryPoint/printBinary.h

@@ -1,17 +1,0 @@
-#ifndef PRINTBINARY_H_INCLUDED
-#define PRINTBINARY_H_INCLUDED
-
-
-
-#endif // PRINTBINARY_H_INCLUDED
-
-void printBinary(unsigned short int val)
-{
-    for(int i = 15; i >= 0; i--) {
-        if( val & ( 1 << i )) {
-            cout << 1;
-        } else {
-            cout << 0;
-        }
-    }
-}

### 2019-11-23 EntryPoint/storageSize.h

@@ -1,28 +1,0 @@
-#ifndef STORAGESIZE_H_INCLUDED
-#define STORAGESIZE_H_INCLUDED
-
-#include <iostream>
-
-#endif // STORAGESIZE_H_INCLUDED
-
-using namespace std;
-
-void display_storage()
-{
-    cout << "The storage allocated for a char is " << sizeof(char)
-        << " bytes" << endl;
-    cout << "The storage allocated for an unsigned short integer is " << sizeof(unsigned short int)
-        << " bytes" << endl;
-    cout << "The storage allocated for an integer is " << sizeof(int)
-        << " bytes" << endl;
-    cout << "The storage allocated for a long integer is " << sizeof(long int)
-        << " bytes" << endl;
-    cout << "The storage allocated for an unsigned long integer is " << sizeof(unsigned long int)
-        << " bytes" << endl;
-    cout << "The storage allocated for a float is " << sizeof(float)
-        << " bytes" << endl;
-    cout << "The storage allocated for a double is " << sizeof(double)
-        << " bytes" << endl;
-    cout << "The storage allocated for a long double is " << sizeof(long double)
-        << " bytes" << endl;
-}

### 2019-11-23 Mathematical Statatistics/ms_ch_02.rmd
MS ch2, CF ch2
MS ch2, CF ch2
mathematical statistics, ch2
mathematical statistics, ch2

@@
 
-Let the random variable X have a Cauchy distribution with $pdf f(x) = \frac{1}{\pi(1 + x(x - \theta)^2))}$ for $-\infty < x < \infty$.
+Let the random variable X have a Cauchy distribution with pdf $f(x) = \frac{1}{\pi(1 + x(x - \theta)^2))}$ for $-\infty < x < \infty$.
 
@@
 c.) Show that $\theta$ is the median of the distribution.
+
+### 2.12
+
+Find:
+
+a.) The $30^{th}$ and $60^{th}$ percentiles for $N(10, 17^2)$.
+
+```{r, echo = T}
+
+# 30th percentile
+qnorm(.3, mean = 10, sd = 17)
+
+# 60th percentile
+qnorm(.6, mean = 10, sd = 17)
+
+```
+
+b.) The 0.10 and 0.90 quantile for $N(25, 32^2)$.
+
+```{r, echo = T}
+
+# .1 quantile
+qnorm(.1, mean = 25, sd = 32)
+
+# .9 quantile
+qnorm(.9, mean = 25, sd = 32)
+
+```
+
+c.) The point that marks off the upper 25% in $N(25, 32^2)$.
+
+```{r, echo = T}
+
+qnorm(.75, mean = 25, sd = 32, lower.tail = T)
+
+```
+
+### 2.13
+
+The cdf of the exponential distribution is $F(t) = 1 - e^{-\lambda t}$
+
+a.) Find an expression for the 0.05 quantile $q_{0.05}$.
+
+\setlength{\leftskip}{2cm}
+
+$q_t = \frac{ln(1 - t)}{-\lambda}$
+
+$q_{0.05} = \frac{0.513}{\lambda}$
+
+\setlength{\leftskip}{0cm}
+
+b.) Let $\lambda = 4$, and use your answer from (a) to find $q_0.05$, and then check your answer in R using __qexp__.
+
+```{r, echo = T}
+qexp(0.05, 4)
+
+-1/4*log(1 - 0.05)
+
+```
+
+### 2.14
+
+Let $X$ be a random variable with cdf $F(x) = \frac{x^2}{a^2}$ for $0 \le x \le a$.
+
+Find an expression for the $\frac{a}{2}$ and $(1 - \frac{a}{2})$ quantiles, where 0 < a < 1.
+
+$\sqrt{\frac{a}{2}}x$, $1 - \sqrt{\frac{a}{2}}x$
+
+### 2.15
+
+Let X be a random variable with cdf $F(x) = 1 - \frac{9}{x^2}$ for $x \ge 3$.
+
+Find an expression for the $q^{th}$ quantile of $X$.
+
+$\frac{3}{\sqrt{1-x}}$
+
+\newpage
+
+### 2.16
+
+Let $X \sim {\sf Binom}(20, 0.3)$ and let __F__ denotes its cdf.
+
+Does there exist a __q__ such that $F(q) = 0.05$?
+
+### 2.17
+
+In this exercise, we investigate normal quantile plots using R.
+
+a.) Draw a random sample of size $n = 15$ from $N(0, 1)$, and plot both a normal quantile plot and a histogram.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+plot_norm_sample <- function( n ) {
+   dat <- data.table( X = rnorm(n) )[, N := .I]
+
+   p1 <- ggplot(dat, aes(sample = X)) +
+      geom_qq() +
+      geom_qq_line() +
+      labs(title = paste0( "Normal Distribution, N=", n ))
+
+   p2 <- ggplot(dat, aes(X)) +
+      geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+      geom_density(col = "darkorange", fill = "darkorange", alpha = .3)
+
+   grid.arrange(p1, p2, nrow = 2)
+}
+
+```
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+plot_norm_sample(15)
+
+```
+
++ Do the points on the quantile plot appear to fall on a straight line? Is the histogram symmetric, unimodal and bell shaped?
+
+No, the lines do not fit the QQ plot and the histogram is skewed all over.
+
++ Do this several times.
+
+b.) Repeat part (a) for samples of size n = 30, n = 60, and n = 100.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+plot_norm_sample(30)
+
+```
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+plot_norm_sample(60)
+
+```
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+plot_norm_sample(100)
+
+```
+
+c.) What lesson do you draw about using graphs to assess whether or not a data set follows a normal distribution?
+
+The random samples from a normal distirbution do not follow the expectation until the sample size is relatively larger (n>30).
+
+### 2.18
+
+Plot by hand the empirical cumulative distribution function for the set of values:
+
+4, 7, 8, 9, 9, 13, 18, 18, 18, 21
+
+### 2.19
+
+The ecdf for a data set with $n = 20$ values is given in figure 2.18.
+
+a.) How many values are less than or equal to 7?
+
+3
+
+b.) How many times does the value 8 occur?
+
+4
+
+c.) In a histogram of these values, how many values fall in the bin (20, 25]?
+
+6
+
+### 2.20
+
+The data set __ChiMarathonMen__ has a sample of times ffor men between 20 and 39 years of age who completed the Chicago Marathon in 2015.
+
+Graph the ecdf's of the times for men in the 25-29-age division and men in the 35-39-age division.
+
+```{r, echo = T, fig.height=4, fig.width=8}
+marathon <- data.table(read.csv(paste0(data.dir, "ChiMarathonMen.csv"),
+                 header = T))
+
+p1 <- ggplot(marathon[Division == "25-29"], aes(FinishMin)) +
+   stat_ecdf(geom = "point") +
+   geom_vline(xintercept = 160, color = "cornflowerblue") +
+   labs(title = "Men 25-29 Finish Times", y = "")
+
+p2 <- ggplot(marathon[Division == "35-39"], aes(FinishMin)) +
+   stat_ecdf(geom = "point") +
+   geom_vline(xintercept = 160, color = "cornflowerblue") +
+   labs(title = "Men 35-39 Finish Times", y = "")
+
+grid.arrange(p1, p2, nrow = 2)
+
+```
+
++ Approximately what proportion of men in these two divisions finished in 160 min or less?
+
+25-29 is around 70%, and 35-39 is around 40%.

### 2019-11-23 Mathematical Statatistics/ms_ch_03.rmd
Mathematical Statistics, ch3
Mathematical Statistics, ch3

--- 2019-11-11 Computational Finance/cf_ch_01.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Exploratory Data Analysis}
-   \rfoot{\color{headergrey}Chapter 2}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Introduction to Hypothesis - Testing Permutation Tests}
+   \rfoot{\color{headergrey}Chapter 3}
    \lfoot{\color{headergrey}}

### 2019-11-24 Mathematical Statatistics/ms_ch_03.r
mathematical statistics, ch3
mathematical statistics, ch3

@@ -1,0 +1,70 @@
+library(data.table)
+library(ggplot2)
+
+### Beer/Wings Consumption
+
+Beerwings <- data.table(read.csv(paste0(data.dir, "Beerwings.csv"),
+                               header = T))
+
+with(Beerwings, {
+  tapply(Hotwings, Gender, mean)
+})
+
+observed <- 14.5333 - 9.3333 # store observed mean difference
+
+hotwings <- Beerwings$Hotwings
+
+N <- 10^5 - 1
+
+result <- numeric(N)
+
+for( i in 1:N)
+{ #sample of size 15, from 1 to 30, without replacement
+  index <- sample(30, size = 15, replace = F)
+  result[i] <- mean(hotwings[index]) - mean(hotwings[-index])
+}
+
+ggplot(data.table(result), aes(result)) +
+  geom_histogram(aes(fill = ..count..)) +
+  geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.3) +
+  scale_y_continuous(labels = comma) +
+  labs(title = "Beer/Wings Resampling")
+
+p <- (sum(result >= observed) + 1) / (N + 1) # p-value
+v <- p*(1 - p) / N # variance of the samples
+
+### Verizon Repair Times
+
+Verizon <- data.table(read.csv(paste0(data.dir, "Verizon.csv"),
+                                 header = T))
+
+with(Verizon, {
+  tapply(Time, Group, mean)
+})
+
+Time.CLEC <- Verizon[Group == "CLEC"]$Time
+Time.ILEC <- Verizon[Group == "ILEC"]$Time
+Time <- c(Time.CLEC, Time.ILEC)
+
+observed <- mean(Time.ILEC) - mean(Time.CLEC) # store the observed mean difference
+
+N <- 10e4 - 1
+
+result <- numeric(N)
+
+sample.size <- length(Time.ILEC) + length(Time.CLEC)
+
+for( i in 1:N)
+{
+  index <- sample(sample.size, length(Time.ILEC), replace = F)
+  result[i] <- mean(Time[index]) - mean(Time[-index])
+}
+
+ggplot(data.table(result), aes(result)) +
+  geom_histogram(aes(fill = ..count..)) +
+  geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.3) +
+  scale_y_continuous(labels = comma) +
[…]

### 2019-11-24 Mathematical Statatistics/ms_ch_03.r
ms, ch3 work
ms, ch3 work

@@ -66,5 +66,118 @@ ggplot(data.table(result), aes(result)) +
   scale_y_continuous(labels = comma) +
   labs(title = "Verizon Repair Times - Resampling")
 
-p <- ( sum(result <= observed) + 1 ) / ( N + 1 )
-v <- p*(1 - p) / N
+p <- ( sum(result <= observed) + 1 ) / ( N + 1 ) # p-value
+p * 100
+
+v <- p*(1 - p) / N # variance of p
+
+### Verizon, Median
+
+with(Verizon, {
+  tapply(Time, Group, median)
+})
+
+observed <- median(Time.ILEC) - median(Time.CLEC)
+
+N <- 10e4 - 1
+
+result <- numeric(N)
+sample.size <- length(Time.ILEC) + length(Time.CLEC)
+
+for( i in 1:N)
+{
+  index <- sample(sample.size, length(Time.ILEC), replace = F)
+  result[i] <- median(Time[index]) - median(Time[-index])
+}
+
+ggplot(data.table(result), aes(result)) +
+  geom_histogram(aes(fill = ..count..)) +
+  geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.3) +
+  scale_y_continuous(labels = comma) +
+  labs(title = "Verizon Repair Times - Resampling")
+
+p <- ( sum( result <= observed ) + 1) / ( N + 1)
+p * 100
+v <- p*(1 - p)/N
+
+### Verizon, Trimmed Mean
+
+with(Verizon, {
+  tapply(Time, Group, function(x) { mean(x, trim = .25)})
+})
+
+observed <- mean(Time.ILEC, trim = .25) - mean(Time.CLEC, trim = .25)
+
+N <- 10e4 - 1
+
+result <- numeric(N)
+sample.size <- length(Time.ILEC) + length(Time.CLEC)
+
+for( i in 1:N)
+{
+  index <- sample(sample.size, length(Time.ILEC), replace = F)
+  result[i] <- mean(Time[index], trim = .25) - mean(Time[-index], trim = .25)
+}
+
+ggplot(data.table(result), aes(result)) +
+  geom_histogram(aes(fill = ..count..)) +
+  geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.3) +
+  scale_y_continuous(labels = comma) +
+  labs(title = "Verizon Repair Times - Resampling")
+
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_01/Exercises.R
Added Statistics and Data Analysis for Financial Engineering work.
Added Statistics and Data Analysis for Financial Engineering work.

@@ -1,0 +1,55 @@
+
+# Suppose that the daily log returns on a stock are i.i.d. and normally distributed with mean 0.001 and a sd of 0.015.
+# Suppose you buy $1,000 worth of this stock.
+
+# What is the probability that after one trading day your investment is worth less than $990?
+
+days <- 1
+ret.avg <- 0.001
+ret.sd <- 0.015
+
+price.threshold <- log(990 / 1000)
+
+round(pnorm(price.threshold, mean = ret.avg * days, sd = ret.sd * sqrt(days)), 4) * 100
+
+# What is the probability after 5 trading days your investment is worth less than $990?
+
+days <- 5
+
+round(pnorm(price.threshold, mean = ret.avg * days, sd = ret.sd * sqrt(days)), 4) * 100
+
+# The yearly log returns on a stock are normally distributed with mean 0.1 and sd 0.2.
+
+ret.avg <- 0.1
+ret.sd <- 0.2
+
+# The stock is selling at $100 today.
+
+# What is the probability that 1 year from now it is selling at $110 or more?
+
+days <- 252
+price.threshold <- log(110 / 100)
+
+round(pnorm(price.threshold, mean = ret.avg, sd = ret.sd, lower.tail = F), 5) * 100
+
+# The yearly log returns on a stock are normally distributed with mean 0.08 and sd 0.15.
+
+years <- 2
+
+ret.avg <- 0.08
+ret.sd <- 0.15
+
+# What is the probability that 2 years from now it is selling at $90 or more?
+price.threshold <- log(90 / 80)
+
+round(pnorm(price.threshold, mean = ret.avg * years, sd = ret.sd * sqrt(years), lower.tail = F), 5) * 100
+
+# Suppose the prices of a stock at times 1, 2 and 3 are:
+
+prices <- c(95, 103, 98)
+
+# Find R(3):
+
+ret <- (prices[3] - prices[1]) / prices[1]
+log(1 + ret)
+

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_01/Module_01.R

@@ -1,0 +1,113 @@
+######################################################
+# Module 1, Monte Carlo Simulations
+# Moretz, Brandon
+######################################################
+
+# Suppose a hedge fund owns $1,000,000 of stock and used $50,000 of its own capital and $950,000 in borrowed money for the purchase.
+# Suppose that if the value of the stock falls below $950,000 at the end of any trading day, then the hedge fund will sell all the stock and repay the loan.
+# This will wipe out its $50,000 investment.
+# The hedge fund is said to be leveraged 20:1 since its position is 220 times the amount of its own capital invested.
+
+# Suppose that the daily log returns on the stock have a mean of 0.05/year and a standard deviation of 0.23/year. 
+# These can be converted to rates per trading day by dividing by 253 and sqrt(253), respectively.
+
+# What is the probability that the value of the stock will be below $950,000 at the close of at least one of the next 45 trading days?
+
+niter <- 1e5 # number of iterations
+set.seed(2009) # reproducible
+
+seed.capital <- 5e4
+initial.investment <- log(1e6)
+profit.threshold <- log(1.1e6)
+loss.threshold <- log(9.5e5)
+target.profit <- 1e5
+
+simulate_market <- function(days) {
+  # generate random returns for N days
+  r <- rnorm(days, mean = 0.05 / 253,
+             sd = 0.23 / sqrt(253))
+  
+  cumsum(r) # return the final log price after N days.
+}
+
+# setup storage
+outcomes <- list(below = rep(0, niter))
+
+# Simulation: Probability dips below $950,000.
+for (i in 1:niter) {
+  
+  logPrice = initial.investment + simulate_market(45) # simulate 45 trading days.
+  
+  minlogP = min(logPrice) # miniumum price over next 45 days
+  
+  outcomes$below[i] = as.numeric(minlogP < loss.threshold)
+}
+
+print(paste0("Probability the value of the stock is below $950,000 at least one of next 45 sessions: ", round(mean(outcomes$below), 3) * 100, "%"))
+
+# reset seed for next simulation.
+
+# Suppose the hedge fund will sell the stock for a profit of at least $100,000 if the value of the stock rises to at least 
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_01/capped_vs_uncapped.R

@@ -1,0 +1,71 @@
+niter <- 1e5 # number of iterations
+set.seed(2009) # reproducible
+
+seed.capital <- 5e4
+initial.investment <- log(1e6)
+profit.threshold <- log(1.1e6)
+loss.threshold <- log(9.5e5)
+target.profit <- 1e5
+
+simulate_market <- function(days) {
+  # generate random returns for N days
+  r <- rnorm(days, mean = 0.05 / 253,
+             sd = 0.23 / sqrt(253))
+  
+  cumsum(r) # return the final log price after N days.
+}
+
+# reset seed for next simulation.
+
+# Suppose the hedge fund will sell the stock for a profit of at least $100,000 if the value of the stock rises to at least 
+# $1,100,000 at the end of one of the first 100 trading days, sell it for a sloss if the value falls below $950,000 at the end of
+# one of the first 100 trading days, or sell it (for "FMV") after 100 trading days if the closing price has stayed between $950,000 and $1,100,000.
+# reproducible
+set.seed(2009)
+
+outcomes <- list(above = rep(0, niter),
+                 below = rep(0, niter),
+                 middle = rep(0, niter),
+                 uncapped_pnl = rep(0, niter),
+                 capped_pnl = rep(0, niter),
+                 ret = rep(0, niter))
+
+for (i in 1:niter) {
+  
+  # simulate 100 trading days.
+  logPrice = initial.investment + simulate_market(100)
+  
+  suppressWarnings({
+    # ignore Inf returned if condition not meet.
+    profit.day <- min(which(logPrice >= profit.threshold))
+    loss.day <- min(which(logPrice <= loss.threshold))
+  })
+  
+  is.market <- profit.day == Inf && loss.day == Inf
+  
+  # What was the exit condition of the position, hince the final price of the stock?
+  days.open <- ifelse(is.market, length(logPrice),
+                      min(profit.day, loss.day))
+  
+  outcomes$above[i] <- min(profit.day) < min(loss.day)
+  outcomes$middle[i] <- is.market
+  outcomes$below[i] <- min(loss.day) < min(profit.day)
+  
+  # p&l = ending value - initial investment
+  pnl <- exp(logPrice[days.open]) - exp(initial.investment)
+  
+  # market pnl = use FMV, otherwise cap p/l
+  outcomes$uncapped_pnl[i] <- pnl
+  
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_01/labs.R

@@ -1,0 +1,205 @@
+library(data.table)
+library(dplyr)
+library(ggplot2)
+library(ggrepel)
+library(GGally)
+library(ggthemes)
+library(scales)
+library(reshape2)
+library(skimr)
+library(gridExtra)
+
+#####################################################################
+######################### Chapter 2 Lab #############################
+#####################################################################
+
+theme_set(theme_sjplot())
+
+path.data <- "D:/Projects/MSDS-RiskAnalytics/datasets"
+setwd(path.data)
+
+# 2.4.1
+# Data Analytics
+
+dat <- read.csv("Stock_bond.csv", header = T)
+
+names(dat)
+attach(dat)
+
+par(mfrow = c(1, 2))
+
+plot(GM_AC, type = "l")
+plot(F_AC, type = "l")
+
+n <- dim(dat)[1]
+GMReturn <- GM_AC[-1] / GM_AC[-n] - 1
+FReturn <- F_AC[-1] / F_AC[-n] - 1
+
+par(mfrow = c(1, 1))
+plot(GMReturn, FReturn)
+
+# Problems
+
+# 1.)
+# Do GM and Ford returns seem positively correlated? Do you notice any outliying returns? 
+
+cor(GMReturn, FReturn) # .61
+
+# If "yes," do outlying GM returns seem to occur with outliying Ford returns?
+
+logGMReturn <- diff(log(GM_AC))
+
+plot(GMReturn, logGMReturn)
+cor(GMReturn, logGMReturn)
+
+# MSFT / MRK
+
+MSReturn <- MSFT_AC[-1] / MSFT_AC[-n] - 1
+logMSReturn <- diff(log(MSFT_AC))
+
+MRKReturn <- MRK_AC[-1] / MRK_AC[-n] - 1
+logMRKReturn <- diff(log(MRK_AC))
+
+plot(MSReturn, MRKReturn)
+
+cor(MSReturn, MRKReturn)
+
+plot(MSReturn, logMSReturn)
+
+# 2.4.2
+# Simulations
+
+niter <- 1e5 # number of iterations
+set.seed(2009) # reproducible
+
+seed.capital <- 5e4
+initial.investment <- log(1e6)
+profit.threshold <- log(1.1e6)
+loss.threshold <- log(9.5e5)
+
+target.profit <- 1e5
+
+ret.avg <- 0.05
+ret.sd <- 0.23
+market.open <- 253
+
+simulate_market <- function(days) {
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_03/Discussion_03.R

@@ -1,0 +1,48 @@
+library(data.table)
+library(ggplot)
+library(ggtheme)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+library(sn)
+
+theme_set(theme_light())
+
+setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "right", legend.title = element_blank())
+
+x <- seq( from = -4, to = 4, by = 0.001)
+y_norm <- dnorm(x)
+y_df1 <- dt(x, df = 1)
+y_df2 <- dt(x, df = 2)
+y_df5 <- dt(x, df = 5)
+y_df10 <- dt(x, df = 10)
+y_df20 <- dt(x, df = 20)
+
+
+p1 <- ggplot(data = data.table(x = x, y = y_norm)) +
+  geom_line(aes(x, y_df1, color = "t-df = 1"), lwd = 1) +
+  geom_line(aes(x, y_df2, color = "t-df = 2"), lwd = 1) +
+  geom_line(aes(x, y_df5, color = "t-df = 5"), lwd = 1) +
+  geom_line(aes(x, y_df10, color = "t-df = 10"), lwd = 1) +
+  geom_line(aes(x, y_df20, color = "t-df = 20"), lwd = 1) +
+  geom_line(aes(x, y_norm, color = "Normal"), lwd = 1) +
+  labs(title = "Normal vs t", x = "", y = "")
+
+p2 <- ggplot(data = data.table(x = x, y = y_norm)) +
+  geom_line(aes(x, y_df1, color = "t-df = 1"), lwd = 1) +
+  geom_line(aes(x, y_df2, color = "t-df = 2"), lwd = 1) +
+  geom_line(aes(x, y_df5, color = "t-df = 5"), lwd = 1) +
+  geom_line(aes(x, y_df10, color = "t-df = 10"), lwd = 1) +
+  geom_line(aes(x, y_df20, color = "t-df = 20"), lwd = 1) +
+  geom_line(aes(x, y_norm, color = "Normal"), lwd = 1) +
+  labs(x = "", y = "", title = "Zoomed Left Tail") +
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_03/earnings_lab.R

@@ -1,0 +1,85 @@
+library(data.table)
+library(ggplot)
+library(ggtheme)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+library(sn)
+
+theme_set(theme_light())
+
+setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+#### Earnings Lab
+
+library("Ecdat")
+
+?CPSch3
+
+data(CPSch3)
+dimnames(CPSch3)[[2]]
+
+male.earnings <- CPSch3[CPSch3[, 3] == "male", 2]
+sqrt.male.earnings <- sqrt(male.earnings)
+log.male.earnings <- log(male.earnings)
+one.fourth.male.earnings = male.earnings^(1/4)
+
+par(mfrow = c(2, 2))
+qqnorm(male.earnings, datax = T, main = "untransformed")
+qqnorm(sqrt.male.earnings, datax = T,
+       main = "square-root transformed")
+qqnorm(log.male.earnings, datax = T,
+       main = "log-transformed")
+qqnorm(one.fourth.male.earnings, datax = T,
+       main = "1/4 power")
+
+par(mfrow = c(2, 2))
+boxplot(male.earnings, datax = T, main = "untransformed")
+boxplot(sqrt.male.earnings, datax = T,
+        main = "square-root transformed")
+boxplot(log.male.earnings, datax = T,
+        main = "log-transformed")
+boxplot(one.fourth.male.earnings, datax = T,
+        main = "1/4 power")
+
+par(mfrow = c(2, 2))
+plot(density(male.earnings), main = "untransformed")
+plot(density(sqrt.male.earnings),
+     main = "square-root transformed")
+plot(density(log.male.earnings),
+     main = "log-transformed")
+plot(density(one.fourth.male.earnings),
+     main = "1/4 power")
+
+
+library(MASS)
+par(mfrow = c(1, 1))
+boxcox(male.earnings ~ 1)
+boxcox(male.earnings ~ 1, lambda = seq(0.3, 0.45, 1 / 100))
+
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_03/equityReturns_lab.R

@@ -1,0 +1,53 @@
+library(data.table)
+library(ggplot)
+library(ggtheme)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+library(sn)
+
+theme_set(theme_light())
+
+setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+#### Equity Returns Lab
+
+berndtInvest <- read.csv("berndtInvest.csv")
+Berndt <- as.matrix(berndtInvest[, 2:5])
+
+cov(Berndt)
+cor(Berndt)
+
+pairs(Berndt)
+
+library(MASS)
+library(mnormt)
+
+df <- seq(2.5, 8, 0.1)
+n <- length(df)
+
+loglik_profile <- rep(0, n)
+for(i in 1:n)
+{
+  fit <- cov.trob(Berndt, nu = df[i])
+  mu <-  as.vector(fit$center)
+  sigma <- matrix(fit$cov, nrow = 4)
+  loglik_profile[i] <- sum(log(dmt(Berndt, mean = fit$center,
+                                   S = fit$cov, df = df[i])))
+}
+
+aic_t <- -max(2 * loglik_profile) + 2 * (4 + 10 + 1)
+z1 <- ( 2 * loglik_profile > 2 * max(loglik_profile) - qchisq(0.95, 1) )
+plot(df, 2 * loglik_profile, type = "l", cex.axis = 1.5,
+     cex.lab = 1.5, ylab = "2 * loglikelihood - 64,000", lwd = 2)
+abline(h = 2 * max(loglik_profile) - qchisq(0.95, 1) - 64000)
+abline(h = 2 * max(loglik_profile) - 64000)
+abline(v = (df[16] + df[17]) / 2)
+abline(v = (df[130] + df[131]) / 2)

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_03/mle_lab.R

@@ -1,0 +1,200 @@
+library(data.table)
+library(ggplot)
+library(ggtheme)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+library(sn)
+
+theme_set(theme_light())
+
+setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+#### Maximum Likelihood Lab
+
+data(Capm, package = "Ecdat")
+x <- diff(Capm$rf)
+fitdistr(x, "t")
+
+# classical t-distribution
+
+n <- length(x)
+start <- c(mean(x), sd(x), 5)
+loglik_t <- function(beta) sum( - dt((x - beta[1]) / beta[2],
+                                     beta[3], log = T) + log(beta[2]))
+fit_t <- optim(start, loglik_t, hessian = T,
+               method = "L-BFGS-B", lower = c(-1, 0.001, 1))
+AIC_t <- 2 * fit_t$value + 2 * 3
+BIC_t <- 2 * fit_t$value + log(n) * 3
+sd_t <- sqrt(diag(solve(fit_t$hessian)))
+
+fit_t$par
+sd_t
+AIC_t
+BIC_t
+
+# standardized t-distribution
+
+loglik_std <- function(beta) sum( - dstd(x, mean = beta[1],
+                                         sd = beta[2], nu = beta[3], log = T))
+fit_std <- optim(start, loglik_std, hessian = T,
+                 method = "L-BFGS-B", lower = c(-0.1, 0.01, 2.1))
+AIC_std <- 2 * fit_std$value + 2 * 3
+BIC_std <- 2 * fit_std$value + log(n) * 3
+sd_std <- sqrt(diag(solve(fit_std$hessian)))
+
+fit_std$par
+sd_std
+AIC_std
+BIC_std
+
+# F-S skewed t-distribution
+
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_04/lab_bootstrap.R

@@ -1,0 +1,45 @@
+library(bootstrap)
+library(MASS)
+
+set.seed("3857")
+
+data(CRSPday, package = "Ecdat")
+
+ge <- CRSPday[, 4]
+
+getRetVsT <- function(returns, df_canidates = c(1, 2, 4, 6, 10, 20)) {
+  plots <- lapply(df_canidates, function(df) {
+    
+    n <- length(returns)
+    q_range <- (1:n) / (n+1)
+    
+    data <- data.table(ret = returns, theoretical = qt(q_range, df))
+    data$theoretical <- sort(data$theoretical)
+    
+    model <- lm(qt(c(0.25,0.75), df = df) ~ quantile(data$ret,c(0.25,0.75)))
+    
+    ggplot(data, aes(x = sort(ret), y = theoretical)) +
+      geom_abline(col = 'cornflowerblue', lwd = 1.3, slope = model$coefficients[2], intercept = model$coefficients[1]) +
+      geom_point() +
+      labs(title = paste("df = ", df), 
+           x = "returns",
+           y = "theoretical")
+  })
+  
+  do.call(grid.arrange, c(plots, top = "QQ-Plot: returns vs t-distribution"))
+}
+
+getRetVsT(ge)
+
+nboot <- 1000
+
+t_mle <- function(x) { as.vector(
+  fitdistr(x, "t", start = list( m = mean(x), s = sd(x), df = 3) )$estimate
+)}
+
+results <- bootstrap(ge, nboot, t_mle)
+
+rowMeans(results$thetastar[ , ])
+apply(results$thetastar[,], 1, sd)
+
+fitdistr(ge, "t")

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_04/lab_flows.R

@@ -1,0 +1,130 @@
+library(data.table)
+library(ggplot)
+library(ggtheme)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+library(sn)
+library(copula)
+
+theme_set(theme_light())
+
+setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+dat <- read.csv("FlowData.csv")
+dat <- dat / 1000
+
+n <- nrow(dat)
+
+x1 <- dat$Flow1
+fit1 <- st.mple(matrix(1, n, 1), y = x1, dp = c(mean(x1), sd(x1), 0, 10))
+est1 <- fit1$dp
+u1 <- pst(x1, dp = est1)
+
+x2 <- dat$Flow2
+fit2 <- st.mple(matrix(1, n, 1), y = x2, dp = c(mean(x2), sd(x2), 0, 10))
+est2 <- fit2$dp
+
+u2 <- pst(x2, dp = est2)
+
+U.hat <- cbind(u1, u2)
+
+z1 <- qnorm(u1)
+z2 <- qnorm(u2)
+
+Z.hat <- cbind(z1, z2)
+
+library(ks)
+
+fhatU <- kde(x = U.hat, H = Hscv(x=U.hat))
+par(mfrow = c(2, 2), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
+hist(u1, main = "(a)", xlab = expression(hat(U)[1]), freq = F)
+hist(u2, main = "(b)", xlab = expression(hat(U)[2]), freq = F)
+plot(u1, u2, main = "(c)", xlab = expression(hat(U)[1]),
+     ylab = expression(hat(U)[2]), mgp = c(2.5, 1, 0))
+plot(fhatU, drawpoints = F, drawlabels = F,
+     cont = seq(10, 80, 10), main = "(d)", xlab = expression(hat(U)[1]),
+     ylab = expression(hat(U)[2]), mgp = c(2.5, 1, 0))
+
+
+cor.test(u1, u2, method = "spearman")
+cor.test(u1, u2, method = "kendall")
+sin(-0.242*pi/2)
+cor.test(u1, u2, method = "pearson")
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_05/time_series_lab.R

@@ -1,0 +1,364 @@
+library(Ecdat)
+library(tseries)
+library(forecast)
+
+setwd("D:/Projects/MSDS-RiskAnalytics/Module_05")
+
+data(Mishkin, package = "Ecdat")
+
+y <- as.vector(Mishkin[, 1])
+
+par(mfrow = c(1,2))
+acf(y)
+acf(diff(y))
+
+Box.test(diff(y), lag = 10, type = "Ljung-Box")
+
+data(bmw, package = "evir")
+
+Box.test(bmw, lag = 5, type = "Ljung-Box")
+
+fitAR1 <- arima(bmw, order = c(1, 0, 0))
+summary(fitAR1)
+
+acf(bmw)
+
+Box.test(residuals(fitAR1), lag = 5, type = "Ljung-Box", fitdf = 1)
+
+fit <- arima(y, order = c(1, 0, 0))
+Box.test(fit$residuals, type = "Ljung", lag = 24, fitdf = 1)
+
+library(forecast)
+
+auto.arima(diff(y), max.p = 20, max.q = 0, d = 0, ic = "aic")
+
+auto.arima(diff(y), max.p = 20, max.q = 0, d = 0, ic = "bic")
+
+par(mfrow = c(1,1))
+plot.ts(y)
+
+fitMA3 <- arima(diff(y), order = c(0, 0, 3))
+fitMA3
+
+set.seed(4631)
+y1 <- arima.sim(n = 500, list(ar = c(0.4)))
+y2 <- cumsum(y1)
+y3 <- cumsum(y2)
+
+par(mfrow = c(3,1))
+plot.ts(y1)
+plot.ts(y2)
+plot.ts(y3)
+
+adf.test(y)
+pp.test(y)
+
+kpss.test(y)
+
+?adf.test
+
+x <- rnorm(1000)  # no unit-root
+adf.test(x)
+
+y <- diffinv(x)   # contains a unit-root
+adf.test(y)
+
+TT <- 100
+wn <- rnorm(TT) # white noise
+tseries::adf.test(wn)
+
+tseries::adf.test(wn, k=0)
+
+intercept <- 1
+wnt <- wn + 1:TT + intercept
+tseries::adf.test(wnt)
+
+rw <- cumsum(rnorm(TT))
+tseries::adf.test(rw)
+
+tseries::adf.test(rw, k=0)
+
+pdf("inflation_AR1_acf.pdf",width=9,height=5)
+#
+par(mfrow=c(1,2))
+acf(y,main="Inflation rate")
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_06/GARCH_lab.R

@@ -1,0 +1,725 @@
+library(xts)
+library(ggplot2)
+library(ggthemes)
+
+theme_set(theme_light())
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
+
+setwd("D:/Projects/MSDS-RiskAnalytics/Module_06")
+
+data(Mishkin,package="Ecdat")
+infl= as.vector(Mishkin[,1])  # pai1 = one-month inflation rate (in percent, annual rate) 
+
+data(SP500,package="Ecdat")
+SPreturn = SP500$r500
+
+library(evir)  # for emplot
+data(Garch,package="Ecdat")
+attach(Garch)
+
+data(Capm,package="Ecdat")
+difflogrf=diff(log(Capm$rf))
+
+diffdm = diff(dm)  #  Deutsch mark
+diffbp = diff(bp)  #  British pound
+diffcd = diff(cd)  #  Canadian dollar
+diffdy = diff(dy)  #  Japanese yen
+
+n1  = length(SPreturn)
+year_SP = 1981 + (1:n1)* (1991.25-1981)/n1
+d1 = seq(as.Date("1981-01-01"), as.Date("1991-04-01"), by=1)
+date_SP = d1[seq(1,length(d1),by=length(d1)/n1)]
+
+n2 = length(diffdm)
+year_dm = 1980 + (1:n2)* (1987.5-1980)/n2
+d2 = seq(as.Date("1980-01-01"), as.Date("1987-07-01"), by=1)
+date_dm = d2[seq(1,length(d2),by=length(d2)/n2)]
+
+n3 = length(difflogrf)
+year_rf = 1960 + (1:n3) * (2003 - 1960)/n3
+d3 = seq(as.Date("1960-01-01"), as.Date("2003-01-01"), by=1)
+date_rf = d3[seq(1,length(d3),by=length(d3)/n3)]
+
+n4 = length(infl)
+year_infl = 1950+1/12 + (1:n4) * (1991-1950-1/12)/n4
+d4 = seq(as.Date("1950-02-01"), as.Date("1991-01-01"), by=1)
+date_infl = d4[seq(1,length(d4),by=length(d4)/n4)]
+
+
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_08/portfolio_selection_lab.R

@@ -1,0 +1,559 @@
+library(data.table)
+library(xts)
+library(ggplot2)
+library(ggthemes)
+
+theme_set(theme_light())
+
+theme_set(theme_light())
+
+# Theme Overrides
+theme_update(axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
+             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
+             legend.position = "right", legend.title.align=0.5,
+             panel.border = element_rect(linetype = "solid", 
+                                         colour = "lightgray"), 
+             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))
+
+data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
+
+setwd("D:/Projects/MSDS-RiskAnalytics/Module_08")
+
+################################################################
+########## Code for figure 16.1   ##############################
+################################################################
+
+mu1 = 0.14
+mu2 = 0.08
+sig1 = 0.2
+sig2 = 0.15
+rho = 0
+rf = 0.06
+w = seq(0, 1, len = 500)
+means = 0.08 + 0.06 * w
+var = sig1^2 * w^2 + sig2^2 * (1 - w)^2
+risk = sqrt(var)
+ind = !(risk > min(risk))
+ind2 = (means > means[ind])
+wt = 0.693
+meant = 0.08 + 0.06 * wt
+riskt = sqrt(sig1^2 * wt^2 + sig2^2 * (1 - wt)^2)
+
+wp = 0.475
+meanp = 0.08 + 0.06 * wp
+riskp = sqrt(sig1^2 * wp^2 + sig2^2 * (1 - wp)^2)
+
+
+pdf("portfolioNew.pdf", width = 6, height = 5)
+
[…]

### 2019-11-25 Statistics and Data Analysis for Financial Engineering/Module_08/util_fns.R

@@ -1,0 +1,205 @@
+#
+# Written by:
+# -- 
+# John L. Weatherwax                2009-04-21
+# 
+# email: wax@alum.mit.edu
+# 
+# Please send comments and especially bug reports to the
+# above email address.
+#
+# Computes the Efficient Frontier.
+#
+#-----
+
+library(quadprog)
+library(linprog)
+
+efficient_frontier = function(returns, muP, mufree=0.0, w_lower_limit=-Inf, w_upper_limit=+Inf){
+  #
+  # w_lower_limit and w_upper_limit (if finite) are bounds on the porfolio weights such that
+  # 
+  # w_lower_limit <= w <= w_upper_limit
+  #
+  # Note: if solve.QP gives errors about convergence one can try a more restricted range for the values of muP:
+  #
+  # muP = seq(min(mean_vect), max(mean_vect), length.out=500)
+  #
+  #--
+  n_stocks = dim(returns)[2]
+  
+  # Extract individual equity (mean return, std return) values: 
+  mean_vect = apply(returns, 2, mean)
+  cov_mat = cov(returns)
+  sd_vect = sqrt(diag(cov_mat))
+  
+  # Portfolio weight constraints:
+  #
+  #     first condition enforce \sum w_j = 1
+  #     second condition enforce \sum mu_j w_j = mu_target
+  #             
+  Amat = cbind(rep(1,n_stocks),mean_vect)
+  bvec = c(1,NaN)
+  if( is.finite(w_lower_limit) ){
+    Amat = cbind(Amat, diag(1, nrow=n_stocks))
+    bvec = c(bvec, w_lower_limit*rep(1,n_stocks))
+  }
+  if( is.finite(w_upper_limit) ){
+    Amat = cbind(Amat, -diag(1, nrow=n_stocks))
+    bvec = c(bvec, -w_upper_limit*rep(1,n_stocks))
+  }
+
+  # storage for results: 
+  sdP = muP
+  weights = matrix(0, nrow=length(muP), ncol=n_stocks)
+  
+  print(bvec)
+  print(Amat)
+  
+  # find the optimal portfolios for each target expected return:
+  #
+  for( i in 1:length(muP) ){
+    bvec[2] = +muP[i] # enforce portfolio mean constraint 
[…]

### 2019-11-25 Mathematical Statatistics/ms_ch_03.r
mathematical statistics, ch3 work
mathematical statistics, ch3 work

@@ -181,3 +181,54 @@ p <- ( sum( result <= observed ) + 1) / ( N + 1)
 p * 100
 v <- p*(1 - p)/N
 
+### Recidivism Data
+
+Recid <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
+                                 header = T))
+
+k <- complete.cases(Recid$Age25)
+
+Recid2 <- Recid[k]
+Recid2[, .(Recid = sum(Recid == "Yes"), Pct = sum(Recid == "Yes") / .N), by = Age25]
+
+observed <- .365 - 0.306
+N <- 10e4 - 1
+
+result <- numeric(N)
+
+RecidR <- ifelse(Recid2$Recid == "Yes", 1, 0)
+
+for(i in 1:N)
+{
+  index <- sample(nrow(Recid2), 3077, replace = F)
+  result[i] <- mean(RecidR[index]) - mean(RecidR[-index]) 
+}
+
+2 * (sum(result >= observed) + 1) / (N + 1)
+
+### Diving Scores
+
+Diving <- data.table(read.csv(paste0(data.dir, "Diving2017.csv"),
+                             header = T))
+
+Diff <- Diving$Final - Diving$Semifinal # Difference in two scores
+observed <- mean(Diff)                  # mean of differences 
+
+N <- 10e5-1
+result <- numeric(N)
+
+for(i in 1:N)
+{
+  Sign <- sample(c(-1,1), 12, replace = T) # Random vector of 1's or -1's
+  Diff2 <- Sign*Diff # random pairs (a-b) -> (b-a)
+  result[i] <- mean(Diff2)
+}
+
+ggplot(data.table(result), aes(result)) +
+  geom_histogram(aes(fill = ..count..)) +
+  geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.3) +
+  scale_y_continuous(labels = comma) +
+  labs(title = "Diving Times - Resampling")
+
+p <- 2 * (sum(result >= observed + 1)) / (N + 1)
+v <- p*(1 - p) / N

### 2019-11-25 Mathematical Statatistics/ms_ch_03.rmd
ms, ch3 exercises

@@
 
-#### Section
+#### Chapter 3 Exercises
 
-### Subsection
+### 3.1
+
+Suppose you conduct an experiment and inject a drug into three mice.
+
+Their times for running a maze are 8, 10, 15 s; the times for two control mice are 5 and 9 s.
+
+a.) Compute the difference in mean times between the treatment group and the control group.
+
+```{r, echo = T}
+```
+
+b.) Write out all the possible permutations of these times to the two groups and calculate the diffence in means.
+
+```{r, echo = T}
+```
+c.) What proportion of the differences are as large or larger than the observed differences in mean times?
+
+```{r, echo = T}
+```
+
+d.) For each permutation, calculate the mean of the treatment group only.
+
+What proportion of these means are as large or larger than the observed mean of the treatment group?
+
+```{r, echo = T}
+```
+
+### 3.2

### 2019-11-25 Mathematical Statistics/ms_ch_02.rmd
ms, ch3
ms, ch3

--- 2019-11-23 Mathematical Statatistics/ms_ch_02.rmd
@@
 library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
+library(EnvStats, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-Does there exist a __q__ such that $F(q) = 0.05$?
++ Does there exist a __q__ such that $F(q) = 0.05$?
+
+No, the discrete nature of this distribution doesn't have a value that falls in this interval.
 

### 2019-11-25 Mathematical Statistics/ms_ch_03.rmd
ms, ch3
ms, ch3

--- 2019-11-25 Mathematical Statatistics/ms_ch_03.rmd
@@
 library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
+library(gtools, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 ```{r, echo = T}
+
+mice.t <- c(8, 10, 15)
+mice.c <- c(5, 9)
+
+observed <- mean(mice.t) - mean(mice.c)
+
+observed
 ```
@@
 ```{r, echo = T}
+
+mice <- c(mice.t, mice.c)
+
+treatment <- combinations(n = 5, r = 3, mice, repeats.allowed = F)
+
+control <- matrix(nrow = 10, ncol = 2)
+for( i in 1:nrow(control))
+{
+  control[i,] <- mice[!mice %in% treatment[i,]]
+}
+
+perms <- data.table(cbind(treatment, control))
+
+colnames(perms) <- c("D1", "D2", "D3", "C1", "C2")
+
+perms$Xd <- (perms$D1 + perms$D2 + perms$D3) / 3
+perms$Xc <- (perms$C1 + perms$C2) / 2
+perms$Diff <- round(perms$Xd - perms$Xc, 2)
+
 ```
+
+\newpage
+
+```{r, echo = F}
+pretty_kable(perms, "Mice Permutations")
+```
+
 c.) What proportion of the differences are as large or larger than the observed differences in mean times?
@@
 ```{r, echo = T}
+
+gte.observed <- perms[Diff >= observed]
+
+pretty_kable(gte.observed, "Greater than or Equal to Observed")
+
+p1c <- nrow(gte.observed) / nrow(perms)
+
 ```
+
+__Proportion of differences greater than or equal to observed: `r p1c*100`%__
 
@@
 ```{r, echo = T}
+
+gte.t <- perms[ Xd >= mean(mice.t),]
+pretty_kable(gte.t, "Mean Treatment Greater than Observed")
+
+p1d <- nrow(gte.t) / nrow(perms)
+
 ```
 
+__Proportion of treatment groups greater than observed: `r p1d*100`%__
+
 ### 3.2
+
+Your statistics professor comes to class with a big urn that she claims contains 9,999 blue marbels and 1 red marble.
+
+You draw our one marble at random and finds that it is red.
+
+Would you be willing to tell your professor that you think she is wrong about the distribution of colors?
+
+Why or why not?
+
+What are you assuming in making your decision?
+
+What if instead, she claims there are nine blue marbles and 1 red one (and you draw out a red marble)?
+
+### 3.3
+
+In a hypothesis test comparing two populations means, $H_0 : \mu_1 = \mu_2\ versus \ H_A : \mu_1 > \mu_2$
+
+a.) Which P-value, 0.03 or 0.006 provides stronger evidence for the alternative hypothesis?
+
+b.) Which P-value, 0.095 or 0.04 provides stronger evidence that chance alone might account for the observed result?
+
+### 3.4
+
+In the algorithms for conducting a permutation test, why do we add 1 to the number of replications N when calculating the P-Value?
+
+### 3.5
+
+In the flight delays case study in Section 1.1, the data contain flight delays for two airlines, American Airlines and United Airlines.
+
+a.) Conduct a two-sided permutation test to see if the difference in mean delay times between the two carriers are statistically significant.
+
+b.) The flights took place in May and June of 2009. Conduct a two-sided permutation test to see if the differences in mean delay times between two months is statistically significant.

### 2019-11-25 Mathematical Statistics/ms_ch_03.r

@@ -1,5 +1,6 @@
 library(data.table)
 library(ggplot2)
+library(gtools)
 
 ### Beer/Wings Consumption
 
@@ -184,7 +185,7 @@ v <- p*(1 - p)/N
 ### Recidivism Data
 
 Recid <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
-                                 header = T))
+                             header = T))
 
 k <- complete.cases(Recid$Age25)
 
@@ -209,7 +210,7 @@ for(i in 1:N)
 ### Diving Scores
 
 Diving <- data.table(read.csv(paste0(data.dir, "Diving2017.csv"),
-                             header = T))
+                              header = T))
 
 Diff <- Diving$Final - Diving$Semifinal # Difference in two scores
 observed <- mean(Diff)                  # mean of differences 
@@ -232,3 +233,30 @@ ggplot(data.table(result), aes(result)) +
 
 p <- 2 * (sum(result >= observed + 1)) / (N + 1)
 v <- p*(1 - p) / N
+
+# Table 3.1
+
+drug.t <- c(30, 25, 20)
+drug.c <- c(18, 21, 22)
+
+drug <- c(drug.t, drug.c)
+
+treatment <- combinations(n = 6, r = 3, drug, repeats.allowed = F)
+
+control <- matrix(nrow = 20, ncol = 3)
+for( i in 1:nrow(c))
+{
+  control[i,] <- drug[!drug %in% treatment[i,]]
+}
+control
+
+perms <- data.table(cbind(treatment, control))
+
+colnames(perms) <- c("D1", "D2", "D3", "C1", "C2", "C3")
+
+perms$Xd <- (perms$D1 + perms$D2 + perms$D3) / 3
+perms$Xc <- (perms$C1 + perms$C2 + perms$C3) / 3
+perms$Diff <- round(perms$Xd - perms$Xc, 2)
+
+perms[Diff >= 4.67]
+

### 2019-11-25 Mathematical Statistics/ms_ch_03.rmd
mathematical statistics, ch3
mathematical statistics, ch3

@@
 
+# 5 choose 3 for treatment
+
 treatment <- combinations(n = 5, r = 3, mice, repeats.allowed = F)
@@
 perms <- data.table(cbind(treatment, control))
+
+stopifnot(nrow(perms) == choose(5, 3))
 
@@
 gte.t <- perms[ Xd >= mean(mice.t),]
-pretty_kable(gte.t, "Mean Treatment Greater than Observed")
+pretty_kable(gte.t, "Mean Treatment Greater than or Equal to Observed")
 
@@
 
++ Yes, a 1/10,000 chance is pretty rare.
+
 What are you assuming in making your decision?
@@
 What if instead, she claims there are nine blue marbles and 1 red one (and you draw out a red marble)?
+
++ A 1/10 chance is fairly common.
 
@@
 
-a.) Which P-value, 0.03 or 0.006 provides stronger evidence for the alternative hypothesis?
+__a.)__ Which P-value, _0.03_ or _0.006_, provides stronger evidence for the alternative hypothesis?
 
-b.) Which P-value, 0.095 or 0.04 provides stronger evidence that chance alone might account for the observed result?
+_0.03_ provides stronger evidence for the alternative hypothesis.
+
+__b.)__ Which P-value, 0.095 or 0.04, provides stronger evidence that chance alone might account for the observed result?
+
+_0.095_ provides stronger evidence that chance alone is responsible for the observed result.
 
@@
 
+__Answer__: We need to account for the original observed result.
+
 ### 3.5
@@
 
+```{r, echo = T}
+
+Flights <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
+                               header = T))
+
+```
+
 a.) Conduct a two-sided permutation test to see if the difference in mean delay times between the two carriers are statistically significant.
 
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+Flights[, .(Delay = mean(Delay)), by = Carrier]
+
+observed <- mean(Flights[Carrier == "UA"]$Delay) - mean(Flights[Carrier == "AA"]$Delay)
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
+   results[i] <- mean(Flights[index]$Delay) - mean(Flights[-index]$Delay)
+}
+
+p <- (sum(results[results >= observed]) + 1) / ( N + 1)
+v <- p*(1 - p) / ( N + 1 )
+
+ggplot(data.table(results)) +
+   geom_histogram(aes(results, fill = ..count..)) +
+   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("Flight Delay Times, UA/AA vs Observed, p=", p ))
+
+```
+
 b.) The flights took place in May and June of 2009. Conduct a two-sided permutation test to see if the differences in mean delay times between two months is statistically significant.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+Flights[, .(Delay = mean(Delay)), by = Month]
+
+observed <- mean(Flights[Month == "May"]$Delay) - mean(Flights[Month == "June"]$Delay)
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Flights), nrow(Flights[Month == "May"]), replace = F)
+   results[i] = mean(Flights[index]$Delay) - mean(Flights[-index]$Delay)
+}
+
+p <- (sum(results[results <= observed]) + 1) / ( N + 1)
+v <- p*(1 - p) / ( N + 1 )
+
+ggplot(data.table(results)) +
+   geom_histogram(aes(results, fill = ..count..)) +
+   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("Flight Delay Times, May/June, vs Observed, p=", round(p, 5)))
+
+```

### 2019-11-25 Mathematical Statistics/ms_ch_03.r

@@ -239,6 +239,8 @@ v <- p*(1 - p) / N
 drug.t <- c(30, 25, 20)
 drug.c <- c(18, 21, 22)
 
+choose(6, 3)
+
 drug <- c(drug.t, drug.c)
 
 treatment <- combinations(n = 6, r = 3, drug, repeats.allowed = F)
@@ -251,6 +253,8 @@ for( i in 1:nrow(c))
 control
 
 perms <- data.table(cbind(treatment, control))
+
+stopifnot(nrow(perms) == choose(6, 3))
 
 colnames(perms) <- c("D1", "D2", "D3", "C1", "C2", "C3")
 

### 2019-11-25 Mathematical Statistics/ms_ch_03.rmd
mathematical statistics, ch3
mathematical statistics, ch3

@@
 
-p <- (sum(results[results >= observed]) + 1) / ( N + 1)
+# two-sided test
+p <- 2 * (sum(results[results >= observed]) + 1) / ( N + 1)
 v <- p*(1 - p) / ( N + 1 )
@@
 ggplot(data.table(results)) +
-   geom_histogram(aes(results, fill = ..count..)) +
+   geom_histogram(aes(results, fill = ..count..), bins = 30) +
    geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
    scale_y_continuous(labels = comma) +
-   labs(title = paste0("Flight Delay Times, UA/AA vs Observed, p=", p ))
+   labs(title = paste0("Flight Delay Times, UA/AA vs Observed, p=", round(p, 5)),
+        subtitle = paste0("Observed Value: ", round(observed, 4)))
 
@@
 
-p <- (sum(results[results <= observed]) + 1) / ( N + 1)
+# two-sided test
+
+p <- 2 * (sum(results[results <= observed]) + 1) / ( N + 1)
 v <- p*(1 - p) / ( N + 1 )
@@
 ggplot(data.table(results)) +
-   geom_histogram(aes(results, fill = ..count..)) +
+   geom_histogram(aes(results, fill = ..count..), bins = 30) +
    geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
    scale_y_continuous(labels = comma) +
-   labs(title = paste0("Flight Delay Times, May/June, vs Observed, p=", round(p, 5)))
+   labs(title = paste0("Flight Delay Times, May/June, vs Observed, p=", round(p, 5)),
+        subtitle = paste0("Observed Value: ", round(observed, 4)))
 
 ```
+
+### 3.6
+
+In the flight delays case study in Section 1.1, the data contains flight delays for two airlines, American and United.
+
+a.) Compute the proportion of times that each carrier's flight was delays more than 20 min.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+Flights[, .(Delay20 = sum(Delay > 20) / .N), by = Carrier]
+
+observed <- as.numeric(Flights[Carrier == "UA", .(Delay = sum(Delay > 20)/.N)] - Flights[Carrier == "AA", .(Delay = sum(Delay > 20)/.N)])
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
+   results[i] <- as.numeric(Flights[index, .(Delay = sum(Delay > 20)/.N)] - Flights[-index, .(Delay = sum(Delay > 20)/.N)])
+}
+
+p <- 2 * (sum(results[results >= observed]) + 1) / ( N + 1)
+v <- p*(1 - p) / ( N + 1 )
+
+ggplot(data.table(results)) +
+   geom_histogram(aes(results, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("Flight Delay Times, Over 20 Minutes, vs Observed, p=", round(p, 5)), 
+        subtitle = paste0("Observed Value: ", round(observed, 4)))
+
+```
+
++ Conduct a two-sided test to see if the difference in these proportions is statistically significant.
+
+__Answer__: There is statistical significance with a P-value < 0.0001.
+
+b.) Compute the variance in the flight delay lengths for each carrier.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+Flights[, .(Variance = var(Delay)), by = Carrier]
+
+observed <- var(Flights[Carrier == "UA"]$Delay) - var(Flights[Carrier == "AA"]$Delay)
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
+   results[i] <- var(Flights[index]$Delay) - var(Flights[-index]$Delay)
+}
+
+p <- min(1, 2 * (sum(results[results >= observed]) + 1) / ( N + 1))
+v <- p*(1 - p) / ( N + 1 )
+
+ggplot(data.table(results)) +
+   geom_histogram(aes(results, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("Flight Delay Time Variance, vs Observed, p=", round(p, 5)), 
+        subtitle = paste0("Observed Value: ", round(observed, 4)))
+
+```
+
++ Conduct a test to see if the variance for United Airlines differes from that of American Airlines.
+
+__Answer__: There does not appear to be a statistically significant difference in the variance in delay times between airlines.
+
+\newpage
+
+### 3.7
+
+In the flight delays case study in Section 1.1, repeat Excercise 3.5 part __(a)__ using three test statistics,
+
++ i.) The mean of the United Airline delay times
++ ii.) The sum of the United Airline delay times
++ iii.) The difference in the means
+
+Compare the P-values.
+
+Make sure all three test statistics are computed within the same __for__ loop.
+
+What do you observe?
+
+```{r, echo = T, fig.height=7.5, fig.width=8}
+
+ua.delay <- Flights[Carrier == "UA"]$Delay
+aa.delay <- Flights[Carrier == "AA"]$Delay
+
+observed.mean <- mean(ua.delay)
+observed.sum <- sum(ua.delay)
+observed.diff <- mean(ua.delay) - mean(aa.delay)
+
+N <- 10e2 - 1
+results <- matrix(nrow = N, ncol = 3)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Flights), length(ua.delay), replace = F)
+
+   results[i, 1] <- mean(Flights[index]$Delay)
+   results[i, 2] <- sum(Flights[index]$Delay)
+   results[i, 3] <- mean(Flights[index]$Delay) - mean(Flights[-index]$Delay)
+}
+
+dt.results <- data.table(results)
+colnames(dt.results) <- c("Mean", "Sum", "MeanDiff")
+
+p.mean <- ( sum( dt.results$Mean >= observed.mean ) + 1 ) / ( N + 1)
+p.sum <- ( sum(dt.results$Sum >= observed.sum) + 1) / ( N + 1 )
+p.diff <- ( sum(dt.results$MeanDiff >= observed.diff) + 1) / ( N + 1)
+
+p1 <- ggplot(dt.results) +
+   geom_histogram(aes(Mean, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = observed.mean, col = "darkorange", lwd = 1.2, linetype = 3) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("United Airlines - Mean Delay Time vs Observed, p=", round(p.mean, 5)), 
+        subtitle = paste0("Observed Value: ", round(observed.mean, 4)))
+
+p2 <- ggplot(dt.results) +
+   geom_histogram(aes(Sum, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = observed.sum, col = "darkorange", lwd = 1.2, linetype = 3) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("United Airlines - Sum Delay Time vs Observed, p=", round(p.sum, 5)), 
+        subtitle = paste0("Observed Value: ", round(observed.sum, 4)))
+
+p3 <- ggplot(dt.results) +
+   geom_histogram(aes(MeanDiff, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = observed.diff, col = "darkorange", lwd = 1.2, linetype = 3) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("Mean Delay Time Difference vs Observed, p=", round(p.diff, 5)), 
+        subtitle = paste0("Observed Value: ", round(observed.diff, 4)))
+
+grid.arrange(p1, p2, p3, nrow = 3)
+
+```

### 2019-11-26 Mathematical Statistics/ms_ch_03.rmd
mathematical statistics, ch3 work
mathematical statistics, ch3 work

@@
 
-ua.delay <- Flights[Carrier == "UA"]$Delay
-aa.delay <- Flights[Carrier == "AA"]$Delay
+UA.Delay <- Flights[Carrier == "UA"]$Delay
+AA.Delay <- Flights[Carrier == "AA"]$Delay
 
-observed.mean <- mean(ua.delay)
-observed.sum <- sum(ua.delay)
-observed.diff <- mean(ua.delay) - mean(aa.delay)
+observed.mean <- mean(UA.Delay)
+observed.sum <- sum(UA.Delay)
+observed.diff <- mean(UA.Delay) - mean(AA.Delay)
 
@@
 {
-   index <- sample(nrow(Flights), length(ua.delay), replace = F)
+   index <- sample(nrow(Flights), length(UA.Delay), replace = F)
 
@@
 ```
+
+\newpage
+
+### 3.8
+
+In the flight delays case study in Section 1.1,
+
+a.) Find the trimmed mean of the delay times for United Airlines and American Airlines.
+
+```{r, echo = T}
+
+trim.amount <- .25
+UA.trimmed <- mean(UA.Delay, trim = trim.amount)
+AA.trimmed <- mean(AA.Delay, trim = trim.amount)
+
+observed <- UA.trimmed - AA.trimmed
+
+pretty_kable(data.table( UA = UA.trimmed, AA = AA.trimmed), "Trimmed Means")
+
+```
+
+b.) Conduct a two-sided test to see if the difference in trimmed means is statistically significant.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+N <- 10e2 - 1
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
+   results[i] <- mean(Flights[index]$Delay, trim = trim.amount) - mean(Flights[-index]$Delay, trim = trim.amount)
+}
+
+p <- min(1, 2 * (sum(results[results >= observed]) + 1) / ( N + 1))
+v <- p*(1 - p) / ( N + 1 )
+
+ggplot(data.table(results)) +
+   geom_histogram(aes(results, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
+   scale_y_continuous(labels = comma) +
+   labs(title = paste0("Flight Delay Time Trimmed Mean vs Observed, p=", round(p, 5)), 
+        subtitle = paste0("Observed Value: ", round(observed, 4)))
+```
+
+### 3.9
+
+In the flight delays case study in Section 1.1,
+
+a.) Compute the proportion of times the flights in May and in June were delayed more than 20 min.
+
+```{r, echo = T}
+
+delay20_month <- Flights[, .(Delay20 = sum(Delay > 20) / .N), by = Month]
+
+pretty_kable(delay20_month, "Delayed by more than 20 min, by month")
+
+```
+
+Conduct a two-sided test for statistical significance.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+observed <- delay20_month[Month == "May"]$Delay20 - delay20_month[Month == "June"]$Delay20
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N)
+{
+   index <- sample(nrow(Flights), nrow(Flights[Month == "May"]), replace = F)
+   results[i] <- as.numeric( Flights[index, .(Delay = sum(Delay > 20) / .N)] - Flights[-index, .(Delay = sum(Delay > 20) / .N)] )
+}
+
+p <- min(1, 2 * (sum(results[results >= observed]) + 1) / ( N + 1))
+v <- p*(1 - p) / ( N + 1 )
+
+```
+
+P-value: `r round(p, 4)`, which is statistically significant.
+
+b.) Compute the ratio of the variances in the flight delay times in May and in June.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+observed <- var(Flights[Month == "May"]$Delay) - var(Flights[Month == "June"]$Delay)
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N)
+{
+   index <- sample(nrow(Flights), nrow(Flights[Month == "May"]), replace = F)
+   results[i] <- var( Flights[index, .(Delay)] ) - var( Flights[-index, .(Delay)] )
+}
+
+p <- min(1, 2 * (sum(results[results >= observed]) + 1) / ( N + 1))
+v <- p*(1 - p) / ( N + 1 )
+
+```
+
+Is this evidence that the true ratio is not equal to 1, or could this be due to chance variability?
+
+The variance appear to be due to random chance, so there does appear to be a stastical significance between the two months.
+
+Conduct a two-sided test to check.
+
+P-value: __`r p`__
+
+### 3.10
+
+In the black spruce case study in Section 1.10, seedlings were planted in plots that were either subject to competition (from other plants), or not.
+
+Use the data set _Spruce_ to conduct a test to see if the mean difference is how much the seedlings grow (in height) over the corse of the study under these two treatments is stastically significant.
+
+```{r, echo = T}
+Spruce <- data.table(read.csv(paste0(data.dir, "Spruce.csv"),
+                               header = T))
+
+observed <- mean( Spruce[Competition == "NC"]$Ht.change ) - mean( Spruce[Competition == "C"]$Ht.change )
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Spruce), nrow(Spruce[Competition == "NC"]), replace = F)
+   resutls <- mean( Spruce[index]$Ht.change ) - mean( Spruce[-index]$Ht.change )
+}
+
+p <- min(1, 2 * (sum(results[results >= observed]) + 1) / ( N + 1))
+
+```
+
+There is stastical significance in between the heights of the two groups (Competition / No-competition).
+
+P-value: __`r p`__
+
+### 3.11
+
+The file _Phillies2009_ contains data from the 2009 season for the baseball team the Philadelphia Phillies.
+
+```{r, echo = T}
+Phillies <- data.table(read.csv(paste0(data.dir, "Phillies2009.csv"),
+                               header = T))
+```
+
+a.) Compare the empirical distribution functions of the number of strike-outs per game (_StrikeOuts_) for games played at home and games played away (_Location_).
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+
+ggplot(Phillies, aes(StrikeOuts, color = Location)) +
+   stat_ecdf(stat = "point") +
+   labs(title = "Strike-outs / Game, by Location", y = "Percent")
+
+```
+
+b.) Find the mean number of strike-outs per game for the home and the away games.
+
+```{r, echo = T}
+strikeouts <- Phillies[, .(StrikeOuts = mean(StrikeOuts)), by = Location]
+
+pretty_kable(strikeouts, "StrikeOuts by Location")
+
+observed <- mean(strikeouts[Location == "Away"]$StrikeOuts) - mean(strikeouts[Location == "Home"]$StrikeOuts)
+
+```
+c.) Perform a permutation tests to see if the differences in means is statistically significant.
+
+```{r, echo = T}
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N)
+{
+   index <- sample(nrow(Phillies), nrow(Phillies[Location == "Home"]), replace = F)
+   results[i] <- mean(Phillies[index]$StrikeOuts) - mean(Phillies[-index]$StrikeOuts)
+}
+
+p <- min(1, ( sum(results >= observed) + 1) / ( N + 1) )
+
+```
+
+There does not appear to be a statistically significat relationship between the number of home and away strikeouts.
+
+P-value: __`r p`__

### 2019-11-26 Mathematical Statistics/ms_ch_03.rmd
Mathematical Statistics, ch 3
Mathematical Statistics, ch 3

@@
 library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
-library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
 library(scales, quietly = TRUE, warn.conflicts = FALSE)
@@
 P-value: __`r p`__
+
+### 3.12
+
+In the Iowa recidivism case study in Section 1.4, offenders had originally been convicted of either a felony or misdemenor.
+
+```{r, echo = T}
+Recidivism <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
+                               header = T))
+```
+
+a.) Use R to create a table displaying the proportion of felons who recidivated and the propportion of those convictted of a misdemeanor who recidivated.
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+
+Recidivism[, Recidivated := is.na(Days)]
+
+Recidivated <- Recidivism[, .(RecidPct = sum(Recidivated) / .N), by = Offense]
+
+pretty_kable(Recidivated, "Recidivated")
+
+```
+
+b.) Determine whether or not the difference in recidivism proportions computed in (a) is statistically significant.
+
+```{r, echo = T}
+
+observed <- as.numeric( Recidivated[Offense == "Felony"]$RecidPct - Recidivated[Offense == "Misdemeanor"]$RecidPct )
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N )
+{
+   index <- sample(nrow(Recidivism), nrow(Recidivism[Offense == "Felony"]), replace = F)
+   results[i] <- as.numeric( Recidivism[index, .(Pct = sum(Recidivated) / .N)] - Recidivism[-index, .(Pct = sum(Recidivated) / .N)] )
+}
+
+p <- ( sum(results >= observed) + 1 ) / ( N + 1 )
+
+```
+
+It appears there is statistical significance in the recidivism rates between felony and misdemeanor convictions.
+
+P-value: `r p`
+
+### 3.13
+
+In the Iowa recidivism case study in Section 1.4, for those offenders who recidivated, we have data on the number of days until they reoffended.
+
+For those offenders who did recidivate, determine if the difference in the mean number of days (__Days__) until recidivism between those under 25 years of age and those 25 years of age and older is statistically significant.
+
+```{r, echo = T}
+
+Recid.Age25 <- Recidivism[, .(RecidDays = mean(Days, na.rm = T)), by = Age25]
+
+pretty_kable(Recid.Age25, "Days Until Recidivism by Age ( < 25 < )")
+
+observed <- as.numeric( Recid.Age25[Age25 == "Under 25"]$RecidDays - Recid.Age25[Age25 == "Over 25"]$RecidDays)
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N )
+{
+   index <- sample(nrow(Recidivism), nrow(Recidivism[Age25 == "Under 25"]), replace = F)
+   results[i] <- mean(Recidivism[index]$Days, na.rm = T) - mean(Recidivism[-index]$Days, na.rm = T)
+}
+
+p <- min(1, ( sum(results >= observed) + 1 ) / ( N + 1 ) )
+
+```
+
+There does not appear to be a statistical difference between the days until recidivism by those over/under 25 years of age.
+
+P-value: `r p`
+
+### 3.14
+
+Does chocolate ice cream have more calories than vanilla ice cream? The data set __IceCream__ contains calorie information for a sample of brands of chocolate and vanilla ice cream.
+
+```{r, echo = T}
+IceCream <- data.table(read.csv(paste0(data.dir, "IceCream.csv"),
+                               header = T))
+```
+
+a.) Inspect the data set, then explain why this is an example of matched pairs data.
+
+```{r, echo = T}
+pretty_kable(IceCream, "Ice Cream Data")
+```
+
+This data set is matched pairs becase there are multiple brands associated with each flavor. They are not independent samples.
+
+b.) Compute summary statistics of the number of calories for the two flavors.
+
+```{r, echo = T}
+
+pretty_kable(IceCream[, .(Vanilla = mean(VanillaCalories), Chocolate = mean(ChocolateCalories))], 
+             "Ice cream Calories by Flavor")
+
+```
+
+c.) Conduct a permutation tests to determine whether or not chocolate ice cream has, on average, more calories than vanilla ice cream.
+
+```{r, echo = T}
+
+diff <- IceCream$VanillaCalories - IceCream$ChocolateCalories
+observed <- mean(diff)
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N )
+{
+   Sign <- sample(c(-1, 1), nrow(IceCream), replace = T)
+   diff2 <- Sign * diff
+   results[i] <- mean(diff2)
+}
+
+p <- ( sum( results <= observed ) + 1 ) / ( N + 1 )
+
+```
+
+The data supports the claim that chocolate ice cream has more calories than vanilla.
+
+P-value: `r p`
+
+### 3.15
+
+Is there a difference in the price of groceries sold by the two retailers Target and Walmart?
+
+The data set Groceries contains a sample of grocery items and their prices advertised on their respective web sites on one specific day.
+
+```{r, echo = T}
+Groceries <- data.table(read.csv(paste0(data.dir, "Groceries.csv"),
+                               header = T))
+
+```
+
+a.) Inspect the data set, then explain why this is an example of matched pairs data.
+
+```{r, echo = T}
+pretty_kable(Groceries, "Target vs Walmart")
+```
+
+b.) Compute summary statistics of the prices for each store.
+
+```{r, echo = T}
+
+diff <- Groceries$Target - Groceries$Walmart
+
+observed <- mean(diff)
+
+```
+
+Observed difference: __`r observed`__
+
+c.) Conduct a permutation test to determine whether or not there is a difference in the mean prices.
+
+```{r, echo = T}
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N )
+{
+   Sign <- sample(c(-1, 1), nrow(Groceries), replace = T)
+   diff2 <- Sign * diff
+   results[i] <- mean(diff2)
+}
+
+p <- min(1, 2 * ( sum(results >= observed ) + 1 ) / ( N + 1) )
+
+```
+
+There does not appear to be a statistical difference in the prices at the two stores.
+
+P-value: `r p`
+
+d.) Create a histogram of the difference in prices.
+
+```{r, echo = T, fig.height=5, fig.width=8}
+
+ggplot(data.table( Product = Groceries$Product, Diff = diff)) +
+   geom_bar(aes(Product, Diff, fill = Product), stat = "identity") +
+   coord_flip() +
+   labs(title = "Grocery Price Difference (Target vs Walmart)") +
+   theme(legend.position = "none")
+
+```
+
+What is unusual about Quaker Oats Life cerial?
+
+__Price difference is a multiple SD outlier.__
+
+e.) Redo the hypothesis test without this observation.
+
+```{r, echo = T}
+
+diff <- Groceries[Product != "Quaker Oats Life Cereal  Original ", .(Diff = Target - Walmart ) ]$Diff
+
+observed <- mean(diff)
+
+N <- 10e2 - 1
+results <- numeric(N)
+
+for( i in 1:N )
+{
+   Sign <- sample(c(-1, 1), length(diff), replace = T)
+   diff2 <- Sign * diff
+   results[i] <- mean(diff2)
+}
+
+p <- min(1, 2 * ( sum(results >= observed ) + 1 ) / ( N + 1) )
+
+```
+
+Do you reach the same conclusion?
+
+__No, there is a statistical difference in the prices of products at the two retailers with the outlier removed.__
+
+P-value: `r p`
+
+### 3.16
+
+In the sampling version of permutation testing, the one-sided P-value is
+
+$\hat{P} = \frac{(X + 1)}{(N + 1)}$, where X is the number of permutation test statistics that are as large or larger than the observed test statistic.
+
+Suppose the true P-value (for the exaustive test, conditional on the observed data) is __p__.
+
+a.) What is the variance of $\hat{P}$?
+
+$\mathbb{V} = p*(1 - p) / (N + 1)$
+
+b.) What is the variance of $\hat{P}_2$ for the two-sided test (assuming that __p__ is not close to 0.5, where __p__ is the smaller true one-side P-value)?

### 2019-11-26 Mathematical Statistics/ms_ch_04.r
Computational Finance, ch1
Computational Finance, ch1

@@ -1,0 +1,57 @@
+N <- 1000
+Xbar <- numeric(N) # space for results
+
+for( i in 1:n)
+{
+  x <- rexp(100, rate = 1/15) # draw random sample of size 100
+  Xbar[i] <- mean(x)
+}
+
+hist(Xbar)
+
+qqnorm(Xbar)
+qqline(Xbar)
+
+mean(Xbar)
+sd(Xbar)
+
+maxY <- numeric(N)
+
+for( i in 1:N)
+{
+  y <- runif(12) # draw random sample of size 12
+  maxY[i] <- max(y)
+}
+
+hist(maxY)
+
+X <- rpois(10^4, 5) # Draw 10^4 values from Pois(5)
+Y <- rpois(10^4, 12) # Draw 10^4 values from Pois(12)
+
+W <- X + Y
+
+hist(W, prob = T) # prob = T, scales hist to 1
+lines(2:35, dpois(2:35, 17), type = "b")
+
+mean(W)
+var(W)
+
+sd(W) / sqrt(N)
+
+Xbar <- numeric(N)
+
+for( i in 1:N )
+{
+  x <- rgamma(30, shape = 5, rate = 2)
+  Xbar[i] <- mean(x)
+}
+
+hist(Xbar)
+
+qqnorm(Xbar)
+qqline(Xbar)
+
+mean(Xbar)
+sd(Xbar)
+
+mean(Xbar > 3)

### 2019-11-27 Computational Finance/cf_ch_02.rmd
mathematical statistics, ch4
mathematical statistics, ch4

--- 2019-11-11 Computational Finance/cf_ch_01.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Exploratory Data Analysis}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistics of Financial Time Series}
    \rfoot{\color{headergrey}Chapter 2}
    \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Finance}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 
-#### Chapter 1
+#### Chapter 2
 
-### Ex 1.1
+### Ex 2.1
 
-We shall use the __getSymbols__ function in the quantmod package to retrieve financial data for General Electric (__GE__).
+Consider the curve drawn by the price history of the Dow Jones Industrial Average (DJIA) from 1960 to 2010.
 
@@
 
-getSymbols("GE", src = "yahoo", from = "2000-01-01", to = "2009-12-30")
+getSymbols("DJIA", src = "yahoo", from = "1960-01-01") ## DJIA from 1896
 
-names(GE)
-
-GE["2000-01-01/2000-01-20"]
-
-geAdj = GE$GE.Adjusted["2000-01-01-/2000-01-20"]; geAdj
-
-max(geAdj); min(geAdj); mean(geAdj)
+plot(DJIA['1960/2010'], main = "DJIA")
 
@@
 ```{r echo = T, fig.height=3.5}
-chartSeries(GE)
+
+getSymbols("AAPL", src = "yahoo")
+aplRd <- periodReturn(AAPL, period = "daily")
+plot(aplRd, main = "APPLE Daily Returns")
 ```
 
-```{r echo = T, fig.height=3.5}
-chartSeries(GE, TA=NULL, subset='2001-01::2001-02')
+```{r, echo = T}
+x <- seq(-4, 4, 0.01)
+plot(x, dnorm(x, mean = 0.5, sd = 1.3), type = "l")
 ```
 
-```{r echo=T}
-saveRDS(GE, file = "GE.rds")
+```{r, echo=T}
+
 ```

### 2019-11-27 Mathematical Statistics/ms_ch_04.rmd
Computational Finance, ch1
Computational Finance, ch1
mathematical statistics, ch4
mathematical statistics, ch4

--- 2019-11-23 Mathematical Statatistics/ms_ch_03.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Introduction to Hypothesis - Testing Permutation Tests}
-   \rfoot{\color{headergrey}Chapter 3}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Sampling Distributions}
+   \rfoot{\color{headergrey}Chapter 4}
    \lfoot{\color{headergrey}}
@@
 
-#### Section
+#### Sampling Distributions
 
-### Subsection
+### 4.1
+
+Consider the population {1, 2, 5, 6, 10, 12}.
+
+Find (and plot) the sampling distribution of medians for samples of size 3 without replacement.
+
+```{r, echo = T, fig.height=4.5, fig.width=8}
+
+p <- c(1, 2, 5, 6, 10, 12)
+c <- combinations(v = p, n = 6, r = 3)
+m <- apply(c, 1, median)
+
+ggplot(data.table(value = m), aes(value, fill = ..count..)) +
+   geom_histogram(bins = 30) +
+   labs(title = "Median Sampling Distribution of p")
+
+```
+
+Compare the median of the population to the mean of the medians.
+
+Median of p = __`r median(p)`__. Mean of Medians of p = __`r mean(m)`__
+
+### 4.2
+
+Consider the population {3, 6, 7, 9, 11, 14}.
+
+For samples of size 3 without replacement, find (and plot) the sampling distribution for the minimum.
+
+What is the mean of the sampling distribution?
+
+The statistic is an estimate of some parameter - what is the value of that parameter?
+
+```{r, echo = T}
+
+p <- c(3, 6, 7, 9, 11, 14)
+
+```

### 2019-11-28 Mathematical Statistics/ms_ch_04.rmd
mathematical statistics
mathematical statistics

@@
 
-#### Sampling Distributions
-
 ### 4.1
@@
 c <- combinations(v = p, n = 6, r = 3)
-m <- apply(c, 1, median)
+t <- apply(c, 1, median)
 
-ggplot(data.table(value = m), aes(value, fill = ..count..)) +
+ggplot(data.table(value = t), aes(value, fill = ..count..)) +
    geom_histogram(bins = 30) +
@@
 
-Median of p = __`r median(p)`__. Mean of Medians of p = __`r mean(m)`__
+Median of p = __`r median(p)`__. Mean of Medians of p = __`r mean(t)`__
+
+\newpage
 
@@
 
-What is the mean of the sampling distribution?
+```{r, echo = T, fig.height=4.5, fig.width=8}
+
+p <- c(3, 6, 7, 9, 11, 14)
+c <- combinations(v = p, n = 6, r = 3)
+t <- apply(c, 1, min)
+
+ggplot(data.table(value = t), aes(value, fill = ..count..)) +
+   geom_histogram(bins = 30) +
+   labs(title = "Minimum Sampling Distribution of p")
+
+```
+
+What is the mean of the sampling distribution? __`r mean(t)`__
 
@@
 
+This is an estimation of the minimum, which is: __`r min(p)`__
+
+\newpage
+
+### 4.3
+
+Let _A_ denote the population {1, 3, 4, 5} and _B_ the population {5, 7, 9}.
+
 ```{r, echo = T}
 
-p <- c(3, 6, 7, 9, 11, 14)
+permutations( v = p, n = 2, r = 2, repeats.allowed = T)
+
+combinations( v = p, n = 3, r = 2, repeats.allowed = T)
+
+A <- c(1, 3, 4, 5)
+B <- c(5, 7, 9)
 
 ```
+
+Let _X_ be a random value from _A_, and _Y_ and random value from _B_.
+
+a.) Find the sampling distribution of _X + Y_.
+
+```{r, echo = T}
+
+p <- c(3, 4, 6, 6)
+
+sample.size <- 2
+perm <- permutations( length(p), sample.size, repeats.allowed = T)
+
+d <- data.table( c1 = p[perm[, 1]], c2 = p[perm[, 2]])
+d[, m := (c1 + c2) / 2]
+
+```
+
+b.) In this example, does the sampling distribution depend on whether you sample with or without replacement?
+
+_No._
+
+Why or why not?
+
+_Because 5 in is both sets._
+
+c.) Compute the mean of the values for each of _A_ and _B_ and the values in the sampling distribution of _X + Y_.
+
+How are the means related?
+
+d.) Suppose you draw a random value from _A_ and a random value from _B_.
+
+What is the probability that the sum is 13 or larger?

### 2019-11-28 Mathematical Statistics/ms_ch_04.r

@@ -55,3 +55,14 @@ mean(Xbar)
 sd(Xbar)
 
 mean(Xbar > 3)
+
+# work
+
+p <- c(3, 4, 6, 6)
+
+sample.size <- 2
+perm <- permutations( length(p), sample.size, repeats.allowed = T)
+
+d <- data.table( c1 = p[perm[, 1]], c2 = p[perm[, 2]])
+d[, m := (c1 + c2) / 2]
+

### 2019-11-28 Mathematical Statistics/ms_ch_04.rmd
mathematical statistics, ch4
mathematical statistics, ch4

@@
 
-permutations( v = p, n = 2, r = 2, repeats.allowed = T)
-
-combinations( v = p, n = 3, r = 2, repeats.allowed = T)
-
 A <- c(1, 3, 4, 5)
@@
 
-p <- c(3, 4, 6, 6)
+result = numeric(12)
+index <- 1
+for(j in 1:length(A))
+{
+   for(k in 1:length(B))
+   {
+      result[index] <- A[j] + B[k]
+      index <- index + 1
+   }
+}
 
-sample.size <- 2
-perm <- permutations( length(p), sample.size, repeats.allowed = T)
-
-d <- data.table( c1 = p[perm[, 1]], c2 = p[perm[, 2]])
-d[, m := (c1 + c2) / 2]
+sort(result)
 
@@
 
+Mean of _A_: __`r mean(A)`__. Mean of _B_: __`r mean(B)`__.
+
+Mean of _A + B_: __`r mean(result)`__
+
 How are the means related?
+
+mean(A) + mean(B) = mean(A + B).
 
@@
 
-What is the probability that the sum is 13 or larger?
+```{r, echo = T}
+
+prob <- sum(result >= 13) / length(result)
+
+```
+
+What is the probability that the sum is 13 or larger? __`r prob*100`%__
+
+\newpage
+
+### 4.4
+
+Consider the population {3, 5, 6, 6, 8, 11, 13, 15, 19, 20}.
+
+a.) Compute the mean and standard deviation and create a dot plot of its distribution.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+p <- c(3, 5, 6, 6, 8, 11, 13, 15, 19, 20)
+
+mu <- mean(p)
+sigma <- sd(p)
+
+ggplot(data.table(value = p)) +
+   geom_dotplot(aes(value, fill = ..count..), binwidth = 1) +
+   labs(title = "Population Dot Plot")
+
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+b.) Simulate the sampling distribution of $\bar{X}$ by taking random samples of size 4 and plot your results.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+N <- 10e2
+results <- numeric(N)
+
+for( i in 1:N)
+{
+   index <- sample(length(p), size = 4, replace = F)
+   results[i] <- mean( p[index] )
+}
+
+ggplot(data.table(value = results)) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs(title = "Sample Means")
+
+xbar <- mean(results)
+se <- sd(results) / sqrt(N)
+
+```
+
+Compute the mean and standard error, and compare to the population mean and standard deviation.
+
+mean: `r xbar`, standard error: `r se`
+
+c.) Use the simulation to find $P(\bar{X} < 11)$.
+
+```{r, echo = T}
+
+prob <- mean(results < 11)
+
+```
+
+$P(\bar{X} < 11) = `r prob*100`$%

### 2019-11-28 Mathematical Statistics/ms_ch_04.rmd
ms ch4 & cf ch2
ms ch4 & cf ch2

@@
 $P(\bar{X} < 11) = `r prob*100`$%
+
+\newpage
+
+### 4.5
+
+Consider two populations A = {3, 5, 7, 9, 10, 16}, B = {8, 10, 11, 15, 18, 25, 28}.
+
+```{r, echo = T}
+A <- c(3, 5, 7, 9, 10, 16)
+B <- c(8, 10, 11, 15, 18, 25, 28)
+```
+
+a.) Using R, draw random samples (without replacement) of size 3 from each population, and simulate the sampling distribution of the sum of their maximums.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+```
+
+Describe the distribution.
+
+b.) Use your simulation to estimate the probability that the sum of the maximums is less than 20.
+
+```{r, echo = T}
+```
+
+c.) Draw random samples of size 3 from each population, and find the maximum of the union of these two sets.
+
+Simulate the sampling distribution of the maximums of this union.
+
+Compare the distribution to part (a).
+
+d.) Use simulation to find the probability that the maximum of the union is less than 20.
+
+### 4.6

### 2019-11-28 Mathematical Statistics/ms_ch_04.rmd
mathematical statistics, ch 4
mathematical statistics, ch 4

@@
 ```{r, echo = T, fig.height=3.5, fig.width=8}
+N <- 10e2
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   samp.a <- sample(A, 3, replace = F)
+   samp.b <- sample(B, 3, replace = F)
+
+   results[i] <- max(samp.a) + max(samp.b) 
+}
+
+ggplot(data.table(value = results)[, index := .I]) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs(title = "Sampling Distribution: max(A) + max(B)")
+
 ```
-
-Describe the distribution.
 
@@
 ```{r, echo = T}
+prob <- mean(results < 20)
 ```
+
+Probability: `r prob*100`%
 
@@
 
-Compare the distribution to part (a).
+```{r, echo = T, fig.height=3.5, fig.width=8}
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   samp.a <- sample(A, 3, replace = F)
+   samp.b <- sample(B, 3, replace = F)
+
+   results[i] <- max(union(samp.a, samp.b))
+}
+
+ggplot(data.table(value = results)[, index := .I]) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs(title = "Sampling Distribution: max(union(A,B))")
+```
 
@@
 
+```{r, echo = T}
+prob <- mean(results < 20)
+```
+
+Probability: `r prob*100`%
+
+\newpage
+
 ### 4.6

### 2019-11-28 Mathematical Statistics/ms_ch_04.rmd
mathematical statistics, ch4
mathematical statistics, ch4

@@
 ### 4.6
+
+The data set _Recidivism_ contains the poopulation of all Iowa offenders convicted of either a felony or misdemeanor who were released in 2010 (case study in Section 1.4).
+
+```{r, echo = T}
+Recidivism <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
+                               header = T))
+```
+
+Of these, 31.6% recidivated and were sent back to prision.
+
+Simulate the sampling distribution of $\hat{p}$, the sample proportion of offeneders who recidivated, for random samples of size 25.
+
+```{r, echo = T}
+
+mean(Recidivism$Recid == "Yes")
+
+N <- 10e2
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- sample(Recidivism$Recid, 25)
+   results[i] <- mean(samp == "Yes")
+}
+
+```
+
+a.) Create a histogram and describe the simulated sampling distribution of $\hat{p}$.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+ggplot(data.table(value = results)) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs(title = "Recidivism Sampling Distribution")
+```
+
+Estimate the mean and standard error.
+
+```{r, echo = T}
+mu <- mean(results)
+se <- sd(results) / sqrt(25)
+```
+
+$\mu = `r mu`$, $\sigma = `r se`$
+
+b.) Compare your estimate of the standard error with the theoretical standard error (_Corollary 4.3.2_).
+
+```{r, echo = T}
+tse <- mu * ( 1 - mu ) / sqrt(25)
+```
+
+Theoretical: `r tse`
+
+c.) Repeat the above using samples of size 250, and compare with the $n = 25$ case.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+N <- 10e2
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- sample(Recidivism$Recid, 250)
+   results[i] <- mean(samp == "Yes")
+}
+
+ggplot(data.table(value = results)) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs(title = "Recidivism Sampling Distribution")
+
+mu <- mean(results)
+se <- sd(results) / sqrt(250)
+
+```
+
+$\mu = `r mu`$, $\sigma = `r se`$
+
+### 4.7
+
+The data set _FlightDelays_ contains the population of all flight departures by United Airlines and American Airlines out of LGA during May and June 2009 (case study in Section 1.1).
+
+```{r, echo = T}
+Flights <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
+                               header = T))
+```
+
+a.) Create a histogram of _Delay_ and describe the distribution.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+ggplot(Flights, aes(Delay)) +
+   geom_histogram(aes(fill = ..count..), bins = 30) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Flight Delays")
+
+```
+
+Compute the mean and standard deviation.
+
+```{r, echo = T}
+mu <- mean(Flights$Delay)
+sigma <- sd(Flights$Delay)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+b.) Simulate the sampling distribution of $\bar{x}$, the sample mean of the length of the flight delays (_Delay_), for sample size 25.
+
+```{r, echo = T}
+
+N <- 10e2
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- sample(Flights$Delay, 25, replace = F)
+   results[i] <- mean(samp)
+}
+
+```
+
+Create a histogram and describe the simulated sampling distribution of $\bar{x}$.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+ggplot(data.table(value = results)) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs(title = "Flight Delay Sampling Distribution")
+```
+
+Estimate the mean and standard error.
+
+```{r, echo = T}
+mu <- mean(results)
+se <- sd(results) / sqrt(25)
+```
+
+$\mu = `r mu`$, $\Sigma = `r se`$
+
+c.) Compare your estimate of the standard error with the theoretical standard error (_Corollary A.4.1_).
+
+```{r, echo = T}
+tse <- var(results) / 25
+```
+
+Theoretical: `r tse`
+
+d.) Repeat with sample size 250.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+N <- 10e2
+results <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- sample(Flights$Delay, 250, replace = F)
+   results[i] <- mean(samp)
+}
+
+ggplot(data.table(value = results)) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs(title = "Flight Delay Sampling Distribution")
+
+mu <- mean(results)
+se <- sd(results) / sqrt(250)
+tse <- var(results) / 250
+```
+
+$\mu = `r mu`$, $\Sigma = `r se`$
+
+Theoretical: `r tse`
+
+### 4.8
+
+Let $X_1, X_2, \ldots, X_{25}$ be a random sample from some distribution and $W = T(X_1, X_2, \ldots, X_n)$ be a statistic.
+
+Suppose the _sampling distribution_ of Y has pdf $f(y) = (3/8)y^2 \ for \ 0 \le y \le 2$.
+
+Find $P(0 \le Y \le \frac{1}{5})$

### 2019-11-28 Mathematical Statistics/ms_ch_04.rmd
ms, ch4
ms, ch4

@@
 
+Suppose the _sampling distribution_ of W has a pdf given by $f(x) = \frac{2}{x^2}$, for 1 < x < 2.
+
+Find $P(w < 1.5)$
+
+__Solution__:
+
+### 4.9
+
+Let $X_1, X_2, \ldots, X_{n}$ be a random sample from some distribution and $Y = T(X_1, X_2, \ldots, X_n)$ be a statistic.
+
 Suppose the _sampling distribution_ of Y has pdf $f(y) = (3/8)y^2 \ for \ 0 \le y \le 2$.
@@
 Find $P(0 \le Y \le \frac{1}{5})$
+
+__Solution__:
+
+### 4.10
+
+Suppose the heights of boys in a certain large city follow a distribution with mean 48 in. and variance $9^2$.
+
+Use the CLT approximation to estimate the probability that in a random sample of 30 boys, the mean height is more than 51 in.
+
+### 4.11
+
+Let $X_1, X_2, \ldots, X_{36} \sim Bern(.55)$ be independent, and let $\hat{p}$ denote the sample proportion.
+
+Use the CLT approximation with continuity correction to find the probability that $\hat{p} \le 0.5$.
+
+### 4.12
+
+A random sample of size $n = 20$ is drawn from a distribution with mean 6 and variance 10.
+
+Use the CLT approximation to estimate $P(\bar{X} \le 4.6)$.

### 2019-11-29 Mathematical Statistics/ms_ch_04.rmd
mathematical statistics, ch4
mathematical statistics, ch4

@@
 
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+f <- function(x) 2 / x^2
+
+x <- seq( from = 1.01, to = 1.99, by = 0.001)
+
+y <- f(x)
+
+ggplot(data.table(x, y)) +
+   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
+   labs(title = "pdf f(x) = 2/x^2")
+
+a <- cumsum(y) / sum(y)
+p <- round( a[x == 1.5], 4 ) * 100
+
+d <- data.table(x, y = a)
+ggplot(d) +
+   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
+   geom_area(aes(x, y), data = d[x < 1.5], fill = "cornflowerblue", alpha = .3) +
+   labs(title = paste("cdf f(x) = 2/x^2, A =", p ))
+
+```
+
+Numerical solution: `r p`%
+
 ### 4.9
@@
 Use the CLT approximation to estimate $P(\bar{X} \le 4.6)$.
+
+### 4.13
+
+A random sample of size $n = 244$ is drawn from a distribution with pdf $f(x) = (3/16)(x - 4)^2, 2 \le x \le 6$.
+
+Use the CLT approximation to estimate $P(X \ge 4.2)$.
+
+### 4.14
+
+According to the 2000 census, 28.6% of the US adult population recieved a high school diploma.
+
+In a random sample of 800 US adults, what is the probability that between 220 and 230 (inclusive) people have a high school deploma?
+
+Use the CLT approximation with continuity correction, and compare with the exact probility.
+
+```{r, echo = T}
+
+```
+
+### 4.15
+
+If $X_1, \ldots, X_n$ are i.i.d. from Unif[0, 1], how larage should n be so that $P(\bar{X} - \frac{1}{2} < 0.05) \ge 0.90$,
+
+that is, is there at least a 90% chance that the sample mean is within 0.05 of $\frac{1}{2}$? Use the CLT approximation.

### 2019-11-29 Mathematical Statistics/ms_ch_04.rmd
mathematical statistics, ch4
mathematical statistics, ch4

@@
 
-x <- seq( from = 1.01, to = 1.99, by = 0.001)
+x <- seq( from = 1.0001, to = 1.999, by = 0.0001)
 
@@
 
+Analytical Solution: ${\int_{1}^{1.5}}\frac{2}{x^2} = \frac{2}{3}$
+
 ### 4.9
@@
 
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+f <- function(x) (3/8)*x**2
+
+x <- seq( from = 0, to = 2, by = 0.001)
+
+y <- f(x)
+
+ggplot(data.table(x, y)) +
+   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
+   labs(title = "pdf f(x) = (3/8)*x**2")
+
+a <- cumsum(y) / sum(y)
+p <- round( a[x == 1/5], 4 ) * 100
+
+d <- data.table(x, y = a)
+
+ggplot(d) +
+   geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
+   geom_area(aes(x, y), data = d[x < 1/5], fill = "cornflowerblue", alpha = .3) +
+   labs(title = paste("cdf f(x) = 2/x^2, A =", p ))
+
+```
+
+Numerical Solution: `r p`%
+
+Analytical Solution: ${\int_{0}^{\frac{1}{5}}}\frac{x^3}{8} = \frac{.008}{8} = .001 = .1$ %
+
 ### 4.10
@@
 
+```{r, echo = T}
+
+z <- (51 - 48) / (9^2 / sqrt(30))
+p <- pnorm(z, lower.tail = F)
+
+```
+
+Probability: __`r round(p, 4)*100`%__
+
 ### 4.11
@@
 Use the CLT approximation with continuity correction to find the probability that $\hat{p} \le 0.5$.
+
+```{r, echo = T}
+
+z <- ( .5 - .55 ) / sqrt(.55 * (1 - .55) / 36)
+
+p <- pnorm(z, lower.tail = T)
+
+```
+
+Probability: `r round(p, 4)*100`%
 

### 2019-11-29 Mathematical Statistics/ms_ch_04.r

@@ -66,3 +66,14 @@ perm <- permutations( length(p), sample.size, repeats.allowed = T)
 d <- data.table( c1 = p[perm[, 1]], c2 = p[perm[, 2]])
 d[, m := (c1 + c2) / 2]
 
+( 3 - 5/2 ) / (sqrt(5/2^2) / sqrt(30))
+
+pnorm(2.44949, lower.tail = F)
+
+
+z <- (0.53333 - 0.5) / 0.0289
+pbinom(z, prob = .5, size = 300, lower.tail = T)
+
+pbinom(160, size = 300, prob = .5, lower.tail = T)
+
+

### 2019-12-02 Mathematical Statistics/ms_ch_04.rmd
ms, ch4
ms, ch4

@@
    geom_point(aes(x, y), color = "cornflowerblue", size = .6) +
-   labs(title = "pdf f(x) = (3/8)*x**2")
+   labs(title = paste("pdf: ", paste0(deparse(f), collapse = " ")))
 
@@
 
+```{r, echo = T}
+
+z <- ( 4.6 - 6 ) / ( 10 * sqrt(20) )
+
+p <- pnorm(z, lower.tail = T)
+
+```
+
+Probability: `r round(p, 4)*100`%
+
 ### 4.13
@@
 Use the CLT approximation to estimate $P(X \ge 4.2)$.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+pdf <- function(x) (3/16)*(x - 4)^2
+x <- seq(from = 2, to = 6, by = 0.001)
+y <- pdf(x)
+
+ggplot(data.table(x,y)) +
+   geom_point(aes(x, y), col = "cornflowerblue", lwd = .8) +
+   labs(title = paste("PDF: ", paste0(deparse(f), collapse = " ")))
+
+cdf <- function(x) (3/8)*(x - 4)
+
+y <- cumsum(y) / sum(y)
+
+ev <- x[min(which(y > .5))]
+
+ggplot(data.table(x,y)) +
+   geom_point(aes(x, y), col = "cornflowerblue", lwd = .8) +
+   geom_vline(xintercept = ev) +
+   labs(title = paste("CDF: ", paste0(deparse(f), collapse = " ")))
+
+z <- ( 4.2 - ev ) / sqrt(244)
+pnorm(z, lower.tail = F)
+
+```
 
@@
 
-If $X_1, \ldots, X_n$ are i.i.d. from Unif[0, 1], how larage should n be so that $P(\bar{X} - \frac{1}{2} < 0.05) \ge 0.90$,
+If $X_1, \ldots, X_n$ are i.i.d. from Unif[0, 1], how large should n be so that $P(\bar{X} - \frac{1}{2} < 0.05) \ge 0.90$,
 
 that is, is there at least a 90% chance that the sample mean is within 0.05 of $\frac{1}{2}$? Use the CLT approximation.
+
+```{r, echo = T}
+
+```
+
+### 4.16
+
+Maria claims that she has drawn a random sample of size 30 from the exponential distribution with $\lambda = 1/10$.
+
+The mean of her sample is 12.
+
+a.) What is the expected value of a sample mean?
+
+b.) Run a simulation by drawing 1000 random samples, each of size 30, from Exp(1/10), and compute the mean for each sample.
+
+What proportion of the sample means is as large or larger than 12?
+
+c.) Is a mean of 12 unusual for a sample of size 30 from Exp(1/10)?
+
+### 4.17
+
+Let $X \sim N(15, 3^2)$ and $Y \sim N(4, 2^2)$ be independent random variables.
+
+a.) What is the exact sampling distribution of $W = X - 2Y$?
+
+b.) Use R to simulate the sampling distribution of $W$ and plot your results.
+
+Check that the simulated mean and standard error are close to the theoretical mean and standard error.
+
+c.) Use the simulated sampling to estimate $P(W \le 10)$, and then check your estimate with an exact calculation.
+
+### 4.18
+
+Let $X \sim Pois(4)$, $Y \sim Pois(12)$, $U \sim Pois(3)$ be independent random variables.
+
+a.) What is the exact sampling distribution of $W = X + Y + U$?
+
+b.) Use R to simulate the sampling distribution of $W$ and plot your results.
+
+Check that the simulated mean and standard error are close to the theoretical mean and standard error.
+
+c.) Use the simulated sampling distribution to estimate $P(W \le 14)$ and then check your estimate with an exact calculation.
+
+### 4.19
+
+Let $X_1, X_2, \ldots, X_{10} \sim^{i.i.d} N(20, 8^2)$ and $Y_1, Y_2, \ldots, Y_{15} \sim^{i.i.d} N(16, 7^2)$.
+
+Let $W = \bar{X} + \bar{Y}$
+
+a.) Give the exact sampling distribution of W.
+
+b.) Simulate the sampling distribution in R and plot your results. Check that the simulated mean and standard error are close to the exact mean and standard error.
+
+c.) Use your simulation to find $P(W < 40)$. Calculate an exact answer and compare.

### 2019-12-02 Computational Finance/cf_ch_01.rmd
mathematical statistics work
computational finance, ch1 work
computational finance, ch1 work

@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Exploratory Data Analysis}
-   \rfoot{\color{headergrey}Chapter 2}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}An Abridged Introduction to Finance}
+   \rfoot{\color{headergrey}Chapter 1}
    \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Finance}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 
-```{r echo = T, fig.height=3.5}
+```{r echo = T, fig.height=3.5, fig.width=8}
 chartSeries(GE)
@@
 
-```{r echo = T, fig.height=3.5}
+```{r echo = T, fig.height=3.5, fig.width=8}
 chartSeries(GE, TA=NULL, subset='2001-01::2001-02')
@@
 ```
+
+### 1.3.4
+
+```{r, echo = T}
+
+symbols <- c('^VLIC', 'GE', 'KO', 'AAPL', 'MCD')
+getSymbols( symbols, src = "yahoo", from = "2012-02-01", to = "2013-02-01")
+
+# obtain Adjusted Close
+
+VLICad <- VLIC$VLIC.Adjusted; GEad <- GE$GE.Adjusted
+KOad <- KO$KO.Adjusted; AAPLad <- AAPL$AAPL.Adjusted
+MCDad <- MCD$MCD.Adjusted
+
+# compute cumulative sum (cumsum) of daily returns (Delt)
+
+vl <- cumsum( (Delt(VLICad) * 100)[-1, ])
+ge <- cumsum( (Delt(GEad) * 100)[-1, ])
+ko <- cumsum( (Delt(KOad) * 100)[-1, ])
+ap <- cumsum( (Delt(AAPLad) * 100)[-1, ])
+md <- cumsum( (Delt(MCDad) * 100)[-1, ])
+
+### Range for the plot
+
+lim <- c(min(vl, ge, ko, ap, md), max(vl, ge, ko, ap, md))
+
+```
+
+```{r, echo = T, fig.width=8, fig.height=4}
+### plot
+
+stocks <- data.table( Date = index(vl),
+                      VLIC = as.numeric(vl), 
+                      GE = as.numeric(ge), 
+                      KO = as.numeric(ko), 
+                      AAPL = as.numeric(ap), 
+                      MCD = as.numeric(md))
+
+stocks.long <- melt(stocks, 
+     id.vars = c("Date"), 
+     variable.name = c("Stock"), 
+     value.name = "Return")
+
+ggplot(stocks.long, aes(Date, Return, col = Stock)) +
+   geom_line() +
+   labs(title = "Stock Performance")
+
+```
+
+### 1.3.6
+
+Using no arbitrage arguments show that for options on stocks:
+
+i.) the stock's price is an upper bound for the price of a call;
+
+ii.) the strike price is an upper bound for the price of a put.
+
+### 1.3.7
+
+The following is a list of well-known investment strategies obtained by different combinations of put and call options on the same underlying asset.
+For each one of these strategies compute the payoff function and draw the profit graph.
+
+Additionally, argue about the situations where the strategy is profitable.
+
+```{r, echo = T}
+
+plot.payoff <- function(payoff, strike) {
+
+   breakeven <- payoff[min(which(payoff$Payoff > 0))]$Strike
+   ggplot(payoff) +
+      geom_point(aes(Strike, Payoff), col = ifelse(payoff$Payoff < 0, "darkred", "darkgreen"), lwd = .8) +
+      geom_hline(yintercept = 0, col = "cornflowerblue", lwd = 1, alpha = .45) +
+      labs(title = paste0("Option Payoff: K=$", strike, ", Break-Even: $", breakeven), 
+           x = "Strike", y = "Payoff") +
+      scale_x_continuous(labels = dollar) +
+      scale_y_continuous(labels = dollar)
+}
+
+```
+
+#### Vanilla Call Payoff
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+call.payoff <- function( strike, initial.price, contracts = 1, contract.size = 100,
+                         lower = -.15, upper = .15) {
+
+   cost <- -(initial.price * contract.size) * contracts
+
+   x <- strike + seq(strike*lower, strike*upper, .01)
+   y <- ( ( x - strike - initial.price ) * contract.size ) * contracts
+
+   payoff <- ifelse(y < cost, cost, y)
+
+   data.table(Strike = x, Payoff = payoff)
+}
+
+K <- 45
+plot.payoff(call.payoff(K, 2.88), K)
+
+```
+
+#### Vanilla Put Payoff
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+put.payoff <- function( strike, initial.price, contracts = 1, contract.size = 100,
+                         lower = -.15, upper = .15) {
+
+   cost <- -(initial.price * contract.size) * contracts
+
+   x <- strike + seq(strike*lower, strike*upper, .01)
+   y <- ( ( strike - x - initial.price ) * contract.size ) * contracts
+
+   payoff <- ifelse(y < cost, cost, y)
+
+   data.table(Strike = x, Payoff = payoff)
+}
+
+K <- 45
+plot.payoff(put.payoff(K, 2.45), K)
+
+```
+
+#### The Straddle:
+
+The straddle strategy consists of two legs, one long and one short, both with the same strike and expiration date.
+
+The main idea behind the straddle is that it is a non-directional long volatility strategy.
+It is generally suitable when you expect the underlying security to be very volatile and move a lot, but you are not sure whether the price move will be up or down.
+The position makes a profit when your expectation is correct and the underlying does make a big move to one side or the other.
+If the underlying price stays more or less the same, the trade makes a loss.
+
+Straddle Payoff:
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+K <- 45
+plot.payoff(call.payoff(K, 2.85) + put.payoff(K, 2.88), K)
+
+```
+
+#### The strangle:
+
+Similar to the straddle.
+It consists on buying a put and a call with the same expiration date, but different strike prices.
+If $K_c$ is the strike price for the call, and $K_p$ is the strike price for the put, then the strategy requires $K_c > K_p$.
+
+```{r, echo = T}
+
+```
+
+#### The strip:
+
+This strategy consists of long positions in one call and two puts, all with the same strike price and expiration date.
+
+```{r, echo = T}
+
+```
+
+#### The strap:
+
+This one consists of long positions in two calls and one put, all with the same strike price and expiration date.
+
+```{r, echo = T}
+
+```
+
+#### The butterfly spread:
+
+This is made with options of the same type.
+Suppose we use calls and athe underlying assset is a stock.
+Then a butterfly spread of calls consists on short selling two calls waith strike price $K_0$ close to the current stock price, and buying two calls, one with strike price $K_0-c$ and the other with strike price $K_0 + c$, where 2c > 0 is the length of the spread chosen by the investor.
+
+```{r, echo = T}
+
+```

### 2019-12-02 Mathematical Statistics/04_sampling_distributions.rmd
ms refactor
ms refactor

--- 2019-12-02 Mathematical Statistics/ms_ch_04.rmd
@@
 
-```{r, echo = T, fig.height=3.5, fig.width=8}
+```{r, echo = T, fig.height=2.6, fig.width=8}
 pdf <- function(x) (3/16)*(x - 4)^2
@@
 
-Use the CLT approximation with continuity correction, and compare with the exact probility.
+Use the CLT approximation with continuity correction, and compare with the exact probability.
 

### 2019-12-03 Mathematical Statistics/04_sampling_distributions.rmd
mathematical statistics, ch4 work
mathematical statistics, ch4 work

@@
 
+__Solution__:
+
+The sampling distribution of $\hat{p}$ is approximately normal with:
+
 ```{r, echo = T}
 
+n <- 800
+mu <- .286
+
+ev <- 800 * mu
+sigma <- sqrt(n*mu*(1-mu))
+
 ```
+
+$\mathbb{E}[X] = `r ev`$ and $\sigma = \sqrt{800(.286)(1 - .286)} = `r round(sigma, 4)`$
+
+```{r, echo = T}
+
+l <- pnorm((ev - 219.5) / sigma)
+h <- pnorm((ev - 230.5) / sigma)
+
+p <- l - h
+
+```
+
+Probability: `r round(p, 4)`
 
@@
 
+$X \sim Exp(\frac{1}{10}), \mathbb{E}(x) = 10$
+
 b.) Run a simulation by drawing 1000 random samples, each of size 30, from Exp(1/10), and compute the mean for each sample.
 
-What proportion of the sample means is as large or larger than 12?
+```{r, echo = T, fig.height=3.5, fig.width=12}
+
+N <- 1000
+result <- numeric(N)
+
+for( i in 1:N)
+{
+   samp <- rexp( n = 30, rate = 1/10)
+   result[i] <- mean(samp)
+}
+
+ggplot(data.table(result), aes(result)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_vline(xintercept = 12, col = "darkorange") +
+   stat_density( kernel = "gaussian", fill = "darkorange", alpha = .3) +
+   labs(title = "Exp(1/10) Sampling Distribution")
+
+p <- mean(result > 12)
+
+```
+
+What proportion of the sample means is as large or larger than 12? __`r p*100`%__
 
 c.) Is a mean of 12 unusual for a sample of size 30 from Exp(1/10)?
+
+__Yes__, only ~13% of the sample means have a value of 12 or higher.
 
@@
 
+$W \sim N(7, 5^2)$
+
 b.) Use R to simulate the sampling distribution of $W$ and plot your results.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+X <- rnorm(10e3, 15, 3)
+Y <- rnorm(10e3, 4, 2)
+
+W <- X - 2*Y
+
+ggplot(data.table(x = W)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = 7, col = "darkorange", lwd = 1.5) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+
+```
 
@@
 
+```{r, echo = T}
+mu <- mean(W)
+sigma <- sd(W)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
 c.) Use the simulated sampling to estimate $P(W \le 10)$, and then check your estimate with an exact calculation.
+
+```{r, echo = T}
+phat <- mean(W <= 10)
+
+p <- pnorm(10, mean = 7, sd = 5)
+
+```
+
+$\hat{p}= `r phat*100`$%
+
+$P(W \le 10) = `r round(p, 4)*100`$%
 
@@
 
+$W \sim Pois(19)$
+
 b.) Use R to simulate the sampling distribution of $W$ and plot your results.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+W <- rpois(10e3, lambda = 19)
+
+ggplot(data.table(x = W)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = 19, col = "darkorange", lwd = 1.5) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+
+```
 
@@
 
+```{r, echo = T}
+mu <- mean(W)
+sigma <- sd(W)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
 c.) Use the simulated sampling distribution to estimate $P(W \le 14)$ and then check your estimate with an exact calculation.
+
+```{r, echo = T}
+phat <- mean(W <= 14)
+
+p <- ppois(14, lambda = 19)
+
+```
+
+$\hat{p} = `r phat*100`$%
+
+$P(W \le 14) = `r round(p, 4)*100`$%
 
@@
 
-b.) Simulate the sampling distribution in R and plot your results. Check that the simulated mean and standard error are close to the exact mean and standard error.
+$W \sim N(36, 3.1^2)$
+
+b.) Simulate the sampling distribution in R and plot your results.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+N <- 10e3
+result <- numeric(N)
+
+for( i in 1:N)
+{
+   X <- rnorm(10, 20, 8)
+   Y <- rnorm(15, 16, 7)
+
+   result[i] <- mean(X) + mean(Y)
+}
+
+ggplot(data.table(x = result)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = 36, col = "darkorange", lwd = 1.5) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+
+```
+
+Check that the simulated mean and standard error are close to the exact mean and standard error.
+
+```{r, echo = T}
+mu <- mean(result)
+sigma <- sd(result)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
 
 c.) Use your simulation to find $P(W < 40)$. Calculate an exact answer and compare.
+
+```{r, echo = T}
+phat <- mean(result <= 40)
+
+p <- pnorm(40, 36, 3)
+```
+
+$\hat{p} = `r phat*100`$%
+
+$P(W < 40) = `r round(p, 4)*100`$%

### 2019-12-03 Mathematical Statistics/04_sampling_distributions.rmd
mathematical statistics, ch 4
mathematical statistics, ch 4

@@
    \fancyhead[C]{\Large\bfseries\color{headergrey}Sampling Distributions}
-   \rfoot{\color{headergrey}Chapter 4}
-   \lfoot{\color{headergrey}}
+   \rfoot{\color{headergrey}\thepage}
+   \lfoot{\color{headergrey}Chapter 4}
    \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
@@
 
+$\sigma = (10 + 15) / sqrt( 10 + 15 - 1) = 3.1$
+
 $W \sim N(36, 3.1^2)$
@@
 $P(W < 40) = `r round(p, 4)*100`$%
+
+### 4.20
+
+Let $X_1, X_2, \ldots, X_9 \sim^{i.i.d.} N(7, 3^2)$, and $Y_1, Y_2, \ldots, X_{12} \sim^{i.i.d.} N(10, 5^2)$.
+
+Let $W = \bar{X} - \bar{Y}$.
+
+a.) Give the sampling distribution of $W$.
+
+$\sigma = (3 + 5) / sqrt( 9 + 12 - 1 ) = 1.79$
+
+$W = N(-3, 1.79^2)$
+
+b.) Simulate the sampling distribution of W in R, and plot your results.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+N <- 10e3
+result <- numeric(N)
+
+for(i in 1:N)
+{
+   X <- rnorm(9, 7, 3)
+   Y <- rnorm(12, 10, 5)
+   result[i] <- mean(X) - mean(Y)
+}
+
+ggplot(data.table(x = result)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   geom_vline(xintercept = -3, col = "darkorange", lwd = 1.5) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+```
+
+Check that the simulated mean and standard error are close to the theoretical mean and standard error.
+
+```{r, echo = T}
+mu <- mean(result)
+sigma <- sd(result)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+c.) Use your simulation to find $P(W < -1.5)$.
+
+```{r, echo = T}
+phat <- mean(result <= -1.5)
+
+p <- pnorm(-1.5, -3, 1.79)
+```
+
+$\hat{p} = `r phat*100`$%
+
+Calculate an exact answer and compare.
+
+$P(W < 1.5) = `r round(p, 4)*100`$%
+
+### 4.21
+
+Let $X_1, X_2, \ldots, X_N$ be a random sample from $N(0, 1)$. Let $W = X^2_1 + X^2_2 + \ldots + X^2_n$
+
+What is the mean and variance of the sampling distribution of W?
+
+$\mu = 0$, $\sigma = 1$
+
+Repeat using N = 4, N = 5.
+
+$N = 4, \sigma = 4 / sqrt(4 - 1) = 2.3$
+
+$N = 5, \sigma = 5 / sqrt(5 - 1) = 2.5$
+
+What observations or conjectures do you have for general __n__?
+
+### 4.22
+
+Let $X$ be a uniform random variable on the interval $[40, 60]$ and Y be a uniform random variable on $[45, 80]$.
+
+Assume that X and Y are independent.
+
+a.) Compute the expected value and variance of $X + Y$.
+
+$\mu = 112.5$, $Var = 1/24*(140 - 85)^2 = 126.04$
+
+b.) Simulate a sampling distribution of $X + Y$.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+X <- runif(1000, 40, 60)
+Y <- runif(1000, 45, 80)
+
+total <- X + Y
+
+ggplot(data.table(value = total)) +
+   geom_histogram(aes(value, fill = ..count..), bins = 30) +
+   labs("X + Y Sampling Distribution")
+
+```
+
+Describe the sampling distribution of $X + Y$. __Approximately Normal__
+
+Compute the mean and variance of the sampling distribution and compare this with the theoretical mean and variance.
+
+```{r, echo = T}
+mu <- mean(total)
+var <- var(total)
+```
+
+$\mu = `r mu`$, $Var = `r var`$
+
+c.) Suppose the time (in minutes) Andy takes to complete his statistics homework is $Unif[40, 60]$ and the time Adam takes is $Unif[45, 80]$.
+
+Assume they work independently.
+
+One day they announce that their total time to finish an assignment was less than 90 minutes.
+
+How likely is this?
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+p <- punif(90, 85, 140)
+```
+
+Probability: __`r round(p, 4)*100`%__
+
+### 4.23
+
+Let $X_1, X_2, \ldots, X_{20} \sim^{i.i.d.} Exp(2)$. Let $X = \sum^{20}_{i=1}X_i$.
+
+a.) Simulate the sampling distribution of $X$ in R.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+
+N <- 10e3
+result <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- rexp(n = 20, rate = 2)
+   result[i] <- sum(samp)
+}
+
+```
+
+b.) From your simulation, find $\mathbb{E}[X]$ and $Var[X]$.
+
+```{r, echo = T}
+mu <- mean(result)
+var <- var(result)
+```
+
+$\mu = `r mu`$, $Var = `r var`$
+
+c.) From your simulation, find $P( X \le 10)$.
+
+```{r, echo = T}
+phat <- mean(result <= 10)
+```
+
+Probablity: _`r round(phat, 4)*100`%_
+
+### 4.24
+
+Let $X_1, X_2, \ldots, X_{30} \sim^{i.i.d.} Exp(1/3)$ and let $\bar{X}$ denote the sample mean.
+
+a.) Simulate the sampling distribution of $\bar{X}$ in R.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+N <- 10e3
+result <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- rexp(n = 30, rate = 1/3)
+   result[i] <- mean(samp)
+}
+```
+
+b.) Find the mean and standard error of the sampling distribution, and compare with the theoretical results.
+
+```{r, echo = T}
+mu <- mean(result)
+sigma <- sd(result)
+```
+
+Sample: $\mu = `r mu`$, $\sigma = `r sigma`$
+
+Theoretical: $\mu = \frac{\lambda}{1} = 3$, $sigma = 3/sqrt(30 - 1) = .56$
+
+c.) From your simulation, find $P(\bar{X} \le 3.5)$.
+
+```{r, echo = T}
+phat <- mean(result <= 3.5)
+```
+
+$\hat{p} = `r phat`$
+
+d.) Estimate $P(\bar{X} \le 3.5)$ by assuming the CLT approximation holds.
+
+Compare this result with the one in part __(c)__.
+
+```{r, echo = T}
+
+z <- (3.5 - 3) / sigma
+p <- pnorm(z)
+```
+
+$p = `r round(p, 4)`$
+
+### 4.25
+
+Consider the exponential distribution with density $f(x) \frac{1}{20}e^{-x/20}$, with mean and standard deviation of 20.
+
+a.) Calculate the median of this distribution.
+
+$0.5 = \int_{m}^{\inf}f(x)dx = -e^{-x/20}$
+
+$\ldots = M = 20*log(2)$
+
+$\ldots = `r 20*log(2)`$
+
+b.) Using R, draw a random sample of size 50 and graph the histogram.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+samp <- rexp( n = 50, rate = 1/20)
+
+ggplot(data.table(x = samp)) +
+   geom_histogram(aes(x, fill = ..count..)) +
+   labs(title = "RExp(50, 1/20)")
+```
+
+What are the mean and standard deviation of your sample?
+
+```{r, echo = T}
+mu <- mean(samp)
+sigma <- sd(samp)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+c.) Run a simulation to find the (approximate) sampling distribution for the median of sample size 50 from the exponential distribution and describe it.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+N <- 10e3
+result <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- rexp( n = 50, rate = 1/20)
+   result[i] <- median(samp)
+}
+
+ggplot(data.table(x = result)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+```
+
+What is the mean and the standard error of this sampling distribution?
+
+```{r, echo = T}
+mu <- mean(result)
+sigma <- sd(result) / sqrt(50)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+d.) Repeat the above but use sample sizes $n = 100, 500 \ and \ 1,000$.
+
+```{r, echo = T, fig.height=3, fig.width=8}
+N <- 10e3
+result <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- rexp( n = 100, rate = 1/20)
+   result[i] <- median(samp)
+}
+
+ggplot(data.table(x = result)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+```
+
+```{r, echo = T}
+mu <- mean(result)
+sigma <- sd(result) / sqrt(50)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+```{r, echo = T, fig.height=3, fig.width=8}
+N <- 10e3
+result <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- rexp( n = 500, rate = 1/20)
+   result[i] <- median(samp)
+}
+
+ggplot(data.table(x = result)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+```
+
+```{r, echo = T}
+mu <- mean(result)
+sigma <- sd(result) / sqrt(50)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+```{r, echo = T, fig.height=3, fig.width=8}
+N <- 10e3
+result <- numeric(N)
+
+for(i in 1:N)
+{
+   samp <- rexp( n = 1000, rate = 1/20)
+   result[i] <- median(samp)
+}
+
+ggplot(data.table(x = result)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   scale_y_continuous(labels = comma) +
+   labs(title = "Sampling Distribution")
+```
+
+```{r, echo = T}
+mu <- mean(result)
+sigma <- sd(result) / sqrt(50)
+```
+
+$\mu = `r mu`$, $\sigma = `r sigma`$
+
+How does sample size affect the sampling distribution?
+
+_The sample mean and standard error converge to the analytical solution with increased sample size._
+
+### 4.26
+
+Prove theorem 4.2.1.
+
+### 4.27
+
+Let $X_1, X_2 \sim^{i.i.d.} F$ with corresponding pdf $f(x) = \frac{2}{x^2}$, $1 \le x \le 2$.
+
+a.) Find the pdf of $X_{max}$.
+
+$F_{max}(x) = 8(\frac{1}{x^2} - \frac{1}{x^3})$
+
+b.) Find the expected value of $X_{max}$.
+
+$\mathbb{E}{[F_{max}]} = 1.545$

### 2019-12-03 Mathematical Statistics/04_sampling_distributions.rmd
mathematical statistics, ch 4 work
mathematical statistics, ch 4 work

@@
 $\mathbb{E}{[F_{max}]} = 1.545$
+
+### 4.28
+
+Let $X_1, X_2, \ldots, X_N \sim^{i.i.d.}$ with corresponding pdf $f(x) = 3x^2, 0 \le x \le 1$.
+
+a.) Find the pdf for $X_{min}$.
+
+b.) Find the pdf for $X_{max}$.
+
+c.) If $n = 10$, find the probability that the largest value, $X_{max}$, is greater than 0.92.
+
+### 4.29
+
+Compute the pdf of the sampling distribution of the maximum samples of size 10 from a population with an exponential distribtuion with $\lambda = 12$.
+
+### 4.30
+
+Let $X_1, X_2, \ldots, X_N \sim^{i.i.d.} Exp(\lambda)$ with pdf $f(x) = \lambda e^{-\lambda x}, \lambda > 0, x > 0$.
+
+a.) Find the pdf $f_{min}(x)$ for the sample minimum $X_{min}$. Recognize this as the pdf of a known distribution.
+
+b.) Simulate in R the sampling distribution of $X_{min}$ of samples of size $n = 25$ from the exponential distribution with $\lambda = 7$.
+
+Compare the theoretical expected value of $X_{min}$ with the simulated expected value.
+
+### 4.31
+
+Let $X_1, X_2, \ldots, X_n \sim^{i.i.d.} Pois(3)$. Let $X = \sum^{10}_{i=1}X_i$.
+
+Find the pdf for the sampling distribution of X.
+
+### 4.32
+
+Let $X_1$ and $X_2$ be independent exponential random variables, both with parameter $\lambda > 0$.
+
+Find the cumulative distribution function for the sampling distribution of $X = X_1 + X_2$.
+
+### 4.33
+
+This simulation illustrates the CLT for a finite population.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+N <- 400
+n <- 5
+
+finpop <- rexp(N, 1/10)
+
+ggplot(data.table(x = finpop)) +
+   geom_histogram(aes(x, fill = ..count..)) +
+   labs(title = "Exp(1/10)")
+
+mean(finpop) #mean (mu) of your pop.
+sd(finpop) # stdev (sigma) of your pop.
+sd(finpop)/sqrt(n) # theoretical standard error of sampling distribution
+
+sd(finpop)/sqrt(n) * sqrt((N-n)/(N-1)) # without replacement
+
+Xbar <- numeric(1000)
+for(i in 1:1000)
+{
+   x <- sample(finpop, n) # Random sample of size n
+                          # (w/o replacement)
+   Xbar[i] <- mean(x)
+}
+
+p1 <- ggplot(data.table(x = Xbar)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   labs(title = "Mean Sampling Distribution Histogram")
+
+p2 <- ggplot(data.table(x = Xbar), aes(sample = x)) +
+   stat_qq() +
+   stat_qq_line() +
+   labs(title = "Mean Sampling Distribution QQ-Plot")
+
+grid.arrange(p1, p2, nrow = 2)
+
+mean(Xbar)
+sd(Xbar) # estimated standard error of sampling distribution
+
+```
+
+a.) Does the sampling distribution of sample means appear approximately normal?
+
+b.) Compare the mean and standard error of your simulated sampling distribution with the theoretical ones.
+
+c.) Calculate $(\sigma/\sqrt(n))(\sqrt{(N-n)/(N-1)}$, where $\sigma$ is the standard deviation of the finite population and compare with the (estimated) standard error of the sampling distribution.
+
+d.) Repeat for larger __n__, say __n__ = 20 and __n__ = 100.
+
+### 4.34
+
+Let $X_1, X_2, \ldots, X_n$ be independent random variables from $N(\mu, \sigma)$.
+
+We are interested in the sampling distribution of the variance.
+
+Run a simulation to draw random samples of size 20 from $N(25, 7^2)$ and calculate the variance for each sample.
+
+```{r, echo = T, fig.height=3.5, fig.width=8}
+W <- numeric(1000)
+
+for(i in 1:1000)
+{
+   x <- rnorm(20, 25, 7)
+   W[i] <- var(x)
+}
+
+mean(W)
+var(W)
+
+p1 <- ggplot(data.table(x = W)) +
+   geom_histogram(aes(x, fill = ..count..), bins = 30) +
+   labs(title = "Variance Sampling Distribution")
+
+p2 <-
+   ggplot(data.table(value = W), aes(sample = value)) +
+   stat_qq() +
+   stat_qq_line() +
+   labs(title = "Mean Sampling Distribution QQ-Plot")
+
+grid.arrange(p1, p2, nrow = 2)
+
+```
+
+Does the sampling distribution appear to be normally distributed?
+
+Repeat with n = 50 and n = 200.
+
+### 4.35
+
+A random sample of size $n = 100$ is drawn from a distribution with pdf $f(x) = 3(1- x)^2, 0 \le x \le 1$.
+
+a.) Use the CLT approximation to estimate $P(\bar{X} \le 0.27)$.
+
+b.) Use the expanded CLT to estimate the same probability (dnorm).
+
+c.) If $X_1, X_2, X_3 \sim^{i.i.d.} Unif[0, 1]$, then the minimum has density __f__ given above.
+
+Use simulation to estimate the probability.

### 2019-12-05 Mathematical Statistics/04_sampling_distributions.r
mathematical statistics, ch 4/5
mathematical statistics, ch 4/5

@@ -1,3 +1,6 @@
+library(data.table)
+library(ggplot2)
+
 N <- 1000
 Xbar <- numeric(N) # space for results
 

### 2019-12-05 Mathematical Statistics/05_conf_int_bootstrap.r

@@ -1,0 +1,170 @@
+library(data.table)
+library(ggplot2)
+
+data.dir <- "D:/Projects/Statistical-Computing/datasets/"
+
+### Bangladesh
+
+Bangladesh <- data.table(read.csv(paste0(data.dir, "Bangladesh.csv"),
+                                 header = T))
+
+Arsenic <- Bangladesh$Arsenic
+
+ggplot(data.table(value = Arsenic)) +
+  geom_histogram(aes(value, fill = ..count..), bins = 30) +
+  scale_x_continuous(labels = comma) +
+  labs(title = "Arsenic")
+
+ggplot(data.table(value = Arsenic), aes(sample = value)) +
+  stat_qq() +
+  stat_qq_line() +
+  labs(title = "Arsenic")
+
+n <- length(Arsenic)
+N <- 10e3
+
+arsenic.mean <- numeric(N)
+
+for(i in 1:N)
+{
+  x <- sample(Arsenic, n, replace = T)
+  arsenic.mean[i] <- mean(x)
+}
+
+ggplot(data.table(value = arsenic.mean)) +
+  geom_histogram(aes(value, fill = ..count..), bins = 30) +
+  geom_vline(xintercept = mean(arsenic.mean), col = "darkorange", alpha = .6, lwd = 2) +
+  scale_y_continuous(labels = comma) +
+  labs(title = "Arsenic Bootstrap Distribution of Means")
+
+ggplot(data.table(value = arsenic.mean), aes(sample = value)) +
+  stat_qq() +
+  stat_qq_line() +
+  labs(title = "Arsenic")
+
+mean(arsenic.mean) # Bootstrap mean
+mean(arsenic.mean) - mean(Arsenic) # bias
+sd(arsenic.mean) # bootstrap SE
+
+uq <- mean(arsenic.mean) - 1.96 * sd(arsenic.mean)
+lq <- mean(arsenic.mean) + 1.96 * sd(arsenic.mean)
+
+sum(arsenic.mean > lq) / N
+sum(arsenic.mean < uq) / N
+
+quantile(arsenic.mean, c(0.025, 0.975)) # confidence intervals
+
+### testosterone study
+
+Skateboard <- data.table(read.csv(paste0(data.dir, "Skateboard.csv"),
+                                  header = T))
+
+testF <- Skateboard[Experimenter == "Female"]$Testosterone
+testM <- Skateboard[Experimenter == "Male"]$Testosterone
+
[…]

### 2019-12-07 Mathematical Statistics/05_conf_int_bootstrap.rmd
mathematical statistics, ch 4/5
mathematical statistics, ch 4/5
Update 05_conf_int_bootstrap.rmd

--- 2019-11-23 Mathematical Statatistics/ms_ch_03.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Introduction to Hypothesis - Testing Permutation Tests}
-   \rfoot{\color{headergrey}Chapter 3}
-   \lfoot{\color{headergrey}}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Confidence Intervals - The Bootstrap}
+   \rfoot{\color{headergrey}\thepage}
+   \lfoot{\color{headergrey}Chapter 5}
    \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
@@
 
-#### Section
+#### 5.1
 
-### Subsection
+Consider the samples 1-6. Use a six-sided die to obtain three different bootstrap samples and their corresponding means.
+
+```{r,echo=T}
+
+```
+
+#### 5.2
+
+Consider the samples 1, 3, 4, and 6 from some distribution.
+
+a.) For one random bootstrap sample, find the probability that the mean is one.
+
+b.) For one random bootstrap sample, find the probability that the maximum is 6.
+
+c.) For one random bootstrap sample, find the probability that exactly two elements in the sample are less than 2.
+
+### 5.3
+
+Consider the sample 1-3.
+
+a.) List all the (ordered) bootstrap samples from this sample. How many are there?
+
+b.) How many unordered bootstrap samples are there? For example, {1, 2, 2} and {2, 1, 2} are considered to be the same.
+
+c.) How many ordered bootstrap samples have one occurrence of 1 and two occurences of 3?
+
+Is this the same number of bootstrap samples that have each of 1, 2 and 3 occuring exactly once?
+
+d.) Is the probability of obtaining a bootstrap sample with one 1 and two 3's the same as the probabiliity of obtaining a bootstrap sample with each of 1, 2 and 3 occuring exactly once?

### 2019-12-07 Statistical Methods/01_basics.rmd
statistical methods, ch1
statistical methods, ch1

--- 2019-11-23 template.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey} - Title}
-   \rfoot{\color{headergrey} -Section}
-   \lfoot{\color{headergrey}}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Methods}
+   \rfoot{\color{headergrey}\thepage}
+   \lfoot{\color{headergrey}Chapter 1}
    \fancyfoot[C]{\rmfamily\color{headergrey} -LF}
@@
 library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
-library(png, quietly = TRUE, warn.conflicts = FALSE)
-library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
-library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
-library(stringr, quietly = TRUE, warn.conflicts = FALSE)
-library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
-library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
-library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
-library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
-library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
-library(MASS, quietly = TRUE, warn.conflicts = FALSE)
-library(copula, quietly = TRUE, warn.conflicts = FALSE)
-library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
-library(forecast, quietly = TRUE, warn.conflicts = FALSE)
-library(tseries, quietly = TRUE, warn.conflicts = FALSE)
-library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
-library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
-library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-#### Section
+#### Chapter 1
 
-### Subsection
+### 1.1
+
+Store the values _-20, -15, -5, 8, 12, 9, 2, 23, 19_ in the R variable _x_ and verify the sum is 33.
+
+```{r, echo = T}
+x <- c(-20, -15, -5, 8, 12, 9, 2, 23, 19)
+
+sum(x)
+```
+
+### 1.2
+
+For the data in E1, verify the average is 3.67.
+
+```{r, echo = T}
+mean(x)
+```
+
+### 1.3
+
+What commands can be used to compute an average without using the R command _mean_?
+
+```{r, echo = T}
+n <- length(x)
+
+sum(x) / n
+
+```
+
+### 1.4
+
+Sum the positive values:
+
+```{r, echo = T}
+sum( x[x >= 0] )
+```
+
+### 1.5
+
+Use _which_ command to get the average of the values ignoring the largest.
+
+```{r, echo = T}
+
+mean( which( x != max(x) ) )
+
+```
+
+### 1.6
+
+Speculate about the values corresponding to _x[ abs(x) >= 8 & x < 8 ]_.
+
+[T, T, F, T, T, F, T, T]
+&
+[T, T, T, F, F, F, F, F]
+=
+[T, T, F, F, F, F, F, F]
+
+Verify.
+
+```{r, echo = T}
+x[ abs(x) >= 8 & x < 8]
+```
+
+### 1.7
+
+Your recorded time to commute for 10 days is:
+
+23, 18, 29, 22, 24, 27, 28, 19, 28, 23
+
+```{r, echo = T}
+commute <- c(23, 18, 29, 22, 24, 27, 28, 19, 28, 23)
+
+mean(commute)
+min(commute)
+max(commute)
+
+```
+
+### 1.8
+
+Verify:
+
+```{r, echo = T}
+y <- c(2, 4, 8)
+z <- c(1, 5, 2)
+2*y
+```
+
+returns the values _4, 8, 16_.
+
+### 1.9
+
+Let _x = c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5)_.
+
+```{r, echo = T}
+x <- c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5)
+
+n <- length(x)
+
+stopifnot( (sum(x) / n) == mean(x))
+
+```
+
+### 1.10
+
+For the values in _E9_, use R to subtract the average from each value, then sum the results.
+
+```{r, echo = T}
+xbar <- mean(x)
+sum(x - xbar)
+```
+
+### 1.11
+
+Imagine a matrix having 100 rows and 2 columns. Further imagine that some of the values in the first column are _NA_.
+
+Use _is.na_ to remove missing.
+
+```{r, echo = T}
+
+values <- matrix(nrow = 100, ncol = 2)
+rval <- runif(100)
+values[, 1] <- ifelse(rval <= .25, NA, rval)
+
+values[!is.na(values[,1]),]
+
+```
+
+### 1.12
+
+R has a built-in data set ChickWeight. Verify the mean.
+
+```{r, echo = T}
+mean(ChickWeight[, 1])
+
+mean(ChickWeight[, 3])
+
+mean(as.numeric(ChickWeight[, 3]))
+
+```
+
+### 1.13
+
+Create a matrix with two rows and five columns with some of the entries stored as NA.
+
+```{r, echo = T}
+
+rval <- runif(10, 0, 1)
+values <- matrix(ifelse(rval <= .25, NA, rval), nrow = 2)
+
+```
+
+### 1.14
+
+Use R to compute the average weight among chicks that were fed horsebean.
+
+```{r, echo = T}
+mean(chickwts[chickwts$feed == "horsebean",]$weight)
+```
+
+### 1.15
+
+Let _x = 1, 8, 2, 6, 3, 8, 5, 5, 5, 5_
+
+Sum values not equal to 2 or 3, two ways.
+
+```{r, echo = T}
+x <- c(1, 8, 2, 6, 3, 8, 5, 5, 5, 5)
+
+sum( x[!x %in% c(2, 3)] )
+
+sum( x[x != 2 & x != 3] )
+
+```
+
+### 1.16
+
+For the values used in the previous exercise, use two different R commands to sum all the values != 5.
+
+```{r, echo = T}
+sum( x[ x != 5] )
+sum( x[ !x %in% c(5) ])
+```
+
+### 1.17
+
+Use a single command to change all values equal to 8 to 7.
+
+```{r, echo = T}
+x[x == 8] <- 7
+x  
+```
+
+### 1.18
+
+Create a matrix with four rows and two columns with the values: _1, 2, 3, 4_ in the first column, and _5, 6, 7, 8_ in the second column.
+
+```{r, echo = T}
+matrix(seq(1, 8, by = 1), nrow = 4, ncol = 2)
+```
+
+### 1.19
+
+Create a matrix with four columns and two rows with the values _1, 2, 3, 4_ in the first row and _11, 12, 13, 14_ in the second row.
+
+```{r, echo = T}
+matrix(c(seq(1,4), seq(11, 14)), ncol = 4, byrow = T)
+```

### 2019-12-07 Statistical Methods/_packages.R

@@ -1,0 +1,1 @@
+install.packages("Rallfun")

### 2019-12-07 Statistical Methods/01_basics.rmd
statistical methods, ch2
statistical methods, ch2

@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Methods}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
    \rfoot{\color{headergrey}\thepage}
    \lfoot{\color{headergrey}Chapter 1}
-   \fancyfoot[C]{\rmfamily\color{headergrey} -LF}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 
-#### Chapter 1
+#### Chapter 2
 

### 2019-12-07 Statistical Methods/02_numerical-summaries.rmd
statistical methods, ch2
statistical methods, ch2

--- 2019-11-23 template.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey} - Title}
-   \rfoot{\color{headergrey} -Section}
-   \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey} -LF}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Numerical Summaries}
+   \rfoot{\color{headergrey}\thepage}
+   \lfoot{\color{headergrey}Chapter 2}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 
-#### Section
+#### Chapter 2
 
-### Subsection
+### 2.1
+
+```{r, echo = T}
+
+```

### 2019-12-08 Mathematical Statistics/03_permutation_testing.rmd
statistical methods, ch2
statistical methods, ch2

--- 2019-11-26 Mathematical Statistics/ms_ch_03.rmd
@@
 library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
-library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
-library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
-library(MASS, quietly = TRUE, warn.conflicts = FALSE)
-library(copula, quietly = TRUE, warn.conflicts = FALSE)
-library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
-library(forecast, quietly = TRUE, warn.conflicts = FALSE)
-library(tseries, quietly = TRUE, warn.conflicts = FALSE)
 library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
-library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
-library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
 library(gtools, quietly = TRUE, warn.conflicts = FALSE)

### 2019-12-08 Mathematical Statistics/05_conf_int_bootstrap.rmd
statistical methods, ch2
statistical methods, ch2

@@
 library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
-library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
-library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
-library(MASS, quietly = TRUE, warn.conflicts = FALSE)
-library(copula, quietly = TRUE, warn.conflicts = FALSE)
-library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
-library(forecast, quietly = TRUE, warn.conflicts = FALSE)
-library(tseries, quietly = TRUE, warn.conflicts = FALSE)
-library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
-library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
-library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
+library(gtools, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 ```{r,echo=T}
+pop <- seq(from = 1, to = 6, by = 1)
+
+n <- 6
+
+s1 <- mean( sample(pop, n, replace = T) )
+s2 <- mean( sample(pop, n, replace = T) )
+s3 <- mean( sample(pop, n, replace = T) )
 
 ```
+
+$\bar{x}^*_1 = `r s1`$, $\bar{x}^*_2 = `r s2`$, $\bar{x}^*_3 = `r s3`$
 
@@
 
+```{r, echo = T}
+
+pop <- c(1, 3, 4, 6)
+
+samples <- permutations(n = 4, r = 4, pop, repeats.allowed = T)
+
+```
+
 a.) For one random bootstrap sample, find the probability that the mean is one.
+
+```{r, echo = T}
+
+means <- apply(samples, 1, mean)
+
+p <- mean( means == 1 )
+
+```
+
+Probability: __`r round(p, 4) *100`%__
 
@@
 
+```{r, echo = T}
+maxes <- apply(samples, 1, max)
+
+p <- mean( maxes == 6 )
+```
+
+Probability: __`r round(p, 4) *100`%__
+
 c.) For one random bootstrap sample, find the probability that exactly two elements in the sample are less than 2.
+
+```{r, echo = T}
+
+lt2 <- apply(t(apply(samples, 1, function(x) { x < 2})), 1, sum)
+
+p <- mean( lt2 == 2 )
+
+```
+
+Probability: __`r round(p, 4) *100`%__
 
@@
 
+```{r, echo = T}
+samples <- permutations(n = 3, r = 3, 1:3, repeats.allowed = T)
+
+n <- nrow(samples)
+```
+
+Samples: = $3^3$ = __`r n`__
+
 b.) How many unordered bootstrap samples are there? For example, {1, 2, 2} and {2, 1, 2} are considered to be the same.
+
+```{r, echo = T}
+samples <- combinations(n = 3, r = 3, 1:3, repeats.allowed = T)
+
+n <- nrow(samples)
+
+choose(3^3, 3)
+
+```
+
+Samples: = $3^3$ = __`r n`__
 
@@
 
+```{r, echo = T}
+
+samples <- permutations(n = 3, r = 3, 1:3, repeats.allowed = T)
+
+n.ones <- apply(t(apply(samples, 1, FUN = function(x) { x == 1 })), 1, function(x) sum(x) )
+n.threes <- apply(t(apply(samples, 1, FUN = function(x) { x == 3 })), 1, function(x) sum(x) )
+
+sum((n.ones == 1 & n.threes == 2) == T)
+```
+
 Is this the same number of bootstrap samples that have each of 1, 2 and 3 occuring exactly once?
 
+```{r, echo = T}
+
+n.ones <- apply(t(apply(samples, 1, FUN = function(x) { x == 1 })), 1, function(x) sum(x) )
+n.twos <- apply(t(apply(samples, 1, FUN = function(x) { x == 2 })), 1, function(x) sum(x) )
+n.threes <- apply(t(apply(samples, 1, FUN = function(x) { x == 3 })), 1, function(x) sum(x) )
+
+sum((n.ones == 1 & n.twos == 1 & n.threes == 1) == T)
+```
+
+No, 3 != 6.
+
 d.) Is the probability of obtaining a bootstrap sample with one 1 and two 3's the same as the probabiliity of obtaining a bootstrap sample with each of 1, 2 and 3 occuring exactly once?
+
+```{r, echo = T}
+
+( sum((n.ones == 1 & n.threes == 2)) / n ) == ( sum((n.ones == 1 & n.twos == 1 &  n.threes == 1)) / n )
+
+```
+
+No, 3% and 6% chances respectfully.

### 2019-12-08 Statistical Methods/02_numerical-summaries.rmd
statistical methods, ch2
statistical methods, ch2

@@
 
+Suppose,
+
+$X_1 = 1, X_2 = 3, X_3 = 0,$
+$X_4 = -2, X_5 = 4, X_6 = -1,$
+$X_7 = 5, X_8 = 2, X_9 = 10.$
+
+```{r, echo = T}
+x <- c(1, 3, 0, -2, 4, -1, 5, 2, 10)
+```
+
+Find:
+
+a.) $\sum{X_i}$
+
+```{r, echo = T}
+sum(x)
+```
+
+b.) $\sum^5_{i=3}X_i$
+
+```{r, echo = T}
+sum( x[3:5] )
+```
+
+c.) $\sum^4_{i=1}X^3_i$
+
+```{r, echo = T}
+sum(x[1:4]^3)
+```
+
+d.) $(\sum{X_i})^2$
+
+```{r, echo = T}
+(sum(x))^2
+```
+
+e.) $\sum{3}$
+
+```{r, echo = T}
+3 * length(x)
+```
+
+f.) $\sum{(X_i - 7)}$
+
+```{r, echo = T}
+sum(x - 7)
+```
+
+g.) $3\sum^5_{i=1}X_i - \sum^9_{i=6}X_i$
+
+```{r, echo = T}
+3 * sum( x[1:5] ) - sum( x[6:9] )
+```
+
+h.) $\sum{10X}$
+
+```{r, echo = T}
+sum( 10 * x)
+```
+
+i.) $\sum^6_{i=2}iX_i$
+
+```{r, echo = T}
+i <- 2:6
+sum( i * x[i]  )
+```
+
+j.) $\sum6$
+
+```{r, echo = T}
+6 * length(x)
+```
+
+### 2.2
+
+Express the following in summation notation.
+
+a.) $X_1 + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$
+
+$\ldots = \frac{X_1}{1} + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$
+
+$\ldots = \sum{^4}_{i=1}{\frac{X_i}{i}}$
+
+b.) $U_1 + U^2_2 + U^3_3 + U^4_4$
+
+$\ldots = U^1_1 + U^2_2 + U^3_3 + U^4_4$
+
+$\ldots = \sum^4_{i=1}{U^i_i}$
+
+c.) $(Y_1 + Y_2 + Y_3)^4$
+
+$(\sum^3_{i=1}{Y_i})^4$
+
+### 2.3
+
+Show by numerical example that $\sum{X^2_i}$ is not necessarily equal to $(\sum{X_i}^2)$.
+
 ```{r, echo = T}
@@
 ```
+
+### 2.4
+
+Find the mean and median of the following sets of numbers.
+
+a.) -1, 0, 3, 0, 2, -5.
+
+b.) 2, 2, 3, 10, 100, 1,000
+
+### 2.5
+
+The final exam scores for 15 students are: 73, 74, 92, 98, 100, 72, 74, 85, 76, 94, 89, 73, 76, 99.
+
+Compute the mean, 20% trimmed mean, and median using R.
+
+```{r, echo = T}
+```
+
+### 2.6
+
+The average of 23 numbers is 14.7. What is the sum of these numbers?
+
+### 2.7
+
+Consider the 10 values: 3, 6, 8, 12, 23, 26, 37, 42, 49, 63.
+
+The mean is $\bar{X} = 26.9$
+
+a.) What is the value of the mean if the largest value, 63, is increased to 100?
+
+b.) What is the mean if 633 is increased to 1,000?
+
+c.) What do these results illustrate about the mean?
+
+### 2.8
+
+Repeat the previous exercise, only compute the median instead.
+
+### 2.9
+
+In general, how many values must be altered to make the sample mean arbitrarily large?
+
+### 2.10
+
+What is the minimum number of values that must be altered to make the 20% trimmed mean and sample median arbitrarily large?

### 2019-12-08 Mathematical Statistics/03_permutation_testing.r

@@ -2,10 +2,12 @@ library(data.table)
 library(ggplot2)
 library(gtools)
 
+data.dir <- "D:/Project/Statistical-Computing/datasets/"
+
 ### Beer/Wings Consumption
 
 Beerwings <- data.table(read.csv(paste0(data.dir, "Beerwings.csv"),
-                               header = T))
+                                 header = T))
 
 with(Beerwings, {
   tapply(Hotwings, Gender, mean)

### 2019-12-08 Mathematical Statistics/05_conf_int_bootstrap.r

@@ -122,7 +122,7 @@ ggplot(data.table(value = result), aes(sample = value)) +
 ### Verizon data
 
 Verizon <- data.table(read.csv(paste0(data.dir, "Verizon.csv"),
-                              header = T))
+                               header = T))
 
 Time.ILEC <- Verizon[Group == "ILEC"]$Time
 Time.CLEC <- Verizon[Group == "CLEC"]$Time

### 2019-12-08 Statistical Methods/_packages.R

@@ -1,1 +1,1 @@
-install.packages("Rallfun")
+install.packages("WRS")

### 2019-12-10 Statistical Methods/02_numerical-summaries.rmd
statistical methods, ch2
statistical methods, ch2

@@
 ```{r, echo = T}
+x <- 1:10
 
+sum(x^2)
+
+(sum(x))^2
 ```
@@
 
+```{r, echo = T}
+x <- c(-1, 0, 3, 0, 2, -5)
+
+mean(x)
+median(x)
+```
+
 b.) 2, 2, 3, 10, 100, 1,000
+
+```{r, echo = T}
+x <- c(2, 2, 3, 10, 100, 1000)
+
+mean(x)
+median(x)
+```
 
@@
 ```{r, echo = T}
+scores <- c(73, 74, 92, 98, 100, 72, 74, 85, 76, 94, 89, 73, 76, 99)
+
+mean(scores)
+mean(scores, trim = .2)
+median(scores)
 ```
@@
 The average of 23 numbers is 14.7. What is the sum of these numbers?
+
+```{r, echo = T}
+23 * 14.7
+```
 
@@
 
+```{r, echo = T}
+x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 63)
+mean(x)
+```
+
 a.) What is the value of the mean if the largest value, 63, is increased to 100?
+
+```{r, echo=T}
+x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 100)
+mean(x)
+```
 
@@
 
+```{r, echo=T}
+x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 1000)
+mean(x)
+```
+
 c.) What do these results illustrate about the mean?
+
+_The mean is very sensitive to outliers._
 
@@
 
+```{r, echo=T}
+x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 100)
+median(x)
+```
+
+b.) What is the mean if 633 is increased to 1,000?
+
+```{r, echo=T}
+x <- c(3, 6, 8, 12, 23, 26, 37, 42, 49, 1000)
+median(x)
+```
+
 ### 2.9
@@
 
+_One._
+
 ### 2.10
@@
 What is the minimum number of values that must be altered to make the 20% trimmed mean and sample median arbitrarily large?
+
+mean = $g = (.2N), g + 1$
+
+median = ~$.5N$
+
+### 2.11
+
+For the values 0, 23, -1, 12, -10, -7, 1, -19, -6, 12, 1, -3 compute the lower and upper quartiles using the ideal fourths.
+
+```{r, echo = T}
+# Used in 11 & 12
+idealf <- function( x ) {
+   n <- length(x)
+   sorted <- sort(x)
+   i <- (n / 4) + 5/12
+   j <- floor(i)
+   h <- i - j
+   q1 <- (1 - h)*sorted[j] + h*sorted[j + 1]
+
+   k <- n - j + 1
+   q2 <- (1 - h)*sorted[k] + h*sorted[k - 1]
+
+   c(x[1], q1, q2, sorted[n])
+}
+
+```
+
+```{r, echo = T}
+x <- c(0, 23, -1, 12, -10, -7, 1, -19, -6, 12, 1, -3)
+
+idealf(x)
+```
+
+### 2.12
+
+For the values: -1, -10, 2, 2, -7, -2, 3, 3, -6, 12, -1, -12, -6, 8, 6 compute the lower and upper quartiles (the ideal fourths).
+
+```{r, echo = T}
+x <- c(-1, -10, 2, 2, -7, -2, 3, 3, -6, 12, -1, -12, -6, 8, 6)
+
+idealf(x)
+```
+
+### 2.13
+
+Approximately how many values must be altered to make _q2_, the estimate of the upper quartile based on the ideal fourths, arbitrarily large?
+
+_About $\frac{1}{4}$_
+
+### 2.14
+
+Argue that the smallest observed value, $X_1$, satisfies the definition of a measure of location.
+
+_The smallest (or largest) value in a set defines the boundaries of the set. This is by definition a measure of location._
+
+### 2.15
+
+The height of 10 plants is measured in inches and found to be 12, 6, 15, 3, 12, 6, 21, 15, 18 and 12.
+
+```{r, echo = T}
+
+x <- c(12, 6, 15, 3, 12, 6, 21, 15, 18, 12)
+
+xbar <- mean(x)
+
+stopifnot( sum( x - xbar ) == 0 )
+
+```
+
+Verify that $\sum(X_i - \bar{X}) = 0$
+
+### 2.16
+
+For the data in the previous exercise, compute the range, variance and standard deviation.
+
+```{r, echo = T}
+
+n <- length(x)
+
+range <- max(x) - min(x)
+variance <- (1/(n-1))*sum( ( x - xbar )^2)
+stdDev <- sqrt(variance)
+
+```
+
+$\bar{X} = `r xbar`$, Range = `r range`, $Var = `r variance`$, $\sigma = `r stdDev`$
+
+\newpage
+
+### 2.17
+
+Use the rules of summation notation to show that it is always the case that $\sum(X_i - \bar{X}) = 0$.
+
+$\bar{X} = \frac{1}{n}\sum{X_i}$
+
+$\sum{X_i} - \frac{1}{n}\sum{X_i}$
+
+$\ldots = \frac{1}{n}\sum{X_i} - \sum{X_i}$
+
+$\ldots = \frac{1}{n}(\sum{X_i - X_i})$
+
+$\ldots = \frac{1}{n}(0)$
+
+$\ldots = 0$
+
+### 2.18
+
+Seven different thermometers were used to measure the temperature of a substance. The reading in degrees Celsius are -4.10, -4.13, -5.09, -4.08, -4.10, -4.09 and -4.12.
+
+```{r, echo = T}
+x <- c(-4.10, -4.13, -5.09, -4.08, -4.10, -4.09, -4.12)
+n <- length(x)
+xbar <- 1/n*sum(x)
+variance <- (1/(n-1))*sum( (x - xbar)^2 )
+stdDev <- sqrt(variance)
+```
+
+Find the variance and standard deviation.
+
+$Var = `r variance`$, $\sigma = `r stdDev`$
+
+### 2.19
+
+A weightlifter's maximum bench press (in pounds) in each of 6 successive weaks was 280, 295, 275, 305, 300, 290.
+
+Find the standard deviation.
+
+```{r, echo = T}
+x <- c(280, 295, 275, 305, 300, 290)
+n <- length(x)
+xbar <- 1/n*(sum(x))
+variance <- (1/(n-1))*sum( (x - xbar)^2)
+sqrt(variance)
+```
+
+### 2.20
+
+For the values,
+
+20, 121, 132, 123, 145, 151, 119, 133, 134, 130, 200
+
+use the classic outlier detection rule to determine whether any outliers exist.
+
+```{r, echo = T}
+x <- c(20, 121, 132, 123, 145, 151, 119, 133, 134, 130, 200)
+
+x[ abs(x - mean(x)) / sd(x) > 2 ]
+
+```
+
+### 2.21
+
+Apply the boxplot rule and the MAD-median rule using the values in the preceding exercise. Note that the results differ, compared to using the classic rule.
+
+```{r, echo = T}
+
+x[ abs( x - median(x) ) / mad(x) > 2.27 ]
+
+```
+
+Explain why this happened.
+
+_The classic method masks the value 200 as an outlier._
+
+### 2.22
+
+Consider the values,
+
+0, 121, 132, 123, 145, 151, 119, 133, 134, 130, 250.
+
+```{r, echo = T}
+x <- c(0, 121, 132, 123, 145, 151, 119, 133, 134, 130, 250)
+
+x[ ( abs(x - mean(x)) /sd(x) ) > 2 ]
+
+```
+
+Are the values 0 and 250 declared outliers using the classic outlier detection rule?
+
+_Yes._
+
+### 2.23
+
+Verify that for the data in the previous exercise, the boxplot rule declares the values 0 and 250 outliers.
+
+```{r, echo = T}
+
+bounds <- quantile(x)[c(2,4)]
+
+lower <- bounds[1] - 1.5*(bounds[2] - bounds[1])
+upper <- bounds[2] + 1.5*(bounds[2] - bounds[1])
+
+x[ x < lower | x > upper]
+
+```
+
+### 2.24
+
+Consider the values
+
+20, 121, 132, 123, 145, 151, 119, 133, 134, 240, 250
+
+Verify that no outliers are found using the classic outlier detection rule.
+
+```{r, echo = T}
+x <- c(20, 121, 132, 123, 145, 151, 119, 133, 134, 240, 250)
+
+stopifnot( length( x[ ( abs( x - mean(x) ) / sd(x) ) > 2] ) == 0 )
+
+```
+
+### 2.25
+
+Verify that for the data in the previous exercise, the boxplot rule declares the values 20, 240 and 250 outliers.
+
+```{r, echo = T}
+x <- c(20, 121, 132, 123, 145, 151, 119, 133, 134, 240, 250)
+
+bounds <- quantile(x)[c(2,4)]
+
+lower <- bounds[1] - 1.5*(bounds[2] - bounds[1])
+upper <- bounds[2] + 1.5*(bounds[2] - bounds[1])
+
+x[ x < lower | x > upper]
+
+```
+
+### 2.26
+
+What do the last three exercises suggest about the boxplot rule versus the classic rule for detecting outliers?
+
+_The classic boxplot rule masks outliers._
+
+### 2.27
+
+What is the typical pulse rate (beats per minute) among adults? Imagine that you sample 21 adults, measure their pulse rate, and get:
+
+80, 85, 81, 75, 77, 79, 74, 86, 79, 55, 82, 89, 73, 79, 83, 82, 88, 79, 77, 81, 82
+
+```{r, echo = T}
+x <- c(80, 85, 81, 75, 77, 79, 74, 86, 79, 55, 82, 89, 73, 79, 83, 82, 88, 79, 77, 81, 82)
+```
+
+compute the 20% trimmed mean.
+
+```{r, echo = T}
+
+n <- length(x)
+o <- sort(x)
+g <- floor(.2*n)
+t <- o[ (g+1):(n-g) ]
+tbar <- 1/(n - 2*g)*sum(t)
+
+stopifnot(tbar == mean(x, trim = .2))
+```
+
+### 2.28
+
+For the observations,
+
+21, 36, 42, 24, 25, 36, 35, 49, 32
+
+```{r, echo = T}
+x <- c(21, 36, 42, 24, 25, 36, 35, 49, 32)
+n <- length(x)
+
+xbar <- 1/n * sum(x)
+
+o <- sort(x)
+g <- floor(.2*n)
+t <- o[ (g+1):(n-g) ]
+tbar <- 1/(n - 2*g)*sum(t)
+
+M <- o[ (n + 1)/2 ]
+
+```
+
+Verify that the sample mean, 20% trimmed mean, and median are $\bar{X} = 33.33, \bar{X}_t = 32.9, and \ M = 35$
+
+$\bar{X} = `r xbar`$, $\bar{X}_t = `r tbar`$, $M = `r M`$

### 2019-12-10 Statistical Methods/02_numerical-summaries.rmd
statistical methods, ch2 wrap-up
statistical methods, ch2 wrap-up

@@
 $\bar{X} = `r xbar`$, $\bar{X}_t = `r tbar`$, $M = `r M`$
+
+### 2.29
+
+The largest observation in the last problem is 49. If 49 is replaced by the value 200, verify that the sample mean is now $\bar{X} = 50.1$ but the 20% trimmed mean and median are not changed.
+
+```{r, echo = T}
+x <- c(21, 36, 42, 24, 25, 36, 35, 200, 32)
+
+xbar <- 1/n * sum(x)
+
+o <- sort(x)
+g <- floor(.2*n)
+t <- o[ (g+1):(n-g) ]
+tbar <- 1/(n - 2*g)*sum(t)
+
+M <- o[ (n + 1)/2 ]
+```
+
+$\bar{X} = `r xbar`$, $\bar{X}_t = `r tbar`$, $M = `r M`$
+
+### 2.30
+
+For the data in Exercise 28, what is the minimum number of observations that must be altered so that the 20% trimmed mean is greather than 1,000?
+
+$g + 1 = `r g + 1`$
+
+### 2.31
+
+Repeat the previous exercise, but use the median instead.
+
+_N = 4_
+
+What does this illustrate about the resistance of the mean, median and 20% trimmed mean?
+
+_The mean has the least resistance, followed by the trimmed mean and then the median._
+
+### 2.32
+
+For the observations,
+
+6, 3, 2, 7, 6, 5, 8, 9, 11
+
+Use R to verify that the sample mean, 20% trimmed mean, and median are $\bar{X} = 6.5, \bar{X}_t = 6.7 \ and \ M = 6.5$, respectfully.
+
+```{r, echo = T}
+x <- c(6, 3, 2, 7, 6, 5, 8, 9, 8, 11)
+
+mean(x)
+
+mean(x, trim = .2)
+
+median(x)
+```
+
+### 2.33
+
+In general, when you have _n_ observations, what is the minimum number of values that must be altered to make the 20% trimmed mean grow as larage as you want?
+
+floor(n*.2) + 1
+
+### 2.34
+
+A class of fourth graders was asked to bring a pumpkin to school. Each of the 29 students counted the number of seeds in their pumpkin, and the results were:
+
+250, 220, 281, 247, 230, 209, 240, 160, 370, 274, 210, 204, 243, 251, 190,
+200, 130, 150, 177, 475, 221, 350, 224, 163, 272, 236, 200, 171, 98
+
+Use R to compute the sample mean, 20% trimmed mean, median and MOM.
+
+```{r, echo = T}
+x <- c(250, 220, 281, 247, 230, 209, 240, 160, 370, 274, 210, 204, 243, 251, 190, 
+       200, 130, 150, 177, 475, 221, 350, 224, 163, 272, 236, 200, 171, 98)
+
+xbar <- mean(x)
+tbar <- mean(x, trim = .2)
+M <- median(x)
+MOM <-  mean( x[ abs( x - median(x) ) / mad(x) < 2.27 ] )
+
+```
+
+$\bar{X} = `r xbar`$, $\bar{X}_t = `r tbar`$, $M = `r M`$, $MOM = `r MOM`$
+
+### 2.35
+
+Compute the 20% Winsorized values for the observations:
+
+21, 36, 42, 24, 25, 36, 35, 49, 32
+
+```{r, echo = T}
+x <- c(21, 36, 42, 24, 25, 36, 35, 49, 32)
+
+winsorize <- function( x, trim = .2) {
+   n <- length(x)
+
+   o <- sort(x)
+   g <- floor(trim*n)
+
+   o[1:(g+1)] <- o[(g+1)]
+   o[(n-g):n] <- o[n-g]
+
+   o
+}
+
+winsorize(x)
+```
+
+### 2.36
+
+For the observations in the pervious problem, use R to verify that the 20% Winsorized variance is 51.36.
+
+```{r, echo = T}
+
+Wvar <- var(winsorize(x))
+
+```
+
+$W_{var} = `r Wvar`$
+
+### 2.37
+
+In the previous problem, would you expect the sample variance to be larger or smaller than 51.36?
+
+_Larger, Winsorizing pulls in the extremes._
+
+Verify your answer.
+
+```{r, echo = T}
+var(x)
+```
+
+### 2.38
+
+In general, will the Winsorized sample variance, $s^2_w$, be less than the sample variance, $s^2$?
+
+_Yes._
+
+### 2.39
+
+For the observations,
+
+6, 3, 2, 7, 6, 5, 8, 9, 8, 11
+
+verify that the sample variance and 20% Winsorized variance are 7.4 and 1.8, respectfully.
+
+```{r, echo = T}
+x <- c(6, 3, 2, 7, 6, 5, 8, 9, 8, 11)
+
+variance <- var(x)
+
+```
+
+$Var = `r variance`$, $Var_w = `r Wvar`$
+
+### 2.40
+
+Consider again the number of pumpkin seeds given in Exercise 34.
+
+Compute the 20% Winsorized variance.
+
+```{r, echo = T}
+x <- c(250, 220, 281, 247, 230, 209, 240, 160, 370, 274, 210, 204, 243, 251, 190, 
+       200, 130, 150, 177, 475, 221, 350, 224, 163, 272, 236, 200, 171, 98)
+
+Wvar <- var( winsorize(x) )
+
+```
+
+$W_{var} = `r Wvar`$
+
+### 2.41
+
+Snedecor and Cochran (1967) report results from an experiment dealing with weight gaim in rats as a function of source and amount of protein.
+
+One of the groups was fed beef with a low amount of protein. The weight gains were:
+
+90, 76, 90, 64, 86, 51, 72, 90, 95, 78
+
+Compute the 20% trimmed mean and 20% Winsorized variance.
+
+```{r, echo = T}
+x <- c(90, 76, 90, 64, 86, 51, 72, 90, 95, 78)
+
+tbar <- mean(x, trim = .2)
+wvar <- var(winsorize(x))
+
+```
+
+$\bar{X}_t = `r tbar`$, $Var_w = `r wvar`$

### 2019-12-10 Statistical Methods/03_graphical-summaries.rmd
statistical methods, ch3
statistical methods, ch3

--- 2019-12-08 Statistical Methods/02_numerical-summaries.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Numerical Summaries}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Graphical Summaries}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 2}
+   \lfoot{\color{headergrey}Chapter 3}
    \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
@@
 
-#### Chapter 2
+#### Chapter 3
 
-### 2.1
+### 3.1
 
-Suppose,
+Based upon a sample of 100 individuals, the values 1, 2, 3, 4, 5 are observed with relative frequencies 0.2, 0.3, 0.1, 0.25, 0.15, respectfully.
 
-$X_1 = 1, X_2 = 3, X_3 = 0,$
-$X_4 = -2, X_5 = 4, X_6 = -1,$
-$X_7 = 5, X_8 = 2, X_9 = 10.$
+Compute the mean, variance and standard deviation.
 
 ```{r, echo = T}
-x <- c(1, 3, 0, -2, 4, -1, 5, 2, 10)
-```
+values <- c(1, 2, 3, 4, 5)
+frequencies <- c(0.2, 0.3, 0.1, 0.25, 0.15)
 
-Find:
+stopifnot(sum(frequencies) == 1)
 
-a.) $\sum{X_i}$
+n <- 100
 
-```{r, echo = T}
-sum(x)
-```
-
-b.) $\sum^5_{i=3}X_i$
-
-```{r, echo = T}
-sum( x[3:5] )
-```
-
-c.) $\sum^4_{i=1}X^3_i$
-
-```{r, echo = T}
-sum(x[1:4]^3)
-```
-
-d.) $(\sum{X_i})^2$
-
-```{r, echo = T}
-(sum(x))^2
-```
-
-e.) $\sum{3}$
-
-```{r, echo = T}
-3 * length(x)
-```
-
-f.) $\sum{(X_i - 7)}$
-
-```{r, echo = T}
-sum(x - 7)
-```
-
-g.) $3\sum^5_{i=1}X_i - \sum^9_{i=6}X_i$
-
-```{r, echo = T}
-3 * sum( x[1:5] ) - sum( x[6:9] )
-```
-
-h.) $\sum{10X}$
-
-```{r, echo = T}
-sum( 10 * x)
-```
-
-i.) $\sum^6_{i=2}iX_i$
-
-```{r, echo = T}
-i <- 2:6
-sum( i * x[i]  )
-```
-
-j.) $\sum6$
-
-```{r, echo = T}
-6 * length(x)
-```
-
-### 2.2
-
-Express the following in summation notation.
-
-a.) $X_1 + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$
-
-$\ldots = \frac{X_1}{1} + \frac{X_2}{2} + \frac{X_3}{3} + \frac{X_4}{4}$
-
-$\ldots = \sum{^4}_{i=1}{\frac{X_i}{i}}$
-
-b.) $U_1 + U^2_2 + U^3_3 + U^4_4$
-
-$\ldots = U^1_1 + U^2_2 + U^3_3 + U^4_4$
-
-$\ldots = \sum^4_{i=1}{U^i_i}$
-
-c.) $(Y_1 + Y_2 + Y_3)^4$
-
-$(\sum^3_{i=1}{Y_i})^4$
-
-### 2.3
-
-Show by numerical example that $\sum{X^2_i}$ is not necessarily equal to $(\sum{X_i}^2)$.
-
-```{r, echo = T}
+xbar <- sum(values * frequencies)
+variance <- n/(n - 1) * sum( ((values - xbar)^2) * frequencies )
+stdDev <- sqrt(variance)
 
@@
 
-### 2.4
+$\bar{X} = `r xbar`$, $Var = `r variance`$, $\sigma = `r stdDev`$
 
-Find the mean and median of the following sets of numbers.
+### 3.2
 
-a.) -1, 0, 3, 0, 2, -5.
+Fifty individual are rated on how open-mided they are. The ratings have the values 1, 2, 3, 4, and the corresponding relative frequencies are 0.2, 0.24, 0.4, 0.16, respectively.
 
-b.) 2, 2, 3, 10, 100, 1,000
-
-### 2.5
-
-The final exam scores for 15 students are: 73, 74, 92, 98, 100, 72, 74, 85, 76, 94, 89, 73, 76, 99.
-
-Compute the mean, 20% trimmed mean, and median using R.
+Compute the mean, variance and standard deviation.
 
@@
 
-### 2.6
+### 3.3
 
-The average of 23 numbers is 14.7. What is the sum of these numbers?
+For the values 0, 1, 2, 3, 4, 5, 6, the corresponding relative frequencies based on a sample of 10,000 observations are 0.015625, 0.093750, 0.234375, 0.312500, 0.234375, 0.312500, 0.234375, 0.093750, 0.015625, respectively.
 
-### 2.7
+Determine the mean, variance and standard deviation.
 
-Consider the 10 values: 3, 6, 8, 12, 23, 26, 37, 42, 49, 63.
+```{r, echo = T}
+```
 
-The mean is $\bar{X} = 26.9$
+### 3.4
 
-a.) What is the value of the mean if the largest value, 63, is increased to 100?
+For a local charity, the donations in dollars recieved during the last month were 5, 10, 15, 20, 25, 50, having the frequencies 20, 30, 10, 40, 50, 5, respectively.
 
-b.) What is the mean if 633 is increased to 1,000?
+Compute the mean, variance and standard deviation.
 
-c.) What do these results illustrate about the mean?
+### 3.5
 
-### 2.8
+The values 1, 5, 10, 20 have the frequencies 10, 20, 40, 30.
 
-Repeat the previous exercise, only compute the median instead.
+Compute the mean, variance and standard deviation.
 
-### 2.9
+### 3.6
 
-In general, how many values must be altered to make the sample mean arbitrarily large?
+For the data in Table 2.1, dealing with changes in cholesterol levels, create a histogram with the R function hist.
 
-### 2.10
+```{r, echo = T}
+data <- read.csv(paste0(data.dir, "ibtable2_1_dat.txt"), header = F)
 
-What is the minimum number of values that must be altered to make the 20% trimmed mean and sample median arbitrarily large?
+```
+
+### 3.7
+
+For the data in Table 2.2, create a histogram using the R function _hist_ and speculate about whether values less than zero are outliers.
+
+```{r, echo = T}
+data <- read.csv(paste0(data.dir, "ibtable2_2_dat.txt"), header = F)
+
+```
+
+Then compare your answer to the result obtained using _outbox_.
+
+### 3.8
+
+The heights of 30 male Egyptian skulls from 4,000 BC are reported by Thomson and Randall-Maciver (1905) to be:
+
+121, 124, 129, 130, 130, 131, 131, 132, 132, 132, 133, 133, 134, 134, 134,
+134, 135, 136, 136, 136, 136, 137, 137, 138, 138, 138, 140, 143.
+
+```{r, echo = T}
+x <- c(121, 124, 129, 129, 130, 130, 131, 131, 132, 132, 132, 133, 133, 134, 134, 134,
+134, 135, 136, 136, 136, 136, 136, 137, 137, 138, 138, 138, 140, 143)
+
+n <- length(x)
+
+stopifnot(n == 30)
+
+```
+
+Create a histogram.
+
+Find outliers with classic rule and MAD-median rule.
+
+### 3.9
+
+For the dta in the previous exercise, verify that the classic outlier detection rule, given by _2.6_, finds three outliers, which match the values flagged as outliers by the MAD-median rule.
+
+Despite this result, what are the concerns with the classic outlier detection rule?
+
+### 3.10
+
+What do Exercises 6 and 7 suggest about using a histogram to detect outliers?
+
+### 3.11
+
+Table 3.5 shows the exam scores for 27 students.
+
+Create a stem-and-leaf display.
+
+### 3.12
+
+If the leaf is the hundredths digit, what is the stem for the number 34.679?
+
+### 3.13
+
+Consider the values 5.134, 5.532, 5.869, 5.809, 5.268, 5.495, 5.142, 5.483, 
+5.329, 5.149, 5.240, 5.823.
+
+If the leaf is taken to be the tenths digit, why would this make an uninteresting stem-and-leaf display?
+
+```{r, echo = T}
+x <- c(5.134, 5.532, 5.869, 5.809, 5.268, 5.495, 5.142, 5.483, 
+5.329, 5.149, 5.240, 5.823)
+
+```
+
+### 3.14
+
+For the boxplot in Figure _3.11_, determine approximately, the quartiles, the interquartile range, and the median.
+
+Approximately how large is the largest value not declared an outlier?
+
+### 3.15
+
+In Figure _3.11_, about how large must a value be to be declared an outlier? How small?
+
+### 3.16
+
+Use R to create both a boxplot and a plot of the relative frequencies using the film data in Table 3.1.
+
+### 3.17
+
+Use R to create a boxplot and a kernel density estimate using the data in table 3.2.
+
+### 3.18
+
+Describe a situation where the sample histogram is likely to give a good indication of the population histogram based on 100 observations.
+
+### 3.19
+
+Comment generally on how large of a sample size is needed to ensure that the sample histogram will likely provide a good indication of the population histogram?
+
+### 3.20
+
+When trying to detect outliers, discuss the relative merits of using a histogram vs a boxplot.
+
+### 3.21
+
+A sample histogram indicates that the data are highly skewed to the right.
+
+Is this a reliable indication that if all individuals of interest could be measured, the resulting histogram would also be highly skewed?

### 2019-12-10 Statistical Methods/03_graphical-summaries.rmd
statistical methods, ch 3
statistical methods, ch 3

@@
 ```{r, echo = T}
+n <- 50
+x <- c(1, 2, 3, 4)
+f <- c(0.2, 0.24, 0.4, 0.16)
+
+stopifnot(sum(f) == 1)
+
+xbar <- sum(x * f)
+variance <- n / ( n - 1) * sum( (x - xbar)^2 * f)
+stdDev <- sqrt(variance)
+
 ```
+
+$\bar{X} = `r xbar`$, $Var = `r variance`$, $\sigma = `r stdDev`$
 
@@
 ```{r, echo = T}
+n <- 10e3
+x <- 0:6
+f <- c(0.015625, 0.093750, 0.234375, 0.312500, 0.234375, 0.093750, 0.015625)
+
+stopifnot(sum(f) == 1)
+
+xbar <- sum(x * f)
+variance <- ( n ) / (n - 1) * sum( (x - xbar) ^2 * f)
+stdDev <- sqrt(variance)
+
 ```
+
+$\bar{X} = `r xbar`$, $Var = `r variance`$, $\sigma = `r stdDev`$
 
@@
 
+```{r, echo = T}
+
+x <- c(5, 10, 15, 20, 25, 50)
+d <- c(20, 30, 10, 40, 50, 5)
+n <- sum(d)
+f <- d / n
+
+stopifnot( sum(f) == 1)
+
+xbar <- sum(x * f)
+variance <- n / (n - 1) * sum( (x - xbar)^2 * f)
+stdDev <- sqrt(variance)
+
+```
+
+$\bar{X} = `r xbar`$, $Var = `r variance`$, $\sigma = `r stdDev`$
+
 ### 3.5
@@
 Compute the mean, variance and standard deviation.
+
+```{r, echo = T}
+x <- c(1, 5, 10, 20)
+c <- c(10, 20, 40, 30)
+n <- sum(c)
+f <- c / n
+
+stopifnot(sum(f) == 1)
+
+xbar <- sum( x * f )
+variance <- n / (n - 1) * sum( (x - xbar)^2 * f)
+stdDev <- sqrt(variance)
+
+```
+
+$\bar{X} = `r xbar`$, $Var = `r variance`$, $\sigma = `r stdDev`$
 

### 2019-12-11 Statistical Methods/03_graphical-summaries.rmd
statistical methods, ch 3
statistical methods, ch 3

@@
 
-For the data in Table 2.1, dealing with changes in cholesterol levels, create a histogram with the R function hist.
+For the data in Table 2.1, dealing with changes in cholesterol levels, create a histogram with R.
 
-```{r, echo = T}
-data <- read.csv(paste0(data.dir, "ibtable2_1_dat.txt"), header = F)
+```{r, echo = T, , fig.width=8, fig.height=4}
+data <- data.table::fread(paste0(data.dir, "ibtable2_1_dat.txt"), fill = T)
+
+suppressWarnings({print(
+   ggplot(data.table(value = as.vector(as.matrix(data))), aes(value)) +
+      geom_histogram(aes(fill = ..count..), bins = 30)
+)})
 
@@
 
-For the data in Table 2.2, create a histogram using the R function _hist_ and speculate about whether values less than zero are outliers.
+For the data in Table 2.2, create a histogram using R, and speculate about whether values less than zero are outliers.
 
-```{r, echo = T}
-data <- read.csv(paste0(data.dir, "ibtable2_2_dat.txt"), header = F)
+```{r, echo = T, fig.width=8, fig.height=4}
+data <- data.table::fread(paste0(data.dir, "ibtable2_2_dat.txt"), fill = T)
+
+values <- data.table(value = as.vector(as.matrix(data)))
+values <- values[!is.na(values$value)]
+
+suppressWarnings({print(
+   ggplot(values, aes(value)) +
+      geom_histogram(aes(fill = ..count..), bins = 30)
+)})
 
@@
 
+Below -45 & above 50 appear to be outliers.
+
 Then compare your answer to the result obtained using _outbox_.
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+# MAD outliers
+
+values[ abs(values$value - median(values$value)) / mad(values$value) > 2.27 ]
+
+# Classical Outliers
+
+values[ abs(values$value - mean(values$value)) / sd(values$value) > 2]
+```
 
@@
 
-```{r, echo = T}
+```{r, echo = T, fig.width=8, fig.height=4}
 x <- c(121, 124, 129, 129, 130, 130, 131, 131, 132, 132, 132, 133, 133, 134, 134, 134,
@@
 
+ggplot(data.table(value = x), aes(x)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(fill = "darkorange", alpha = .35)
+
 ```
@@
 
+```{r, echo = T}
+
+x[ abs(x - median(x)) / mad(x) > 2.27 ]
+
+# Classical Outliers
+
+x[ abs(x - mean(x)) / sd(x) > 2]
+
+```
+
 ### 3.9
 
-For the dta in the previous exercise, verify that the classic outlier detection rule, given by _2.6_, finds three outliers, which match the values flagged as outliers by the MAD-median rule.
+For the data in the previous exercise, verify that the classic outlier detection rule, given by _2.6_, finds three outliers, which match the values flagged as outliers by the MAD-median rule.
+
+_Confirmed 3 outliers with both rules_
 
 Despite this result, what are the concerns with the classic outlier detection rule?
+
+_There appears to be masking going on, as there are clearly more than 3 outliers._
 

### 2019-12-11 Statistical Methods/03_graphical-summaries.rmd
statistical methods, ch 3
statistical methods, ch 3
finished graphical summaries
finished graphical summaries

@@
 
+_Histograms are not always useful in detecting outliers_
+
 ### 3.11
@@
 
+```{r, echo = T}
+scores <- c(83, 69, 82, 72, 63, 88, 92, 81, 54,
+            57, 79, 84, 99, 74, 86, 71, 94, 71,
+            80, 51, 68, 81, 84, 92, 63, 99, 91)
+```
+
 Create a stem-and-leaf display.
+
+```{r, echo = T}
+stem(scores)
+```
 
@@
 If the leaf is the hundredths digit, what is the stem for the number 34.679?
+
+_34.6_
 
@@
 
+There would only be one stem:
+
+```{r, echo = T}
+stem(x, scale = 1/10)
+```
+
 ### 3.14
 
-For the boxplot in Figure _3.11_, determine approximately, the quartiles, the interquartile range, and the median.
+For the boxplot in Figure _3.11_, determine approximately, the quartiles, the interquartile range, and the median. _M = 80, Q1/Q2 = 50/121, IQR = 71_
 
-Approximately how large is the largest value not declared an outlier?
+Approximately how large is the largest value not declared an outlier? _215_
 
@@
 
-In Figure _3.11_, about how large must a value be to be declared an outlier? How small?
+In Figure _3.11_, about how large must a value be to be declared an outlier? _x > 227.5_
+
+How small? _x < -56.5_
 
@@
 
+```{r, echo = T, fig.width=8, fig.height=4}
+data <- data.table::fread(paste0(data.dir, "film_dat.txt"), fill = T)
+
+values <- data.table(value = as.vector(as.matrix(data)))
+values <- values[!is.na(values$value)]
+
+p1 <- ggplot(values, aes(y = value)) +
+   geom_boxplot(fill = "cornflowerblue") +
+   labs(title = "Film Ratings")
+
+p2 <- ggplot(values, aes(x = value)) +
+   geom_bar(aes(y = ..prop..), fill = "cornflowerblue", width = 0.05) +
+   labs(title = "Film Ratings")
+
+grid.arrange(p1, p2, nrow = 2)
+
+```
+
 ### 3.17
@@
 Use R to create a boxplot and a kernel density estimate using the data in table 3.2.
+
+```{r, echo = T, fig.width=8, fig.height=4}
+data <- data.table::fread(paste0(data.dir, "mismatch_dat.txt"), fill = T)
+
+values <- data.table(value = as.vector(as.matrix(data)))
+values <- values[!is.na(values$value)]
+
+p1 <- ggplot(values, aes(y = value)) +
+   geom_boxplot(fill = "cornflowerblue") +
+   labs(title = "T5 Mistmatch")
+
+p2 <- ggplot(values, aes(x = value)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(fill = "darkorange", alpha = .3) +
+   labs(title = "T5 Mistmatch Density")
+
+grid.arrange(p1, p2, nrow = 2)
+
+```
 
@@
 
+_The population follows a symmetric distribution and outliers are rare._
+
 ### 3.19
@@
 
+_There are numerical estimators avalaible using bootstrapping techniques, however, a general rule of thumb is > 100._
+
 ### 3.20
@@
 When trying to detect outliers, discuss the relative merits of using a histogram vs a boxplot.
+
+_A boxplot is generally better at showing outliers_.
 
@@
 Is this a reliable indication that if all individuals of interest could be measured, the resulting histogram would also be highly skewed?
+
+_Not always. A population could be symetric and a given sample could be skewed._

### 2019-12-11 Statistical Methods/04_probability.rmd
statistical methods exercises
statistical methods exercises

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Probability and Related Concepts}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 4}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 4

### 4.1

If the possible values for x are 0, 1, 2, 3, 4, 5, and the corresponding values for $P(x)$ are 0.2, 0.2, 0.15, 0.3, 0.35, 0.2, 0.1, respectively, does $P(x)$ qualify as a probability function?

```{r, echo = T}
x <- c(0, 1, 2, 3, 4, 5)
p <- c(0.2, 0.2, 0.15, 0.3, 0.35, 0.2, 0.1)

sum(p)
```

_No. P > 1._

### 4.2

If the possible values for x are 2, 3, 4, and the corresponding values for $P(x)$ are 0.2, -0.1, 0.9, respectively, does $P(x)$ qualify as a probability function?

_No. $0 \le x \le 1$_

### 4.3

If the possible values for x are 1, 2, 3, 4, and the corresponding values for $P(x)$ are 0.1, 0.15, 0.5, 0.25, respectively, does $P(x)$ qualify as a probability function?

```{r, echo = T}
x <- c(1, 2, 3, 4)
p <- c(0.1, 0.15, 0.5, 0.25)

sum(p)

```

Yes, this is a valid probability function.

### 4.4

If the possible values for x are 2, 3, 4, 5, and the corresponding values for $P(x)$ are 0.2, 0.3, 0.4, 0.1, respectively, what is the probability of observing a value less than or equal to 3.4?

```{r, echo = T}
x <- c(2, 3, 4, 5)
p <- c(0.2, 0.3, 0.4, 0.1)

stopifnot( sum(p) == 1 )

prob <- sum( p[ x <= 3.4 ] )

```

Probability: __`r prob*100`%__

### 4.5

For the previous distribution, what is the probability of observing a 1?

_Zero._

### 4.6

For the previous distribution, what is the probability of observing a value greater than 3?

```{r, echo = T}
prob <- sum( p[ x > 3 ])
```

Probability: __`r prob*100`%__

### 4.7

For the previous distribution, what is the probability of observing a value greater than or equal to 3?

```{r, echo = T}
prob <- sum( p[ x >= 3 ])
```

Probability: __`r prob*100`%__

### 4.8

If the probability of observing a value less than or equal to 6 is 0.3, what is the probability of observing a value greater than 6?

```{r, echo = T}
prob <- 1 - .3
```

Probability: __`r prob*100`%__

### 4.9

For the probability function:

$x \ : \ 0, 1$

$P(x) \ : 0.7, 0.3$

Verify that the mean and variance are 0.3 and 0.21, respectively.

```{r, echo = T}
x <- c(0, 1)
p <- c(0.7, 0.3)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$

What is the probability of getting a value less than the mean?

_50%_

### 4.10

Imagine that an auto manufacturer wants to evaluate how potential customers will rate handling for a new car being considered for production. Also, suppose that if all potential customers were to rate handling on a four-point scale, 1 being poor and 4 being excellent, the corresponding probabilities associated with these ratings would be:

$P(1) = 0.2, P(2) = 0.4, P(3) = 0.3, P(4) = 0.1$

Determine the population mean, variance and standard deviation.

```{r, echo = T}
x <- 1:4
p <- c(0.2, 0.4, 0.3, 0.1)

stopifnot(sum(p) == 1)

mu <- sum(x * p)
variance <- sum( (x - mu)^2 * p)
stdDev <- sqrt(variance)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.11

If the possible values for x are 1, 2, 3, 4, 5, with probabilities 0.2, 0.1, 0.1, 0.5, 0.1, respectfully, what are the population mean, variance, and standard deviation?

```{r, echo = T}
x <- 1:5
p <- c(0.2, 0.1, 0.1, 0.5, 0.1)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p )
sigma <- sqrt(variance)
```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.12

In the previous exercise, determine the probability of getting a value within one standard deviation of the mean.

That is, $\mu - \sigma \le x \le \mu + \sigma$

```{r, echo = T}

vals <- mu + c(-1, 1)*sigma
round(vals, 4)

sum( p[ x >= vals[1] & x <= vals[2] ] )

```

### 4.13

If the possible values for x are 1, 2, 3, with probabilities 0.2, 0.6, and 0.2, respectively, what is the mean and standard deviation?

```{r, echo = T}
x <- 1:3
p <- c(0.2, 0.6, 0.2)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p)
stdDev <- sqrt(variance)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.14

In the previous excersize, suppose the possible values for x are now 0, 2, 4 with the same probabilities as before.

Will the standard deviation increase, decrease or stay the same?

_Increase._

```{r, echo = T}
x <- c(0, 2, 4)
p <- c(0.2, 0.6, 0.2)

mu <- sum( x * p )
variance <- sum( (x - mu)^2 * p)
stdDev <- sqrt(variance)

```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.15

For the probability function:

$x : 1, 2, 3, 4, 5$
$P(x) : 0.15, 0.2, 0.3, 0.2, 0.15$

Determine the mean, the variance, and the probability that a value is less than the mean.

```{r, echo = T}
x <- 1:5
p <- c(0.15, 0.2, 0.3, 0.2, 0.15)

mu <- sum( x * p)
variance <- sum( ( x - mu)^2 * p )
stdDev <- sqrt(variance)

sum( p[x < mu] )
```

$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$

### 4.16

For the probability function:

$x : 1, 2, 3, 4, 5$
$P(x) : 0.1, 0.25, 0.3, 0.25, 0.1$

Would you expect the variance to be larger or smaller than the previous pdf?

_Larger._

### 4.17

For the probability function:

$x : 1, 2, 3, 4, 5$
$P(x) : 0.2, 0.2, 0.2, 0.2, 0.2$

Would you expect the variance to be larger or smaller than the previous pdf?

_Smaller._

### 4.18

For the following probabilities:

\begin{center}
 \begin{tabular}{||c c c c||}
 \hline
 \multicolumn{4}{|c|}{Income} \\ 
 \hline
 Age & High & Medium & Low \\ [0.5ex] 
 \hline\hline
 < 30 & 0.030 & 0.180 & 0.090 \\ 
 \hline
 30-50 & 0.052 & 0.312 & 0.156 \\
 \hline
  Over 50 & 0.018 & 0.108 & 0.054 \\
 \hline
 \end{tabular}
\end{center}

a.) The probability that someone is under 30.

_.03 + 0.18 + 0.09 = .30_

b.) The probability that someone has a high income given that they are under 30.

_.03 / .3 = .01_

c.) The probability of someone having a low income given that they are under 30.

_0.09 / .3 = 0.3_

d.) The probability of a medium income given that they are over 50.

_0.018 + 0.108 + 0.054 = .18_

_.108 / .18 = .6_

### 4.19

For the previous data, are income and age independent?

_Yes._

### 4.20

\begin{center}
 \begin{tabular}{||c c c||}
 \hline
 \multicolumn{3}{|c|}{Attitude} \\ 
 \hline
 Member & 1 & 0 \\ [0.5ex] 
 \hline\hline
 Yes & 757 & 496 \\ 
 \hline
 No & 1,071 & 1,074 \\
 \hline
 \end{tabular}
\end{center}

```{r, echo = T}
d <- matrix(c(757, 496, 1071, 1074), nrow = 2, byrow = T)

prop.table(data.table(d))

```

a.) Probability of boy choosing "yes".

.4

b.) P(yes|1)

.22

c.) P(1|yes)

.41

d.) is yes independent of attitude?

No, the probabilities are disproportionate

### 4.21

Let Y be the cost of a home and let X be a measure of the crime rate. If the variance of the cost of a home changes with X, does this mean that the cost of a home and the crime rate are dependent?

_Yes, this can only happen when the conditional probabilites change when told X._

### 4.22

If the probability of Y < 6 is .4 given that X = 2, and if the probability of Y < 6 is .3 given that X = 4, does this mean that X and Y are dependent?

_Yes._

### 4.23

If the range of possible Y values varies with X, does this mean that X and Y are dependent?

_Absolutely._

### 4.24

```{r, echo = T}

```

### 4.25

```{r, echo = T}

```

### 4.26

```{r, echo = T}

```

### 4.27

```{r, echo = T}

```

### 4.28

```{r, echo = T}

```

### 4.29

```{r, echo = T}

```

### 4.30

```{r, echo = T}

```

### 4.31

```{r, echo = T}

```

### 4.33

```{r, echo = T}

```

### 4.34

```{r, echo = T}

```

### 4.35

```{r, echo = T}

```

### 4.36

```{r, echo = T}

```

### 4.37

```{r, echo = T}

```

### 4.38

```{r, echo = T}

```

### 4.39

```{r, echo = T}

```

### 4.40

```{r, echo = T}

```

### 4.41

```{r, echo = T}

```

### 4.42

```{r, echo = T}

```

### 4.43

```{r, echo = T}

```

### 4.44

```{r, echo = T}

```

### 4.45

```{r, echo = T}

```

### 4.46

```{r, echo = T}

```

### 4.47

```{r, echo = T}

```

### 4.48

```{r, echo = T}

```

### 4.49

```{r, echo = T}

```

### 4.50

```{r, echo = T}

```

### 4.51

```{r, echo = T}

```

### 4.52

```{r, echo = T}

```

### 4.53

```{r, echo = T}

```

### 4.54

```{r, echo = T}

```

### 4.55

```{r, echo = T}

```

### 4.56

```{r, echo = T}

```

### 4.57

```{r, echo = T}

```

### 4.58

```{r, echo = T}

```

### 4.59

```{r, echo = T}

```

### 4.60

```{r, echo = T}

```

### 4.61

```{r, echo = T}

```

### 4.62

```{r, echo = T}

```

### 2019-12-12 Statistical Methods/04_probability.rmd
statistical methods
statistical methods
statistical methods, ch4
statistical methods, ch4

@@
 
+For a binomial with n = 10 and p = .4, determine:
+
+a.) $P(0)$
+
 ```{r, echo = T}
+dbinom(0, size = 10, prob = .4)
+```
+
+b.) $P(X \le 3)$
+
+```{r, echo = T}
+pbinom(3, size = 10, prob = .4)
+```
+
+c.) $P(X < 3)$
+
+```{r, echo = T}
+pbinom(2, size = 10, prob = .4)
+```
+
+d.) $P(X > 4)$
+
+```{r, echo = T}
+1 - pbinom(4, size = 10, prob = .4)
+```
+
+e.) $P(2 \le X \le 5)$
+
+```{r, echo = T}
+pbinom(5, size = 10, prob = .4) - pbinom(1, size = 10, prob = .4)
+```
+
+### 4.25
+
+For a binomial with n = 15 and p = 0.3, determine.
+
+a.) $P(0)$
+
+```{r, echo = T}
+dbinom(x = 0, prob = .3, size = 15)
+```
+
+b.) $P( X \le 3 )$
+
+```{r, echo = T}
+pbinom( q = 3, prob = .3, size = 15)
+```
+
+c.) $P(X < 3)$
+
+```{r, echo = T}
+pbinom(2, size = 15, prob = .3)
+```
+
+d.) $P(X > 4)$
+
+```{r, echo = T}
+pbinom(4, size = 15, prob = .3, lower.tail = F)
+```
+
+e.) $P(2 \le X \le 5)$
+
+```{r, echo = T}
+pbinom(5, size = 15, prob = .3) - pbinom(1, size = 15, prob = .3)
+```
+
+### 4.26
+
+For a binomial with n = 15, p = 0.6 determine the probability of exactly 10 successes.
+
+```{r, echo = T}
+dbinom(10, size = 15, prob = .6)
+```
+
+### 4.27
+
+For a binomial with n = 7 and p = 0.35, what is the probability of exactly 2 successes?
+
+```{r, echo = T}
+dbinom(2, size = 7, p = .35)
+```
+
+### 4.28
+
+For a binomial with n = 18 and p = 0.6, determine the mean, variance of X, the total number of successes.
+
+```{r, echo = T}
+n <- 18
+p <- 0.6
+q <- 1 - p
+
+mu <- n * p
+variance <- mu * q
 
@@
 
-### 4.25
+$\mu = `r mu`$, $\sigma^2 = `r variance`$
+
+### 4.29
+
+For a binomial with n = 22 and p = .2, determine the mean and variance of X, the total number of successes.
+
+```{r, echo = T}
+n <- 22
+p <- .2
+q <- 1 - p
+
+mu <- n * p
+variance <- mu * q
+```
+
+$\mu = `r mu`$, $\sigma^2 = `r variance`$
+
+### 4.30
+
+For a binomial with n = 20 and p = .7, determine the mean and variance of $\hat{p}$, the proportion of observed success.
+
+```{r, echo = T}
+n <- 20
+p <- .7
+q <- 1 - p
+
+mu <- n * p
+variance <- mu * q
+```
+
+### 4.31
+
+For a binomial with n = 30 and p = 0.3, determine the mean and variance of $\hat{p}$.
+
+```{r, echo = T}
+n <- 30
+p <- .3
+q <- 1 - p
+
+phat <- p / n
+variance <- p*q / n
+```
+
+$\hat{p} = `r phat`$, $\sigma^2 = `r variance`$
+
+### 4.32
+
+For a binomial with n = 10 and p = 0.8, determine:
+
+```{r, echo = T}
+n <- 10
+p <- 0.8
+q <- 1 - p
+
+variance <- p*q / n
+```
+
+a.) the probability that $\hat{p}$ is less than or equal to 0.7.
 
@@
 
-### 4.26
+b.) the probability that $\hat{p}$ is greater than or equal to 0.8.
 
@@
 
-### 4.27
-
-```{r, echo = T}
-
-```
-
-### 4.28
-
-```{r, echo = T}
-
-```
-
-### 4.29
-
-```{r, echo = T}
-
-```
-
-### 4.30
-
-```{r, echo = T}
-
-```
-
-### 4.31
+c.) the probability that $\hat{p}$ is exactly equal to 0.8.
 
@@
 
+A coin is rigged so that when it is flipped, the probability of a head is 0.7. If the coin is flipped three times, which is the more likely outcome, exactly three heads or two heads and a tail?
+
 ```{r, echo = T}
+
+dbinom(3, 3, .7) # 3 heads
+
+dbinom(2, 3, .7) # 2 heads 1 tail
 
@@
 
+_Two heads, 1 tail._
+
 ### 4.34
 
+Imagine that the probability of heads when flipping a coin is given by the binomial probability function with p = 0.5.
+
+If you flip the coin nine times and get nine heads, what is the probability of a head on the 10th flip?
+
 ```{r, echo = T}
-
+# independent events.
+dbinom(1, 1, .5)
 ```
@@
 
+The Department of Agriculture of the United States reports that 75% of all people who invest in the futures market lose money. Based on the binomial probability function, with n = 5, determine:
+
+a.) the probability that all 5 lose money.
+
+$P(x) = 0$
+
 ```{r, echo = T}
+dbinom(5, size = 5, prob = .75)
+```
+
+b.) the probability that all five make money.
+
+```{r, echo = T}
+dbinom(5, size = 5, prob = .25)
 
@@
 
+c.) the probability that at least two lose money.
+
+```{r, echo = T}
+pbinom(q = 3, size = 5, prob = .25)
+```
+
 ### 4.36
+
+If for a binomial distribution p = 0.4 and n = 25, determine:
+
+a.) $P(X < 11)$
+
+b.) $P(X \le 11)$
+
+c.) $P(X > 9)$
+
+d.) $P(X \ge 9)$
 
@@
 
+In the previous problem, determine the mean of X, the variance of X, the mean of $\hat{p}$, and the variance of $\hat{p}$.
+
 ```{r, echo = T}
@@
 
+Given that Z has a standard normal distribution, determine:
+
+a.) $P(Z \ge 1.5)$
+
+b.) $P(Z \le -2.5)$
+
+c.) $P(Z < -2.5)$
+
+d.) $P(-1 \le Z \le 1)$
+
 ```{r, echo = T}
@@
 ### 4.39
+
+If Z has a standard normal distribution, determine:
+
+a.) $P(Z \le 0.5)$
+
+b.) $P(Z > 1.25)$
+
+c.) $P(-1.2 \le Z \le 1.2)$
+
+d.) $P(-1.8 \le Z \le 1.8)$
 

### 2019-12-14 Statistical Methods/04_probability.rmd
statistical methods, ch 4
statistical methods, ch 4
statistical methods, ch3

@@
 
+```{r, echo = T}
+n <- 25
+p <- .4
+q <- 1 - p
+
+```
+
 a.) $P(X < 11)$
+
+```{r, echo = T}
+pbinom(10, size = n, prob = p)
+```
 
@@
 
+```{r, echo = T}
+pbinom(11, size = n, prob = p)
+```
+
 c.) $P(X > 9)$
+
+```{r, echo = T}
+pbinom(9, size = n, prob = p, lower.tail = F)
+```
 
@@
 ```{r, echo = T}
-
+pbinom(8, size = n, prob = p, lower.tail = F)
 ```
@@
 
+mu <- n * p
+variance <- mu * q
+
+phat <- p
+v <- p*q/n
 ```
+
+$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\hat{p} = `r phat`$, $\sigma^2 = `r v`$
 
@@
 
+```{r, echo = T}
+pnorm(1.5, lower.tail = F)
+```
+
 b.) $P(Z \le -2.5)$
 
+```{r, echo = T}
+pnorm(-2.5)
+```
+
 c.) $P(Z < -2.5)$
+
+```{r, echo = T}
+pnorm(-2.5)
+```
 
@@
 ```{r, echo = T}
+# P(Z > -1) - P(Z > 1)
+pnorm(-1, lower.tail = F) - pnorm(1, lower.tail = F)
 
+# 1 - 2*tail_area
+1 - 2 * pnorm(1, lower.tail = F)
 ```
@@
 
-b.) $P(Z > 1.25)$
+```{r, echo = T}
+pnorm(0.5)
+```
 
-c.) $P(-1.2 \le Z \le 1.2)$
+b.) $P(Z > -1.25)$
+
+```{r, echo = T}
+pnorm(-1.25, lower.tail = F)
+```
+
+c.) $P(-1.2 < Z < 1.2)$
+
+```{r, echo = T}
+1 - 2 *pnorm(1.2, lower.tail = F)
+```
 
 d.) $P(-1.8 \le Z \le 1.8)$
+
+```{r, echo = T}
+pnorm(1.8, lower.tail = T) - pnorm(-1.8)
+```
+
+### 4.40
+
+If Z has a standard normal distribution, determine:
+
+a.) $P(Z < -.5)$
+
+```{r, echo = T}
+1 - pnorm(-.5, lower.tail = F)
+```
+
+b.) $P(Z < 1.2)$
+
+```{r, echo = T}
+1 - pnorm(1.2, lower.tail = F)
+```
+
+c.) $P(Z > 2.1)$
+
+```{r, echo = T}
+pnorm(2.1, lower.tail = F)
+```
+
+d.) $P(-.28 < Z < 0.28)$
+
+```{r, echo = T}
+pnorm(-.28, lower.tail = F) - pnorm(.28, lower.tail = F)
+```
+
+### 4.41
+
+If Z has a standard normal distribution, find _c_ such that:
+
+a.) $P(Z \le c) = 0.0099$
+
+```{r, echo = T}
+qnorm(0.0099)
+```
+
+b.) $P(Z < c) = .9732$
+
+```{r, echo = T}
+qnorm(.9732)
+```
+
+c.) $P(Z > c) = 0.5691$
+
+```{r, echo = T}
+qnorm(.5691, lower.tail = F)
+```
+
+d.) $P(-c \le Z \le c) = 0.2358$
+
+```{r, echo = T}
+qnorm( (1 + 0.2358) / 2)
+```
+
+### 4.42
+
+If X has a standard normal distribution with, determine:
+
+a.) $P(Z > c) = 0.0764$
 
@@
 
-### 4.40
+b.) $P(Z > c) = 0.5040$
 
@@
 
-### 4.41
+c.) $P(-c \le Z \le c) = 0.9108$
 
@@
 
-### 4.42
+d.) $P(-c \le Z \le c) = 0.8$
 
@@
 
+If X has a normal distribution with mean $\mu = 50$ and standard deviation $\sigma = 9$
+
+a.) $P(x \le 40)$
+
+```{r, echo = T}
+
+```
+
+b.) $P(X < 55)$
+
+```{r, echo = T}
+
+```
+
+c.) $P(X > 60)$
+
+```{r, echo = T}
+
+```
+
+d.) $P(40 \le X \le 60)$
+
 ```{r, echo = T}
@@
 ### 4.44
+
+If X has a normal distribution with $\mu = 20$ and $\sigma = 9$, determine:
+
+a.) $P(x < 22)$
+
+```{r, echo = T}
+
+```
+
+b.) $P(X > 17)$
+
+```{r, echo = T}
+
+```
+
+c.) $P(X > 15)$
+
+```{r, echo = T}
+
+```
+
+d.) $P(2 \le X \le 38)$
 
@@
 
+If X has a normal distribution with mean $\mu = .75$ and standard deviation $\sigma = 0.5$, determine:
+
+a.) $P(0.5 < X < 1)$
+
+```{r, echo = T}
+
+```
+
+b.) $P(0.25 < X < 1.25)$
+
 ```{r, echo = T}
@@
 ### 4.46
+
+If X has a normal distribution, determine _c_ such that:
+
+$P(\mu - c\sigma < X < \mu + c\sigma) = .95$
 
@@
 
+If X has a normal distribution, determine _c_ such that:
+
+$P(\mu - c\sigma < X < \mu + c\sigma) = .8$
+
 ```{r, echo = T}
@@
 ### 4.48
+
+Assuming that the scores on a math achievement test are normally distributed with $\mu = 69$ and standard deviation $\sigma = 10$, what is the probability of getting a score greater than 78?
 
@@
 
+In the previous problem, how high must someone score to be in the top 5%?
+
+That is, determine _c_ such that $P(X > c) = 0.05$
+
 ```{r, echo = T}
@@
 ### 4.50
+
+A manufacturer of car batteries claims that the life of their batteries is normally distributed with mean $\mu = 58$ and $\sigma = 3$.
+
+Determien the probability that a randomly selected battery will last at least 62 months.
 
@@
 
+Assume that the income of pediatricials is normally distributed with mean $\mu = \$100,000$ and $\sigma = 10,000$.
+
+Determine the probability of observing an income between $85,000 and $115,000.
+
 ```{r, echo = T}
@@
 
+Suppose the winnings of gamblers at Las Vegas are normally distributed with $\mu = -300$ and $\sigma = 100$.
+
+Determine the probability that a gambler does not lose any money.
+
 ```{r, echo = T}
@@
 ### 4.53
+
+A large computer company claims that their salaries are normally distributed with $\mu = \$50,000$ and $\sigma = \$10,000$.
+
+What is the probability of observing an income between $40,000 and $60,000?
 

### 2019-12-14 Statistical Methods/04_probability.rmd
statistical method, ch 4
statistical method, ch 4

@@
 
-If X has a standard normal distribution with, determine:
+If Z has a standard normal distribution with, determine:
 
@@
 ```{r, echo = T}
-
+qnorm(0.0764, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+qnorm(0.504, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+qnorm((1 + 0.9108)/2)
 ```
@@
 ```{r, echo = T}
-
+qnorm((1+.8)/2)
 ```
@@
 
-a.) $P(x \le 40)$
-
-```{r, echo = T}
-
+a.) $P(X \le 40)$
+
+```{r, echo = T}
+pnorm(40, mean = 50, sd = 9)
 ```
@@
 ```{r, echo = T}
-
+1 - pnorm(55, mean = 50, sd = 9, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+pnorm(60, mean = 50, sd = 9, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+pnorm(60, mean = 50, sd = 9) - pnorm(40, mean = 50, sd = 9)
 ```
@@
 
-a.) $P(x < 22)$
-
-```{r, echo = T}
-
+a.) $P(X < 22)$
+
+```{r, echo = T}
+1 - pnorm(22, mean = 20, sd = 9, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+pnorm(17, mean = 20, sd = 9, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+pnorm(15, mean = 20, sd = 9, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+pnorm(38, mean = 20, sd = 5) - pnorm(2, mean = 20, sd = 9)
 ```
@@
 ```{r, echo = T}
-
+pnorm(1, mean = .75, sd = .5) - pnorm(.5, mean = .75, sd = .5)
 ```
@@
 ```{r, echo = T}
-
+pnorm(1.25, mean = .75, sd = .5) - pnorm(.25, mean = .75, sd = .5)
 ```
@@
 ```{r, echo = T}
-
+qnorm((1 +.95)/2)
 ```
@@
 ```{r, echo = T}
-
+qnorm((1 +.8)/2)
 ```
@@
 
-Assuming that the scores on a math achievement test are normally distributed with $\mu = 69$ and standard deviation $\sigma = 10$, what is the probability of getting a score greater than 78?
-
-```{r, echo = T}
-
+Assuming that the scores on a math achievement test are normally distributed with $\mu = 68$ and standard deviation $\sigma = 10$, what is the probability of getting a score greater than 78?
+
+```{r, echo = T}
+pnorm(78, mean = 68, sd = 10, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+qnorm(1 - 0.05, mean = 68, sd = 10)
 ```
@@
 
-Determien the probability that a randomly selected battery will last at least 62 months.
-
-```{r, echo = T}
-
+Determine the probability that a randomly selected battery will last at least 62 months.
+
+```{r, echo = T}
+pnorm(62, mean = 58, sd = 3, lower.tail = F) # more than 62
 ```
@@
 ```{r, echo = T}
-
+pnorm(1.15, mean = 1, sd = .1) - pnorm(.85, mean = 1, sd = .1)
 ```
@@
 ```{r, echo = T}
-
+pnorm(0, mean = -300, sd = 100, lower.tail = F)
 ```
@@
 ```{r, echo = T}
-
+pnorm(6, mean = 5, sd = 1) - pnorm(4, mean = 5, sd = 1)
 ```
@@
 
-```{r, echo = T}
-
+Suppose the daily amount of solar radiation in Los Angeles is normally distributed with mean 450 and sd 50.
+
+Determine the probability that for a given day the radiation is between 350 and 550.
+
+```{r, echo = T}
+pnorm(5.5, mean = 4.5, sd = .5) - pnorm(3.5, mean = 4.5, sd = .5)
 ```
@@
 
-```{r, echo = T}
-
+If the cholesterol levels of adults are normally distributed with mean 230 and standard deviation 25, what is the probability that a randomly sampled adult has a cholesterol level greater than 260?
+
+```{r, echo = T}
+pnorm(2.6, mean = 2.3, sd = .25, lower.tail = F)
 ```
@@
 
-```{r, echo = T}
-
+If after 1 year, the annual mileage of privately owned cars is normally distributed with mean 14,000 miles and sd 3,500, what is the probability a car has greater than 20,000 miles?
+
+```{r, echo = T}
+pnorm(20, mean = 14, sd = 3.5, lower.tail = F)
 ```
@@
 
-```{r, echo = T}
-
-```
+Can small changes in the tails of a distribution result in large changes in the population mean, $\mu$, relative to the changes in median?
+
+_Yes, the mean is heavly influenced by the tails, where as the median is not._
 
@@
 
-```{r, echo = T}
-
-```
+Explain in what sense the population variance is sensitive to small changes in a distribution.
+
+_The variance is sensetive to small changes in the tail._
 
@@
 
-```{r, echo = T}
-
-```
+For normal random variables, the probability of being within one standard deviation of the mean is .68. That is, $P(\mu - \sigma \le X \le \mu + \sigma) = .68$, if X has a normal distribution.
+
+For nonnormal distribution, is it safe to assume that this probability is again .68?
+
+_No. The AUC (and therefore the mean/variance relationship) for a distribution is defined by its pdf, $P(X) = F_x$, which will be unique per distribution._
 
@@
 
-```{r, echo = T}
-
-```
+If a distribution appears to be bell-shaped and symmetric about its mean, can we assume that the probability of being within one sd of the mean is .68?
+
+_No._
 
@@
 
-```{r, echo = T}
-
-```
+Can two distribution differ by a large amount yet have equal means and variances?
+
+_Yes._
 
@@
 
-```{r, echo = T}
-
-```
+If a distribution is skewed, is it possible that the mean exceedes the .85 quantile?
+
+_Yes._

### 2019-12-14 Statistical Methods/05_sampling-distributions.rmd
statistical methods, ch5
statistical methods, ch5

--- 2019-12-12 Statistical Methods/04_probability.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Probability and Related Concepts}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Sampling Distributions}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 4}
+   \lfoot{\color{headergrey}Chapter 5}
    \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
@@
 
-#### Chapter 4
+#### Chapter 5
 
-### 4.1
-
-If the possible values for x are 0, 1, 2, 3, 4, 5, and the corresponding values for $P(x)$ are 0.2, 0.2, 0.15, 0.3, 0.35, 0.2, 0.1, respectively, does $P(x)$ qualify as a probability function?
-
-```{r, echo = T}
-x <- c(0, 1, 2, 3, 4, 5)
-p <- c(0.2, 0.2, 0.15, 0.3, 0.35, 0.2, 0.1)
-
-sum(p)
-```
-
-_No. P > 1._
-
-### 4.2
-
-If the possible values for x are 2, 3, 4, and the corresponding values for $P(x)$ are 0.2, -0.1, 0.9, respectively, does $P(x)$ qualify as a probability function?
-
-_No. $0 \le x \le 1$_
-
-### 4.3
-
-If the possible values for x are 1, 2, 3, 4, and the corresponding values for $P(x)$ are 0.1, 0.15, 0.5, 0.25, respectively, does $P(x)$ qualify as a probability function?
-
-```{r, echo = T}
-x <- c(1, 2, 3, 4)
-p <- c(0.1, 0.15, 0.5, 0.25)
-
-sum(p)
-
-```
-
-Yes, this is a valid probability function.
-
-### 4.4
-
-If the possible values for x are 2, 3, 4, 5, and the corresponding values for $P(x)$ are 0.2, 0.3, 0.4, 0.1, respectively, what is the probability of observing a value less than or equal to 3.4?
-
-```{r, echo = T}
-x <- c(2, 3, 4, 5)
-p <- c(0.2, 0.3, 0.4, 0.1)
-
-stopifnot( sum(p) == 1 )
-
-prob <- sum( p[ x <= 3.4 ] )
-
-```
-
-Probability: __`r prob*100`%__
-
-### 4.5
-
-For the previous distribution, what is the probability of observing a 1?
-
-_Zero._
-
-### 4.6
-
-For the previous distribution, what is the probability of observing a value greater than 3?
-
-```{r, echo = T}
-prob <- sum( p[ x > 3 ])
-```
-
-Probability: __`r prob*100`%__
-
-### 4.7
-
-For the previous distribution, what is the probability of observing a value greater than or equal to 3?
-
-```{r, echo = T}
-prob <- sum( p[ x >= 3 ])
-```
-
-Probability: __`r prob*100`%__
-
-### 4.8
-
-If the probability of observing a value less than or equal to 6 is 0.3, what is the probability of observing a value greater than 6?
-
-```{r, echo = T}
-prob <- 1 - .3
-```
-
-Probability: __`r prob*100`%__
-
-### 4.9
-
-For the probability function:
-
-$x \ : \ 0, 1$
-
-$P(x) \ : 0.7, 0.3$
-
-Verify that the mean and variance are 0.3 and 0.21, respectively.
-
-```{r, echo = T}
-x <- c(0, 1)
-p <- c(0.7, 0.3)
-
-mu <- sum( x * p )
-variance <- sum( (x - mu)^2 * p)
-
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$
-
-What is the probability of getting a value less than the mean?
-
-_50%_
-
-### 4.10
-
-Imagine that an auto manufacturer wants to evaluate how potential customers will rate handling for a new car being considered for production.
-Also, suppose that if all potential customers were to rate handling on a four-point scale, 1 being poor and 4 being excellent, the corresponding probabilities associated with these ratings would be:
-
-$P(1) = 0.2, P(2) = 0.4, P(3) = 0.3, P(4) = 0.1$
-
-Determine the population mean, variance and standard deviation.
-
-```{r, echo = T}
-x <- 1:4
-p <- c(0.2, 0.4, 0.3, 0.1)
-
-stopifnot(sum(p) == 1)
-
-mu <- sum(x * p)
-variance <- sum( (x - mu)^2 * p)
-stdDev <- sqrt(variance)
-
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$
-
-### 4.11
-
-If the possible values for x are 1, 2, 3, 4, 5, with probabilities 0.2, 0.1, 0.1, 0.5, 0.1, respectfully, what are the population mean, variance, and standard deviation?
-
-```{r, echo = T}
-x <- 1:5
-p <- c(0.2, 0.1, 0.1, 0.5, 0.1)
-
-mu <- sum( x * p )
-variance <- sum( (x - mu)^2 * p )
-sigma <- sqrt(variance)
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$
-
-### 4.12
-
-In the previous exercise, determine the probability of getting a value within one standard deviation of the mean.
-
-That is, $\mu - \sigma \le x \le \mu + \sigma$
-
-```{r, echo = T}
-
-vals <- mu + c(-1, 1)*sigma
-round(vals, 4)
-
-sum( p[ x >= vals[1] & x <= vals[2] ] )
-
-```
-
-### 4.13
-
-If the possible values for x are 1, 2, 3, with probabilities 0.2, 0.6, and 0.2, respectively, what is the mean and standard deviation?
-
-```{r, echo = T}
-x <- 1:3
-p <- c(0.2, 0.6, 0.2)
-
-mu <- sum( x * p )
-variance <- sum( (x - mu)^2 * p)
-stdDev <- sqrt(variance)
-
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$
-
-### 4.14
-
-In the previous excersize, suppose the possible values for x are now 0, 2, 4 with the same probabilities as before.
-
-Will the standard deviation increase, decrease or stay the same?
-
-_Increase._
-
-```{r, echo = T}
-x <- c(0, 2, 4)
-p <- c(0.2, 0.6, 0.2)
-
-mu <- sum( x * p )
-variance <- sum( (x - mu)^2 * p)
-stdDev <- sqrt(variance)
-
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$
-
-### 4.15
-
-For the probability function:
-
-$x : 1, 2, 3, 4, 5$
-$P(x) : 0.15, 0.2, 0.3, 0.2, 0.15$
-
-Determine the mean, the variance, and the probability that a value is less than the mean.
-
-```{r, echo = T}
-x <- 1:5
-p <- c(0.15, 0.2, 0.3, 0.2, 0.15)
-
-mu <- sum( x * p)
-variance <- sum( ( x - mu)^2 * p )
-stdDev <- sqrt(variance)
-
-sum( p[x < mu] )
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$, $\sigma = `r stdDev`$
-
-### 4.16
-
-For the probability function:
-
-$x : 1, 2, 3, 4, 5$
-$P(x) : 0.1, 0.25, 0.3, 0.25, 0.1$
-
-Would you expect the variance to be larger or smaller than the previous pdf?
-
-_Larger._
-
-### 4.17
-
-For the probability function:
-
-$x : 1, 2, 3, 4, 5$
-$P(x) : 0.2, 0.2, 0.2, 0.2, 0.2$
-
-Would you expect the variance to be larger or smaller than the previous pdf?
-
-_Smaller._
-
-### 4.18
-
-For the following probabilities:
-
-\begin{center}
- \begin{tabular}{||c c c c||}
- \hline
- \multicolumn{4}{|c|}{Income} \\ 
- \hline
- Age & High & Medium & Low \\ [0.5ex] 
- \hline\hline
- < 30 & 0.030 & 0.180 & 0.090 \\ 
- \hline
- 30-50 & 0.052 & 0.312 & 0.156 \\
- \hline
-  Over 50 & 0.018 & 0.108 & 0.054 \\
- \hline
- \end{tabular}
-\end{center}
-
-a.) The probability that someone is under 30.
-
-_.03 + 0.18 + 0.09 = .30_
-
-b.) The probability that someone has a high income given that they are under 30.
-
-_.03 / .3 = .01_
-
-c.) The probability of someone having a low income given that they are under 30.
-
-_0.09 / .3 = 0.3_
-
-d.) The probability of a medium income given that they are over 50.
-
-_0.018 + 0.108 + 0.054 = .18_
-
-_.108 / .18 = .6_
-
-### 4.19
-
-For the previous data, are income and age independent?
-
-_Yes._
-
-### 4.20
-
-\begin{center}
- \begin{tabular}{||c c c||}
- \hline
- \multicolumn{3}{|c|}{Attitude} \\ 
- \hline
- Member & 1 & 0 \\ [0.5ex] 
- \hline\hline
- Yes & 757 & 496 \\ 
- \hline
- No & 1,071 & 1,074 \\
- \hline
- \end{tabular}
-\end{center}
-
-```{r, echo = T}
-d <- matrix(c(757, 496, 1071, 1074), nrow = 2, byrow = T)
-
-prop.table(data.table(d))
-
-```
-
-a.) Probability of boy choosing "yes".
-
-.4
-
-b.) P(yes|1)
-
-.22
-
-c.) P(1|yes)
-
-.41
-
-d.) is yes independent of attitude?
-
-No, the probabilities are disproportionate
-
-### 4.21
-
-Let Y be the cost of a home and let X be a measure of the crime rate. If the variance of the cost of a home changes with X, does this mean that the cost of a home and the crime rate are dependent?
-
-_Yes, this can only happen when the conditional probabilites change when told X._
-
-### 4.22
-
-If the probability of Y < 6 is .4 given that X = 2, and if the probability of Y < 6 is .3 given that X = 4, does this mean that X and Y are dependent?
-
-_Yes._
-
-### 4.23
-
-If the range of possible Y values varies with X, does this mean that X and Y are dependent?
-
-_Absolutely._
-
-### 4.24
-
-For a binomial with n = 10 and p = .4, determine:
-
-a.) $P(0)$
-
-```{r, echo = T}
-dbinom(0, size = 10, prob = .4)
-```
-
-b.) $P(X \le 3)$
-
-```{r, echo = T}
-pbinom(3, size = 10, prob = .4)
-```
-
-c.) $P(X < 3)$
-
-```{r, echo = T}
-pbinom(2, size = 10, prob = .4)
-```
-
-d.) $P(X > 4)$
-
-```{r, echo = T}
-1 - pbinom(4, size = 10, prob = .4)
-```
-
-e.) $P(2 \le X \le 5)$
-
-```{r, echo = T}
-pbinom(5, size = 10, prob = .4) - pbinom(1, size = 10, prob = .4)
-```
-
-### 4.25
-
-For a binomial with n = 15 and p = 0.3, determine.
-
-a.) $P(0)$
-
-```{r, echo = T}
-dbinom(x = 0, prob = .3, size = 15)
-```
-
-b.) $P( X \le 3 )$
-
-```{r, echo = T}
-pbinom( q = 3, prob = .3, size = 15)
-```
-
-c.) $P(X < 3)$
-
-```{r, echo = T}
-pbinom(2, size = 15, prob = .3)
-```
-
-d.) $P(X > 4)$
-
-```{r, echo = T}
-pbinom(4, size = 15, prob = .3, lower.tail = F)
-```
-
-e.) $P(2 \le X \le 5)$
-
-```{r, echo = T}
-pbinom(5, size = 15, prob = .3) - pbinom(1, size = 15, prob = .3)
-```
-
-### 4.26
-
-For a binomial with n = 15, p = 0.6 determine the probability of exactly 10 successes.
-
-```{r, echo = T}
-dbinom(10, size = 15, prob = .6)
-```
-
-### 4.27
-
-For a binomial with n = 7 and p = 0.35, what is the probability of exactly 2 successes?
-
-```{r, echo = T}
-dbinom(2, size = 7, p = .35)
-```
-
-### 4.28
-
-For a binomial with n = 18 and p = 0.6, determine the mean, variance of X, the total number of successes.
-
-```{r, echo = T}
-n <- 18
-p <- 0.6
-q <- 1 - p
-
-mu <- n * p
-variance <- mu * q
-
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$
-
-### 4.29
-
-For a binomial with n = 22 and p = .2, determine the mean and variance of X, the total number of successes.
-
-```{r, echo = T}
-n <- 22
-p <- .2
-q <- 1 - p
-
-mu <- n * p
-variance <- mu * q
-```
-
-$\mu = `r mu`$, $\sigma^2 = `r variance`$
-
-### 4.30
-
-For a binomial with n = 20 and p = .7, determine the mean and variance of $\hat{p}$, the proportion of observed success.
-
-```{r, echo = T}
-n <- 20
-p <- .7
-q <- 1 - p
-
-mu <- n * p
-variance <- mu * q
-```
-
-### 4.31
-
-For a binomial with n = 30 and p = 0.3, determine the mean and variance of $\hat{p}$.
-
-```{r, echo = T}
-n <- 30
-p <- .3
-q <- 1 - p
-
-phat <- p / n
-variance <- p*q / n
-```
-
-$\hat{p} = `r phat`$, $\sigma^2 = `r variance`$
-
-### 4.32
-
-For a binomial with n = 10 and p = 0.8, determine:
-
-```{r, echo = T}
-n <- 10
-p <- 0.8
-q <- 1 - p
-
-variance <- p*q / n
-```
-
-a.) the probability that $\hat{p}$ is less than or equal to 0.7.
-
-```{r, echo = T}
-
-```
-
-b.) the probability that $\hat{p}$ is greater than or equal to 0.8.
-
-```{r, echo = T}
-
-```
-
-c.) the probability that $\hat{p}$ is exactly equal to 0.8.
-
-```{r, echo = T}
-
-```
-
-### 4.33
-
-A coin is rigged so that when it is flipped, the probability of a head is 0.7. If the coin is flipped three times, which is the more likely outcome, exactly three heads or two heads and a tail?
-
-```{r, echo = T}
-
-dbinom(3, 3, .7) # 3 heads
-
-dbinom(2, 3, .7) # 2 heads 1 tail
-
-```
-
-_Two heads, 1 tail._
-
-### 4.34
-
-Imagine that the probability of heads when flipping a coin is given by the binomial probability function with p = 0.5.
-
-If you flip the coin nine times and get nine heads, what is the probability of a head on the 10th flip?
-
-```{r, echo = T}
-# independent events.
-dbinom(1, 1, .5)
-```
-
-### 4.35
-
-The Department of Agriculture of the United States reports that 75% of all people who invest in the futures market lose money. Based on the binomial probability function, with n = 5, determine:
-
-a.) the probability that all 5 lose money.
-
-$P(x) = 0$
-
-```{r, echo = T}
-dbinom(5, size = 5, prob = .75)
-```
-
-b.) the probability that all five make money.
-
-```{r, echo = T}
-dbinom(5, size = 5, prob = .25)
-
-```
-
-c.) the probability that at least two lose money.
-
-```{r, echo = T}
-pbinom(q = 3, size = 5, prob = .25)
-```
-
-### 4.36
-
-If for a binomial distribution p = 0.4 and n = 25, determine:
-
-a.) $P(X < 11)$
-
-b.) $P(X \le 11)$
-
-c.) $P(X > 9)$
-
-d.) $P(X \ge 9)$
-
-```{r, echo = T}
-
-```
-
-### 4.37
-
-In the previous problem, determine the mean of X, the variance of X, the mean of $\hat{p}$, and the variance of $\hat{p}$.
-
-```{r, echo = T}
-
-```
-
-### 4.38
-
-Given that Z has a standard normal distribution, determine:
-
-a.) $P(Z \ge 1.5)$
-
-b.) $P(Z \le -2.5)$
-
-c.) $P(Z < -2.5)$
-
-d.) $P(-1 \le Z \le 1)$
-
-```{r, echo = T}
-
-```
-
-### 4.39
-
-If Z has a standard normal distribution, determine:
-
-a.) $P(Z \le 0.5)$
-
-b.) $P(Z > 1.25)$
-
-c.) $P(-1.2 \le Z \le 1.2)$
-
-d.) $P(-1.8 \le Z \le 1.8)$
-
-```{r, echo = T}
-
-```
-
-### 4.40
-
-```{r, echo = T}
-
-```
-
-### 4.41
-
-```{r, echo = T}
-
-```
-
-### 4.42
-
-```{r, echo = T}
-
-```
-
-### 4.43
-
-```{r, echo = T}
-
-```
-
-### 4.44
-
-```{r, echo = T}
-
-```
-
-### 4.45
-
-```{r, echo = T}
-
-```
-
-### 4.46
-
-```{r, echo = T}
-
-```
-
-### 4.47
-
-```{r, echo = T}
-
-```
-
-### 4.48
-
-```{r, echo = T}
-
-```
-
-### 4.49
-
-```{r, echo = T}
-
-```
-
-### 4.50
-
-```{r, echo = T}
-
-```
-
-### 4.51
-
-```{r, echo = T}
-
-```
-
-### 4.52
-
-```{r, echo = T}
-
-```
-
-### 4.53
-
-```{r, echo = T}
-
-```
-
-### 4.54
-
-```{r, echo = T}
-
-```
-
-### 4.55
-
-```{r, echo = T}
-
-```
-
-### 4.56
-
-```{r, echo = T}
-
-```
-
-### 4.57
-
-```{r, echo = T}
-
-```
-
-### 4.58
-
-```{r, echo = T}
-
-```
-
-### 4.59
-
-```{r, echo = T}
-
-```
-
-### 4.60
-
-```{r, echo = T}
-
-```
-
-### 4.61
-
-```{r, echo = T}
-
-```
-
-### 4.62
-
-```{r, echo = T}
-
-```
+### 5.1

### 2019-12-16 Mathematical Statistics/03_permutation_testing.tex
statistical methods
statistical methods

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{report}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Arial}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdfborder={0 0 0},
  breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[left = 1cm, right = 1cm, top = 2cm, bottom = 3cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-2}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\PassOptionsToPackage{table}{xcolor}
\usepackage{caption}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage[table]{xcolor}
\usepackage{fancyhdr}
\usepackage{boldline}
\usepackage{tipa} \definecolor{headergrey}{HTML}{545454} \definecolor{msdblue}{HTML}{1C93D1} \pagestyle{fancy} \setlength\headheight{30pt} \rhead{\color{headergrey}\today} \fancyhead[L]{\color{headergrey}Moretz, Brandon} \fancyhead[C]{\Large\bfseries\color{headergrey}Introduction to Hypothesis - Testing Permutation Tests} \rfoot{\color{headergrey}Chapter 3} \lfoot{\color{headergrey}} \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\author{}
\date{\vspace{-2.5em}15 December, 2019}

\begin{document}

\hypertarget{chapter-3-exercises}{%
\subsubsection{Chapter 3 Exercises}\label{chapter-3-exercises}}

\hypertarget{section}{%
\subsection{3.1}\label{section}}

Suppose you conduct an experiment and inject a drug into three mice.

Their times for running a maze are 8, 10, 15 s; the times for two
control mice are 5 and 9 s.

a.) Compute the difference in mean times between the treatment group and
the control group.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mice.t <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{8}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{15}\NormalTok{)}
\NormalTok{mice.c <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{9}\NormalTok{)}

\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(mice.t) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(mice.c)}

\NormalTok{observed}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 4
\end{verbatim}

b.) Write out all the possible permutations of these times to the two
groups and calculate the diffence in means.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mice <-}\StringTok{ }\KeywordTok{c}\NormalTok{(mice.t, mice.c)}

\CommentTok{# 5 choose 3 for treatment}

\NormalTok{treatment <-}\StringTok{ }\KeywordTok{combinations}\NormalTok{(}\DataTypeTok{n =} \DecValTok{5}\NormalTok{, }\DataTypeTok{r =} \DecValTok{3}\NormalTok{, mice, }\DataTypeTok{repeats.allowed =}\NormalTok{ F)}

\NormalTok{control <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{nrow =} \DecValTok{10}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{)}
\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\KeywordTok{nrow}\NormalTok{(control))}
\NormalTok{\{}
\NormalTok{  control[i,] <-}\StringTok{ }\NormalTok{mice[}\OperatorTok{!}\NormalTok{mice }\OperatorTok{%in%}\StringTok{ }\NormalTok{treatment[i,]]}
\NormalTok{\}}

\NormalTok{perms <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(}\KeywordTok{cbind}\NormalTok{(treatment, control))}

\KeywordTok{stopifnot}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(perms) }\OperatorTok{==}\StringTok{ }\KeywordTok{choose}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{3}\NormalTok{))}

\KeywordTok{colnames}\NormalTok{(perms) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"D1"}\NormalTok{, }\StringTok{"D2"}\NormalTok{, }\StringTok{"D3"}\NormalTok{, }\StringTok{"C1"}\NormalTok{, }\StringTok{"C2"}\NormalTok{)}

\NormalTok{perms}\OperatorTok{$}\NormalTok{Xd <-}\StringTok{ }\NormalTok{(perms}\OperatorTok{$}\NormalTok{D1 }\OperatorTok{+}\StringTok{ }\NormalTok{perms}\OperatorTok{$}\NormalTok{D2 }\OperatorTok{+}\StringTok{ }\NormalTok{perms}\OperatorTok{$}\NormalTok{D3) }\OperatorTok{/}\StringTok{ }\DecValTok{3}
\NormalTok{perms}\OperatorTok{$}\NormalTok{Xc <-}\StringTok{ }\NormalTok{(perms}\OperatorTok{$}\NormalTok{C1 }\OperatorTok{+}\StringTok{ }\NormalTok{perms}\OperatorTok{$}\NormalTok{C2) }\OperatorTok{/}\StringTok{ }\DecValTok{2}
\NormalTok{perms}\OperatorTok{$}\NormalTok{Diff <-}\StringTok{ }\KeywordTok{round}\NormalTok{(perms}\OperatorTok{$}\NormalTok{Xd }\OperatorTok{-}\StringTok{ }\NormalTok{perms}\OperatorTok{$}\NormalTok{Xc, }\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\newpage

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-3}Mice Permutations}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|r}
\hline
D1 & D2 & D3 & C1 & C2 & Xd & Xc & Diff\\
\hline
5 & 8 & 9 & 10 & 15 & 7.33 & 12.5 & -5.17\\
\hline
5 & 8 & 10 & 15 & 9 & 7.67 & 12.0 & -4.33\\
\hline
5 & 8 & 15 & 10 & 9 & 9.33 & 9.5 & -0.17\\
\hline
5 & 9 & 10 & 8 & 15 & 8.00 & 11.5 & -3.50\\
\hline
5 & 9 & 15 & 8 & 10 & 9.67 & 9.0 & 0.67\\
\hline
5 & 10 & 15 & 8 & 9 & 10.00 & 8.5 & 1.50\\
\hline
8 & 9 & 10 & 15 & 5 & 9.00 & 10.0 & -1.00\\
\hline
8 & 9 & 15 & 10 & 5 & 10.67 & 7.5 & 3.17\\
\hline
8 & 10 & 15 & 5 & 9 & 11.00 & 7.0 & 4.00\\
\hline
9 & 10 & 15 & 8 & 5 & 11.33 & 6.5 & 4.83\\
\hline
\end{tabular}
\end{table}

c.) What proportion of the differences are as large or larger than the
observed differences in mean times?

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{gte.observed <-}\StringTok{ }\NormalTok{perms[Diff }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]}

\KeywordTok{pretty_kable}\NormalTok{(gte.observed, }\StringTok{"Greater than or Equal to Observed"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-4}Greater than or Equal to Observed}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|r}
\hline
D1 & D2 & D3 & C1 & C2 & Xd & Xc & Diff\\
\hline
8 & 10 & 15 & 5 & 9 & 11.00 & 7.0 & 4.00\\
\hline
9 & 10 & 15 & 8 & 5 & 11.33 & 6.5 & 4.83\\
\hline
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p1c <-}\StringTok{ }\KeywordTok{nrow}\NormalTok{(gte.observed) }\OperatorTok{/}\StringTok{ }\KeywordTok{nrow}\NormalTok{(perms)}
\end{Highlighting}
\end{Shaded}

\textbf{Proportion of differences greater than or equal to observed:
20\%}

d.) For each permutation, calculate the mean of the treatment group
only.

What proportion of these means are as large or larger than the observed
mean of the treatment group?

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{gte.t <-}\StringTok{ }\NormalTok{perms[ Xd }\OperatorTok{>=}\StringTok{ }\KeywordTok{mean}\NormalTok{(mice.t),]}
\KeywordTok{pretty_kable}\NormalTok{(gte.t, }\StringTok{"Mean Treatment Greater than or Equal to Observed"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-5}Mean Treatment Greater than or Equal to Observed}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|r}
\hline
D1 & D2 & D3 & C1 & C2 & Xd & Xc & Diff\\
\hline
8 & 10 & 15 & 5 & 9 & 11.00 & 7.0 & 4.00\\
\hline
9 & 10 & 15 & 8 & 5 & 11.33 & 6.5 & 4.83\\
\hline
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{p1d <-}\StringTok{ }\KeywordTok{nrow}\NormalTok{(gte.t) }\OperatorTok{/}\StringTok{ }\KeywordTok{nrow}\NormalTok{(perms)}
\end{Highlighting}
\end{Shaded}

\textbf{Proportion of treatment groups greater than observed: 20\%}

\hypertarget{section-1}{%
\subsection{3.2}\label{section-1}}

Your statistics professor comes to class with a big urn that she claims
contains 9,999 blue marbels and 1 red marble.

You draw our one marble at random and finds that it is red.

Would you be willing to tell your professor that you think she is wrong
about the distribution of colors?

Why or why not?

\begin{itemize}
\tightlist
\item
  Yes, a 1/10,000 chance is pretty rare.
\end{itemize}

What are you assuming in making your decision?

What if instead, she claims there are nine blue marbles and 1 red one
(and you draw out a red marble)?

\begin{itemize}
\tightlist
\item
  A 1/10 chance is fairly common.
\end{itemize}

\hypertarget{section-2}{%
\subsection{3.3}\label{section-2}}

In a hypothesis test comparing two populations means,
\(H_0 : \mu_1 = \mu_2\ versus \ H_A : \mu_1 > \mu_2\)

\textbf{a.)} Which P-value, \emph{0.03} or \emph{0.006}, provides
stronger evidence for the alternative hypothesis?

\emph{0.03} provides stronger evidence for the alternative hypothesis.

\textbf{b.)} Which P-value, 0.095 or 0.04, provides stronger evidence
that chance alone might account for the observed result?

\emph{0.095} provides stronger evidence that chance alone is responsible
for the observed result.

\hypertarget{section-3}{%
\subsection{3.4}\label{section-3}}

In the algorithms for conducting a permutation test, why do we add 1 to
the number of replications N when calculating the P-Value?

\textbf{Answer}: We need to account for the original observed result.

\hypertarget{section-4}{%
\subsection{3.5}\label{section-4}}

In the flight delays case study in Section 1.1, the data contain flight
delays for two airlines, American Airlines and United Airlines.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Flights <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(data.dir, }\StringTok{"FlightDelays.csv"}\NormalTok{),}
                               \DataTypeTok{header =}\NormalTok{ T))}
\end{Highlighting}
\end{Shaded}

a.) Conduct a two-sided permutation test to see if the difference in
mean delay times between the two carriers are statistically significant.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Flights[, .(}\DataTypeTok{Delay =} \KeywordTok{mean}\NormalTok{(Delay)), by =}\StringTok{ }\NormalTok{Carrier]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
   Carrier    Delay
1:      UA 15.98308
2:      AA 10.09738
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "AA"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{nrow}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[index]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{Delay)}
\NormalTok{\}}

\CommentTok{# two-sided test}
\NormalTok{p <-}\StringTok{ }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{)}
\NormalTok{v <-}\StringTok{ }\NormalTok{p}\OperatorTok{*}\NormalTok{(}\DecValTok{1} \OperatorTok{-}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}

\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.table}\NormalTok{(results)) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(results, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Flight Delay Times, UA/AA vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p, }\DecValTok{5}\NormalTok{)),}
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed, }\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-7-1.png}

b.) The flights took place in May and June of 2009. Conduct a two-sided
permutation test to see if the differences in mean delay times between
two months is statistically significant.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Flights[, .(}\DataTypeTok{Delay =} \KeywordTok{mean}\NormalTok{(Delay)), by =}\StringTok{ }\NormalTok{Month]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
   Month     Delay
1:   May  8.884442
2:  June 14.547783
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[Month }\OperatorTok{==}\StringTok{ "May"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[Month }\OperatorTok{==}\StringTok{ "June"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{nrow}\NormalTok{(Flights[Month }\OperatorTok{==}\StringTok{ "May"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] =}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[index]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{Delay)}
\NormalTok{\}}

\CommentTok{# two-sided test}

\NormalTok{p <-}\StringTok{ }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{<=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{)}
\NormalTok{v <-}\StringTok{ }\NormalTok{p}\OperatorTok{*}\NormalTok{(}\DecValTok{1} \OperatorTok{-}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}

\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.table}\NormalTok{(results)) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(results, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Flight Delay Times, May/June, vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p, }\DecValTok{5}\NormalTok{)),}
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed, }\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-8-1.png}

\hypertarget{section-5}{%
\subsection{3.6}\label{section-5}}

In the flight delays case study in Section 1.1, the data contains flight
delays for two airlines, American and United.

a.) Compute the proportion of times that each carrier's flight was
delays more than 20 min.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Flights[, .(}\DataTypeTok{Delay20 =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{.N), by =}\StringTok{ }\NormalTok{Carrier]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
   Carrier   Delay20
1:      UA 0.2128228
2:      AA 0.1693049
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{, .(}\DataTypeTok{Delay =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{)}\OperatorTok{/}\NormalTok{.N)] }\OperatorTok{-}\StringTok{ }\NormalTok{Flights[Carrier }\OperatorTok{==}\StringTok{ "AA"}\NormalTok{, .(}\DataTypeTok{Delay =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{)}\OperatorTok{/}\NormalTok{.N)])}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{nrow}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(Flights[index, .(}\DataTypeTok{Delay =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{)}\OperatorTok{/}\NormalTok{.N)] }\OperatorTok{-}\StringTok{ }\NormalTok{Flights[}\OperatorTok{-}\NormalTok{index, .(}\DataTypeTok{Delay =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{)}\OperatorTok{/}\NormalTok{.N)])}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{)}
\NormalTok{v <-}\StringTok{ }\NormalTok{p}\OperatorTok{*}\NormalTok{(}\DecValTok{1} \OperatorTok{-}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}

\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.table}\NormalTok{(results)) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(results, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Flight Delay Times, Over 20 Minutes, vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p, }\DecValTok{5}\NormalTok{)), }
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed, }\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-9-1.png}

\begin{itemize}
\tightlist
\item
  Conduct a two-sided test to see if the difference in these proportions
  is statistically significant.
\end{itemize}

\textbf{Answer}: There is statistical significance with a P-value
\textless{} 0.0001.

b.) Compute the variance in the flight delay lengths for each carrier.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Flights[, .(}\DataTypeTok{Variance =} \KeywordTok{var}\NormalTok{(Delay)), by =}\StringTok{ }\NormalTok{Carrier]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
   Carrier Variance
1:      UA 2037.525
2:      AA 1606.457
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{var}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{var}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "AA"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{nrow}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{(Flights[index]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{var}\NormalTok{(Flights[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{Delay)}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{))}
\NormalTok{v <-}\StringTok{ }\NormalTok{p}\OperatorTok{*}\NormalTok{(}\DecValTok{1} \OperatorTok{-}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}

\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.table}\NormalTok{(results)) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(results, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Flight Delay Time Variance, vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p, }\DecValTok{5}\NormalTok{)), }
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed, }\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-10-1.png}

\begin{itemize}
\tightlist
\item
  Conduct a test to see if the variance for United Airlines differes
  from that of American Airlines.
\end{itemize}

\textbf{Answer}: There does not appear to be a statistically significant
difference in the variance in delay times between airlines.

\newpage

\hypertarget{section-6}{%
\subsection{3.7}\label{section-6}}

In the flight delays case study in Section 1.1, repeat Excercise 3.5
part \textbf{(a)} using three test statistics,

\begin{itemize}
\tightlist
\item
  i.) The mean of the United Airline delay times
\item
  ii.) The sum of the United Airline delay times
\item
  iii.) The difference in the means
\end{itemize}

Compare the P-values.

Make sure all three test statistics are computed within the same
\textbf{for} loop.

What do you observe?

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{UA.Delay <-}\StringTok{ }\NormalTok{Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay}
\NormalTok{AA.Delay <-}\StringTok{ }\NormalTok{Flights[Carrier }\OperatorTok{==}\StringTok{ "AA"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay}

\NormalTok{observed.mean <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(UA.Delay)}
\NormalTok{observed.sum <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(UA.Delay)}
\NormalTok{observed.diff <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(UA.Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(AA.Delay)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{nrow =}\NormalTok{ N, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{length}\NormalTok{(UA.Delay), }\DataTypeTok{replace =}\NormalTok{ F)}

\NormalTok{   results[i, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[index]}\OperatorTok{$}\NormalTok{Delay)}
\NormalTok{   results[i, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(Flights[index]}\OperatorTok{$}\NormalTok{Delay)}
\NormalTok{   results[i, }\DecValTok{3}\NormalTok{] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[index]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{Delay)}
\NormalTok{\}}

\NormalTok{dt.results <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(results)}
\KeywordTok{colnames}\NormalTok{(dt.results) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"Mean"}\NormalTok{, }\StringTok{"Sum"}\NormalTok{, }\StringTok{"MeanDiff"}\NormalTok{)}

\NormalTok{p.mean <-}\StringTok{ }\NormalTok{( }\KeywordTok{sum}\NormalTok{( dt.results}\OperatorTok{$}\NormalTok{Mean }\OperatorTok{>=}\StringTok{ }\NormalTok{observed.mean ) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ ) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{)}
\NormalTok{p.sum <-}\StringTok{ }\NormalTok{( }\KeywordTok{sum}\NormalTok{(dt.results}\OperatorTok{$}\NormalTok{Sum }\OperatorTok{>=}\StringTok{ }\NormalTok{observed.sum) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}
\NormalTok{p.diff <-}\StringTok{ }\NormalTok{( }\KeywordTok{sum}\NormalTok{(dt.results}\OperatorTok{$}\NormalTok{MeanDiff }\OperatorTok{>=}\StringTok{ }\NormalTok{observed.diff) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{)}

\NormalTok{p1 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(dt.results) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(Mean, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed.mean, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"United Airlines - Mean Delay Time vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p.mean, }\DecValTok{5}\NormalTok{)), }
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed.mean, }\DecValTok{4}\NormalTok{)))}

\NormalTok{p2 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(dt.results) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(Sum, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed.sum, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"United Airlines - Sum Delay Time vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p.sum, }\DecValTok{5}\NormalTok{)), }
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed.sum, }\DecValTok{4}\NormalTok{)))}

\NormalTok{p3 <-}\StringTok{ }\KeywordTok{ggplot}\NormalTok{(dt.results) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(MeanDiff, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed.diff, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Mean Delay Time Difference vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p.diff, }\DecValTok{5}\NormalTok{)), }
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed.diff, }\DecValTok{4}\NormalTok{)))}

\KeywordTok{grid.arrange}\NormalTok{(p1, p2, p3, }\DataTypeTok{nrow =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-11-1.png}

\newpage

\hypertarget{section-7}{%
\subsection{3.8}\label{section-7}}

In the flight delays case study in Section 1.1,

a.) Find the trimmed mean of the delay times for United Airlines and
American Airlines.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{trim.amount <-}\StringTok{ }\FloatTok{.25}
\NormalTok{UA.trimmed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(UA.Delay, }\DataTypeTok{trim =}\NormalTok{ trim.amount)}
\NormalTok{AA.trimmed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(AA.Delay, }\DataTypeTok{trim =}\NormalTok{ trim.amount)}

\NormalTok{observed <-}\StringTok{ }\NormalTok{UA.trimmed }\OperatorTok{-}\StringTok{ }\NormalTok{AA.trimmed}

\KeywordTok{pretty_kable}\NormalTok{(}\KeywordTok{data.table}\NormalTok{( }\DataTypeTok{UA =}\NormalTok{ UA.trimmed, }\DataTypeTok{AA =}\NormalTok{ AA.trimmed), }\StringTok{"Trimmed Means"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-12}Trimmed Means}
\centering
\begin{tabular}[t]{r|r}
\hline
UA & AA\\
\hline
-0.8 & -2.57\\
\hline
\end{tabular}
\end{table}

b.) Conduct a two-sided test to see if the difference in trimmed means
is statistically significant.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{nrow}\NormalTok{(Flights[Carrier }\OperatorTok{==}\StringTok{ "UA"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[index]}\OperatorTok{$}\NormalTok{Delay, }\DataTypeTok{trim =}\NormalTok{ trim.amount) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Flights[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{Delay, }\DataTypeTok{trim =}\NormalTok{ trim.amount)}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{))}
\NormalTok{v <-}\StringTok{ }\NormalTok{p}\OperatorTok{*}\NormalTok{(}\DecValTok{1} \OperatorTok{-}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}

\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.table}\NormalTok{(results)) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_histogram}\NormalTok{(}\KeywordTok{aes}\NormalTok{(results, }\DataTypeTok{fill =}\NormalTok{ ..count..), }\DataTypeTok{bins =} \DecValTok{30}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =}\NormalTok{ observed, }\DataTypeTok{col =} \StringTok{"darkorange"}\NormalTok{, }\DataTypeTok{linetype =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lwd =} \FloatTok{1.2}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{labels =}\NormalTok{ comma) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Flight Delay Time Trimmed Mean vs Observed, p="}\NormalTok{, }\KeywordTok{round}\NormalTok{(p, }\DecValTok{5}\NormalTok{)), }
        \DataTypeTok{subtitle =} \KeywordTok{paste0}\NormalTok{(}\StringTok{"Observed Value: "}\NormalTok{, }\KeywordTok{round}\NormalTok{(observed, }\DecValTok{4}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-13-1.png}

\hypertarget{section-8}{%
\subsection{3.9}\label{section-8}}

In the flight delays case study in Section 1.1,

a.) Compute the proportion of times the flights in May and in June were
delayed more than 20 min.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{delay20_month <-}\StringTok{ }\NormalTok{Flights[, .(}\DataTypeTok{Delay20 =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{.N), by =}\StringTok{ }\NormalTok{Month]}

\KeywordTok{pretty_kable}\NormalTok{(delay20_month, }\StringTok{"Delayed by more than 20 min, by month"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-14}Delayed by more than 20 min, by month}
\centering
\begin{tabular}[t]{l|r}
\hline
Month & Delay20\\
\hline
May & 0.17\\
\hline
June & 0.20\\
\hline
\end{tabular}
\end{table}

Conduct a two-sided test for statistical significance.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\NormalTok{delay20_month[Month }\OperatorTok{==}\StringTok{ "May"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay20 }\OperatorTok{-}\StringTok{ }\NormalTok{delay20_month[Month }\OperatorTok{==}\StringTok{ "June"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay20}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{nrow}\NormalTok{(Flights[Month }\OperatorTok{==}\StringTok{ "May"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{( Flights[index, .(}\DataTypeTok{Delay =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{.N)] }\OperatorTok{-}\StringTok{ }\NormalTok{Flights[}\OperatorTok{-}\NormalTok{index, .(}\DataTypeTok{Delay =} \KeywordTok{sum}\NormalTok{(Delay }\OperatorTok{>}\StringTok{ }\DecValTok{20}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{.N)] )}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{))}
\NormalTok{v <-}\StringTok{ }\NormalTok{p}\OperatorTok{*}\NormalTok{(}\DecValTok{1} \OperatorTok{-}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}
\end{Highlighting}
\end{Shaded}

P-value: 0.0017, which is statistically significant.

b.) Compute the ratio of the variances in the flight delay times in May
and in June.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{var}\NormalTok{(Flights[Month }\OperatorTok{==}\StringTok{ "May"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay) }\OperatorTok{-}\StringTok{ }\KeywordTok{var}\NormalTok{(Flights[Month }\OperatorTok{==}\StringTok{ "June"}\NormalTok{]}\OperatorTok{$}\NormalTok{Delay)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Flights), }\KeywordTok{nrow}\NormalTok{(Flights[Month }\OperatorTok{==}\StringTok{ "May"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{( Flights[index, .(Delay)] ) }\OperatorTok{-}\StringTok{ }\KeywordTok{var}\NormalTok{( Flights[}\OperatorTok{-}\NormalTok{index, .(Delay)] )}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{))}
\NormalTok{v <-}\StringTok{ }\NormalTok{p}\OperatorTok{*}\NormalTok{(}\DecValTok{1} \OperatorTok{-}\StringTok{ }\NormalTok{p) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}
\end{Highlighting}
\end{Shaded}

Is this evidence that the true ratio is not equal to 1, or could this be
due to chance variability?

The variance appear to be due to random chance, so there does appear to
be a stastical significance between the two months.

Conduct a two-sided test to check.

P-value: \textbf{1}

\hypertarget{section-9}{%
\subsection{3.10}\label{section-9}}

In the black spruce case study in Section 1.10, seedlings were planted
in plots that were either subject to competition (from other plants), or
not.

Use the data set \emph{Spruce} to conduct a test to see if the mean
difference is how much the seedlings grow (in height) over the corse of
the study under these two treatments is stastically significant.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Spruce <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(data.dir, }\StringTok{"Spruce.csv"}\NormalTok{),}
                               \DataTypeTok{header =}\NormalTok{ T))}

\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{( Spruce[Competition }\OperatorTok{==}\StringTok{ "NC"}\NormalTok{]}\OperatorTok{$}\NormalTok{Ht.change ) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{( Spruce[Competition }\OperatorTok{==}\StringTok{ "C"}\NormalTok{]}\OperatorTok{$}\NormalTok{Ht.change )}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{(i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Spruce), }\KeywordTok{nrow}\NormalTok{(Spruce[Competition }\OperatorTok{==}\StringTok{ "NC"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   resutls <-}\StringTok{ }\KeywordTok{mean}\NormalTok{( Spruce[index]}\OperatorTok{$}\NormalTok{Ht.change ) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{( Spruce[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{Ht.change )}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{(}\KeywordTok{sum}\NormalTok{(results[results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed]) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

There is stastical significance in between the heights of the two groups
(Competition / No-competition).

P-value: \textbf{0.002}

\hypertarget{section-10}{%
\subsection{3.11}\label{section-10}}

The file \emph{Phillies2009} contains data from the 2009 season for the
baseball team the Philadelphia Phillies.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Phillies <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(data.dir, }\StringTok{"Phillies2009.csv"}\NormalTok{),}
                               \DataTypeTok{header =}\NormalTok{ T))}
\end{Highlighting}
\end{Shaded}

a.) Compare the empirical distribution functions of the number of
strike-outs per game (\emph{StrikeOuts}) for games played at home and
games played away (\emph{Location}).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(Phillies, }\KeywordTok{aes}\NormalTok{(StrikeOuts, }\DataTypeTok{color =}\NormalTok{ Location)) }\OperatorTok{+}
\StringTok{   }\KeywordTok{stat_ecdf}\NormalTok{(}\DataTypeTok{stat =} \StringTok{"point"}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Strike-outs / Game, by Location"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Percent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: Ignoring unknown parameters: stat
\end{verbatim}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-19-1.png}

b.) Find the mean number of strike-outs per game for the home and the
away games.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{strikeouts <-}\StringTok{ }\NormalTok{Phillies[, .(}\DataTypeTok{StrikeOuts =} \KeywordTok{mean}\NormalTok{(StrikeOuts)), by =}\StringTok{ }\NormalTok{Location]}

\KeywordTok{pretty_kable}\NormalTok{(strikeouts, }\StringTok{"StrikeOuts by Location"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-20}StrikeOuts by Location}
\centering
\begin{tabular}[t]{l|r}
\hline
Location & StrikeOuts\\
\hline
Home & 6.95\\
\hline
Away & 7.31\\
\hline
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(strikeouts[Location }\OperatorTok{==}\StringTok{ "Away"}\NormalTok{]}\OperatorTok{$}\NormalTok{StrikeOuts) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(strikeouts[Location }\OperatorTok{==}\StringTok{ "Home"}\NormalTok{]}\OperatorTok{$}\NormalTok{StrikeOuts)}
\end{Highlighting}
\end{Shaded}

c.) Perform a permutation tests to see if the differences in means is
statistically significant.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N)}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Phillies), }\KeywordTok{nrow}\NormalTok{(Phillies[Location }\OperatorTok{==}\StringTok{ "Home"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Phillies[index]}\OperatorTok{$}\NormalTok{StrikeOuts) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Phillies[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{StrikeOuts)}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, ( }\KeywordTok{sum}\NormalTok{(results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) )}
\end{Highlighting}
\end{Shaded}

There does not appear to be a statistically significat relationship
between the number of home and away strikeouts.

P-value: \textbf{0.192}

\hypertarget{section-11}{%
\subsection{3.12}\label{section-11}}

In the Iowa recidivism case study in Section 1.4, offenders had
originally been convicted of either a felony or misdemenor.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Recidivism <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(data.dir, }\StringTok{"Recidivism.csv"}\NormalTok{),}
                               \DataTypeTok{header =}\NormalTok{ T))}
\end{Highlighting}
\end{Shaded}

a.) Use R to create a table displaying the proportion of felons who
recidivated and the propportion of those convictted of a misdemeanor who
recidivated.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Recidivism[, Recidivated }\OperatorTok{:}\ErrorTok{=}\StringTok{ }\KeywordTok{is.na}\NormalTok{(Days)]}

\NormalTok{Recidivated <-}\StringTok{ }\NormalTok{Recidivism[, .(}\DataTypeTok{RecidPct =} \KeywordTok{sum}\NormalTok{(Recidivated) }\OperatorTok{/}\StringTok{ }\NormalTok{.N), by =}\StringTok{ }\NormalTok{Offense]}

\KeywordTok{pretty_kable}\NormalTok{(Recidivated, }\StringTok{"Recidivated"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-23}Recidivated}
\centering
\begin{tabular}[t]{l|r}
\hline
Offense & RecidPct\\
\hline
Felony & 0.69\\
\hline
Misdemeanor & 0.67\\
\hline
\end{tabular}
\end{table}

b.) Determine whether or not the difference in recidivism proportions
computed in (a) is statistically significant.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{( Recidivated[Offense }\OperatorTok{==}\StringTok{ "Felony"}\NormalTok{]}\OperatorTok{$}\NormalTok{RecidPct }\OperatorTok{-}\StringTok{ }\NormalTok{Recidivated[Offense }\OperatorTok{==}\StringTok{ "Misdemeanor"}\NormalTok{]}\OperatorTok{$}\NormalTok{RecidPct )}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N )}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Recidivism), }\KeywordTok{nrow}\NormalTok{(Recidivism[Offense }\OperatorTok{==}\StringTok{ "Felony"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{( Recidivism[index, .(}\DataTypeTok{Pct =} \KeywordTok{sum}\NormalTok{(Recidivated) }\OperatorTok{/}\StringTok{ }\NormalTok{.N)] }\OperatorTok{-}\StringTok{ }\NormalTok{Recidivism[}\OperatorTok{-}\NormalTok{index, .(}\DataTypeTok{Pct =} \KeywordTok{sum}\NormalTok{(Recidivated) }\OperatorTok{/}\StringTok{ }\NormalTok{.N)] )}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\NormalTok{( }\KeywordTok{sum}\NormalTok{(results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ ) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}
\end{Highlighting}
\end{Shaded}

It appears there is statistical significance in the recidivism rates
between felony and misdemeanor convictions.

P-value: 0.028

\hypertarget{section-12}{%
\subsection{3.13}\label{section-12}}

In the Iowa recidivism case study in Section 1.4, for those offenders
who recidivated, we have data on the number of days until they
reoffended.

For those offenders who did recidivate, determine if the difference in
the mean number of days (\textbf{Days}) until recidivism between those
under 25 years of age and those 25 years of age and older is
statistically significant.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Recid.Age25 <-}\StringTok{ }\NormalTok{Recidivism[, .(}\DataTypeTok{RecidDays =} \KeywordTok{mean}\NormalTok{(Days, }\DataTypeTok{na.rm =}\NormalTok{ T)), by =}\StringTok{ }\NormalTok{Age25]}

\KeywordTok{pretty_kable}\NormalTok{(Recid.Age25, }\StringTok{"Days Until Recidivism by Age ( < 25 < )"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-25}Days Until Recidivism by Age ( < 25 < )}
\centering
\begin{tabular}[t]{l|r}
\hline
Age25 & RecidDays\\
\hline
Under 25 & 449.07\\
\hline
Over 25 & 479.72\\
\hline
NA & NaN\\
\hline
\end{tabular}
\end{table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{observed <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{( Recid.Age25[Age25 }\OperatorTok{==}\StringTok{ "Under 25"}\NormalTok{]}\OperatorTok{$}\NormalTok{RecidDays }\OperatorTok{-}\StringTok{ }\NormalTok{Recid.Age25[Age25 }\OperatorTok{==}\StringTok{ "Over 25"}\NormalTok{]}\OperatorTok{$}\NormalTok{RecidDays)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N )}
\NormalTok{\{}
\NormalTok{   index <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{nrow}\NormalTok{(Recidivism), }\KeywordTok{nrow}\NormalTok{(Recidivism[Age25 }\OperatorTok{==}\StringTok{ "Under 25"}\NormalTok{]), }\DataTypeTok{replace =}\NormalTok{ F)}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Recidivism[index]}\OperatorTok{$}\NormalTok{Days, }\DataTypeTok{na.rm =}\NormalTok{ T) }\OperatorTok{-}\StringTok{ }\KeywordTok{mean}\NormalTok{(Recidivism[}\OperatorTok{-}\NormalTok{index]}\OperatorTok{$}\NormalTok{Days, }\DataTypeTok{na.rm =}\NormalTok{ T)}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, ( }\KeywordTok{sum}\NormalTok{(results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ ) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ ) )}
\end{Highlighting}
\end{Shaded}

There does not appear to be a statistical difference between the days
until recidivism by those over/under 25 years of age.

P-value: 0.998

\hypertarget{section-13}{%
\subsection{3.14}\label{section-13}}

Does chocolate ice cream have more calories than vanilla ice cream? The
data set \textbf{IceCream} contains calorie information for a sample of
brands of chocolate and vanilla ice cream.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{IceCream <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(data.dir, }\StringTok{"IceCream.csv"}\NormalTok{),}
                               \DataTypeTok{header =}\NormalTok{ T))}
\end{Highlighting}
\end{Shaded}

a.) Inspect the data set, then explain why this is an example of matched
pairs data.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{pretty_kable}\NormalTok{(IceCream, }\StringTok{"Ice Cream Data"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-27}Ice Cream Data}
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r}
\hline
Brand & VanillaCalories & VanillaFat & VanillaSugar & ChocolateCalories & ChocolateFat & ChocolateSugar\\
\hline
Baskin Robbins & 260 & 16.0 & 26.0 & 260 & 14.0 & 31.0\\
\hline
Ben \& Jerry's & 240 & 16.0 & 19.0 & 260 & 16.0 & 22.0\\
\hline
Blue Bunny & 140 & 7.0 & 12.0 & 130 & 7.0 & 14.0\\
\hline
Breyers & 140 & 7.0 & 13.0 & 140 & 8.0 & 16.0\\
\hline
Brigham's & 190 & 12.0 & 17.0 & 200 & 12.0 & 18.0\\
\hline
Bulla & 234 & 13.5 & 21.8 & 266 & 15.0 & 22.6\\
\hline
Carvel & 240 & 14.0 & 21.0 & 250 & 13.0 & 25.0\\
\hline
Cass-Clay & 130 & 7.0 & 11.0 & 150 & 7.0 & 16.0\\
\hline
Chapman's & 120 & 6.0 & 11.0 & 120 & 5.0 & 12.0\\
\hline
Cold Stone & 270 & 15.5 & 23.0 & 264 & 16.2 & 23.6\\
\hline
Culver's & 222 & 13.0 & 19.0 & 205 & 10.0 & 20.0\\
\hline
Dairy Queen & 140 & 4.5 & 19.0 & 150 & 5.0 & 17.0\\
\hline
Dove & 240 & 15.0 & 20.0 & 290 & 17.0 & 27.0\\
\hline
Dreamery & 260 & 15.0 & 24.0 & 280 & 12.0 & 33.0\\
\hline
Edy's Grand & 140 & 8.0 & 13.0 & 150 & 8.0 & 15.0\\
\hline
Emack \& Bolio's & 160 & 9.0 & 12.0 & 170 & 9.0 & 13.0\\
\hline
Good Humor & 120 & 6.0 & 12.0 & 120 & 6.0 & 14.0\\
\hline
Graeter's & 260 & 16.0 & 24.0 & 260 & 16.0 & 24.0\\
\hline
Green and Black & 194 & 11.6 & 18.0 & 227 & 12.8 & 22.7\\
\hline
Green's & 150 & 8.0 & 17.0 & 140 & 8.0 & 15.0\\
\hline
Haagen Dazs & 270 & 18.0 & 21.0 & 270 & 18.0 & 21.0\\
\hline
Hershey's & 140 & 9.0 & 14.0 & 140 & 8.0 & 13.0\\
\hline
Hill Station & 226 & 15.6 & 16.8 & 235 & 14.3 & 21.2\\
\hline
Kemp's & 130 & 7.0 & 13.0 & 140 & 6.0 & 17.0\\
\hline
Klein's & 130 & 8.0 & 15.0 & 140 & 8.0 & 14.0\\
\hline
Oberweis Dairy & 307 & 21.0 & 23.0 & 320 & 21.0 & 19.0\\
\hline
Our Family & 130 & 7.0 & 11.0 & 130 & 6.0 & 15.0\\
\hline
Perry's & 140 & 8.0 & 15.0 & 140 & 7.0 & 15.0\\
\hline
Ronnybrook Farm & 240 & 16.0 & 20.0 & 260 & 19.0 & 21.0\\
\hline
Ruggles & 150 & 8.0 & 12.0 & 150 & 8.0 & 16.0\\
\hline
Sara Lee & 242 & 15.5 & 21.5 & 234 & 14.4 & 20.9\\
\hline
Schwan's & 140 & 7.0 & 12.0 & 140 & 7.0 & 12.0\\
\hline
Sheer Bliss & 300 & 19.0 & 27.0 & 320 & 19.0 & 29.0\\
\hline
Smith's & 150 & 8.0 & 13.0 & 150 & 8.0 & 13.0\\
\hline
Stonyfield Farm & 240 & 16.0 & 20.0 & 250 & 17.0 & 20.0\\
\hline
Tillamook & 160 & 9.0 & 10.0 & 170 & 9.0 & 13.0\\
\hline
Turkey Hill & 140 & 8.0 & 16.0 & 150 & 8.0 & 19.0\\
\hline
Value Choice & 130 & 6.0 & 12.0 & 130 & 6.0 & 15.0\\
\hline
Whitey's & 250 & 14.0 & 23.0 & 250 & 13.0 & 25.0\\
\hline
\end{tabular}
\end{table}

This data set is matched pairs becase there are multiple brands
associated with each flavor. They are not independent samples.

b.) Compute summary statistics of the number of calories for the two
flavors.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{pretty_kable}\NormalTok{(IceCream[, .(}\DataTypeTok{Vanilla =} \KeywordTok{mean}\NormalTok{(VanillaCalories), }\DataTypeTok{Chocolate =} \KeywordTok{mean}\NormalTok{(ChocolateCalories))], }
             \StringTok{"Ice cream Calories by Flavor"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-28}Ice cream Calories by Flavor}
\centering
\begin{tabular}[t]{r|r}
\hline
Vanilla & Chocolate\\
\hline
191.41 & 198.74\\
\hline
\end{tabular}
\end{table}

c.) Conduct a permutation tests to determine whether or not chocolate
ice cream has, on average, more calories than vanilla ice cream.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{diff <-}\StringTok{ }\NormalTok{IceCream}\OperatorTok{$}\NormalTok{VanillaCalories }\OperatorTok{-}\StringTok{ }\NormalTok{IceCream}\OperatorTok{$}\NormalTok{ChocolateCalories}
\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(diff)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N )}
\NormalTok{\{}
\NormalTok{   Sign <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\KeywordTok{nrow}\NormalTok{(IceCream), }\DataTypeTok{replace =}\NormalTok{ T)}
\NormalTok{   diff2 <-}\StringTok{ }\NormalTok{Sign }\OperatorTok{*}\StringTok{ }\NormalTok{diff}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(diff2)}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\NormalTok{( }\KeywordTok{sum}\NormalTok{( results }\OperatorTok{<=}\StringTok{ }\NormalTok{observed ) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ ) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ )}
\end{Highlighting}
\end{Shaded}

The data supports the claim that chocolate ice cream has more calories
than vanilla.

P-value: 0.002

\hypertarget{section-14}{%
\subsection{3.15}\label{section-14}}

Is there a difference in the price of groceries sold by the two
retailers Target and Walmart?

The data set Groceries contains a sample of grocery items and their
prices advertised on their respective web sites on one specific day.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Groceries <-}\StringTok{ }\KeywordTok{data.table}\NormalTok{(}\KeywordTok{read.csv}\NormalTok{(}\KeywordTok{paste0}\NormalTok{(data.dir, }\StringTok{"Groceries.csv"}\NormalTok{),}
                               \DataTypeTok{header =}\NormalTok{ T))}
\end{Highlighting}
\end{Shaded}

a.) Inspect the data set, then explain why this is an example of matched
pairs data.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{pretty_kable}\NormalTok{(Groceries, }\StringTok{"Target vs Walmart"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-31}Target vs Walmart}
\centering
\begin{tabular}[t]{l|l|r|r}
\hline
Product & Size & Target & Walmart\\
\hline
Kellogg NutriGrain Bars & 8 bars & 2.50 & 2.78\\
\hline
Quaker Oats Life Cereal  Original & 18oz & 3.19 & 6.01\\
\hline
General Mills Lucky Charms & 11.50z & 3.19 & 2.98\\
\hline
Quaker Oats Old Fashioned & 18oz & 2.82 & 2.68\\
\hline
Nabisco Oreo Cookies & 14.3oz & 2.99 & 2.98\\
\hline
Nabisco Chips Ahoy & 13oz & 2.64 & 1.98\\
\hline
Doritos Nacho Cheese Chips & 10oz & 3.99 & 2.50\\
\hline
Cheez-it Original Baked & 21oz & 4.79 & 4.79\\
\hline
Swiss Miss Hot Chocolate & 10 count & 1.49 & 1.28\\
\hline
Tazo Chai Classic Latte Black Tea & 32 oz & 3.49 & 2.98\\
\hline
Annie's Macaroni \& Cheese & 6oz & 1.79 & 1.72\\
\hline
Rice A Roni Chicken & 6.9oz & 1.00 & 1.00\\
\hline
Zatarain's Jambalaya Rice Mix & 8oz & 1.62 & 1.54\\
\hline
SPAM Original Lunch Meat & 12oz & 2.79 & 2.64\\
\hline
Campbell's Chicken Noodle Soup & 10.75oz & 0.99 & 1.58\\
\hline
Dinty Moore Hearty Meals Beef Stew & 15oz & 1.99 & 1.98\\
\hline
Hormel  Chili with Beans & 15oz & 1.94 & 1.88\\
\hline
Dole Pineapple Chunks & 20 oz & 1.59 & 1.47\\
\hline
Skippy Creamy Peanut Butter & 16.3oz & 2.59 & 2.58\\
\hline
Smucker's Strawberry Preserve & 18oz & 2.99 & 2.84\\
\hline
Heinz Tomato Ketchup & 32oz & 2.99 & 2.88\\
\hline
Near East Couscous Toasted Pine Nuts mix & 5.6oz & 2.12 & 1.98\\
\hline
Barilla Angel Hair Pasta & 16oz & 1.42 & 1.38\\
\hline
Betty Crocker Super Moist Chocolate Fudge Cake Mix & 15.25oz & 1.22 & 1.17\\
\hline
Kraft Jet-Puffed Marshmllows & 16oz & 1.99 & 1.96\\
\hline
Dunkin' Donuts Original Blend Medium Roast Ground Coffee & 12oz & 7.19 & 6.98\\
\hline
Dove Promises Milk Chocolate & 8.87oz & 3.19 & 3.50\\
\hline
Skittles & 41oz & 7.99 & 6.98\\
\hline
Vlasic Kosher Dill Pickle Spears & 24oz & 2.39 & 2.18\\
\hline
Vlasic Old Fashioned Sauerkraut & 32oz & 1.99 & 1.97\\
\hline
\end{tabular}
\end{table}

b.) Compute summary statistics of the prices for each store.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{diff <-}\StringTok{ }\NormalTok{Groceries}\OperatorTok{$}\NormalTok{Target }\OperatorTok{-}\StringTok{ }\NormalTok{Groceries}\OperatorTok{$}\NormalTok{Walmart}

\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(diff)}
\end{Highlighting}
\end{Shaded}

Observed difference: \textbf{0.0566667}

c.) Conduct a permutation test to determine whether or not there is a
difference in the mean prices.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N )}
\NormalTok{\{}
\NormalTok{   Sign <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\KeywordTok{nrow}\NormalTok{(Groceries), }\DataTypeTok{replace =}\NormalTok{ T)}
\NormalTok{   diff2 <-}\StringTok{ }\NormalTok{Sign }\OperatorTok{*}\StringTok{ }\NormalTok{diff}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(diff2)}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{( }\KeywordTok{sum}\NormalTok{(results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed ) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ ) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) )}
\end{Highlighting}
\end{Shaded}

There does not appear to be a statistical difference in the prices at
the two stores.

P-value: 0.672

d.) Create a histogram of the difference in prices.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{data.table}\NormalTok{( }\DataTypeTok{Product =}\NormalTok{ Groceries}\OperatorTok{$}\NormalTok{Product, }\DataTypeTok{Diff =}\NormalTok{ diff)) }\OperatorTok{+}
\StringTok{   }\KeywordTok{geom_bar}\NormalTok{(}\KeywordTok{aes}\NormalTok{(Product, Diff, }\DataTypeTok{fill =}\NormalTok{ Product), }\DataTypeTok{stat =} \StringTok{"identity"}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{coord_flip}\NormalTok{() }\OperatorTok{+}
\StringTok{   }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{title =} \StringTok{"Grocery Price Difference (Target vs Walmart)"}\NormalTok{) }\OperatorTok{+}
\StringTok{   }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{03_permutation_testing_files/figure-latex/unnamed-chunk-34-1.png}

What is unusual about Quaker Oats Life cerial?

\textbf{Price difference is a multiple SD outlier.}

e.) Redo the hypothesis test without this observation.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{diff <-}\StringTok{ }\NormalTok{Groceries[Product }\OperatorTok{!=}\StringTok{ "Quaker Oats Life Cereal  Original "}\NormalTok{, .(}\DataTypeTok{Diff =}\NormalTok{ Target }\OperatorTok{-}\StringTok{ }\NormalTok{Walmart ) ]}\OperatorTok{$}\NormalTok{Diff}

\NormalTok{observed <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(diff)}

\NormalTok{N <-}\StringTok{ }\FloatTok{10e2} \OperatorTok{-}\StringTok{ }\DecValTok{1}
\NormalTok{results <-}\StringTok{ }\KeywordTok{numeric}\NormalTok{(N)}

\ControlFlowTok{for}\NormalTok{( i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\NormalTok{N )}
\NormalTok{\{}
\NormalTok{   Sign <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{), }\KeywordTok{length}\NormalTok{(diff), }\DataTypeTok{replace =}\NormalTok{ T)}
\NormalTok{   diff2 <-}\StringTok{ }\NormalTok{Sign }\OperatorTok{*}\StringTok{ }\NormalTok{diff}
\NormalTok{   results[i] <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(diff2)}
\NormalTok{\}}

\NormalTok{p <-}\StringTok{ }\KeywordTok{min}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2} \OperatorTok{*}\StringTok{ }\NormalTok{( }\KeywordTok{sum}\NormalTok{(results }\OperatorTok{>=}\StringTok{ }\NormalTok{observed ) }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{ ) }\OperatorTok{/}\StringTok{ }\NormalTok{( N }\OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{) )}
\end{Highlighting}
\end{Shaded}

Do you reach the same conclusion?

\textbf{No, there is a statistical difference in the prices of products
at the two retailers with the outlier removed.}

P-value: 0.036

\hypertarget{section-15}{%
\subsection{3.16}\label{section-15}}

In the sampling version of permutation testing, the one-sided P-value is

\(\hat{P} = \frac{(X + 1)}{(N + 1)}\), where X is the number of
permutation test statistics that are as large or larger than the
observed test statistic.

Suppose the true P-value (for the exaustive test, conditional on the
observed data) is \textbf{p}.

a.) What is the variance of \(\hat{P}\)?

\(\mathbb{V} = p*(1 - p) / (N + 1)\)

b.) What is the variance of \(\hat{P}_2\) for the two-sided test
(assuming that \textbf{p} is not close to 0.5, where \textbf{p} is the
smaller true one-side P-value)?

\end{document}

### 2019-12-16 Statistical Methods/05_sampling-distributions.rmd
statistical methods
statistical methods

@@
 
-setwd("D:/Projects/Statistical-Computing/RDS")
-
 ```
@@
 ### 5.1
+
+For a binomial with $n = 25$ and $p = 0.5$, determine:
+
+a.) $P(\hat{p} \le 15/25)$
+
+```{r, echo = T}
+pbinom(25 * 15/25, size = 25, prob = .5)
+```
+
+b.) $P(\hat{p} > 15/25)$
+
+```{r, echo = T}
+pbinom(15/25 * 25, size = 25, prob = .5, lower.tail = F)
+```
+
+c.) $P(10/25 \le \hat{p} \le 15/25)$
+
+```{r, echo = T}
+pbinom(25 * 15/25, size = 25, prob = .5) - pbinom(25 * 10/25, size = 25, prob = .5)
+```
+
+### 5.2
+
+Many research teams intend to conduct a study regarding the proportion of people who have colon cancer.
+If a random sample of 10 individuals could be obtained, and if the probability of having colon cancer is 0.05, what is the probability that a research team will get $\hat{p} = 0.1$?
+
+```{r, echo = T}
+dbinom( 10 * .1, size = 10, prob = 0.05)
+```
+
+### 5.3
+
+In the previous problem, what is the probability of $\hat{p} = 0.05$?
+
+```{r, echo = T}
+dbinom( 10 * .05, size = 10, prob = 0.05)
+```
+
+### 5.4
+
+Someone claims that the probability of losing money, when using an investment strategy for buying and selling commodities, is 0.1.
+If this claim is correct, what is the probability of getting $\hat{p} \le 0.05$ based on a random sample of 25 investments?
+
+```{r, echo = T}
+pbinom( 25 * .05, size = 25, prob = .1)
+```
+
+### 5.5
+
+You interview a married couple and ask the wife whether she supports the current leader of their country.
+Her husband is asked the same question.
+Describe why it might be unreasonable to view these two responses as a random sample.
+
+_The liklihood of the two married people having the same political views is high._
+
+### 5.6
+
+Inagine that 1,000 research teams draw a random sample from a binomial distribution with $p = 0.4$, with each study based on a sample size of 30.
+So, this would result in 1,000 $\hat{p}$ values.
+If these 1,000 values were averages, what, approximately, would be the result?
+
+```{r, echo = T}
+p <- .4
+n <- 30
+
+N <- 1e3
+result <- numeric(N)
+for( i in 1:N )
+{
+   result[i] <- mean( rbinom(n, size = 1, prob = p) )
+}
+
+mean(result)
+
+```
+
+### 5.7
+
+In the previous problem, if you computed the sample variance of the $\hat{p}$ values, what, approximatley, would be the result?
+
+```{r, echo = T}
+N <- 1e3
+result <- numeric(N)
+for( i in 1:N )
+{
+   result[i] <- mean( rbinom(n, size = 1, prob = p) )
+}
+
+var(result) # estimated
+
+(p * (1 - p)) / n # exact
+```
+
+### 5.8
+
+Suppose $n = 16, \sigma = 2, \mu = 30$.
+
+Assume normality and determine:
+
+a.) $P(\bar{X} \le 29)$
+
+```{r, echo = T}
+pnorm( 29, mean = 30, sd = sqrt(2^2 / 16))
+```
+
+b.) $P(\bar{X} \ge 30.5)$
+
+```{r, echo = T}
+pnorm( 30.5, mean = 30, sd = sqrt(2^2 / 16), lower.tail = F)
+```
+
+c.) $P( 29 \le \bar{X} \le 31)$
+
+```{r, echo = T}
+pnorm(31, mean = 30, sd = sqrt(2^2 / 16)) - pnorm( 29, mean = 30, sd = sqrt(2^2 / 16))
+```
+
+### 5.9
+
+Suppose $n = 25, \sigma = 5, \mu = 5$
+
+Assume normality and determine:
+
+a.) $P(\bar{X} \le 4)$
+
+```{r, echo = T}
+
+```
+
+b.) $P(\bar{X} \le 7)$
+
+```{r, echo = T}
+
+```
+
+c.) $P(3 \le \bar{X} \le 7)$
+
+```{r, echo = T}
+
+```

### 2019-12-16 packages.R

@@ -1,0 +1,15 @@
+install.packages('tinytex')
+
+tinytex::install_tinytex()
+
+# tinytex:::is_tinytex()
+
+library(tinytex)
+tlmgr_search('/times.sty')   # search for times.sty
+tlmgr_install('psnfss')      # install the psnfss package
+tlmgr_update()               # update everything
+
+
+tinytex::install_tinytex()
+
+tinytex::uninstall_tinytex()

### 2019-12-16 Statistical Methods/05_sampling-distributions.rmd
statistical methods
statistical methods

@@
 ```{r, echo = T}
-
+pnorm(4, mean = 5, sd = sqrt(5^2 / 25))
 ```
@@
 ```{r, echo = T}
-
+pnorm(7, mean = 5, sd = sqrt(5^2 / 25), lower.tail = F)
 ```
@@
 ```{r, echo = T}
+pnorm(7, mean = 5, sd = sqrt(5^2 / 25)) - pnorm(3, mean = 5, sd = sqrt(5^2 / 25))
+```
+
+### 5.10
+
+Someone claims that within a certain neighborhood, the average cost of a house is $\mu = \$100,000$ with a standard deviation $\sigma = \$10,000$.
+Suppose that based on n = 16 homes, you find that the average cost of a house is $\bar{X} = \$95,000$.
+Assuming normality, what is the probability of getting a sample mean this low or lower if the claims about the mean and standard deviation are true?
+
+```{r, echo = T}
+pnorm(95e3, mean = 1e5, 1e4 / sqrt(16) )
+```
+
+### 5.11
+
+In the previous problem, what is the probabiility of getting a sample mean between $97,500 and $102,500?
+
+```{r, echo = T}
+pnorm(1025e2, mean = 1e5, 1e4 / sqrt(16), lower.tail = T) - pnorm(975e2, mean = 1e5, 1e4 / sqrt(16))
+```
+
+### 5.12
+
+A company claims that the premiums paid by its clients for auto insurance has a normal distribution with mean $\mu = \$750$ dollars and standard deviation $\sigma = 100$.
+
+Assuming normality, what is the probability that for $n = 9$ randomly sampled clients, the sample mean will have a value between $700 and $800?
+
+```{r, echo = T}
+s <- 100 / sqrt(9)
+
+pnorm(800, mean = 750, sd = s) - pnorm(700, mean = 750, sd = s)
+```
+
+### 5.13
+
+Imagine you are a health professional interested in the effects of medication on the diastolic blood pressure of adult women.
+For a particular drug being investigated, you find that for $n = 9$ women, the sample mean is $\bar{X} = 85$ and the sample variance is $s^2 = 160.78$.
+
+Estimate the standard error of the sample mean, assuming random sampling.
+
+```{r, echo = T}
+sqrt(160.78 / 9)
+```
+
+### 5.14
+
+You randomly sample 16 observations from a discrete distribution with mean $\mu = 36$ and variance $\sigma^2 = 25$.
+
+```{r, echo = T}
+n <- 16; mu <- 36; variance <- 25
+
+s <- sqrt(variance / n)
 
 ```
+
+Use the central limit therom to determine:
+
+a.) $P(\bar{X} \le 34)$
+
+```{r, echo = T}
+pnorm(34, mean = mu, sd = s)
+```
+
+b.) $P(\bar{X} \le 37)$
+
+```{r, echo = T}
+pnorm(37, mean = mu, sd = s)
+```
+
+c.) $P(\bar{X} \ge 33)$
+
+```{r, echo = T}
+pnorm(33, mean = mu, sd = s, lower.tail = F)
+```
+
+d.) $P(34 \le \bar{X} \le 37)$
+
+```{r, echo = T}
+pnorm(37, mean = mu, sd = s) - pnorm(34, mean = mu, sd = s)
+```
+
+### 5.15
+
+You sample 25 observations from a nonnormal distribution with mean $\mu = 25$ and variance $\sigma^2 = 9$.
+
+Use the central limit therom to determine:
+
+a.) $P(\bar{X} \le 24)$
+
+```{r, echo = T}
+```
+
+b.) $P(\bar{X} \le 26)$
+
+```{r, echo = T}
+```
+
+c.) $P(\bar{X} \le 24)$
+
+```{r, echo = T}
+```
+
+d.) $P(24 \le \bar{X} \le 26)$
+
+```{r, echo = T}
+```
+
+### 5.16
+
+Describe a situation where reliance on the central limit theorm to determine $P(\bar{X} \le 24)$ might be unsatisfactory.
+
+### 5.17
+
+Describe situations where a normal distribution provides a good approximation of the sampling distribution of the mean.
+
+### 5.18
+
+For the values 4, 8, 23, 43, 12, 11, 32, 15, 6, 29, verify that the McKean-Schrader estimate of the standard error of the mean is 7.57.
+
+### 5.19
+
+In the previous exercise, how would you argue that the method used to estimate the standard error of the median is a reasonable approach?
+
+### 5.20
+
+For the values 5, 7, 2, 3, 4, 5, 2, 6, 7, 3, 4, 6, 1, 7, 4, verify that the McKean-Schrader estimate of the standard error of the median is 0.97.
+
+### 5.21
+
+In the previous excersize, how would you argue that the method used to estimate the standard error of th median might be highly inaccurate?
+
+### 5.22
+
+In Exercise 20, would it be advisable to approximate $P(M \le 4)$, the probability that the sample median is less than or equal to 4, using Equation (5.7)?
+
+### 5.23
+
+For the values, 2, 3, 5, 6, 8, 12, 14, 18, 19, 22, 201, why would you suspect that the McKean-Schrader estiamte of the standard error of the median will be smaller than the standard error of the mean?

### 2019-12-17 Statistical Methods/05_sampling-distributions.rmd
statistical methods, ch 5
statistical methods, ch 5

@@
 
+```{r, echo = T}
+n <- 25; mu <- 25; var <- 9
+
+s <- sqrt(var / n)
+```
+
 Use the central limit therom to determine:
@@
 ```{r, echo = T}
+pnorm(24, mean = mu, sd = s)
 ```
@@
 ```{r, echo = T}
+pnorm(26, mean = mu, sd = s)
 ```
 
-c.) $P(\bar{X} \le 24)$
+c.) $P(\bar{X} \ge 24)$
 
 ```{r, echo = T}
+pnorm(24, mean = mu, sd = s, lower.tail = F)
 ```
@@
 ```{r, echo = T}
+pnorm(26, mean = mu, sd = s) - pnorm(24, mean = mu, sd = s)
 ```
@@
 
+_Skewed, heavy-tailed distributions (mixed normal)._
+
 ### 5.17
@@
 Describe situations where a normal distribution provides a good approximation of the sampling distribution of the mean.
+
+_Symetric, light-tailed distributions._
 
@@
 
+```{r, echo = T}
+x <- c(4, 8, 23, 43, 12, 11, 32, 15, 6, 29)
+
+msmedse <- function(x) {
+   n <- length(x)
+   vals <- sort(x)
+
+   k <- floor( (n + 1)/2 - 2.5758*sqrt(n/4) )
+   m <- ((vals[n - k + 1] - vals[k]) / 5.1517)^2
+
+   sqrt(m)
+}
+
+msmedse(x)
+```
+
 ### 5.19
@@
 In the previous exercise, how would you argue that the method used to estimate the standard error of the median is a reasonable approach?
+
+_There are not any tied values._
 
@@
 
+```{r, echo = T}
+x <- c(5, 7, 2, 3, 4, 5, 2, 6, 7, 3, 4, 6, 1, 7, 4)
+msmedse(x)
+```
+
 ### 5.21
@@
 In the previous excersize, how would you argue that the method used to estimate the standard error of th median might be highly inaccurate?
+
+_Several tied values._
 
@@
 
+_No, there are ties._
+
 ### 5.23
 
-For the values, 2, 3, 5, 6, 8, 12, 14, 18, 19, 22, 201, why would you suspect that the McKean-Schrader estiamte of the standard error of the median will be smaller than the standard error of the mean?
+For the values, 2, 3, 5, 6, 8, 12, 14, 18, 19, 22, 201, why would you suspect that the McKean-Schrader estimate of the standard error of the median will be smaller than the standard error of the mean?
+
+_There is an outlier._
+
+```{r, echo = T}
+x <- c(2, 3, 5, 6, 8, 12, 14, 18, 19, 22, 201)
+
+msmedse(x)
+
+sqrt(var(x) / length(x))
+
+```
+
+### 5.24
+
+Summarize when it would and would not be reasonable to assume that the sampling distribution of M, the sample median, is normal.
+
+_When the distribution is not symetric the sampling distribution of M will not be normal._
+
+### 5.25
+
+For the values:
+
+59, 106, 174, 207, 219, 237, 313, 365, 485, 497, 515, 529, 557, 615, 625, 645, 973, 1065, 3215
+
+Estimate the standard error of the 20% trimmed mean.
+
+```{r, echo = T}
+x <- c(59, 106, 174, 207, 219, 237, 313, 365, 485, 497, 515, 529, 557, 615, 625, 645, 973, 1065, 3215)
+
+winsorize <- function( x, trim = .2) {
+   n <- length(x)
+
+   o <- sort(x)
+   g <- floor(trim*n)
+
+   o[1:(g+1)] <- o[(g+1)]
+   o[(n-g):n] <- o[n-g]
+
+   o
+}
+
+n <- length(x)
+
+sqrt( var(winsorize(x)) / ( .36 * n ) )
+
+```
+
+### 5.26
+
+For the data in Exercise 25, why would you suspect that the standard error of the sample mean will be larger than the standard error of the 20% trimmed mean?
+
+_Because the mean doesn't drop any of the larger observations._
+
+```{r, echo = T}
+sd(x) / sqrt(n)
+```
+
+Verify this speculation.
+
+### 5.27
+
+The ideal estimator of location would have a smaller standard error than any other estimator we might use. Explain why such an estimator does not exist.
+
+_There are going to be tradeoffs involved in any approach. If there wasn't, we would just have a universal estimator._
+
+### 5.28
+
+Under normality, the sample mean has a smaller standard error than the 20% trimmed mean or median.
+If observations are sampled from a distribution that appears to be normal, does this suggest that the mean should be preferred over the trimmed mean and median?
+
+_It depends on the tail behavior and symmetry._
+
+### 5.29
+
+If the sample mean and 20% trimmed mean are nearly identical, it might be thought that for future studies, it will make little difference which measure of location is used.
+Comment on why this is not the case.
+
+_This situation is completely concidental, future studies are not reflective of past studies._
+
+### 5.30
+
+Observations can be generated from the distribution in Figure 5.4a using the R command rlnorm.
+for example, rlnorm(25) would generate 25 observations.
+Indicate the R commands for determining the sampling distribution of the sample median, M.
+In particular, how would you determine the probability that M will have a value less than 1.5?
+
+```{r, echo = T}
+N <- 1e4
+results <- numeric(N)
+
+for( i in 1:N )
+{
+   sample <- rlnorm(25)
+   results[i] <- median(sample)
+}
+
+mean(results < 1.5)
+```

### 2019-12-17 Statistical Methods/05_sampling-distributions.rmd
statistical computing
statistical computing

--- 2019-12-16 Statistical Methods/05_sampling-distributions.rmd
@@
 
+_No, there are ties._
+
 ### 5.23
 
-For the values, 2, 3, 5, 6, 8, 12, 14, 18, 19, 22, 201, why would you suspect that the McKean-Schrader estiamte of the standard error of the median will be smaller than the standard error of the mean?
+For the values, 2, 3, 5, 6, 8, 12, 14, 18, 19, 22, 201, why would you suspect that the McKean-Schrader estimate of the standard error of the median will be smaller than the standard error of the mean?
+
+_There is an outlier._
+
+```{r, echo = T}
+x <- c(2, 3, 5, 6, 8, 12, 14, 18, 19, 22, 201)
+
+msmedse(x)
+
+sqrt(var(x) / length(x))
+
+```
+
+### 5.24
+
+Summarize when it would and would not be reasonable to assume that the sampling distribution of M, the sample median, is normal.
+
+### 5.25
+
+For the values:
+
+59, 106, 174, 207, 219, 237, 313, 365, 485, 497, 515, 529, 557, 615, 625, 645, 973, 1065, 3215
+
+Estimate the standard error of the 20% trimmed mean.
+
+```{r, echo = T}
+x <- c(59, 106, 174, 207, 219, 237, 313, 365, 485, 497, 515, 529, 557, 615, 625, 645, 973, 1065, 3215)
+
+```

### 2019-12-18 Statistical Methods/06_confidence-intervals.rmd
statistical methods, ch 6
statistical methods

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Confidence Intervals}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 6}
   \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(here, quietly = T, warn.conflicts = F)
library(ggplot2, quietly = T, warn.conflicts = F)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 6

### 6.1

Explain the meaning of a .95 confidence interval.

_A confidence interval is a statistical measure of the probability coverage a given lies within some interval based upon the observed data._

### 6.2

If the goal is to compute a .80, or .92, or a .98 confidence interval for $\mu$ when $\sigma$ is known, and sampling is from a normal distribution, what values for c should be used in Equation (6.4)?

```{r, echo = T}
pretty_kable(
   data.table( 
      Confidence = c(.8, .92, .98) )[, 
         Val := (1 - (1 - Confidence)/2)][, C := qnorm(Val)] , 
   "Confidence Values")
```

### 6.3

```{r, echo = T}
conf <- function(alpha) {
   c( Lower = (1 - alpha)/2, Upper = 1 - (1 - alpha)/2)   
}
```

Assuming random sampling is from a normal distribution with standard deviation $\sigma = 5$, if the sample mean is $\bar{X} = 45$ based on $n = 25$ participants, what is the 0.95 confidence interval for $\mu$?

```{r, echo = T}
qnorm(conf(.95), mean = 45, sd = 5/sqrt(25))
```

### 6.4

Repeat the previous example, only compute a .99 confidence interval instead.

```{r, echo = T}
qnorm(conf(.99), mean = 45, sd = 5/sqrt(25))
```

### 6.5

A manufacturer claims that their light bulbs have an average life span that follows a normal distribution with $\mu = 1,200$ hours and a standard deviation of $\sigma = 25$. If you randomly test 36 light bulbs and find that their average life span is $\bar{X} = 1,150$, does a .95 confidence interval for $mu$ suggest that the claim $mu = 1,200$ is reasonable?

```{r, echo = T}
qnorm(conf(.95), mean = 1150, sd = 25/sqrt(36))
```

No, 1,200 is outside the bounds of a 95% confidence interval.

### 6.6

Compute a .95 confidence interval for the mean, assuming normality, for the following situations:

a.) $n = 12, \sigma = 22, \bar{X} = 65$

```{r, echo = T}
qnorm(conf(.95), mean = 65, sd = 22/sqrt(12))
```

b.) $n = 22, \sigma = 10, \bar{X} = 185$

```{r, echo = T}
qnorm(conf(.95), mean = 185, sd = 10/sqrt(22))
```

c.) $n = 50, \sigma = 30, \bar{X} = 19$

```{r, echo = T}
qnorm(conf(.95), mean = 19, sd = 30/sqrt(50))
```

\newpage

### 6.7

What happens to the length of a confidence interval for the mean of a normal distribution when the sample size is doubled? In particular, what is the ratio of the lengths?

What is the ratio of the lengths if the sample size is quadrupled?

__Answer__:

In general, For some _x_,

$CI = \mu \pm C\frac{\sigma}{x\sqrt{n}}$

As _x_ increases, the size of the confidence intervals decrease (closer approximations).

### 6.8

The length of a bolt made by a machine parts company has a normal distribution with standard deviation $\sigma$ equal to 0.01 mm. The lengths of four randomly selected bolts are as follows:

20.01, 19.88, 20.00, 19.99

a.) Compute a 0.95 confidence interval for the mean.

```{r, echo = T}
x <- c(20.01, 19.88, 20.00, 19.99)
n <- length(x)
xbar <- mean(x)

qnorm(conf(.95), mean = xbar, sd = 0.01/sqrt(n))
```

b.) Specifications requre a mean lengths $\mu$ of 20.00 mm for the population of bolts. Do the data indicate that this specification is being met?

_No, the 95% confidence interval for the mean does not include 20._

c.) Given that the 0.95 confidence interval contains the value 20, why might it be inappropriate to conclude that the specification is being met?

_Just because the confidence interval includes some value, that does not mean that the population paramter is that value._

### 6.9

The weight of trout sold at a trout farm has a standard deviation of 0.25. Based on a sample of 10 trout, the average weight is 2.10 lbs.

Assume normality and Compute a .99 confidence interval for the mean.

```{r, echo = T}
qnorm(conf(.99), mean = 2.10, sd = .25/sqrt(10))
```

### 6.10

The average bounce of 45 randomly selected tennis balls is found to be $\bar{X} = 1.70$.

Assuming that the standard deviation of the bounce is .30, compute a 0.90 confidence interval for the average bounce assuming normality.

```{r, echo = T}
qnorm(conf(.90), mean = 1.70, sd = .30/sqrt(45))
```

### 6.11

Assuming that the degrees of freedom are 20, find the value t for which:

a.) $P(T \le t) = 0.995$

```{r, echo = T}
qt(.995, df = 20)
```

b.) $P(T \ge t) = 0.025$

```{r, echo = T}
1 - qt(.025, df = 20)
```

c.) $P(-t \le T \le t) = 0.90$

```{r, echo = T}
qt( (1 + .90) / 2, df = 20)
```

### 6.12

Compute a 0.95 confidence interval if

a.) $n = 10, \bar{X} = 26, s = 9$

```{r, echo = T}
26 + qt(conf(.95), df = 9) * 9/sqrt(10)
```

b.) $n = 18, \bar{X} = 132, s = 20$

```{r, echo = T}
132 + qt(conf(.95), df = 17) * 20/sqrt(18)
```

c.) $n = 25, \bar{X} = 52, s = 12$

```{r, echo = T}
52 + qt(conf(.95), df = 24) * 12/sqrt(25)
```

### 6.13

Compute a 0.99 confidence interval if

a.) $n = 10, \bar{X} = 26, s = 9$

```{r, echo = T}
26 + qt(conf(.99), df = 9) * 9/sqrt(10)
```

b.) $n = 18, \bar{X} = 132, s = 20$

```{r, echo = T}
132 + qt(conf(.99), df = 17) * 20/sqrt(18)
```

c.) $n = 25, \bar{X} = 52, s = 12$

```{r, echo = T}
52 + qt(conf(.99), df = 24) * 12/sqrt(25)
```

### 6.14

For a study on self-awareness, the observed values for one of the groups were:

77, 87, 88, 114, 151, 210, 219, 246, 253, 262, 296, 299, 306, 376, 428, 515, 666, 1310, 2611.

Compute a .95 confidence interval for the mean assuming normality.

```{r, echo = T}
x <- c(77,87,88,114,151,210,219,246,253,262,296,299,306,376,428,515, 666,1310,2611)
xbar <- mean(x)
n <- length(x)

xbar + qt(conf(.95), df = n - 1) * sd(x)/sqrt(n)
```

Compare the result to the bootstrap-t confidence interval.

```{r, echo = T}

winsorize <- function( x, trim = .2) {
   n <- length(x)

   o <- sort(x)
   g <- floor(trim*n)

   o[1:(g+1)] <- o[(g+1)]
   o[(n-g):n] <- o[n-g]

   o
}

trimse<-function(x,tr=.2,na.rm=FALSE){
   if(na.rm)x<-x[!is.na(x)]
   trimse <- sqrt(var(winsorize(x,tr)))/((1-2*tr)*sqrt(length(x)))
   trimse
}

nboot <- 2000
alpha <- .95

sims <- matrix(sample(x, size = x*nboot, replace =T), nrow = nboot)
data <- sims - xbar

top <- apply(data, 1, mean)
bot <- apply(data, 1, trimse)

tval <- top/bot
tval <- sort(tval)

icrit <- round((1-alpha)*nboot)

c(Lower = mean(x)+tval[icrit]*trimse(x),
   Upper = mean(x)-tval[icrit]*trimse(x))
```

Why do they differ?

_Student's T gives a shorter confidence interval, however the probability coverage suggest less than 95% coverage._

### 6.15

Rats are subjected to a drug that might affect aggression. Suppose that for a random sample of rats, measures of aggression are found to be:

5, 12, 23, 24, 18, 9, 18, 11, 36, 15

Compute a 0.95 confidence interval for the mean assuming the scores are from a normal distribution.

### 6.16

Suppose $M = 34$ and the McKean-Schrader estimate of the standard error of M is $S_M = 3$.

Compute a .95 confidence interval for the population median.

### 6.17

### 6.18

### 6.19

### 6.20

### 6.21

### 6.22

### 6.23

### 6.24

### 6.25

### 6.26

### 6.27

### 6.28

### 6.29

### 6.30

### 6.31

### 6.32

### 6.33

### 6.34

### 6.35

### 6.36

### 6.37

### 6.38

### 6.39

### 6.40

### 6.41

### 6.42

### 2019-12-18 Statistical Methods/06_confidence-intervals.rmd
statistical methods, ch6
statistical methods, ch6

@@
 
+```{r, echo = T}
+x <- c(5, 12, 23, 24, 18, 9, 18, 11, 36, 15)
+
+xbar <- mean(x); n <- length(x)
+
+xbar + qt(conf(.95), df = n - 1) * sd(x)/sqrt(n)
+```
+
 ### 6.16
@@
 
+```{r, echo = T}
+34 + qnorm(conf(.95)) * 3
+```
+
 ### 6.17
+
+For the data in E14, the McKean-Schrader estimate of the standard error of _M_ is $S_M = 77.8$ and the sample median is 262.
+
+Compute a .99 confidence interval for the population median.
+
+```{r, echo = T}
+x <- c(77,87,88,114,151,210,219,246,253,262,296,299,306,376,428,515, 666,1310,2611)
+M <- median(x); n <- length(x)
+
+M + qnorm(conf(.99)) * 77.8
+```
 
@@
 
+If $n = 10$ and a confidence interval for the median is computed using $(X_k, X_{n-k})$ as described in 6.3.4, what is the probability coverage if $k=2$?
+
+```{r, echo = T}
+1 - 2*pbinom(1, size = 10, prob = .5)
+```
+
 ### 6.19
+
+From the previous, what is the probability coverage when  $n = 15$ and $k = 4$?
+
+```{r, echo = T}
+1 - 2*pbinom(3, size = 15, prob = .5)
+```
 
@@
 
+For the data in E14, if you use (88, 515) as a confidence interval for the median, what is the probability this interval contains the population median?
+
+```{r, echo = T}
+x <- c(77, 87, 88, 114, 151, 210, 219, 246, 253, 262, 296, 299, 306, 376, 428, 515, 666, 1310, 2611)
+
+k <- 3
+
+1 - 2*pbinom(k, size = n - k - 1, prob = .5)
+```
+
 ### 6.21
+
+You observe the following successes and failures:
+
+1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0
+
+Compute a .95 confidence interval for $p$ using E6.15, as well as the Agresti-Coull method.
+
+```{r, echo = T}
+x <- c(1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)
+n <- length(x)
+
+alpha <- 0.05
+
+phat <- sum(x) / n
+
+phat + qnorm(conf(1 - alpha)) * sqrt(phat*(1 - phat)/n)
+
+c <- qnorm(1 - alpha/2)
+nhat <- n + c^2
+xhat <- sum(x) + c^2/2
+phat <- xhat / nhat
+
+c(Lower = phat - c * sqrt(phat*(1 - phat)/nhat),
+  Upper = phat + c * sqrt(phat*(1 - phat)/nhat))
+```
 
@@
 
+Given the following results for a sample from a binomial distribution, compute the squared standard error of $\hat{p}$.
+
+a.) $n = 25, X = 5$
+
+```{r, echo = T}
+n <- 25; success <- 5
+phat <- success/n
+
+phat*(1 - phat)/n
+```
+
+b.) $n = 48, X = 12$
+
+```{r, echo = T}
+n <- 48; success <- 12
+phat <- success/n
+
+phat*(1 - phat)/n
+```
+
+c.) $n = 100, X = 80$
+
+```{r, echo = T}
+n <- 100; success <- 80
+phat <- success/n
+
+phat*(1 - phat)/n
+```
+
+d.) $n = 300, X = 160$
+
+```{r, echo = T}
+n <- 300; success <- 160
+phat <- success/n
+
+phat*(1 - phat)/n
+```
+
 ### 6.23
+
+Among 100 randomly sampled adults, 10 were found to be unemployed.
+
+Compute a 0.95 confidence interval for the percentage of adults unemployed using E6.15.
+
+Compare this result to the Agresti-Coull confidence interval.
 
@@
 
+A sample of 1,000 fishes was obtained from a lake. It was found that 290 were members of the bass family.
+
+Compute a 0.95 confidence interval for the percenttage of bass in the lake using E6.15.
+
 ### 6.25
+
+Among a random sample of 1,000 adults, 60 reported never having any legal problems.
+Use Equation 6.15 to compute the .95 confidence interval for the percentage of adults who have never had legal problems.
 
@@
 
+Among a sample of 600 items, only one was found to be defective. Explain why E6.15 might be unsatisfactory when computing a confidence interval for the probability that an item is defective.
+
 ### 6.27
+
+In the previous exercise, compute a 0.90 confidence interval for the probability that an item is defective.
 
@@
 
+One-fourth of 300 persons in a large city stated that they are opposed to a certain political issue favored by the mayor.
+
+Using E6.15, calculate a .99 confidence interval for the fraction of individuals opposed to this issue.
+
 ### 6.29
 
+Consider a method for detecting a type of cancer that is highly accurate but very expensive and invasive, resulting in a great deal of discomfort for the patient.
+An alternative test to detect this type of cancer has been developed, and it is of interest to know the probability of a false-negative indication, meaning that the test fails to detect cancer when it is present.
+The test is conducted on 250 patients known to have cancer, and 5 tests fail to show its presence.
+Use equation 6.15 to determine a 0.95 confidence interval for the probability of a false-negative indication.
+
 ### 6.30
+
+In the previous exercise, imagine that 0 false-negative indications were found.
+
+Determine a .99 confidence interval for the probability of a false-negative indication.
 

### 2019-12-19 Statistical Methods/06_confidence-intervals.rmd
statistical methods, ch 6 wrap-up
statistical methods, ch 6 wrap-up

@@
 library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
+library(WRS, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
+install.packages("WRS")
 ```
@@
 
+```{r, echo = T}
+n <- 100; success <- 10
+phat <- success/n
+
+phat + qnorm(conf(.95))*sqrt(phat*(1-phat)/n)
+```
+
 Compute a 0.95 confidence interval for the percentage of adults unemployed using E6.15.
@@
 Compare this result to the Agresti-Coull confidence interval.
+
+```{r, echo = T}
+alpha <- .05
+
+c <- qnorm(1 - alpha/2)
+nhat <- n + c^2
+xhat <- success + c^2/2
+phat <- xhat / nhat
+
+c(Lower = phat - c * sqrt(phat*(1 - phat)/nhat),
+  Upper = phat + c * sqrt(phat*(1 - phat)/nhat))
+```
 
@@
 
-Compute a 0.95 confidence interval for the percenttage of bass in the lake using E6.15.
+Compute a 0.95 confidence interval for the percentage of bass in the lake using E6.15.
+
+```{r, echo = T}
+n <- 1000; success <- 290
+phat <- success/n
+
+phat + qnorm(conf(.95))*sqrt(phat*(1-phat)/n)
+```
 
@@
 
+```{r, echo = T}
+n <- 1000; success <- 60
+
+phat <- success/n
+phat + qnorm(conf(.95))*sqrt(phat*(1-phat)/n)
+```
+
 ### 6.26
@@
 
+_When the probability is very low (or high) the equation in 6.15 performs poorly._
+
 ### 6.27
@@
 In the previous exercise, compute a 0.90 confidence interval for the probability that an item is defective.
+
+```{r, echo = T}
+n <- 600; success <- 1
+phat <- success / n
+
+alpha <- .10
+
+round(c(Lower = 1 - (1 - alpha/2)^(success/n),
+      Upper = 1 - (alpha/2)^(success/n)), 5)
+```
 
@@
 Using E6.15, calculate a .99 confidence interval for the fraction of individuals opposed to this issue.
+
+```{r, echo = T}
+n <- 300; success <- n * .25
+phat <- success/n
+
+phat + qnorm(conf(.99)) * sqrt(phat*(1-phat)/n)
+```
 
@@
 
+```{r, echo = T}
+n <- 250; success <- 5
+phat <- success/n
+
+phat + qnorm(conf(.95)) * sqrt(phat*(1-phat)/n)
+```
+
 ### 6.30
@@
 
+```{r, echo = T}
+n <- 250;
+
+alpha <- 0.01
+
+phat <- 1 - alpha^(1/250)
+
+phat
+
+round(c(Lower = 1 - (1 - alpha/2)^(success/n),
+      Upper = 1 - (alpha/2)^(success/n)), 5)
+```
+
 ### 6.31
 
+A cosmetic company found that 180 of 1,000 randomly selected women in New York city have seen the company's latest television ad.
+Use E6.15 to compute a .95 confidence interval for the percentage of women in New York city who have seen the ad.
+
+```{r, echo = T}
+n <- 1000; success <- 180
+phat <- success/n
+
+phat + qnorm(conf(.95)) * sqrt(phat*(1-phat)/n)
+```
+
 ### 6.32
+
+Describe in general terms how nonnormality can affect Student's T distribution.
+
+_The actual probability coverage can be substantially over or understated depending on the skewness._
 
@@
 
+Listed here are the average LSAT scores for the 1973 entering class of 15 American law schools.
+
+545,555,558,572,575,576,578,580,
+594,605,635,651,653,661,666
+
+The .95 confidence interval for the population mean $\mu$ is (577.1, 623.4).
+
+Use a boxplot to verify that there are not outliers and that the data appear have a skewed distribution.
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+x <- c(545,555,558,572,575,576,578,580,594,605,635,651,653,661,666)
+
+data <- data.table(values = x)
+
+ggplot(data, aes(y  = values, group = "")) +
+   geom_boxplot(fill = "cornflowerblue") +
+   geom_hline(yintercept = 577.1, color = "darkred") +
+   geom_hline(yintercept = 623.4, color = "darkred")
+
+```
+
 ### 6.35
+
+Compute a 0.95 confidence interval for the 20% trimmed mean. if:
+
+a.) $n = 24, s^2_w = 12, \bar{X}_t = 52$
+
+```{r, echo = T}
+xbar <- 52; wt <- 12; n <- 24
+
+g <- floor(n * .2)
+
+xbar + qt(conf(.95), df = n - 2*g - 1) * sqrt(wt) / (.6*sqrt(n))
+```
+
+b.) $n = 36, s^2_w = 30, \bar{X}_t = 10$
+
+```{r, echo = T}
+xbar <- 10; wt <- 30; n <- 36
+
+g <- floor(n * .2)
+
+xbar + qt(conf(.95), df = n - 2*g - 1) * sqrt(wt) / (.6*sqrt(n))
+```
+
+c.) $n = 12, s^2_w = 9, \bar{X}_t = 16$
+
+```{r, echo = T}
+xbar <- 16; wt <- 9; n <- 12
+
+g <- floor(n * .2)
+
+xbar + qt(conf(.95), df = n - 2*g - 1) * sqrt(wt) / (.6*sqrt(n))
+```
 
@@
 
+Repeat with .99 confidence.
+
+a.) $n = 24, s^2_w = 12, \bar{X}_t = 52$
+
+```{r, echo = T}
+xbar <- 52; wt <- 12; n <- 24
+
+g <- floor(n * .2)
+
+xbar + qt(conf(.99), df = n - 2*g - 1) * sqrt(wt) / (.6*sqrt(n))
+```
+
+b.) $n = 36, s^2_w = 30, \bar{X}_t = 10$
+
+```{r, echo = T}
+xbar <- 10; wt <- 30; n <- 36
+
+g <- floor(n * .2)
+
+xbar + qt(conf(.99), df = n - 2*g - 1) * sqrt(wt) / (.6*sqrt(n))
+```
+
+c.) $n = 12, s^2_w = 9, \bar{X}_t = 16$
+
+```{r, echo = T}
+xbar <- 16; wt <- 9; n <- 12
+
+g <- floor(n * .2)
+
+xbar + qt(conf(.99), df = n - 2*g - 1) * sqrt(wt) / (.6*sqrt(n))
+```
+
 ### 6.37
+
+The last example at the end of Section 6.3.1 used data from a study of self-awareness. In another portion of the study, a group of participants had the follow values.
+
+59, 106, 174, 207, 219, 237, 313, 365, 458, 497, 515, 529, 557, 615, 625,
+645, 973, 1065, 3215
+
+Compute a .95 confidence interval for both the population mean and 20% trimmed mean.
+
+```{r, echo = T}
+x <- c(59, 106, 174, 207, 219, 237, 313, 365, 458, 497, 515, 529, 557, 615, 625,
+645, 973, 1065, 3215)
+
+xbar <- mean(x); n <- length(x)
+
+xbar + qt(conf(.95), df = n - 1) * sqrt(sd(x) / n)
+
+g <- floor(n * .2)
+
+xt <- x[g:n-g]
+
+xtbar <- mean(xt); nt <- length(xt)
+
+xtbar + qt(conf(.95), df = nt - 1) * sd(xt) / sqrt(nt)
+```
 
@@
 
+The ideal method for computing a confidence interval for the mean would have a shorter length than any other method that might be used.
+Explain why such a method does not exist even when sampling from a perfectly symetric distribution.
+
+_Because outliers._
+
 ### 6.39
+
+For the values:
+
+5, 60, 43, 56, 32, 43, 47, 79, 39, 41
+
+compute a .95 confidence interval for the 20% trimmed mean and compare the results to the .95 confidence interval for the mean.
+
+```{r, echo = T}
+x <- c(5, 60, 43, 56, 32, 43, 47, 79, 39, 41)
+
+xbar <- mean(x); n <- length(x)
+xbar + qt(conf(.95), df = n - 1) * sqrt(sd(x)/n)
+
+g <- floor(n * .2)
+
+x <- sort(x)
+x <- x[g:n-g]
+
+xbar <- mean(x); n <- length(x)
+xbar + qt(conf(.95), df = n - 1) * sqrt(sd(x)/n)
+
+```
 
@@
 
+In the previous exercise, the confidence interval for the 20% trimmed mean is shorter than the confidence interval for the mean.
+
+```{r, echo = T}
+outbox(x)
+```
+
+Explain why this is not suprising.
+
 ### 6.41
 
+The following code for estimating the distribution of Student's T when data are sampled from a skewed, light-tailed distribution that has a population mean $\mu = 1.649$:
+
+```{r, echo = T}
+t.vals <- NULL
+for(study in 1:1000) {
+   x <- rlnorm(25)
+   suppressMessages({
+      t.vals[study] <- trimciv2(x, tr=0, null.value=1.649)$test.stat})
+}
+```
+
 ### 6.42
+
+Modify the previous code to use actual data from a study to compute a confidence interval.
+
+```{r, echo = T, fig.width=8, fig.height=3}
+t.vals=NULL
+for(i in 1:1000){
+z=sample(x,25,replace=T)
+t.vals[i]=trimci(z,tr=0,null.value=mean(x),pr=F)$test.stat
+}
+
+ggplot(data.table(values = x), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(fill = "darkorange", alpha = .3)
+```

### 2019-12-19 Statistical Methods/07_hypothesis-testing.rmd
statistical methods, ch 6 wrap-up
statistical methods, ch 6 wrap-up

--- 2019-12-18 Statistical Methods/06_confidence-intervals.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Confidence Intervals}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Hypothesis Testing}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 6}
+   \lfoot{\color{headergrey}Chapter 7}
    \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
@@
 
-#### Chapter 6
+#### Chapter 7
 
-### 6.1
-
-Explain the meaning of a .95 confidence interval.
-
-_A confidence interval is a statistical measure of the probability coverage a given lies within some interval based upon the observed data._
-
-### 6.2
-
-If the goal is to compute a .80, or .92, or a .98 confidence interval for $\mu$ when $\sigma$ is known, and sampling is from a normal distribution, what values for c should be used in Equation (6.4)?
-
-```{r, echo = T}
-pretty_kable(
-   data.table( 
-      Confidence = c(.8, .92, .98) )[, 
-         Val := (1 - (1 - Confidence)/2)][, C := qnorm(Val)] , 
-   "Confidence Values")
-```
-
-### 6.3
-
-```{r, echo = T}
-conf <- function(alpha) {
-   c( Lower = (1 - alpha)/2, Upper = 1 - (1 - alpha)/2)   
-}
-```
-
-Assuming random sampling is from a normal distribution with standard deviation $\sigma = 5$, if the sample mean is $\bar{X} = 45$ based on $n = 25$ participants, what is the 0.95 confidence interval for $\mu$?
-
-```{r, echo = T}
-qnorm(conf(.95), mean = 45, sd = 5/sqrt(25))
-```
-
-### 6.4
-
-Repeat the previous example, only compute a .99 confidence interval instead.
-
-```{r, echo = T}
-qnorm(conf(.99), mean = 45, sd = 5/sqrt(25))
-```
-
-### 6.5
-
-A manufacturer claims that their light bulbs have an average life span that follows a normal distribution with $\mu = 1,200$ hours and a standard deviation of $\sigma = 25$.
-If you randomly test 36 light bulbs and find that their average life span is $\bar{X} = 1,150$, does a .95 confidence interval for $mu$ suggest that the claim $mu = 1,200$ is reasonable?
-
-```{r, echo = T}
-qnorm(conf(.95), mean = 1150, sd = 25/sqrt(36))
-```
-
-No, 1,200 is outside the bounds of a 95% confidence interval.
-
-### 6.6
-
-Compute a .95 confidence interval for the mean, assuming normality, for the following situations:
-
-a.) $n = 12, \sigma = 22, \bar{X} = 65$
-
-```{r, echo = T}
-qnorm(conf(.95), mean = 65, sd = 22/sqrt(12))
-```
-
-b.) $n = 22, \sigma = 10, \bar{X} = 185$
-
-```{r, echo = T}
-qnorm(conf(.95), mean = 185, sd = 10/sqrt(22))
-```
-
-c.) $n = 50, \sigma = 30, \bar{X} = 19$
-
-```{r, echo = T}
-qnorm(conf(.95), mean = 19, sd = 30/sqrt(50))
-```
-
-\newpage
-
-### 6.7
-
-What happens to the length of a confidence interval for the mean of a normal distribution when the sample size is doubled? In particular, what is the ratio of the lengths?
-
-What is the ratio of the lengths if the sample size is quadrupled?
-
-__Answer__:
-
-In general, For some _x_,
-
-$CI = \mu \pm C\frac{\sigma}{x\sqrt{n}}$
-
-As _x_ increases, the size of the confidence intervals decrease (closer approximations).
-
-### 6.8
-
-The length of a bolt made by a machine parts company has a normal distribution with standard deviation $\sigma$ equal to 0.01 mm. The lengths of four randomly selected bolts are as follows:
-
-20.01, 19.88, 20.00, 19.99
-
-a.) Compute a 0.95 confidence interval for the mean.
-
-```{r, echo = T}
-x <- c(20.01, 19.88, 20.00, 19.99)
-n <- length(x)
-xbar <- mean(x)
-
-qnorm(conf(.95), mean = xbar, sd = 0.01/sqrt(n))
-```
-
-b.) Specifications requre a mean lengths $\mu$ of 20.00 mm for the population of bolts. Do the data indicate that this specification is being met?
-
-_No, the 95% confidence interval for the mean does not include 20._
-
-c.) Given that the 0.95 confidence interval contains the value 20, why might it be inappropriate to conclude that the specification is being met?
-
-_Just because the confidence interval includes some value, that does not mean that the population paramter is that value._
-
-### 6.9
-
-The weight of trout sold at a trout farm has a standard deviation of 0.25. Based on a sample of 10 trout, the average weight is 2.10 lbs.
-
-Assume normality and Compute a .99 confidence interval for the mean.
-
-```{r, echo = T}
-qnorm(conf(.99), mean = 2.10, sd = .25/sqrt(10))
-```
-
-### 6.10
-
-The average bounce of 45 randomly selected tennis balls is found to be $\bar{X} = 1.70$.
-
-Assuming that the standard deviation of the bounce is .30, compute a 0.90 confidence interval for the average bounce assuming normality.
-
-```{r, echo = T}
-qnorm(conf(.90), mean = 1.70, sd = .30/sqrt(45))
-```
-
-### 6.11
-
-Assuming that the degrees of freedom are 20, find the value t for which:
-
-a.) $P(T \le t) = 0.995$
-
-```{r, echo = T}
-qt(.995, df = 20)
-```
-
-b.) $P(T \ge t) = 0.025$
-
-```{r, echo = T}
-1 - qt(.025, df = 20)
-```
-
-c.) $P(-t \le T \le t) = 0.90$
-
-```{r, echo = T}
-qt( (1 + .90) / 2, df = 20)
-```
-
-### 6.12
-
-Compute a 0.95 confidence interval if
-
-a.) $n = 10, \bar{X} = 26, s = 9$
-
-```{r, echo = T}
-26 + qt(conf(.95), df = 9) * 9/sqrt(10)
-```
-
-b.) $n = 18, \bar{X} = 132, s = 20$
-
-```{r, echo = T}
-132 + qt(conf(.95), df = 17) * 20/sqrt(18)
-```
-
-c.) $n = 25, \bar{X} = 52, s = 12$
-
-```{r, echo = T}
-52 + qt(conf(.95), df = 24) * 12/sqrt(25)
-```
-
-### 6.13
-
-Compute a 0.99 confidence interval if
-
-a.) $n = 10, \bar{X} = 26, s = 9$
-
-```{r, echo = T}
-26 + qt(conf(.99), df = 9) * 9/sqrt(10)
-```
-
-b.) $n = 18, \bar{X} = 132, s = 20$
-
-```{r, echo = T}
-132 + qt(conf(.99), df = 17) * 20/sqrt(18)
-```
-
-c.) $n = 25, \bar{X} = 52, s = 12$
-
-```{r, echo = T}
-52 + qt(conf(.99), df = 24) * 12/sqrt(25)
-```
-
-### 6.14
-
-For a study on self-awareness, the observed values for one of the groups were:
-
-77, 87, 88, 114, 151, 210, 219, 246, 253, 262, 296, 299, 306, 376, 428, 515, 666, 1310, 2611.
-
-Compute a .95 confidence interval for the mean assuming normality.
-
-```{r, echo = T}
-x <- c(77,87,88,114,151,210,219,246,253,262,296,299,306,376,428,515, 666,1310,2611)
-xbar <- mean(x)
-n <- length(x)
-
-xbar + qt(conf(.95), df = n - 1) * sd(x)/sqrt(n)
-```
-
-Compare the result to the bootstrap-t confidence interval.
-
-```{r, echo = T}
-
-winsorize <- function( x, trim = .2) {
-   n <- length(x)
-
-   o <- sort(x)
-   g <- floor(trim*n)
-
-   o[1:(g+1)] <- o[(g+1)]
-   o[(n-g):n] <- o[n-g]
-
-   o
-}
-
-trimse<-function(x,tr=.2,na.rm=FALSE){
-   if(na.rm)x<-x[!is.na(x)]
-   trimse <- sqrt(var(winsorize(x,tr)))/((1-2*tr)*sqrt(length(x)))
-   trimse
-}
-
-nboot <- 2000
-alpha <- .95
-
-sims <- matrix(sample(x, size = x*nboot, replace =T), nrow = nboot)
-data <- sims - xbar
-
-top <- apply(data, 1, mean)
-bot <- apply(data, 1, trimse)
-
-tval <- top/bot
-tval <- sort(tval)
-
-icrit <- round((1-alpha)*nboot)
-
-c(Lower = mean(x)+tval[icrit]*trimse(x),
-   Upper = mean(x)-tval[icrit]*trimse(x))
-```
-
-Why do they differ?
-
-_Student's T gives a shorter confidence interval, however the probability coverage suggest less than 95% coverage._
-
-### 6.15
-
-Rats are subjected to a drug that might affect aggression. Suppose that for a random sample of rats, measures of aggression are found to be:
-
-5, 12, 23, 24, 18, 9, 18, 11, 36, 15
-
-Compute a 0.95 confidence interval for the mean assuming the scores are from a normal distribution.
-
-### 6.16
-
-Suppose $M = 34$ and the McKean-Schrader estimate of the standard error of M is $S_M = 3$.
-
-Compute a .95 confidence interval for the population median.
-
-### 6.17
-
-### 6.18
-
-### 6.19
-
-### 6.20
-
-### 6.21
-
-### 6.22
-
-### 6.23
-
-### 6.24
-
-### 6.25
-
-### 6.26
-
-### 6.27
-
-### 6.28
-
-### 6.29
-
-### 6.30
-
-### 6.31
-
-### 6.32
-
-### 6.33
-
-### 6.34
-
-### 6.35
-
-### 6.36
-
-### 6.37
-
-### 6.38
-
-### 6.39
-
-### 6.40
-
-### 6.41
-
-### 6.42
+### 7.1

### 2019-12-19 Statistical Methods/07_hypothesis-testing.rmd
statistical methods
statistical methods

@@
 ### 7.1
+
+Given that $\bar{X} = 78, \sigma^2 =25, n = 10, \alpha = 0.05$ test $H_0 : \mu \ge 80$, assuming that observations are randomly sampled from a normal distribution.
+
+```{r, echo = T}
+xbar <- 78; sigma <- 5; n <- 10
+mu <- 80; alpha <- 0.05
+
+# mu > 80 = left tail test
+
+crit <- qnorm(alpha)
+
+Z <- sqrt(n)*(xbar - mu)/sigma
+
+ifelse(Z > crit, "Fail to reject the null", "Reject the null")
+
+```
+
+### 7.2
+
+Repeat the previous exercise, but test $H_0 : \mu = 80$
+
+```{r, echo = T}
+xbar <- 78; sigma <- 5; n <- 10
+mu <- 80; alpha <- 0.05
+
+# mu = 80, two-tailed test
+
+crit <- qnorm(1 - alpha/2)
+
+Z <- sqrt(n)*(xbar - mu)/sigma
+
+ifelse(abs(Z) <= crit, "Fail to reject the null", "Reject the null")
+
+```
+
+### 7.3
+
+For the data in Exercise 1, compute a 0.95 confidence interval and verify that this interval is consistent with y our decision about whether to reject the null hypothesis $H_0 : \mu = 80$.
+
+```{r, echo = T}
+xbar <- 78; sigma <- 5; n <- 10
+mu <- 80; alpha <- 0.05
+
+xbar + qnorm(c(Lower = alpha/2, Upper = 1 - (alpha/2))) * sigma/sqrt(n)
+```
+
+_80_ is in the confidence interval, so fail to reject.
+
+### 7.4
+
+For exercise 1, determine the p-value.
+
+### 7.5
+
+For exercise 2, determine the p-value.
+
+### 7.6
+
+Given that $\bar{X} = 120, \sigma = 5, n = 49, \alpha = 0.05$, test $H_0 : \mu \ge 130$, assuming that observations are randomly sampled from a normal distribution.
+
+### 7.7
+
+Repeat the previous exercise but test $H_0 : \mu = 130$.
+
+### 7.8
+
+For the previous exercise, compute a 0.95 confidence interval and compare the result with your decision about whether to reject $H_0$.
+
+### 7.9
+
+If $\bar{X} = 23$ and $\alpha = 0.025$, can you make a decision about whether to reject $H_0 : \mu \le 25$ without knowing $\sigma$?
+
+### 7.10
+### 7.11
+### 7.12
+### 7.13
+### 7.14
+### 7.15
+### 7.16
+### 7.17
+### 7.18
+### 7.19
+
+### 7.20
+### 7.21
+### 7.22
+### 7.23
+### 7.24
+### 7.25
+### 7.26
+### 7.27
+### 7.28
+### 7.29
+
+### 7.30
+### 7.31
+### 7.32
+### 7.33
+### 7.34
+### 7.35
+### 7.36
+### 7.37

### 2019-12-19 Statistical Methods/_packages.R

@@ -1,1 +1,13 @@
-install.packages("WRS")
+# first: install dependent packages
+install.packages(c("MASS", "akima", "robustbase"))
+
+# second: install suggested packages
+install.packages(c("cobs", "robust", "mgcv", "scatterplot3d", "quantreg", "rrcov", "lars", "pwr", "trimcluster", "mc2d", "psych", "Rfit", "DepthProc", "class", "fda"))
+
+# third: install an additional package which provides some C functions
+install.packages("devtools")
+library("devtools")
+install_github("mrxiaohe/WRScpp")
+
+# fourth: install WRS
+install_github("nicebread/WRS", subdir="pkg")

### 2019-12-20 Statistical Methods/07_hypothesis-testing.rmd
statistical methods
statistical methods

@@
 ```{r, echo = T}
-xbar <- 78; sigma <- 5; n <- 10
-mu <- 80; alpha <- 0.05
+xbar <- 78; var <- 25; n <- 10
+alpha <- 0.05; h0 <- 80
 
-# mu > 80 = left tail test
+# greater than = left tail test
 
@@
 
-Z <- sqrt(n)*(xbar - mu)/sigma
+Z <- sqrt(n) * (xbar - h0) / sqrt( var )
 
-ifelse(Z > crit, "Fail to reject the null", "Reject the null")
-
+ifelse(Z < crit, "Reject the null", "Cannot reject null")
 ```
@@
 ```{r, echo = T}
-xbar <- 78; sigma <- 5; n <- 10
-mu <- 80; alpha <- 0.05
 
-# mu = 80, two-tailed test
+# equals = two-tailed test
 
-crit <- qnorm(1 - alpha/2)
+crit <- qnorm(alpha/2)
 
-Z <- sqrt(n)*(xbar - mu)/sigma
+Z <- sqrt(n) * (xbar - h0) / sqrt( var )
 
-ifelse(abs(Z) <= crit, "Fail to reject the null", "Reject the null")
+ifelse(Z < crit || crit > Z, "Reject the null", "Cannot reject the null")
 
@@
 ```{r, echo = T}
-xbar <- 78; sigma <- 5; n <- 10
-mu <- 80; alpha <- 0.05
 
-xbar + qnorm(c(Lower = alpha/2, Upper = 1 - (alpha/2))) * sigma/sqrt(n)
+xbar + qnorm(c(alpha/2, 1 - alpha/2)) * sqrt( var / n )
+
 ```
@@
 
+```{r, echo = T}
+pnorm(Z)
+```
+
 ### 7.5
@@
 For exercise 2, determine the p-value.
+
+```{r, echo = T}
+2 * (1 - pnorm(Z, lower.tail = F))
+```
 
@@
 
+```{r, echo = T}
+xbar <- 120; sigma <- 5; n <- 49
+alpha <- 0.05; h0 <- 130
+
+crit <- qnorm(alpha)
+
+Z <- sqrt(n) * (xbar - h0) / sigma
+
+ifelse( Z < crit, "Reject null hypothesis", "Cannot reject null")
+```
+
 ### 7.7
@@
 Repeat the previous exercise but test $H_0 : \mu = 130$.
+
+```{r, echo = T}
+h0 <- 130
+
+# two tailed test
+crit <- qnorm(c(alpha/2, 1 - alpha/2))
+
+ifelse(Z < crit[1] || Z > crit[2], "Reject null hypothesis", "Cannot Reject Null")
+```
 
@@
 
+```{r, echo = T}
+xbar + qnorm(c(Lower = alpha/2, Upper = 1 - alpha/2)) * sigma/sqrt(n)
+```
+
+130 is not in the confidence region, reject null-hypothesis.
+
 ### 7.9
@@
 
+_Yes, the hypothesis is consistent with the sample mean._
+
 ### 7.10
+
+An electronics firm mass-produces a component for which there is a standard measure of quality.
+Based on testing vast numbers of these components, the company has found that the average quality is $\mu = 232, \sigma = 4$.
+
+However, in recent years, the quality has not been checked, so management asks you to check their claim with the goal of being reasonable certain that an average quality of less than 232 can be ruled out.
+That is, assume that the quality is poor and in fact less than 232 with the goal of empirically establishing that this assumption is unlikely.
+
+You get $\bar{X} = 240$ based on a sample of $n = 25$ components, and you want the probability of a Type I error to be less than 0.01.
+
+State the null hypothesis and perform the appropriate test, assuming normality.
+
+#### Hypothesis:
+
+$H_0 : \mu \le 232$
+
+$H_A : \mu \ge 232$
+
+```{r, echo = T}
+xbar <- 240; sigma <- 4; n <- 25
+alpha <- 0.01; h0 <- 232
+
+# less than = upper-tailed test
+
+crit <- qnorm(1 - alpha)
+
+Z <- sqrt(n) * ( xbar - h0 ) / sigma
+
+ifelse(Z > crit, "Reject the null", "Cannot reject the null")
+```
+
 ### 7.11

### 2019-12-21 Statistical Methods/07_hypothesis-testing.rmd
statistical methods
statistical methods

@@
 ### 7.11
+
+An antipollution device for cars is claimed to have an average effectiveness of exactly 546. Based on a test of 20 such devices, you find that $\bar{X} = 565$.
+
+Assuming normality and that $\sigma = 40$, would you rule out the claim with a Type I error probability of 0.05?
+
+```{r, echo = T}
+xbar <- 565; sigma <- 40; n <- 20
+alpha <- 0.05; h0 <- 546
+
+# equality = two tailed
+
+crit <- qnorm(c(alpha/2, 1 - alpha/2))
+
+Z <- sqrt(n) * (xbar - h0) / sigma
+
+ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
+```
+
 ### 7.12
+
+Comment on the relative merits of using a 0.95 confidence interval for addressing the effectiveness of the antipollution device in the previous exercise rather than test the hypothesis: $H_0 : \mu = 232$
+
+```{r, echo = T}
+xbar + qnorm(c(alpha/2, 1 - alpha/2)) * sigma/sqrt(n)
+
+2 * ( 1 - pnorm(Z) )
+```
+
+The hypothesised value doesn't fall within the confidence interval, however the formalized test gives us a degree of strength on which we can reject the null.
+Here our p-value is .03, so we reject with confidence.
+
 ### 7.13
+
 ### 7.14

### 2019-12-21 Statistical Methods/07_hypothesis-testing.rmd
statistical methods, ch7
statistical methods, ch7

@@
 The hypothesised value doesn't fall within the confidence interval, however the formalized test gives us a degree of strength on which we can reject the null.
-Here our p-value is .03, so we reject with confidence.
+Here our p-value is .03, so we reject with slight confidence.
+
+\newpage
 
@@
 
+For $n = 25, \alpha = 0.01, \sigma = 5, H_0 : \mu \ge 60$, verify that power is .95 when $\mu = 56$
+
+```{r, echo = T}
+
+mu <- 56; sigma <- 5; n <- 25
+alpha <- 0.01; h0 <- 60
+
+# is mu greater than 60? gt = lower tail test
+
+crit <- qnorm(alpha)
+
+Z <- sqrt(n) * ( mu - h0) / sigma
+
+pnorm(crit - Z) # power
+```
+
 ### 7.14
+
+For $n = 36, \alpha = 0.025, \sigma = 8, H_0 : \mu \le 100$, verify that power is .61 when $\mu = 103$
+
+```{r, echo = T}
+mu <- 103; sigma <- 8; n <- 36
+alpha <- 0.025; h0 <- 100
+
+# less than, upper tailed test
+
+crit <- qnorm(alpha, lower.tail = F)
+
+Z <- sqrt(n) * (mu - h0) / sigma
+
+ifelse( Z < crit, "Reject null", "Cannot reject null")
+
+pnorm(crit - Z, lower.tail = F) # power
+```
+
+\newpage
+
 ### 7.15
+
+For $n = 49, \alpha = 0.05, \sigma = 10, H_0 : \mu = 50$, verify that power is approximately .56 when $\mu = 47$
+
+```{r, echo = T}
+mu <- 47; sigma <- 10; n <- 49
+alpha <- 0.05; h0 <- 50
+
+# mu = 50? two-tailed test
+
+crit <- qnorm(c(Lower = alpha/2, Upper = 1 - alpha/2))
+
+Z <- sqrt(n) * (mu - h0) / sigma
+
+ifelse( Z < crit[1] || Z > crit[2], "Reject null hypothesis", "Cannot reject null")
+
+as.vector(pnorm(crit[1] - Z) + pnorm(crit[2] - Z, lower.tail = F)) # power
+```
+
 ### 7.16
+
+A manufacturer of medication for migraine headaches knows that their product can damage the stomach if taken too often.
+Imagine that by a standard measuring process, the average damage is $\mu = 48$.
+A modification of their product is being contemplated, and based on 10 trials, it is found that $\bar{X} = 46$.
+Assuming $\sigma = 5$ they test $H_0 : \mu \ge 48$, the idea being that if they reject, there is convincing evidence that the average amount of damage is less than 48.
+Then,
+
+$Z = \frac{46 - 48}{5/\sqrt10} = -1.3$
+
+With $\alpha = 0.05$, the critical value is -1.645, so they do not reject Z is less than the critical value.
+
+What might be wrong with accepting $H_0$ and concluding that the modificaiton results in an average amount of damage greater than or equal to 48?
+
+```{r, echo = T}
+mu <- 46; sigma <- 5; n <- 10
+alpha <- 0.05; h0 <- 48
+
+crit <- qnorm(alpha)
+
+Z <- (mu - h0) / sigma * sqrt(n)
+
+pnorm(crit - Z)
+```
+
+Low power in the test, might commit a Type II error.
+
 ### 7.17
+
+For the previous exercise, verify that the power is 0.35 if $\mu = 46$
+
+```{r, echo = T}
+pnorm(crit - Z)
+```
+
 ### 7.18
+
+The previous exercise indicates that power is relatively low with only $n = 10$ observations.
+Imagine that you want power to be at least .8.
+One way of getting more power is to increase the sample size, _n_.
+
+Verify that for sample sizes of 20, 30 and 40, power is .56, 0.71, and .81, respectively.
+
+```{r, echo = T}
+n <- c(20, 30, 40)
+
+Z <- (mu - h0) / sigma * sqrt(n)
+
+pnorm(crit - Z)
+```
+
 ### 7.19
 
+For the previous exercise, rather than increase the sample size, what else might you do to increase the power? What is the negative consequence for this strategy?
+
+_An alternative to increasing the sample size is to increase alpha, which will increase the probability of a type I error._
+
 ### 7.20
+
+Given the following values for $\bar{X}$ and s:
+
+test the hypothesis $H_0 : \mu = 42, \alpha = 0.05, n = 25$
+
+a.) $\bar{X} = 44, s = 10$
+
+```{r, echo = T}
+xbar <- 44; s <- 10; n <- 25
+alpha <- 0.05; h0 <- 42
+
+crit <- qt(c(alpha/2, 1 - alpha/2), df = n - 1)
+
+Z <- (xbar - h0) / s * sqrt(n)
+
+ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
+```
+
+b.) $\bar{X} = 43, s = 10$
+
+```{r, echo = T}
+xbar <- 43; s <- 10; n <- 25
+alpha <- 0.05; h0 <- 42
+
+crit <- qt(c(alpha/2, 1 - alpha/2), df = n - 1)
+
+Z <- (xbar - h0) / s * sqrt(n)
+
+ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
+```
+
+c.) $\bar{X} = 43, s = 2$
+
+```{r, echo = T}
+xbar <- 43; s <- 2; n <- 25
+alpha <- 0.05; h0 <- 42
+
+crit <- qt(c(alpha/2, 1 - alpha/2), df = n - 1)
+
+Z <- (xbar - h0) / s * sqrt(n)
+
+ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
+```
+
 ### 7.21
+
+For part b of the last exercise, you fail to reject but you reject for the situation in part c. What does this illustrate about power?
+
+_Power is higher with smaller variance._
+
 ### 7.22
+
+Test the hypothesis $H_0 : \mu \le 42$ with $\alpha = 0.05, n = 16$ for the following values of $\bar{X}$ and s:
+
+a.) $\bar{X} = 44, s = 10$
+
+```{r, echo = T}
+xbar <- 44; s <- 10; n <- 16
+alpha <- 0.05; h0 <- 42
+
+# less than, upper tailed test
+crit <- qt(alpha, df = n - 1, lower.tail = F)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse( Tval > crit, "reject the null", "fail to reject")
+```
+
+b.) $\bar{X} = 43, s = 10$
+
+```{r, echo = T}
+xbar <- 43; s <- 10
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse( Tval > crit, "reject the null", "fail to reject")
+```
+
+c.) $\bar{X} = 43, s = 2$
+
+```{r, echo = T}
+xbar <- 43; s <- 2
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse(Tval > crit, "reject the null", "fail to reject")
+```
+
 ### 7.23
+
+Repeat the previous for $H_0 : \mu \ge 42$
+
+a.) $\bar{X} = 44, s = 10$
+
+```{r, echo = T}
+xbar <- 44; s <- 10
+h0 <- 42
+
+# greater than, lower tailed test
+crit <- qt(alpha, df = n - 1)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse( Tval < crit, "reject the null", "fail to reject null")
+```
+
+b.) $\bar{X} = 43, s = 10$
+
+```{r, echo = T}
+xbar <- 43; s <- 10
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse( Tval < crit, "reject the null", "fail to reject null")
+```
+
+c.) $\bar{X} = 43, s = 2$
+
+```{r, echo = T}
+xbar <- 43; s <- 2
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse( Tval < crit, "reject the null", "fail to reject null")
+```
+
 ### 7.24
+
+A company claims that on average, when their toothpaste is used, 45% of all bacteria related to gingivitus is killed.
+
+For 10 individuals, it is found that the percentages of bacteria killed are 38, 44, 62, 72, 43, 40, 43, 42, 39, 41.
+
+The mean and standard deviation of these values are $\bar{X} = 46.4, s = 11.27$, respectively.
+
+Assuming normality, test the hypothesis that the average percentage is 45 with $\alpha = 0.05$
+
+```{r, echo = T}
+x <- c(38, 44, 62, 72, 43, 40, 43, 42, 39, 41)
+
+xbar <- mean(x); s <- sd(x); n <- length(x)
+alpha <- 0.05; h0 <- 45
+
+crit <- qt(1 - alpha/2, df = n - 1)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse(abs(Tval) > crit, "Reject the null", "Fail to reject the null")
+
+pt(Tval, df = n - 1) # p-value
+```
+
+\newpage
+
 ### 7.25
+
+A portion of a study by Wechsler (1958) reports that for 100 males taking the Wechsler Adult Intelligent Scale (WAIS), the sample mean and variance on picture completion are $\bar{X} = 9.79$ and $s = 2.72$, respectively.
+
+Test the hypothesis $H_0 : \mu \ge 10.5, \alpha = 0.025$
+
+```{r, echo = T}
+xbar <- 9.79; s <- 2.72; n <- 100
+alpha <- 0.025; h0 <- 10.5
+
+crit <- qt(alpha, df = n - 1)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse(Tval < crit, "reject the null", "fail to reject")
+
+pt(Tval, df = n - 1) # p-value
+```
+
 ### 7.26
+
+Given that $n = 16, \bar{X} = 40, and s = 4$, test that $H_0 : \mu \le 38, \alpha = 0.01$
+
+```{r, echo = T}
+xbar <- 40; s <- 4; n <- 16
+alpha <- 0.01; h0 <- 38
+
+crit <- qt(alpha, df = n - 1, lower.tail = F)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse(Tval > crit, "Reject null", "Fail to reject null")
+
+pt(Tval, df = n - 1)
+```
+
 ### 7.27
+
+Given that $n = 9, \bar{X} = 33, s = 4$ test $H_0 : \mu = 32, \alpha = 0.05$
+
+```{r, echo = T}
+xbar <- 33; s <- 4; n <- 9
+alpha <- 0.05; h0 <- 32
+
+crit <- qt(1 - alpha/2, df = n - 1)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse(abs(Tval) > crit, "Reject null", "Cannot reject null")
+```
+
 ### 7.28
+
+An engineer believes it takes an average of 150 man-hours to assemble a portion of an automobile. As a check, the time to assemble 10 such parts was ascertained, yielding $\bar{X} = 146, s = 2.5$.
+
+Test the engineer's belief with $\alpha = 0.05$
+
+```{r, echo = T}
+xbar <- 146; s <- 2.5; n <- 10
+alpha <- 0.05; h0 <- 150
+
+crit <- qt(1 - alpha/2, df = n - 1)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse(abs(Tval) > crit, "Reject the null", "Fail to Reject")
+```
+
 ### 7.29
 
+In a study of court administration, the following times to disposition were determined for 20 cases and found to be:
+
+42, 90, 84, 87, 119, 95, 86, 99, 93, 92, 121, 71, 66, 98, 79, 102, 60, 112, 105, 98
+
+Test the hypothesis that the average time to disposition is less than or equal to 80 at $\alpha = 0.01$.
+
+```{r, echo = T}
+x <- c(42, 90, 84, 87, 119, 95, 86, 99, 93, 92, 121, 71, 66, 98, 79, 102, 60, 112, 105, 98)
+
+xbar <- mean(x); s <- sd(x); n <- length(x)
+alpha <- 0.01; h0 <- 80
+
+crit <- qt(alpha, df = n - 1, lower.tail = F)
+
+Tval <- (xbar - h0) / s * sqrt(n)
+
+ifelse(Tval > crit, "Reject the null", "Cannot reject the null")
+
+pt(Tval, df = n - 1, lower.tail = F) # p-value
+
+```
+
+Or, simply:
+
+```{r, echo = T}
+t.test(x, alternative = 'greater', conf.level = 1 - alpha, mu = h0)
+```
+
 ### 7.30
+
+Assuming 20% trimming, test the hypothesis $H_0 : \mu = 42, \alpha = 0.05, n = 20$ given the following values for $\bar{X}$ and $s_w$:
+
+a.) $\bar{X}_t = 44, s_w = 9$
+
+b.) $\bar{X}_t = 43, s_w = 9$
+
+c.) $\bar{X}_t = 43, s_w = 3$
+
 ### 7.31
+
+Repeat the previous exercise, only test the hypothesis $H_0 : \mu_t \le 42, \alpha = 0.05$
+
+a.) $\bar{X}_t = 44, s_w = 9$
+
+b.) $\bar{X}_t = 43, s_w = 9$
+
+c.) $\bar{X}_t = 43, s_w = 3$
+
 ### 7.32
+
+For the data in Exercise 24, the 20% trimmed mean is $\bar{X}_t = 42.17$ with a Winsorized standard deviation of $S_w = 1.73$
+
+Test the hypothesis that the population trimmed mean is 45 with $\alpha = 0.05$
+
 ### 7.33

### 2019-12-21 Statistical Methods/07_hypothesis-testing.rmd
statistical methods, ch7
statistical methods, ch7

@@
 
-Assuming 20% trimming, test the hypothesis $H_0 : \mu = 42, \alpha = 0.05, n = 20$ given the following values for $\bar{X}$ and $s_w$:
+Assuming 20% trimming, test the hypothesis $H_0 : \mu = 42, \alpha = 0.05, n = 20$ given the following values for $\bar{X}_t$ and $s_w$:
 
@@
 
+```{r, echo = T}
+xbar <- 44; s <- 9; n <- 20
+alpha <- 0.05; h0 <- 42
+
+g <- floor(.2*n)
+
+df <- n - g*2 - 1
+
+crit <- qt(1 - alpha/s, df)
+
+Tval <- .6 * (xbar - h0) / s * sqrt(n)
+
+ifelse(abs(Tval) > crit, "Reject the null", "Cannot reject null")
+```
+
 b.) $\bar{X}_t = 43, s_w = 9$
+
+```{r, echo = T}
+xbar <- 43; s <- 9;
+
+Tval <- .6 * (xbar - h0) / s * sqrt(n)
+
+ifelse(abs(Tval) > crit, "Reject the null", "Cannot reject null")
+```
 
@@
 
+```{r, echo = T}
+xbar <- 43; s <- 3;
+
+Tval <- .6 * (xbar - h0) / s * sqrt(n)
+
+ifelse(abs(Tval) > crit, "Reject the null", "Cannot reject null")
+```
+
 ### 7.31
 
-Repeat the previous exercise, only test the hypothesis $H_0 : \mu_t \le 42, \alpha = 0.05$
+Repeat the previous exercise, only test the hypothesis $H_0 : \mu_t \le 42, \alpha = 0.05, n = 16$
 
@@
 
+```{r, echo = T}
+xbar <- 44; s <- 9; n <- 16
+alpha <- 0.05
+
+g <- floor(.2*n)
+
+df <- n - g*2 - 1
+
+crit <- qt(1 - alpha, df)
+
+Tval <- .6 * (xbar - h0) / s * sqrt(n)
+
+ifelse(Tval > crit, "Reject the null", "Cannot reject null")
+```
+
 b.) $\bar{X}_t = 43, s_w = 9$
 
+```{r, echo = T}
+xbar <- 43; s <- 9;
+
+Tval <- .6 * (xbar - h0) / s * sqrt(n)
+
+ifelse(Tval > crit, "Reject the null", "Cannot reject null")
+```
+
 c.) $\bar{X}_t = 43, s_w = 3$
+
+```{r, echo = T}
+xbar <- 43; s <- 3;
+
+Tval <- .6 * (xbar - h0) / s * sqrt(n)
+
+ifelse(Tval > crit, "Reject the null", "Cannot reject null")
+```
 
@@
 ### 7.33
+
+A standard measure of aggression in 7-year-old children has been found to have a 20% trimmed mean of 4.8 based on years of experience.
+A psychologiest wants to know whether the trimmed mean for children with divorced parents is greater than or less than 4.8.
+
+Suppose $\bar{X}_t = 5.1, s_w = 7$ based on $n = 25$
+
+Test $H_0 : \mu_t = 4.8, \alpha = 0.01$
+
+```{r, echo = T}
+
+```
+
 ### 7.34
+
+Summarize the relative merits of using a percentile bootstrap method.
+
+```{r, echo = T}
+
+```
+
 ### 7.35
+
+For the data in R variable _x_, imagine that you want to compute Z with the goal of testing the hypothesis that $\mu = 6$.
+Assuming that the R variable _null.value_ contains the hypothesized value and that _sigma_ contains the standard deviation, indicate the R command for computing Z.
+
+test <- (mean(x) - null.value) / sigma * length(x)
+
 ### 7.36
+
+R has a built-in data set called _ToothGrowth_.
+Information about the data can be obtained with the R command ?ToothGrowth.
+The first column contains measures of tooth growth in guinea pigs and the third column indicates the dose levels of vitaime C that were used in the experiment.
+
+Consider the claim that the typical growth is 8 when the vitamine C level is 0.5 mg.
+Using R, verify that with Student's T and $\alpha = 0.05$, it would be decided that the population mean is larger than 8.
+
+Verify that no decision would be made based on the Tukey-McLaughlin method with a 20% trimmed mean.
+
+Plot the data and comment why the two methods give different results.
+
 ### 7.37
+
+Based on the results for the previous exercise, speculate about whether you would reject using a median or MOM.
+
+Use R to check your speculation.

### 2019-12-21 Statistical Methods/07_hypothesis-testing.rmd
statistical methods, ch7 wrap-up
statistical methods, ch7 wrap-up

@@
 
+```{r, echo = T}
+x <- c(38, 44, 62, 72, 43, 40, 43, 42, 39, 41)
+
+trim <- .2; alpha <- 0.05; h0 <- 45; n <- length(x)
+xbar <- mean(x, tr = trim)
+sw <- 1.73
+
+g <- floor( trim * n )
+df <- n - 2*g - 1
+
+crit <- qt(1 - alpha/2, df)
+
+score <- ( 1 - 2*trim ) * (xbar - h0) / sw * sqrt(n)
+
+ifelse(abs(score) > crit, "Reject", "Fail to reject")
+```
+
 ### 7.33
@@
 ```{r, echo = T}
+xbar <- 5.1; sw <- 7; n <- 25
+alpha <- 0.01; h0 <- 4.8; trim <- .2
 
+g <- floor(n*trim)
+
+df <- n - 2*g - 1
+
+crit <- qt(1 - alpha/2, df)
+
+score <- (1 - 2*trim) * (xbar - h0) / sw * sqrt(n)
+
+ifelse(abs(score) > crit, "Reject", "Fail to reject")
 ```
@@
 Summarize the relative merits of using a percentile bootstrap method.
-
-```{r, echo = T}
-
-```
 
@@
 
+```{r, echo = T}
+flag=which(ToothGrowth[,3]==0.5)
+trimci(ToothGrowth[flag,1],null.value=8,tr=0)
+```
+
 Verify that no decision would be made based on the Tukey-McLaughlin method with a 20% trimmed mean.
 
+```{r, echo = T}
+trimci(ToothGrowth[flag,1],tr=0.2,null.value=8)
+```
+
 Plot the data and comment why the two methods give different results.
+
+```{r, echo = T}
+akerd(ToothGrowth[flag,1])
+```
 
@@
 Use R to check your speculation.
+
+```{r, echo = T}
+id=which(ToothGrowth[,3]==0.5)
+momci(ToothGrowth[id,1],null.value=8)
+```

### 2019-12-23 Statistical Methods/08_Regression.rmd
statistical methods
statistical methods

--- 2019-12-21 Statistical Methods/07_hypothesis-testing.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Hypothesis Testing}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Regression}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 7}
+   \lfoot{\color{headergrey}Chapter 8}
    \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
@@
 
-#### Chapter 7
+#### Chapter 8
 
-### 7.1
+### 8.1
+### 8.2
+### 8.3
+### 8.4
+### 8.5
+### 8.6
+### 8.7
+### 8.8
+### 8.9
+### 8.10
 
-Given that $\bar{X} = 78, \sigma^2 =25, n = 10, \alpha = 0.05$ test $H_0 : \mu \ge 80$, assuming that observations are randomly sampled from a normal distribution.
+### 8.21
+### 8.22
+### 8.23
+### 8.24
+### 8.25
+### 8.26
+### 8.27
+### 8.28
+### 8.29
+### 8.30
 
-```{r, echo = T}
-xbar <- 78; var <- 25; n <- 10
-alpha <- 0.05; h0 <- 80
-
-# greater than = left tail test
-
-crit <- qnorm(alpha)
-
-Z <- sqrt(n) * (xbar - h0) / sqrt( var )
-
-ifelse(Z < crit, "Reject the null", "Cannot reject null")
-```
-
-### 7.2
-
-Repeat the previous exercise, but test $H_0 : \mu = 80$
-
-```{r, echo = T}
-
-# equals = two-tailed test
-
-crit <- qnorm(alpha/2)
-
-Z <- sqrt(n) * (xbar - h0) / sqrt( var )
-
-ifelse(Z < crit || crit > Z, "Reject the null", "Cannot reject the null")
-
-```
-
-### 7.3
-
-For the data in Exercise 1, compute a 0.95 confidence interval and verify that this interval is consistent with y our decision about whether to reject the null hypothesis $H_0 : \mu = 80$.
-
-```{r, echo = T}
-
-xbar + qnorm(c(alpha/2, 1 - alpha/2)) * sqrt( var / n )
-
-```
-
-_80_ is in the confidence interval, so fail to reject.
-
-### 7.4
-
-For exercise 1, determine the p-value.
-
-```{r, echo = T}
-pnorm(Z)
-```
-
-### 7.5
-
-For exercise 2, determine the p-value.
-
-```{r, echo = T}
-2 * (1 - pnorm(Z, lower.tail = F))
-```
-
-### 7.6
-
-Given that $\bar{X} = 120, \sigma = 5, n = 49, \alpha = 0.05$, test $H_0 : \mu \ge 130$, assuming that observations are randomly sampled from a normal distribution.
-
-```{r, echo = T}
-xbar <- 120; sigma <- 5; n <- 49
-alpha <- 0.05; h0 <- 130
-
-crit <- qnorm(alpha)
-
-Z <- sqrt(n) * (xbar - h0) / sigma
-
-ifelse( Z < crit, "Reject null hypothesis", "Cannot reject null")
-```
-
-### 7.7
-
-Repeat the previous exercise but test $H_0 : \mu = 130$.
-
-```{r, echo = T}
-h0 <- 130
-
-# two tailed test
-crit <- qnorm(c(alpha/2, 1 - alpha/2))
-
-ifelse(Z < crit[1] || Z > crit[2], "Reject null hypothesis", "Cannot Reject Null")
-```
-
-### 7.8
-
-For the previous exercise, compute a 0.95 confidence interval and compare the result with your decision about whether to reject $H_0$.
-
-```{r, echo = T}
-xbar + qnorm(c(Lower = alpha/2, Upper = 1 - alpha/2)) * sigma/sqrt(n)
-```
-
-130 is not in the confidence region, reject null-hypothesis.
-
-### 7.9
-
-If $\bar{X} = 23$ and $\alpha = 0.025$, can you make a decision about whether to reject $H_0 : \mu \le 25$ without knowing $\sigma$?
-
-_Yes, the hypothesis is consistent with the sample mean._
-
-### 7.10
-
-An electronics firm mass-produces a component for which there is a standard measure of quality.
-Based on testing vast numbers of these components, the company has found that the average quality is $\mu = 232, \sigma = 4$.
-
-However, in recent years, the quality has not been checked, so management asks you to check their claim with the goal of being reasonable certain that an average quality of less than 232 can be ruled out.
-That is, assume that the quality is poor and in fact less than 232 with the goal of empirically establishing that this assumption is unlikely.
-
-You get $\bar{X} = 240$ based on a sample of $n = 25$ components, and you want the probability of a Type I error to be less than 0.01.
-
-State the null hypothesis and perform the appropriate test, assuming normality.
-
-#### Hypothesis:
-
-$H_0 : \mu \le 232$
-
-$H_A : \mu \ge 232$
-
-```{r, echo = T}
-xbar <- 240; sigma <- 4; n <- 25
-alpha <- 0.01; h0 <- 232
-
-# less than = upper-tailed test
-
-crit <- qnorm(1 - alpha)
-
-Z <- sqrt(n) * ( xbar - h0 ) / sigma
-
-ifelse(Z > crit, "Reject the null", "Cannot reject the null")
-```
-
-### 7.11
-
-An antipollution device for cars is claimed to have an average effectiveness of exactly 546. Based on a test of 20 such devices, you find that $\bar{X} = 565$.
-
-Assuming normality and that $\sigma = 40$, would you rule out the claim with a Type I error probability of 0.05?
-
-```{r, echo = T}
-xbar <- 565; sigma <- 40; n <- 20
-alpha <- 0.05; h0 <- 546
-
-# equality = two tailed
-
-crit <- qnorm(c(alpha/2, 1 - alpha/2))
-
-Z <- sqrt(n) * (xbar - h0) / sigma
-
-ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
-```
-
-### 7.12
-
-Comment on the relative merits of using a 0.95 confidence interval for addressing the effectiveness of the antipollution device in the previous exercise rather than test the hypothesis: $H_0 : \mu = 232$
-
-```{r, echo = T}
-xbar + qnorm(c(alpha/2, 1 - alpha/2)) * sigma/sqrt(n)
-
-2 * ( 1 - pnorm(Z) )
-```
-
-The hypothesised value doesn't fall within the confidence interval, however the formalized test gives us a degree of strength on which we can reject the null.
-Here our p-value is .03, so we reject with slight confidence.
-
-\newpage
-
-### 7.13
-
-For $n = 25, \alpha = 0.01, \sigma = 5, H_0 : \mu \ge 60$, verify that power is .95 when $\mu = 56$
-
-```{r, echo = T}
-
-mu <- 56; sigma <- 5; n <- 25
-alpha <- 0.01; h0 <- 60
-
-# is mu greater than 60? gt = lower tail test
-
-crit <- qnorm(alpha)
-
-Z <- sqrt(n) * ( mu - h0) / sigma
-
-pnorm(crit - Z) # power
-```
-
-### 7.14
-
-For $n = 36, \alpha = 0.025, \sigma = 8, H_0 : \mu \le 100$, verify that power is .61 when $\mu = 103$
-
-```{r, echo = T}
-mu <- 103; sigma <- 8; n <- 36
-alpha <- 0.025; h0 <- 100
-
-# less than, upper tailed test
-
-crit <- qnorm(alpha, lower.tail = F)
-
-Z <- sqrt(n) * (mu - h0) / sigma
-
-ifelse( Z < crit, "Reject null", "Cannot reject null")
-
-pnorm(crit - Z, lower.tail = F) # power
-```
-
-\newpage
-
-### 7.15
-
-For $n = 49, \alpha = 0.05, \sigma = 10, H_0 : \mu = 50$, verify that power is approximately .56 when $\mu = 47$
-
-```{r, echo = T}
-mu <- 47; sigma <- 10; n <- 49
-alpha <- 0.05; h0 <- 50
-
-# mu = 50? two-tailed test
-
-crit <- qnorm(c(Lower = alpha/2, Upper = 1 - alpha/2))
-
-Z <- sqrt(n) * (mu - h0) / sigma
-
-ifelse( Z < crit[1] || Z > crit[2], "Reject null hypothesis", "Cannot reject null")
-
-as.vector(pnorm(crit[1] - Z) + pnorm(crit[2] - Z, lower.tail = F)) # power
-```
-
-### 7.16
-
-A manufacturer of medication for migraine headaches knows that their product can damage the stomach if taken too often.
-Imagine that by a standard measuring process, the average damage is $\mu = 48$.
-A modification of their product is being contemplated, and based on 10 trials, it is found that $\bar{X} = 46$.
-Assuming $\sigma = 5$ they test $H_0 : \mu \ge 48$, the idea being that if they reject, there is convincing evidence that the average amount of damage is less than 48.
-Then,
-
-$Z = \frac{46 - 48}{5/\sqrt10} = -1.3$
-
-With $\alpha = 0.05$, the critical value is -1.645, so they do not reject Z is less than the critical value.
-
-What might be wrong with accepting $H_0$ and concluding that the modificaiton results in an average amount of damage greater than or equal to 48?
-
-```{r, echo = T}
-mu <- 46; sigma <- 5; n <- 10
-alpha <- 0.05; h0 <- 48
-
-crit <- qnorm(alpha)
-
-Z <- (mu - h0) / sigma * sqrt(n)
-
-pnorm(crit - Z)
-```
-
-Low power in the test, might commit a Type II error.
-
-### 7.17
-
-For the previous exercise, verify that the power is 0.35 if $\mu = 46$
-
-```{r, echo = T}
-pnorm(crit - Z)
-```
-
-### 7.18
-
-The previous exercise indicates that power is relatively low with only $n = 10$ observations.
-Imagine that you want power to be at least .8.
-One way of getting more power is to increase the sample size, _n_.
-
-Verify that for sample sizes of 20, 30 and 40, power is .56, 0.71, and .81, respectively.
-
-```{r, echo = T}
-n <- c(20, 30, 40)
-
-Z <- (mu - h0) / sigma * sqrt(n)
-
-pnorm(crit - Z)
-```
-
-### 7.19
-
-For the previous exercise, rather than increase the sample size, what else might you do to increase the power? What is the negative consequence for this strategy?
-
-_An alternative to increasing the sample size is to increase alpha, which will increase the probability of a type I error._
-
-### 7.20
-
-Given the following values for $\bar{X}$ and s:
-
-test the hypothesis $H_0 : \mu = 42, \alpha = 0.05, n = 25$
-
-a.) $\bar{X} = 44, s = 10$
-
-```{r, echo = T}
-xbar <- 44; s <- 10; n <- 25
-alpha <- 0.05; h0 <- 42
-
-crit <- qt(c(alpha/2, 1 - alpha/2), df = n - 1)
-
-Z <- (xbar - h0) / s * sqrt(n)
-
-ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
-```
-
-b.) $\bar{X} = 43, s = 10$
-
-```{r, echo = T}
-xbar <- 43; s <- 10; n <- 25
-alpha <- 0.05; h0 <- 42
-
-crit <- qt(c(alpha/2, 1 - alpha/2), df = n - 1)
-
-Z <- (xbar - h0) / s * sqrt(n)
-
-ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
-```
-
-c.) $\bar{X} = 43, s = 2$
-
-```{r, echo = T}
-xbar <- 43; s <- 2; n <- 25
-alpha <- 0.05; h0 <- 42
-
-crit <- qt(c(alpha/2, 1 - alpha/2), df = n - 1)
-
-Z <- (xbar - h0) / s * sqrt(n)
-
-ifelse(Z < crit[1] || Z > crit[2], "Reject the null", "Cannot reject the null")
-```
-
-### 7.21
-
-For part b of the last exercise, you fail to reject but you reject for the situation in part c. What does this illustrate about power?
-
-_Power is higher with smaller variance._
-
-### 7.22
-
-Test the hypothesis $H_0 : \mu \le 42$ with $\alpha = 0.05, n = 16$ for the following values of $\bar{X}$ and s:
-
-a.) $\bar{X} = 44, s = 10$
-
-```{r, echo = T}
-xbar <- 44; s <- 10; n <- 16
-alpha <- 0.05; h0 <- 42
-
-# less than, upper tailed test
-crit <- qt(alpha, df = n - 1, lower.tail = F)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse( Tval > crit, "reject the null", "fail to reject")
-```
-
-b.) $\bar{X} = 43, s = 10$
-
-```{r, echo = T}
-xbar <- 43; s <- 10
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse( Tval > crit, "reject the null", "fail to reject")
-```
-
-c.) $\bar{X} = 43, s = 2$
-
-```{r, echo = T}
-xbar <- 43; s <- 2
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse(Tval > crit, "reject the null", "fail to reject")
-```
-
-### 7.23
-
-Repeat the previous for $H_0 : \mu \ge 42$
-
-a.) $\bar{X} = 44, s = 10$
-
-```{r, echo = T}
-xbar <- 44; s <- 10
-h0 <- 42
-
-# greater than, lower tailed test
-crit <- qt(alpha, df = n - 1)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse( Tval < crit, "reject the null", "fail to reject null")
-```
-
-b.) $\bar{X} = 43, s = 10$
-
-```{r, echo = T}
-xbar <- 43; s <- 10
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse( Tval < crit, "reject the null", "fail to reject null")
-```
-
-c.) $\bar{X} = 43, s = 2$
-
-```{r, echo = T}
-xbar <- 43; s <- 2
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse( Tval < crit, "reject the null", "fail to reject null")
-```
-
-### 7.24
-
-A company claims that on average, when their toothpaste is used, 45% of all bacteria related to gingivitus is killed.
-
-For 10 individuals, it is found that the percentages of bacteria killed are 38, 44, 62, 72, 43, 40, 43, 42, 39, 41.
-
-The mean and standard deviation of these values are $\bar{X} = 46.4, s = 11.27$, respectively.
-
-Assuming normality, test the hypothesis that the average percentage is 45 with $\alpha = 0.05$
-
-```{r, echo = T}
-x <- c(38, 44, 62, 72, 43, 40, 43, 42, 39, 41)
-
-xbar <- mean(x); s <- sd(x); n <- length(x)
-alpha <- 0.05; h0 <- 45
-
-crit <- qt(1 - alpha/2, df = n - 1)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse(abs(Tval) > crit, "Reject the null", "Fail to reject the null")
-
-pt(Tval, df = n - 1) # p-value
-```
-
-\newpage
-
-### 7.25
-
-A portion of a study by Wechsler (1958) reports that for 100 males taking the Wechsler Adult Intelligent Scale (WAIS), the sample mean and variance on picture completion are $\bar{X} = 9.79$ and $s = 2.72$, respectively.
-
-Test the hypothesis $H_0 : \mu \ge 10.5, \alpha = 0.025$
-
-```{r, echo = T}
-xbar <- 9.79; s <- 2.72; n <- 100
-alpha <- 0.025; h0 <- 10.5
-
-crit <- qt(alpha, df = n - 1)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse(Tval < crit, "reject the null", "fail to reject")
-
-pt(Tval, df = n - 1) # p-value
-```
-
-### 7.26
-
-Given that $n = 16, \bar{X} = 40, and s = 4$, test that $H_0 : \mu \le 38, \alpha = 0.01$
-
-```{r, echo = T}
-xbar <- 40; s <- 4; n <- 16
-alpha <- 0.01; h0 <- 38
-
-crit <- qt(alpha, df = n - 1, lower.tail = F)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse(Tval > crit, "Reject null", "Fail to reject null")
-
-pt(Tval, df = n - 1)
-```
-
-### 7.27
-
-Given that $n = 9, \bar{X} = 33, s = 4$ test $H_0 : \mu = 32, \alpha = 0.05$
-
-```{r, echo = T}
-xbar <- 33; s <- 4; n <- 9
-alpha <- 0.05; h0 <- 32
-
-crit <- qt(1 - alpha/2, df = n - 1)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse(abs(Tval) > crit, "Reject null", "Cannot reject null")
-```
-
-### 7.28
-
-An engineer believes it takes an average of 150 man-hours to assemble a portion of an automobile. As a check, the time to assemble 10 such parts was ascertained, yielding $\bar{X} = 146, s = 2.5$.
-
-Test the engineer's belief with $\alpha = 0.05$
-
-```{r, echo = T}
-xbar <- 146; s <- 2.5; n <- 10
-alpha <- 0.05; h0 <- 150
-
-crit <- qt(1 - alpha/2, df = n - 1)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse(abs(Tval) > crit, "Reject the null", "Fail to Reject")
-```
-
-### 7.29
-
-In a study of court administration, the following times to disposition were determined for 20 cases and found to be:
-
-42, 90, 84, 87, 119, 95, 86, 99, 93, 92, 121, 71, 66, 98, 79, 102, 60, 112, 105, 98
-
-Test the hypothesis that the average time to disposition is less than or equal to 80 at $\alpha = 0.01$.
-
-```{r, echo = T}
-x <- c(42, 90, 84, 87, 119, 95, 86, 99, 93, 92, 121, 71, 66, 98, 79, 102, 60, 112, 105, 98)
-
-xbar <- mean(x); s <- sd(x); n <- length(x)
-alpha <- 0.01; h0 <- 80
-
-crit <- qt(alpha, df = n - 1, lower.tail = F)
-
-Tval <- (xbar - h0) / s * sqrt(n)
-
-ifelse(Tval > crit, "Reject the null", "Cannot reject the null")
-
-pt(Tval, df = n - 1, lower.tail = F) # p-value
-
-```
-
-Or, simply:
-
-```{r, echo = T}
-t.test(x, alternative = 'greater', conf.level = 1 - alpha, mu = h0)
-```
-
-### 7.30
-
-Assuming 20% trimming, test the hypothesis $H_0 : \mu = 42, \alpha = 0.05, n = 20$ given the following values for $\bar{X}$ and $s_w$:
-
-a.) $\bar{X}_t = 44, s_w = 9$
-
-b.) $\bar{X}_t = 43, s_w = 9$
-
-c.) $\bar{X}_t = 43, s_w = 3$
-
-### 7.31
-
-Repeat the previous exercise, only test the hypothesis $H_0 : \mu_t \le 42, \alpha = 0.05$
-
-a.) $\bar{X}_t = 44, s_w = 9$
-
-b.) $\bar{X}_t = 43, s_w = 9$
-
-c.) $\bar{X}_t = 43, s_w = 3$
-
-### 7.32
-
-For the data in Exercise 24, the 20% trimmed mean is $\bar{X}_t = 42.17$ with a Winsorized standard deviation of $S_w = 1.73$
-
-Test the hypothesis that the population trimmed mean is 45 with $\alpha = 0.05$
-
-### 7.33
-### 7.34
-### 7.35
-### 7.36
-### 7.37
+### 8.31
+### 8.32
+### 8.33
+### 8.34
+### 8.35
+### 8.36

### 2019-12-24 Statistical Methods/08_Regression.rmd
statistical methods
statistical methods

@@
 
-data.dir <- "D:/Projects/Statistical-Computing/datasets/"
+data.dir <- paste0(here::here(), "/datasets/")
 
-install.packages("WRS")
 ```
@@
 ### 8.1
+
+For the following data, use R to verify that the least squares regresion line is $\hat{Y} = 1.8X - 8.5$
+
+X: 5, 8, 9, 7, 14
+Y: 3, 1, 6, 7, 19
+
+```{r, echo = T}
+dat <- data.table(
+   X = c(5, 8, 9, 7, 14),
+   Y = c(3, 1, 6, 7, 19))
+
+lm(Y ~ X, data = dat)
+```
+
+Also verify that the Theil-Sen estimator, the slope is estimated to be 1.746 and the intercept is estimated to be -7.968.
+
+```{r, echo = T}
+tsreg(dat$X, dat$Y)$coef
+```
+
 ### 8.2
+
+Using the R function _lsfit_, compute the residuals using the data in E1,
+
+Verify that if you square and sum the residusls, you get 46.585.
+
+```{r, echo = T}
+res <- lsfit(dat$X, dat$Y)$resid
+
+sum(res^2)
+```
+
 ### 8.3
+
+Verify that for the data in E1, if you use $\hat{Y} = 2X - 9$, the sum of the squared residuals is greater than 46.584.
+
+```{r, echo = T}
+Yhat <- 2*dat$X - 9
+res <- sum( (dat$Y - Yhat)^2 )
+stopifnot(res > 46.583)
+res
+
+```
+
+Why would you expect a value greater than 46.584?
+
+_The x coefficent increased._
+
 ### 8.4
+
+Suppose that based on $n = 25$ values, $s^2_x = 12$ and $\sum(X_i - \bar{X})(Y_i - \bar{Y}) = 144$.
+
+What is the slope of the least squares regression?
+
+$A = 144, C = (n - 1)s^2_x = 288, b_1 = A / C = 144/288 = .5$
+
 ### 8.5
+
+The following table reports breast cancer rates plus levels of solar radiation (in calories per day) for various cities in the United States. The data are stored in the file cancer_rate_dat.txt.
+
+```{r, echo =T}
+
+dat <- data.table::fread(paste0(data.dir, "cancer_rate_dat.txt"), fill = T, sep = "&")
+
+dat
+```
+
+Fit a OLS regression to predict cancer rates and comment on what this suggests.
+
+```{r, echo = T}
+fit <- lm(Rate ~ calories, data = dat)
+
+plot(fit)
+```
+
 ### 8.6
+
+For the following data, use R to compute the least squares regresion line for predicting GPA given SAT.
+
+SAT: 500, 530, 590, 660, 610, 700, 570, 640
+GPA: 2.3, 3.1, 2.6, 3.0, 2.4, 3.3, 2.6, 3.5
+
+```{r, echo = T}
+dat <- data.table(
+   SAT = c(500, 530, 590, 660, 610, 700, 570, 640),
+   GPA = c(2.3, 3.1, 2.6, 3.0, 2.4, 3.3, 2.6, 3.5)
+)
+
+fit <- lm(GPA ~ SAT, data = dat)
+coef(fit)
+```
+
 ### 8.7
+
+Compute the residuals for the data used in the previous problem and verify that sum to zero.
+
+```{r, echo = T}
+round(sum((dat$GPA - fit$fitted.values)), 5)
+```
+
 ### 8.8
+
+For the following data, use R to compute the least squares regression line for predicting Y from X.
+
+X: 40, 41, 42, 43, 44, 45, 46
+Y: 1.62, 1.63, 1.90, 2.64, 2.05, 2.13, 1.94
+
+```{r, echo = T}
+dat <- data.table(
+   X = c(40, 41, 42, 43, 44, 45, 46),
+   Y = c(1.62, 1.63, 1.90, 2.64, 2.05, 2.13, 1.94)
+)
+
+summary(fit <- lm(Y ~ X, data = dat))
+```
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+ggplot(dat, aes(X, Y)) +
+   geom_point() +
+   geom_smooth(method = "lm")
+```
+
 ### 8.9

### 2019-12-24 Statistical Methods/08_Regression.rmd
statistical methods
statistical methods

@@
 ### 8.9
+
+In exercise 5, what would be the least squares estimate of the cancer rate given a solar radiation of 600?
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+dat <- data.table::fread(paste0(data.dir, "cancer_rate_dat.txt"), fill = T, sep = "&")
+
+ggplot(dat, aes(calories, Rate)) +
+   geom_point() +
+   geom_smooth(method = "lm")
+
+fit <- lm(Rate ~ calories, data = dat)
+
+coef(fit)
+
+39.99 -0.037*600
+```
+Why might this be unreasonable?
+
+_Because 600 is outside of the bounds of seen values (extrapolation)._
+
 ### 8.10

### 2019-12-27 Statistical Methods/08_Regression.rmd
statistical methods
statistical methods

@@
 
+Maximal oxygen uptake (MOU) is a measure of an individual's physical fitness. You want to know how MOU is related to how fast someone can run a mile. Suppose you randomly sample six athletes and get:
+
+MOU (kl/kg): 63.3, 60.1, 53.6, 58.8, 67.5, 62.5
+Time (seconds): 241.5, 249.8, 246.1, 232.4, 237.2, 238.4
+
+Compute the least squres regression line and comment on what the results suggest.
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+dat <- data.table(
+   MOU = c(63.3, 60.1, 53.6, 58.8, 67.5, 62.5),
+   Time = c(241.5, 249.8, 246.1, 232.4, 237.2, 238.4)
+)
+
+fit <- lm(Time ~ MOU, data = dat)
+
+ggplot(dat, aes(MOU, Time)) +
+   geom_point() +
+   geom_smooth(method = "lm") +
+   labs(title = "Time vs. MOU")
+
+```
+
+Generally, time decreases as MOU increases.
+
+### 8.11
+
+Verify that for the following pairs of points, the least squares regression line has a slope of zero. Plot the points and comment on the assumption that the regression line is straight.
+
+X: 1, 2, 3, 4, 5, 6
+Y: 1, 4, 7, 7, 4, 1
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+dat <- data.table(
+   X = c(1, 2, 3, 4, 5, 6),
+   Y = c(1, 4, 7, 7, 4, 1)
+)
+
+fit <- lm(Y ~ X, data = dat)
+
+coef(fit)
+```
+
+### 8.12
+
+Repeat the last exercise, only for points:
+
+X: 1, 2, 3, 4, 5, 6
+Y: 4, 5, 6, 7, 8, 2
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+dat <- data.table(
+   X = c(1, 2, 3, 4, 5, 6),
+   Y = c(4, 5, 6, 7, 8, 2)
+)
+
+fit <- lm(Y ~ X, data = dat)
+
+coef(fit)
+```
+
+### 8.13
+
+Vitamin A is required for good health. However, one bite of polar bear liver results in death becasue it contains a high concentration of vitamin A.
+
+Comment on this fact in terms of extrapolation.
+
+_Vitaim A as a function of health is a bounded variable. Predicting a response when the independent variables are far from what has been observed can lead to undefined results (death)._
+
+### 8.14
+
+Socket et al.
+(1987) report data related to patterns of residual insulin secretion in children.
+A portion of the study was concerned with whether age can be used to predict the logarithm of C-peptide concentrations at diagnosis.
+The observed values are (data file):
+
+Replicate the _LOESS_ smoothed curve in fig 8.4.1.
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+dat <- data.table::fread(paste0(data.dir, "diabetes_sockett_dat.txt"), fill = T)
+
+ggplot(dat, aes(age, pep)) +
+   geom_point() +
+   geom_smooth(method = "loess")
+
+```
+
+### 8.15
+
+### 8.16
+### 8.17
+### 8.18
+### 8.19
+### 8.20
+
 ### 8.21

### 2019-12-27 Statistical Methods/08_Regression.rmd
statistical methods
statistical methods

@@
 
-Vitamin A is required for good health. However, one bite of polar bear liver results in death becasue it contains a high concentration of vitamin A.
-
-Comment on this fact in terms of extrapolation.
+Vitamin A is required for good health. However, one bite of polar bear liver results in death becasue it contains a high concentration of vitamin A. Comment on this fact in terms of extrapolation.
 
@@
 A portion of the study was concerned with whether age can be used to predict the logarithm of C-peptide concentrations at diagnosis.
-The observed values are (data file):
+The observed values are (data file), Replicate the _LOESS_ smoothed curve in fig 8.4.1.
 
-Replicate the _LOESS_ smoothed curve in fig 8.4.1.
-
-```{r, echo = T, fig.width=8, fig.height=3.5}
+```{r, echo = T, fig.width=8, fig.height=2.5}
 dat <- data.table::fread(paste0(data.dir, "diabetes_sockett_dat.txt"), fill = T)
@@
 
+For the data in the last exercise, use R to verify that a least squares regression line using only X values (age) less than or equal to 7 yields a p-value equal to 0.026 when using the R function _olsch4_.
+Also verify that the p-value, when using the Theil-Sen estimator, is 0.0233.
+
+```{r, echo = T}
+dat2 <- dat[age <= 7]
+
+WRS::olshc4(dat2$age, dat2$pep)
+
+WRS::regci(dat2$age, dat2$pep)
+```
+
 ### 8.16
+
+For the reading data in Table 8.5, verify that the R function spearci returns a p-value equal to 0.014 and that scorci returns a p-value equal to 0.002.
+Based on the plot returned by scorci, why is it not suprising that these two function give similar results?
+
+```{r, echo = T, fig.width=8, fig.height=3.5}
+dat <- data.table::fread(paste0(data.dir, "read_dat.txt"), fill = T, skip = 13)
+
+spearci( dat[,4], dat[,8])
+
+scorciMC( dat[,4], dat[,8])
+
+```
+
 ### 8.17
+
+Given that $b_1 = -1.5, n = 10, s^2_{y,x} = 35, \sum(X_i - \bar{X})^2 = 140$, assume normality and homosecedasiticy and compute a 0.95 confidence interval for slope, $\beta_1$.
+
+```{r, echo = T}
+b1 <- -1.5; n <- 10; se <- sqrt(35 / 140)
+alpha <- .05
+
+b1 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * se
+```
+
 ### 8.18
+
+Repeat the previous problem, only compute a 0.98 confidence interval.
+
+```{r, echo = T}
+b1 <- -1.5; n <- 10; se <- sqrt(35 / 140)
+alpha <- 1 - .98
+
+b1 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * se
+```
+
 ### 8.19
+
+Based on results convered in the previous chapters, speculate about why the confidence intervals computed in the last two problems might be inaccurate.
+
+_Least squares regression can be negatively impacted by non-normality, heteroscedasticity and outliers._
+
 ### 8.20
 
+Assume normality and homoscedasticity and suppose $n = 30, \sum{(X_i)} = 15, \sum(Y_i) = 30, \sum(X_i - \bar{X})(Y_i - \bar{Y}) = 30, \sum(X_i - \bar{X})^2 = 10$
+
+```{r, echo = T}
+n <- 30; A <- 30; C <- 10;
+xbar <- 15 / 30; ybar <- 30 / 30
+
+slope <- A / C
+intercept <- ybar - xbar*slope
+
+```
+
+Determine the least squares estimates of the slope and intercept.
+
+$\beta_1 = `r slope`, \beta_0 = `r intercept`$
+
 ### 8.21
+
+Assume normality and homoscedasticity and suppose $n = 38, \bar{Y} = 20, \sum{X^2_i} = 1,922, \bar{X} = 7, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 180, \sum{X_i - \bar{X}^2} = 60, s^2_{X, Y} = 121.$
+
+a.) Determine the least squares estimate of the slope and intercept.
+
+```{r, echo = T}
+```
+
+b.) Test the hypothesis $H_0 : \beta_0 = 0, \alpha = 0.05$
+
+```{r, echo = T}
+```
+
+c.) Compute a 0.9 confidence interval for $\beta_1$
+
+```{r, echo = T}
+```
+
 ### 8.22
+
+Assume normality and homoscedasticity and suppose $n = 41, \bar{Y} = 10, \bar{X} = 12, \sum(X_i - \bar{X})(Y_i - \bar{Y}) = 100, \sum(X_i - \bar{X})^2 = 400, s^2_{Y.X} = 144$.
+
+a.) Determine the least squares regression line
+
+b.) Compute a .9 confidence interval for $\beta_1$
+
 ### 8.23
+
+Assume normality and homoscedasticity and suppose $n = 18, \beta_1 = 3.1, \sum{(X_i - \bar{X}})^2 = 144, s^2_{X.Y} = 36$.
+
+Compute a 0.95 confidence interval for $\beta_1$.
+
+Would you conclude $\beta_1$ > 2?
+
 ### 8.24
+
+Assume normality and homoscedasticity and suppose $n = 20, \beta_0 = 6, \sum{X^2_i} = 169, S^2_{Y.X} = 25, \sum{(X_i - \bar{X}})^2 = 90$.
+
+Compute a .95 confidence interval for $\beta_0$.
+
 ### 8.25
+
+Given the following quantities, find the sample correlation coefficent, r, and test $H_0 : \rho = 0$ at the indicated value for $\alpha$.
+
+a.) $n = 27, \sum{(Y_i - \bar{Y})}^2 = 100, \sum{(X_i - \bar{X})^2} = 625, \sum(X_i -\bar{X})(Y_i - \bar{Y})^2 = 200, \alpha = 0.01$
+
+b.) $n = 5, \sum{(Y_i - \bar{Y})}^2 = 16, \sum{(X_i - \bar{X})^2} = 25, \sum(X_i -\bar{X})(Y_i - \bar{Y})^2 = 10, \alpha = 0.05$
+
 ### 8.26
+
 ### 8.27
+
 ### 8.28
+
 ### 8.29
+
 ### 8.30

### 2019-12-27 Statistical Methods/08_Regression.rmd
statistical methods, ch8
statistical methods

@@
 
-Assume normality and homoscedasticity and suppose $n = 38, \bar{Y} = 20, \sum{X^2_i} = 1,922, \bar{X} = 7, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 180, \sum{X_i - \bar{X}^2} = 60, s^2_{X, Y} = 121.$
+Assume normality and homoscedasticity and suppose $n = 38, \bar{Y} = 20, \sum{X^2_i} = 1,922, \bar{X} = 7, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 180, \sum{(X_i - \bar{X})^2} = 60, s^2_{X.Y} = 121.$
 
@@
 ```{r, echo = T}
+n <- 38; ybar <- 20; xbar <- 7
+variance <- 121; A <- 180; C <- 60
+ssx <- 1922
+
+slope <- A / C
+intercept <- ybar - xbar*slope
+
 ```
 
-b.) Test the hypothesis $H_0 : \beta_0 = 0, \alpha = 0.05$
+$\beta_1 = `r slope`, \beta_0 = `r intercept`$
+
+b.) Test the hypothesis $H_0 : \beta_0 = 0, \alpha = 0.02$
 
 ```{r, echo = T}
+alpha <- 0.02
+
+slope + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt(variance / n)
+
+Tval <- intercept * sqrt( ((n - 2) * C) / (variance  * ssx) )
+
+crit <- qt(alpha/2, df = n - 2)
+
+ifelse(abs(Tval) < crit, "Reject Null", "Fail to Reject")
 ```
@@
 ```{r, echo = T}
+alpha <- 1 - .9
+
+slope + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( variance / C)
 ```
@@
 
+```{r, echo = T}
+n <- 41; ybar <- 10; xbar <- 12; 
+ssr <- 100; ssx <- 400; variance <- 144
+
+b1 <- ssr / ssx
+
+b0 <- ybar - xbar*b1
+```
+
+$\beta_0 = `r b0`, \beta_1 = `r b1`$
+
 b.) Compute a .9 confidence interval for $\beta_1$
+
+```{r, echo = T}
+alpha <- 1 - .9
+
+b1 + qt(c( Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( variance / ssx)
+```
 
@@
 
+```{r, echo = T}
+n <- 18; b1 <- 3.1; ssx <- 144; variance <- 36
+alpha <- 1 - .95
+
+b1 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( variance / ssx )
+
+```
+
 Would you conclude $\beta_1$ > 2?
+
+_Given the confidence interval above, it seems reasonable._
 
@@
 
+```{r, echo = T}
+n <- 20; b0 <- 6; ssx <- 169
+variance <- 25; xvar <- 90
+alpha <- 1 - .95
+
+b0 + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df = n - 2) * sqrt( (variance*ssx) / (n * xvar))
+```
+
 ### 8.25
@@
 
+```{r, echo = T}
+n <- 27; yvar <- 100; xvar <- 625
+ssr <- 200; alpha <- 0.01
+
+r <- ssr / sqrt( yvar * xvar )
+
+test.stat.T <- r * sqrt( (n-2)/ (1 - r^2))
+
+crit <- qt(1 - alpha/2, df = n - 2)
+
+ifelse(abs(test.stat.T) >= crit, "Reject Null", "Fail to Reject")
+```
+
 b.) $n = 5, \sum{(Y_i - \bar{Y})}^2 = 16, \sum{(X_i - \bar{X})^2} = 25, \sum(X_i -\bar{X})(Y_i - \bar{Y})^2 = 10, \alpha = 0.05$
+
+```{r, echo = T}
+n <- 5; yvar <- 16; xvar <- 25
+ssr <- 10; alpha <- 0.05
+
+r <- ssr / sqrt( yvar * xvar )
+
+test.stat.T <- r * sqrt( (n-2)/ (1 - r^2))
+
+crit <- qt(1 - alpha/2, df = n - 2)
+
+ifelse(abs(test.stat.T) >= crit, "Reject Null", "Fail to Reject")
+```
 
@@
 
+The high school grade-point average (X) and college grade-point (Y) for 29 randomly sampled college freshmen yielded the following results:
+
+$\sum{(Y_i - \bar{Y})}^2 = 64, \sum{(X_i - \bar{X})}^2 = 100, \sum{(X_i - \bar{X})(Y_i - \bar{Y})} = 40$
+
+Test $H_0 : \rho = 0, \alpha = 0.1$ and interprest the results.
+
+```{r, echo = T}
+n <- 29; yvar <- 64; xvar <- 100; ssr <- 40
+alpha <- 0.1
+
+r <- ssr / sqrt( yvar * xvar )
+
+crit <- qt(1 - alpha/2, df = n - 2)
+
+test.stat.T <- r * sqrt( (n - 2) / (1 - r^2) )
+
+ifelse( abs(test.stat.T) >= crit, "Reject null", "Fail to reject")
+```
+
 ### 8.27
+
+For the previous exercise, answer the following questions:
+
+a.) Is it reasonable to conclude that the least squares regression line has a positive slope?
+
+_Yes, it appears there is a positive correlation._
+
+b.) Is it possible that despite the value for r, as high school grade-point average increases, college grade-point average decreases? Explain.
+
+_It is possible, although unlikely._
+
+c.) What might you do, beyond considering r, to decide it is reasonable to conclude that as high school grade-point averages increase, college grade point averages increase as well?
+
+_Plot the data to assess the visual indication of the relationship._
 
@@
 
+Using R, determine what happens to Pearson's correlation between X and Y if the Y values are multiplied by 3.
+Argue that if Y is multiplied by any constant c != 0, Pearson's correlation does not change.
+
+```{r, echo = T}
+set.seed(46)
+x <- rnorm(30); y <- rnorm(30)
+cor(x, y)
+
+cor(3*x, y)
+
+```
+
 ### 8.29
+
+Repeat the previous problem, only determine what happens to the slope of the least squares regresion line.
+
+```{r, echo = T}
+set.seed(46)
+x <- rnorm(30); y <- rnorm(30)
+
+lsfit(x, y)$coef
+
+lsfit(x, 3*y)$coef
+```
+
+The absolute value of the slope gets larger.
 
@@
 
+Consider a least squares regression line $Y = 0.52X + 2$, assume homoscedasiticy as consider the situation where the common variance is $\sigma^2 = 1$?
+What happens to the correlation coefficent between X and Y if instead $\sigma^2 = 2$?
+
+_It will increase the variance so the correlation will decrease._
+
 ### 8.31
+
+The numerator of the coefficient of determination is $\sum{(Y_i - \bar{Y}) - \sum{(Y_i - \hat{Y})}}$. Based on the least squares principal, why is the value always greater than zero?
+
+_If $\bar{Y}$ is used to predict Y, $\sum{(Y_i - \bar{Y})}$ will be larger than $\sum{(Y_i - \hat{Y})}$._
+
 ### 8.32
+
+Imagine a study where the correlation between some amount of an experimental drug and liver damage yields a value for r close to zero and the hypothesis $H_0 : \rho = 0$ is not rejected.
+Why might it be unreasonable to conclude that the two variabels are independent?
+
+_The magnitude of the residuals, curvature, outliers._
+
 ### 8.33
+
+Suppose $r^2 = 0.95$. Explain why this does not provide convincing evidence that the least squares line provides a good fit to a scatterplot of the points.
+
+_Outliers can cause large $r^2$, however, be a poor fit._
+
 ### 8.34
+
+Imagine a situation where points are removed for which the X values are judged to be outliers.
+Note that this restricts the range of X values.
+Without looking at the data, can someone predict whether Pearson's correlation will increase or decrease after these points are removed?
+
+_No. You need to look at the data. Restricting the range of X can increase as well as decrease r._
+
 ### 8.35
+
+If the normality assumption is violated, what effect might this have when computing confidence intervals for the slope and intercept as described in 8.4.1?
+
+_The condence interval can be relatively long and is potentially inaccurate._
+
 ### 8.36
+
+If the homoscedasticity assumption is violated, what effect might this have when computing confidence intervals as described in 8.4.1?
+
+_The condence interval can be relatively inaccurate due to using the wrong standard error._

### 2019-12-27 Statistical Methods/09_Comparing_Two_Independent_Groups.rmd
statistical methods, ch9/10
statistical methods, ch9/10

--- 2019-12-27 Statistical Methods/08_Regression.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Regression}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Comparing Two Independent Groups}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 8}
+   \lfoot{\color{headergrey}Chapter 9}
    \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
@@
 
-#### Chapter 8
+#### Chapter 9
 
-### 8.1
-
-For the following data, use R to verify that the least squares regresion line is $\hat{Y} = 1.8X - 8.5$
-
-X: 5, 8, 9, 7, 14
-Y: 3, 1, 6, 7, 19
-
-```{r, echo = T}
-dat <- data.table(
-   X = c(5, 8, 9, 7, 14),
-   Y = c(3, 1, 6, 7, 19))
-
-lm(Y ~ X, data = dat)
-```
-
-Also verify that the Theil-Sen estimator, the slope is estimated to be 1.746 and the intercept is estimated to be -7.968.
-
-```{r, echo = T}
-tsreg(dat$X, dat$Y)$coef
-```
-
-### 8.2
-
-Using the R function _lsfit_, compute the residuals using the data in E1,
-
-Verify that if you square and sum the residusls, you get 46.585.
-
-```{r, echo = T}
-res <- lsfit(dat$X, dat$Y)$resid
-
-sum(res^2)
-```
-
-### 8.3
-
-Verify that for the data in E1, if you use $\hat{Y} = 2X - 9$, the sum of the squared residuals is greater than 46.584.
-
-```{r, echo = T}
-Yhat <- 2*dat$X - 9
-res <- sum( (dat$Y - Yhat)^2 )
-stopifnot(res > 46.583)
-res
-
-```
-
-Why would you expect a value greater than 46.584?
-
-_The x coefficent increased._
-
-### 8.4
-
-Suppose that based on $n = 25$ values, $s^2_x = 12$ and $\sum(X_i - \bar{X})(Y_i - \bar{Y}) = 144$.
-
-What is the slope of the least squares regression?
-
-$A = 144, C = (n - 1)s^2_x = 288, b_1 = A / C = 144/288 = .5$
-
-### 8.5
-
-The following table reports breast cancer rates plus levels of solar radiation (in calories per day) for various cities in the United States. The data are stored in the file cancer_rate_dat.txt.
-
-```{r, echo =T}
-
-dat <- data.table::fread(paste0(data.dir, "cancer_rate_dat.txt"), fill = T, sep = "&")
-
-dat
-```
-
-Fit a OLS regression to predict cancer rates and comment on what this suggests.
-
-```{r, echo = T}
-fit <- lm(Rate ~ calories, data = dat)
-
-plot(fit)
-```
-
-### 8.6
-
-For the following data, use R to compute the least squares regresion line for predicting GPA given SAT.
-
-SAT: 500, 530, 590, 660, 610, 700, 570, 640
-GPA: 2.3, 3.1, 2.6, 3.0, 2.4, 3.3, 2.6, 3.5
-
-```{r, echo = T}
-dat <- data.table(
-   SAT = c(500, 530, 590, 660, 610, 700, 570, 640),
-   GPA = c(2.3, 3.1, 2.6, 3.0, 2.4, 3.3, 2.6, 3.5)
-)
-
-fit <- lm(GPA ~ SAT, data = dat)
-coef(fit)
-```
-
-### 8.7
-
-Compute the residuals for the data used in the previous problem and verify that sum to zero.
-
-```{r, echo = T}
-round(sum((dat$GPA - fit$fitted.values)), 5)
-```
-
-### 8.8
-
-For the following data, use R to compute the least squares regression line for predicting Y from X.
-
-X: 40, 41, 42, 43, 44, 45, 46
-Y: 1.62, 1.63, 1.90, 2.64, 2.05, 2.13, 1.94
-
-```{r, echo = T}
-dat <- data.table(
-   X = c(40, 41, 42, 43, 44, 45, 46),
-   Y = c(1.62, 1.63, 1.90, 2.64, 2.05, 2.13, 1.94)
-)
-
-summary(fit <- lm(Y ~ X, data = dat))
-```
-
-```{r, echo = T, fig.width=8, fig.height=3.5}
-ggplot(dat, aes(X, Y)) +
-   geom_point() +
-   geom_smooth(method = "lm")
-```
-
-### 8.9
-
-In exercise 5, what would be the least squares estimate of the cancer rate given a solar radiation of 600?
-
-```{r, echo = T, fig.width=8, fig.height=3.5}
-dat <- data.table::fread(paste0(data.dir, "cancer_rate_dat.txt"), fill = T, sep = "&")
-
-ggplot(dat, aes(calories, Rate)) +
-   geom_point() +
-   geom_smooth(method = "lm")
-
-fit <- lm(Rate ~ calories, data = dat)
-
-coef(fit)
-
-39.99 -0.037*600
-```
-Why might this be unreasonable?
-
-_Because 600 is outside of the bounds of seen values (extrapolation)._
-
-### 8.10
-
-Maximal oxygen uptake (MOU) is a measure of an individual's physical fitness. You want to know how MOU is related to how fast someone can run a mile. Suppose you randomly sample six athletes and get:
-
-MOU (kl/kg): 63.3, 60.1, 53.6, 58.8, 67.5, 62.5
-Time (seconds): 241.5, 249.8, 246.1, 232.4, 237.2, 238.4
-
-Compute the least squres regression line and comment on what the results suggest.
-
-```{r, echo = T, fig.width=8, fig.height=3.5}
-dat <- data.table(
-   MOU = c(63.3, 60.1, 53.6, 58.8, 67.5, 62.5),
-   Time = c(241.5, 249.8, 246.1, 232.4, 237.2, 238.4)
-)
-
-fit <- lm(Time ~ MOU, data = dat)
-
-ggplot(dat, aes(MOU, Time)) +
-   geom_point() +
-   geom_smooth(method = "lm") +
-   labs(title = "Time vs. MOU")
-
-```
-
-Generally, time decreases as MOU increases.
-
-### 8.11
-
-Verify that for the following pairs of points, the least squares regression line has a slope of zero. Plot the points and comment on the assumption that the regression line is straight.
-
-X: 1, 2, 3, 4, 5, 6
-Y: 1, 4, 7, 7, 4, 1
-
-```{r, echo = T, fig.width=8, fig.height=3.5}
-dat <- data.table(
-   X = c(1, 2, 3, 4, 5, 6),
-   Y = c(1, 4, 7, 7, 4, 1)
-)
-
-fit <- lm(Y ~ X, data = dat)
-
-coef(fit)
-```
-
-### 8.12
-
-Repeat the last exercise, only for points:
-
-X: 1, 2, 3, 4, 5, 6
-Y: 4, 5, 6, 7, 8, 2
-
-```{r, echo = T, fig.width=8, fig.height=3.5}
-dat <- data.table(
-   X = c(1, 2, 3, 4, 5, 6),
-   Y = c(4, 5, 6, 7, 8, 2)
-)
-
-fit <- lm(Y ~ X, data = dat)
-
-coef(fit)
-```
-
-### 8.13
-
-Vitamin A is required for good health. However, one bite of polar bear liver results in death becasue it contains a high concentration of vitamin A.
-
-Comment on this fact in terms of extrapolation.
-
-_Vitaim A as a function of health is a bounded variable. Predicting a response when the independent variables are far from what has been observed can lead to undefined results (death)._
-
-### 8.14
-
-Socket et al.
-(1987) report data related to patterns of residual insulin secretion in children.
-A portion of the study was concerned with whether age can be used to predict the logarithm of C-peptide concentrations at diagnosis.
-The observed values are (data file):
-
-Replicate the _LOESS_ smoothed curve in fig 8.4.1.
-
-```{r, echo = T, fig.width=8, fig.height=3.5}
-dat <- data.table::fread(paste0(data.dir, "diabetes_sockett_dat.txt"), fill = T)
-
-ggplot(dat, aes(age, pep)) +
-   geom_point() +
-   geom_smooth(method = "loess")
-
-```
-
-### 8.15
-
-### 8.16
-### 8.17
-### 8.18
-### 8.19
-### 8.20
-
-### 8.21
-### 8.22
-### 8.23
-### 8.24
-### 8.25
-### 8.26
-### 8.27
-### 8.28
-### 8.29
-### 8.30
-
-### 8.31
-### 8.32
-### 8.33
-### 8.34
-### 8.35
-### 8.36
+### 9.1

### 2019-12-29 Statistical Methods/09_Comparing_Two_Independent_Groups.rmd
statistical methods
statistical methods

@@
 
-#### Chapter 9
+### Chapter 9
 
-### 9.1
+#### 9.1
+
+Suppose that the sample means and variances are $\bar{X}_1 = 15, \bar{X}_2 = 12, s^2_1 = 8, s^2_2 = 24, n_1 = 20, n_2 = 10$
+
+Verify that $s^2_p = 13.14, T = 2.14$ and that Student's T test rejects the hypothesis of equal means with $\alpha = 0.05$.
+
+```{r, echo = T}
+xbar1 <- 15; xbar2 <- 12; s1sq <- 8; s2sq <- 24
+n1 <- 20; n2 <- 10
+alpha <- 0.05
+
+variance <- ( (n1 - 1)*s1sq + (n2 - 1)*s2sq ) / ( n1 + n2 - 2 )
+variance
+
+t.stat <- (xbar1 - xbar2) / sqrt( variance*(1/n1 + 1/n2) )
+t.stat
+
+crit <- qt(alpha/2, df = n1 + n2 - 2)
+
+ifelse(abs(t.stat) >= crit, "Reject Null", "Fail to Reject")
+```
+
+#### 9.2
+
+For two independent groups of subjects, $\bar{X}_1 = 45, \bar{X}_2 = 36, s^2_1 = 4, s^2_2 = 16, n_1 = 20, n_2 = 30$.
+Assume that the population variances of the two groups are equal and verify that the estimate of this common variance is 11.25.
+
+```{r, echo = T}
+xbar1 <- 45; xbar2 <- 36; s1sq <- 4; s2sq <- 16
+n1 <- 20; n2 <- 30
+
+variance <- ( (n1 - 1)*s1sq + (n2 - 1)*s2sq ) / ( n1 + n2 - 2)
+
+stopifnot(variance == 11.25)
+```
+
+#### 9.3
+
+Still assuming equal variances, test the hypothesis of equal means using Student's T test and the data in the last exercise. Use $\alpha = 0.05$
+
+```{r, echo = T}
+alpha <- 0.05
+
+t.stat <- (xbar1 - xbar2 ) / sqrt( variance * (1/n1 + 1/n2) )
+
+crit <- qt(alpha/2, df = (n1 + n2 - 2))
+
+ifelse(abs(t.stat) >= crit, "Reject null", "Fail to Reject")
+```
+
+#### 9.4
+
+Repeat the last problem, only use Welch's test for comparing means.
+
+```{r, echo = T}
+w.stat <- (xbar1 - xbar2) / sqrt(s1sq/n1 + s2sq/n2)
+
+q1 <- s1sq/n1; q2 <- s2sq/n2
+df <- (q1 + q2)^2 / ( q1^2/(n1 - 1) + q2^2/(n2 - 1) )
+
+crit <- qt(1 - alpha/2, df)
+
+ifelse(abs(w.stat) >= crit, "Reject Null", "Fail to Reject")
+```
+
+#### 9.5
+
+Comparing the results for the last two problems, what do they suggest regarding the power of Welch's test compared to STudent's T test when the sample variances differ sufficiently.
+
+```{r, echo = T}
+t.stat
+w.stat
+```
+
+_Roughtly, the W statistic has more power because it is larger than the T statistic._
+
+#### 9.6
+
+For two independent groups $\bar{X}_1 = 86, \bar{X}_2 = 80, s^2_1 = s^2_2 = 25, n_1 = n_2 = 20$.
+Assume that the population variances of the two groups are equal and verify that Student's T rejects with $\alpha = 0.01$.
+
+```{r, echo = T}
+xbar1 <- 86; xbar2 <- 80; s1sq <- s2sq <- 25
+n1 <- n2 <- 20; alpha <- 0.01
+
+t.stat <- (xbar1 - xbar2) / sqrt(2 * s1sq/n1 )
+
+crit <- qt(alpha/2, df = n1 + n2 - 2)
+
+ifelse(abs(t.stat) >= crit, "Reject", "Fail to Reject")
+```
+
+#### 9.7
+
+Repeat the last exercise using Welch's method.
+
+```{r, echo = T}
+
+w.stat <- (xbar1 - xbar2) / sqrt( 2 * s1sq/n1 )
+
+q1 <- s1sq/n1; q2 <- s2sq/n2
+
+df <- (q1 + q2)^2 / ( q1^2/(n1 - 1) + q2^2/(n2 - 1))
+
+crit <- qt(alpha/2, df = 38)
+
+ifelse(abs(w.stat) >= crit, "Reject", "Fail to Reject")
+```
+
+#### 9.8
+
+Comparing the results of the last two problems, what do they suggest about using Student's T versus Welch's method when the sample variances are approximately equal?
+
+```{r, echo = T}
+t.stat
+w.stat
+```
+
+_With equal sample variances, Student's T and Welch give exactly the same result, suggesting that when the sample variances are approximately equal, the choice between T and W makes little difference._
+
+#### 9.9
+
+For $\bar{X}_1 = 10, \bar{X}_2 = 5, s^2_1 = 21, s^2_2 = 29, n_1 = n_2 = 16$, compute a 0.95 conffidence interval for the difference between the means using Welch's method and state whether a decision can be made about which group has the larger popluation mean.
+
+```{r, echo = T}
+xbar1 <- 10; xbar2 <- 5; s1sq <- 21; s2sq <- 29
+n1 <- n2 <- 16; alpha <- 0.05
+
+q1 <- s1sq/n1; q2 <- s2sq/n2
+df <- (q1 + q2)^2 / (q1^2 / (n1 - 1) + q2^2 / (n2 -1) )
+
+variance <- sqrt(s1sq/n1 + s2sq/n2)
+
+(xbar1 - xbar2) + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), df)*variance
+```
+
+#### 9.10
+
+Repeat the last problem, only use Student's T instead.
+
+```{r, echo = T}
+sqp <- (s1sq+s2sq) / 2
+
+(xbar1 - xbar2) + qt(c(Lower = alpha/2, Upper = 1 - alpha/2), n1 + n2 - 2) * sqrt(sqp * (1/n1 + 1/n2))
+```
+
+#### 9.11
+
+Two methods for training accountants are to be compared.
+Students are randomly assigned to one of the two methods.
+At the end of the course, each student is asked to prepare a tax return for the same individual.
+The amounts of the refund reported by the students are:
+
+M1: 132, 204, 603, 50, 125, 90, 185, 134
+M2: 92, -42, 121, 63, 182, 101, 294, 36
+
+Using Welch's test would you conclude that the methods differ in terms of the average return? Use $\alpha = 0.05$.
+
+```{r, echo = T}
+
+```
+
+#### 9.12
+
+Responses to stress are govered by the hypothalamus.
+Imagine you have two randomly sampled groups of individuals between the ages of 60 and 65.
+The first shows signs of heart disease and the other does not.
+You want to determine whether the groups differ in terms of some measure associated with the hypothalamus.
+For the first group of subjects with no heart disease, the measures are:
+
+11.1, 12.2, 15.5, 17.6, 13.0, 7.5, 9.1, 6.6, 9.5, 18.0, 12.6
+
+For the other group with heart disease, the kmeasures are:
+
+18.2, 14.1, 13.8, 12.1, 34.1, 12.0, 14.1, 14.5, 12.6, 12.5, 19.8, 13.4, 16.8, 14.1, 12.9
+
+Determine whether the groups differ based on Welch's test. Use $\alpha = 0.05$.
+
+```{r, echo = T}
+
+```
+
+#### 9.13
+
+The 0.95 confidence interval for the difference between the means, using Student's T, is (2.2, 20.5).
+
+What are the practical concerns with this confidence interval?
+
+### 10.14
+
+For the first of two binomial distributions, there are 15 successes amoung 24 observations. For the second, there are 23 successes among 42 observations.
+
+Test $H_0 : p_1 = p_2$ with a Type I error probability of 0.05 using the Storer-Kim method, followed by Beal's method.
+
+Note the p-values and comment on what this suggests in terms of power.

### 2019-12-30 Statistical Methods/09_Comparing_Two_Independent_Groups.rmd
statistical methods
statistical methods

@@
 ```{r, echo = T}
+dat <- data.table(
+   M1 = c(132, 204, 603, 50, 125, 90, 185, 134),
+   M2 = c(92, -42, 121, 63, 182, 101, 294, 36))
 
+t.test(dat$M1, dat$M2, conf.level = .95)
 ```
+
+The p-value is 0.258, so no, fail to reject and make no decision about which group has the larger population mean.
 
@@
 ```{r, echo = T}
+dat <- data.table(
+   M1 = c(11.1, 12.2, 15.5, 17.6, 13.0, 7.5, 9.1, 6.6, 9.5, 18.0, 12.6),
+   M2 = c(18.2, 14.1, 13.8, 12.1, 34.1, 12.0, 14.1, 14.5, 12.6, 12.5, 19.8, 13.4, 16.8, 14.1, 12.9))
 
+t.test(dat$M1, dat$M2, conf.level = .95)
 ```
+
+_Fail to reject because the p-value is greater than 0.05 and make no decision about which group has the larger population mean._
 
@@
 
-### 10.14
+_The actual probability coverage could differ substantially from 0.95._
+
+#### 9.14
 
@@
 Note the p-values and comment on what this suggests in terms of power.
+
+```{r, echo = T}
+# Storer-Kim
+twobinom(15, 24, 23, 42)
+# Beal's
+twobicipv(15, 24, 23, 42)
+```
+
+_In this particular case, the p-value returned by Beal's method is substantially larger than the p-value returned by the Storer Kim method, suggesting that generally the Storer Kim method will have more power._
+
+#### 9.15
+
+A film producer wants to know which of two versions of a particular scene is more likely to be viewed as disturbing.
+A group of 98 individuals views the first version and 20 say that it is disturbing.
+Another group sees the second version and 30 of 70 people say that it is disturbing.
+
+What is the probability these two probabilities are equal?
+
+```{r, echo = T}
+# Storer-Kim
+twobinom(20, 98, 30, 70)
+
+# Beal's
+twobicipv(20, 98, 30, 70)
+```
+
+Both reject, again Storer Kim method has more power.
+
+#### 9.16
+
+It is found that amoung 121 individuals who take a training program on investing in commodities, 20 make money during the next year and the rest do not.
+With another training program, 15 of 80 make money.
+
+Test the hypothesis that the probability of making money is the same for both training programs with a Type I error probability of 0.05.
+
+```{r, echo = T}
+p1 <- 20/121; p2 <- 15/80
+Z <- (p1-p2)/sqrt(p1*(1-p1)/121 + p2*(1-p2)/80)
+crit <- qnorm(0.05/2)
+ifelse(abs(Z) >= crit, "Reject Null", "Fail To Reject")
+```
+
+#### 9.17
+
+In a study dealing with violence between factions in the Middle East, one goal was to compare measures of depression for two groups of young males.
+In the first group, no family member was wounded or killed by someone belonging to the opposite faction, and the measures were:
+
++ 22, 23, 12, 11, 30, 22, 7, 42, 24, 33, 28, 19, 4, 34, 15, 26, 50, 27, 20, 30, 14, 42
+
+The second group considted of young males who had a family member killed or wounded. The observed values were:
+
++ 17, 22, 16, 16, 14, 29, 20, 20, 19, 14, 10, 8, 26, 9, 14, 17, 21, 16, 14, 11, 14, 11, 29, 13, 4, 16, 16, 7, 21
+
+Test the hypothesis of equal means with Student's T test and $\alpha = 0.05$, and compute a .95 confidence interval.
+
+```{r, echo = T}
+dat <- data.table(
+   X1 = c(22, 23, 12, 11, 30, 22, 7, 42, 24, 33, 28, 19, 4, 34, 15, 26, 50, 27, 20, 30, 14, 42),
+   X2 = c(17, 22, 16, 16, 14, 29, 20, 20, 19, 14, 10, 8, 26, 9, 14, 17, 21, 16, 14, 11, 14, 11, 29, 13, 4, 16, 16, 7, 21)
+)
+
+t.test(dat$X1, dat$X2)
+t.test(dat$X1, dat$X2, var.equal = T)
+```
+
+_Student's T yields a shorter confidence interval, but the confidence interval based on Student's T can be less accurate compared to the confidence interval based on Welch's method._
+
+#### 9.18
+
+For the data in the last problem, the difference between the medians is -7.5 and the corresponding McKean-Schrader estimate of the standard error is $\sqrt(S^2_1 + S^2_2) = 3.93$.
+
+Verify that you do not reject the hypothesis of equal medians with $\alpha = 0.05$
+
+```{r, echo = T}
+msmed(dat$X1, dat$X2)
+```
+
+#### 9.19
+
+Using the data in Excercise 17, use a percentile bootstrap method to compare the medians.
+Note that the p-value is 0.01.
+This is in contrast to the result in Exercise 18, where the hypothesis of equal medians is not rejected at the 0.05 level.
+
+```{r, echo = T}
+medpb2(dat$X1, dat$X2)
+```
+
+_The p-value indicates that the hypothesis of equal medians would be rejected at the 0.01 level, in contrast to the method based on theMcKean-Schrader estimator, which has a p-value equal to 0.057._
+
+#### 9.20
+
+Repeat the analysis of the sexual attitude data, the second example in Section 9.10.2, only now use _twobinom_ and _twobicipv_.
+
+```{r, echo = T}
+twobinom(49,105,101,156)
+
+twobicipv(49,105,101,156)
+```
+
+#### 9.21
+
+Does the consumption of alcohol limit our attention span?
+An article in the July 10, 2006 Los Angeles Times described a study conducted at the University of Washington where 23 people were given enough alcohol to reach a BAC of 0.04%.
+A second group of 23 people drank no alcohol.
+The researchers then showed amembers of both groups a 25-second video clip in which two teams passed a ball back and forth and asked them to count the number of times one team passed the ball.
+During the clip, a person in a gorilla suit walks through the crowd, thumps its chest, and walks off.
+
+Researchers found that 11 of the 23 participants in the control group saw the gorilla, compared to 10 in the alcohol group.
+
+Verify that you do not reject the null hypothesis that the two groups have the same probability of seeing the gorilla using the Storer-Kim method with $\alpha = 0.05$.
+
+```{r, echo = T}
+twobinom(11,23,10,23)
+```
+
+#### 9.22
+
+Despite any problems it might have, summarize how you would justify using Student's T test to compare two indepentent groups.
+
+_Student's T controls the Type I error probability when comparing identical distributions.
+It is possible that Student's T accurately detects a difference in the distributions that might be missed by other methods._
+
+#### 9.23
+
+Summarize any practical conccerns about Student's T test and comment on how they might be addressed.
+
+_In terms of understanding the nature of any difference that might exist, Student's T can be highly inaccurate in some situations.
+Confidence intervals for the difference between the means can be inaccurate when distributions differ.
+Problems with heteroscedasticity can be reduced by using a method that allows heteroscedasticity, such as Welch's test or a bootstrap-t method.
+But inaccurate confidence intervals, as well as relatively low power due to outliers, are always a concern when using means.
+Other methods might have substantially more power, such as techniques based on the median or 20% trimmed mean.
+Even with discrete distributions where outliers are rare, the methods in Section 9.11 might have more power.
+The only safe way to determine whether alternative methods make a difference is to try them._
+
+#### 9.24
+
+Summarize the relative merits of comparing groups with medians.
+
+_The median better reflects the typical value for skewed distributions, it can have relatively high power when outliers are common.
+Accurate confidence intervals can be computed over a fairly broad range of situations, much broader than methods based on means.
+But for skewed distributions, it is possible that means have more power, even when the Type I error probability is controlled reasonably well, because the difference between the means might be larger.
+As noted in Chapter 2, there are situations where the mean can be argued to be a better measure of location.
+Also, when dealing with distributions that have light tails (outliers are relatively uncommon), the mean or some smaller amount of trimming can have a smaller standard error._
+
+#### 9.25
+
+For each of two independent groups, 1,000 bootstrap stamples are generated and it is found that there are 10 instances where the bootstrap sample trimmed mean for the first group is less than the bootstrap sample trimmed mean for the second.
+There are two instances where the bootstrap sample trimmed means are equal.
+Determine a p-value when testing the hypothesis of equal population trimmed means.
+
+B = 1000, A = 10, C = 2,
+Q = (A + 0:5C)=B = 0:01 + 0:5(0:002) = 0:011. Set P equal to Q or 1 Q, whichever is smaller, 
+so P = 0:011 and the p-value is 2P = 2(0:011) = 0:022
+
+#### 9.26
+
+Repeat the analysis of Salk's data in section 9.1.6 using Welch's method.
+
+```{r, echo = T}
+salk <- data.table::fread(paste0(data.dir, "salk_dat.txt"), fill = T)
+
+t.test(salk[,1],salk[,2])
+```
+
+#### 9.27
+
+For the self-awareness data, apply Yuen's method with 20% trimming.
+
+```{r, echo = T}
+dana <- data.table::fread(paste0(data.dir, "dana_dat.txt"), fill = T)
+
+yuen(dana[,1],dana[,2])
+```
+
+#### 9.28
+
+Duplicate the results for shoulder pain.
+
+```{r, echo = T}
+g1=c(2, 4, 4, 2, 2, 2, 4, 3, 2, 4, 2, 3, 2, 4, 3, 2, 2, 3, 5, 5, 2, 2)
+g2=c(5, 1, 4, 4, 2, 3, 3, 1, 1, 1, 1, 2, 2, 1, 1, 5, 3, 5)
+disc2com(g1,g2)
+```

### 2020-01-17 Statistical Methods/10_Comparing_Independent_Groups.rmd
statistical methods, ch9/10
statistical methods, ch9/10
statistical methods
statistical methods
statistical methods, ch10
statistical methods, ch10

--- 2019-12-27 Statistical Methods/09_Comparing_Two_Independent_Groups.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Comparing Two Independent Groups}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Comparing N Independent Groups}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 9}
+   \lfoot{\color{headergrey}Chapter 10}
    \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
@@
 
-#### Chapter 9
+### Chapter 10
 
-### 9.1
+#### 10.1
+
+Consider the following data:
+
+Group 1: 3, 5, 2, 4, 8, 4, 3, 9
+
+Group 2: 4, 4, 3, 8, 7, 4, 2, 5
+
+Group 3: 6, 7, 8, 6, 7, 9, 10, 9
+
+```{r}
+g1 <- c(3, 5, 2, 4, 8, 4, 3, 9)
+g2 <- c(4, 4, 3, 8, 7, 4, 2, 5)
+g3 <- c(6, 7, 8, 6, 7, 9, 10, 9)
+
+data <- data.table(g1, g2, g3)
+
+means <- colMeans(data)
+sds <- apply(data, 2, sd) ** 2
+```
+
+Assume the three groups have a common population variance, $\sigma^2_p$. Estimate $\sigma^2_p$.
+
+```{r}
+# mean squares between groups:
+sum(sds)/ncol(data)
+```
+
+#### 10.2
+
+For the previous data, use R to test the hypothesis of equal means using the ANOVA F test with $\alpha = 0.05$.
+
+```{r}
+anova1(data)
+```
+
+#### 10.3
+
+For the previous data, verify that Welch's test statistic is $F_w = 7.7$ with degrees of freedom $v_1 = 2$, $v_2 = 13.4$
+
+```{r}
+test <- t1way(data, tr = 0)
+
+assertthat::are_equal(round(test$TEST, 2), 7.77)
+assertthat::are_equal(test$nu1, 2)
+assertthat::are_equal(round(test$nu2, 2), 13.4)
+
+```
+
+#### 10.4
+#### 10.5
+#### 10.6
+#### 10.7
+#### 10.8
+#### 10.9
+#### 10.10
+#### 10.11
+#### 10.12
+#### 10.13
+#### 10.15
+#### 10.16
+#### 10.17
+#### 10.18
+#### 10.19
+#### 10.20
+#### 10.21
+#### 10.22
+#### 10.23
+#### 10.24
+#### 10.25
+#### 10.26
+#### 10.27
+#### 10.28
+#### 10.29
+#### 10.30

### 2020-01-17 Statistical Methods/10_Comparing_Independent_Groups.rmd
statistical methods, group comparisons
statistical methods, group comparisons

@@
 
-# DO NOT ADD OR REVISE CODE HERE
-knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
+knitr::opts_chunk$set(
+   echo = T, 
+   eval = TRUE, 
+   dev = 'png', 
+   fig.width = 9, 
+   fig.height = 3.5)
+
 options(knitr.table.format = "latex")
@@
 ```{r}
-anova1(data)
+anova <- anova1(data)
+
+anova
+
+ifelse(anova$p.value < 0.05, "Reject", "Fail to Reject")
+
 ```
@@
 #### 10.4
+
+Using R, test the hypothesis of equal means using the ANOVA F test for the following data:
+
+Group 1: 15, 17, 22
+
+Group 2: 9, 12, 15
+
+Group 3: 17, 20, 23
+
+Group 4: 13, 12, 17
+
+```{r}
+g1 <- c(15, 17, 22)
+g2 <- c(9, 12, 15)
+g3 <- c(17, 20, 23)
+g4 <- c(13, 12, 17)
+
+data <- data.table(g1, g2, g3, g4)
+
+anova <- anova1(data)
+
+ggplot(stack(data)) +
+   geom_boxplot(aes(ind, values, fill = ind))
+
+anova
+
+ifelse(anova$p.value < 0.05, "Reject Null", "Fail to Reject")
+```
+
 #### 10.5
+
+Using the previous data, compute the common assumed variance.
+
+```{r}
+mean( apply(data, 2, var) )
+```
+
 #### 10.6
+
+Use R to verify the p-value for the previous data based on Welch's test is 0.124.
+
+```{r}
+w <- t1way(data)
+
+assertthat::are_equal(round(w$p.value, 3), 0.124)
+```
+
 #### 10.7
+
+For the data in E4, explain why you would get the same result using a 20% trimmed mean.
+
+_Sample size would not reduce with 3 values_
+
 #### 10.8
+
+Why would you not recommend the strategy of testing for equal variances, and if not significant, using kthe ANOVA F test rather than the Welch's method?
+
+_Don't know when power is high enough._
+
 #### 10.9
+
+Five independent groups are compared with n = 15 observations for each group, SSBG = 20 and SSWG = 150.
+
+Perform the ANOVA F test with $\alpha = 0.05$
+
+```{r}
+n <- 15; j <- 5
+SSBG <- 20; SSWG <- 150
+
+MSBG <- SSBG / (j -1); MSWG <- SSWG / (n * j - j)
+
+f.stat <- MSBG / MSWG
+
+v1 <- j - 1; v2 <- j *n - j
+crit.val <- qf(1 - 0.05, v1, v2)
+
+ifelse(f.stat > crit.val, "Reject", "Fail to Reject")
+
+vals <- data.table(x = seq(0, 10, by = .001))
+vals[, y := df(vals$x, v1, v2)]
+
+ggplot(vals, aes(x, y)) +
+   geom_point(size = .7) +
+   geom_vline(xintercept = crit.val, col = "darkred") +
+   geom_vline(xintercept = f.stat, col = "cornflowerblue")
+
+```
+
 #### 10.10
+
+Use the R function anova1 to verify that for the following data, MSGB = 14.4 and MSGW = 12.59.
+
+G1: 9, 10, 15
+
+G2: 16, 8, 13, 6
+
+G3: 7, 6, 9
+
+```{r}
+g1 <- c(9, 10, 15)
+g2 <- c(16, 8, 13, 6)
+g3 <- c(7, 6, 9)
+
+data <- list(g1, g2, g3)
+
+anova1(data)
+```
+
 #### 10.11
+
+Consider five groups (J = 5) with population means 3, 4, 5, 6 and 7, and a common variance $\sigma^2_p = 2$.
+If the number of observations in each group is 10 (n = 10), indicate what is being estimated by MSBG, and based on the information given, determine its value.
+That is, if the population means and common variance were known, what is the value being estimated by MSBG?
+
+MSBG estimates $\sigma^2_p + n\sigma^2_\mu$, the variation among the population means.
+
+```{r}
+var(c(3, 4, 5, 6, 7))
+```
+
+How does this differ from the value estimated by MSBG?
+
 #### 10.12

### 2020-01-19 Statistical Methods/10_Comparing_Independent_Groups.rmd
statistical methods, ch10
statistical methods, ch10

@@
 
-How does this differ from the value estimated by MSBG?
+#### 10.12
 
-#### 10.12
+For the following data, verify that ANOVA F test does not reject with $\alpha = 0.05$, but Welch's test _does reject_.
+
+Group 1: 10, 11, 12, 9, 8, 7
+
+Group 2: 10, 66, 15, 32, 22, 51
+
+Group 3: 1, 12, 42, 31, 55, 19
+
+```{r}
+g1 <- c(10, 11, 12, 9, 8, 7)
+g2 <- c(10, 66, 15, 32, 22, 51)
+g3 <- c(1, 12, 42, 31, 55, 19)
+
+data <- list(g1,g2,g3)
+
+anova1(data) # p-value 0.082, fail to reject
+
+t1way(data, tr=0) # p-value 0.046, reject
+```
+
 #### 10.13
+
+Using the ANOVA F test, you see that $v_1 = 5$, and $v_2 = 8$. Verify that the number of groups is $J = 4$ and the total number of observations is $N = 12$.
+
+$v_1$ = J - 1, J = 4
+
+$v_2$ = N - J = 13
+
+#### 10.14
+
+A researacher reports a p-value of 0.001 with the ANOVA F test. Describe what conclusions are reasonable based upon this result.
+
+_The distributions differ, suggesting that in particular the means differ._
+
 #### 10.15
+
+Summarize the reasons you might fail to reject with the ANOVA F test.
+
+_Low power due to outliers, violating the equal variance assumption_
+
 #### 10.16
+
+Someone tests for equal variances and fails to reject. Does this justify the use of the ANOVA F test?
+
+_No, this depends on power._
+
 #### 10.17
+
+A researcher reports that a test for normality was performed and that based on this test, no evidence of nonnormality was found. Why might it be unreasonable to assume normality despite this result?
+
+_Unclear whether the test has enough power to detect a departure from normality._
+
 #### 10.18
+
+Outline how you might construct an example where sampling is from normal distributions, Welch's test rejects, but the ANOVA F test does not.
+
+_Generate data so that groups have unequal variances._
+
 #### 10.19
+
+Consider a 2-by-2 design with population means:
+
+$\mu_1 = 110, \mu_2 = 70$
+
+$\mu_3 = 80, \mu_4 = 40$
+
+```{r}
+vals <- matrix(c(110, 70, 80, 40), nrow = 2, byrow = F)
+
+apply(vals, 2, sum) # main effect
+
+apply(vals, 1, sum) # interaction effect
+```
+
 #### 10.20
+
+Consider a 2-by-2 design with population means.
+
+$\mu_1 = 10, \mu_2 = 20$
+
+$\mu_3 = 40, \mu_4 = 10$
+
+```{r}
+vals <- matrix(c(10, 20, 40, 10), nrow = 2, byrow = F)
+
+apply(vals, 2, sum) # main effect
+
+apply(vals, 1, sum) # interaction effect
+```
+
 #### 10.21
+
+For the data in the last example of 10.1, use med1way to test hyhpothesis that population medians are equal.
+
+```{r}
+x=list()
+x[[1]]=c( 7, 9, 8, 12, 8, 7, 4, 10, 9, 6)
+x[[2]]=c(10, 13, 9, 11, 5, 9, 8, 10, 8, 7)
+x[[3]]=c(12, 11, 15, 7, 14, 10, 12, 12, 13, 14)
+
+med1way(x)
+```
+
+p < 0.05, reject
+
 #### 10.22
-#### 10.23
-#### 10.24
-#### 10.25
-#### 10.26
-#### 10.27
-#### 10.28
-#### 10.29
-#### 10.30
+
+Repeat previous with Qanvoa:
+
+```{r}
+Qanova(x)
+```
+
+p < 0.01, reject

### 2020-01-23 Statistical Methods/12_Multiple_Comparisons.rmd
statistical methods, ch11
statistical methods, ch11
bootstrap work
bootstrap work

--- 2019-12-27 Statistical Methods/09_Comparing_Two_Independent_Groups.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Comparing Two Independent Groups}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Multiple Comparisons}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 9}
+   \lfoot{\color{headergrey}Chapter 12}
    \fancyfoot[C]{\rmfamily\color{headergrey}Statistical Methods}
@@
 library(grid, quietly = TRUE, warn.conflicts = FALSE)
-library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
+library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)+
 library(WRS, quietly = TRUE, warn.conflicts = FALSE)
@@
 
-#### Chapter 9
+### Chapter 12
 
-### 9.1
+#### 12.1
+#### 12.2
+#### 12.3
+#### 12.4
+#### 12.5
+#### 12.6
+#### 12.7
+#### 12.8
+#### 12.9
+#### 12.10
+#### 12.11
+#### 12.12
+#### 12.13
+#### 12.14
+#### 12.15
+#### 12.16
+#### 12.17
+#### 12.18
+#### 12.19
+#### 12.20
+#### 12.21
+#### 12.22
+#### 12.23
+#### 12.24
+#### 12.25
+#### 12.26
+#### 12.27
+#### 12.28
+#### 12.29
+#### 12.30

### 2020-01-23 Mathematical Statistics/05_conf_int_bootstrap.r

@@ -1,7 +1,31 @@
 library(data.table)
 library(ggplot2)
+library(here)
+library(boot) # alternate to verify long-hand calcs
 
-data.dir <- "D:/Projects/Statistical-Computing/datasets/"
+data.dir <- paste0(here::here(), "/datasets/")
+
+# gamma example
+
+my.sample <- rgamma(16, 1, 1/2)
+N <- 10e5
+
+my.boot <- numeric(N)
+
+for(i in 1:N)
+{
+  x <- sample(x = my.sample, size = length(my.sample), replace = T) 
+  my.boot[i] <- mean(x)
+}
+
+ggplot(data.table(value = my.sample), aes(value, fill = ..count..)) +
+  geom_histogram(bins = 8)
+
+ggplot(data.table(value = my.boot), aes(value, fill = ..count..)) +
+  geom_histogram(bins = 30)
+
+mean(my.boot)
+sd(my.boot)
 
 ### Bangladesh
 
@@ -9,6 +33,8 @@ Bangladesh <- data.table(read.csv(paste0(data.dir, "Bangladesh.csv"),
                                  header = T))
 
 Arsenic <- Bangladesh$Arsenic
+
+mean(Arsenic) #  sample mean
 
 ggplot(data.table(value = Arsenic)) +
   geom_histogram(aes(value, fill = ..count..), bins = 30) +
@@ -46,13 +72,59 @@ mean(arsenic.mean) # Bootstrap mean
 mean(arsenic.mean) - mean(Arsenic) # bias
 sd(arsenic.mean) # bootstrap SE
 
-uq <- mean(arsenic.mean) - 1.96 * sd(arsenic.mean)
-lq <- mean(arsenic.mean) + 1.96 * sd(arsenic.mean)
+# this is not accurate: CLT cannot be used here
+lq <- mean(arsenic.mean) - 1.96 * sd(arsenic.mean)
+uq <- mean(arsenic.mean) + 1.96 * sd(arsenic.mean)
 
-sum(arsenic.mean > lq) / N
-sum(arsenic.mean < uq) / N
+sum(arsenic.mean < lq) / N
+sum(arsenic.mean > uq) / N
 
-quantile(arsenic.mean, c(0.025, 0.975)) # confidence intervals
+alpha <- .05
+quantile(arsenic.mean, c(alpha/2, 1 - alpha/2)) # confidence intervals
+
+boot.fn <- function(data, index) {
[…]

### 2020-01-27 Mathematical Statistics/05_conf_int_bootstrap.rmd
bootstrap work
bootstrap work

@@
 ```{r knitr_setup, include = FALSE}
+knitr::opts_chunk$set(
+   echo = T, 
+   eval = TRUE, 
+   dev = 'png', 
+   fig.width = 9, 
+   fig.height = 3.5)
 
-# DO NOT ADD OR REVISE CODE HERE
-knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
 options(knitr.table.format = "latex")
-
 ```
@@
 library(gtools, quietly = TRUE, warn.conflicts = FALSE)
+library(here, quietly = T, warn.conflicts = F)
+library(boot, quietly = T, warn.conflicts = F)
 
@@
 
-data.dir <- "D:/Projects/Statistical-Computing/datasets/"
-
-setwd("D:/Projects/Statistical-Computing/RDS")
+setwd(paste0(here::here(), "/datasets/"))
 
@@
 p <- mean( means == 1 )
-
 ```
 
-Probability: __`r round(p, 4) *100`%__
+Probability: __`r round(p, 4) * 100`%__
 
@@
 
-Probability: __`r round(p, 4) *100`%__
+Probability: __`r round(p, 4) * 100`%__
 
@@
 p <- mean( lt2 == 2 )
-
 ```
@@
 
-```{r, echo = T}
+```{r}
 samples <- combinations(n = 3, r = 3, 1:3, repeats.allowed = T)
@@
 
-choose(3^3, 3)
-
+assertthat::are_equal(n, choose(3 + 3 - 1, 3))
 ```
 
-Samples: = $3^3$ = __`r n`__
+Samples: = $5 \choose 3$ = __`r n`__
 
@@
 
-```{r, echo = T}
-
+```{r}
 samples <- permutations(n = 3, r = 3, 1:3, repeats.allowed = T)
@@
 
-```{r, echo = T}
+```{r}
 
@@
 No, 3% and 6% chances respectfully.
+
+### 5.4
+
+Consider the samples 1, 3, 3, and 5 from some distribution.
+
+```{r}
+samples <- c(1, 3, 3, 5)
+```
+
+a.) How many bootstrap samples are there?
+
+```{r}
+boot <- permutations(n = 3, r = 4, v = samples, repeats.allowed = T)
+
+n <- nrow(boot)
+```
+
+_3 unique items to pick from, 4 places to put each item._
+
+Number of permutations: *$3^4$ = `r n`*
+
+b.) List the distinct bootstrap samples assuming order does not matter.
+
+```{r}
+combinations(n = 3, r = 4, v = samples, repeats.allowed = T)
+```
+
+### 5.5
+
+We determine the number of distinct bootstrap samples from a finite set.
+
+a.) A bakery sells five types of cookies: sugar, chocolate chip, oatmeal, peanut butter, and ginger snap. Show that the number of ways to order 5 cookies is $9 \choose 5$
+
+Unordered samping with replacement: $n + k - 1 \choose k$, $n = 5, k = 5$
+
+```{r}
+choose(5 + 5 - 1, 5)
+```
+
+b.) Show that the number of sets of size n (order does not matter) drawn with replacement from the (distinct) $a_1, a_2, \ldots, a_n$ is ${2n-1}\choose n$
+
+Conclude that the number of distinct bootstrap samples from the set $[a_1, a_2, \ldots, a_n]$ is ${2n - 1} \choose  n$
+
+### 5.6
+
+Let $k_1, k_2, \ldots, k_n$ denote non-negative integers satisfying $k_1 + k_2 + \ldots + k_n = n$, and suppose the elements in the set $a_1, a_2, \ldots, a_n$ are distinct.
+
+a.) Show that the number of bootstrap samples with $k_1$ occurrences of $a_1, k_2$ occurrences of $a_2, \ldots, k_n$ occurrences of $a_n$ is $n \choose k_1, k_2, \ldots, k_n$
+
+b.) Compute the probability that a randomly drawn bootstrap sample will have $k_i$ occurrences of $a_i, i = 1, 2, \ldots, n$
+
+### 5.7
+
+Refer to Example 5.4 and the remark at the end of the example.
+
+a.) What might account for the fact that there were more missing values for the men who skateboarded in front of the male experimenter? How might this bias the outcome?
+
+_It could be that the approached skateboarders refused to participate in the study of performing tricks in front of other men._
+
+b.) Why do you suppose it was important that the two experimenters were blinded to the purpose of the study?
+
+_The female could have flurted or otherwise influenced skateboarders who were performing tricks if they knew the intent of the study._
+
+### 5.8
+
+Consider a population that has a normal distribution with mean $\mu = 36$, standard deviation $\sigma = 8$.
+
+```{r}
+mu <- 36; sigma <- 8; n <- 200
+
+se <- mu / sqrt(n)
+```
+
+a.) The sampling distribution of $\bar{X}$ for samples of size 200 will have what mean, standard error, and shape?
+
+_$\mu = 36$, SE = 36 / sqrt(200) = `r se`, shape will be approximately normal (CLT)._
+
+b.) Use R to draw a random sample of size 200 from this population. Conduct EDA on your sample.
+
+```{r}
+set.seed(123)
+
+samp <- data.table(values = rnorm(200, mean = 36, sd = 8))
+
+ggplot(samp, aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), color = "darkorange") +
+   geom_vline(xintercept = mu, col = "black", height = .8, lty = 2) +
+   geom_vline(xintercept = mu - se, col = "darkred", lty = 2) +
+   geom_vline(xintercept = mu + se, col = "darkred", lty = 2)
+
+```
+
+c.) Compute the bootstrap distribution for your sample, and note the bootstrap mean and standard error.
+
+```{r}
+boot.fn <- function(data, index){
+   mean(data[index]$values)
+}
+
+I <- 10e3
+
+boot(samp, boot.fn, R = I) # boot pkg
+
+cst.boot <- function(values, n) {
+   bootstrap <- numeric(I)
+
+   for(i in 1:I)
+   {
+      bootstrap[i] <- mean( sample(values, n, replace = T) )
+   }
+
+   alpha <- 0.05
+
+   observed <- mean(values)
+
+   boot.mean <- mean(bootstrap)
+   boot.bias <- observed - boot.mean
+   boot.se <- sd(bootstrap)
+
+   list(observed = observed,
+         mean = boot.mean,
+         bias = boot.bias,
+         se = boot.se,
+         conf = quantile(bootstrap, c(alpha/2, 1 - alpha/2)))
+}
+
+```
+
+d.) Compare the bootstrap distribution to the theoretical sampling distribution by creating a table like Table 5.2:
+
+```{r}
+n.200 <- cst.boot(samp$values, 200)
+
+tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
+                  Mean = c(mu, mu, n.200$observed, n.200$mean),
+                  SD = c(sigma, mu/sqrt(n), sd(samp$values), n.200$se))
+
+pretty_kable(tbl, "Sampling Statistics")
+```
+
+e.) Repeat for sample sizes $n = 50$ and $n = 10$. Carefully describe yyour observations about the effects of sample size on the bootstrap distribuiton.
+
+```{r}
+n.50 <- cst.boot(samp$values, 50)
+
+tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
+                  Mean = c(mu, sigma, n.50$observed, n.50$mean),
+                  SD = c(sigma, mu/sqrt(n), sd(samp$values), n.50$se))
+
+pretty_kable(tbl, "Sampling Statistics")
+
+```
+
+```{r}
+n.10 <- cst.boot(samp$values, 10)
+
+tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
+                  Mean = c(mu, sigma, n.10$observed, n.10$mean),
+                  SD = c(sigma, mu/sqrt(n), sd(samp$values), n.10$se))
+
+pretty_kable(tbl, "Sampling Statistics")
+```
+
+_The center of the bootstrap distribution doesn't vary much with smaller n, however, confidence intervals (the sd of the bootstrap dist) vary wildy._
+
+### 5.9
+
+Consider a population that has a gamma distribution with parameters r = 5, $\lambda = 4$.
+
+a.) Use simulation (with $n = 200$) to generate an approximate sampling distribution of the mean; plot and describe the distribution.
+
+b.) Now, draw one random sample of size 200 from this population. Create a histogram of your sample, and find the mean and standard deviation.
+
+c.) Compute the bootstrap distribution of the mean for you sample, plot it, and note the bootstrap mean and standard error.
+
+d.) Compare the bootstrap distribution to the approximate theoretical sampling distribution by creating a table like Table 5.2.
+
+e.) Repeat (a-e) for sample sizes of n = 50, and n = 10. Describe carefully your observations about the effects of sample size on the bootstrap distribution.
+
+### 5.10
+
+We investigate the bootstrap distribution of the median. Create random sample of size n for various n and bootstrap the median. Describe the bootstrap distribution.
+
+```{r}
+
+```

### 2020-01-27 Mathematical Statistics/05_conf_int_bootstrap.r

@@ -265,12 +265,15 @@ for(i in 1:N)
   result[i] <- mean(Verizon[index]$Time) / mean(Verizon[-index]$Time)
 }
 
-mean(time.ratio.mean)
-sd(time.ratio.mean)
+observed - mean(time.ratio.mean) # bias
+sd(time.ratio.mean) # standard error
 
 quantile(time.ratio.mean, c(0.025, .975))
 
-mean(time.ratio.mean) - mean(Time.ILEC)/mean(Time.CLEC) # Bias
+time.ratio.bias <- mean(time.ratio.mean) - mean(Time.ILEC)/mean(Time.CLEC) # Bias
+
+time.ratio.bias / sd(time.ratio.mean)
+
 
 # Significance
 p <- (sum(result >= observed) + 1) / ( N + 1)

### 2020-01-27 Mathematical Statistics/05_conf_int_bootstrap.rmd
bootstrapping
bootstrapping

@@
 library(boot, quietly = T, warn.conflicts = F)
+library(ggExtra, quietly = T, warn.conflicts = F)
 
@@
 
-setwd(paste0(here::here(), "/datasets/"))
-
+data.dir <- paste0(here::here(), "/datasets/")
 ```
@@
    geom_density(aes(values), color = "darkorange") +
-   geom_vline(xintercept = mu, col = "black", height = .8, lty = 2) +
+   geom_vline(xintercept = mu, col = "black", lty = 2) +
    geom_vline(xintercept = mu - se, col = "darkred", lty = 2) +
@@
 ```{r}
+
 boot.fn <- function(data, index){
@@
 
-cst.boot <- function(values, n) {
+cst.boot <- function(values, n, I = 10e3, alpha = 0.05) {
    bootstrap <- numeric(I)
@@
 
-   alpha <- 0.05
-
    observed <- mean(values)
@@
 
-   list(observed = observed,
+   list(bootstrap = bootstrap,
+         observed = observed,
          mean = boot.mean,
@@
 ```{r}
-n.50 <- cst.boot(samp$values, 50)
+n <- 50
+samp <- data.table(values = rnorm(n, mean = mu, sd = sigma))
+n.50 <- cst.boot(samp$values, n)
 
@@
 ```{r}
+n <- 10
+samp <- data.table(values = rnorm(n, mean = mu, sd = sigma))
 n.10 <- cst.boot(samp$values, 10)
@@
 
+```{r}
+set.seed(123)
+
+n <- 200; r <- 5; lambda <- 4; mu <- lambda/r
+
+pop <- data.table(values = rgamma(n, shape = lambda, rate = r))
+
+ggplot(pop, aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange")
+```
+
 b.) Now, draw one random sample of size 200 from this population. Create a histogram of your sample, and find the mean and standard deviation.
+
+```{r}
+samp <- data.table(values = sample(pop$values, n, replace = T))
+
+ggplot(samp, aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange")
+
+xbar <- mean(samp$values); sd <- sd(samp$values)
+
+xbar; sd
+```
 
@@
 
+```{r}
+I <- 10e3
+
+boot.fn <- function(data, index) {
+   mean(data[index]$values)
+}
+
+boot(samp, boot.fn, R = I)
+
+n.200 <- cst.boot(samp$values, n)
+
+ggplot(data.table(values = n.200$bootstrap), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange") +
+   labs(title = "N=200, bootstrapped mean (Gamma r=5, lambda = 4)")
+```
+
 d.) Compare the bootstrap distribution to the approximate theoretical sampling distribution by creating a table like Table 5.2.
 
+```{r}
+tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
+                  Mean = c(mu, sd(samp$values), n.200$observed, n.200$mean),
+                  SD = c(mu, mu/sqrt(n), sd(samp$values), n.200$se))
+
+pretty_kable(tbl, "Sampling Statistics")
+```
+
 e.) Repeat (a-e) for sample sizes of n = 50, and n = 10. Describe carefully your observations about the effects of sample size on the bootstrap distribution.
+
+```{r}
+n <- 50
+
+pop <- data.table(values = rgamma(n, shape = lambda, rate = r))
+
+ggplot(pop, aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange")
+
+samp <- data.table(values = sample(pop$values, n, replace = T))
+
+ggplot(samp, aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange")
+
+xbar <- mean(samp$values); sd <- sd(samp$values)
+
+xbar; sd
+
+n.50 <- cst.boot(samp$values, n)
+
+ggplot(data.table(values = n.50$bootstrap), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange") +
+   labs(title = "N=50, bootstrapped mean (Gamma r=5, lambda = 4)")
+
+tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
+                  Mean = c(mu, sd(samp$values), n.50$observed, n.50$mean),
+                  SD = c(mu, mu/sqrt(n), sd(samp$values), n.50$se))
+
+pretty_kable(tbl, "Sampling Statistics")
+```
+
+```{r}
+n <- 10
+
+pop <- data.table(values = rgamma(n, shape = lambda, rate = r))
+
+ggplot(pop, aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange")
+
+samp <- data.table(values = sample(pop$values, n, replace = T))
+
+ggplot(samp, aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange")
+
+xbar <- mean(samp$values); sd <- sd(samp$values)
+
+xbar; sd
+
+n.50 <- cst.boot(samp$values, n)
+
+ggplot(data.table(values = n.50$bootstrap), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange") +
+   labs(title = "N=50, bootstrapped mean (Gamma r=5, lambda = 4)")
+
+tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
+                  Mean = c(mu, sd(samp$values), n.50$observed, n.50$mean),
+                  SD = c(mu, mu/sqrt(n), sd(samp$values), n.50$se))
+
+pretty_kable(tbl, "Sampling Statistics")
+```
 
@@
 ```{r}
+ne <- 14 # n even
+no <- 15 # n odd
 
+wwe <- rnorm(ne) # draw samples of size ne
+wwo <- rnorm(no) # draw random samples of size no
+
+N <- 10^4
+
+even.boot <- numeric(N) # save space
+odd.boot <- numeric(N)
+
+for(i in 1:N)
+{
+   x.even <- sample(wwe, ne, replace = T)
+   x.odd <- sample(wwo, no, replace = T)
+
+   even.boot[i] <- median(x.even)
+   odd.boot[i] <- median(x.odd)
+}
+
+boot <- data.table(even = even.boot, odd = odd.boot)
+
+p1 <- ggplot(boot, aes(even)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange") +
+   labs(title = "even")
+
+p2 <- ggplot(boot, aes(odd)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange") +
+   labs(title = "odd")
+
+gridExtra::grid.arrange(p1, p2)
 ```
+
+Change the sample sizes to 36 and 37; 200 and 201; and 10,000 and 10,001.
+
+Note the similarities/dissimalarities, trends, and so on. Why does the parity of the sample size matter? (_Note: Adjust the x limits in the plots as needed._)
+
+```{r}
+ne <- 36 # n even
+no <- 37 # n odd
+
+wwe <- rnorm(ne) # draw samples of size ne
+wwo <- rnorm(no) # draw random samples of size no
+
+N <- 10^4
+
+even.boot <- numeric(N) # save space
+odd.boot <- numeric(N)
+
+for(i in 1:N)
+{
+   x.even <- sample(wwe, ne, replace = T)
+   x.odd <- sample(wwo, no, replace = T)
+
+   even.boot[i] <- median(x.even)
+   odd.boot[i] <- median(x.odd)
+}
+
+boot <- data.table(even = even.boot, odd = odd.boot)
+
+p1 <- ggplot(boot, aes(even)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange")
+
+p2 <- ggplot(boot, aes(odd)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange")
+
+gridExtra::grid.arrange(p1, p2)
+```
+
+```{r}
+ne <- 200 # n even
+no <- 201 # n odd
+
+wwe <- rnorm(ne) # draw samples of size ne
+wwo <- rnorm(no) # draw random samples of size no
+
+N <- 10^4
+
+even.boot <- numeric(N) # save space
+odd.boot <- numeric(N)
+
+for(i in 1:N)
+{
+   x.even <- sample(wwe, ne, replace = T)
+   x.odd <- sample(wwo, no, replace = T)
+
+   even.boot[i] <- median(x.even)
+   odd.boot[i] <- median(x.odd)
+}
+
+boot <- data.table(even = even.boot, odd = odd.boot)
+
+p1 <- ggplot(boot, aes(even)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange")
+
+p2 <- ggplot(boot, aes(odd)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange")
+
+gridExtra::grid.arrange(p1, p2)
+```
+
+```{r}
+ne <- 10000 # n even
+no <- 10001 # n odd
+
+wwe <- rnorm(ne) # draw samples of size ne
+wwo <- rnorm(no) # draw random samples of size no
+
+N <- 10^4
+
+even.boot <- numeric(N) # save space
+odd.boot <- numeric(N)
+
+for(i in 1:N)
+{
+   x.even <- sample(wwe, ne, replace = T)
+   x.odd <- sample(wwo, no, replace = T)
+
+   even.boot[i] <- median(x.even)
+   odd.boot[i] <- median(x.odd)
+}
+
+boot <- data.table(even = even.boot, odd = odd.boot)
+
+p1 <- ggplot(boot, aes(even)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange")
+
+p2 <- ggplot(boot, aes(odd)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(even), col = "darkorange")
+
+gridExtra::grid.arrange(p1, p2)
+```
+
+For odd n, median will be one of the sample points. For smaller n, there will be only n possible values for the median, so the sampling distribution is more "granular" than when n is even.
+
+### 5.11
+
+Import the data from data set Bangladesh. In addition to aresnic concentrations for 271 wells, the data set contains cobolt and chlorine concentrations.
+
+a.) Conduct EDA on the chlorine concentrations and describe the salient features.
+
+```{r}
+Bangladesh <- data.table(read.csv(paste0(data.dir, "Bangladesh.csv"),
+                                 header = T))
+
+head(Bangladesh)
+
+ggplot(Bangladesh, aes(Chlorine)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(Chlorine), col = "darkorange")
+
+GGally::ggpairs(Bangladesh)
+```
+
+b.) Bootstrap the mean.
+
+```{r}
+N <- 10e3
+
+boot.fn <- function(data, index){
+   mean(data[index], na.rm = T)
+}
+
+boot(Bangladesh$Chlorine, boot.fn, R = N)
+
+observed <- mean(Bangladesh$Chlorine, na.rm = T)
+
+bootstrap <- numeric(N)
+
+for(i in 1:N)
+{
+   bootstrap[i] <- mean(sample(Bangladesh$Chlorine, size = nrow(Bangladesh), replace = T), na.rm = T)   
+}
+
+ggplot(data.table(values = bootstrap), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(aes(values), col = "darkorange")
+```
+
+c.) Find and interpret the 95% bootstrap percentile confidence interval.
+
+```{r}
+alpha <- 0.05
+quantile(bootstrap, c(alpha/2, 1 - alpha/2))
+```
+
+The spread on the confidence interval is extremely large, which is unsuprising given the heavly skewed distribtion of the sample.
+
+d.) What is the bootstrap estimate of the bias? What fraction of the bootstrap standard error does it represent?
+
+```{r}
+bias <- observed - mean(bootstrap)
+
+sd(bootstrap) / bias
+```

### 2020-01-27 Mathematical Statistics/05_conf_int_bootstrap.r

@@ -223,10 +223,10 @@ N.CLEC <- length(Time.CLEC)
 
 for(i in 1:N)
 {
-    ilec <- sample(Time.ILEC, N.ILEC, replace = T)
-    clec <- sample(Time.CLEC, N.CLEC, replace = T)
-    
-    results[i] <- mean(ilec) - mean(clec)
+  ilec <- sample(Time.ILEC, N.ILEC, replace = T)
+  clec <- sample(Time.CLEC, N.CLEC, replace = T)
+  
+  results[i] <- mean(ilec) - mean(clec)
 }
 
 alpha <- 0.05

### 2020-01-28 Mathematical Statistics/05_conf_int_bootstrap.r
pre/post reconstruct
pre/post reconstruct

@@ -30,7 +30,7 @@ sd(my.boot)
 ### Bangladesh
 
 Bangladesh <- data.table(read.csv(paste0(data.dir, "Bangladesh.csv"),
-                                 header = T))
+                                  header = T))
 
 Arsenic <- Bangladesh$Arsenic
 

### 2020-01-28 Mathematical Statistics/05_conf_int_bootstrap.rmd
pre/post reconstruct
pre/post reconstruct
bootstrap work
bootstrap work

@@
 combinations(n = 3, r = 4, v = samples, repeats.allowed = T)
+
+choose(4 + 3 - 1, 4)
 ```
@@
                                  header = T))
-
 head(Bangladesh)
@@
 ```
+
+### 5.12
+
+Consider Bangladesh chlorine (concentration). Bootstrap the trimmed mean (say, trim the upper and lower 25%), and compare your results with the usual mean (previous result).
+
+```{r}
+values <- Bangladesh[!is.na(Chlorine)]$Chlorine
+
+n <- length(values); N <- 10e3; trim <- .25
+
+observed <- mean(values, trim = trim)
+bootstrap <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   bootstrap[i] <- mean( sample(values, n, replace = T), trim = trim )
+}
+
+ggplot(data.table(value = bootstrap), aes(value)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange")
+
+alpha <- 0.05
+
+boot.mean <- mean(bootstrap)
+boot.bias <- observed - boot.mean
+boot.se <- sd(bootstrap)
+
+quantile(bootstrap, c(alpha/2, 1 - alpha/2))
+
+# boot pkg
+
+boot.fn <- function(data, index){
+   mean( data[index], trim = trim)
+}
+
+boot(values, boot.fn, R = N)
+```
+
+The bootstrap 20% trimmed mean is substantially smaller than the regular mean. The confidence intervals are also tighter, which is unsuprising due to the heavy presence of outliers in our sample.
+
+### 5.13
+
+The data set _FishMercury_ contains mercury levels (parts per million) for 30 fish caught in lakes in Minnesota.
+
+```{r}
+FishMercury <- data.table(read.csv(paste0(data.dir, "FishMercury.csv"),
+                                 header = T))
+```
+
+a.) Create a histogram or boxplot of the data. What do you observe?
+
+```{r}
+ggplot(FishMercury, aes(Mercury)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_rug(col = "darkred")
+
+FishMercury[Mercury > 1.5]
+
+ggplot(FishMercury[Mercury < 1.5], aes(Mercury)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_rug(col = "darkred")
+
+FishMercury[Mercury > 1.5] / sd(FishMercury$Mercury) # 6 sd outlier
+
+```
+
+One extreme (6 SD) outlier in the data.
+
+b.) Bootstrap the mean, and record the bootstrap standard error and the 95% bootstrap percentile interval.
+
+```{r}
+n <- nrow(FishMercury); N <- 10e3
+
+observed <- mean(FishMercury$Mercury)
+
+bootstrap <- vector(mode = "numeric", length = n)
+
+for(i in 1:N)
+{
+   bootstrap[i] <- mean( sample(FishMercury$Mercury, n, replace = T) )
+}
+
+boot.mean <- mean(bootstrap)
+boot.bias <- observed - boot.mean
+boot.se <- sd(bootstrap)
+
+alpha <- 0.05
+
+quantile(bootstrap, c(alpha/2, 1 - alpha/2))
+
+# boot pkg
+
+boot.fn <- function(data, index) {
+   mean( data[index] )
+}
+
+boot(FishMercury$Mercury, boot.fn, R = N)
+```
+
+c.) Remove the outlier and bootstrap the mean of the remaining data. Record the bootstrap standard error and the 95% bootstrap percentile interval.
+
+```{r}
+mercury <- FishMercury[Mercury < 1.5]$Mercury
+
+n <- length(mercury)
+bootstrap.values <- vector(mode = "numeric", length = n)
+
+observed <- mean(mercury)
+
+for( i in 1:N)
+{
+   bootstrap.values[i] <- mean( sample(mercury, size = n, replace = T) )   
+}
+
+mean(bootstrap.values)
+
+boot(mercury, boot.fn, R = N)
+
+quantile(bootstrap.values, c(alpha/2, 1 - alpha/2))
+```
+
+The standard error reduced drastically, and the confidence intervals are much more narrow.
+
+### 5.14
+
+In Section 3.3, we performed a permutation test to determine if men and women consumed, on average, different amounts of hot wings.
+
+```{r}
+BeerWings <- data.table(read.csv(paste0(data.dir, "Beerwings.csv"),
+                                 header = T))
+
+wings <- BeerWings$Hotwings
+
+?stat_ecdf
+
+ggplot(BeerWings, aes(Hotwings, group = Gender, col = Gender)) +
+   stat_ecdf(geom = "step")
+
+N <- 10e3; alpha <- 0.05; n <- length(wings)
+
+wings.m <- BeerWings[Gender == "M"]$Hotwings
+wings.f <- BeerWings[Gender == "F"]$Hotwings
+
+observed <- mean(wings.m) - mean(wings.f)
+
+permutation.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp <- sample(1:n, length(wings.m), replace = F)
+
+   samp.m <- wings[samp]; samp.w <- wings[-samp]
+
+   permutation.values[i] <- mean(samp.m) - mean(samp.w)
+}
+
+ggplot(data.table(values = permutation.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+ggplot(data.table(value = permutation.values), aes(value)) +
+   stat_ecdf(geom = "step") +
+   geom_vline(xintercept = observed, col = "darkred") +
+   scale_y_continuous(labels = comma)
+
+p <- (sum(permutation.values >= observed) + 1) / (N + 1)
+
+var <- p*(1 - p)/N
+```
+
+a.) Bootstrap the difference in means and describe the bootstrap distribution.
+
+```{r}
+bootstrap.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp.m <- sample(wings.m, size = length(wings.m), replace = T)
+   samp.f <- sample(wings.f, size = length(wings.f), replace = T)
+
+   bootstrap.values[i] <- mean(samp.m) - mean(samp.f)
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30)
+
+boot.mean <- mean(bootstrap.values)
+boot.se <- sd(bootstrap.values)
+
+boot.bias <- observed - boot.mean
+
+boot.se; boot.bias
+```
+
+b.) Find a 95% bootstrap percentile confidence interval for the difference of means, and give a sentence interpreting this interval.
+
+```{r}
+quantile(bootstrap.values, c(alpha/2, 1 - alpha/2))
+```
+
+The 95% confidence interval does not contain 0, further supporting our permutation test results.
+
+c.) How do the bootstrap and permutation distributions differ?
+
+The permutation distribution is sampled without replacement, and the bootstrap distribution is sampled using replacement.
+
+### 5.15
+
+A high school student was curious about the total number of minutes devoted to commercials during any given half-hour time period on basic and extended cable TV channels.
+(B.
+Rodgers and T.
+Robinson, personal communication).
+
+Import the TV data.
+
+```{r}
+TV <- data.table(read.csv(paste0(data.dir, "TV.csv"),
+                                 header = T))
+```
+
+a.) Perform some exploratory data analysis and obtain summary statistics on the commercial times on basic and extended cable TV channels (do seperate analysis for each type of channel).
+
+```{r}
+ggplot(TV, aes(Cable, Times, fill = Cable)) +
+   geom_boxplot()
+
+ggplot(TV, aes(Times, group = Cable)) +
+   geom_density(aes(y = ..density.., fill = Cable), alpha = .45)
+
+aov(TV$Times ~ TV$Cable)
+
+tv.basic <- TV[Cable == "Basic"]$Times
+tv.extended <- TV[Cable == "Extended"]$Times
+tv.pooled <- c(tv.basic, tv.extended)
+
+observed <- mean(tv.basic) - mean(tv.extended)
+
+mean(tv.basic); mean(tv.extended)
+```
+
+b.) Bootstrap the difference in mean times, plot the distribution, and give summary statistics of the bootstrap distribution.
+Obtain a 95% bootstrap percentile confidence interval, and interpret this interval.
+
+```{r}
+N <- 10e3; n <- length(TV$Times); alpha <- 0.05
+
+bootstrap.values <- vector(mode = "numeric", N)
+
+for(i in 1:N)
+{
+   samp.basic <- sample(tv.basic, size = length(tv.basic), replace = T)
+   samp.extended <- sample(tv.extended, size = length(tv.extended), replace = T)
+
+   bootstrap.values[i] <- mean(samp.basic) - mean(samp.extended)
+}
+
+boot.mean <- mean(bootstrap.values)
+boot.se <- sd(bootstrap.values)
+boot.bias <- observed - boot.mean
+
+quantile(bootstrap.values, c(alpha/2, 1 - alpha/2)) # bootstrap interval
+```
+
+The 95% boostrap interval does not contain 0, suggesting there is in fact a difference in the times between basic and extended (basic being longer).
+
+c.) What is the bootstrap estimate of the bias?
+
+```{r}
+boot.bias
+
+boot.bias / boot.se # about 1% of the standard error
+```
+
+d.) Conduct a permutation test to see if the difference in mean commercial times is statistically significant, and state your conclusion.
+
+```{r}
+permutation.values <- vector(mode = "numeric", n = N)
+
+for(i in 1:N)
+{
+   samp <- sample(length(tv.pooled), length(tv.basic), replace = F)
+
+   samp.basic <- tv.pooled[samp]; samp.extended <- tv.pooled[-samp]
+
+   permutation.values[i] <- mean(samp.basic) - mean(samp.extended)
+}
+
+ggplot(data.table(values = permutation.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_vline(xintercept = observed, col = "darkred")
+
+p <- ( sum(permutation.values >= observed) + 1) / ( N + 1)
+var <- p*(1 - p)/N
+```
+
+Permutation test supports the alternative hypothesis that there is a difference in commercial times.

### 2020-01-29 Mathematical Statistics/05_conf_int_bootstrap.rmd
more bootstrapping
more bootstrapping

@@
 library(ggExtra, quietly = T, warn.conflicts = F)
+library(GGally, quietly = T, warn.conflicts = F)
 
@@
    boot.mean <- mean(bootstrap)
-   boot.bias <- observed - boot.mean
+   boot.bias <- boot.mean - observed
    boot.se <- sd(bootstrap)
@@
 ```{r}
-bias <- observed - mean(bootstrap)
-
-sd(bootstrap) / bias
-```
+bias <-  mean(bootstrap) -observed
+
+bias / sd(bootstrap)
+```
+
+roughly 1% of the standard error.
 
@@
 boot.mean <- mean(bootstrap)
-boot.bias <- observed - boot.mean
+boot.bias <- boot.mean - observed
 boot.se <- sd(bootstrap)
@@
 
-boot.bias <- observed - boot.mean
+boot.bias <- boot.mean - observed
 
@@
 boot.se <- sd(bootstrap.values)
-boot.bias <- observed - boot.mean
+boot.bias <- boot.mean - observed
 
@@
 ```{r}
-permutation.values <- vector(mode = "numeric", n = N)
+permutation.values <- vector(mode = "numeric", length = N)
 
@@
 Permutation test supports the alternative hypothesis that there is a difference in commercial times.
+
+### 5.16
+
+Researchers conducted a study of primary and early secondary school children in Italy to examine gender differences in math anxiety.
+One of the measures used to understand math anxiety is the _Abbreviated Math Anxiety Scale (AMAS)_, a self-reported math anxiety questionnaire.
+A higher score indicates more max anxiety.
+The data set MathAnxiety contains the results for a subset of the children in the original study.
+
+```{r}
+MathAnxiety <- data.table(read.csv(paste0(data.dir, "MathAnxiety.csv"),
+                                 header = T))
+```
+
+a.) Perform some exploratory analysis and obtain summary statistics of the AMAS scores for the boys and girls (do seperate analysis for each gender).
+
+```{r}
+ggplot(MathAnxiety, aes(Grade, AMAS)) +
+   geom_boxplot(aes(fill = Grade))
+
+ggplot(MathAnxiety, aes(Age, AMAS)) +
+   geom_point()
+
+ggplot(MathAnxiety, aes(RCMAS, AMAS)) +
+   geom_point(aes(color = Gender))
+
+ggplot(MathAnxiety, aes(Gender, AMAS)) +
+   geom_boxplot(aes(color = Gender))
+
+ggpairs(MathAnxiety)
+```
+
+b.) Bootstrap the difference in mean times, plot the distribution, and give summary statistics of the bootstrap distribution. Obtain a 95% confidence interval, and interperet this interval.
+
+```{r}
+amas_boy <- MathAnxiety[Gender == "Boy"]$AMAS
+amas_girl <- MathAnxiety[Gender == "Girl"]$AMAS
+
+N <- 10e3; n <- nrow(MathAnxiety); alpha <- 0.05
+
+observed <- mean(amas_boy) - mean(amas_girl)
+
+bootstrap.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp_boy <- sample(amas_boy, length(amas_boy), replace = T)
+   samp_girl <- sample(amas_girl, length(amas_girl), replace = T)
+
+   bootstrap.values[i] <- mean(samp_boy) - mean(samp_girl)
+}
+
+boot.mean <- mean(bootstrap.values)
+boot.bias <- boot.mean - observed
+boot.se <- sd(bootstrap.values)
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(color = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+boot.bias; boot.se
+
+quantile(bootstrap.values, c(alpha/2, 1 - alpha/2)) # conf interval
+```
+
+The 95% confidence interval is -2.8 ~ -0.7, which does not include zero.
+The data suggests that there is a statistical difference between the self-reported AMAS scores between boys and girls (boys being lower).
+
+c.) What is the bootstrap estimate of the bias? What fraction of the bootstrap; standard error does this represent?
+
+```{r}
+boot.bias
+
+boot.bias / boot.se
+```
+
+Bootstrap bias is .006, which is less than 1% of the se.
+
+d.) Conduct a permutation test to see if the difference in mean AMAS scores is statistically significant, and state your conclustion.
+
+```{r}
+amas_pooled <- c(amas_boy, amas_girl)
+
+n <- length(amas_pooled); observed <- mean(amas_boy) - mean(amas_girl)
+
+permutation.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp <- sample(1:n, length(samp_boy), replace = F)
+
+   samp_boys <- amas_pooled[samp]; samp_girl <- amas_pooled[-samp]
+
+   permutation.values[i] <- mean(samp_boy) - mean(samp_girl)
+}
+
+mean(permutation.values)
+
+ggplot(data.table(value = permutation.values), aes(value)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed)
+
+p <- ( sum(permutation.values <= observed) + 1 ) / (N + 1)
+var <- p*(1 - p)/ N
+
+t.test(amas_boy, amas_girl, alternative = "less")
+```
+
+The p-value for the permutation test is less than 1%, we reject the null hypothesis that there is no difference in self-reported AMAS scores between boys and girls.
+This conclustion is in-line with our results from the bootstraped difference in mean test.
+
+### 5.17
+
+Import the data from Girls2004.
+
+```{r}
+Girls2004 <- data.table(read.csv(paste0(data.dir, "Girls2004.csv"),
+                                 header = T))
+```
+
+a.) Perform some exploratory data analysis, and obtain summary statistics on the weights of baby girls born in Wyoming and Alaska.
+
+```{r}
+ggpairs(Girls2004)
+
+ggplot(Girls2004, aes(Smoker, Weight, fill = State)) +
+   geom_boxplot() +
+   scale_y_continuous(labels = scales::comma)
+
+ggplot(Girls2004, aes(MothersAge, Weight, fill = State)) +
+   geom_boxplot()
+```
+
+b.) Bootstrap the difference in means, plot the distribution, and give the summary statistics. Obtain a 95% confidence boostrap percentile confidence interval and interpret this interval.
+
+```{r}
+weights_wy <- Girls2004[State == "WY"]$Weight
+weights_ak <- Girls2004[State == "AK"]$Weight
+
+observed <- mean(weights_wy) - mean(weights_ak); N <- 10e3
+alpha <- 0.05
+
+bootstrap.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp_wy <- sample(weights_wy, size = length(weights_wy), replace = T)
+   samp_ak <- sample(weights_ak, size = length(weights_ak), replace = T)
+
+   bootstrap.values[i] <- mean(samp_wy) - mean(samp_ak)
+}
+
+bootstrap.mean <- mean(bootstrap.values)
+bootstrap.bias <- bootstrap.mean - observed
+bootstrap.se <- sd(bootstrap.values)
+
+ggplot(data.table(value = bootstrap.values), aes(value)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1- alpha/2))
+```
+
+There is a statistically significant difference in the mean birth weights of girls born in Wyoming and Alaska (Wyoming girls weighing less on average), by between 530 and 96 oz less.
+
+c.) What is the bootstrap estimate of the bias? What fraction of thh boostrap standard error does it represent?
+
+```{r}
+bootstrap.bias
+
+bootstrap.bias / bootstrap.se
+```
+
+d.) Conduct a permutation test to see if the difference in mean weights is statistically significant, and state your conclusion.
+
+```{r}
+observed <- mean(weights_wy) - mean(weights_ak)
+weights_pooled <- c(weights_wy, weights_ak)
+
+permutation.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp <- sample(length(weights_pooled), size = length(weights_ak), replace = F)
+
+   samp_wy <- weights_pooled[samp]; samp_ak <- weights_pooled[-samp]
+
+   permutation.values[i] <- mean(samp_wy) - mean(samp_ak)
+}
+
+mean(permutation.values)
+
+ggplot(data.table(values = permutation.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+p <- (sum(permutation.values <= observed) + 1) / (N + 1)
+
+var <- p*(1 - p)/N
+```
+
+The permutation test further supports the claim that the weights of baby girls born in Wyoming and Alaska are different.
+We reject the null hypothesis that the means are the same at the 0.05% level (p = 0.003).
+
+e.) For what population(s), if any, does this conclusion hold?
+
+_I do not
+### 5.18
+
+Is there a difference in the price of groceries sold by the two retailers Target and Walmart?
+The data set _Groceries_ contain, a sample of grocery items and their prices advertised on their respective websites on one specific day.
+
+```{r}
+Groceries <- data.table(read.csv(paste0(data.dir, "Groceries.csv"),
+                                 header = T))
+```
+
+a.) Compute summary statistics of the prices for each store.
+
+```{r, fig.height=7}
+groceries <- melt(Groceries %>% select(-Size), 
+                  id.vars = c("Product"), 
+                  value.name = "Price", 
+                  variable.name = "Store")
+
+setorder(groceries, -Product)
+
+ggplot(groceries, aes(Product, Price, col = Store, group = Store)) +
+   geom_point(aes(color = Store)) +
+   coord_flip()
+```
+
+b.) Use the bootstrap to determine whether or not there is a difference in the mean prices.
+
+```{r}
+prices.target <- groceries[Store == "Target"]$Price
+prices.walmart <- groceries[Store == "Walmart"]$Price
+
+N <- 10e3
+
+observed <- mean(prices.target) - mean(prices.walmart)
+bootstrap.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp.target <- sample(prices.target, size = length(prices.target), replace = T)
+   samp.walmart <- sample(prices.walmart, size = length(prices.walmart), replace = T)
+
+   bootstrap.values[i] <- mean(samp.target) - mean(samp.walmart)
+}
+
+boot.mean <- mean(bootstrap.values)
+boot.bias <- boot.mean - observed
+boot.se <- sd(bootstrap.values)
+
+boot.bias / boot.se
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1 - alpha/2))
+```
+
+There doesn't appear to be a statistically significant difference in the average prices at the stores.
+
+c.) Create a histogram of the difference in prices. What is unusal about Quaker Oats Life cereal?
+
+```{r}
+Groceries[Product == "Quaker Oats Life Cereal  Original ",]
+```
+
+It's almost double the price.
+
+d.) Recompute the bootstrap percentile interval without this observation. What do you conclude?
+
+```{r}
+groceries_sans_quaker <- groceries[Product != "Quaker Oats Life Cereal  Original ",]
+
+stopifnot(nrow(groceries_sans_quaker) == nrow(groceries) - 2) # ensure removeed
+
+prices.target <- groceries_sans_quaker[Store == "Target"]$Price
+prices.walmart <- groceries_sans_quaker[Store == "Walmart"]$Price
+
+N <- 10e3
+
+observed <- mean(prices.target) - mean(prices.walmart)
+bootstrap.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   samp.target <- sample(prices.target, size = length(prices.target), replace = T)
+   samp.walmart <- sample(prices.walmart, size = length(prices.walmart), replace = T)
+
+   bootstrap.values[i] <- mean(samp.target) - mean(samp.walmart)
+}
+
+boot.mean <- mean(bootstrap.values)
+boot.bias <- boot.mean - observed
+boot.se <- sd(bootstrap.values)
+
+boot.bias / boot.se
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1 - alpha/2))
+```
+
+The bootstrap confidence intervals include zero, so there is not sufficent evidence here to support the claim that the prices have a statistically significant difference.

### 2020-01-30 Mathematical Statistics/05_conf_int_bootstrap.rmd
never enough bootstrap
(NEED MORE BOOTSTRAP)

@@
 ggplot(MathAnxiety, aes(Gender, AMAS)) +
-   geom_boxplot(aes(color = Gender))
+   geom_boxplot(aes(fill = Gender))
 
@@
 {
-   samp.target <- sample(prices.target, size = length(prices.target), replace = T)
-   samp.walmart <- sample(prices.walmart, size = length(prices.walmart), replace = T)
+   index <- sample(length(prices.target), size = length(prices.target), replace = T)
+
+   samp.target <- prices.target[index]
+   samp.walmart <- prices.walmart[index]
 
@@
 {
-   samp.target <- sample(prices.target, size = length(prices.target), replace = T)
-   samp.walmart <- sample(prices.walmart, size = length(prices.walmart), replace = T)
+   # paired data, must use the same products from each company
+   index <- sample(length(prices.target), size = length(prices.target), replace = T)
+
+   samp.target <- prices.target[index]
+   samp.walmart <- prices.walmart[index]
 
@@
 
-The bootstrap confidence intervals include zero, so there is not sufficent evidence here to support the claim that the prices have a statistically significant difference.
+The bootstrap confidence intervals suggest that there is a statistically significant difference in prices (Target being more expensive than Walmart). The Quaker Oats cereal made a difference.
+
+### 5.19
+
+Do chocolate and vanilla ice cream have the same number of calories?
+
+The data set Ice Cream contains calorie information for a sample of brands of chocolate and vanilla ice cream.
+
+```{r}
+IceCream <- data.table(read.csv(paste0(data.dir, "IceCream.csv"),
+                                 header = T))
+```
+
+a.) Compute summary statistics of the calories for the two flavors.
+
+```{r}
+
+icecream <- IceCream %>%
+   mutate(Vanilla = VanillaCalories,
+          Chocolate = ChocolateCalories) %>%
+   select(Vanilla, Chocolate) %>%
+   melt(variable.name = "Flavor", value.name = "Calories") %>%
+   as.data.table()
+
+stopifnot(nrow(icecream) == nrow(IceCream) * 2)
+
+ggplot(icecream, aes(Calories, group = Flavor)) +
+   geom_density(aes(fill = Flavor), alpha = .35)
+
+ggplot(icecream, aes(Flavor, Calories, fill = Flavor)) +
+   geom_boxplot()
+```
+
+b.) Use the bootstrap to determine whether or not there is a difference in the mean number of calories.
+
+```{r}
+calories.vanilla <- icecream[Flavor == "Vanilla"]$Calories
+calories.chocolate <- icecream[Flavor == "Chocolate"]$Calories
+
+N <- 10e3
+
+observed <- mean(calories.chocolate) - mean(calories.vanilla)
+
+bootstrap.values <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   # paired data, must use the same samples from chocolate and vanilla (ie, same manufacture)
+   index <- sample(length(calories.vanilla), length(calories.vanilla), replace = T)
+
+   v.samp <- calories.vanilla[index]
+   c.samp <- calories.chocolate[index]
+
+   bootstrap.values[i] <- mean(c.samp) - mean(v.samp)
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed)
+
+bootstrap.mean <- mean(bootstrap.values)
+bootstrap.bias <- bootstrap.mean - observed
+bootstrap.se <- sd(bootstrap.values)
+
+bootstrap.bias / bootstrap.se
+
+quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1 - alpha/2))
+```
+
+There appers to be a statistically difference in the amount of calories (Chocolate having more, by approx 4-12 on average).
+
+### 5.20
+
+In a remark at the end of Section 5.4.1, we mentioned that the procedure for bootstrapping the difference of medians is different than for thee mean.
+
+Import the data set Diving2017.
+
+```{r}
+
+```

### 2020-01-30 Mathematical Statistics/05_conf_int_bootstrap.rmd
bootstrap
bootstrap

@@
 ```{r}
-
 icecream <- IceCream %>%
@@
 ```{r}
-
-```
+Diving2017 <- data.table(read.csv(paste0(data.dir, "Diving2017.csv"),
+                                 header = T))
+```
+
+a.) Compute the difference in the median scores in the final and semi-final rounds.
+
+```{r}
+observed <- median(Diving2017$Final) - median(Diving2017$Semifinal)
+```
+
+b.) Calculate a 95% bootstrap interval for this statistic.
+
+```{r}
+N <- 10^5
+
+result <- vector(mode = "numeric", length = N)
+
+for(i in 1:N)
+{
+   index <- sample(nrow(Diving2017), replace = T)
+
+   Dive.boot <- Diving2017[index, ] # resample pairs
+   result[i] <- median(Dive.boot$Final) - median(Dive.boot$Semifinal)
+}
+
+ggplot(data.table(values = result), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange")
+```
+
+### 5.21
+
+Two college students collecteed data on the price of hardcover textbooks from two disciplinary areas: Mathematics and the Natural Sciences and the Social Sciences. The data are in the file BookPrices.
+
+```{r}
+BookPrices <- data.table(read.csv(paste0(data.dir, "BookPrices.csv"),
+                                 header = T))
+```
+
+a.) Perform some exploratory data analysis on book prices for each of the two diciplinary areas.
+
+```{r}
+ggpairs(BookPrices)
+
+ggplot(BookPrices, aes(Area, Price, group = Area)) +
+   geom_boxplot(aes(fill = Area))
+```
+
+b.) Bootstrap the mean of book price for each area separately, and describe the distributions.
+
+```{r}
+
+prices.math <- BookPrices[Area == "Math & Science"]$Price
+prices.science <- BookPrices[Area == "Social Sciences"]$Price
+
+N <- 10e4
+
+bootstrap.values <- vector(mode = "numeric", length = N); n <- length(prices.math)
+observed <- mean(prices.math)
+
+for(i in 1:N)
+{
+   bootstrap.values[i] <- mean( sample(prices.math, n, replace = T) )
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+bootstrap.mean <- mean(bootstrap.values)
+
+```
+
+```{r}
+N <- 10e4
+
+bootstrap.values <- vector(mode = "numeric", length = N); n <- length(prices.science)
+observed <- mean(prices.science)
+
+for(i in 1:N)
+{
+   bootstrap.values[i] <- mean( sample(prices.science, n, replace = T) )
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+bootstrap.mean <- mean(bootstrap.values)
+```
+
+_Both of the distributions are relatively symetrical and approximately normal._
+
+c.) Bootstrap the ratio of means. Provide plots of the bootstrap distribution and comment.
+
+```{r}
+N <- 10e4
+
+bootstrap.values <- vector(mode = "numeric", length = N); n <- length(prices.science)
+observed <- mean(prices.math) / mean(prices.science)
+
+for(i in 1:N)
+{
+   samp.math <- sample(prices.math, length(prices.math), replace = T)
+   samp.science <- sample(prices.science, length(prices.science), replace = T)
+
+   bootstrap.values[i] <- mean(samp.math) / mean(samp.science)
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed, col = "darkred")
+
+bootstrap.mean <- mean(bootstrap.values)
+```
+
+_The ratio of means distribution is heavly skewed with a long right tail._
+
+d.) Find the 95% bootstrap percentile interval for the ratio of means. Interpret this interval.
+
+```{r}
+alpha <- 0.05
+
+quantile(bootstrap.values, c(alpha/2, 1 - alpha/2))
+```
+
+_There is statistically significant evidence that the cost of math books are higher than science books._
+
+e.) What is the bootstrap estimate of the bias? What fraction of the bootstrap standard error does it represent?
+
+```{r}
+bootstrap.bias <- mean(bootstrap.values) - observed
+bootstrap.se <- sd(bootstrap.values)
+
+bootstrap.bias
+
+bootstrap.bias / bootstrap.se
+```
+
+The bias in the bootstrap is quite high at 5%. Also, the bias is almost 16% of the SE which is extremely high.
+
+### 5.22
+
+Import the data from flight delays case study in S1.1.
+
+For this study, we will consider the ratio of means.
+
+```{r}
+FlightDelays <- data.table(read.csv(paste0(data.dir, "FlightDelays.csv"),
+                                 header = T))
+```
+
+a.) Perform some exploratory data analysis on flight delay lengths.
+
+```{r}
+ggpairs(FlightDelays)
+
+ggplot(FlightDelays, aes(Carrier, Delay, col = Carrier)) +
+   geom_boxplot()
+```
+
+b.) Bootstrap the mean of flight delay lengths for each airline separately, and describe the distribution.
+
+```{r}
+ua.delay <- FlightDelays[Carrier == "UA"]$Delay
+aa.delay <- FlightDelays[Carrier == "AA"]$Delay
+
+n <- length(ua.delay); N <- 10e4
+bootstrap.values <- vector(mode = "numeric", length = n)
+
+for(i in 1:N)
+{
+   bootstrap.values[i] <- mean( sample(ua.delay, n, replace = T))   
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange")
+
+n <- length(aa.delay); N <- 10e4
+bootstrap.values <- vector(mode = "numeric", length = n)
+
+for(i in 1:N)
+{
+   bootstrap.values[i] <- mean( sample(aa.delay, n, replace = T))   
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange")
+
+```
+
+c.) Bootstrap the ratio of means. Provide plots of the bootstrap distribution and describe the distribution.
+
+```{r}
+n <- length(aa.delay); N <- 10e4
+bootstrap.values <- vector(mode = "numeric", length = n)
+
+observed <- mean(aa.delay) / mean(ua.delay)
+
+for(i in 1:N)
+{
+   bootstrap.values[i] <- mean( sample(aa.delay, n, replace = T)) / mean( sample(ua.delay, n, replace = T))   
+}
+
+ggplot(data.table(values = bootstrap.values), aes(values)) +
+   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
+   geom_density(col = "darkorange") +
+   geom_vline(xintercept = observed)
+```
+
+d.) Find and interpret the 95% confidence bootstrap interval for the ratio of means.
+
+```{r}
+quantile(bootstrap.values, c(alpha/2, 1 - alpha/2))
+```
+
+Distribution is approximately normal, and the interval contains zero so we can't rule out chance in the flight delay data.
+
+e.) What is the bootstrap estimate of the bias? What fraction of the standard error does it represent?
+
+```{r}
+boot.bias <- mean(bootstrap.values) - observed
+
+boot.bias / sd(bootstrap.values)
+```
+
+The se is approximately 3% of the bias.

### 2020-01-30 Mathematical Statistics/05_conf_int_bootstrap.rmd
wrap-up bootstrap
wrap-up bootstrap

@@
    geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed)
+   geom_vline(xintercept = observed, col = "darkred")
 ```

### 2020-02-11 Mathematical Statistics/06_Estimation.rmd
estimation / regularization
estimation / regularization

--- 2020-01-30 Mathematical Statistics/05_conf_int_bootstrap.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Confidence Intervals - The Bootstrap}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Estimation}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 5}
+   \lfoot{\color{headergrey}Chapter 6}
    \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
@@
 
-#### 5.1
-
-Consider the samples 1-6. Use a six-sided die to obtain three different bootstrap samples and their corresponding means.
-
-```{r,echo=T}
-pop <- seq(from = 1, to = 6, by = 1)
-
-n <- 6
-
-s1 <- mean( sample(pop, n, replace = T) )
-s2 <- mean( sample(pop, n, replace = T) )
-s3 <- mean( sample(pop, n, replace = T) )
-
-```
-
-$\bar{x}^*_1 = `r s1`$, $\bar{x}^*_2 = `r s2`$, $\bar{x}^*_3 = `r s3`$
-
-#### 5.2
-
-Consider the samples 1, 3, 4, and 6 from some distribution.
-
-```{r, echo = T}
-
-pop <- c(1, 3, 4, 6)
-
-samples <- permutations(n = 4, r = 4, pop, repeats.allowed = T)
-
-```
-
-a.) For one random bootstrap sample, find the probability that the mean is one.
-
-```{r, echo = T}
-
-means <- apply(samples, 1, mean)
-
-p <- mean( means == 1 )
-```
-
-Probability: __`r round(p, 4) * 100`%__
-
-b.) For one random bootstrap sample, find the probability that the maximum is 6.
-
-```{r, echo = T}
-maxes <- apply(samples, 1, max)
-
-p <- mean( maxes == 6 )
-```
-
-Probability: __`r round(p, 4) * 100`%__
-
-c.) For one random bootstrap sample, find the probability that exactly two elements in the sample are less than 2.
-
-```{r, echo = T}
-
-lt2 <- apply(t(apply(samples, 1, function(x) { x < 2})), 1, sum)
-
-p <- mean( lt2 == 2 )
-```
-
-Probability: __`r round(p, 4) *100`%__
-
-### 5.3
-
-Consider the sample 1-3.
-
-a.) List all the (ordered) bootstrap samples from this sample. How many are there?
-
-```{r, echo = T}
-samples <- permutations(n = 3, r = 3, 1:3, repeats.allowed = T)
-
-n <- nrow(samples)
-```
-
-Samples: = $3^3$ = __`r n`__
-
-b.) How many unordered bootstrap samples are there? For example, {1, 2, 2} and {2, 1, 2} are considered to be the same.
-
-```{r}
-samples <- combinations(n = 3, r = 3, 1:3, repeats.allowed = T)
-
-n <- nrow(samples)
-
-assertthat::are_equal(n, choose(3 + 3 - 1, 3))
-```
-
-Samples: = $5 \choose 3$ = __`r n`__
-
-c.) How many ordered bootstrap samples have one occurrence of 1 and two occurences of 3?
-
-```{r}
-samples <- permutations(n = 3, r = 3, 1:3, repeats.allowed = T)
-
-n.ones <- apply(t(apply(samples, 1, FUN = function(x) { x == 1 })), 1, function(x) sum(x) )
-n.threes <- apply(t(apply(samples, 1, FUN = function(x) { x == 3 })), 1, function(x) sum(x) )
-
-sum((n.ones == 1 & n.threes == 2) == T)
-```
-
-Is this the same number of bootstrap samples that have each of 1, 2 and 3 occuring exactly once?
-
-```{r}
-
-n.ones <- apply(t(apply(samples, 1, FUN = function(x) { x == 1 })), 1, function(x) sum(x) )
-n.twos <- apply(t(apply(samples, 1, FUN = function(x) { x == 2 })), 1, function(x) sum(x) )
-n.threes <- apply(t(apply(samples, 1, FUN = function(x) { x == 3 })), 1, function(x) sum(x) )
-
-sum((n.ones == 1 & n.twos == 1 & n.threes == 1) == T)
-```
-
-No, 3 != 6.
-
-d.) Is the probability of obtaining a bootstrap sample with one 1 and two 3's the same as the probabiliity of obtaining a bootstrap sample with each of 1, 2 and 3 occuring exactly once?
-
-```{r, echo = T}
-
-( sum((n.ones == 1 & n.threes == 2)) / n ) == ( sum((n.ones == 1 & n.twos == 1 &  n.threes == 1)) / n )
-
-```
-
-No, 3% and 6% chances respectfully.
-
-### 5.4
-
-Consider the samples 1, 3, 3, and 5 from some distribution.
-
-```{r}
-samples <- c(1, 3, 3, 5)
-```
-
-a.) How many bootstrap samples are there?
-
-```{r}
-boot <- permutations(n = 3, r = 4, v = samples, repeats.allowed = T)
-
-n <- nrow(boot)
-```
-
-_3 unique items to pick from, 4 places to put each item._
-
-Number of permutations: *$3^4$ = `r n`*
-
-b.) List the distinct bootstrap samples assuming order does not matter.
-
-```{r}
-combinations(n = 3, r = 4, v = samples, repeats.allowed = T)
-
-choose(4 + 3 - 1, 4)
-```
-
-### 5.5
-
-We determine the number of distinct bootstrap samples from a finite set.
-
-a.) A bakery sells five types of cookies: sugar, chocolate chip, oatmeal, peanut butter, and ginger snap. Show that the number of ways to order 5 cookies is $9 \choose 5$
-
-Unordered samping with replacement: $n + k - 1 \choose k$, $n = 5, k = 5$
-
-```{r}
-choose(5 + 5 - 1, 5)
-```
-
-b.) Show that the number of sets of size n (order does not matter) drawn with replacement from the (distinct) $a_1, a_2, \ldots, a_n$ is ${2n-1}\choose n$
-
-Conclude that the number of distinct bootstrap samples from the set $[a_1, a_2, \ldots, a_n]$ is ${2n - 1} \choose  n$
-
-### 5.6
-
-Let $k_1, k_2, \ldots, k_n$ denote non-negative integers satisfying $k_1 + k_2 + \ldots + k_n = n$, and suppose the elements in the set $a_1, a_2, \ldots, a_n$ are distinct.
-
-a.) Show that the number of bootstrap samples with $k_1$ occurrences of $a_1, k_2$ occurrences of $a_2, \ldots, k_n$ occurrences of $a_n$ is $n \choose k_1, k_2, \ldots, k_n$
-
-b.) Compute the probability that a randomly drawn bootstrap sample will have $k_i$ occurrences of $a_i, i = 1, 2, \ldots, n$
-
-### 5.7
-
-Refer to Example 5.4 and the remark at the end of the example.
-
-a.) What might account for the fact that there were more missing values for the men who skateboarded in front of the male experimenter? How might this bias the outcome?
-
-_It could be that the approached skateboarders refused to participate in the study of performing tricks in front of other men._
-
-b.) Why do you suppose it was important that the two experimenters were blinded to the purpose of the study?
-
-_The female could have flurted or otherwise influenced skateboarders who were performing tricks if they knew the intent of the study._
-
-### 5.8
-
-Consider a population that has a normal distribution with mean $\mu = 36$, standard deviation $\sigma = 8$.
-
-```{r}
-mu <- 36; sigma <- 8; n <- 200
-
-se <- mu / sqrt(n)
-```
-
-a.) The sampling distribution of $\bar{X}$ for samples of size 200 will have what mean, standard error, and shape?
-
-_$\mu = 36$, SE = 36 / sqrt(200) = `r se`, shape will be approximately normal (CLT)._
-
-b.) Use R to draw a random sample of size 200 from this population. Conduct EDA on your sample.
-
-```{r}
-set.seed(123)
-
-samp <- data.table(values = rnorm(200, mean = 36, sd = 8))
-
-ggplot(samp, aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), color = "darkorange") +
-   geom_vline(xintercept = mu, col = "black", lty = 2) +
-   geom_vline(xintercept = mu - se, col = "darkred", lty = 2) +
-   geom_vline(xintercept = mu + se, col = "darkred", lty = 2)
-
-```
-
-c.) Compute the bootstrap distribution for your sample, and note the bootstrap mean and standard error.
-
-```{r}
-
-boot.fn <- function(data, index){
-   mean(data[index]$values)
-}
-
-I <- 10e3
-
-boot(samp, boot.fn, R = I) # boot pkg
-
-cst.boot <- function(values, n, I = 10e3, alpha = 0.05) {
-   bootstrap <- numeric(I)
-
-   for(i in 1:I)
-   {
-      bootstrap[i] <- mean( sample(values, n, replace = T) )
-   }
-
-   observed <- mean(values)
-
-   boot.mean <- mean(bootstrap)
-   boot.bias <- boot.mean - observed
-   boot.se <- sd(bootstrap)
-
-   list(bootstrap = bootstrap,
-         observed = observed,
-         mean = boot.mean,
-         bias = boot.bias,
-         se = boot.se,
-         conf = quantile(bootstrap, c(alpha/2, 1 - alpha/2)))
-}
-
-```
-
-d.) Compare the bootstrap distribution to the theoretical sampling distribution by creating a table like Table 5.2:
-
-```{r}
-n.200 <- cst.boot(samp$values, 200)
-
-tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
-                  Mean = c(mu, mu, n.200$observed, n.200$mean),
-                  SD = c(sigma, mu/sqrt(n), sd(samp$values), n.200$se))
-
-pretty_kable(tbl, "Sampling Statistics")
-```
-
-e.) Repeat for sample sizes $n = 50$ and $n = 10$. Carefully describe yyour observations about the effects of sample size on the bootstrap distribuiton.
-
-```{r}
-n <- 50
-samp <- data.table(values = rnorm(n, mean = mu, sd = sigma))
-n.50 <- cst.boot(samp$values, n)
-
-tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
-                  Mean = c(mu, sigma, n.50$observed, n.50$mean),
-                  SD = c(sigma, mu/sqrt(n), sd(samp$values), n.50$se))
-
-pretty_kable(tbl, "Sampling Statistics")
-
-```
-
-```{r}
-n <- 10
-samp <- data.table(values = rnorm(n, mean = mu, sd = sigma))
-n.10 <- cst.boot(samp$values, 10)
-
-tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
-                  Mean = c(mu, sigma, n.10$observed, n.10$mean),
-                  SD = c(sigma, mu/sqrt(n), sd(samp$values), n.10$se))
-
-pretty_kable(tbl, "Sampling Statistics")
-```
-
-_The center of the bootstrap distribution doesn't vary much with smaller n, however, confidence intervals (the sd of the bootstrap dist) vary wildy._
-
-### 5.9
-
-Consider a population that has a gamma distribution with parameters r = 5, $\lambda = 4$.
-
-a.) Use simulation (with $n = 200$) to generate an approximate sampling distribution of the mean; plot and describe the distribution.
-
-```{r}
-set.seed(123)
-
-n <- 200; r <- 5; lambda <- 4; mu <- lambda/r
-
-pop <- data.table(values = rgamma(n, shape = lambda, rate = r))
-
-ggplot(pop, aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange")
-```
-
-b.) Now, draw one random sample of size 200 from this population. Create a histogram of your sample, and find the mean and standard deviation.
-
-```{r}
-samp <- data.table(values = sample(pop$values, n, replace = T))
-
-ggplot(samp, aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange")
-
-xbar <- mean(samp$values); sd <- sd(samp$values)
-
-xbar; sd
-```
-
-c.) Compute the bootstrap distribution of the mean for you sample, plot it, and note the bootstrap mean and standard error.
-
-```{r}
-I <- 10e3
-
-boot.fn <- function(data, index) {
-   mean(data[index]$values)
-}
-
-boot(samp, boot.fn, R = I)
-
-n.200 <- cst.boot(samp$values, n)
-
-ggplot(data.table(values = n.200$bootstrap), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange") +
-   labs(title = "N=200, bootstrapped mean (Gamma r=5, lambda = 4)")
-```
-
-d.) Compare the bootstrap distribution to the approximate theoretical sampling distribution by creating a table like Table 5.2.
-
-```{r}
-tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
-                  Mean = c(mu, sd(samp$values), n.200$observed, n.200$mean),
-                  SD = c(mu, mu/sqrt(n), sd(samp$values), n.200$se))
-
-pretty_kable(tbl, "Sampling Statistics")
-```
-
-e.) Repeat (a-e) for sample sizes of n = 50, and n = 10. Describe carefully your observations about the effects of sample size on the bootstrap distribution.
-
-```{r}
-n <- 50
-
-pop <- data.table(values = rgamma(n, shape = lambda, rate = r))
-
-ggplot(pop, aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange")
-
-samp <- data.table(values = sample(pop$values, n, replace = T))
-
-ggplot(samp, aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange")
-
-xbar <- mean(samp$values); sd <- sd(samp$values)
-
-xbar; sd
-
-n.50 <- cst.boot(samp$values, n)
-
-ggplot(data.table(values = n.50$bootstrap), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange") +
-   labs(title = "N=50, bootstrapped mean (Gamma r=5, lambda = 4)")
-
-tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
-                  Mean = c(mu, sd(samp$values), n.50$observed, n.50$mean),
-                  SD = c(mu, mu/sqrt(n), sd(samp$values), n.50$se))
-
-pretty_kable(tbl, "Sampling Statistics")
-```
-
-```{r}
-n <- 10
-
-pop <- data.table(values = rgamma(n, shape = lambda, rate = r))
-
-ggplot(pop, aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange")
-
-samp <- data.table(values = sample(pop$values, n, replace = T))
-
-ggplot(samp, aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange")
-
-xbar <- mean(samp$values); sd <- sd(samp$values)
-
-xbar; sd
-
-n.50 <- cst.boot(samp$values, n)
-
-ggplot(data.table(values = n.50$bootstrap), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange") +
-   labs(title = "N=50, bootstrapped mean (Gamma r=5, lambda = 4)")
-
-tbl <- data.table(Data = c("Population", "Sampling Distribution", "Sample", "Bootstrap Distribution"),
-                  Mean = c(mu, sd(samp$values), n.50$observed, n.50$mean),
-                  SD = c(mu, mu/sqrt(n), sd(samp$values), n.50$se))
-
-pretty_kable(tbl, "Sampling Statistics")
-```
-
-### 5.10
-
-We investigate the bootstrap distribution of the median. Create random sample of size n for various n and bootstrap the median. Describe the bootstrap distribution.
-
-```{r}
-ne <- 14 # n even
-no <- 15 # n odd
-
-wwe <- rnorm(ne) # draw samples of size ne
-wwo <- rnorm(no) # draw random samples of size no
-
-N <- 10^4
-
-even.boot <- numeric(N) # save space
-odd.boot <- numeric(N)
-
-for(i in 1:N)
-{
-   x.even <- sample(wwe, ne, replace = T)
-   x.odd <- sample(wwo, no, replace = T)
-
-   even.boot[i] <- median(x.even)
-   odd.boot[i] <- median(x.odd)
-}
-
-boot <- data.table(even = even.boot, odd = odd.boot)
-
-p1 <- ggplot(boot, aes(even)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange") +
-   labs(title = "even")
-
-p2 <- ggplot(boot, aes(odd)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange") +
-   labs(title = "odd")
-
-gridExtra::grid.arrange(p1, p2)
-```
-
-Change the sample sizes to 36 and 37; 200 and 201; and 10,000 and 10,001.
-
-Note the similarities/dissimalarities, trends, and so on. Why does the parity of the sample size matter? (_Note: Adjust the x limits in the plots as needed._)
-
-```{r}
-ne <- 36 # n even
-no <- 37 # n odd
-
-wwe <- rnorm(ne) # draw samples of size ne
-wwo <- rnorm(no) # draw random samples of size no
-
-N <- 10^4
-
-even.boot <- numeric(N) # save space
-odd.boot <- numeric(N)
-
-for(i in 1:N)
-{
-   x.even <- sample(wwe, ne, replace = T)
-   x.odd <- sample(wwo, no, replace = T)
-
-   even.boot[i] <- median(x.even)
-   odd.boot[i] <- median(x.odd)
-}
-
-boot <- data.table(even = even.boot, odd = odd.boot)
-
-p1 <- ggplot(boot, aes(even)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange")
-
-p2 <- ggplot(boot, aes(odd)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange")
-
-gridExtra::grid.arrange(p1, p2)
-```
-
-```{r}
-ne <- 200 # n even
-no <- 201 # n odd
-
-wwe <- rnorm(ne) # draw samples of size ne
-wwo <- rnorm(no) # draw random samples of size no
-
-N <- 10^4
-
-even.boot <- numeric(N) # save space
-odd.boot <- numeric(N)
-
-for(i in 1:N)
-{
-   x.even <- sample(wwe, ne, replace = T)
-   x.odd <- sample(wwo, no, replace = T)
-
-   even.boot[i] <- median(x.even)
-   odd.boot[i] <- median(x.odd)
-}
-
-boot <- data.table(even = even.boot, odd = odd.boot)
-
-p1 <- ggplot(boot, aes(even)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange")
-
-p2 <- ggplot(boot, aes(odd)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange")
-
-gridExtra::grid.arrange(p1, p2)
-```
-
-```{r}
-ne <- 10000 # n even
-no <- 10001 # n odd
-
-wwe <- rnorm(ne) # draw samples of size ne
-wwo <- rnorm(no) # draw random samples of size no
-
-N <- 10^4
-
-even.boot <- numeric(N) # save space
-odd.boot <- numeric(N)
-
-for(i in 1:N)
-{
-   x.even <- sample(wwe, ne, replace = T)
-   x.odd <- sample(wwo, no, replace = T)
-
-   even.boot[i] <- median(x.even)
-   odd.boot[i] <- median(x.odd)
-}
-
-boot <- data.table(even = even.boot, odd = odd.boot)
-
-p1 <- ggplot(boot, aes(even)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange")
-
-p2 <- ggplot(boot, aes(odd)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(even), col = "darkorange")
-
-gridExtra::grid.arrange(p1, p2)
-```
-
-For odd n, median will be one of the sample points. For smaller n, there will be only n possible values for the median, so the sampling distribution is more "granular" than when n is even.
-
-### 5.11
-
-Import the data from data set Bangladesh. In addition to aresnic concentrations for 271 wells, the data set contains cobolt and chlorine concentrations.
-
-a.) Conduct EDA on the chlorine concentrations and describe the salient features.
-
-```{r}
-Bangladesh <- data.table(read.csv(paste0(data.dir, "Bangladesh.csv"),
-                                 header = T))
-head(Bangladesh)
-
-ggplot(Bangladesh, aes(Chlorine)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(Chlorine), col = "darkorange")
-
-GGally::ggpairs(Bangladesh)
-```
-
-b.) Bootstrap the mean.
-
-```{r}
-N <- 10e3
-
-boot.fn <- function(data, index){
-   mean(data[index], na.rm = T)
-}
-
-boot(Bangladesh$Chlorine, boot.fn, R = N)
-
-observed <- mean(Bangladesh$Chlorine, na.rm = T)
-
-bootstrap <- numeric(N)
-
-for(i in 1:N)
-{
-   bootstrap[i] <- mean(sample(Bangladesh$Chlorine, size = nrow(Bangladesh), replace = T), na.rm = T)   
-}
-
-ggplot(data.table(values = bootstrap), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(aes(values), col = "darkorange")
-```
-
-c.) Find and interpret the 95% bootstrap percentile confidence interval.
-
-```{r}
-alpha <- 0.05
-quantile(bootstrap, c(alpha/2, 1 - alpha/2))
-```
-
-The spread on the confidence interval is extremely large, which is unsuprising given the heavly skewed distribtion of the sample.
-
-d.) What is the bootstrap estimate of the bias? What fraction of the bootstrap standard error does it represent?
-
-```{r}
-bias <-  mean(bootstrap) -observed
-
-bias / sd(bootstrap)
-```
-
-roughly 1% of the standard error.
-
-### 5.12
-
-Consider Bangladesh chlorine (concentration). Bootstrap the trimmed mean (say, trim the upper and lower 25%), and compare your results with the usual mean (previous result).
-
-```{r}
-values <- Bangladesh[!is.na(Chlorine)]$Chlorine
-
-n <- length(values); N <- 10e3; trim <- .25
-
-observed <- mean(values, trim = trim)
-bootstrap <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   bootstrap[i] <- mean( sample(values, n, replace = T), trim = trim )
-}
-
-ggplot(data.table(value = bootstrap), aes(value)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange")
-
-alpha <- 0.05
-
-boot.mean <- mean(bootstrap)
-boot.bias <- boot.mean - observed
-boot.se <- sd(bootstrap)
-
-quantile(bootstrap, c(alpha/2, 1 - alpha/2))
-
-# boot pkg
-
-boot.fn <- function(data, index){
-   mean( data[index], trim = trim)
-}
-
-boot(values, boot.fn, R = N)
-```
-
-The bootstrap 20% trimmed mean is substantially smaller than the regular mean. The confidence intervals are also tighter, which is unsuprising due to the heavy presence of outliers in our sample.
-
-### 5.13
-
-The data set _FishMercury_ contains mercury levels (parts per million) for 30 fish caught in lakes in Minnesota.
-
-```{r}
-FishMercury <- data.table(read.csv(paste0(data.dir, "FishMercury.csv"),
-                                 header = T))
-```
-
-a.) Create a histogram or boxplot of the data. What do you observe?
-
-```{r}
-ggplot(FishMercury, aes(Mercury)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_rug(col = "darkred")
-
-FishMercury[Mercury > 1.5]
-
-ggplot(FishMercury[Mercury < 1.5], aes(Mercury)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_rug(col = "darkred")
-
-FishMercury[Mercury > 1.5] / sd(FishMercury$Mercury) # 6 sd outlier
-
-```
-
-One extreme (6 SD) outlier in the data.
-
-b.) Bootstrap the mean, and record the bootstrap standard error and the 95% bootstrap percentile interval.
-
-```{r}
-n <- nrow(FishMercury); N <- 10e3
-
-observed <- mean(FishMercury$Mercury)
-
-bootstrap <- vector(mode = "numeric", length = n)
-
-for(i in 1:N)
-{
-   bootstrap[i] <- mean( sample(FishMercury$Mercury, n, replace = T) )
-}
-
-boot.mean <- mean(bootstrap)
-boot.bias <- boot.mean - observed
-boot.se <- sd(bootstrap)
-
-alpha <- 0.05
-
-quantile(bootstrap, c(alpha/2, 1 - alpha/2))
-
-# boot pkg
-
-boot.fn <- function(data, index) {
-   mean( data[index] )
-}
-
-boot(FishMercury$Mercury, boot.fn, R = N)
-```
-
-c.) Remove the outlier and bootstrap the mean of the remaining data. Record the bootstrap standard error and the 95% bootstrap percentile interval.
-
-```{r}
-mercury <- FishMercury[Mercury < 1.5]$Mercury
-
-n <- length(mercury)
-bootstrap.values <- vector(mode = "numeric", length = n)
-
-observed <- mean(mercury)
-
-for( i in 1:N)
-{
-   bootstrap.values[i] <- mean( sample(mercury, size = n, replace = T) )   
-}
-
-mean(bootstrap.values)
-
-boot(mercury, boot.fn, R = N)
-
-quantile(bootstrap.values, c(alpha/2, 1 - alpha/2))
-```
-
-The standard error reduced drastically, and the confidence intervals are much more narrow.
-
-### 5.14
-
-In Section 3.3, we performed a permutation test to determine if men and women consumed, on average, different amounts of hot wings.
-
-```{r}
-BeerWings <- data.table(read.csv(paste0(data.dir, "Beerwings.csv"),
-                                 header = T))
-
-wings <- BeerWings$Hotwings
-
-?stat_ecdf
-
-ggplot(BeerWings, aes(Hotwings, group = Gender, col = Gender)) +
-   stat_ecdf(geom = "step")
-
-N <- 10e3; alpha <- 0.05; n <- length(wings)
-
-wings.m <- BeerWings[Gender == "M"]$Hotwings
-wings.f <- BeerWings[Gender == "F"]$Hotwings
-
-observed <- mean(wings.m) - mean(wings.f)
-
-permutation.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   samp <- sample(1:n, length(wings.m), replace = F)
-
-   samp.m <- wings[samp]; samp.w <- wings[-samp]
-
-   permutation.values[i] <- mean(samp.m) - mean(samp.w)
-}
-
-ggplot(data.table(values = permutation.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed, col = "darkred")
-
-ggplot(data.table(value = permutation.values), aes(value)) +
-   stat_ecdf(geom = "step") +
-   geom_vline(xintercept = observed, col = "darkred") +
-   scale_y_continuous(labels = comma)
-
-p <- (sum(permutation.values >= observed) + 1) / (N + 1)
-
-var <- p*(1 - p)/N
-```
-
-a.) Bootstrap the difference in means and describe the bootstrap distribution.
-
-```{r}
-bootstrap.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   samp.m <- sample(wings.m, size = length(wings.m), replace = T)
-   samp.f <- sample(wings.f, size = length(wings.f), replace = T)
-
-   bootstrap.values[i] <- mean(samp.m) - mean(samp.f)
-}
-
-ggplot(data.table(values = bootstrap.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30)
-
-boot.mean <- mean(bootstrap.values)
-boot.se <- sd(bootstrap.values)
-
-boot.bias <- boot.mean - observed
-
-boot.se; boot.bias
-```
-
-b.) Find a 95% bootstrap percentile confidence interval for the difference of means, and give a sentence interpreting this interval.
-
-```{r}
-quantile(bootstrap.values, c(alpha/2, 1 - alpha/2))
-```
-
-The 95% confidence interval does not contain 0, further supporting our permutation test results.
-
-c.) How do the bootstrap and permutation distributions differ?
-
-The permutation distribution is sampled without replacement, and the bootstrap distribution is sampled using replacement.
-
-### 5.15
-
-A high school student was curious about the total number of minutes devoted to commercials during any given half-hour time period on basic and extended cable TV channels.
-(B.
-Rodgers and T.
-Robinson, personal communication).
-
-Import the TV data.
-
-```{r}
-TV <- data.table(read.csv(paste0(data.dir, "TV.csv"),
-                                 header = T))
-```
-
-a.) Perform some exploratory data analysis and obtain summary statistics on the commercial times on basic and extended cable TV channels (do seperate analysis for each type of channel).
-
-```{r}
-ggplot(TV, aes(Cable, Times, fill = Cable)) +
-   geom_boxplot()
-
-ggplot(TV, aes(Times, group = Cable)) +
-   geom_density(aes(y = ..density.., fill = Cable), alpha = .45)
-
-aov(TV$Times ~ TV$Cable)
-
-tv.basic <- TV[Cable == "Basic"]$Times
-tv.extended <- TV[Cable == "Extended"]$Times
-tv.pooled <- c(tv.basic, tv.extended)
-
-observed <- mean(tv.basic) - mean(tv.extended)
-
-mean(tv.basic); mean(tv.extended)
-```
-
-b.) Bootstrap the difference in mean times, plot the distribution, and give summary statistics of the bootstrap distribution.
-Obtain a 95% bootstrap percentile confidence interval, and interpret this interval.
-
-```{r}
-N <- 10e3; n <- length(TV$Times); alpha <- 0.05
-
-bootstrap.values <- vector(mode = "numeric", N)
-
-for(i in 1:N)
-{
-   samp.basic <- sample(tv.basic, size = length(tv.basic), replace = T)
-   samp.extended <- sample(tv.extended, size = length(tv.extended), replace = T)
-
-   bootstrap.values[i] <- mean(samp.basic) - mean(samp.extended)
-}
-
-boot.mean <- mean(bootstrap.values)
-boot.se <- sd(bootstrap.values)
-boot.bias <- boot.mean - observed
-
-quantile(bootstrap.values, c(alpha/2, 1 - alpha/2)) # bootstrap interval
-```
-
-The 95% boostrap interval does not contain 0, suggesting there is in fact a difference in the times between basic and extended (basic being longer).
-
-c.) What is the bootstrap estimate of the bias?
-
-```{r}
-boot.bias
-
-boot.bias / boot.se # about 1% of the standard error
-```
-
-d.) Conduct a permutation test to see if the difference in mean commercial times is statistically significant, and state your conclusion.
-
-```{r}
-permutation.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   samp <- sample(length(tv.pooled), length(tv.basic), replace = F)
-
-   samp.basic <- tv.pooled[samp]; samp.extended <- tv.pooled[-samp]
-
-   permutation.values[i] <- mean(samp.basic) - mean(samp.extended)
-}
-
-ggplot(data.table(values = permutation.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_vline(xintercept = observed, col = "darkred")
-
-p <- ( sum(permutation.values >= observed) + 1) / ( N + 1)
-var <- p*(1 - p)/N
-```
-
-Permutation test supports the alternative hypothesis that there is a difference in commercial times.
-
-### 5.16
-
-Researchers conducted a study of primary and early secondary school children in Italy to examine gender differences in math anxiety.
-One of the measures used to understand math anxiety is the _Abbreviated Math Anxiety Scale (AMAS)_, a self-reported math anxiety questionnaire.
-A higher score indicates more max anxiety.
-The data set MathAnxiety contains the results for a subset of the children in the original study.
-
-```{r}
-MathAnxiety <- data.table(read.csv(paste0(data.dir, "MathAnxiety.csv"),
-                                 header = T))
-```
-
-a.) Perform some exploratory analysis and obtain summary statistics of the AMAS scores for the boys and girls (do seperate analysis for each gender).
-
-```{r}
-ggplot(MathAnxiety, aes(Grade, AMAS)) +
-   geom_boxplot(aes(fill = Grade))
-
-ggplot(MathAnxiety, aes(Age, AMAS)) +
-   geom_point()
-
-ggplot(MathAnxiety, aes(RCMAS, AMAS)) +
-   geom_point(aes(color = Gender))
-
-ggplot(MathAnxiety, aes(Gender, AMAS)) +
-   geom_boxplot(aes(fill = Gender))
-
-ggpairs(MathAnxiety)
-```
-
-b.) Bootstrap the difference in mean times, plot the distribution, and give summary statistics of the bootstrap distribution. Obtain a 95% confidence interval, and interperet this interval.
-
-```{r}
-amas_boy <- MathAnxiety[Gender == "Boy"]$AMAS
-amas_girl <- MathAnxiety[Gender == "Girl"]$AMAS
-
-N <- 10e3; n <- nrow(MathAnxiety); alpha <- 0.05
-
-observed <- mean(amas_boy) - mean(amas_girl)
-
-bootstrap.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   samp_boy <- sample(amas_boy, length(amas_boy), replace = T)
-   samp_girl <- sample(amas_girl, length(amas_girl), replace = T)
-
-   bootstrap.values[i] <- mean(samp_boy) - mean(samp_girl)
-}
-
-boot.mean <- mean(bootstrap.values)
-boot.bias <- boot.mean - observed
-boot.se <- sd(bootstrap.values)
-
-ggplot(data.table(values = bootstrap.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(color = "darkorange") +
-   geom_vline(xintercept = observed, col = "darkred")
-
-boot.bias; boot.se
-
-quantile(bootstrap.values, c(alpha/2, 1 - alpha/2)) # conf interval
-```
-
-The 95% confidence interval is -2.8 ~ -0.7, which does not include zero.
-The data suggests that there is a statistical difference between the self-reported AMAS scores between boys and girls (boys being lower).
-
-c.) What is the bootstrap estimate of the bias? What fraction of the bootstrap; standard error does this represent?
-
-```{r}
-boot.bias
-
-boot.bias / boot.se
-```
-
-Bootstrap bias is .006, which is less than 1% of the se.
-
-d.) Conduct a permutation test to see if the difference in mean AMAS scores is statistically significant, and state your conclustion.
-
-```{r}
-amas_pooled <- c(amas_boy, amas_girl)
-
-n <- length(amas_pooled); observed <- mean(amas_boy) - mean(amas_girl)
-
-permutation.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   samp <- sample(1:n, length(samp_boy), replace = F)
-
-   samp_boys <- amas_pooled[samp]; samp_girl <- amas_pooled[-samp]
-
-   permutation.values[i] <- mean(samp_boy) - mean(samp_girl)
-}
-
-mean(permutation.values)
-
-ggplot(data.table(value = permutation.values), aes(value)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed)
-
-p <- ( sum(permutation.values <= observed) + 1 ) / (N + 1)
-var <- p*(1 - p)/ N
-
-t.test(amas_boy, amas_girl, alternative = "less")
-```
-
-The p-value for the permutation test is less than 1%, we reject the null hypothesis that there is no difference in self-reported AMAS scores between boys and girls.
-This conclustion is in-line with our results from the bootstraped difference in mean test.
-
-### 5.17
-
-Import the data from Girls2004.
-
-```{r}
-Girls2004 <- data.table(read.csv(paste0(data.dir, "Girls2004.csv"),
-                                 header = T))
-```
-
-a.) Perform some exploratory data analysis, and obtain summary statistics on the weights of baby girls born in Wyoming and Alaska.
-
-```{r}
-ggpairs(Girls2004)
-
-ggplot(Girls2004, aes(Smoker, Weight, fill = State)) +
-   geom_boxplot() +
-   scale_y_continuous(labels = scales::comma)
-
-ggplot(Girls2004, aes(MothersAge, Weight, fill = State)) +
-   geom_boxplot()
-```
-
-b.) Bootstrap the difference in means, plot the distribution, and give the summary statistics. Obtain a 95% confidence boostrap percentile confidence interval and interpret this interval.
-
-```{r}
-weights_wy <- Girls2004[State == "WY"]$Weight
-weights_ak <- Girls2004[State == "AK"]$Weight
-
-observed <- mean(weights_wy) - mean(weights_ak); N <- 10e3
-alpha <- 0.05
-
-bootstrap.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   samp_wy <- sample(weights_wy, size = length(weights_wy), replace = T)
-   samp_ak <- sample(weights_ak, size = length(weights_ak), replace = T)
-
-   bootstrap.values[i] <- mean(samp_wy) - mean(samp_ak)
-}
-
-bootstrap.mean <- mean(bootstrap.values)
-bootstrap.bias <- bootstrap.mean - observed
-bootstrap.se <- sd(bootstrap.values)
-
-ggplot(data.table(value = bootstrap.values), aes(value)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed, col = "darkred")
-
-quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1- alpha/2))
-```
-
-There is a statistically significant difference in the mean birth weights of girls born in Wyoming and Alaska (Wyoming girls weighing less on average), by between 530 and 96 oz less.
-
-c.) What is the bootstrap estimate of the bias? What fraction of thh boostrap standard error does it represent?
-
-```{r}
-bootstrap.bias
-
-bootstrap.bias / bootstrap.se
-```
-
-d.) Conduct a permutation test to see if the difference in mean weights is statistically significant, and state your conclusion.
-
-```{r}
-observed <- mean(weights_wy) - mean(weights_ak)
-weights_pooled <- c(weights_wy, weights_ak)
-
-permutation.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   samp <- sample(length(weights_pooled), size = length(weights_ak), replace = F)
-
-   samp_wy <- weights_pooled[samp]; samp_ak <- weights_pooled[-samp]
-
-   permutation.values[i] <- mean(samp_wy) - mean(samp_ak)
-}
-
-mean(permutation.values)
-
-ggplot(data.table(values = permutation.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed, col = "darkred")
-
-p <- (sum(permutation.values <= observed) + 1) / (N + 1)
-
-var <- p*(1 - p)/N
-```
-
-The permutation test further supports the claim that the weights of baby girls born in Wyoming and Alaska are different.
-We reject the null hypothesis that the means are the same at the 0.05% level (p = 0.003).
-
-e.) For what population(s), if any, does this conclusion hold?
-
-_I do not
-### 5.18
-
-Is there a difference in the price of groceries sold by the two retailers Target and Walmart?
-The data set _Groceries_ contain, a sample of grocery items and their prices advertised on their respective websites on one specific day.
-
-```{r}
-Groceries <- data.table(read.csv(paste0(data.dir, "Groceries.csv"),
-                                 header = T))
-```
-
-a.) Compute summary statistics of the prices for each store.
-
-```{r, fig.height=7}
-groceries <- melt(Groceries %>% select(-Size), 
-                  id.vars = c("Product"), 
-                  value.name = "Price", 
-                  variable.name = "Store")
-
-setorder(groceries, -Product)
-
-ggplot(groceries, aes(Product, Price, col = Store, group = Store)) +
-   geom_point(aes(color = Store)) +
-   coord_flip()
-```
-
-b.) Use the bootstrap to determine whether or not there is a difference in the mean prices.
-
-```{r}
-prices.target <- groceries[Store == "Target"]$Price
-prices.walmart <- groceries[Store == "Walmart"]$Price
-
-N <- 10e3
-
-observed <- mean(prices.target) - mean(prices.walmart)
-bootstrap.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   index <- sample(length(prices.target), size = length(prices.target), replace = T)
-
-   samp.target <- prices.target[index]
-   samp.walmart <- prices.walmart[index]
-
-   bootstrap.values[i] <- mean(samp.target) - mean(samp.walmart)
-}
-
-boot.mean <- mean(bootstrap.values)
-boot.bias <- boot.mean - observed
-boot.se <- sd(bootstrap.values)
-
-boot.bias / boot.se
-
-ggplot(data.table(values = bootstrap.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed, col = "darkred")
-
-quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1 - alpha/2))
-```
-
-There doesn't appear to be a statistically significant difference in the average prices at the stores.
-
-c.) Create a histogram of the difference in prices. What is unusal about Quaker Oats Life cereal?
-
-```{r}
-Groceries[Product == "Quaker Oats Life Cereal  Original ",]
-```
-
-It's almost double the price.
-
-d.) Recompute the bootstrap percentile interval without this observation. What do you conclude?
-
-```{r}
-groceries_sans_quaker <- groceries[Product != "Quaker Oats Life Cereal  Original ",]
-
-stopifnot(nrow(groceries_sans_quaker) == nrow(groceries) - 2) # ensure removeed
-
-prices.target <- groceries_sans_quaker[Store == "Target"]$Price
-prices.walmart <- groceries_sans_quaker[Store == "Walmart"]$Price
-
-N <- 10e3
-
-observed <- mean(prices.target) - mean(prices.walmart)
-bootstrap.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   # paired data, must use the same products from each company
-   index <- sample(length(prices.target), size = length(prices.target), replace = T)
-
-   samp.target <- prices.target[index]
-   samp.walmart <- prices.walmart[index]
-
-   bootstrap.values[i] <- mean(samp.target) - mean(samp.walmart)
-}
-
-boot.mean <- mean(bootstrap.values)
-boot.bias <- boot.mean - observed
-boot.se <- sd(bootstrap.values)
-
-boot.bias / boot.se
-ggplot(data.table(values = bootstrap.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed, col = "darkred")
-
-quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1 - alpha/2))
-```
-
-The bootstrap confidence intervals suggest that there is a statistically significant difference in prices (Target being more expensive than Walmart). The Quaker Oats cereal made a difference.
-
-### 5.19
-
-Do chocolate and vanilla ice cream have the same number of calories?
-
-The data set Ice Cream contains calorie information for a sample of brands of chocolate and vanilla ice cream.
-
-```{r}
-IceCream <- data.table(read.csv(paste0(data.dir, "IceCream.csv"),
-                                 header = T))
-```
-
-a.) Compute summary statistics of the calories for the two flavors.
-
-```{r}
-
-icecream <- IceCream %>%
-   mutate(Vanilla = VanillaCalories,
-          Chocolate = ChocolateCalories) %>%
-   select(Vanilla, Chocolate) %>%
-   melt(variable.name = "Flavor", value.name = "Calories") %>%
-   as.data.table()
-
-stopifnot(nrow(icecream) == nrow(IceCream) * 2)
-
-ggplot(icecream, aes(Calories, group = Flavor)) +
-   geom_density(aes(fill = Flavor), alpha = .35)
-
-ggplot(icecream, aes(Flavor, Calories, fill = Flavor)) +
-   geom_boxplot()
-```
-
-b.) Use the bootstrap to determine whether or not there is a difference in the mean number of calories.
-
-```{r}
-calories.vanilla <- icecream[Flavor == "Vanilla"]$Calories
-calories.chocolate <- icecream[Flavor == "Chocolate"]$Calories
-
-N <- 10e3
-
-observed <- mean(calories.chocolate) - mean(calories.vanilla)
-
-bootstrap.values <- vector(mode = "numeric", length = N)
-
-for(i in 1:N)
-{
-   # paired data, must use the same samples from chocolate and vanilla (ie, same manufacture)
-   index <- sample(length(calories.vanilla), length(calories.vanilla), replace = T)
-
-   v.samp <- calories.vanilla[index]
-   c.samp <- calories.chocolate[index]
-
-   bootstrap.values[i] <- mean(c.samp) - mean(v.samp)
-}
-
-ggplot(data.table(values = bootstrap.values), aes(values)) +
-   geom_histogram(aes(y = ..density.., fill = ..count..), bins = 30) +
-   geom_density(col = "darkorange") +
-   geom_vline(xintercept = observed)
-
-bootstrap.mean <- mean(bootstrap.values)
-bootstrap.bias <- bootstrap.mean - observed
-bootstrap.se <- sd(bootstrap.values)
-
-bootstrap.bias / bootstrap.se
-
-quantile(bootstrap.values, c(Lower = alpha/2, Upper = 1 - alpha/2))
-```
-
-There appers to be a statistically difference in the amount of calories (Chocolate having more, by approx 4-12 on average).
-
-### 5.20
-
-In a remark at the end of Section 5.4.1, we mentioned that the procedure for bootstrapping the difference of medians is different than for thee mean.
-
-Import the data set Diving2017.
-
-```{r}
-
-```
+#### 6.1

### 2020-02-11 Mathematical Statistics/05_conf_int_bootstrap.r

@@ -91,7 +91,7 @@ arsenic.boot <- boot(Arsenic, boot.fn, R = N)
 ### NC birth weights
 
 NCBirths <- data.table(read.csv(paste0(data.dir, "NCBirths2004.csv"),
-                                  header = T))
+                                header = T))
 
 weights <- NCBirths$Weight
 

### 2020-02-11 Mathematical Statistics/06_Estimation.R

@@ -1,0 +1,15 @@
+# Suppose X1, X2, ... Xn are a random sample from the Cauchy distribution with pdf f(x;theta) = 1/(pi(1 + (x - theta)^2))
+
+x <- c(1, 2, 2, 3)
+
+g <- function(theta) sum(log( 1 + (x-theta)^2))
+
+obj <- optimize(g, interval = c(0, 4))
+
+obj$minimum
+
+obj$objective
+
+logL <- function(theta) sum(log(dcauchy(x, theta)))
+
+optimize(logL, interval = c(0, 4), maximum = T)

### 2020-02-19 HOML/05_Logistic-Regression.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Logistic Regression}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 5}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init()

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Logistic Regression

### Data Set

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)
```
### Overview

Linear Regression vs Logistic Regression

```{r, fig.height=3}
p1 <- ISLR::Default %>%
   mutate(prob = ifelse(default == "Yes", 1, 0)) %>%
   ggplot(aes(balance, prob)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "lm") +
   ggtitle("Linear Regression model fit") +
   ylab("Balance") + xlab("Probability of Default")

p2 <- ISLR::Default %>%
   mutate(prob = ifelse(default == "Yes", 1, 0)) %>%
   ggplot(aes(balance, prob)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial")) +
   ggtitle("Logistic regression model fit") +
   xlab("Balance") + ylab("Probability of Default")

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Create training (70%) and test (30%) tests.

```{r, echo = T}
set.seed(123)

churn_split <- initial_split(attrition, 
                             prop = 0.7, strata = "Attrition")

churn_train <- training(churn_split)
churn_test <- testing(churn_split)
```

### Simple Logistic Regression

Fit two generalized linear models to predict attrition.

```{r}
model1 <- glm(Attrition ~ MonthlyIncome, family = "binomial",
              data = churn_train)

model2 <- glm(Attrition ~ OverTime, family = "binomial",
              data = churn_train)

```

```{r}
churn_train2 <- churn_train %>% mutate(prob = ifelse(Attrition == "Yes", 1, 0))
churn_train2 <- broom::augment(model2, churn_train2) %>% mutate(.fitted = exp(.fitted))

p1 <- ggplot(churn_train2, aes(MonthlyIncome, prob)) +
   geom_point(alpha = 0.15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial")) +
   ggtitle("Predicted probabilities for model1") +
   xlab("Monthly Income") +
   ylab("Probability of Attrition")

p2 <- ggplot(churn_train2, aes(OverTime, .fitted, color = OverTime)) +
   geom_boxplot(show.legend = F) +
   geom_rug(sides = "b", position = "jitter", alpha = 0.2, show.legend = F) +
   ggtitle("Predicted probabilities for model2") +
   xlab("Over Time") +
   scale_y_continuous("Probability of Attrition", limits = c(0, 1))

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Model Diagnostics

```{r}
tidy(model1)

tidy(model2)
```

```{r}
exp(coef(model1))

exp(coef(model2))
```

```{r}
confint(model1)

confint(model2)
```

### Multiple Logistic Regression

```{r}
model3 <- glm(
   Attrition ~ MonthlyIncome + OverTime,
   family = "binomial",
   data = churn_train
)

tidy(model3)
```

```{r}
churn_train3 <- churn_train %>% mutate(prob = ifelse(Attrition == "Yes", 1, 0))
churn_train3 <- broom::augment(model3, churn_train3) %>% mutate(.fitted = exp(.fitted))

ggplot(churn_train3, aes(MonthlyIncome, prob, color = OverTime)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial"), se = F) +
   ggtitle("Predicted probabilites for model3") +
   xlab("Monthly Income") +
   ylab("Probability of Attrition")
```

### Assessing Model Accuracy

```{r}
set.seed(123)

cv_model1 <- train(
   Attrition ~ MonthlyIncome,
   data = churn_train,
   method = "glm",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

set.seed(123)

cv_model2 <- train(
   Attrition ~ MonthlyIncome + OverTime,
   data = churn_train,
   method = "glm",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

set.seed(123)

cv_model3 <- train(
   Attrition ~ .,
   data = churn_train,
   method = "glm",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

# extract out of sample performance measurse

summary(
   resamples(
      list(
         model_1 = cv_model1,
         model_2 = cv_model2,
         model_3 = cv_model3
      )
   )
)$statistics$Accuracy
```

```{r}
# predicted class
pred_class <- predict(cv_model3, churn_train)

# create confusion matrix
confusionMatrix(
   data = relevel(pred_class, ref = "Yes"),
   reference = relevel(churn_train$Attrition, ref = "Yes")
)
```

No-information rate

```{r}
table(churn_train$Attrition) %>% prop.table()
```

Basically, this is saying if we just predicted "No" for every instance we would have 83.8% accuracy.

```{r}
# Compute predicted probabilities
m1_prob <- predict(cv_model1, churn_train, type = "prob")$Yes
m3_prob <- predict(cv_model3, churn_train, type ="prob")$Yes

# Compute AUC metrics for cv_model1 and cv_model3
perf1 <- prediction(m1_prob, churn_train$Attrition) %>%
   performance(measure = "tpr", x.measure = "fpr")

perf2 <- prediction(m3_prob, churn_train$Attrition) %>%
   performance(measure = "tpr", x.measure = "fpr")

# Plot ROC curves for cv_model1 and cv_model3
plot(perf1, col = "black", lty = 2)
plot(perf2, col = "blue", add = T)

legend(0.8, 0.2, legend = c("cv_model1", "cv_model3"),
       col = c("black", "blue"), lty = 2:1, cex = 0.6)
```

```{r}
set.seed(123)
cv_model_pls <- train(
   Attrition ~ .,
   data = churn_train,
   method = "pls",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10),
   preProcess = c("zv", "center", "scale"),
   tuneLength = 16
)

cv_model_pls$bestTune

ggplot(cv_model_pls)
```

### Feature Interpretation

```{r}
vip::vip(cv_model3, num_features = 20)
```

```{r}
pred.fun <- function(object, newdata) {
   Yes <- mean(predict(object, newdata, type = "prob")$Yes)
   as.data.frame(Yes)
}

p1 <- pdp::partial(cv_model3, pred.var = "OverTime", pred.fun = pred.fun ) %>%
   autoplot(rug = T) + ylim(c(0, 1))

p2 <- pdp::partial(cv_model3, pred.var = "JobSatisfaction", pred.fun = pred.fun) %>%
   autoplot() + ylim(c(0, 1))

p3 <- pdp::partial(cv_model3, pred.var = "NumCompaniesWorked", pred.fun = pred.fun, gr = 10) %>%
   autoplot() + scale_x_continuous(breaks = 0:9) + ylim(c(0, 1))

p4 <- pdp::partial(cv_model3, pred.var = "EnvironmentSatisfaction", pred.fun = pred.fun) %>%
   autoplot() + ylim(c(0, 1))

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-02-19 HOML/07_Multivariate-Adaptive-Regression-Splines.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Multivariate Adaptive Regression Splines}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 7}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(mda, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Multivariate Adaptive Regression Splines

### Data Sets

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)
```

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

### Overview

Linear models assume the underlying phenomena we are modeling is intrinsically linear which is not usually true. Multivariate adaptive regression splines (MARS) allow us to model non-linear relationships.

Basic strategies for modeling non-linear fits include polynomial regression and step-wise models.

Visually:

```{r}
set.seed(123)

x <- seq(from = 0, to = 2 * pi, length = 500)
y <- sin(x) + rnorm(length(x), sd = .3)

df <- data.table(x, y) %>%
   filter(x < 6)

p1 <- ggplot(df, aes(x, y)) +
   geom_point(alpha = .25) +
   geom_smooth(method = "lm", se = F) +
   ggtitle("(A) Assumed Linear Relationship")

p2 <- ggplot(df, aes(x, y)) +
   geom_point(alpha = .25) +
   geom_smooth(method = "lm", se = F, formula = y ~ poly(x, 2, raw = T)) +
   ggtitle("(B) Degree-2 Polynomial Regression")

p3 <- ggplot(df, aes(x, y)) +
   geom_point(alpha = .25) +
   geom_smooth(method = "lm", se = F, formula = y ~ poly(x, 3, raw = T)) +
   ggtitle("(C) Degree-3 Polynomial Regression")

# fit step function model (6 steps)

step_fit <- lm(y ~ cut(x, 5), data = df)
step_pred <- predict(step_fit, df)

p4 <- ggplot(df, aes(x, y)) +
   geom_point(alpha = .25) +
   geom_line(aes(y = step_pred), size = 1, color = "blue") +
   ggtitle("(D) Step Function Regression")

gridExtra::grid.arrange(p1, p2, p3, p4, nrow = 2)
```

### Multivariate Adaptive Regression Splines (MARS)

Similar to the step-wise approach, we can access "knots" in the data to model this behavior.

MARS models with _mda_ package.

```{r}
mars1 <- mda::mars(df$x, df$y, nk = 3, prune = F)

p1 <- df %>%
   mutate(predicted = as.vector(mars1$fitted.values)) %>%
   ggplot(aes(x, y)) +
   geom_point(size = 1, alpha = .2) +
   geom_line(aes(y = predicted), size = 1, color = "blue") +
   ggtitle("(A) One Knot")

mars2 <- mda::mars(df$x, df$y, nk = 5, prune = F)

p2 <- df %>%
   mutate(predicted = as.vector(mars2$fitted.values)) %>%
   ggplot(aes(x, y)) +
   geom_point(size = 1, alpha = .2) +
   geom_line(aes(y = predicted), size = 1, color = "blue") +
   ggtitle("(B) Two Knots")

mars3 <- mda::mars(df$x, df$y, nk = 7, prune = F)

p3 <- df %>%
   mutate(predicted = as.vector(mars3$fitted.values)) %>%
   ggplot(aes(x, y)) +
   geom_point(size = 1, alpha = .2) +
   geom_line(aes(y = predicted), size = 1, color = "blue") +
   ggtitle("(C) Three Knots")

mars4 <- mda::mars(df$x, df$y, nk = 9, prune = F)

p4 <- df %>%
   mutate(predicted = as.vector(mars4$fitted.values)) %>%
      ggplot(aes(x, y)) +
   geom_point(size = 1, alpha = .2) +
   geom_line(aes(y = predicted), size = 1, color = "blue") +
   ggtitle("(D) Four Knots")

gridExtra::grid.arrange(p1, p2, p3, p4, nrow = 2)
```

### Fitting a basic MARS model

```{r}
mars1 <- earth(
   Sale_Price ~ .,
   data = ames.train
)

print(mars1)
```

```{r}
summary(mars1) %>% .$coefficients %>% head(10)
```

```{r}
plot(mars1, which = 1)
plot(mars1, which = 2)
```

```{r}
mars2 <- earth(
   Sale_Price ~.,
   data = ames.train,
   degree = 2
)

summary(mars2) %>% .$coefficients %>% head(10)
```

```{r}
plot(mars2, which = 1)
```

### Tuning

As always, we will use a cross-validated grid search procedure to tune the hyperparameters.

First pass:

```{r}
hyper.grid <- expand.grid(
   degree = 1:3,
   nprune = seq(2, 100, length.out = 10) %>% floor()
)

head(hyper.grid)
```

```{r}
set.seed(123)

suppressWarnings(print({
cv.mars <- train(
   x = ames.train %>% select(-Sale_Price),
   y = ames.train$Sale_Price,
   method = "earth",
   metric = "RMSE",
   trControl = trainControl(method = "cv", number = 10),
   tuneGrid = hyper.grid
)}))

cv.mars$bestTune

cv.mars$results %>% 
   as_tibble() %>% 
   arrange(RMSE)

```

```{r}
ggplot(cv.mars) +
   scale_y_continuous(labels = scales::comma)
```

Refinement:

```{r}
refine.grid <- expand.grid(
   degree = 1,
   nprune = seq(from = 30, to = 45)
)

suppressWarnings(print({
   cv.mars2 <- train(
      x = ames.train %>% select(-Sale_Price),
      y = ames.train$Sale_Price,
      method = "earth",
      metric = "RMSE",
      trControl = trainControl(method = "cv", number = 10),
      tuneGrid = refine.grid
   )
}))

cv.mars2$bestTune

ggplot(cv.mars2) +
   scale_y_continuous(labels = scales::comma)

```

### Feature Interpretation

Variable Importance Plots

```{r, fig.height=6}
p1 <- vip::vip(cv.mars, num_features = 40, bar = F, value = "gcv") +
   ggtitle("GCV")

p2 <- vip::vip(cv.mars2, num_features = 40, bar = F, value = "rss") +
   ggtitle("RSS")

gridExtra::grid.arrange(p1, p2, nrow = 2)
```

```{r, fig.height=6}
p1 <- vip::vip(cv.mars2, num_features = 40, bar = F, value = "gcv") +
   ggtitle("GCV")

p2 <- vip::vip(cv.mars2, num_features = 40, bar = F, value = "rss") +
   ggtitle("RSS")

gridExtra::grid.arrange(p1, p2, nrow = 2)
```

```{r}
# extract coefficients, covert to tidy & filter for interaction

cv.mars2$finalModel %>%
   coef() %>%
   broom::tidy()
```

```{r}
# Construct partial dependence plots
p1 <- pdp::partial(cv.mars2, pred.var = "Gr_Liv_Area", grid.resolution = 10) %>% 
  autoplot() +
   scale_y_continuous(labels = scales::comma)

p2 <- pdp::partial(cv.mars2, pred.var = "Year_Built", grid.resolution = 10) %>% 
  autoplot() +
   scale_y_continuous(labels = scales::comma)

p3 <- pdp::partial(cv.mars2, pred.var = c("Gr_Liv_Area", "Year_Built"), 
              grid.resolution = 10) %>% 
  plotPartial(levelplot = FALSE, zlab = "yhat", drape = TRUE, colorkey = TRUE, 
              screen = list(z = -20, x = -60))

# Display plots side by side
gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
```

### Attrition data

```{r}
df <- rsample::attrition %>% mutate_if(is.ordered, factor, order = F)

# Create training (70%) and test (30%) sets for the attrition data.
set.seed(123)

churn.split <- initial_split(df, prop = .7, strata = "Attrition")
churn.train <- training(churn.split)
churn.test <- testing(churn.split)

set.seed(123)

suppressWarnings(print({
tuned.mars <- train(
   x = subset(churn.train, select = -Attrition),
   y = churn.train$Attrition,
   method = "earth",
   trControl = trainControl(method = "cv", number = 10),
   tuneGrid = hyper.grid
)}))

tuned.mars$bestTune
```

```{r}
ggplot(tuned.mars)
```

```{r}
# train logistic regression model
set.seed(123)

glm.mod <- train(
  Attrition ~ ., 
  data = churn.train, 
  method = "glm",
  family = "binomial",
  preProc = c("zv", "center", "scale"),
  trControl = trainControl(method = "cv", number = 10)
)

# train regularized logistic regression model
set.seed(123)

penalized.mod <- train(
  Attrition ~ ., 
  data = churn.train, 
  method = "glmnet",
  family = "binomial",
  preProc = c("zv", "center", "scale"),
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)

# extract out of sample performance measures
summary(resamples(list(
  Logistic_model = glm.mod, 
  Elastic_net = penalized.mod,
  MARS_model = tuned.mars
  )))$statistics$Accuracy %>%
  kableExtra::kable(caption = "Cross-validated accuracy results for tuned MARS and regression models.") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))

```

```{r}
# clean up
rm(list = ls())
```

### 2020-02-19 HOML/08_K-Nearest-Neighbors.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}K-Nearest Neighbors}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 8}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(mda, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Multivariate Adaptive Regression Splines

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")
churn.train <- training(churn)
churn.test <- testing(churn)

```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

MNIST

```{r}
mnist <- dslabs::read_mnist()
names(mnist)
```

### Overview

K-nearest neighbors (KNN) is a very simple algorithm in which each observation is predicted based on its "similarity" to other observations.

```{r}
df <- recipe(Sale_Price ~., data = ames.train) %>%
   step_nzv(all_nominal()) %>%
   step_integer(matches("Qual|Cond|QC|Qu")) %>%
   step_center(all_numeric(), -all_outcomes()) %>%
   step_scale(all_numeric(), -all_outcomes()) %>%
   step_dummy(all_nominal(), -all_outcomes(), one_hot = T) %>%
   prep(training = ames.train, retain = T) %>%
   juice() %>%
   select(-Sale_Price)

home <- 30
k <- 10
index <- as.vector(FNN::knnx.index(df[-home, ], df[home, ], k = k))
knn.homes <- ames.train[c(home, index), ]

knn.homes %>%
   select(Longitude, Latitude) %>%
   mutate(desc = factor(c('House of Interest', rep('Closest Neighbors', k)),
                        levels = c("House of interest", "Closest Neighbors"))) %>%
   qmplot(Longitude, Latitude, data = .,
          maptype = "toner-background", darken = .7, color = desc, size = I(2.5)) +
   theme(legend.position = "top",
         legend.title = element_blank())
```

### Distance Measures

Euclidean: $\sqrt{\sum^{P}_{j=1}(x_{aj} - x_{bj})^2}$

Manhattan: $\sum^{P}_{j=1}|x_{aj} - x_{bj}|$

Example:

```{r}
two.homes <- ames.train %>%
   select(Gr_Liv_Area, Year_Built) %>%
   sample_n(2)

two.homes

# Euclidean
dist(two.homes, method = "euclidean")

# Manhattan
dist(two.homes, method = "manhattan")
```

Visually:

```{r}
p1 <- ggplot(two.homes, aes(Gr_Liv_Area, Year_Built)) +
  geom_point() +
  geom_line(lty = "dashed") +
  ggtitle("(A) Euclidean distance")

p2 <- ggplot(two.homes, aes(Gr_Liv_Area, Year_Built)) +
  geom_point() +
  geom_step(lty = "dashed") +
  ggtitle("(B) Manhattan distance")

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

### Pre-processing

```{r}
home1 <- ames %>%
  mutate(id = row_number()) %>%
  select(Bedroom_AbvGr, Year_Built, id) %>%
  filter(Bedroom_AbvGr == 4 & Year_Built == 2008) %>%
  slice(1) %>%
  mutate(home = "home1") %>%
  select(home, everything())

home2 <- ames %>%
  mutate(id = row_number()) %>%
  select(Bedroom_AbvGr, Year_Built, id) %>%
  filter(Bedroom_AbvGr == 2 & Year_Built == 2008) %>%
  slice(1) %>%
  mutate(home = "home2") %>%
  select(home, everything())

home3 <- ames %>%
  mutate(id = row_number()) %>%
  select(Bedroom_AbvGr, Year_Built, id) %>%
  filter(Bedroom_AbvGr == 3 & Year_Built == 1998) %>%
  slice(1) %>%
  mutate(home = "home3") %>%
  select(home, everything())

features <- c("Bedroom_AbvGr", "Year_Built")

# distance between home 1 and 2
dist(rbind(home1[,features], home2[,features]))

# distance between home 1 and 3
dist(rbind(home1[,features], home3[,features]))
```

```{r}
scaled.ames <- recipe(Sale_Price ~., ames.train) %>%
   step_center(all_numeric()) %>%
   step_scale(all_numeric()) %>%
   prep(training = ames, retain = T) %>%
   juice()

home1.std <- scaled.ames %>%
   mutate(id = row_number()) %>%
   filter(id == home1$id) %>%
   select(Bedroom_AbvGr, Year_Built, id) %>%
   mutate(home = "home1") %>%
   select(home, everything())

home2.std <- scaled.ames %>%
   mutate(id = row_number()) %>%
   filter(id == home2$id) %>%
   select(Bedroom_AbvGr, Year_Built, id) %>%
   mutate(home = "home2") %>%
   select(home, everything())

home3.std <- scaled.ames %>%
   mutate(id = row_number()) %>%
   filter(id == home1$id) %>%
   select(Bedroom_AbvGr, Year_Built, id) %>%
   mutate(home = "home3") %>%
   select(home, everything())

dist(rbind(home1.std[, features], home2.std[, features]))

dist(rbind(home1.std[, features], home3.std[, features]))
```

### Choosing K

Blueprint

```{r}
blueprint <- recipe(Attrition ~., data = churn.train) %>%
   step_nzv(all_nominal()) %>%
   step_integer(contains("Satisfaction")) %>%
   step_integer(WorkLifeBalance) %>%
   step_integer(JobInvolvement) %>%
   step_dummy(all_nominal(), -all_outcomes(), one_hot = T) %>%
   step_center(all_numeric(), -all_outcomes()) %>%
   step_scale(all_numeric(), -all_outcomes())

# Create a resampling method
cv <- trainControl(
   method = "repeatedcv",
   number = 10,
   repeats = 5,
   classProbs = T,
   summaryFunction = twoClassSummary
)

# Create a hyperparameter grid search
hyper.grid <- expand.grid(
   k = floor(seq(1, nrow(churn.train)/3, length.out = 20))
)

# Fit knn model and perform grid search
knn.grid <- train(
   blueprint,
   data = churn.train,
   method = "knn",
   trControl = cv,
   tuneGrid = hyper.grid,
   metric = "ROC"
)

ggplot(knn.grid)
```

MNIST

KNN doesn't work well with zero (or near zero) variance features. This is a particularly bad problem with MNIST:

```{r}
set.seed(123)

index <- sample(nrow(mnist$train$images), size = 10e3)

mnist.x <- mnist$train$images[index, ]
mnist.y <- factor(mnist$train$labels[index])

mnist.x %>%
   as.data.frame() %>%
   map_df(sd) %>%
   gather(feature, sd) %>%
   ggplot(aes(sd)) +
   geom_histogram(binwidth = 1)
```

```{r}
nzv <- nearZeroVar(mnist.x)

par(mfrow = c(1, 4))
i <- 2
image(1:28, 1:28, matrix(mnist$test$images[i,], nrow=28)[ , 28:1], 
      col = gray(seq(0, 1, 0.05)), xlab = "", ylab="", 
      xaxt="n", yaxt="n", main = "(A) Example image \nfor digit 2")
i <- 7
image(1:28, 1:28, matrix(mnist$test$images[i,], nrow=28)[ , 28:1], 
      col = gray(seq(0, 1, 0.05)), xlab = "", ylab="", 
      xaxt="n", yaxt="n", main = "(B) Example image \nfor digit 4")

i <- 9
image(1:28, 1:28, matrix(mnist$test$images[i,], nrow=28)[ , 28:1], 
      col = gray(seq(0, 1, 0.05)), xlab = "", ylab="", 
      xaxt="n", yaxt="n", main = "(C) Example image \nfor digit 5")
image(matrix(!(1:784 %in% nzv), 28, 28), col = gray(seq(0, 1, 0.05)), 
      xaxt="n", yaxt="n", main = "(D) Typical variability \nin images.")
```

Clean-up features w/ carat

```{r}
colnames(mnist.x) <- paste0("V", 1:ncol(mnist.x))

# Remove near zero variance

nzv <- nearZeroVar(mnist.x)
index <- setdiff(1:ncol(mnist.x), nzv)
mnist.x <- mnist.x[, index]
```

hyper-parameter search

```{r}
cv <- trainControl(
   method = "LGOCV",
   p = 0.7,
   number = 1,
   savePredictions = T
)

hyper.grid <- expand.grid(k = seq(3, 25, by = 2))

knn.mnist <- train(
   mnist.x,
   mnist.y,
   method = "knn",
   tuneGrid = hyper.grid,
   preProc = c("center", "scale"),
   trControl = cv
)

ggplot(knn.mnist)
```

Diagnostics

```{r}
cm <- confusionMatrix(knn.mnist$pred$pred, knn.mnist$pred$obs)
cm$byClass[, c(1:2, 11)] # sensitivity, specificity & accuracy
```

```{r}
vi <- varImp(knn.mnist)
vi
```

Most influential areas of the image:

```{r}
par(mfrow = c(1,1))
# Get median value for feature importance
imp <- vi$importance %>%
   rownames_to_column(var = "feature") %>%
   gather(response, imp, -feature) %>%
   group_by(feature) %>%
   summarize(imp = median(imp))

# Create tibble for all edge pixels
edges <- tibble(
   feature = paste0("V", nzv),
   imp = 0
)

# Combine and plot
imp <- rbind(imp, edges) %>%
   mutate(ID = as.numeric(str_extract(feature, "\\d+"))) %>%
   arrange(ID)

image(matrix(imp$imp, 28, 28), col = gray(seq(0, 1, 0.05)),
      xaxt = "n", yaxt = "n")
```

```{r}
# Get a few accurate predictions
set.seed(9)
good <- knn_mnist$pred %>%
  filter(pred == obs) %>%
  sample_n(4)

# Get a few inaccurate predictions
set.seed(9)
bad <- knn_mnist$pred %>%
  filter(pred != obs) %>%
  sample_n(4)

combine <- bind_rows(good, bad)

# Get original feature set with all pixel features
set.seed(123)
index <- sample(nrow(mnist$train$images), 10000)
X <- mnist$train$images[index,]

# Plot results
par(mfrow = c(4, 2), mar=c(1, 1, 1, 1))
layout(matrix(seq_len(nrow(combine)), 4, 2, byrow = FALSE))
for(i in seq_len(nrow(combine))) {
  image(matrix(X[combine$rowIndex[i],], 28, 28)[, 28:1], 
        col = gray(seq(0, 1, 0.05)),
        main = paste("Actual:", combine$obs[i], "  ", 
                     "Predicted:", combine$pred[i]),
        xaxt="n", yaxt="n") 
}
```

```{r}
# clean up
rm(list = ls())
```

### 2020-02-19 HOML/09_Decision-Trees.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Decision Trees}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 9}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(rpart, quietly = TRUE, warn.conflicts = F)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

##  Decision Trees

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")
churn.train <- training(churn)
churn.test <- testing(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

### Decision Trees

A decision tree is a flowchart-like structure in which each internal node represents a "test" on an attribute (e.g. whether a coin flip comes up heads or tails), each branch represents the outcome of the test, and each leaf node represents a class label (decision taken after computing all attributes). The most common type of decision tree is a *c*lassification *a*nd *r*egresion *t*ree (CART).

### Partitioning

CART uses _binary recursive partitioning_, where the objective at each node is to find the "best" feature ($x_i$) to partition the remaining data into one of two regions ($R_1, R_2$) such that the overall error between the actual response ($y_i$) and the predicted constant ($c_i$) is minimized.

For regression, the objective is to minimize the total SSE:

$SSE = \sum_{i \in R_1}{(y_i - c_1)^2} + \sum_{i \in R_2}{(y_i - c_2)^2}$

For classification problems, the partitioning is usually made to maximize the reduction in cross-enthropy or the Gini index.

For example, say we have data generated from a simple sin function with Gaussian noise: $Y_i \stackrel{i.i.d}{\sim} N(sin(X_i, \sigma^2)$

Data & model:

```{r}
set.seed(1112)

df <- tibble::tibble(
   x = seq(from = 0, to = 2 * pi, length = 500),
   y = sin(x) + rnorm(length(x), sd = .5),
   truth = sin(x)
)

# run decision stump model
ctrl <- list(cp = 0, minbucket = 5, maxdepth = 1)
fit <- rpart(y ~ x, data = df, control = ctrl)

# plot tree
rpart.plot(fit)
```

Decision Boundry:

```{r}
df %>%
   mutate(pred = predict(fit, df)) %>%
   ggplot(aes(x, y)) +
      geom_point(alpha = .2, size = 1) +
      geom_line(aes(x, y = truth), color = "blue", size = .75) +
      geom_line(aes(y = pred), color = "red", size = .75) +
      geom_segment(x = 3.1, xend = 3.1, y = -Inf, yend = -.95,
                   arrow = arrow(length = unit(0.25, "cm")), size = .25) +
      annotate("text", x = 3.1, y = -Inf, label = "split", hjust = 1.2, vjust = -1, size = 3) +
      geom_segment(x = 5.5, xend = 6, y = 2, yend = 2, size = .75, color = "blue") +
      geom_segment(x = 5.5, xend = 6, y = 1.7, yend = 1.7, size = .75, color = "red") +
      annotate("text", x = 5.3, y = 2, label = "truth", hjust = 1, size = 3, color = "blue") +
      annotate("text", x = 5.3, y = 1.7, label = "decision boundary", hjust = 1, size = 3, color = "red")
```

Depth 3 decision tree:

```{r}
# fit depth 3 decision tree
ctrl <- list(cp = 0, minbucket = 5, maxdepth = 3)
fit <- rpart(y ~ x, data = df, control = ctrl)
rpart.plot(fit)
```

Decision Boundry:

```{r}
# plot decision boundary
df %>%
  mutate(pred = predict(fit, df)) %>%
  ggplot(aes(x, y)) +
  geom_point(alpha = .2, size = 1) +
  geom_line(aes(x, y = truth), color = "blue", size = .75) +
  geom_line(aes(y = pred), color = "red", size = .75)
```

IRIS dataset:

```{r}
# decision tree
iris_fit <- rpart(Species ~ Sepal.Length + Sepal.Width, data = iris)
rpart.plot(iris_fit)
```

```{r}
# decision boundary
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species, shape = Species)) +
  geom_point(show.legend = FALSE) +
  annotate("rect", xmin = -Inf, xmax = 5.44, ymin = 2.8, ymax = Inf, alpha = .75, fill = "orange") +
  annotate("text", x = 4.0, y = 4.4, label = "setosa", hjust = 0, size = 3) +
  annotate("rect", xmin = -Inf, xmax = 5.44, ymin = 2.79, ymax = -Inf, alpha = .75, fill = "grey") +
  annotate("text", x = 4.0, y = 2, label = "versicolor", hjust = 0, size = 3) +
  annotate("rect", xmin = 5.45, xmax = 6.15, ymin = 3.1, ymax = Inf, alpha = .75, fill = "orange") +
  annotate("text", x = 6, y = 4.4, label = "setosa", hjust = 1, vjust = 0, size = 3) +
  annotate("rect", xmin = 5.45, xmax = 6.15, ymin = 3.09, ymax = -Inf, alpha = .75, fill = "grey") +
  annotate("text", x = 6.15, y = 2, label = "versicolor", hjust = 1, vjust = 0, fill = "grey", size = 3) +
  annotate("rect", xmin = 6.16, xmax = Inf, ymin = -Inf, ymax = Inf, alpha = .75, fill = "green") +
  annotate("text", x = 8, y = 2, label = "virginica", hjust = 1, vjust = 0, fill = "green", size = 3)
```

### How deep?

If we keep increasing the depth of the tree, we will eventually overfit the training data.

```{r}
ctrl <- list(cp = 0, minbucket = 1, maxdepth = 50)
fit <- rpart(y ~ x, data = df, control = ctrl)
rpart.plot(fit)
```

```{r}
df %>%
  mutate(pred = predict(fit, df)) %>%
  ggplot(aes(x, y)) +
  geom_point(alpha = .2, size = 1) +
  geom_line(aes(x, y = truth), color = "blue", size = 0.75) +
  geom_line(aes(y = pred), color = "red", size = 0.75)
```

There are two basic strategies for finding the optimal depth of the tree, early stopping and pruning:

```{r}
hyper.grid <- expand.grid(
   maxdepth = c(1, 5, 15),
   minbucket = c(1, 5, 15)
)

results <- data.frame(NULL)

for(i in seq_len(nrow(hyper.grid))) {
   ctrl <- list(cp = 0, maxdepth = hyper.grid$maxdepth[i], minbucket = hyper.grid$minbucket[i])
   fit <- rpart(y ~ x, data = df, control = ctrl)

   predictions <- mutate(
      df,
      minbucket = factor(paste("Min node size =", hyper.grid$minbucket[i]), ordered = T),
      maxdepth = factor(paste("Max tree depth =", hyper.grid$maxdepth[i]), ordered = T)
   )

   predictions$pred <- predict(fit, df)
   results <- rbind(results, predictions)
}

ggplot(results, aes(x, y)) +
   geom_point(alpha = .2, size = 1) +
   geom_line(aes(x, y = truth), color = "blue", size = .75) +
   geom_line(aes(y = pred), color = "red", size = 1) +
   facet_grid(minbucket ~ maxdepth)
```

### Pruning

Alternative to specifying a max depth, build the most complicated tree and then prune it back for generalizability.

We find the optimal subtree by using a _cost complexity parameter_ ($\alpha$) that penalizes our objective function:

$minimize\{SST+\alpha|T|\}$

(Similar to Lasso regression)

```{r}
ctrl <- list(cp = 0, minbucket = 1, maxdepth = 50)
fit <- rpart(y ~ x, data = df, control = ctrl)

p1 <- df %>%
  mutate(pred = predict(fit, df)) %>%
  ggplot(aes(x, y)) +
  geom_point(alpha = .3, size = 2) +
  geom_line(aes(x, y = truth), color = "blue", size = 1) +
  geom_line(aes(y = pred), color = "red", size = 1)

fit2 <- rpart(y ~ x, data = df)

p2 <- df %>%
  mutate(pred2 = predict(fit2, df)) %>%
  ggplot(aes(x, y)) +
  geom_point(alpha = .3, size = 2) +
  geom_line(aes(x, y = truth), color = "blue", size = 1) +
  geom_line(aes(y = pred2), color = "red", size = 1)

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

### Example: AMES Housing Data

```{r}
ames.dt1 <- rpart(
   formula = Sale_Price ~ .,
   data = ames.train,
   method = "anova"
)

ames.dt1

rpart.plot(ames.dt1)
```

```{r}
plotcp(ames.dt1)
```

```{r}
ames.dt2 <- rpart(
    formula = Sale_Price ~ .,
    data    = ames.train,
    method  = "anova", 
    control = list(cp = 0, xval = 10)
)

plotcp(ames.dt2)
abline(v = 11, lty = "dashed")
```

```{r}
ames.dt1$cptable
```

Cross-validated parameter search:

```{r}
ames.dt3 <- train(
   Sale_Price ~ .,
   data = ames.train,
   method = "rpart",
   trControl = trainControl(method = "cv", number = 10),
   tuneLength = 20
)

ggplot(ames.dt3)
```

### Feature Interpretation

```{r}
vip(ames.dt3, num_features = 40, geom = "point")
```

```{r}
# Construct partial dependence plots
p1 <- pdp::partial(ames.dt3, pred.var = "Gr_Liv_Area") %>% autoplot()
p2 <- pdp::partial(ames.dt3, pred.var = "Year_Built") %>% autoplot()
p3 <- pdp::partial(ames.dt3, pred.var = c("Gr_Liv_Area", "Year_Built")) %>% 
  plotPartial(levelplot = FALSE, zlab = "yhat", drape = TRUE, 
              colorkey = TRUE, screen = list(z = -20, x = -60))

# Display plots side by side
gridExtra::grid.arrange(p1, p2, p3, ncol = 3)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-02-19 HOML/10_Bagging.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Bagging}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 10}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)
library(foreach, quietly = T, warn.conflicts = F)
library(doParallel, quietly = T, warn.conflicts = F)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(rpart, quietly = TRUE, warn.conflicts = F)
library(ipred, quietly = T, warn.conflicts = F)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

##  Bagging

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")
churn.train <- training(churn)
churn.test <- testing(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

### Bagging Overview

Bagging (Bootstrap AGGregatING) is a procedure that uses bootstrapping techniques to create an ensemble of predictions to improve the accuracy of regression and classification methods.

### How Bagging Works

Bagging is a fairly straightforward algorithm that makes _b_ bootstrap copies of the original training data, to which the (regression or classification) algorithm is applied (referred to as the "Base Learner"), and then the predictions are averaged together from the individual base learners.

For each record, X, we want to predict:

$\hat{f_{bag}}$ is the bagged prediction, and $\hat{f_1(X)}, \hat{f_2(X)}, \ldots, \hat{f_b(X)}$ are the predictions from the individual base learners.

So that,

$\hat{f(bag)} = \hat{f_1(X)} + \hat{f_2(X)} + \ldots + \hat{f_b(X)}$

Bagging effectively reduces the variance of an individual base learning through the aggrigation process. So, this process works especially well for high variance base learners (Decision Tress, KNN (small k)).

This is an example of the "Wisdom of the Crowd" approach to learners.

Example of "Bagging":

```{r, fig.height=7}
# Simulate some nonlinear monotonic data
set.seed(123)  # for reproducibility

x <- seq(from = 0, to = 2 * pi, length = 500)
y <- sin(x) + rnorm(length(x), sd = 0.3)
df <- data.frame(x, y) %>%
  filter(x < 4.5)

# bootstrapped polynomial model fit
bootstrap_n <- 100
bootstrap_results <- NULL
for(i in seq_len(bootstrap_n)) {
  # reproducible sampled data frames
  set.seed(i)
  index <- sample(seq_len(nrow(df)), nrow(df), replace = TRUE)
  df_sim <- df[index, ]

  # fit model and add predictions to results data frame
  fit <- lm(y ~ I(x^3), data = df_sim)
  df_sim$predictions <- predict(fit, df_sim)
  df_sim$model <- paste0("model", i)
  df_sim$ob <- index
  bootstrap_results <- rbind(bootstrap_results, df_sim)
}

p1 <- ggplot(bootstrap_results, aes(x, predictions)) +
  geom_point(data = df, aes(x, y), alpha = .25) +
  geom_line(aes(group = model), show.legend = FALSE, size = .5, alpha = .2) +
  stat_summary(fun.y = "mean", colour = "red", size = 1, geom = "line") +
  scale_y_continuous("Response", limits = c(-2, 2), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 5), expand = c(0, 0)) +
  ggtitle("A) Polynomial regression")

# bootstrapped MARS model fit
bootstrap_n <- 100
bootstrap_results <- NULL
for(i in seq_len(bootstrap_n)) {
  # reproducible sampled data frames
  set.seed(i)
  index <- sample(seq_len(nrow(df)), nrow(df), replace = TRUE)
  df_sim <- df[index, ]

  # fit model and add predictions to results data frame
  fit <- earth::earth(y ~ x, data = df_sim)
  df_sim$predictions <- predict(fit, df_sim)
  df_sim$model <- paste0("model", i)
  df_sim$ob <- index
  bootstrap_results <- rbind(bootstrap_results, df_sim)
}

p2 <- ggplot(bootstrap_results, aes(x, predictions)) +
  geom_point(data = df, aes(x, y), alpha = .25) +
  geom_line(aes(group = model), show.legend = FALSE, size = .5, alpha = .2) +
  stat_summary(fun.y = "mean", colour = "red", size = 1, geom = "line") +
  scale_y_continuous(NULL, limits = c(-2, 2), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 5), expand = c(0, 0)) +
  ggtitle("B) MARS")

# bootstrapped decision trees fit
bootstrap_n <- 100
bootstrap_results <- NULL
for(i in seq_len(bootstrap_n)) {
  # reproducible sampled data frames
  set.seed(i)
  index <- sample(seq_len(nrow(df)), nrow(df), replace = TRUE)
  df_sim <- df[index, ]

  # fit model and add predictions to results data frame
  fit <- rpart::rpart(y ~ x, data = df_sim)
  df_sim$predictions <- predict(fit, df_sim)
  df_sim$model <- paste0("model", i)
  df_sim$ob <- index
  bootstrap_results <- rbind(bootstrap_results, df_sim)
}

p3 <- ggplot(bootstrap_results, aes(x, predictions)) +
  geom_point(data = df, aes(x, y), alpha = .25) +
  geom_line(aes(group = model), show.legend = FALSE, size = .5, alpha = .2) +
  stat_summary(fun.y = "mean", colour = "red", size = 1, geom = "line") +
  scale_y_continuous(NULL, limits = c(-2, 2), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 5), expand = c(0, 0)) +
  ggtitle("C) Decision trees")

gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
```

Optimal performance is typically found using 50-500 trees.

Interestingly, bagging decision trees does not tend to overfit, however, using a CV approach with bagging does begin to become computationally burdensome.

The out-of-bag (OOB) is an internal estimate of the predictive performance. Think of this as a "free" CV statistic.

### Implementation

Here we will use a bag of unprunded (we don't prune the individual trees because the bagging process will reduce the variance, and therefore have the largest impact) decision trees on the ames data set.

```{r}
set.seed(123)

# train bagged model
ames.bag1 <- bagging(
  formula = Sale_Price ~ .,
  data = ames.train,
  nbagg = 100,
  coob = T,
  control = rpart.control(minsplit = 2, cp = 0)
)

ames.bag1
```

Error curve for bagging 1-200 deep, unpruned decision trees:

```{r}
n.tree <- seq(1, 200, by = 2)

rmse <- vector(mode = "numeric", length = length(n.tree))

for(i in seq_along(n.tree))
{
  set.seed(123)

  # perform bagged model
  model <- ranger::ranger(
    formula = Sale_Price ~ .,
    data = ames.train,
    num.trees = n.tree[i],
    mtry = ncol(ames.train) - 1,
    min.node.size = 1
  )

  rmse[i] <- sqrt(model$prediction.error)
}

bagging.errors <- data.table(n.tree, rmse)

ggplot(bagging.errors, aes(n.tree, rmse)) +
  geom_line() +
  geom_hline(yintercept = 41019, lty = "dashed", color = "grey50") +
  annotate("text", x = 100, y = 41385, label = "Best individual pruned tree", vjust = 0, hjust = 0, color = "grey50") +
  annotate("text", x = 100, y = 26750, label = "Bagged trees", vjust = 0, hjust = 0) +
  scale_y_continuous(labels = comma) +
  ylab("RMSE") +
  xlab("Number of trees")
```

```{r}
ames.bag2 <- train(
  Sale_Price ~ .,
  data = ames.train,
  method = "treebag",
  trControl = trainControl(method = "cv", number = 10),
  nbagg = 200,
  control = rpart.control(minsplit = 2, cp = 0)
)

ames.bag2
```

Since each tree is trained on independent samples of the data, we can easily parallelize this process.

```{r}
cl <- makeCluster(8)

registerDoParallel(cl)

# fit trees in parallel and compute predictions on the test set

predictions <- foreach(
  icount(160),
  .packages = "rpart",
  .combine = cbind
) %dopar% {
    # bootstrap copy of training data
    index <- sample(nrow(ames.train), replace = T)
    ames.train.boot <- ames.train[index, ]

    # fit tree to bootstrap copy
    bagged.tree <- rpart(
      Sale_Price ~ .,
      control = rpart.control(minsplit = 2, cp = 0),
      data = ames.train.boot
    )

    predict(bagged.tree, newdata = ames.test)
}

predictions[1:5, 1:7]

```

```{r}
predictions %>%
  as.data.frame() %>%
  mutate(
    observation = 1:n(),
    actual = ames.test$Sale_Price) %>%
  tidyr::gather(tree, predicted, -c(observation, actual)) %>%
  group_by(observation) %>%
  mutate(tree = stringr::str_extract(tree, '\\d+') %>% as.numeric()) %>%
  ungroup() %>%
  arrange(observation, tree) %>%
  group_by(observation) %>%
  mutate(avg_prediction = cummean(predicted)) %>%
  group_by(tree) %>%
  summarize(RMSE = RMSE(avg_prediction, actual)) %>%
  ggplot(aes(tree, RMSE)) +
  geom_line() +
  xlab('Number of trees')
```

```{r}
stopCluster(cl)
```

### Feature Interpretation

Unfortunatly, due to the many trees involved feature interpretation becomes tricky. Typically, the bagging process will use many features, at lower levels of importance.

```{r, fig.height=8}
vip::vip(ames.bag2, num_features = 40, bar = FALSE)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-02-19 HOML/15_Stacked-Models.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Stacked Models}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 15}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)
library(svmpath, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(kernlab, quietly = TRUE, warn.conflicts = FALSE)
library(mlbench, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Stacked Models

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

set.seed(123)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

h2o.init(max_mem_size = "10g", strict_version_check = F)

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

```{r}
# ensure consistent categorical levels
blueprint <- recipe(Sale_Price ~., data = ames_train) %>%
  step_other(all_nominal(), threshold = 0.005)

# Create training / test h2o frames
train_h2o <- prep(blueprint, training = ames_train, retain = T) %>%
  juice() %>%
  as.h2o()

test_h2o <- prep(blueprint, training = ames_train) %>%
  bake(new_data = ames_test) %>%
  as.h2o()

Y <- "Sale_Price"
X <- setdiff(names(ames_train), Y)
```

### Stacking Overview

Stacking is the process of combining multiple "base" learners (RF, GBM, GLM, etc.) into a "Super Learner" that uses the individual learners to make the final prediction.

The Super Learner Algorithm

1.) Setup the Ensemble

  Specify a list of L base learners (with a specific set of model parameters)

  Specify a meta learning algorithm. This can be any one of the algorithms discussed in the previous chapters, but most often is some form of regularized regression.

2.) Train the ensemble

  Train each of the L base learners on the training set.

  Perform k-fold CV on each of the base learners and collect the cross-validated predictions from each (the same k-fold must be used for each base learner). These predicted values represent $p_1, ..., p_l$.

  The N cross-validated predicted values from each of the L algorithms can be combined to form a new N X L feature matrix (Z)

  Train the meta learning algorithm on level-one data (y = f(Z)). The "ensemble model" consista of the L base learning models and the meta learning model, which can then be used to generate predictions on new data.

3.) Predict on new data.

  To generate ensemble predictions, first generate predictions from the base learners.

  Feed those predictions into the meta learner to generate the ensemble prediction.

### Stacking Existing Models

Restrictions:

1.) All models must be trained on the same training set

2.) All models must be trained with the same number of CV folds.

3.) All models must use the same fold assignment to ensure the same observations are used.

4.) The cross-validated predictions from all of the models must be preserved by setting keep_cross_validation_predictions = T.

```{r}
# Train & cross-validate a GLM model
best_glm <- h2o.glm(
  x = X, y = Y, training_frame = train_h2o, alpha = 0.1,
  remove_collinear_columns = TRUE, nfolds = 10, fold_assignment = "Modulo",
  keep_cross_validation_predictions = TRUE, seed = 123
)

# Train & cross-validate a RF model
best_rf <- h2o.randomForest(
  x = X, y = Y, training_frame = train_h2o, ntrees = 1000, mtries = 20,
  max_depth = 30, min_rows = 1, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate a GBM model
best_gbm <- h2o.gbm(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.01,
  max_depth = 7, min_rows = 5, sample_rate = 0.8, nfolds = 10,
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,es

  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
  stopping_tolerance = 0
)

# Train & cross-validate an XGBoost model
best_xgb <- h2o.xgboost(
  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
  nfolds = 10, fold_assignment = "Modulo", 
  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
  stopping_metric = "RMSE", stopping_tolerance = 0
)
```

Train the ensemble

```{r}
# Train a stacked tree ensemble
ensemble_tree <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble",
  base_models = list(best_glm, best_rf, best_gbm, best_xgb),
  metalearner_algorithm = "drf"
)
```

```{r}
get_rmse <- function(model) {
  results <- h2o.performance(model, newdata = test_h2o)
  results@metrics$RMSE
}

list(best_glm, best_rf, best_glm, best_xgb) %>%
  purrr::map_dbl(get_rmse)
```

Stacked Results

```{r}
h2o.performance(ensemble_tree, newdata = test_h2o)@metrics$RMSE
```

```{r}
data.frame(
  GLM_pred = as.vector(h2o.getFrame(best_glm@model$cross_validation_holdout_predictions_frame_id$name)),
  RF_pred = as.vector(h2o.getFrame(best_rf@model$cross_validation_holdout_predictions_frame_id$name)),
  GBM_pred = as.vector(h2o.getFrame(best_gbm@model$cross_validation_holdout_predictions_frame_id$name)),
  XGB_pred = as.vector(h2o.getFrame(best_xgb@model$cross_validation_holdout_predictions_frame_id$name))
) %>% cor()
```

### Stacking a Search Grid

```{r}
# Define GBM hyperparameter grid
hyper_grid <- list(
  max_depth = c(1, 3, 5),
  min_rows = c(1, 5, 10),
  learn_rate = c(0.01, 0.05, 0.1),
  learn_rate_annealing = c(0.99, 1),
  sample_rate = c(0.5, 0.75, 1),
  col_sample_rate = c(0.8, 0.9, 1)
)

# Define random grid search criteria
search_criteria <- list(
  strategy = "RandomDiscrete",
  max_models = 25
)

# Build random grid search 
random_grid <- h2o.grid(
  algorithm = "gbm", grid_id = "gbm_grid", x = X, y = Y,
  training_frame = train_h2o, hyper_params = hyper_grid,
  search_criteria = search_criteria, ntrees = 5000, stopping_metric = "RMSE",
  stopping_rounds = 10, stopping_tolerance = 0, nfolds = 10, 
  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
  seed = 123
)
```

```{r}
# Sort results by RMSE
h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "rmse"
)
```

```{r}
# Grab the model_id for the top model, chosen by validation error
best_model_id <- random_grid@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)
h2o.performance(best_model, newdata = test_h2o)
```

```{r}
# Train a stacked ensemble using the GBM grid
ensemble <- h2o.stackedEnsemble(
  x = X, y = Y, training_frame = train_h2o, model_id = "ensemble_gbm_grid",
  base_models = random_grid@model_ids, metalearner_algorithm = "gbm"
)

# Eval ensemble performance on a test set
h2o.performance(ensemble, newdata = test_h2o)
```

### Auto ML

```{r}
# Use AutoML to find a list of candidate models (i.e., leaderboard)
auto_ml <- h2o.automl(
  x = X, y = Y, training_frame = train_h2o, nfolds = 5, 
  max_runtime_secs = 60 * 120, max_models = 50,
  keep_cross_validation_predictions = TRUE, sort_metric = "RMSE", seed = 123,
  stopping_rounds = 50, stopping_metric = "RMSE", stopping_tolerance = 0
)
```

```{r}
# Assess the leader board; the following truncates the results to show the top 
# 25 models. You can get the top model with auto_ml@leader
auto_ml@leaderboard %>% 
  as.data.frame() %>%
  dplyr::select(model_id, rmse) %>%
  dplyr::slice(1:25)
```

#### Clean-up

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-02-19 ISLR/02_Statistical-Learning.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 2}
   \fancyfoot[C]{\rmfamily\color{headergrey}Introduction to Statistical Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

```

## Chapter 2

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### R Lab

```{r, echo = T, fig.width=8, fig.height=3.5}
x <- seq(-pi, pi, length.out = 50); y <- x

f <- outer(x, y, function(x, y) cos(y)/(1+x^2))

contour(x, y, f)

contour(x, y, f, nlevels = 45, add = T)

fa <- (f - t(f)) / 2

contour(x, y, fa, nlevels = 15)
```

```{r, echo = T, fig.width=8, fig.height=3.5}
image(x, y, fa)

persp(x, y, fa)

persp(x, y, fa, theta = 30)

persp(x, y, fa, theta = 30, phi = 20)
persp(x, y, fa, theta = 30, phi = 70)
persp(x, y, fa, theta = 30, phi = 40)
```

### Conceptual

#### 1.)

For each of parts (a) through (d), indicate whether i. or ii. is correct, and explain your answer. In general, do we expect the performance of a flexible statistical learning method to perform better or worse than an inflexible method when:

a.)The sample size n is extremely large, and the number of predictors p is small ?

_Better. A flexible method will fit the data closer and with the large sample size, would perform better than an inflexible approach._

b.) The number of predictors p is extremely large, and the number of observations n is small ?

_Worse. A flexible method would overfit the small number of observations._

c.) The relationship between the predictors and response is highly non-linear ?

_Better. With more degrees of freedom, a flexible method would fit better than an inflexible one._

d.) The variance of the error terms, i.e. σ2=Var(ε), is extremely high ?

_Worse. A flexible method would fit to the noise in the error terms and increase variance._

#### 2.)

Explain whether each scenario is a classification or regression problem, and indicate whether we are most interested in inference or prediction. Finally, provide n and p.

a.) We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of employees, industry and the CEO salary. We are interested in understanding which factors affect CEO salary.

_Regression. n = 500, p = 3_

b.) We are considering launching a new product and wish to know whether it will be a success or a failure. We collect data on 20 similar products that were previously launched. For each product we have recorded whether it was a success or failure, price charged for the product, marketing budget, competition price, and ten other variables.

_Classification. n = 20, p = 13_

c.) We are interesting in predicting the % change in the US dollar in relation to the weekly changes in the world stock markets. Hence we collect weekly data for all of 2012. For each week we record the % change in the dollar, the % change in the US market, the % change in the British market, and the % change in the German market.

_Regression. n = 52, p = 3_

#### 3.)

We now revisit the bias-variance decomposition.

a.) Provide a sketch of typical (squared) bias, variance, training error, test error, and Bayes (or irreducible) error curves, on a single plot, as we go from less flexible statistical learning methods towards more flexible approaches. The x-axis should represent the amount of flexibility in the method, and the y-axis should represent the values for each curve. There should be five curves. Make sure to label each one.

```{r, echo = T, fig.width=8, fig.height=4}
mu <- 2

Z <- rnorm(20000, mu)

MSE <- function(estimate, mu) {
   return(sum((estimate - mu)^2) / length(estimate))
}

n <- 50
shrink <- seq(0,0.5, length=n)
test.mse <- numeric(n)
train.mse <- numeric(n)
bias <- numeric(n)
variance <- numeric(n)

for (i in 1:n) {
   test.mse[i] <- MSE((1 - shrink[i]) * Z, mu)
   bias[i] <- mu * shrink[i]
   variance[i] <- (1 - shrink[i])^2
   train.mse[i] <- (variance[i] - bias[i] ) ^2
}

data.table(x = shrink, var = variance, bias = bias^2, test.mse = test.mse, train.mse = train.mse) %>%
   ggplot(data = .) +
   geom_line(aes(x, var, col = "Variance"), lwd = .8, lty = 2,) +
   geom_line(aes(x, bias, col = "Bias^2"), lwd = .8, lty = 2) +
   geom_line(aes(x, test.mse, col = "Test MSE"), lwd = .8) +
   geom_line(aes(x, train.mse, col = "Train MSE"), lwd = .8) +
   geom_hline(aes(yintercept = .8, col = "Irreduciable"), lty = 3) +
   scale_colour_manual(values=c("cornflowerblue", "black", "darkorange", "darkred", "darkgreen")) +
   labs(title = "Bias vs Variance Trade-off", y = "Variance", x = "Flexibility") +
   theme(legend.title = element_blank(),
      axis.text.x=element_blank(), axis.text.y=element_blank())

```

b.) Explain why each of the five curves has the shape displayed in part (a

#### 4.)

You will now think of some real-life applications for statistical learning.

a.) Describe three real-life applications in which classification might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction ? Explain your answer.

+ Infering if an e-mail is spam/ham. Y = Spam {Y|No}, Y = {words in email, subject, from addr}.

+ Predicting if a customer will redeemd a coupon. Y = Redeem {Y||No}, X = {purchase history, coupon value, frequency of store visit}.

+ Predicting if an inmate will recidivate. Y = Recid {Y|N}, X = {crime type, age, release date, time served}

b.) Describe three real-life applications in which regression might be useful. Describe the response, as well as the predictors. Is the goal of each application inference or prediction ? Explain your answer.

+ Predicting the sale price of a home. Y = sale price, X = {year built, sq footage, quality}.

+ Predicting the next day return of a stock. Y = log(Return), X = {prior returns}

+ Predicting the customer annual spend on clothing. Y = {Spend? $}, X = {num of items purchased last 1 year, inventory}

#### 5.)

What are the advantages and disadvantages of a very flexible (versus a less flexible) approach for regression or classification ? Under what circumstances might a more flexible approach be preferred to a less flexible approach? When might a less flexible approach be preferred ?

_The advantages of a very flexible approach are that it may give a better fit for non-linear models and it decreases the bias.

The disadvantages of a very flexible approach are that it requires estimating a greater number of parameters, it follows the noise too closely (overfit) and it increases the variance.

A more flexible approach would be preferred to a less flexible approach when we are interested in prediction and not the interpretability of the results.

A less flexible approach would be preferred to a more flexible approach when we are interested in inference and the interpretability of the results._

#### 6.)

Describe the differences between a parametric and a non-parametric statistical learning approach. What are the advantages of a parametric approach to regression or classification (as opposed to a nonparametric approach) ? What are its disadvantages ?

_A parametric approach reduces the problem of estimating f down to one of estimating a set of parameters because it assumes a form for f.

A non-parametric approach does not assume a patricular form of f and so requires a very large sample to accurately estimate f.

The advantages of a parametric approach to regression or classification are the simplifying of modeling f to a few parameters and not as many observations are required compared to a non-parametric approach.

The disadvantages of a parametric approach to regression or classification are a potentially inaccurate estimate f if the form of f assumed is wrong or to overfit the observations if more flexible models are used._

#### 7.)

The table below provides a training data set containing 6 observations, 3 predictors, and 1 qualitative response variable. Suppose we wish to use this data set to make a prediction for Y when X1 = X2 = X3 = 0 using K-nearest neighbors.

```{r, echo = T}
dat <- data.table(Obs = 1:6, 
           X1 = c(0, 2, 0, 0, -1, 1),
           X2 = c(3, 0, 1, 1, 0, 1),
           X3 = c(0, 0, 3, 2, 1, 1),
           Y = c("Red", "Red", "Red", "Green", "Green", "Red"))

dat
```

a.) Compute the Euclidean distance between each observation and the test point, X1 = X2 = X3 = 0.

```{r, echo = T}
dat[, Distance := ( (X1 - 0)^2 + (X2 - 0)^2 + (X3 - 0)^2 )^.5 ]
dat
```

b.) What is our prediction with K=1 ? Why ?

If $K = 1$ then $X_5 \in N_0$, so that:

$P(Y = Red | X = x_0) = \frac{1}{1}\sum_{i\in N_0} I(y_i = Red) = 0$

and

$P(Y = Green | X = x_0) = \frac{1}{1}\sum_{i\in N_0} I(y_i = Greeen) = 1$

or:

```{r, echo = T}
setorder(dat, Distance)

dat[1]$Y
```

Our prediction is _green._

c.) What is our prediction with K=3 ? Why ?

```{r, echo = T}
dat[1:3]
```

_2 out of the closest 3 points are red, so we would predict red._

d.) If the Bayes decision boundary in this problem is highly nonlinear, then would we expect the best value for K to be large or small ? Why ?

_As K becomes larger, the boundary becomes inflexible (linear). So in this case we would expect the best value for K to be small._

### Applied

#### 8.)

This exercise relates to the “College” data set, which can be found in the file “College.csv”. It contains a number of variables for 777 different universities and colleges in the US.

a.) Use the read.csv() function to read the data into R. Call the loaded data “college”. Make sure that you have the directory set to the correct location for the data.

```{r, echo = T}
data(College)
```

b.) Look at the data using the fix() function. You should notice that the first column is just the name of each university. We don’t really want R to treat this as data. However, it may be handy to have these names for later.

```{r, echo = T}
# fix(College)

head(College)
```

c.) Use the summary() function to produce a numerical summary of the variables in the data set.

```{r, echo = T}
summary(College)
```

Use the pairs() function to produce a scatterplot matrix of the first ten columns or variables of the data.

```{r, echo = T, fig.width=8, fig.height=7}
ggpairs(College)
```

Use the plot() function to produce side-by-side boxplots of “Outstate” versus “Private”.

```{r, echo = T, fig.width=8, fig.height=3.5}
ggplot(College, aes(Private, Outstate)) +
   geom_boxplot(fill = "cornflowerblue") +
   scale_y_continuous(label = dollar) +
   labs(title = "Private Schools Tuition")
```

Create a new qualitative variable, called “Elite”, by binning the “Top10perc” variable. Use the summary() function to see how many elite universities there are. Now use the plot() function to produce side-by-side boxplots of “Outstate” versus “Elite”.

```{r, echo = T, fig.width=8, fig.height=3.5}
college <- as.data.table(College)
college[, Elite := ifelse(Top10perc > 50, "Yes", "No")]

summary(college)

ggplot(college, aes(Elite, Outstate)) +
   geom_boxplot(fill = "cornflowerblue") +
   scale_y_continuous(label = dollar) +
   labs(title = "Elite Schools Tuition")
```

Use the hist() function to produce some histograms with numbers of bins for a few of the quantitative variables.

```{r, echo=T, fig.width=8, fig.height=8}

p1 <- ggplot(college) +
   geom_histogram(aes(Books), fill = "darkred", bins = 30) +
   labs(title = "Books")

p2 <-ggplot(college) +
   geom_histogram(aes(PhD), fill = "cornflowerblue", bins = 30) +
   labs(title = "Ph.D")

p3 <-ggplot(college) +
   geom_histogram(aes(Grad.Rate), fill = "darkgreen", bins = 30) +
   labs(title = "Graduation Rate")

p4 <- ggplot(college) +
   geom_histogram(aes(perc.alumni), fill = "darkorange", bins = 30) +
   labs(title = "% Alumni")

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

#### 9.)

This exercise involves the “Auto” data set studied in the lab. Make sure the missing values have been removed from the data.

a.) Which of the predictors are quantitative, and which are qualitative ?

```{r, echo = T}
auto <- ISLR::Auto
str(auto)
```

_All variables except “horsepower” and “name” are quantitative._

b.) What is the range of each quantitative predictor ?

```{r, echo = T}
summary(auto[, -c(4, 9)])
```

c.) What is the mean and standard deviation of each quantitative predictor ?

```{r, echo = T}
sapply(auto[, -c(4, 9)], mean)
sapply(auto[, -c(4, 9)], sd)
```

d.) Now remove the 10th through 85th observations. What is the range, mean, and standard deviation of each predictor in the subset of the data that remains ?

```{r, echo = T}
subset <- auto[-c(10:85), -c(4,9)]
sapply(subset, range)

sapply(subset, mean)

sapply(subset, sd)
```

e.) Using the full data set, investigate the predictors graphically, using scatterplots or other tools of your choice. Create some plots highlighting the relationships among the predictors. Comment on your findings.

```{r, echo = T, fig.width=8, fig.height=4}
auto$cylinders <- as.factor(auto$cylinders)
auto$year <- as.factor(auto$year)
auto$origin <- as.factor(auto$origin)

ggpairs(auto[, c("cylinders", "year", "origin")])
```

f.) Suppose that we wish to predict gas mileage (“mpg”) on the basis of other variables. Do your plots suggest that any of the other variables might be useful in predicting “mpg” ?

From the plots above, the cylinders, horsepower, year and origin can be used as predictors. Displacement and weight were not used because they are highly correlated with horespower and with each other.

```{r, echo = T}
auto$horsepower <- as.numeric(auto$horsepower)
cor(auto$weight, auto$horsepower)

cor(auto$weight, auto$displacement)

cor(auto$displacement, auto$horsepower)
```

#### 10.)

This exercise involves the “Boston” housing data set.

a.) To begin, load in the “Boston” data set.

```{r, echo = T}
boston <- MASS::Boston

boston$chas <- as.factor(boston$chas)
dim(Boston)
```

b.) Make some pairwise scatterplots of the predictors in this data set.

```{r, echo=T, fig.width=8, fig.height=4}

p1 <- ggplot(boston) +
   geom_point(aes(nox, crim))

p2 <- ggplot(boston) +
   geom_point(aes(rm, crim))

p3 <- ggplot(boston) +
   geom_point(aes(age, crim))

p4 <- ggplot(boston) +
   geom_point(aes(dis, crim))

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

c.) Are any of the predictors associated with per capita crime rate ?

```{r, echo = T, fig.width=8, fig.height=3.5}
ggplot(boston) +
   geom_histogram(aes(crim, fill = ..count..), bins = 40)
```

Most suburbs do not have any crime (80% of data falls in crim < 20).

```{r, echo = T, fig.width=8, fig.height=3.5}
ggpairs(boston[boston$crim < 20,])
```

d.) Do any of the suburbs of Boston appear to have particularly high crime rates ? Tax rates ? Pupil-teacher ratios ?

```{r, echo = T, fig.width=8, fig.height=3.5}
nrow(Boston[Boston$tax == 666, ])
```

e.) How many of the suburbs in this data set bound the Charles river ?

```{r, echo = T, fig.width=8, fig.height=3.5}
nrow(Boston[Boston$chas == 1, ])
```

f.) What is the median pupil-teacher ratio among the towns in this data set ?

```{r, echo = T, fig.width=8, fig.height=3.5}
median(Boston$ptratio)
```

g.) Which suburb of Boston has lowest median value of owner-occupied homes ? What are the values of the other predictors for that suburb, and how do those values compare to the overall ranges for those predictors ?

```{r, echo = T, fig.width=8, fig.height=3.5}
row.names(Boston[min(Boston$medv), ])
range(Boston$tax)
boston[min(boston$medv), ]$tax
```

h.) In this data set, how many of the suburbs average more than seven rooms per dwelling ? More than eight rooms per dwelling ?

```{r, echo = T, fig.width=8, fig.height=3.5}
row.names(Boston[min(boston$medv), ])

nrow(boston[boston$rm > 7, ])

nrow(boston[boston$rm > 8, ])
```

### 2020-02-19 ISLR/06_Model-Selection.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 6}
   \fancyfoot[C]{\rmfamily\color{headergrey}Model Selection}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(leaps, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(pls, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select
```

## Chapter 6

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Lab

```{r}
hitters <- as.data.table(ISLR::Hitters)

# ggpairs(hitters)
```

```{r}
sum(is.na(hitters$Salary))
full <- nrow(hitters)
```

```{r}
hitters <- hitters[ complete.cases(hitters), ]

1 - (nrow(hitters) / full) # drop about 18% of the data

regfit.full <- regsubsets(Salary ~., hitters)
summary(regfit.full)
```

```{r}
regfit.full <- regsubsets(Salary ~., data = hitters, nvmax = 19)
reg.summary <- summary(regfit.full)

names(reg.summary)
```

```{r}
reg.summary$rsq
```

```{r}
par(mfrow = c(2,2))
plot(reg.summary$rss, xlab = "Number of Variables", ylab = "RSS")
which.min(reg.summary$rss)
points(19, reg.summary$rss[19], col="red", cex=2, pch=20)
plot(reg.summary$adjr2, xlab = "Number of Variables", ylab = "Adj. R^2")
which.max(reg.summary$adjr2)
points(11, reg.summary$adjr2[11], col="red", cex=2, pch=20)
plot(reg.summary$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
which.min(reg.summary$cp)
points(10, reg.summary$cp[10], col="red", cex=2, pch=20)
which.min(reg.summary$bic)
plot(reg.summary$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
points(6, reg.summary$bic[6], col = "red", cex = 2, pch = 20)
```

```{r}
plot(regfit.full, scale = "r2")
plot(regfit.full, scale = "adjr2")
plot(regfit.full, scale = "Cp")
plot(regfit.full, scale = "bic")
```

```{r}
coef(regfit.full, 6)
```

```{r}
regfit.fwd <- regsubsets(Salary ~., data = hitters, nvmax = 19)
summary(regfit.full)
```

```{r}
regfit.bwd <- regsubsets(Salary ~., data = hitters, nvmax = 19)
summary(regfit.bwd)
```

Cross-Validation using best subset selection:

```{r}
train <- sample(c(T, F), size = nrow(hitters), replace = T)
test <- (!train)
```

```{r}
regfit.best <- regsubsets(Salary ~., data = hitters[train,], nvmax = 19)
test_mat <- model.matrix(Salary ~., data = hitters)

val_errors <- rep(NA, 19)
for(i in 1:19)
{
   coefi <- coef(regfit.best, id = i)
   pred <- test_mat[, names(coefi)] %*% coefi
   val_errors[i] <- mean((hitters$Salary[test] - pred)^2)
}

which.min(val_errors)
```

```{r}
predict_regsubsets <- function(object, newdata, id, ...)
{
   form <- as.formula(object$call[[2]])
   mat <- model.matrix(form, newdata)
   coefi <- coef(object, id = id)
   xvars <- names(coefi)
   mat[, xvars] %*% coefi
}

regfit.best <- regsubsets(Salary ~ ., data = hitters, nvmax = 19)
coef(regfit.best, 10)
```

```{r}
k <- 10
set.seed(1)
folds <- sample(1:k, nrow(hitters), replace = T)
cv.errors <- matrix(NA, k, 19, dimnames = list(NULL, paste(1:19)))

for(j in 1:k)
{
   best.fit <- regsubsets(Salary ~ ., data = hitters[folds != j, ], nvmax = 19)
   for(i in 1:19) {
      pred <- predict_regsubsets(best.fit, hitters[folds == j, ], id = i)
      cv.errors[j, i] <- mean( (hitters$Salary[folds == j] - pred)^2)
   }
}

mean.cv.errors <- apply(cv.errors, 2, mean)
mean.cv.errors

par(mfrow = c(1,1))
plot(mean.cv.errors, type ='b')
```

### Ridge Regression

```{r}
x <- model.matrix(Salary ~ ., hitters)[, -1]
y <- hitters$Salary

grid <- 10^seq(10, -2, length = 100)

ridge.mod <- glmnet(x, y, alpha = 0, lambda = grid)
plot(ridge.mod)

dim(coef(ridge.mod))
```

```{r}
ridge.mod$lambda[50]
coef(ridge.mod)[,50]

sqrt(sum(coef(ridge.mod)[-1, 50])^2)
```

```{r}
index <- 60
ridge.mod$lambda[index]
coef(ridge.mod)[, index]

sqrt(sum(coef(ridge.mod)[-1, index])^2)
```

```{r}
set.seed(1)
train <- sample(1:nrow(x), nrow(x)/2)
test <- (-train)

y.test <- y[test]

ridge.mod <- glmnet(x[train,], y[train], alpha = 0, lambda = grid, threshold = 1e-12)

ridge.pred <- predict(ridge.mod, s = 4, newx = x[test,])
mean((ridge.pred - y.test)^2)

```

### Lasso

```{r}
lasso.mod <- glmnet(x[train,], y[train], alpha = 1, lambda = grid)
plot(lasso.mod)
```

```{r}
set.seed(1)

cv.out <- cv.glmnet(x[train, ], y[train], alpha = 1)

plot(cv.out)

bestlam <- cv.out$lambda.min
lasso.pred <- predict(lasso.mod, s = bestlam, newx = x[test,])

mean((lasso.pred - y.test)^2)
```

```{r}
out <- glmnet(x, y, alpha = 1, lambda = grid)
lasso.coef <- predict(out, type = "coefficients", s = bestlam)[1:20,]
lasso.coef
```

### Principal Components Regression

```{r}
set.seed(2)

pcr.fit <- pcr(Salary ~., data = hitters, scale = T, validation = "CV")

summary(pcr.fit)

validationplot(pcr.fit, val.type = "MSEP")
```

```{r}
set.seed(1)

pcr.fit <- pcr(Salary ~ ., data = hitters, subset = train, scale = T,
               validation = "CV")
validationplot(pcr.fit, val.type = "MSEP")
```

```{r}
pcr.pred <- predict(pcr.fit, x[test,], ncomp = 7)
mean((pcr.pred - y.test)^2)
```

### Partial Least Squares

```{r}
set.seed(1)

pls.fit <- plsr(Salary ~ ., data = hitters, subset = train, scale = T,
                validation = "CV")

summary(pls.fit)

pls.pred <- predict(pls.fit, x[test, ], ncomp = 2)
mean((pls.pred - y.test)^2)

pls.fit <- plsr(Salary ~ ., data = hitters, scale = T, ncomp = 2)
summary(pls.fit)
```

## Conceptual

### 1.)

We perform best subset, forward stepwise, and backward stepwise selection on a single data set. For each approach, we obtain $p + 1$ models, containing $0, 1, 2, \ldots, p$ predictors.

a.) Which of the three models with k predictors has the smallest *training* RSS?

_The model with the smallest training RSS will be the $C{k \choose p}$ model with p = k._

b.) Which of the three models with k predictors has the smallest *test* RSS?

_Difficult to say. The best subset selection technique looks at more models, however, forward or backward selection could pick a better model by chance._

c.) T/F

i.) The predictors in the k-variable model identified by foward stepwise selection are a subset of the predictors in the (k+1)-variable model identified by forward stepwise selection.

__T__

ii.) The predictors isn the k-variable model identified by backward stepwise are a subset of the predictors in the (k+1)-variable model identified by backward stepwise selection.

__T__

iii.) The predictors in the k-variable model identified by backward stepwise are a subset of the predictors in the (k+1)-variable model identified by forward stepwise selection.

__F__. There is no link between these models predictors.

iv.) The predictors in the k-variable model identifed by forward stepwise are a subset of the predictors in the (k+1)-variable model identifed by backward stepwise selection.

__F__. There is no link between these models predictors.

v.) The predictors in the k-variable model identified by best subset are a subset of the predictors in the (k+1)-variable model identifed by best subset selection.

__F__. There is no link between these models predictors.

### 2.)

For parts (a) through (c), indicate which of i through iv. is correct.

a.) The lasso, relative to least squares, is:

i.) More flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance.

__T__

ii.) More flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias.

__F__

iii.) Less flexible and hence will give improved prediction accuracyt when its increase in bias is less than its decrease in variance.

__F__

iv.) Less flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias.

__F__

### 3.)

Suppose we estimate the regression coefficients in a linear regression model by minimizing:

$\sum^n_{i=1}{(y_i - \beta_0 - \sum^p_{j=1}\beta_jx_{i,j})}$

subject to,

$\sum^p_{j=1}(|\beta_j|) \le s$

a.) As we increase s from 0, the training RSS will:

Steadily decrease. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes a steady decrease in the training RSS.

b.) Test RSS will:

Decrease initially, and then eventually start increasing in a U shape. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes at first a decrease in the test RSS before increasing again after that in a typical U shape.

c.) Variance will:

Steadily increase. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes a steady increase in variance.

d.) (squared) bias will:

Steadily decrease. As we increase s from 0, we are restricting the βj coefficients less and less (the coefficients will increase to their least squares estimates), and so the model is becoming more and more flexible which provokes a steady decrease in bias.

e.) irreducible error will:

remain unchanged.

### 4.)

Suppose we estimate the regression coefficients in a linear regression model by minimizing:

$\sum^n_{i=1}(y_i - \beta_0 - \sum^p_{j=1}\beta_jx_{ij}) + \lambda\sum^p_{j=1}\beta^2_j$

for a particular value of $\lambda$.

a.) As we increase $\lambda$ from 0, the training RSS will:

Steadily increase.

b.) Test RSS will:

Decrease initially, then turn to a U shape.

c.) Variance will:

Steadly decrease.

d.) (squared) bias will:

Steadily increase

e.) Irreducible error will:

Remain constant.

### 5.)

It is well-known that ridge regression tends to give similar coefficient values to correlated variables, whereas the lasso may give quite different coefficient values to correlated variables. We will now explore this property in a very simple setting.

Suppose that $n = 2, p = 2, x_{11} = x_{12} = x_{21} = x_{22}$. Furthermore, suppose that $y_1 + y_2 = 0$ and $x_{11} + x_{21} = 0$ so the estimate for the intercept in a least squares, ridge regresion or lasso model is zero: $\hat{\beta}_0 = 0$

Write out the ridge regression optimization problem in this setting:

$(y_1 - \hat{\beta_1}x_1 - \hat{\beta_2})^2 + (y_2 - \hat{\beta_1}x_2 - \hat{\beta_2}x_2)^2 + \lambda(\hat{\beta^2_1} + \hat{\beta^2_2})$

### 6.)

We will now explore (6.12) and (6.13) further.

a.) Consider (6.12) with p=1. For some choice of y1 and λ>0, plot (6.12) as a function of β1. Your plot should confirm that (6.12) is solved by (6.14).

```{r}
y <- 3
lambda <- 2
beta <- seq(-10, 10, 0.1)
plot(beta, (y - beta)^2 + lambda * beta^2, pch = 20, xlab = "beta", ylab = "Ridge optimization")
beta.est <- y / (1 + lambda)
points(beta.est, (y - beta.est)^2 + lambda * beta.est^2, col = "red", pch = 4, lwd = 5)
```

We may see that the function is minimized at β=y/(1+λ).

b.) Consider (6.13) with p=1. For some choice of y1 and λ>0, plot (6.13) as a function of β1. Your plot should confirm that (6.13) is solved by (6.15).

```{r}
y <- 3
lambda <- 2
beta <- seq(-10, 10, 0.1)
plot(beta, (y - beta)^2 + lambda * abs(beta), pch = 20, xlab = "beta", ylab = "Lasso optimization")
beta.est <- y - lambda / 2
points(beta.est, (y - beta.est)^2 + lambda * abs(beta.est), col = "red", pch = 4, lwd = 5)
```

We may see that the function is minimized at β=y−λ/2 as y>λ/2.

## Applied

### 8.)

In this exercise, we will generate simulated data, and will then use this data to perform best subset selection.

a.) Use the rnorm() function to generate a predictor X of length n=100, as well as a noise vector ε of length n=100.

```{r}
set.seed(1)

n <- 100

x = rnorm(n)
eps = rnorm(n)
```

b.) Generate a response vector Y of length n=100 according to the model

Y=β0+β1X+β2X2+β3X3+ε

where β0, β1, β2 and β3 are constants of your choice.

```{r}
b0 <- 2
b1 <- 3
b2 <- 0.3
b3 <- -1

y <- b0 + b1 * x + b2 * x^2 + b3 * x^3 + eps
```

c.) Use the regsubsets() function to perform best subset selection in order to choose the best model containing the predictors X,X2,⋯,X10. What is the best model obtained according to Cp, BIC, and adjusted R2 ? Show some plots to provide evidence for your answer, and report the coefficients of the best model obtained. Note you will need to use the data.frame() function to create a single data set containing both X and Y.

```{r}
data.full <- data.frame(y = y, x = x)
regfit.full <- regsubsets(y ~ x + I(x^2) + I(x^3) + I(x^4) + I(x^5) + I(x^6) + I(x^7) + I(x^8) + I(x^9) + I(x^10), data = data.full, nvmax = 10)
reg.summary <- summary(regfit.full)

par(mfrow = c(2, 2))
plot(reg.summary$cp, xlab = "Number of variables", ylab = "C_p", type = "l")
points(which.min(reg.summary$cp), reg.summary$cp[which.min(reg.summary$cp)], col = "red", cex = 2, pch = 20)
plot(reg.summary$bic, xlab = "Number of variables", ylab = "BIC", type = "l")
points(which.min(reg.summary$bic), reg.summary$bic[which.min(reg.summary$bic)], col = "red", cex = 2, pch = 20)
plot(reg.summary$adjr2, xlab = "Number of variables", ylab = "Adjusted R^2", type = "l")
points(which.max(reg.summary$adjr2), reg.summary$adjr2[which.max(reg.summary$adjr2)], col = "red", cex = 2, pch = 20)
```

d.) Repeat (c), using forward stepwise selection and also using backwards stepwise selection. How does your answer compare to the results in (c) ?

```{r}
regfit.fwd <- regsubsets(y ~ x + I(x^2) + I(x^3) + I(x^4) + I(x^5) + I(x^6) + I(x^7) + I(x^8) + I(x^9) + I(x^10), data = data.full, nvmax = 10, method = "forward")
reg.summary.fwd <- summary(regfit.fwd)

par(mfrow = c(2, 2))
plot(reg.summary.fwd$cp, xlab = "Number of variables", ylab = "C_p", type = "l")
points(which.min(reg.summary.fwd$cp), reg.summary.fwd$cp[which.min(reg.summary.fwd$cp)], col = "red", cex = 2, pch = 20)
plot(reg.summary.fwd$bic, xlab = "Number of variables", ylab = "BIC", type = "l")
points(which.min(reg.summary.fwd$bic), reg.summary.fwd$bic[which.min(reg.summary.fwd$bic)], col = "red", cex = 2, pch = 20)
plot(reg.summary.fwd$adjr2, xlab = "Number of variables", ylab = "Adjusted R^2", type = "l")
points(which.max(reg.summary.fwd$adjr2), reg.summary.fwd$adjr2[which.max(reg.summary.fwd$adjr2)], col = "red", cex = 2, pch = 20)
mtext("Plots of C_p, BIC and adjusted R^2 for forward stepwise selection", side = 3, line = -2, outer = TRUE)
```

e.) Now fit a lasso model to the simulated data, again using $X, X^2, \ldots, x^{10}$

```{r}
xmat <- model.matrix(y ~ x + I(x^2) + I(x^3) + I(x^4) + I(x^5) + I(x^6) + I(x^7) + I(x^8) + I(x^9) + I(x^10), data = data.full)

cv.lasso <- cv.glmnet(xmat, y, alpha = 1)

par(mfrow = c(1, 1))
plot(cv.lasso)

bestlam <- cv.lasso$lambda.min
bestlam

fit.lasso <- glmnet(xmat, y, alpha = 1)
predict(fit.lasso, s = bestlam, type = "coefficients")[1:11, ]
```

The lasso method picks X, X2, X3 and X5 as variables for the model.

#### 9.)

In this exercise, we will predict the number of applications recieved using the other variables in the **College** data set.

a.) Split the data into a training set and a test set.

```{r}
college <- as.data.table(ISLR::College)

unif <- runif(nrow(college))

train <- unif < .7
test <- !(train)

train <- college[train]
test <- college[test]
```

b.) Fit a linear model using least squares on the training set, and repor the test error obtained.

```{r}
lm.fit <- lm(Apps ~ ., data = train)

mean((test$Apps - predict(lm.fit, newdata = test))^2)
```

c.) Fit a ridge regression on the training set, with $\lambda$ chosen by cross-validation. Report the error.

```{r}
train.mat <- model.matrix(Apps ~ ., data = train)

cv.ridge <- cv.glmnet(train.mat, train$Apps, data = train, alpha = 0)

bestlam <- cv.ridge$lambda.min
bestlam

test.mat <- model.matrix(Apps ~ ., data = test)

fit.ridge <- glmnet(test.mat, test$Apps, alpha = 1, lambda = bestlam)
pred <- predict(fit.ridge, s = bestlam, newx = test.mat, type = "response")

mean( (test$Apps - pred )^2 )

```

d.) Fit a lasso regresion on the training set, with $\lambda$ chosen by cross-validation. Report the error.

```{r}
cv.lasso <- cv.glmnet(train.mat, train$Apps, data = train, alpha = 1)

bestlam <- cv.lasso$lambda.min
bestlam

fit.lasso <- glmnet(test.mat, test$Apps, alpha = 1, lambda = bestlam)

pred <- predict(fit.lasso, test.mat, s = bestlam)

mean( (test$Apps - pred)^2)
```

e.) Fit a PCR model on the training set, with M chosen by cross-validation. Report the test error, along with the value of M.

```{r}
fit.pcr <- pcr(Apps ~ ., data = train, scale = T, validation = "CV")
validationplot(fit.pcr)

fit.pcr$ncomp

pred.pcr <- predict(fit.pcr, test, ncomp = 10)
mean((pred.pcr - test$Apps)^2)
```

f.) Fit a PLS model on the training set, with M chosen by cross-validation. Report the test error and M.

```{r}
pls.fit <- plsr(Apps ~ ., data = train, scale = T,
                validation = "CV")

validationplot(pls.fit)

pred.pls <- predict(pls.fit, test, ncomp = 10)
mean((pred.pls - test$Apps)^2)
```

#### 10.)

We have seen that as the number of features used in a model increases, the training error will necessarily decrease, but the test error may not. We will now explore this in a simulated data set.

a.) Generate a data set with p = 20 features, n = 1,000 observations, and an associated quantitative response vector generated according to the model:

$Y = X\beta + \epsilon$

```{r}
set.seed(1)
x <- matrix(rnorm(1000 * 20), 1000, 20)
b <- rnorm(20)
b[3] <- 0
b[4] <- 0
b[9] <- 0
b[19] <- 0
b[10] <- 0
eps <- rnorm(1000)
y <- x %*% b + eps
```

b.) Split your data set into a training set containing 100 observations and a test set containing 900 observations.

```{r}
train <- sample(seq(1000), 100, replace = FALSE)
test <- -train
x.train <- x[train, ]
x.test <- x[test, ]
y.train <- y[train]
y.test <- y[test]
```

c.) Perform best subset selection on the training set, and plot the training set MSE associated with the best model of each size.

```{r}
data.train <- data.frame(y = y.train, x = x.train)
regfit.full <- regsubsets(y ~ ., data = data.train, nvmax = 20)
train.mat <- model.matrix(y ~ ., data = data.train, nvmax = 20)
val.errors <- rep(NA, 20)
for (i in 1:20) {
    coefi <- coef(regfit.full, id = i)
    pred <- train.mat[, names(coefi)] %*% coefi
    val.errors[i] <- mean((pred - y.train)^2)
}
plot(val.errors, xlab = "Number of predictors", ylab = "Training MSE", pch = 19, type = "b")
```

d.) Plot the test MSE associated with the best model of each size.

```{r}
data.test <- data.frame(y = y.test, x = x.test)
test.mat <- model.matrix(y ~ ., data = data.test, nvmax = 20)
val.errors <- rep(NA, 20)
for (i in 1:20) {
    coefi <- coef(regfit.full, id = i)
    pred <- test.mat[, names(coefi)] %*% coefi
    val.errors[i] <- mean((pred - y.test)^2)
}
plot(val.errors, xlab = "Number of predictors", ylab = "Test MSE", pch = 19, type = "b")
```

e.) For which model size does the test set MSE take on its minimum value ? Comment on your results. It it takes on its minimum value for a model containing only an intercept or a model containing all the features, then play around with the way that you are generating the data in (a) until you come up with a scenario in which the test MSE is minimized for an intermediate model size.

```{r}
which.min(val.errors)
```

f.) How does the model at which the test set MSE is minimized compare to the true model used to generate the data ? Comment on the coefficient values.

```{r}
coef(regfit.full, which.min(val.errors))
```

g.) Create a plot displaying:

$\sqrt({\sum^p_{j=1}(\beta_j - \hat{\beta}_j^r)^2})$

```{r}
val.errors <- rep(NA, 20)
x_cols = colnames(x, do.NULL = FALSE, prefix = "x.")
for (i in 1:20) {
    coefi <- coef(regfit.full, id = i)
    val.errors[i] <- sqrt(sum((b[x_cols %in% names(coefi)] - coefi[names(coefi) %in% x_cols])^2) + sum(b[!(x_cols %in% names(coefi))])^2)
}
plot(val.errors, xlab = "Number of coefficients", ylab = "Error between estimated and true coefficients", pch = 19, type = "b")
```

#### 11.)

We will now try to predict per capita crime rate in the “Boston” data set.

a.) Try out some of the regression methods explored in this chapter, such as best subset selection, the lasso, ridge regression and PCR. Present and discuss results for the approaches that you consider.

```{r}
data(Boston)
set.seed(1)

predict.regsubsets <- function(object, newdata, id, ...) {
    form <- as.formula(object$call[[2]])
    mat <- model.matrix(form, newdata)
    coefi <- coef(object, id = id)
    xvars <- names(coefi)
    mat[, xvars] %*% coefi
}

k = 10
folds <- sample(1:k, nrow(Boston), replace = TRUE)
cv.errors <- matrix(NA, k, 13, dimnames = list(NULL, paste(1:13)))
for (j in 1:k) {
    best.fit <- regsubsets(crim ~ ., data = Boston[folds != j, ], nvmax = 13)
    for (i in 1:13) {
        pred <- predict(best.fit, Boston[folds == j, ], id = i)
        cv.errors[j, i] <- mean((Boston$crim[folds == j] - pred)^2)
    }
}

mean.cv.errors <- apply(cv.errors, 2, mean)
plot(mean.cv.errors, type = "b", xlab = "Number of variables", ylab = "CV error")
```

```{r}
x <- model.matrix(crim ~ ., Boston)[, -1]
y <- Boston$crim
cv.out <- cv.glmnet(x, y, alpha = 1, type.measure = "mse")
plot(cv.out)
```

```{r}
cv.out <- cv.glmnet(x, y, alpha = 0, type.measure = "mse")
plot(cv.out)
```

### 2020-02-19 ISLR/08_Tree-Based-Methods.rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 8}
   \fancyfoot[C]{\rmfamily\color{headergrey}Tree Based Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(tree, quietly = TRUE, warn.conflicts = FALSE)
library(randomForest, quietly = TRUE, warn.conflicts = FALSE)
library(gbm, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select

```

## Chapter 8

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Lab

### Fitting Classification Trees

```{r}
carseats <- as.data.table(ISLR::Carseats)

carseats[, High := as.factor(ifelse(Sales <= 8, "No", "Yes"))]

tree.carseats <- tree(formula = High ~ .-Sales, data = carseats)

summary(tree.carseats)

plot(tree.carseats)
text(tree.carseats, pretty = 0)
```

```{r}
set.seed(2)

train <- sample(1:nrow(carseats), 200)

carseats.test <- carseats[-train]
high.test <- carseats[-train]$High

tree.carseats <- tree(High ~.-Sales, data = carseats, subset = train)
tree.pred <- predict(tree.carseats, carseats.test, type = "class")

table(tree.pred, high.test)
```

```{r}
set.seed(3)

cv.carseats <- cv.tree(tree.carseats, FUN = prune.misclass)
names(cv.carseats)

cv.carseats
```

```{r}
par(mfrow = c(1, 2))
plot(cv.carseats$size, cv.carseats$dev, type = "b")
plot(cv.carseats$k, cv.carseats$dev, type="b")
```

```{r}
par(mfrow = c(1,1))
prune.carseats <- prune.misclass(tree.carseats, best = 9)
plot(prune.carseats)
text(prune.carseats, pretty = 0)
```

```{r}
tree.pred <- predict(prune.carseats, carseats.test, type = "class")
table(tree.pred, high.test)
```

```{r}
prune.carseats <- prune.misclass(tree.carseats, best = 15)
plot(prune.carseats)
text(prune.carseats, pretty = 0)

tree.pred <- predict(prune.carseats, carseats.test, type = "class")
table(tree.pred, high.test)
```

### Regression Trees

```{r}

boston <- as.data.table(Boston)

N <- nrow(boston)

set.seed(1)

train <- sample(1:N, N /2)

tree.boston <- tree(medv ~ ., boston, subset = train)

summary(tree.boston)
```

```{r}
plot(tree.boston)
text(tree.boston, pretty = 0)
```

```{r}
cv.boston <- cv.tree(tree.boston)
plot(cv.boston$size, cv.boston$dev, type = 'b')
```

```{r}
prune.boston <- prune.tree(tree.boston, best = 5)
plot(prune.boston)
text(prune.boston, pretty = 0)
```

```{r}
yhat <- predict(tree.boston, newdata = boston[-train,])
boston.test <- boston[-train]$medv

plot(yhat, boston.test)
abline(0, 1)
mean((yhat - boston.test)^2)
```

### Bagging and Boosting

```{r}
set.seed(1)

bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, importance = T)
bag.boston
```

```{r}
yhat.bag <- predict(bag.boston, newdata = boston[-train])
plot(yhat.bag, boston.test)
abline(0, 1)
mean((yhat.bag-boston.test)^2)
```

```{r}
bag.boston <- randomForest(medv ~ ., data = boston, subset = train, mtry = 13, ntree = 25)
yhat.bag <- predict(bag.boston, newdata = boston[-train,])
mean((yhat.bag - boston.test)^2)
```

```{r}
set.seed(1)

rf.boston <- randomForest(medv ~  ., data = boston, subset = train, mtry = 6, importance = T)
yhat.rf <- predict(rf.boston, newdata = boston[-train,])
mean((yhat.rf-boston.test)^2)
```

```{r}
importance(rf.boston)
```

```{r}
varImpPlot(rf.boston)
```

### Boosting

```{r}
boost.boston <- gbm(medv ~ ., data = boston[-train,], distribution = "gaussian", n.trees = 5000, interaction.depth = 4)

summary(boost.boston)
```

```{r}
par(mfrow = c(1, 2))
plot(boost.boston, i = "rm")
plot(boost.boston, i = "lstat")
```

```{r}
yhat.boost <- predict(boost.boston, newdata = boston[-train,], n.trees = 5000)
mean((yhat.boost - boston.test)^2)
```

## Applied

In the lab, we applied random forests to the *Boston* data using mtry = 6 and using ntree = 25 and ntree = 500. Create a plot displaying the test error resulting from random forests on this data set for a more comprehensive range of values for mtry ntree.

```{r}
N <- nrow(boston)

train <- sample(1:N, N * .7)

boston.train <- boston[train]
boston.test <- boston[!train]

ntrees <- seq(0, 500, 10)

results <- numeric(length(ntrees))

for(i in 1:length(ntrees))
{
   tree.count <- ntrees[i]

   rf <- randomForest(medv ~ ., data = boston.train, mtry = 6, n.trees = tree.count)

   pred <- predict(rf, newdata = boston.test)

   results[i] <- sqrt(mean((pred - boston.test$medv)^2)) # store the rmse
}

lowest.error <- which.min(results)

rf.results <- data.table(trees = ntrees, errors = results)[, lowest := .I == lowest.error]

ggplot(rf.results, aes(trees, errors, fill = lowest)) +
   geom_line()

```

In the lab, a classification tree was applied to the Carseats data set after converting Sales into a qualitative response variable. Now we will seek to predict Sales using regression trees and related approaches, treating the response as a quantitaive variable.

a.) Split the data into a t training set and a test set.

```{r}
N <- nrow(carseats)

index <- sample(1:N, N * .7)

carseats.train <- carseats[index]
carseats.test <- carseats[!index]
```

b.) Fit a regression tree to the training set. Plot the tree, and interpret the results. What test MSE do you obtain?

```{r}
carseat.tree <- tree(Sales ~., data = carseats.train)
pred <- predict(carseat.tree, newdata = carseats.test)
mse <- mean((pred - carseats.test$Sales)^2)

mse

par(mfrow = c(1,1))
plot(rf)
text(rf, pretty = 0)
```

c.) Use cross-validation in order to determine the optimal level of tree complexity. Does pruning the tree improve the test MSE?

```{r}
carseat.cv <- cv.tree(carseat.tree)

plot(carseat.cv$size, carseat.cv$dev, type = "b", col = "red")
plot(carseat.cv$size, carseat.cv$k, type = "b", col = "red")

prune.carseats <- prune.tree(carseat.tree, best = 5)

plot(prune.carseats)
text(prune.carseats)
```

d.) Use the bagging approach in order to analyze this data. What test MSE do you obtain? Use the importance() function to determine which variables are most important.

```{r}

```

```{r}

```

### 2020-02-19 Mathematical Statistics/06_Estimation-Notes.Rmd
merged machine learning
merged machine learning

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Estimation}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 6}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematical Statistics}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
library(ggplot2)
library(dplyr)
library(stats)
```

### Wind speed case study: fitting the Weibull distribution

This function takes input shape parameter k and  the data to compute
 $(1/k)+ (1/n)*\sum (\log(x_i)) +(1/\alpha)\sum x_i^k\log(x_i)=0$  where $\alpha= \sum x_i^k$.

```{r}
weibull.shape <- function(k, data)
{
  numer <- colSums(outer(data, k, "^") * log(data))
  denom <- colSums(outer(data, k, "^"))
  numer/denom - 1/k - mean(log(data))
}
```

 This function takes input shape parameter k
 and data to compute   $k^{th}$ root of $(1/n) \sum x_i^k$
 where $n=$number of data values.

```{r}
weibull.scale <- function(k, data)
{
  mean(data^k)^(1/k)
}
```

 <tt>`uniroot`</tt> is a built-in function which estimates the roots of a function.
 Provide function, any arguments needed for function,
 and a guess of values two values around root.
 Function values must be opposite signs at lower
 and upper guess.

Now, we do the data specific commands:

```{r}
Turbine <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Turbine.csv")
wind <- Turbine$AveSpeed

#alternatively, wind <- subset(Turbine, select = AveSpeed, drop = TRUE)
```
Estimate the shape parameter k:
```{r}
uniroot(weibull.shape, data = wind, lower = 1,upper = 5)
```
With estimate of shape parameter, now find estimate of scale parameters lambda:

```{r}
weibull.scale(3.169, wind)
```

Next, plot histogram with density curve overlap.
The <tt>`prob = TRUE`</tt> argument scales histogram to area 1.

```{r}
myfun <- function(x){
  dweibull(x, 3.169, 7.661)
}

df <- data.frame(wind)
ggplot(df, aes(wind)) + geom_histogram(aes(y = stat(density)), bins=12) + labs(title="Distribution of average wind speeds", x = "meters/sec") + stat_function(fun = myfun, colour="blue")

ggplot(df, aes(wind)) + stat_ecdf() + stat_function(fun=pweibull, args=list(shape=3.169, scale = 7.661), colour = "blue") + labs(title = "ECDF of wind data")
```

###Goodness-of-fit

Now for the chi-square goodness of fit test
Get the deciles of the Weibull distribution with the given parameters and also, the range of the wind data:

```{r}
q <- qweibull(seq(.1, .9, by = .1), 3.169, 7.661)

range(wind)
```
Now, encompass the range of wind and get the counts in each sub-interval. The
<tt>`plot = FALSE`</tt> argument suppresses the histogram and just gives statistics:
```{r}
q <- c(0, q, 14)

hist(wind, breaks = q, plot = F)$counts

#Now store the output
count <- hist(wind,breaks = q, plot = F)$counts

expected <- length(wind)*.1
```
Compute chi-square test statistic:
```{r}
(count-expected)^2/expected
```

### Example 6.13
Simulation comparing two estimators for uniform:

```{r}
Xbar <- numeric(1000)
maxY <- numeric(1000)

#set.seed(100)
for (i in 1:1000)
 {
   x <- runif(25,0,12) #sample n=25 from Unif[0,12]
   Xbar[i] <- 2*mean(x)
   maxY[i] <- 26/25*max(x)
 }

 #mean and standard deviation of the method of moments estimate
 mean(Xbar)
 sd(Xbar)
range(Xbar)  
 #mean and sd of estimate from MLE
 mean(maxY)
 sd(maxY)
 range(maxY)
```

To plot the results:

```{r}
library(gridExtra)

p1 <- ggplot() + geom_histogram(aes(Xbar), breaks=seq(7, 16.5, by =.5))  + ylim(c(0, 650)) + labs(title = "2*Sample mean", x = "means")
p2 <- ggplot() + geom_histogram(aes(maxY), breaks=seq(7, 16.5, by = .5)) + ylim(c(0, 650)) +labs(title = "25/24*maximum", x = "maximums")
grid.arrange(p1, p2)

```

### Example 6.14

Relative efficiency for the Wind Speed Case Study
```{r}
# Theta = P(wind > 5), nonparametric and parametric estimates
n <- length(wind)
theta1 <- mean(wind > 5)
theta2 <- pweibull(5, 3.169, 7.661, lower.tail = FALSE)
theta1 # 0.75
theta2 # 0.7720827
eta1 <- quantile(wind, .1)
eta2 <- qweibull(.1, 3.169, 7.661)
eta1 # 3.77
eta2 # 3.766038

# nonparametric bootstrap to find standard errors
set.seed(23)
N <- 10^4
boot.theta1 <- numeric(N)
boot.shape <-  numeric(N)
boot.scale <-  numeric(N)
boot.theta2 <- numeric(N)
boot.eta1 <- numeric(N)
boot.eta2 <- numeric(N)
for (i in 1:N)
{
  boot.wind <- wind[sample(1:n, replace=TRUE)]
  boot.theta1[i] <- mean(boot.wind > 5)
  boot.shape[i] <- uniroot(weibull.shape, data=boot.wind, lower=1,upper=5)$root
  boot.scale[i] <- weibull.scale(boot.shape[i], boot.wind)
  boot.theta2[i] <- pweibull(5, boot.shape[i], boot.scale[i], lower.t = FALSE)
  boot.eta1[i] <- quantile(boot.wind, .1)
  boot.eta2[i] <- qweibull(.1, boot.shape[i], boot.scale[i])
}

ggplot() + stat_qq(aes(sample = boot.theta1)) # discrete, normal
ggplot() + stat_qq(aes(sample = boot.theta2)) # continuous, normal
ggplot() + stat_qq(aes(sample = boot.eta1)) # discrete, irregular, roughly normal
ggplot() + stat_qq(aes(sample = boot.eta2)) # continuous, very normal

sd(boot.theta1) # 0.03331767
sd(boot.theta2) # 0.02342771
var(boot.theta1) / var(boot.theta2) # 2.022504

# Formula standard error, for comparison
sqrt(theta1 * (1-theta1) / n) # standard error = 0.03340766

sd(boot.eta1) # 0.2161054
sd(boot.eta2) # 0.1839509
var(boot.eta1) / var(boot.eta2) # 1.380154

# parametric bootstrap to find standard errors
set.seed(23)
pboot.theta1 <- numeric(N)
pboot.shape <-  numeric(N)
pboot.scale <-  numeric(N)
pboot.theta2 <- numeric(N)
pboot.eta1 <- numeric(N)
pboot.eta2 <- numeric(N)
for (i in 1:N)
{
  boot.wind <- rweibull(n, 3.169, 7.661)
  pboot.theta1[i] <- mean(boot.wind > 5)
  pboot.shape[i] <- uniroot(weibull.shape, data=boot.wind, lower=1,upper=5)$root
  pboot.scale[i] <- weibull.scale(pboot.shape[i], boot.wind)
  pboot.theta2[i] <- pweibull(5, pboot.shape[i], pboot.scale[i], lower.t=FALSE)
  pboot.eta1[i] <- quantile(boot.wind, .1)
  pboot.eta2[i] <- qweibull(.1, pboot.shape[i], pboot.scale[i])
}
etaRange <- range(boot.eta1, boot.eta2, pboot.eta1, pboot.eta2)

ggplot() + stat_qq(aes(sample = pboot.theta1)) # discrete, slight neg skewness
ggplot() + stat_qq(aes(sample = pboot.theta2)) # continuous, normal
ggplot() + stat_qq(aes(sample = pboot.eta1)) # very normal, continuous
ggplot() + stat_qq(aes(sample = pboot.eta2)) # very normal, continuous

sd(pboot.theta1) # 0.03198990
sd(pboot.theta2) # 0.02604409
var(pboot.theta1) / var(pboot.theta2) # 1.508716
# Similar to the nonparametric bootstrap, though relative efficiency differs

sd(pboot.eta1) # 0.2823103
sd(pboot.eta2) # 0.2100147
var(pboot.eta1) / var(pboot.eta2) # 1.806982
```

```{r}

```

### 2020-02-19 h2o.R

@@ -1,0 +1,16 @@
+# The following two commands remove any previously installed H2O packages for R.
+if ("package:h2o" %in% search()) { detach("package:h2o", unload=TRUE) }
+if ("h2o" %in% rownames(installed.packages())) { remove.packages("h2o") }
+
+# Next, we download packages that H2O depends on.
+pkgs <- c("RCurl","jsonlite")
+for (pkg in pkgs) {
+  if (! (pkg %in% rownames(installed.packages()))) { install.packages(pkg) }
+}
+
+# Now we download, install and initialize the H2O package for R.
+install.packages("h2o", type="source", repos="http://h2o-release.s3.amazonaws.com/h2o/rel-yu/1/R")
+
+# Finally, let's load H2O and start up an H2O cluster
+library(h2o)
+h2o.init()

### 2020-02-19 keras.R

@@ -1,0 +1,2 @@
+devtools::install_github("rstudio/keras")
+install_tensorflow(gpu = T)

### 2020-02-19 ISLR/08_Tree-Based-Methods.rmd
tree work
tree work

@@
 par(mfrow = c(1,1))
-plot(rf)
-text(rf, pretty = 0)
+plot(carseat.tree)
+text(carseat.tree, pretty = 0)
 ```
@@
 ```{r}
+p <- ncol(carseats)
+
+ntrees <- seq(0, 500, 25)
+rf.error <- numeric(length(ntrees))
+
+for(i in 1:length(ntrees))
+{
+   carseats.rf <- randomForest(Sales ~ ., data = carseats.train, mtry = p, n.trees = ntrees[i])
+
+   pred <- predict(carseats.rf, newdata = carseats.test)
+
+   rf.error[i] <- mean((pred - carseats.test$Sales)^2)
+}
+
+lowest.error <- which.min(rf.error)
+
+results <- data.table(trees = ntrees, error = rf.error)[, lowest := .I == lowest.error]
+
+ggplot(results, aes(trees, error, fill = lowest)) +
+   geom_bar(stat = "identity") +
+   labs(title = paste0("lowest error: #", ntrees[lowest.error]))
 
@@
 
+e.) Use random forest to analyize this data.
+What test MSE do you obtain?
+Use the importance() fucntion to determine which variables are most important.
+Describe the effect of m, the number of variables considered at each split, on the error obtained.
+
 ```{r}
+p <- ncol(carseats)
 
+mtry <- seq(1, p)
+
+rf.error <- numeric(p)
+
+for(i in 1:p)
+{
+   carseats.rf <- randomForest(Sales ~ ., data = carseats.train, mtry = mtry[i], n.trees = 250)
+
+   pred <- predict(carseats.rf, newdata = carseats.test)
+
+   rf.error[i] <- mean((pred - carseats.test$Sales)^2)
+}
+
+lowest.error <- which.min(rf.error)
+
+results <- data.table(mtry = mtry, error = rf.error, lowest = mtry == lowest.error)
+
+ggplot(results, aes(mtry, error, fill = lowest)) +
+   geom_bar(stat = "identity")
 ```

### 2020-02-22 ISLR/08_Tree-Based-Methods.rmd
wrapped up trees
wrapped up trees

@@
 library(gbm, quietly = TRUE, warn.conflicts = FALSE)
+library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
+library(vip, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 ```{r}
-yhat.bag <- predict(bag.boston, newdata = boston[-train])
-plot(yhat.bag, boston.test)
+boston.test <- boston[!train]
+yhat.bag <- predict(bag.boston, newdata = boston.test)
+plot(yhat.bag, boston.test$medv)
 abline(0, 1)
-mean((yhat.bag-boston.test)^2)
+mean((yhat.bag-boston.test$medv)^2)
 ```
@@
 yhat.bag <- predict(bag.boston, newdata = boston[-train,])
-mean((yhat.bag - boston.test)^2)
+mean((yhat.bag - boston.test$medv)^2)
 ```
@@
 ntrees <- seq(0, 500, 10)
+n.features <- ncol(boston[, !"medv"])
 
-results <- numeric(length(ntrees))
+mtry <- round(c(n.features, n.features/2, sqrt(n.features)))
+
+results <- matrix(vector(mode = "numeric", length(ntrees) * length(mtry)), nrow = 3)
 
@@
 {
+   for(j in 1:length(mtry))
+   {
       tree.count <- ntrees[i]
 
-   rf <- randomForest(medv ~ ., data = boston.train, mtry = 6, n.trees = tree.count)
+      rf <- randomForest(medv ~ ., data = boston.train, mtry = mtry[j], n.trees = tree.count)
 
@@
 
-   results[i] <- sqrt(mean((pred - boston.test$medv)^2)) # store the rmse
+      results[j, i] <- sqrt(mean((pred - boston.test$medv)^2)) # store the rmse
+   }
 }
 
-lowest.error <- which.min(results)
+df_results <- as.data.table(results)
+colnames(df_results) <- paste0(ntrees)
 
-rf.results <- data.table(trees = ntrees, errors = results)[, lowest := .I == lowest.error]
+df_results <- melt(cbind(mtry, df_results), id.vars = "mtry")
+df_results$mtry <- as.factor(df_results$mtry)
 
-ggplot(rf.results, aes(trees, errors, fill = lowest)) +
-   geom_line()
+ggplot(df_results, aes(variable, value, group = mtry)) +
+   geom_line(aes(col = mtry))
 
 ```
+
+### Carseats
 
@@
 ```
+
+### OJ
+
+This problem involves the **OJ** data set which is part of the ISLR package.
+
+a.) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.
+
+```{r}
+oj <- as.data.table(ISLR::OJ)
+n <- nrow(oj)
+
+index <- sample(1:n, n *.7, replace = F)
+
+oj.train <- oj[index]
+oj.test <- oj[!index]
+
+head(train)
+```
+
+b.) Fit a tree to the training data, which **Purchases** as the first response and the other variables as predictors.
+Use the **summary** function to produce summary statistics about the tree, and describe the results obtained.
+
+```{r}
+
+oj.tree <- tree(Purchase ~., data = oj.train)
+
+plot(oj.tree)
+text(oj.tree)
+
+summary(oj.tree)
+
+```
+
+What is the training error rate? _0.7469_
+
+How many terminal nodes? _8_
+
+c.) Type in the name of the tree object in order to get a detailed text output. Pick one of the terminal nodes, and interpret the information displayed.
+
+```{r}
+summary(oj.tree)
+```
+
+d.) Create a plot of the tree.
+
+```{r}
+plot(oj.tree)
+text(oj.tree)
+```
+
+e.) Predict the response on the test data, and produce a confusion matrix comparing the test labels to the predicted test labels.
+
+```{r}
+pred <- predict(oj.tree, newdata = oj.test, type = "class")
+
+confusionMatrix(oj.test$Purchase, pred)
+```
+
+f.) Apply the cv.tree function to the training data set in order to determine the optimal tree size.
+
+```{r}
+oj.cv <- cv.tree(oj.tree)
+```
+
+g.) Produce a plot with the tree size and cross-validated error rate.
+
+```{r}
+plot(oj.cv$size, oj.cv$dev, type ='b', col =  "cornflowerblue") # 5
+```
+
+h.) What tree size corresponds to the lowest cv error rate?
+
+**5**
+
+i.) Produce a pruned tree corresponding to the optimal tree size obtained using cross-validation.
+
+```{r}
+oj.pruned <- prune.tree(oj.tree, k = 5)
+```
+
+j.) Compare the training error rates between the pruned and unpruned trees.
+
+```{r}
+
+summary(oj.tree)
+summary(oj.pruned)
+```
+
+Same.
+
+k.) Compare the test error rates.
+
+```{r}
+pruned.pred <- predict(oj.pruned, newdata = oj.test, type = "class")
+
+confusionMatrix(oj.test$Purchase, pred)
+
+confusionMatrix(oj.test$Purchase, pruned.pred)
+```
+
+### Hitters
+
+We now use boosting to predict **Salary** in the hitters data set.
+
+a.) Remove the observations for whom the salary information is unknown, and then log-transform the salaries.
+
+```{r}
+hitters <- as.data.table(ISLR::Hitters)
+hitters <- hitters[complete.cases(hitters)]
+hitters$Salary <- log(hitters$Salary)
+
+```
+
+b.) Create a training set consisting of the first 200 observations and a test set containing the rest.
+
+```{r}
+n <- nrow(hitters)
+
+index <- 1:200
+
+hitters.train <- hitters[index]
+hitters.test <- hitters[!index]
+```
+
+c.) Perform boosting on the training set with 1,000 trees for a range of values of the shrinkage parameter $\lambda$. Produce a plot with the different shrinkage values and the training MSE.
+
+```{r}
+shrinkage <- seq(0.001, 0.5, by = 0.0025)
+iters <- length(shrinkage)
+
+results <- vector(mode = "numeric", length = iters)
+
+for(i in 1:iters)
+{
+   gbm <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[i], distribution = "gaussian")
+   results[i] <- mean(gbm$train.error^2)
+}
+
+lowest.error <- which.min(results)
+
+df_results <- data.table(shrinkage, error = results)[, lowest := .I == lowest.error]
+
+ggplot(df_results, aes(shrinkage, error, fill = lowest)) +
+   geom_bar(stat = "identity")
+```
+
+d.) Produce a plot with different shrinkage methods and the corresponding test mse.
+
+```{r}
+shrinkage <- seq(0.001, 0.5, by = 0.0025)
+iters <- length(shrinkage)
+
+results <- vector(mode = "numeric", length = iters)
+
+for(i in 1:iters)
+{
+   gbm <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[i], distribution = "gaussian")
+
+   pred <- predict(gbm, n.trees = 1000, newdata = hitters.test, type = "response")
+
+   results[i] <- mean((hitters.test$Salary - pred)^2)
+}
+
+lowest.error <- which.min(results)
+
+df_results <- data.table(shrinkage, error = results)[, lowest := .I == lowest.error]
+
+ggplot(df_results, aes(shrinkage, error, fill = lowest)) +
+   geom_bar(stat = "identity")
+
+tree.mse <- results[lowest.error]
+```
+
+e.) Compare the test MSE of boosting to the test MSE that results from applying two of the regression approaches seen in chapters 3 & 6.
+
+```{r}
+
+train.mat <- model.matrix(Salary ~., data = hitters.train)
+test.mat <- model.matrix(Salary ~., data = hitters.test)
+
+cv.ridge <- cv.glmnet(train.mat, hitters.train$Salary, data = hitters.train, alpha = 0)
+
+ridge.fit <- glmnet(test.mat, hitters.test$Salary, alpha = 1, lambda = cv.ridge$lambda.min)
+ridge.pred <- predict(ridge.fit, newx = test.mat, type = "response")
+ridge.error <- mean((hitters.test$Salary - ridge.pred)^2)
+
+cv.lasso <- cv.glmnet(train.mat, hitters.train$Salary, data = hitters.train, alpha = 1)
+
+lasso.fit <- glmnet(test.mat, hitters.test$Salary, alpha = 0, lambda = cv.lasso$lambda.min)
+lasso.pred <- predict(lasso.fit, newx = test.mat, type = "response")
+lasso.error <- mean((hitters.test$Salary - lasso.pred)^2)
+
+results <- data.table(type = c("Tree", "Ridge", "Lasso"), error = c(tree.mse, ridge.error, lasso.error))
+
+ggplot(results, aes(type, error, fill = type)) +
+   geom_bar(stat = "identity")
+```
+
+f.) Which variables appear to be the most important predictors in the boosted model?
+
+```{r}
+
+best.tree <- gbm(Salary ~., data = hitters.train, n.trees = 1000, shrinkage = shrinkage[lowest.error], distribution = "gaussian")
+
+vip::vip(best.tree)
+```
+
+### Caravan
+
+This question uses the caravan data set.
+
+a.) Create a training set consisting of the first 1,000 observations, and a test set consisting of the remaining observations.
+
+```{r}
+caravan <- as.data.table(ISLR::Caravan)
+caravan$Purchase <- ifelse(caravan$Purchase == "Yes", 1, 0)
+index <- 1:1000
+
+caravan.train <- caravan[index]
+caravan.test <- caravan[-index]
+
+head(caravan.train)
+```
+
+b.) Fit a boosting model to the training set with **Purchases** as the response and the other variables as predictors. Use 1,000 trees and a shrinkage value of .01.
+
+```{r}
+boost.fit <- gbm(Purchase ~ ., data = caravan.train, n.trees = 1000, shrinkage = 0.1, distribution = "bernoulli")
+```
+
+What predictors appear to be most important?
+
+```{r}
+vip::vip(boost.fit)
+```
+
+c.) Use the boosting model to predict the response on the test data. Predict that a person will make a purchase if the estimated probablity of purchase is greater than 20%.
+
+```{r}
+probs <- exp(predict.gbm(boost.fit, newdata = caravan.test, type = "response", n.trees = 1000)) - 1
+
+pred <- ifelse(probs > .2, 1, 0)
+
+confusionMatrix(as.factor(caravan.test$Purchase), as.factor(pred))
+```

### 2020-02-29 ISLR/10_Unsupervised-Learning.rmd
islr, support vector machines
islr, support vector machines

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 9}
   \fancyfoot[C]{\rmfamily\color{headergrey}Support Vector Machines}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(e1071, quietly = TRUE, warn.conflicts = FALSE)
library(LiblineaR, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select

```

## Chapter 9

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## R Quiz

```{r}
set.seed(1001)
counts = 100
errate = rep(0, counts)
for(i in 1:counts){
  x = matrix(rnorm(100 * 10), ncol = 10)
  y = c(rep(0, 50), rep(1, 50))
  x[y == 1, 1:5] = x[y == 1, 1:5] + 1
  dat = data.frame(x = x, y = as.factor(y))
  svm.fit = svm(y ~ ., data = dat, kernel = "linear", cost = 1)

  xtest = matrix(rnorm(100 * 10), ncol = 10)
  ytest = sample(c(0, 1), 100, rep = TRUE)
  xtest[ytest == 1,] = xtest[ytest == 1,] + 1
  testdat = data.frame(x = xtest, y = as.factor(ytest))

  ypred = predict(svm.fit, testdat)
  result = table(predict = ypred, truth = testdat$y)
  errate[i] = 1 - (result[1] + result[4]) / 100
}
mean(errate)
```

## Lab

### Support Vector Classifier

```{r}
set.seed(1)

x <- matrix(rnorm(20*2), ncol = 2)
y <- c(rep(-1, 10), rep(1, 10))
x[y == 1,] <- x[y==1,] + 1
plot(x, col=(3-y))
```

```{r}
dat <- data.table(x1 = x[,1], x2 = x[, 2], y = as.factor(y))
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 10, scale = F)
plot(svmfit, dat)
```

```{r}
svmfit$index
```

```{r}
summary(svmfit)
```

```{r}
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 0.1, scale = F)
plot(svmfit, dat)
svmfit$index
```

```{r}
set.seed(1)

tune.out <- tune(svm, y ~ ., data = dat, kernel = "linear",
                 ranges = list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
summary(tune.out)
```

```{r}
bestmod <- tune.out$best.model
summary(bestmod)
```

```{r}
xtest <- matrix(rnorm(20*2), ncol = 2)
ytest <- sample(c(-1, 1), 20, rep = T)
xtest[ytest==1,] = xtest[ytest==1,] + 1
testdata <- data.table(x1 = xtest[, 1], x2 = xtest[, 2], y = as.factor(ytest))
```

```{r}
ypred <- predict(bestmod, testdata)
table(predict = ypred, truth = testdata$y)
```

```{r}
svmfit <- svm( y ~ ., data = dat, kernel = "linear", cost = .01, scale = F)
ypred <- predict(svmfit, testdata)
table(predict = ypred, truth = testdata$y)
```

```{r}
x[y ==1,] <- x[y == 1,] + 0.5
plot(x, col = (y + 5)/2, pch = 19)
```

```{r}
dat <- data.table(x = x, y = as.factor(y))
names(dat)[1:2] <- c("x1", "x2")

svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 1e5)
summary(svmfit)
plot(svmfit, dat)
```

```{r}
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 1)
summary(svmfit)
plot(svmfit, dat)
```

```{r}
set.seed(1)

x <- matrix(rnorm(200*2), ncol = 2)
x[1:100,] = x[1:100,] + 2
x[101:150,] = x[101:150, ] - 2
y <- c(rep(1, 150), rep(2, 50))

dat = data.frame(x = x, y = as.factor(y))

par(mfrow = c(1, 1))
plot(x, col = y)
```

### Support Vecctor Machines

```{r}
train <- sample(200, 100)
svmfit <- svm(y ~ ., data = dat[train,], kernel = "radial", gamma = 1, cost = 1)

plot(svmfit, dat[train,])
```

```{r}
summary(svmfit)
```

```{r}
svmfit <- svm(y ~ ., data = dat[train,], kernel = "radial", gamma = 1,
              cost = 1e5)
plot(svmfit, dat)
```

```{r}
set.seed(1)

tune.out <- tune(svm, y ~ ., data = dat[train,], kernel = "radial",
                 ranges = list(cost = c(0.1, 1, 10, 100, 1000),
                              gamma = c(0.5, 1, 2, 3, 4)))

summary(tune.out)
```

```{r}
table(true = dat[-train, "y"], pred = predict(tune.out$best.model, newdata = dat[-train,]))
```

### ROC Curves

```{r}
rocplot <- function(pred, truth, ...) {
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf, ...)
}
```

```{r}
svmfit.opt <- svm(y ~ ., data = dat[train,], kernel = "radial",
                  gamma = 2, cost = 1, decision.values = T)

fitted <- attributes(predict(svmfit.opt, dat[train,], decision.values = T))$decision.values

par(mfrow = c(1, 2))

rocplot(fitted, dat[train, "y"], main = "Training Data")
svmfit.flex <- svm(y ~ ., data = dat[train,], kernel = "radial", gamma = 50, cost = 1, decision.values = T)
fitted <- attributes(predict(svmfit.flex, dat[train,], decision.values = T))$decision.values
rocplot(fitted, dat[train, "y"], add = T, col = "red")
```

```{r}
fitted <- attributes(predict(svmfit.opt, dat[-train, ], decision.values = T))$decision.values
rocplot(fitted, dat[-train, "y"], main = "Test Data")
fitted <- attributes(predict(svmfit.flex, dat[-train,], decision.values = T))$decision.values
rocplot(fitted, dat[-train, "y"], add = T, col = "red")
```

### SVM w/ Multiple Classes

```{r}
set.seed(1)

x <- rbind(x, matrix(rnorm(50*2), ncol = 2))
y <- c(y, rep(0, 50))

x[y == 0, 2] = x[y == 0, 2] + 2
dat = data.frame(x = x, y = as.factor(y))

par(mfrow = c(1, 1))
plot(x, col = (y + 1))
```

```{r}
svmfit <- svm(y ~ ., data = dat, kernel = "radial", cost = 10, gamma = 1)
plot(svmfit, dat)
```

### Gene Expression Data

```{r}
khan <- ISLR::Khan
```

```{r}
table(khan$ytrain)
table(khan$ytest)
```

```{r}
dat <- data.frame(x = khan$xtrain, y = as.factor(khan$ytrain))
out <- svm(y ~ ., data = dat, kernel = "linear", cost = 10)

summary(out)
```

```{r}
table(out$fitted, dat$y)
```

```{r}
dat.te <- data.frame(x = khan$xtest, y = as.factor(khan$ytest))
pred.te <- predict(out, newdata = dat.te)

table(pred.te, dat.te$y)
```

### Applied

4.) Generate a simulated two-class data set with 100 observations and two features in which there is a visible but non-linear separation between the two classes. Show that in this setting, a support vector machine with a polynomial kernel (with degree greater than 1) or a radial kernel will outperform a support vector classifier on the training data. Which technique performs best on the test data? Make plots and report training and test error rates in order to back up your assertions.

Generate data and plot:

```{r}
set.seed(1)

transl <- 3
X <- matrix(rnorm(100 * 2), ncol = 2)
X[1:30, ] <- X[1:30, ] + transl
X[31:60, ] <- X[31:60, ] - transl
y <- c(rep(0, 60), rep(1, 40))
dat <- data.frame(x = X, y = as.factor(y))

par(mfrow = c(1,1))
plot(X, col = y + 1)
```

Split to training and test set:

```{r}
train <- sample(100, 80)
dat.train <- dat[train, ]
dat.test <- dat[-train, ]
```

```{r}
svm.lin <- svm(y ~ ., data = dat.train, kernel = 'linear', scale = FALSE)
plot(svm.lin, data = dat.train)
```

```{r}
summary(svm.lin)
```

Calculate the training error of the support vector classifier:

```{r}
table(predict = svm.lin$fitted, truth = dat.train$y)
```

Fit with polynomial kernel and calculate the training error rate:

```{r}
svm.poly <- svm(y ~ ., data = dat.train, kernel = 'polynomial', scale = FALSE)
plot(svm.poly, data = dat.train)
```

```{r}
table(predict = svm.poly$fitted, truth = dat.train$y)
```

Fit with radial kernel and calculate the traing error rate:

```{r}
svm.rad <- svm(y ~ ., data = dat.train, kernel = 'radial', scale = FALSE)
plot(svm.rad, data = dat.train)
```

```{r}
table(predict = svm.rad$fitted, truth = dat.train$y)
```

Compare the test errors of the 3 kernels:

```{r}
lin.pred <- predict(svm.lin, dat.test)
table(predict = lin.pred, truth = dat.test$y)
```

```{r}
poly.pred <- predict(svm.poly, dat.test)
table(predict = poly.pred, truth = dat.test$y)
```

```{r}
rad.pred <- predict(svm.rad, dat.test)
table(predict = rad.pred, truth = dat.test$y)
```

5.) We have seen that we can fit an SVM with a non-linear kernel in order to perform classification using a non-linear decision boundary. We will now see that we can also obtain a non-linear decision boundary by performing logistic regression using non-linear transformations of the features.

a.) Generate a data set with n = 500 and p = 2, such that the observations belong to two classes with a quadratic decision boundary between them:

```{r}
set.seed(1)

x1 <- runif(500) - 0.5
x2 <- runif(500) - 0.5
y <- as.integer(x1 ^ 2 - x2 ^ 2 > 0)
```

b.) Plot the observations, colored according to their class labels. Your plot should display X 1 on the x-axis, and X 2 on the y-axis:

```{r}
plot(x1[y == 0], x2[y == 0], col = "red", xlab = "X1", ylab = "X2")
points(x1[y == 1], x2[y == 1], col = "blue")
```

Fit a logistic regression model to the data, using X1 and X2 as predictors.

```{r}
dat <- data.frame(x1 = x1, x2 = x2, y = as.factor(y))
lr.fit <- glm(y ~ ., data = dat, family = 'binomial')
```

d.) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be linear.

```{r}
lr.prob <- predict(lr.fit, newdata = dat, type = 'response')
lr.pred <- ifelse(lr.prob > 0.5, 1, 0)
plot(dat$x1, dat$x2, col = lr.pred + 2)
```

e.) Now ﬁt a logistic regression model to the data using non-linear functions of X1 and X2 as predictors (e.g. X21 , X1×X2, log(X2), and so forth).

```{r}
lr.nl <- glm(y ~ poly(x1, 2) + poly(x2, 2), data = dat, family = 'binomial')
```

f.) Apply this model to the training data in order to obtain a predicted class label for each training observation. Plot the observations, colored according to the predicted class labels. The decision boundary should be obviously non-linear. If it is not, then repeat (a)-(e) until you come up with an example in which the predicted class labels are obviously non-linear.

```{r}
lr.prob.nl <- predict(lr.nl, newdata = dat, type = 'response')
lr.pred.nl <- ifelse(lr.prob.nl > 0.5, 1, 0)
plot(dat$x1, dat$x2, col = lr.pred.nl + 2)
```

g.) Fit a support vector classiﬁer to the data with X1 and X2 as predictors. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.

```{r}
svm.lin <- svm(y ~ ., data = dat, kernel = 'linear', cost = 0.01)
plot(svm.lin, dat)
```

h.) Fit a SVM using a non-linear kernel to the data. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.

```{r}
svm.nl <- svm(y ~ ., data = dat, kernel = 'radial', gamma = 1)
plot(svm.nl, data = dat)
```

6.) At the end of Section 9.6.1, it is claimed that in the case of data that is just barely linearly separable, a support vector classiﬁer with a small value of cost that misclassiﬁes a couple of training observations may perform better on test data than one with a huge value of cost that does not misclassify any training observations. You will now investigate this claim.

a.) Generate two-class data with p = 2 in such a way that the classes are just barely linearly separable.

```{r}
set.seed(1)
obs = 1000
x1 <- runif(obs, min = -4, max = 4)
x2 <- runif(obs, min = -1, max = 16)
y <- ifelse(x2 > x1 ^ 2, 0, 1)
dat <- data.frame(x1 = x1, x2 = x2, y = as.factor(y))
train <- sample(obs, obs/2)
dat.train <- dat[train, ]
dat.test <- dat[-train, ]
par(mfrow = c(1,2))
plot(dat.train$x1, dat.train$x2, col = as.integer(dat.train$y) + 1, main = 'training set')
plot(dat.test$x1, dat.test$x2, col = as.integer(dat.test$y) + 1, main = 'test set')
```

b.) Compute the cross-validation error rates for support vector classiﬁers with a range of cost values. How many training errors are misclassiﬁed for each value of cost considered, and how does this relate to the cross-validation errors obtained?

```{r}
set.seed(1)
cost.grid <- c(0.001, 0.01, 0.1, 1, 5, 10, 100, 10000)
tune.out <- tune(svm, y ~., data = dat.train, kernel = 'linear', ranges = list(cost = cost.grid))
```

```{r}
summary(tune.out)
```

c.) Generate an appropriate trest data set, and compute the test errors corresponding to each of the values of cost considered. Which value of cost leads to the fewest test errors, and how does this compare to the values of cost that generate the fewest train errors?

```{r}
err.rate.train <- rep(NA, length(cost.grid))
for (cost in cost.grid) {
  svm.fit <- svm(y ~ ., data = dat.train, kernel = 'linear', cost = cost)
  plot(svm.fit, data = dat.train)
  res <- table(prediction = predict(svm.fit, newdata = dat.train), truth = dat.train$y)
  err.rate.train[match(cost, cost.grid)] <- (res[2,1] + res[1,2]) / sum(res)
}
```

```{r}
paste('The cost', cost.grid[which.min(err.rate.train)], 'has the minimum training error:', min(err.rate.train))
```

7.) In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the Auto data set.

```{r}
auto <- as.data.table(ISLR::Auto)
```

a.) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage with below the median.

```{r}
auto$eco <- as.factor(ifelse(auto$mpg < median(auto$mpg), 0, 1))
training <- auto[, !"mpg"]
```

b.) Fit a support vector classificer to the data with various values of cost, in order to predict whetehr a car gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter.

```{r}
cost.grid <- c(0.001, 0.1, 1, 100)
set.seed(1)
tune.res <- tune(svm, eco ~ ., data = training, kernel = 'linear', ranges = list(cost = cost.grid))
summary(tune.res)

plot(tune.res$best.model, training, cylinders ~ displacement)
plot(tune.res$best.model, training, horsepower ~ weight)
```

c.) Now repeat this with SVMs with radial and polynomial basis kernels, which different values of gamma and degree.

```{r}
cost.grid <- c(0.001, 0.1, 1, 100)
set.seed(1)
tune.res <- tune(svm, eco ~ ., data = training, kernel = 'radial', ranges = list(cost = cost.grid))
summary(tune.res)

plot(tune.res$best.model, training, cylinders ~ displacement)
plot(tune.res$best.model, training, horsepower ~ weight)
plot(tune.res$best.model, training, acceleration ~ year)
```

```{r}
cost.grid <- c(0.001, 0.1, 1, 100)
set.seed(1)
tune.res <- tune(svm, eco ~ ., data = training, kernel = 'polynomial', ranges = list(cost = cost.grid))
summary(tune.res)

plot(tune.res$best.model, training, cylinders ~ displacement)
plot(tune.res$best.model, training, horsepower ~ weight)
plot(tune.res$best.model, training, acceleration ~ year)
```

8.) This problem involves the OJ data set which is part of the ISLR package.

```{r}
oj <- as.data.table(ISLR::OJ)
```

a.) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.

```{r}
train <- sample(1:nrow(oj), 800)

oj.train <- oj[train,]
oj.test <- oj[-train,]
```

b.) Fit a support vector machine classifier to the training data using cost = 0.01, with purchase as the response and other variables as the predictors.

```{r}
svmfit <- svm(Purchase ~ ., data = oj.train, kernel = "linear", cost = 0.01)

summary(svmfit)
```

c.) What are the train and test error rates?

Test error rate:

```{r}
table(oj.train$Purchase, predict(svmfit))

mean(oj.train$Purchase == predict(svmfit))
```

Test error rate:

```{r}
table(oj.test$Purchase, predict(svmfit, newdata = oj.test))

mean(oj.test$Purchase == predict(svmfit, newdata = oj.test))
```

d.) Use the tune() function to select an optimal model.

```{r}
cost.grid <- c(0.001, 0.1, 1, 100, 250)
out <- tune(svm, Purchase ~ ., data = oj.train, kernel = "linear", ranges = list(cost = cost.grid))

summary(out)
```

e.) Compute the training and test error rates using this new value for cost:

```{r}
svmfit <- out$best.model

table(oj.train$Purchase, predict(svmfit))

mean(oj.train$Purchase == predict(svmfit))
```

```{r}
table(oj.test$Purchase, predict(svmfit, newdata = oj.test))

mean(oj.test$Purchase == predict(svmfit, newdata = oj.test))
```

f.) Repeat parts b-d using a radial kernel.

```{r}
cost.grid <- c(0.001, 0.1, 1, 100, 250)
out <- tune(svm, Purchase ~ ., data = oj.train, kernel = "radial", ranges = list(cost = cost.grid))

summary(out)

svmfit <- out$best.model

table(oj.train$Purchase, predict(svmfit))
mean(oj.train$Purchase == predict(svmfit))

table(oj.test$Purchase, predict(svmfit, newdata = oj.test))
mean(oj.test$Purchase == predict(svmfit, newdata = oj.test))
```

h.) Overall, which approach seems to give the best results?

The linear kernel has the best test error rate.

```{r}
rm(list = ls())
```

### 2020-02-29 ISLR/10_Unsupervised-Learning.rmd
completed ISLR course & book
completed ISLR course & book

@@
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 9}
-   \fancyfoot[C]{\rmfamily\color{headergrey}Support Vector Machines}
+   \lfoot{\color{headergrey}Chapter 10}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Unsupervised Learning}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 library(caret, quietly = TRUE, warn.conflicts = FALSE)
-library(class, quietly = TRUE, warn.conflicts = FALSE)
 library(boot, quietly = TRUE, warn.conflicts = FALSE)
 library(vip, quietly = TRUE, warn.conflicts = FALSE)
-library(e1071, quietly = TRUE, warn.conflicts = FALSE)
-library(LiblineaR, quietly = TRUE, warn.conflicts = FALSE)
-library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-## Chapter 9
+## Chapter 10
 
@@
 
-## R Quiz
+## Lab
 
 ```{r}
-set.seed(1001)
-counts = 100
-errate = rep(0, counts)
-for(i in 1:counts){
-  x = matrix(rnorm(100 * 10), ncol = 10)
-  y = c(rep(0, 50), rep(1, 50))
-  x[y == 1, 1:5] = x[y == 1, 1:5] + 1
-  dat = data.frame(x = x, y = as.factor(y))
-  svm.fit = svm(y ~ ., data = dat, kernel = "linear", cost = 1)
-
-  xtest = matrix(rnorm(100 * 10), ncol = 10)
-  ytest = sample(c(0, 1), 100, rep = TRUE)
-  xtest[ytest == 1,] = xtest[ytest == 1,] + 1
-  testdat = data.frame(x = xtest, y = as.factor(ytest))
-
-  ypred = predict(svm.fit, testdat)
-  result = table(predict = ypred, truth = testdat$y)
-  errate[i] = 1 - (result[1] + result[4]) / 100
-}
-mean(errate)
+load(file.path(here::here(), "ISLR", "10.R.RData"))
 ```
 
-## Lab
+Suppose we want to fit a linear regression, but the number of variables is much larger than the number of observations.
+In some cases, we may improve the fit by reducing the dimension of the features before.
 
-### Support Vector Classifier
+In this problem, we use a data set with n = 300 and p = 200, so we have more observations than variables, but not by much. Load the data x, y, x.test, and y.test from 10.R.RData.
+
+First, concatenate x and x.test using the rbind functions and perform a principal components analysis on the concatenated data frame (use the "scale=TRUE" option).
+To within 10% relative error, what proportion of the variance is explained by the first five principal components?
+
+```{r}
+dat <- rbind(x, x.test)
+
+pca.out <- prcomp(dat, scale = T)
+
+summary(pca.out) # 0.3499
+
+sum(head((pca.out$sdev)^2/ sum(pca.out$sdev^2), length = 5)) # 0.3566807
+
+par(mfrow = c(1,1))
+plot(pca.out)
+```
+
+The previous answer suggests that a relatively small number of "latent variables" account for a substantial fraction of the features' variability.
+We might believe that these latent variables are more important than linear combinations of the features that have low variance.
+
+We can try forgetting about the raw features and using the first five principal components (computed on rbind(x,x.test)) instead as low-dimensional derived features.
+What is the mean-squared test error if we regress y on the first five principal components, and use the resulting model to predict y.test?
+
+```{r}
+xols <- pca.out$x[1:300,1:5]
+fit0 <- lm(y ~ xols)
+summary(fit0)
+
+yhat0 <- predict(fit0, x.test)
+mean((yhat0-y.test)^2)
+```
+
+### Lab
+
+```{r}
+states <- row.names(USArrests)
+states
+```
+
+```{r}
+names(USArrests)
+```
+
+```{r}
+apply(USArrests, 2, mean)
+```
+
+```{r}
+apply(USArrests, 2, var)
+```
+
+```{r}
+pr.out <- prcomp(USArrests, scale = T)
+```
+
+```{r}
+names(pr.out)
+```
+
+```{r}
+pr.out$center
+```
+
+```{r}
+pr.out$rotation
+```
+
+```{r}
+dim(pr.out$x)
+```
+
+```{r}
+pr.out$rotation =-pr.out$rotation
+pr.out$x =-pr.out$x
+biplot(pr.out, scale = 0)
+```
+
+```{r}
+pr.out$sdev
+```
+
+```{r}
+pr.var = pr.out$sdev^2
+pr.out$sdev
+```
+
+```{r}
+pve <- pr.var/sum(pr.var)
+pve
+```
+
+```{r}
+plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0, 1), type = "b")
+plot(cumsum(pve), xlab = "Principal Component", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0, 1), type = 'b')
+```
+
+```{r}
+set.seed(2)
+x <- matrix(rnorm(50*2), ncol = 2)
+x[1:25, 1] = x[1:25, 1] + 3
+x[1:25, 2] = x[1:25, 2] - 4
+```
+
+```{r}
+km.out <- kmeans(x, 2, nstart = 20)
+```
+
+```{r}
+km.out$cluster
+```
+
+```{r}
+plot(x, col = (km.out$cluster + 1), main = "K-Means Clustering Results with K=2", pch = 20, cex = 2)
+```
+
+```{r}
+set.seed(4)
+
+km.out <- kmeans(x, 3, nstart = 20)
+km.out
+```
+
+```{r}
+set.seed(3)
+
+km.out <- kmeans(x, 3, nstart = 1)
+km.out$tot.withinss
+
+km.out <- kmeans(x, 3, nstart = 20)
+km.out$tot.withinss
+```
+
+```{r}
+hc.complete <- hclust(dist(x), method = "complete")
+hc.average <- hclust(dist(x), method = "average")
+hc.single <- hclust(dist(x), method = "single")
+```
+
+```{r}
+par(mfrow = c(1, 3))
+plot(hc.complete, main = "Complete Linkage", cex = .9)
+plot(hc.average, main = "Average Linkage", cex = .9)
+plot(hc.single, main = "Single Linkage", cex = .9)
+```
+
+```{r}
+xsc <- scale(x)
+
+par(mfrow = c(1, 1))
+plot(hclust(dist(xsc), method = "complete"), main = "Hierarchial Clustering with Scaled Features")
+```
+
+```{r}
+x <- matrix(rnorm(30*3), ncol = 3)
+dd <- as.dist(1 - cor(t(x)))
+
+plot(hclust(dd, method = "complete"))
+```
+
+3.) In this problem, you will perform K-means clustering manually, with K=2, on a small example with n=6 observations and p=2 features. The observations are as follows.
+
+a.) Plot the observations.
+
+```{r}
+x <- cbind(c(1, 1, 0, 5, 6, 4), c(4, 3, 4, 1, 2, 0))
+plot(x[,1], x[,2])
+```
+
+b.) Randomly assign a cluster label to each observation. Report the cluster labels for each observation.
 
@@
 set.seed(1)
-
-x <- matrix(rnorm(20*2), ncol = 2)
-y <- c(rep(-1, 10), rep(1, 10))
-x[y == 1,] <- x[y==1,] + 1
-plot(x, col=(3-y))
+labels <- sample(2, nrow(x), replace = T)
+labels
 ```
@@
 ```{r}
-dat <- data.table(x1 = x[,1], x2 = x[, 2], y = as.factor(y))
-svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 10, scale = F)
-plot(svmfit, dat)
+plot(x[, 1], x[, 2], col = (labels + 1), pch = 20, cex = 2)
+```
+
+c.) Compute the centroid for each cluster.
+
+```{r}
+centroid1 <- c(mean(x[labels == 1, 1]), mean(x[labels == 1, 2]))
+centroid2 <- c(mean(x[labels == 2, 1]), mean(x[labels == 2, 2]))
+plot(x[,1], x[,2], col=(labels + 1), pch = 20, cex = 2)
+points(centroid1[1], centroid1[2], col = 2, pch = 4)
+points(centroid2[1], centroid2[2], col = 3, pch = 4)
+```
+
+d.) Assign each observation to the centroid to which it is closest, in terms of Euclidean distance. Report the cluster labels for each observation.
+
+```{r}
+labels <- c(1, 1, 1, 2, 2, 2)
+plot(x[, 1], x[, 2], col = (labels + 1), pch = 20, cex = 2)
+points(centroid1[1], centroid1[2], col = 2, pch = 4)
+points(centroid2[1], centroid2[2], col = 3, pch = 4)
+```
+
+e.) Repeat (c) and (d) until the answers obtained stop changing.
+
+```{r}
+centroid1 <- c(mean(x[labels == 1, 1]), mean(x[labels == 1, 2]))
+centroid2 <- c(mean(x[labels == 2, 1]), mean(x[labels == 2, 2]))
+plot(x[,1], x[,2], col=(labels + 1), pch = 20, cex = 2)
+points(centroid1[1], centroid1[2], col = 2, pch = 4)
+points(centroid2[1], centroid2[2], col = 3, pch = 4)
+```
+
+f.) In your plot from (a), color the observations according to the clusters labels obtained.
+
+```{r}
+plot(x[, 1], x[, 2], col=(labels + 1), pch = 20, cex = 2)
+```
+
+In words, describe the results that you would expect if you performed K-means clustering of the eight shoppers in Figure 10.14, on the basis of their sock and computer purchases, with K=2.
+Give three answers, one for each of the variable scalings displayed.
+Explain.
+
+```{r}
+socks <- c(8, 11, 7, 6, 5, 6, 7, 8)
+computers <- c(0, 0, 0, 0, 1, 1, 1, 1)
+x <- cbind(socks, computers)
+labels <- c(1, 1, 2, 2, 2, 2, 1, 1)
+plot(x[, 1], x[, 2], col=(labels + 1), pch = 20, cex = 2, asp = 1)
+```
+
+If we take into consideration the scaled variables, the number of computers plays a much larger role than the number of socks, so we have the clusters {5,6,7,8} (purchased computer) and {1,2,3,4} (no computer purchased).
+
+```{r}
+x <- cbind(scale(socks, center = FALSE), scale(computers, center = FALSE))
+sd(computers)
 ```
@@
 ```{r}
-svmfit$index
+labels <- c(1, 1, 2, 2, 2, 2, 1, 1)
+plot(x[, 1], x[, 2], col=(labels + 1), pch = 20, cex = 2, asp = 1)
+```
+
+A researcher collects expression measurements for 1000 genes in 100 tissue samples.
+The data can be written as a 1000x1000 matrix, which we call X, in which each row represents a gene and each column a tissue sample.
+Each tissue sample was processed on a different day, and the columns of X are ordered so that the samples that were processed earliest are on the left, and the samples that were processed later are on the right.
+The tissue samples belong to two groups : control (C) and treatment (T).
+The C and T samples were processed in a random order across the days.
+The researcher wishes to determine whether each gene’s expression measurements differ between the treatment and control groups.
+
+As a pre-analysis (before comparing T versus C), the researcher performs a principal component analysis of the data, and finds that the first principal component (a vector of length 100) has a strong linear trend from left to right, and explains 10% of the variation.
+The researcher now remembers that each patient sample was run on one of two machines, A and B, and machine A was used more often in the earlier times while B was used more often later.
+The researcher has a record of which sample was run on which machine.
+
+Explain what it means that the first principal component “explains 10% of the variation”.
+The first principal component “explains 10% of the variation” means 90% of the information in the gene data set is lost by projecting the tissue sample observations onto the first principal component.
+Another way of explaining it is 90% of the variance in the data is not contained in the first principal component.
+
+The researcher decides to replace the (i,j)th element of X with
+xij−zi1ϕj1
+where zi1 is the ith score, and ϕj1 is the jth loading, for the first principal component.
+He will then perform a two-sample t-test on each gene in this new data set in order to determine whether its expression differs between the two conditions.
+Critique this idea, and suggest a better approach.
+Given the flaw shown in pre-analysis of a time-wise linear trend amongst the tissue samples’ first principal component, I would advise the researcher to include the machine used (A vs B) as a feature of the data set.
+This should enhance the PVE of the first principal component before applying the two-sample t-test.
+
+Design and run a small simulation experiment to demonstrate the superiority of your idea.
+
+```{r}
+set.seed(1)
+Control <- matrix(rnorm(50 * 1000), ncol = 50)
+Treatment <- matrix(rnorm(50 * 1000), ncol = 50)
+X <- cbind(Control, Treatment)
+X[1, ] <- seq(-18, 18 - .36, .36) # linear trend in one dimension
+pr.out <- prcomp(scale(X))
+summary(pr.out)$importance[, 1]
+```
+
+We have 9.911% variance explained by the first principal component. Now, adding in A vs B via 10 vs 0 encoding.
+
+```{r}
+X <- rbind(X, c(rep(10, 50), rep(0, 50)))
+pr.out <- prcomp(scale(X))
+summary(pr.out)$importance[, 1]
+```
+
+7.) In the chapter, we mentioned the use of correlation-based distance and Euclidean distance as dissimilarity measures for hierarchical clustering.
+It turns out that these two measures are almost equivalent : if each observation has been centered to have mean zero and standard deviation one, and if we let rij denote the correlation between the ith and jth observations, then the quantity 1−rij is proportional to the squared Euclidean distance between the ith and jth observations.
+On the “USArrests” data, show that this proportionality holds.
+
+```{r}
+set.seed(1)
+dsc <- scale(USArrests)
+d1 <- dist(dsc)^2
+d2 <- as.dist(1 - cor(t(dsc)))
+summary(d2 / d1)
+```
+
+In Section 10.2.3, a formula for calculating PVE was given in Equation 10.8.
+We also saw that the PVE can be obtained using the “sdev” output of the “prcomp()” function.
+On the “USArrests” data, calculate PVE in two ways :
+
+a.) Using the “sdev” output of the “prcomp()” function, as was done in Section 10.2.3.
+
+```{r}
+pr.out <- prcomp(USArrests, scale = TRUE)
+pr.var <- pr.out$sdev^2
+pve <- pr.var / sum(pr.var)
+sum(pr.var)
+```
+
+b.) By applying Equation 10.8 directly. That is, use the “prcomp()” function to compute the principal component loadings. Then, use those loadings in Equation 10.8 to obtain the PVE.
+
+```{r}
+loadings <- pr.out$rotation
+USArrests2 <- scale(USArrests)
+sumvar <- sum(apply(as.matrix(USArrests2)^2, 2, sum))
+apply((as.matrix(USArrests2) %*% loadings)^2, 2, sum) / sumvar
+```
+
+9.) Consider the “USArrests” data. We will now perform hierarchical clustering on the states.
+
+a.) Using hierarchical clustering with complete linkage and Euclidean distance, cluster the states.
+
+```{r}
+set.seed(2)
+hc.complete <- hclust(dist(USArrests), method = "complete")
+plot(hc.complete)
+```
+
+b.) Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters ?
+
+```{r}
+cutree(hc.complete, 3)
+```
+
+c.) Hierachically cluster the states using complete linkage and Euclidean distance, after scaling the variables to have standard deviation one.
+
+```{r}
+sd.data <- scale(USArrests)
+hc.complete.sd <- hclust(dist(sd.data), method = "complete")
+plot(hc.complete.sd)
+```
+
+d.) What effect does scaling the variables have on the hierarchical clustering obtained ?
+In your opinion, should the variables be scaled before the inter-observation dissimilarities are computed ?
+Provide a justification for your answer.
+
+```{r}
+cutree(hc.complete.sd, 3)
 ```
@@
 ```{r}
-summary(svmfit)
+table(cutree(hc.complete, 3), cutree(hc.complete.sd, 3))
+```
+
+10.) In this problem, you will generate simulated data, and then perform PCA and K-means clustering on the data.
+
+a.) Generate a simulated data set with 20 observations in each of three classes (i.e. 60 observations total), and 50 variables.
+
+```{r}
+set.seed(2)
+x <- matrix(rnorm(20 * 3 * 50, mean = 0, sd = 0.001), ncol = 50)
+x[1:20, 2] <- 1
+x[21:40, 1] <- 2
+x[21:40, 2] <- 2
+x[41:60, 1] <- 1
+true.labels <- c(rep(1, 20), rep(2, 20), rep(3, 20))
+```
+
+b.) Perform PCA on the 60 observations and plot the first two principal component score vectors.
+Use a different color to indicate the observations in each of the three classes.
+If the three classes appear separated in this plot, then continue on to part (c).
+If not, the return to part (a) and modify the simulation so that there is greater separation between the three classes.
+Do not continue to part (c) until the three classes show at least some separation in the first two principal component score vectors.
+
+```{r}
+pr.out <- prcomp(x)
+plot(pr.out$x[, 1:2], col = 1:3, xlab = "Z1", ylab = "Z2", pch = 19)
+```
+
+c.) Perform K-means clustering of the observations with K=3. How well do the clusters that you obtained in K-means clustering compare to the true class labels ?
+
+```{r}
+km.out <- kmeans(x, 3, nstart = 20)
+table(true.labels, km.out$cluster)
+```
+
+d.) Perform K-means clustering with K=2. Describe your results.
+
+```{r}
+km.out <- kmeans(x, 2, nstart = 20)
+table(true.labels, km.out$cluster)
+```
+
+e.) Now perform K-means clustering with K=4, and describe your results.
+
+```{r}
+km.out <- kmeans(x, 4, nstart = 20)
+table(true.labels, km.out$cluster)
+```
+
+f.) Now perform K-means clustering with K=3 on the first two principal component score vectors, rather than on the raw data.
+That is, perform K-means clustering on the 60x2 matrix of which the first column is the first principal component score vector, and the second column is the second principal component score vector.
+Comment on the results.
+
+```{r}
+km.out <- kmeans(pr.out$x[, 1:2], 3, nstart = 20)
+table(true.labels, km.out$cluster)
+```
+
+g.) Using the “scale()” function, perform K-means clustering with K=3 on the data after scaling each variable to have standard deviation one.
+How do these results compare to those obtained in (b) ?
+Explain.
+
+```{r}
+km.out <- kmeans(scale(x), 3, nstart = 20)
+table(true.labels, km.out$cluster)
+```
+
+11.) On the book website, there is a gene expression data set that consists of 40 tissue samples with measurements on 1000 genes.
+The first 20 samples are from healthy patients, while the second 20 are from a diseased group.
+
+a.) Load the data using “read.csv()”. You will need to select “header = F”.
+
+```{r}
+genes <- read.csv(file.path(here::here(), "ISLR", "Ch10Ex11.csv"), header = FALSE)
+```
+
+b.) Apply hierarchical clustering to the samples using correlation-based distance, and plot the dendrogram.
+Do the genes separate the samples into two groups ?
+Do your results depend on the type of linkage used ?
+
+```{r}
+hc.complete <- hclust(as.dist(1 - cor(genes)), method = "complete")
+plot(hc.complete)
 ```
@@
 ```{r}
-svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 0.1, scale = F)
-plot(svmfit, dat)
-svmfit$index
+hc.single <- hclust(as.dist(1 - cor(genes)), method = "single")
+plot(hc.single)
 ```
@@
 ```{r}
-set.seed(1)
+hc.average <- hclust(as.dist(1 - cor(genes)), method = "average")
+plot(hc.average)
+```
 
-tune.out <- tune(svm, y ~ ., data = dat, kernel = "linear",
-                 ranges = list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
-summary(tune.out)
+The results are pretty different when using different linkage methods as we obtain two clusters for complete and single linkages or three clusters for average cluster.
+
+c.) Your collaborator wants to know which genes differ the most across the two groups. Suggest a way to answer this question, and apply it here.
+We may use PCA to see which genes differ the most. We will examine the absolute values of the total loadings for each gene as it characterizes the weight of each gene.
+
+```{r}
+pr.out <- prcomp(t(genes))
+head(pr.out$rotation)
 ```
@@
 ```{r}
-bestmod <- tune.out$best.model
-summary(bestmod)
+total.load <- apply(pr.out$rotation, 1, sum)
+index <- order(abs(total.load), decreasing = TRUE)
+index[1:10]
 ```
-
-```{r}
-xtest <- matrix(rnorm(20*2), ncol = 2)
-ytest <- sample(c(-1, 1), 20, rep = T)
-xtest[ytest==1,] = xtest[ytest==1,] + 1
-testdata <- data.table(x1 = xtest[, 1], x2 = xtest[, 2], y = as.factor(ytest))
-```
-
-```{r}
-ypred <- predict(bestmod, testdata)
-table(predict = ypred, truth = testdata$y)
-```
-
-```{r}
-svmfit <- svm( y ~ ., data = dat, kernel = "linear", cost = .01, scale = F)
-ypred <- predict(svmfit, testdata)
-table(predict = ypred, truth = testdata$y)
-```
-
-```{r}
-x[y ==1,] <- x[y == 1,] + 0.5
-plot(x, col = (y + 5)/2, pch = 19)
-```
-
-```{r}
-dat <- data.table(x = x, y = as.factor(y))
-names(dat)[1:2] <- c("x1", "x2")
-
-svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 1e5)
-summary(svmfit)
-plot(svmfit, dat)
-```
-
-```{r}
-svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 1)
-summary(svmfit)
-plot(svmfit, dat)
-```
-
-```{r}
-set.seed(1)
-
-x <- matrix(rnorm(200*2), ncol = 2)
-x[1:100,] = x[1:100,] + 2
-x[101:150,] = x[101:150, ] - 2
-y <- c(rep(1, 150), rep(2, 50))
-
-dat = data.frame(x = x, y = as.factor(y))
-
-par(mfrow = c(1, 1))
-plot(x, col = y)
-```
-
-### Support Vecctor Machines
-
-```{r}
-train <- sample(200, 100)
-svmfit <- svm(y ~ ., data = dat[train,], kernel = "radial", gamma = 1, cost = 1)
-
-plot(svmfit, dat[train,])
-```
-
-```{r}
-summary(svmfit)
-```
-
-```{r}
-svmfit <- svm(y ~ ., data = dat[train,], kernel = "radial", gamma = 1,
-              cost = 1e5)
-plot(svmfit, dat)
-```
-
-```{r}
-set.seed(1)
-
-tune.out <- tune(svm, y ~ ., data = dat[train,], kernel = "radial",
-                 ranges = list(cost = c(0.1, 1, 10, 100, 1000),
-                              gamma = c(0.5, 1, 2, 3, 4)))
-
-summary(tune.out)
-```
-
-```{r}
-table(true = dat[-train, "y"], pred = predict(tune.out$best.model, newdata = dat[-train,]))
-```
-
-### ROC Curves
-
-```{r}
-rocplot <- function(pred, truth, ...) {
-  predob = prediction(pred, truth)
-  perf = performance(predob, "tpr", "fpr")
-  plot(perf, ...)
-}
-```
-
-```{r}
-svmfit.opt <- svm(y ~ ., data = dat[train,], kernel = "radial",
-                  gamma = 2, cost = 1, decision.values = T)
-
-fitted <- attributes(predict(svmfit.opt, dat[train,], decision.values = T))$decision.values
-
-par(mfrow = c(1, 2))
-
-rocplot(fitted, dat[train, "y"], main = "Training Data")
-svmfit.flex <- svm(y ~ ., data = dat[train,], kernel = "radial", gamma = 50, cost = 1, decision.values = T)
-fitted <- attributes(predict(svmfit.flex, dat[train,], decision.values = T))$decision.values
-rocplot(fitted, dat[train, "y"], add = T, col = "red")
-```
-
-```{r}
-fitted <- attributes(predict(svmfit.opt, dat[-train, ], decision.values = T))$decision.values
-rocplot(fitted, dat[-train, "y"], main = "Test Data")
-fitted <- attributes(predict(svmfit.flex, dat[-train,], decision.values = T))$decision.values
-rocplot(fitted, dat[-train, "y"], add = T, col = "red")
-```
-
-### SVM w/ Multiple Classes
-
-```{r}
-set.seed(1)
-
-x <- rbind(x, matrix(rnorm(50*2), ncol = 2))
-y <- c(y, rep(0, 50))
-
-x[y == 0, 2] = x[y == 0, 2] + 2
-dat = data.frame(x = x, y = as.factor(y))
-
-par(mfrow = c(1, 1))
-plot(x, col = (y + 1))
-```
-
-```{r}
-svmfit <- svm(y ~ ., data = dat, kernel = "radial", cost = 10, gamma = 1)
-plot(svmfit, dat)
-```
-
-### Gene Expression Data
-
-```{r}
-khan <- ISLR::Khan
-```
-
-```{r}
-table(khan$ytrain)
-table(khan$ytest)
-```
-
-```{r}
-dat <- data.frame(x = khan$xtrain, y = as.factor(khan$ytrain))
-out <- svm(y ~ ., data = dat, kernel = "linear", cost = 10)
-
-summary(out)
-```
-
-```{r}
-table(out$fitted, dat$y)
-```
-
-```{r}
-dat.te <- data.frame(x = khan$xtest, y = as.factor(khan$ytest))
-pred.te <- predict(out, newdata = dat.te)
-
-table(pred.te, dat.te$y)
-```
-
-### Applied
-
-4.) Generate a simulated two-class data set with 100 observations and two features in which there is a visible but non-linear separation between the two classes.
-Show that in this setting, a support vector machine with a polynomial kernel (with degree greater than 1) or a radial kernel will outperform a support vector classifier on the training data.
-Which technique performs best on the test data?
-Make plots and report training and test error rates in order to back up your assertions.
-
-Generate data and plot:
-
-```{r}
-set.seed(1)
-
-transl <- 3
-X <- matrix(rnorm(100 * 2), ncol = 2)
-X[1:30, ] <- X[1:30, ] + transl
-X[31:60, ] <- X[31:60, ] - transl
-y <- c(rep(0, 60), rep(1, 40))
-dat <- data.frame(x = X, y = as.factor(y))
-
-par(mfrow = c(1,1))
-plot(X, col = y + 1)
-```
-
-Split to training and test set:
-
-```{r}
-train <- sample(100, 80)
-dat.train <- dat[train, ]
-dat.test <- dat[-train, ]
-```
-
-```{r}
-svm.lin <- svm(y ~ ., data = dat.train, kernel = 'linear', scale = FALSE)
-plot(svm.lin, data = dat.train)
-```
-
-```{r}
-summary(svm.lin)
-```
-
-Calculate the training error of the support vector classifier:
-
-```{r}
-table(predict = svm.lin$fitted, truth = dat.train$y)
-```
-
-Fit with polynomial kernel and calculate the training error rate:
-
-```{r}
-svm.poly <- svm(y ~ ., data = dat.train, kernel = 'polynomial', scale = FALSE)
-plot(svm.poly, data = dat.train)
-```
-
-```{r}
-table(predict = svm.poly$fitted, truth = dat.train$y)
-```
-
-Fit with radial kernel and calculate the traing error rate:
-
-```{r}
-svm.rad <- svm(y ~ ., data = dat.train, kernel = 'radial', scale = FALSE)
-plot(svm.rad, data = dat.train)
-```
-
-```{r}
-table(predict = svm.rad$fitted, truth = dat.train$y)
-```
-
-Compare the test errors of the 3 kernels:
-
-```{r}
-lin.pred <- predict(svm.lin, dat.test)
-table(predict = lin.pred, truth = dat.test$y)
-```
-
-```{r}
-poly.pred <- predict(svm.poly, dat.test)
-table(predict = poly.pred, truth = dat.test$y)
-```
-
-```{r}
-rad.pred <- predict(svm.rad, dat.test)
-table(predict = rad.pred, truth = dat.test$y)
-```
-
-5.) We have seen that we can fit an SVM with a non-linear kernel in order to perform classification using a non-linear decision boundary.
-We will now see that we can also obtain a non-linear decision boundary by performing logistic regression using non-linear transformations of the features.
-
-a.) Generate a data set with n = 500 and p = 2, such that the observations belong to two classes with a quadratic decision boundary between them:
-
-```{r}
-set.seed(1)
-
-x1 <- runif(500) - 0.5
-x2 <- runif(500) - 0.5
-y <- as.integer(x1 ^ 2 - x2 ^ 2 > 0)
-```
-
-b.) Plot the observations, colored according to their class labels. Your plot should display X 1 on the x-axis, and X 2 on the y-axis:
-
-```{r}
-plot(x1[y == 0], x2[y == 0], col = "red", xlab = "X1", ylab = "X2")
-points(x1[y == 1], x2[y == 1], col = "blue")
-```
-
-Fit a logistic regression model to the data, using X1 and X2 as predictors.
-
-```{r}
-dat <- data.frame(x1 = x1, x2 = x2, y = as.factor(y))
-lr.fit <- glm(y ~ ., data = dat, family = 'binomial')
-```
-
-d.) Apply this model to the training data in order to obtain a predicted class label for each training observation.
-Plot the observations, colored according to the predicted class labels.
-The decision boundary should be linear.
-
-```{r}
-lr.prob <- predict(lr.fit, newdata = dat, type = 'response')
-lr.pred <- ifelse(lr.prob > 0.5, 1, 0)
-plot(dat$x1, dat$x2, col = lr.pred + 2)
-```
-
-e.) Now ﬁt a logistic regression model to the data using non-linear functions of X1 and X2 as predictors (e.g. X21 , X1×X2, log(X2), and so forth).
-
-```{r}
-lr.nl <- glm(y ~ poly(x1, 2) + poly(x2, 2), data = dat, family = 'binomial')
-```
-
-f.) Apply this model to the training data in order to obtain a predicted class label for each training observation.
-Plot the observations, colored according to the predicted class labels.
-The decision boundary should be obviously non-linear.
-If it is not, then repeat (a)-(e) until you come up with an example in which the predicted class labels are obviously non-linear.
-
-```{r}
-lr.prob.nl <- predict(lr.nl, newdata = dat, type = 'response')
-lr.pred.nl <- ifelse(lr.prob.nl > 0.5, 1, 0)
-plot(dat$x1, dat$x2, col = lr.pred.nl + 2)
-```
-
-g.) Fit a support vector classiﬁer to the data with X1 and X2 as predictors.
-Obtain a class prediction for each training observation.
-Plot the observations, colored according to the predicted class labels.
-
-```{r}
-svm.lin <- svm(y ~ ., data = dat, kernel = 'linear', cost = 0.01)
-plot(svm.lin, dat)
-```
-
-h.) Fit a SVM using a non-linear kernel to the data. Obtain a class prediction for each training observation. Plot the observations, colored according to the predicted class labels.
-
-```{r}
-svm.nl <- svm(y ~ ., data = dat, kernel = 'radial', gamma = 1)
-plot(svm.nl, data = dat)
-```
-
-6.) At the end of Section 9.6.1, it is claimed that in the case of data that is just barely linearly separable, a support vector classiﬁer with a small value of cost that misclassiﬁes a couple of training observations may perform better on test data than one with a huge value of cost that does not misclassify any training observations.
-You will now investigate this claim.
-
-a.) Generate two-class data with p = 2 in such a way that the classes are just barely linearly separable.
-
-```{r}
-set.seed(1)
-obs = 1000
-x1 <- runif(obs, min = -4, max = 4)
-x2 <- runif(obs, min = -1, max = 16)
-y <- ifelse(x2 > x1 ^ 2, 0, 1)
-dat <- data.frame(x1 = x1, x2 = x2, y = as.factor(y))
-train <- sample(obs, obs/2)
-dat.train <- dat[train, ]
-dat.test <- dat[-train, ]
-par(mfrow = c(1,2))
-plot(dat.train$x1, dat.train$x2, col = as.integer(dat.train$y) + 1, main = 'training set')
-plot(dat.test$x1, dat.test$x2, col = as.integer(dat.test$y) + 1, main = 'test set')
-```
-
-b.) Compute the cross-validation error rates for support vector classiﬁers with a range of cost values.
-How many training errors are misclassiﬁed for each value of cost considered, and how does this relate to the cross-validation errors obtained?
-
-```{r}
-set.seed(1)
-cost.grid <- c(0.001, 0.01, 0.1, 1, 5, 10, 100, 10000)
-tune.out <- tune(svm, y ~., data = dat.train, kernel = 'linear', ranges = list(cost = cost.grid))
-```
-
-```{r}
-summary(tune.out)
-```
-
-c.) Generate an appropriate trest data set, and compute the test errors corresponding to each of the values of cost considered.
-Which value of cost leads to the fewest test errors, and how does this compare to the values of cost that generate the fewest train errors?
-
-```{r}
-err.rate.train <- rep(NA, length(cost.grid))
-for (cost in cost.grid) {
-  svm.fit <- svm(y ~ ., data = dat.train, kernel = 'linear', cost = cost)
-  plot(svm.fit, data = dat.train)
-  res <- table(prediction = predict(svm.fit, newdata = dat.train), truth = dat.train$y)
-  err.rate.train[match(cost, cost.grid)] <- (res[2,1] + res[1,2]) / sum(res)
-}
-```
-
-```{r}
-paste('The cost', cost.grid[which.min(err.rate.train)], 'has the minimum training error:', min(err.rate.train))
-```
-
-7.) In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the Auto data set.
-
-```{r}
-auto <- as.data.table(ISLR::Auto)
-```
-
-a.) Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage with below the median.
-
-```{r}
-auto$eco <- as.factor(ifelse(auto$mpg < median(auto$mpg), 0, 1))
-training <- auto[, !"mpg"]
-```
-
-b.) Fit a support vector classificer to the data with various values of cost, in order to predict whetehr a car gets high or low gas mileage.
-Report the cross-validation errors associated with different values of this parameter.
-
-```{r}
-cost.grid <- c(0.001, 0.1, 1, 100)
-set.seed(1)
-tune.res <- tune(svm, eco ~ ., data = training, kernel = 'linear', ranges = list(cost = cost.grid))
-summary(tune.res)
-
-plot(tune.res$best.model, training, cylinders ~ displacement)
-plot(tune.res$best.model, training, horsepower ~ weight)
-```
-
-c.) Now repeat this with SVMs with radial and polynomial basis kernels, which different values of gamma and degree.
-
-```{r}
-cost.grid <- c(0.001, 0.1, 1, 100)
-set.seed(1)
-tune.res <- tune(svm, eco ~ ., data = training, kernel = 'radial', ranges = list(cost = cost.grid))
-summary(tune.res)
-
-plot(tune.res$best.model, training, cylinders ~ displacement)
-plot(tune.res$best.model, training, horsepower ~ weight)
-plot(tune.res$best.model, training, acceleration ~ year)
-```
-
-```{r}
-cost.grid <- c(0.001, 0.1, 1, 100)
-set.seed(1)
-tune.res <- tune(svm, eco ~ ., data = training, kernel = 'polynomial', ranges = list(cost = cost.grid))
-summary(tune.res)
-
-plot(tune.res$best.model, training, cylinders ~ displacement)
-plot(tune.res$best.model, training, horsepower ~ weight)
-plot(tune.res$best.model, training, acceleration ~ year)
-```
-
-8.) This problem involves the OJ data set which is part of the ISLR package.
-
-```{r}
-oj <- as.data.table(ISLR::OJ)
-```
-
-a.) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.
-
-```{r}
-train <- sample(1:nrow(oj), 800)
-
-oj.train <- oj[train,]
-oj.test <- oj[-train,]
-```
-
-b.) Fit a support vector machine classifier to the training data using cost = 0.01, with purchase as the response and other variables as the predictors.
-
-```{r}
-svmfit <- svm(Purchase ~ ., data = oj.train, kernel = "linear", cost = 0.01)
-
-summary(svmfit)
-```
-
-c.) What are the train and test error rates?
-
-Test error rate:
-
-```{r}
-table(oj.train$Purchase, predict(svmfit))
-
-mean(oj.train$Purchase == predict(svmfit))
-```
-
-Test error rate:
-
-```{r}
-table(oj.test$Purchase, predict(svmfit, newdata = oj.test))
-
-mean(oj.test$Purchase == predict(svmfit, newdata = oj.test))
-```
-
-d.) Use the tune() function to select an optimal model.
-
-```{r}
-cost.grid <- c(0.001, 0.1, 1, 100, 250)
-out <- tune(svm, Purchase ~ ., data = oj.train, kernel = "linear", ranges = list(cost = cost.grid))
-
-summary(out)
-```
-
-e.) Compute the training and test error rates using this new value for cost:
-
-```{r}
-svmfit <- out$best.model
-
-table(oj.train$Purchase, predict(svmfit))
-
-mean(oj.train$Purchase == predict(svmfit))
-```
-
-```{r}
-table(oj.test$Purchase, predict(svmfit, newdata = oj.test))
-
-mean(oj.test$Purchase == predict(svmfit, newdata = oj.test))
-```
-
-f.) Repeat parts b-d using a radial kernel.
-
-```{r}
-cost.grid <- c(0.001, 0.1, 1, 100, 250)
-out <- tune(svm, Purchase ~ ., data = oj.train, kernel = "radial", ranges = list(cost = cost.grid))
-
-summary(out)
-
-svmfit <- out$best.model
-
-table(oj.train$Purchase, predict(svmfit))
-mean(oj.train$Purchase == predict(svmfit))
-
-table(oj.test$Purchase, predict(svmfit, newdata = oj.test))
-mean(oj.test$Purchase == predict(svmfit, newdata = oj.test))
-```
-
-h.) Overall, which approach seems to give the best results?
-
-The linear kernel has the best test error rate.
 

### 2020-03-01 Hands-on Machine Learning/01_Data-Sets.rmd
sln clean-up.
sln clean-up.

--- 2019-11-23 template.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey} - Title}
-   \rfoot{\color{headergrey} -Section}
-   \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey} -LF}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}The data sets}
+   \rfoot{\color{headergrey}\thepage}
+   \lfoot{\color{headergrey}Chapter 1}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
-library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
-library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
-library(MASS, quietly = TRUE, warn.conflicts = FALSE)
-library(copula, quietly = TRUE, warn.conflicts = FALSE)
-library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
-library(forecast, quietly = TRUE, warn.conflicts = FALSE)
-library(tseries, quietly = TRUE, warn.conflicts = FALSE)
-library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
-library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
-library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-data.dir <- "D:/Projects/Statistical-Computing/datasets/"
-
-setwd("D:/Projects/Statistical-Computing/RDS")
+data.dir <- "D:/Projects/Machine-Learning/datasets/"
 
@@
 
-#### Section
+#### Data Sets
 
-### Subsection
+### Ames
+
+```{r, echo = T}
+
+ames <- AmesHousing::make_ames()
+
+dim(ames)
+
+head(ames)
+
+```
+
+### Attrition
+
+```{r, echo = T}
+attrition <- rsample::attrition
+
+dim(attrition)
+
+head(attrition)
+```
+
+### MNIST
+
+```{r, echo = T}
+mnist <- dslabs::read_mnist()
+
+dim(mnist$train$images)
+
+head(mnist$train$labels)
+```
+
+### Grocery
+
+```{r, echo = T}
+
+url <- "https://koalaverse.github.io/homlr/data/my_basket.csv"
+
+my_basket <- readr::read_csv(url)
+
+dim(my_basket)
+
+head(my_basket)
+```

### 2020-03-01 Hands-on Machine Learning/02_Modeling-Process.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Modeling Process}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 2}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init()

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Modeling Process

### Data Sets

#### Ames

```{r, echo = T}
ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)
```

#### Attrition

```{r, echo = T}
attrition <- rsample::attrition

churn <- attrition %>%
   mutate_if(is.ordered, .funs = factor, ordered = F)

churn <- as.h2o(churn)
```

### Splitting

```{r, echo = T}
set.seed(123)

# Base R

index_1 <- sample(1:nrow(ames), round(nrow(ames) * .7), replace = F)
train_1 <- ames[index_1, ]
test_1 <- ames[-index_1, ]

# Using caret package

set.seed(123)

index_2 <- createDataPartition(ames$Sale_Price, p = 0.7,
                               list = F)
train_2 <- ames[index_2, ]
test_2 <- ames[-index_2, ]

# Using rsample package

set.seed(123)

split_1 <- initial_split(ames, prop = 0.7)
train_3 <- training(split_1)
test_3 <- testing(split_1)

# Using h20 package

split_2 <- h2o.splitFrame(ames.h2o, ratios = 0.7,
                          seed = 123)

train_4 <- split_2[[1]]
test_4 <- split_2[[2]]

```

```{r, echo = T, fig.width=8, fig.height=4}

p1 <- ggplot() +
   geom_density(data = train_1, aes(Sale_Price), col = "cornflowerblue") +
   geom_density(data = test_1, aes(Sale_Price), col = "darkred") +
   labs(title = "Base R")

p2 <- ggplot() +
   geom_density(data = train_2, aes(Sale_Price), col = "cornflowerblue") +
   geom_density(data = test_2, aes(Sale_Price), col = "darkred") +
   labs(title = "caret")

p3 <- ggplot() +
   geom_density(data = train_3, aes(Sale_Price), col = "cornflowerblue") +
   geom_density(data = test_3, aes(Sale_Price), col = "darkred") +
   labs(title = "rsample")

p4 <- ggplot() +
   geom_density(data = as.data.table(train_4), aes(Sale_Price), col = "cornflowerblue") +
   geom_density(data = as.data.table(test_4), aes(Sale_Price), col = "darkred") +
   labs(title = "h2o")

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

### Stratified sampling

Original Response Distribution

```{r, echo = T}
table(as.data.table(churn)$Attrition) %>% prop.table()
```

Stratified Sampling with the rsample package

```{r, echo = T}
set.seed(123)

split_strat <- initial_split(attrition, prop = 0.7,
                             strata = "Attrition")

train_strat <- training(split_strat)
test_strat <- testing(split_strat)

table(train_strat$Attrition) %>% prop.table()
table(test_strat$Attrition) %>% prop.table()
```

### Class Imbalances

Imbalanced data can have significant impact on model perfomance and predictions. A class imbalance is when there is a disproportionate distribution of a specific value.

Example: Defaults (5%) vs Non-defaults (95%)

Generally, the way to resolve is with either up-sampling or down-sampling.

_Down-sampling_ balances the dataset by reducing the size of the abundant class(es) to match the frequencies in the least prevalant class. This method is used when the quantity of the data is sufficent.

_Up-sampling_ is is used with the quantity of the data is insufficent for down-sampling. This technique balances the dataset by increasing the size of the rarer samples.

### Cross-Validation

Cross-validation is a resampling method that randomly divides the training data into _k_ groups (aka folds) of approximately equal size. The model is fit on k - 1 folds and then the remaining folds are used for performance evaluation. This procedure is repeated _k_ times; each time a different fold is treated as the validtion set. After the model runs on all folds, the results are averaged across all runs for the final indicator.

+ Rules of thumb for size of _k_: 5, 10
+ Most extreme k = n (or leave-one-out cross validation)

Examples:

+ h2o CV

```{r, echo = T}
h2o.cv <- h2o.glm(
   x = c("Sale_Condition"),
   y = c("Sale_Price"),
   training_frame = ames.h2o,
   nfolds = 10 # perform 10-fold CV
)
```

+ rsample CV

```{r, echo = T}
rsample::vfold_cv(ames, v = 10) # nested df
```

### Bootstrapping

Bootstrapping is another resampling technique that uses random sampling _with replacement_. A bootstrap sample is constructed using the same size as the origional sample.

Observations not contained in a particular bootstrap sample are called out-of-bag (__OOB__).

r+ sample bootstrap

```{r, echo = T}
bootstraps(ames, times = 10)
```

### Bias / Variance trade-off

__Bias__ is the difference between the expected (or average) prediction of our model and the correct value which we are trying to predict. It measures how far off in general a model's predictions are from the correct value.

Models with high bias are rarely affected by the noise introduced by resampling.

__Variance__ is the error due to the variabiliy of the data vs a models prediction at that point.

_High variance models are more prone to overfitting, using resampling procedures are critical to reduce this risk._

### Hyperparameter Tuning

Hyperparameters (aka _tuning parameters_) are the "knobs to twiddle" to control the complexity of a machine learning algorithm, and therefor the bias/variance trade-off.

+ One way to perform hyperparameter tuning is to fiddle with hyperparamters manually until you find a great combination.
+ There are several automated mechanism that can help here: cartesian grid search, random grid search, etc.

### Model Evaluation

The traditional approach to assessing model fit is to assess the residuals of the model and goodness-of-fit ($R^2$). However, this can lead to misleading conclusions.

A more robust approach is to use a loss fuction. Loss fuctions are metrics that compare the predicted value to the actual value, the output of this function is often referred to as the _error_ or the _pseudo residual_.

### Regression Models

+ __MSE__: Mean squared error is the average of the squared error (MSE = $\frac{1}{n}\sum^{n}_{i=1}(Y_i - \hat{Y}_i)$). The sqquared component results in larger errors having larger penalties. This (along with RMSE) is the most common error metric to use. __Objective: minimize__

+ __RMSE__: Root mean squared error. This simply takes the square root of the MSE metric (RMSE = $\sqrt{\frac{1}{n}\sum^n_{i=1}{(y_i - \hat{y}_i)}^2}$) so that your error is in the same units as your response variable. If your response variable units are dollars, the units of MSE are dollars-squared, but the RMSE will be in dollars. __Objective: minimize__

+ __Deviance__: Short for mean residual deviance. In essence, it provides a degree to which a model explains the variation in the set of data when using maximum likelihood estimation. Essentially, this computes a saturated model (i.e., fully featured model) to an unsaturaded model (i.e., the intercept model or the average model). If the response variable distribution is Gaussian, then it will be approximately equal to MSE. When not, it usually gives a more usefull estimate of error. Deviance is often used with classification models. __Objective: minimize__

+ __MAE__: Mean absolute error. Similar to MSE but rather than squaring, it just takes the mean absolute difference between the actual and predicted values (MAE = $\frac{1}{n}\sum^n_{i=1}{({|y_i - \hat{y}_i|})}$). This results in less emphasis on larger errors than MSE. __Objective: minimize__

+ __RMSLE__: Root mean squared logarithmic error. Similar to RMSE but it performs a _log()_ transform on the actual and predicted values prior to computing the difference (RMSLE = $\sqrt{\frac{1}{n}\sum^n_{i=1}{(log(y_i + 1) - log(\hat{y}_i + 1)})^2}$). When your response variable has a wide range of values, large response values with large errors can dominate the MSE/RMSE metric. RMSLE minimizes this impact so that small response values with large errors can have just as meaningful of an impact as large response values with large errors. __Objective: minimize__

+ $R^2$: This is a popular metric that represents the proportion of the variance in the dependent variable that is predictable from the independent variable(s). Unfortunately, it has several limitations. For example, two models built from two different data sets could have the exact same RMSE, but if one has less variability in the response variable, then it would have a lower $R^2$ than the other. Should not place too much emphasis on this metric. __Objective: maximize__

### Classification Models

+ __Misclassification:__ This is the overall error. For example, say you are predicting 3 classes (_high, medium, low_) and each class has 25, 30, 35 observations, respectfully (90 total). If you misclassify 3 observations of class _high_, 6 of class _medium_, and 4 of class _low_, then you misclassifiedd _13_ out of 90 observations resulting in a 14% misclassification rate. __Objective: minimize__

+ Mean per class error: This is the average error rate for each class. For the above example, this would be the mean of $\frac{3}{25},\frac{6}{30},\frac{34}{35}$, which is 14.5%. If your classes are balanced this will be identical to misclassification. __Objective: minimize__

+ MSE: Mean squared error. Computes the distance from 1.0 to the probability suggested. So, say we have three classes, A, B and C and your model predicts a probability of 0.91 for A, 0.07 for B, and 0.02 for C. For example, if the correct answer was A the MSE = $0.09^2 = 0.0081$. __Objective: minimize__

+ __Cross-enthropy (aka Log Loss or Deviance)__: Similar to MSE but it incorporates a log of the predicted probability multiplied by the true class. Consequently, this metric disproportionately punished predictions where we predict a small probability for the true class, which is another way of saying having high confidence in the wrong answer is really bad. __Objective: minimize__

+ __Gini index__: Mainly used with tree-based methods and commonly referred to as a measure of _purity_ where a small value indicates that a node contains predominately observations from a single class. __Objective: minimize__

+ __Accuracy__: Overall, how often is the classifier correct? Opposite of misclassification above. Example: $\frac{TP+TN}{total}$. __Objective: maximize__

+ __Precision__: How accurately does the classifier predict events? This metric is concerned with maximizing the true positives to false positive ratio. In other words, for the number of predictions that we made, how many were correct? Example: $\frac{TP}{TP+FP}$. __Objective: maximize__

+ __Sensitivity (aka recall)__: How accurately does the classifier classify actual events? This metric is concerned with maximizing the true positives to false negatives ratio. In other words, for the events that occurred, how many did we predict? Example: $\frac{TP}{TP+FN}$. __Objective: maximize__

+ __Specificity:__ How accurately does the classifier classify actual non-events? Example: $\frac{TN}{TN + FP}$. __Objective: maximize__

+ __AUC__: Area under the curve. A good classifier will have high precision and sensitivity. This means the classifier does well when it predicts and event will and will not occur, which minimizes false positives and false negatives. __Objective: maximize__

### Putting it together

Stratified Sample:

```{r, echo = T}
set.seed(123)

split <- initial_split(ames, prop = 0.7,
                       strata = "Sale_Price")

ames_train <- training(split)
ames_test <- testing(split)
```

Use a k-nearest neighbor regressor (via caret)

+ Resample method: 10-fold CV
+ Grid search: hyperparameter k
+ Model training & validation: train a k-nn model using our pre-specified resampling procedure (trControl = cv)

```{r, echo = T}

# Resampling strategy
cv <- trainControl(
   method = "repeatedcv",
   number = 10,
   repeats = 5
)

# Create grid of hyperparameter values
hyper_grid <- expand.grid(k = seq(2, 25, by = 1))

# Tune a knn model using grid search
knn_fit <- train(
   Sale_Price ~.,
   data = ames_train,
   method = "knn",
   trControl = cv,
   tuneGrid = hyper_grid,
   metric = "RMSE"
)

```

```{r, echo = T}
knn_fit
```

```{r, echo = T, fig.width=8, fig.height=4}
ggplot(knn_fit) + labs(title = "KNN Grid Search")
```

### 2020-03-01 Hands-on Machine Learning/03_Feature-Target-Engineering.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Feature and Target Engineering}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 3}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init()

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Feature and Target Engineering

### Data Set

h2o

```{r, echo = T}
ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)
```

stratified (_Sale_Price_) training sample

```{r, echo = T}
set.seed(123)

split <- initial_split(ames, prop = 0.7,
                       strata = "Sale_Price")

ames_train <- training(split)
ames_test <- testing(split)
```

log transformation (Sale_Price)

```{r, echo = T}
ames_recipe <- recipe(Sale_Price ~ ., data = ames_train) %>%
   step_log(all_outcomes())

ames_recipe
```

Box-Cox transformation (example)

```{r, echo = T}
lambda <- 3

y <- forecast::BoxCox(10, lambda)

inv_box_cox <- function(x, lambda) {
   # for Box-Cox, lambda = 0 -> log transform
   if(lambda == 0) exp(x) else (lambda*x + 1)^(1/lambda)
}

inv_box_cox(y, lambda)
```

```{r engineering-distribution-comparison, echo=TRUE, message=FALSE, warning=FALSE, fig.cap="Response variable transformations.", fig.height=3, fig.width=9}
# Log transformation
train_log_y <- log(ames_train$Sale_Price)
test_log_y  <- log(ames_train$Sale_Price)

# Box Cox transformation
lambda  <- forecast::BoxCox.lambda(ames_train$Sale_Price)
train_bc_y <- forecast::BoxCox(ames_train$Sale_Price, lambda)
test_bc_y  <- forecast::BoxCox(ames_test$Sale_Price, lambda)

# Plot differences
levs <- c("Normal", "Log_Transform", "BoxCox_Transform")
data.frame(
  Normal = ames_train$Sale_Price,
  Log_Transform = train_log_y,
  BoxCox_Transform = train_bc_y
) %>%
  gather(Transform, Value) %>%
  mutate(Transform = factor(Transform, levels = levs)) %>% 
  ggplot(aes(Value, fill = Transform)) +
    geom_histogram(show.legend = FALSE, bins = 40) +
    facet_wrap(~ Transform, scales = "free_x")
```

## Missing Values

```{r, echo = T}
sum(is.na(AmesHousing::ames_raw))
```

```{r, echo = T, fig.width=8, fig.height=3.5}
AmesHousing::ames_raw %>%
   is.na() %>%
   reshape2::melt() %>%
   ggplot(aes(Var2, Var1, fill = value)) +
      geom_raster() +
      coord_flip() +
      scale_y_continuous(NULL, expand = c(0,0)) +
      scale_fill_grey(name = "",
                      labels = c("Present",
                                 "Missing")) +
   xlab("Observations") +
   theme(axis.text.y = element_text(size = 4))
```

Missing Garage?

```{r, echo = T}
AmesHousing::ames_raw %>%
   filter(is.na(`Garage Type`)) %>%
   select(starts_with("Garage"))

```

Missing values w/cluster (_visdat_)

```{r, echo = T, fig.width=8, fig.height=4}
vis_miss(AmesHousing::ames_raw, cluster = T)
```

### Missing Value Imputation

basic descriptive statistic

```{r, echo = T}
ames_recipe %>%
   step_medianimpute(Gr_Liv_Area)
```

KNN approach (typical k = 5-10)

```{r, echo = T}
ames_recipe %>%
   step_knnimpute(all_predictors(), neighbors = 6)
```

```{r engineering-imputation-examples, echo=TRUE, fig.cap="Comparison of three different imputation methods. The red points represent actual values which were removed and made missing and the blue points represent the imputed values. Estimated statistic imputation methods (i.e. mean, median) merely predict the same value for each observation and can reduce the signal between a feature and the response; whereas KNN and tree-based procedures tend to maintain the feature distribution and relationship."}
impute_ames <- ames_train
set.seed(123)
index <- sample(seq_along(impute_ames$Gr_Liv_Area), 50)
actuals <- ames_train[index, ]
impute_ames$Gr_Liv_Area[index] <- NA

p1 <- ggplot() +
  geom_point(data = impute_ames, aes(Gr_Liv_Area, Sale_Price), alpha = .2) +
  geom_point(data = actuals, aes(Gr_Liv_Area, Sale_Price), color = "red") +
  scale_x_log10(limits = c(300, 5000)) +
  scale_y_log10(limits = c(10000, 500000)) +
  ggtitle("Actual values")

# Mean imputation
mean_juiced <- recipe(Sale_Price ~ ., data = impute_ames) %>%
  step_meanimpute(Gr_Liv_Area) %>%
  prep(training = impute_ames, retain = TRUE) %>%
  juice()
mean_impute <- mean_juiced[index, ]

p2 <- ggplot() +
  geom_point(data = actuals, aes(Gr_Liv_Area, Sale_Price), color = "red") +
  geom_point(data = mean_impute, aes(Gr_Liv_Area, Sale_Price), color = "blue") +
  scale_x_log10(limits = c(300, 5000)) +
  scale_y_log10(limits = c(10000, 500000)) +
  ggtitle("Mean Imputation")

# KNN imputation
knn_juiced <- recipe(Sale_Price ~ ., data = impute_ames) %>%
  step_knnimpute(Gr_Liv_Area) %>%
  prep(training = impute_ames, retain = TRUE) %>%
  juice()
knn_impute <- knn_juiced[index, ]

p3 <- ggplot() +
  geom_point(data = actuals, aes(Gr_Liv_Area, Sale_Price), color = "red") +
  geom_point(data = knn_impute, aes(Gr_Liv_Area, Sale_Price), color = "blue") +
  scale_x_log10(limits = c(300, 5000)) +
  scale_y_log10(limits = c(10000, 500000)) +
  ggtitle("KNN Imputation")

# Bagged imputation
bagged_juiced <- recipe(Sale_Price ~ ., data = impute_ames) %>%
  step_bagimpute(Gr_Liv_Area) %>%
  prep(training = impute_ames, retain = TRUE) %>%
  juice()
bagged_impute <- bagged_juiced[index, ]

p4 <- ggplot() +
  geom_point(data = actuals, aes(Gr_Liv_Area, Sale_Price), color = "red") +
  geom_point(data = bagged_impute, aes(Gr_Liv_Area, Sale_Price), color = "blue") +
  scale_x_log10(limits = c(300, 5000)) +
  scale_y_log10(limits = c(10000, 500000)) +
  ggtitle("Bagged Trees Imputation")

gridExtra::grid.arrange(p1, p2, p3, p4, nrow = 2)
```

Increase in training time by model type:

```{r, echo = T, fig.width=9, fig.height=4}
model_results <- read_csv(paste0(data.dir, "feature-selection-impacts-results.csv")) %>%
   mutate(type = case_when(
      model %in% c("lm", "pcr", "pls", "glmnet", "lasso") ~ "Linear Models",
      model %in% c("earth", "svmLinear", "nn") ~ "Non-linear Models",
      TRUE ~ "Tree-based Models"
   )) %>%
   mutate(model = case_when(
      model == "lm" ~ "Linear regression",
      model == "earth" ~ "Multivariate adaptive regression splines",
      model == "gbm" ~ "Gradient boosting machines",
      model == "glmnet" ~ "Elastic net",
      model == "lasso" ~ "Lasso",
      model == "nn" ~ "Neural net",
      model == "pcr" ~ "Principal component regression",
      model == "pls" ~ "Partial least squares",
      model == "ranger" ~ "Random forest",
      TRUE ~ "Support vector machine"
  ))

ggplot(model_results, aes(NIP, RMSE, color = model, lty = model)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ type, nrow = 1) +
  xlab("Number of additional non-informative predictors")
```

```{r engineering-impact-on-time, echo=TRUE, fig.width=10, fig.height=3.5, fig.cap="Impact in model training time as non-informative predictors are added."}
model_results %>%
  group_by(model) %>%
  mutate(
    time_impact = time / first(time),
    time_impact = time_impact - 1
  ) %>%
  ggplot(aes(NIP, time_impact, color = model, lty = model)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ type, nrow = 1) +
    scale_y_continuous("Percent increase in training time", 
                       labels = scales::percent) +
    xlab("Number of additional non-informative predictors")
```

Rules of thumb for zero variance features:

+ The fraction of unique values over the sample size is low (say < 10%)
+ The ratio of the frequency of the most prevalent value to the frequency of the second most prevalent value is large (say > 20%)

If both of these criteria are met, then it is often advantageous to remove them from the model.

```{r, echo = T}
caret::nearZeroVar(ames_train, saveMetrics = T) %>%
   rownames_to_column() %>%
   filter(nzv)
```

## Numeric Feature Engineering

Skewness can have a drastic impact on the performance of GLMs & regularized models.

Non-parametric models are rarely affected by skewed features; however, normalizing features will not have a negative effect on these models' performance. For example, normalizing features will only shift the optimal split points in tree-based algoirthms. Consequently, when in doubt, normalize.

### Skewness

```{r engineering-normalizing}
# Normalize all numeric columns
recipe(Sale_Price ~ ., data = ames_train) %>%
  step_YeoJohnson(all_numeric())                 
```

### Standardization

```{r, echo = T}
ames_recipe %>%
   step_center(all_numeric(), -all_outcomes()) %>%
   step_scale(all_numeric(), -all_outcomes())
```

```{r engineering-standardizing, echo=TRUE, fig.height=3, fig.cap="Standardizing features allows all features to be compared on a common value scale regardless of their real value differences."}
set.seed(123)
x1 <- tibble(
   variable = "x1",
   `Real Value` = runif(25, min = -30, max = 5),
   `Standardized Value` = scale(`Real Value`) %>% as.numeric()

)

set.seed(456)
x2 <- tibble(
  variable = "x2",
  `Real value` = rlnorm(25, log(25)),
  `Standardized value` = scale(`Real value`) %>% as.numeric()
)

set.seed(789)
x3 <- tibble(
  variable = "x3",
  `Real value` = rnorm(25, 150, 15),
  `Standardized value` = scale(`Real value`) %>% as.numeric()
)

x1 %>%
  bind_rows(x2) %>%
  bind_rows(x3) %>%
  gather(key, value, -variable) %>%
  mutate(variable = factor(variable, levels = c("x3", "x2", "x1"))) %>%
  ggplot(aes(value, variable)) +
    geom_point(alpha = .6) +
    facet_wrap(~ key, scales = "free_x") +
    ylab("Feature") +
    xlab("Value")
```

## Categorical Feature Engineering

### Lumping

When a feature contains levels that have few observations.

For example:

```{r, echo = T}
count(ames_train, Neighborhood) %>% arrange(n)
```

```{r, echo = T}
count(ames_train, Screen_Porch) %>% arrange(n)
```

We can benefit from lumping these together into an "other" category when they contain less than 10% of the training sample.

__Note: This can have an adverse effect on performance__

```{r, echo = T}
lumping <- recipe(Sale_Price ~., data = ames_train) %>%
   step_other(Neighborhood, threshold = 0.01,
              other = "other") %>%
   step_other(Screen_Porch, threshold = 0.1,
              other = ">0")

apply_2_training <- prep(lumping, training = ames_train) %>%
  bake(ames_train)

# New distribution of Neighborhood
count(apply_2_training, Neighborhood) %>% arrange(n)

# New distribution of Screen_Porch
count(apply_2_training, Screen_Porch) %>% arrange(n)
```

```{r, echo = T}
dat <- data.table(id = 1:9, x = rep(c("a", "b" , "c"), 3))
dat

# full-rank
dat[, .(id, 
        `X = a` = as.numeric(x == "a"), 
        `X = b` = as.numeric(x == "b"), 
        `X = c` = as.numeric(x == "c")) ]

# one-hot (leave one out)
dat[, .(id, 
        `X = a` = as.numeric(x == "a"), 
        `X = b` = as.numeric(x == "b")) ]

```

```{r, echo = T}
# Lump levels for two features
recipe(Sale_Price ~., data = ames_train) %>%
   step_dummy(all_nominal(), one_hot = T)
```

### Label Encoding

_Label encoding_ is a pure numeric conversion of the levels of a categorical variable.

For example, the _MS_SubClass_ variable has 16 levels, which we can reencode numerically.

__Important:__ The features will be treated as ordered (ordnal encoding), so if the feature is not natually ordered, this will have a poor impact on the model.

```{r, echo = T}
count(ames_train, MS_SubClass)
```

```{r, echo = T}
# Label encoded
recipe(Sale_Price ~ ., data = ames_train) %>%
   step_integer(MS_SubClass) %>%
   prep(ames_train) %>%
   bake(ames_train) %>%
   count(MS_SubClass)
```

Examples of ordnal features:

```{r, echo = T}
ames_train %>% select(contains("Qual"))
```

```{r, echo = T}
count(ames_train, Overall_Qual)
```

```{r, echo = T}
# Label encoded
recipe(Sale_Price ~., data = ames_train) %>%
   step_integer(Overall_Qual) %>%
   prep(ames_train) %>%
   bake(ames_train) %>%
   count(Overall_Qual)
```

### Alternatives

Target encoding:

```{r engineering-target-encoding, echo=TRUE, fig.cap='Example of target encoding the Neighborhood feature of the Ames housing data set.'}
ames_train %>%
  group_by(Neighborhood) %>%
  summarize(`Avg Sale_Price` = mean(Sale_Price, na.rm = TRUE)) %>%
  head(10) %>%
  kable(caption = "Example of target encoding the Neighborhood feature of the Ames housing data set.") %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE)
```

```{r engineering-proportion-encoding, echo=TRUE, fig.cap='Example of categorical proportion encoding the Neighborhood feature of the Ames housing data set.'}
ames_train %>%
  count(Neighborhood) %>%
  mutate(Proportion = n / sum(n)) %>%
  select(-n) %>%
  head(10) %>%
  kable(caption = 'Example of categorical proportion encoding the Neighborhood feature of the Ames housing data set.') %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE)
```

## Dimension Reduction

Example PCA using _resample_ package.

```{r engineering-pca}
recipe(Sale_Price ~ ., data = ames_train) %>%
  step_center(all_numeric()) %>%
  step_scale(all_numeric()) %>%
  step_pca(all_numeric(), threshold = .95)
```

## Full Recipe

Full blueprint recipe applied to training and test data.

```{r, echo = T}
blueprint <- recipe(Sale_Price ~ ., data = ames_train) %>%
   step_nzv(all_nominal()) %>%
   step_integer(matches("Qual|Cond|QC|Qu")) %>%
   step_center(all_numeric(), -all_outcomes()) %>%
   step_scale(all_numeric(), -all_outcomes()) %>%
   step_pca(all_numeric(), -all_outcomes())

blueprint

prepare <- prep(blueprint, training = ames_train)
prepare

baked_train <- bake(prepare, new_data = ames_train)
baked_test <- bake(prepare, new_data = ames_test)

baked_train
```

Full recipe with cross-validation & grid search using carat.

```{r engineering-knn-blueprint}
blueprint <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_nzv(all_nominal()) %>%
  step_integer(matches("Qual|Cond|QC|Qu")) %>%
  step_center(all_numeric(), -all_outcomes()) %>%
  step_scale(all_numeric(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE)
```

```{r engineering-knn-with-blueprint}
# Specify resampling plan
cv <- trainControl(
  method = "repeatedcv", 
  number = 10, 
  repeats = 5
)

# Construct grid of hyperparameter values
hyper_grid <- expand.grid(k = seq(2, 25, by = 1))

# Tune a knn model using grid search
knn_fit2 <- train(
  blueprint, 
  data = ames_train, 
  method = "knn", 
  trControl = cv, 
  tuneGrid = hyper_grid,
  metric = "RMSE"
)
```

```{r engineering-knn-with-blueprint-assess, fig.height=3, fig.cap="Results from the same grid search performed in Section 2.7 but with feature engineering performed within each resample."}
# print model results
knn_fit2
# plot cross validation results
ggplot(knn_fit2)
```

### 2020-03-01 Hands-on Machine Learning/04_Linear-Regression.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Linear Regression}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 4}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

# Additional Data
library(AppliedPredictiveModeling, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init()

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Linear Regression

### Data Set

h2o

```{r, echo = T}
ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)
```

stratified (_Sale_Price_) training sample

```{r, echo = T}
set.seed(123)

split <- initial_split(ames, prop = 0.7,
                       strata = "Sale_Price")

ames_train <- training(split)
ames_test <- testing(split)
```

## Simple Linear Model

```{r, echo = T}
model1 <- lm(Sale_Price ~ Gr_Liv_Area, data = ames_train)
```

```{r, echo = T, fig.height = 4.5}
# Fitted regression line (full training)
p1 <- model1 %>%
   broom::augment() %>%
   ggplot(aes(Gr_Liv_Area, Sale_Price)) +
   geom_point(size = 1, alpha = 0.3) +
   geom_smooth(se = F, method = "lm") +
   scale_y_continuous(labels = scales::dollar) +
   ggtitle("Fitted regression line")

# Fitted regression line (restricted range)
p2 <- model1 %>%
   broom::augment() %>%
   ggplot(aes(Gr_Liv_Area, Sale_Price)) +
   geom_segment(aes(x = Gr_Liv_Area, y = Sale_Price,
                    xend = Gr_Liv_Area, yend = .fitted),
                alpha = .3) +
   geom_point(size = 1, alpha = 0.3) +
   geom_smooth(se = F, method = "lm") +
   scale_y_continuous(labels = scales::dollar) +
   ggtitle("Fitted regression line (with residuals)")

grid.arrange(p1, p2, nrow = 2)
```

```{r, echo=T}
summary(model1)
```

```{r}
sigma(model1) # RMSE, also Residual Standard Error in summary()
sigma(model1)^2 # MSE
```

### Inference

The variability of an estimate is its _standard error (SE)_, the square root of its variance.

t-test for the coefficents are simply the estimated coefficent divided by the standard error (t value = Estimate / Std. Error)

t-test measure the number of standard deviations each coefficent is away from zero (basically abs(T) > 2 is significant at 95% conf)

The confidence interval for coefficents is:

\[\hat{\beta_j} \pm t_{1-\alpha/2, n-p}\hat{SE}(\hat{\beta_j})\]

```{r, echo = T}
confint(model1, level = .95)
```

Interpretation: We are 95% confident that each one unit increase in Gr_Liv_Area adds between 109.9 and 119.8 dollars to the sale price.

Linear Regression Assumptions:

+ 1.) Independent observations
+ 2.) The random errors have mean zero, and constant variance
+ 3.) The random errors are normally distributed

## Multiple Linear Regression

```{r, echo = T}
(model2 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train))
# Equivalent
(model2 <- update(model1, . ~ . + Year_Built))

round(coef(model2), 3)
```

```{r, echo = T}
summary(model3 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built + Gr_Liv_Area:Year_Built, data = ames_train))

round(coef(model3), 3)
```

```{r, echo = T, fig.height=5}
fit1 <- lm(Sale_Price ~ Gr_Liv_Area + Year_Built, data = ames_train)
fit2 <- lm(Sale_Price ~ Gr_Liv_Area * Year_Built, data = ames_train)

# Regression plane
plot_grid <- expand.grid(
   Gr_Liv_Area = seq(from = min(ames_train$Gr_Liv_Area), to = max(ames_train$Gr_Liv_Area),
                     length = 100),
   Year_Built = seq(from = min(ames_train$Year_Built), to = max(ames_train$Year_Built),
                     length = 100)
)

plot_grid$y1 <- predict(fit1, newdata = plot_grid)
plot_grid$y2 <- predict(fit2, newdata = plot_grid)

# Level plots
p1 <- ggplot(plot_grid, aes(x = Gr_Liv_Area, y = Year_Built,
                      z = y1, fill = y1)) +
   geom_tile() +
   geom_contour(color = "white") +
   viridis::scale_fill_viridis(name = "Predicted\nvalue", option = "inferno") +
   theme_bw() +
   ggtitle("Main Effects Only")

p2 <- ggplot(plot_grid, aes(x = Gr_Liv_Area, y = Year_Built,
                      z = y2, fill = y2)) +
   geom_tile() +
   geom_contour(color = "white") +
   viridis::scale_fill_viridis(name = "Predicted\nvalue", option = "inferno") +
   theme_bw() +
   ggtitle("Main effect with two-way interaction")

gridExtra::grid.arrange(p1, p2, nrow = 2)
```

Full Model

```{r, echo = T}
model3 <- lm(Sale_Price ~ ., data = ames_train)
broom::tidy(model3)
```

### Assessing Model Accuracy

Models 1/2/3:

```{r, echo = T}
# Train model using 10-fold cross-validation

set.seed(123) # for reproducibility

(cv_model <- train(
   form = Sale_Price ~ Gr_Liv_Area,
   data = ames_train,
   method = "lm",
   trControl = trainControl(method = "cv", number = 10)
))

set.seed(123)

cv_model2 <- train(
   Sale_Price ~ Gr_Liv_Area + Year_Built,
   data = ames_train,
   method = "lm",
   trControl = trainControl(method = "cv", number = 10)
)

set.seed(123)

suppressWarnings({
   cv_model3 <- train(
      Sale_Price ~ .,
      data = ames_train,
      method = "lm",
      trControl = trainControl(method = "cv", number = 10)
   )
})
```

__Accuracy__:

```{r}
summary(resamples(list(
   model1 = cv_model,
   model2 = cv_model2,
   model3 = cv_model3
)))
```

### Model Concerns / Assumptions

#### 1.) Linear Relationships

Possible solution to non-linear relationship is by variable transformations:

```{r, echo = T}

p1 <- ggplot(ames_train, aes(Year_Built, Sale_Price)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F) +
   scale_y_continuous(labels = scales::dollar) +
   xlab("Year Built") +
   ggtitle(paste("Non-transformed variables with a \n", "non-linear relationship"))

p2 <- ggplot(ames_train, aes(Year_Built, Sale_Price)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F, method = "lm") +
   scale_y_log10("Sale Price", labels = scales::dollar, breaks = seq(0, 400000, by = 100000)) +
   xlab("Year Built") +
   ggtitle(paste("Transforming variables can provide a \n", "near-linear relationship"))

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

#### 2.) Constant variance among residuals

Linear models assume constant variance amoung error terms (_homoscedasticity_).

Notice the cone shape in Model 1:

```{r, echo = T}
df1 <- broom::augment(cv_model$finalModel, data = ames_train)

p1 <- ggplot(df1, aes(.fitted, .resid)) +
   geom_point(size = 1, alpha = .4) +
   xlab("Predicted Values") +
   ylab("Residuals") +
   ggtitle("Model 1", subtitle = "Sale_Price ~ Gr_Liv_Area")

df2 <- broom::augment(cv_model3$finalModel, data = ames_train)

p2 <- ggplot(df2, aes(.fitted, .resid)) +
   geom_point(size = 1, alpha = .4) +
   xlab("Predicted Values") +
   ylab("Residuals") +
   ggtitle("Model 3", subtitle = "Sale_Price ~ .")

gridExtra::grid.arrange(p1, p2, nrow =1)
```

#### 3.) No autocorrelation

Residuals should be uncorrelated and i.i.d.

```{r}
df1 <- mutate(df1, id = row_number())
df2 <- mutate(df2, id = row_number())

p1 <- ggplot(df1, aes(id, .resid)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F) +
   xlab("Row ID") +
   ylab("Residuals") +
   ggtitle("Model 1", subtitle = "Correlated Residuals")

p2 <- ggplot(df2, aes(id, .resid)) +
   geom_point(size = 1, alpha = .4) +
   geom_smooth(se = F) +
   xlab("Row ID") +
   ylab("Residuals") +
   ggtitle("Model 3", subtitle = "Uncorrelated Residuals")

gridExtra::grid.arrange(p1, p2, nrow = 1)

```

Model 1 has a distict pattern to the residuals (due to being ordered by neighborhood, which is unaccounted for in the model)

#### 4.) More observations than predictors

The number of features cannot exceed the number of observations.

(not a problem in this example)

#### 5.)

No or little multicollinearity.

Collinearity refers to the situation where two or more predictor variables are closely related to one another.

```{r}
cor(ames_train$Garage_Area, ames_train$Garage_Cars)
```

Example: _Garage_Area_ and _Garage_Cars_ are highly correlated.

```{r}
summary(cv_model3) %>%
   broom::tidy() %>%
   filter(term %in% c("Garage_Area", "Garage_Cars"))
```

Only _Garage_Area_ is significant (p < 0.05),

However, refit w/o _Garage_Area_ term, 
```{r}
# model without Garage Area
set.seed(123)

mod_wo_Garage_Cars <- train(
   Sale_Price ~ .,
   data = select(ames_train, -Garage_Area),
   method = "lm",
   trControl = trainControl(method = "cv", number = 10)
)

summary(mod_wo_Garage_Cars) %>%
   broom::tidy() %>%
   filter(term == "Garage_Cars")
```

_Garage_Cars_ becomes significant ( p < 0.001 )

Best to check VIF scores (variance inflation factor)

### Principal Component Analysis

PCA -> Linear factorization of features

PCR -> Principal Component Regression

```{r}
# performs 10-fold cross validation on a PCR model tuning the
# number of principal components to use as predictors from 1-20

set.seed(123)

cv_model_pcr <- train(
   Sale_Price ~ .,
   data = ames_train,
   method = "pcr",
   trControl = trainControl(method = "cv", number = 10),
   preProcess = c("zv", "center", "scale"),
   tuneLength = 20
)

cv_model_pcr$bestTune

ggplot(cv_model_pcr)

summary(cv_model_pcr)
```

### Partial least squares

PLS can be viewed as a supervised dimension reduction procedure.

Below is a comparision between PCR and PLS:

```{r}
data(solubility)

df <- cbind(solTrainX, solTrainY)

pca_df <- recipe(solTrainY ~ ., data = df) %>%
   step_center(all_predictors()) %>%
   step_scale(all_predictors()) %>%
   step_pca(all_predictors()) %>%
   prep(training = df, retain = T) %>%
   juice() %>%
   select(PC1, PC2, solTrainY) %>%
   rename(`PCR Component 1` = "PC1", `PCR Component 2` = "PC2") %>%
   gather(component, value, -solTrainY)

pls_df <- recipe(solTrainY ~., data = df) %>%
   step_center(all_predictors()) %>%
   step_scale(all_predictors()) %>%
   step_pls(all_predictors(), outcome = "solTrainY") %>%
   prep(training = df, retain = T) %>%
   juice() %>%
   rename(`PLS Component 1` = "PLS1", `PLS Component 2` = "PLS2") %>%
   gather(component, value, -solTrainY)

pca_df %>%
   bind_rows(pls_df) %>%
   ggplot(aes(value, solTrainY)) +
   geom_point(alpha = .15) +
   geom_smooth(method = "lm", se = F, lty = "dashed") +
   facet_wrap(~ component, scales = "free") +
   labs(x = "PC Eigenvalues", y = "Response")
```

Fit PLS to ames:

```{r}

# perform 10-fold cross validation on a PLS model tuning the
# number of principal compoents to use as predictors from 1-20

set.seed(123)

cv_model_pls <- train(
   Sale_Price ~ .,
   data = ames_train,
   method = "pls",
   trControl = trainControl(method = "cv", number = 10),
   preProcess = c("zv", "center", "scale"),
   tuneLength = 20
)

# model with lowest RMSE

cv_model_pls$bestTune

ggplot(cv_model_pls)

```

### Feature interpretation

Linear models are monotonic, meaning 1 unit change in X means a constant change in Y.

In PLS we can measure the importance of features by an absolute weighted sum of the regression coefficents.

We can use vip  to extract and plot the most important variables:

```{r}
vip(cv_model_pls, num_features = 20, method = "model")
```

Alternatively, we can use PDP (partial dependence plots) to summarize the relationship.

```{r}
p1 <- pdp::partial(cv_model_pls, pred.var = "Gr_Liv_Area", grid.resolution = 20) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p2 <- pdp::partial(cv_model_pls, pred.var = "First_Flr_SF", grid.resolution = 20) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p3 <- pdp::partial(cv_model_pls, pred.var = "Total_Bsmt_SF", grid.resolution = 20) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p4 <- pdp::partial(cv_model_pls, pred.var = "Garage_Cars", grid.resolution = 4) %>% 
  autoplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

We see all 4 of the top predictive features have a positive linear relationship.

```{r}
# clean up
rm(list = ls())
```

### 2020-03-01 Hands-on Machine Learning/06_Regularized-Regression.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Regularized Regression}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 6}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init()

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Regularized Regression

### Data Sets

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)
```

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

### Overview

Regularization methods provide a means to constrain or regularize the estimated coefficients, which can reduce the variance and decrease out of sample error.

### Why regularize?

Linear Model:

```{r}
ames_sub <- ames.train %>%
   filter(Gr_Liv_Area > 1000 & Gr_Liv_Area < 3000) %>%
   sample_frac(.5)

model1 <- lm(Sale_Price ~ Gr_Liv_Area, data = ames_sub)

model1 %>%
   broom::augment() %>%
   ggplot(aes(Gr_Liv_Area, Sale_Price)) +
   geom_segment(aes(x = Gr_Liv_Area, y = Sale_Price,
                    xend = Gr_Liv_Area, yend = .fitted),
                alpha = 0.3) +
   geom_point(size = 1, color = "red") +
   geom_smooth(se = F, method = "lm") +
   scale_y_continuous(labels = scales::dollar)
```

Linear assumptions:

1.) Linear relationship

2.) More observations than features (n > p)

3.) Little to no multicolinearity

4.) Homoscedasticity (constant error variance)

When linear assumptions break-down, especially with large p, regularization methods are useful.

+ Linear OLS: min SSE = $\sum_{i=1}^n{(y_i - \hat{y})^2}$

### Feature Selection

In OLS, we have "hard threshold" methods for variable selection (forward selection, backward elimination, step-wise)

More modern approach is called "soft thresholds", which slowly pushes the effects of irrelevant features toward zero, and in some cases will zero out entire coefficients.

With wide data sets (or ones that exhibit multicolinearity) we have regularization methods (or penalized models / shrinkage methods).

**Reduces variance at expense being unbiased.**

Regularization parameter:

+ min SSE = $\sum_{i=1}^n{(y_i - \bar{y})^2} + P$

Basically, we can think of the regularization parameter as a constraint to the size of the coefficients such that the only way the coefficients can increase is if we experience a comparable decrease in the model's loss function.

Three common regularization methods:

1.) Ridge

2.) Lasso (or LASSO)

3.) Elastic net (or ENET), which is a combination of ridge and lasso.

### Ridge Penalty

Ridge penalty: $\lambda\sum^p_{j=1}\beta^2_j$

So the minimization function becomes:

$\sum_{i=1}^n{(y_i - \hat{y})^2} + \lambda\sum^p_{j=1}\beta^2_j$

The size of this penalty, referred to as $L^2$ (or Euclidiean) norm, can take on a wide range of values, which is controlled by the tuning parameter $\lambda$.

When $L^2$ is zero, there is no effect (reverts back to OLS).

However, as $\lambda \to \inf$, the penalty becomes large and forces the coefficients toward zero (but not all the way).

Ridge Penalty visually:

```{r}
boston_train_x <- model.matrix(cmedv ~ ., pdp::boston)[, -1]
boston_train_y <- pdp::boston$cmedv

# model
boston_ridge <- glmnet::glmnet(
   x = boston_train_x,
   y = boston_train_y,
   alpha = 0
)

lam <- boston_ridge$lambda %>%
   as.data.frame() %>%
   mutate(penalty = boston_ridge$a0 %>% names()) %>%
   rename(lambda = ".")

results <- boston_ridge$beta %>%
   as.matrix() %>%
   as.data.frame() %>%
   rownames_to_column() %>%
   gather(penalty, coefficients, -rowname) %>%
   left_join(lam)

result_labels <- results %>%
   group_by(rowname) %>%
   filter(lambda == min(lambda)) %>%
   ungroup() %>%
   top_n(5, wt = abs(coefficients)) %>%
   mutate(var = paste0("x", 1:5))

ggplot() +
   geom_line(data = results, aes(lambda, coefficients, group = rowname, color = rowname), show.legend = F) +
   scale_x_log10() +
   geom_text(data = result_labels, aes(lambda, coefficients, label = var, color = rowname), nudge_x = -0.06, show.legend = F)
```

In essence, the ridge penalty pushes muticolinear features together rather than allowing one to be wildy positive and one widly negative. Additionally, many of the less-imporant features also get pushed toward zero.

What ridge regression is **NOT**:

A feature selection technique. It will retain all features in the final model. Therefore, ridge regression is appropriate if you need to retain all the features, yet reduce the noise that less influential variables may create.

### Lasso penalty

Lasso stands for **L**east **A**bsolute **S**hrinkage and **S**election **O**perator.

Lasso penalty: $\lambda\sum^p_{j=1}{|\beta_j|}$

Lasso Regresion Function: $\sum_{i=1}^n{(y_i - \hat{y})^2} + \lambda\sum^p_{j=1}{|\beta_j|}$

As $\lambda \to \inf$, the lasso penalty will actually push feature coefficents to zero.

This model serves as a sort of automatic feature selection.

Visually:

```{r}

# model
boston_lasso <- glmnet::glmnet(
   x = boston_train_x,
   y = boston_train_y,
   alpha = 1
)

lam <- boston_lasso$lambda %>%
   as.data.frame() %>%
   mutate(penalty = boston_lasso$a0 %>% names()) %>%
   rename(lambda = ".")

results <- boston_lasso$beta %>%
   as.matrix() %>%
   as.data.frame() %>%
   rownames_to_column() %>%
   gather(penalty, coefficients, -rowname) %>%
   left_join(lam)

result_labels <- results %>%
   group_by(rowname) %>%
   filter(lambda == min(lambda)) %>%
   ungroup() %>%
   top_n(5, wt = abs(coefficients)) %>%
   mutate(var = paste0("x", 1:5))

ggplot() +
   geom_line(data = results, aes(lambda, coefficients, group = rowname, color = rowname), show.legend = F) +
   scale_x_log10() +
   geom_text(data = result_labels, aes(lambda, coefficients, label = var, color = rowname), nudge_x = -0.05, show.legend = F)
```

We can see that as $\lambda$ grows, the number of features retained decreases.

Lasso regression can be a good tool to extract the most consistent features.

### Elastic nets

A generalization of the ridge and lasso penalties, called _elastic nets_, combines the two penalties.

Elastic net function: min $\sum_{i=1}^n{(y_i - \hat{y})^2} + \lambda\sum^p_{j=1}\beta^2_j + \lambda\sum^p_{j=1}{|\beta_j|}$

Visually:

```{r}
# model
boston_elastic <- glmnet::glmnet(
   x = boston_train_x,
   y = boston_train_y,
   alpha = .2
)

lam <- boston_elastic$lambda %>%
   as.data.frame() %>%
   mutate(penalty = boston_elastic$a0 %>% names()) %>%
   rename(lambda = ".")

results <- boston_elastic$beta %>%
   as.matrix() %>%
   as.data.frame() %>%
   rownames_to_column() %>%
   gather(penalty, coefficients, -rowname) %>%
   left_join(lam)

result_labels <- results %>%
   group_by(rowname) %>%
   filter(lambda == min(lambda)) %>%
   ungroup() %>%
   top_n(5, wt = abs(coefficients)) %>%
   mutate(var = paste0("x", 1:5))

ggplot() +
   geom_line(data = results, aes(lambda, coefficients, group = rowname, color = rowname), show.legend = F) +
   scale_x_log10() +
   geom_text(data = result_labels, aes(lambda, coefficients, label = var, color = rowname), nudge_x = -0.05, show.legend = F)
```

### Implementation

glmnet: *alpha* parameter = penalty weight

$\alpha = 0$, ridge

$\alpha = 1$, lasso

$0 < \alpha < 1$, elastic net

```{r}
X <- model.matrix(Sale_Price ~., data = ames.train)[, -1]

Y <- log(ames.train$Sale_Price)
```

We need to ensure that all features are on a common scale (otherwise large magnitude features will have more weight).

(standardization is done automatically with _glmnet_)

```{r}
# Apply ridge regression to ames data

ridge <- glmnet(
   x = X,
   y = Y,
   alpha = 0
)
```

glmnet automatically fits models across a wide range of $\lambda$ values.

```{r}
plot(ridge, xvar = "lambda")
```

coefficients:

```{r}

params <- coef(ridge)

ridge$lambda %>% head()

coef(ridge)[c("Latitude", "Overall_QualVery_Excellent"), 100]

coef(ridge)[c("Latitude", "Overall_QualVery_Excellent"), 1]
```

### Tuning

To help find the optimal value of $\lambda$ we use k-fold cross validation.

Note: by default glmnet::cv.glmnet uses MSE as the loss function, this can be changed by altering the 'type.measurement'

```{r}
# ?glmnet::cv.glmnet

ridge <- cv.glmnet(
   x = X,
   y = Y,
   alpha = 0
)

lasso <- cv.glmnet(
   x = X,
   y = Y,
   alpha = 1
)

# plot results

par(mfrow = c(1, 2))
plot(ridge, main = "Ridge penalty\n\n")
plot(lasso, main = "Lasso penalty\n\n")
```

```{r}
# Ridge model
min(ridge$cvm)

ridge$lambda.min # lambda for this min MSE

ridge$cvm[ridge$lambda == ridge$lambda.1se] # 1-SE rule

ridge$lambda.1se # lambda for this MSE

# Lasso model
min(lasso$cvm)

lasso$lambda.min

lasso$cvm[lasso$lambda == lasso$lambda.1se]
```

Feature Reduction

```{r}
ridge_min <- glmnet(
   x = X,
   y = Y,
   alpha = 0
)

lasso_min <- glmnet(
   x = X,
   y = Y,
   alpha = 1
)

par(mfrow = c(1, 2))

# plot ridge model
plot(ridge_min, xvar = "lambda", main = "Ridge penalty\n\n")
abline(v = log(ridge$lambda.min), col = "red", lty = "dashed")
abline(v = log(ridge$lambda.1se), col = "blue", lty = "dashed")

# plot lasso model
plot(lasso_min, xvar = "lambda", main = "Lasso penalty\n\n")
abline(v = log(lasso$lambda.min), col = "red", lty = "dashed")
abline(v = log(lasso$lambda.1se), col = "blue", lty = "dashed")
```

```{r}

lasso    <- glmnet(X, Y, alpha = 1.0) 
elastic1 <- glmnet(X, Y, alpha = 0.25) 
elastic2 <- glmnet(X, Y, alpha = 0.75) 
ridge    <- glmnet(X, Y, alpha = 0.0)

par(mfrow = c(2, 2))
plot(lasso, xvar = "lambda", main = "Lasso (Alpha = 1)\n\n\n")
plot(elastic1, xvar = "lambda", main = "Elastic Net (Alpha = .25)\n\n\n")
plot(elastic2, xvar = "lambda", main = "Elastic Net (Alpha = .75)\n\n\n")
plot(ridge, xvar = "lambda", main = "Ridge (Alpha = 0)\n\n\n")

```

Grid Search for auto-tune

```{r}
set.seed(123)

cv_glmnet <- train(
   x = X,
   y = Y,
   method = "glmnet",
   preProc = c("zv", "center", "scale"),
   trControl = trainControl(method = "cv", number = 10),
   tuneLength = 10
)

# model with lowest RMSE
cv_glmnet$bestTune

ggplot(cv_glmnet)
```

Evaluate performance:

```{r}
pred <- predict(cv_glmnet, X)

# compute RMSE of transformed predictors
RMSE(exp(pred), exp(Y))
```

### Feature Interpretation

Feature interpretation is roughly the same as in PSL.

```{r}
vip(cv_glmnet, num_features = 20, bar = F)
```

```{r}
p1 <- pdp::partial(cv_glmnet, pred.var = "Gr_Liv_Area", grid.resolution = 20) %>%
  mutate(yhat = exp(yhat)) %>%
  ggplot(aes(Gr_Liv_Area, yhat)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p2 <- pdp::partial(cv_glmnet, pred.var = "Overall_QualExcellent") %>%
  mutate(
    yhat = exp(yhat),
    Overall_QualExcellent = factor(Overall_QualExcellent)
    ) %>%
  ggplot(aes(Overall_QualExcellent, yhat)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p3 <- pdp::partial(cv_glmnet, pred.var = "First_Flr_SF", grid.resolution = 20) %>%
  mutate(yhat = exp(yhat)) %>%
  ggplot(aes(First_Flr_SF, yhat)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

p4 <- pdp::partial(cv_glmnet, pred.var = "Garage_Cars") %>%
  mutate(yhat = exp(yhat)) %>%
  ggplot(aes(Garage_Cars, yhat)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

```{r}
pdp::partial(cv_glmnet, pred.var = "Overall_QualPoor") %>%
  mutate(
    yhat = exp(yhat),
    Overall_QualPoor = factor(Overall_QualPoor)
    ) %>%
  ggplot(aes(Overall_QualPoor, yhat)) +
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 300000), labels = scales::dollar)
```

### Attrition

```{r}
set.seed(123)

churn_split <- initial_split(attrition, prop = .7, strata = "Attrition")
churn.train <- training(churn_split)
churn.test <- testing(churn_split)

glm_mod <- train(
   Attrition ~.,
   data = churn.train,
   method = "glm",
   family = "binomial",
   preProc = c("zv", "center", "scale"),
   trControl = trainControl(method = "cv", number = 10)
)

# train regularized logistic regression model

set.seed(123)

penalized_mod <- train(
   Attrition ~.,
   data = churn.train,
   method = "glmnet",
   family = "binomial",
   preProc = c("zv", "center", "scale"),
   trControl = trainControl(method = "cv", number = 10),
   tuneLength = 10
)

summary(resamples(list(
   logistic_model = glm_mod,
   penalized_model = penalized_mod
)))$statistics$Accuracy

```

```{r}
# clean up
rm(list = ls())
```

### 2020-03-01 Hands-on Machine Learning/11_Random-Forests.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Random Forests}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 11}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)
library(foreach, quietly = T, warn.conflicts = F)
library(doParallel, quietly = T, warn.conflicts = F)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(ranger, quietly = T, warn.conflicts = F)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

##  Random Forests

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition.h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")
churn.train <- training(churn)
churn.test <- testing(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames.h2o <- as.h2o(ames)

ames.split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames.train <- training(ames.split)
ames.test <- testing(ames.split)
```

### Random Forest Overview

Random forests are modifications of bagged decision trees that build a large collection of _de-correlated_ trees to further improve the predictive performance.

### Extended Bagging

The bootstrap aggregation procedure (bagging) has a limited effect on the variance reduction of decision trees.

Random forests help reduce tree correlation by injecting more randomness into the tree-growing process. More specifically, while growing a decision tree during the bagging process, random forests perform split-variable randomization where each time a split is to be peformed, the search for the split variable is limited to a random subset of $m_{try}$ of the original p features.

Typical default: $m_{try} = \frac{p}{3}$ (regression) and $m_{try} = \sqrt{p}$ for classification.

Basic algorithm is as follows::

+ 1.) Given a training data set

+ 2.) Select number of trees to build (n_trees)

+ 3.) for i = 1 to n_trees do:

+ 4.) Generate a bootstrap sample of the original data

+ 5.) Grow a regression/classification tree to the bootstrapped data

+ 6.) for each split

+ 7.) Select m_try variables at random for all p variables

+ 8.) Pick the best variable/split-point among the m_try

+ 9.) split the node into two child nodes

+ 10.) end

+ 11.) Use typical tree model stopping criteria to determine when a tree is complete (do not prune)

+ 12.) end

+ 13.) output ensemble of trees

Note: when $m_{try} = p$, the algorithm is equivalent to bagging decission trees.

### Out-of-the-box Performance

```{r}
# number of features
n.features <- length(setdiff(names(ames.train), "Sale_Price"))

# train a default random forest model
ames.rf1 <- ranger(
  Sale_Price ~ ., 
  data = ames.train,
  mtry = floor(n.features / 3),
  respect.unordered.factors = "order",
  seed = 123
)

# get OOB RMSE
(default.rmse <- sqrt(ames.rf1$prediction.error))
```

### Hyperparamteres

+ 1.) The number of trees in the forest

+ 2.) The number of features to consider at any given split

+ 3.) The complexity of each tree

+ 4.) The sampling scheme

+ 5.) The splitting rule to use during tree construction

#### Number of Trees

Typically start with 10x the number of features

```{r}
# number of features
n.features <- ncol(ames.train) - 1

# tuning grid
tuning.grid <- expand.grid(
  trees = seq(10, 1000, by = 20),
  rmse  = NA
)

for(i in seq_len(nrow(tuning.grid))) {

  # Fit a random forest
  fit <- ranger(
    formula = Sale_Price ~ ., 
    data = ames.train, 
    num.trees = tuning.grid$trees[i],
    mtry = floor(n.features / 3),
    respect.unordered.factors = 'order',
    verbose = FALSE,
    seed = 123
  )

  # Extract OOB RMSE
  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)
}

ggplot(tuning.grid, aes(trees, rmse)) +
  geom_line(size = 1) +
  ylab("OOB Error (RMSE)") +
  xlab("Number of trees")
```

#### $m_{try}$

Hyperparameter that control the split-variable randomization feature of random forests.

With a high number of predictive features a high value of $m_{try}$ is likely to perform better. When many features are relevant, lower might perform better.

```{r}
tuning.grid <- expand.grid(
  trees = seq(10, 1000, by = 20),
  mtry  = floor(c(seq(2, 80, length.out = 5), 26)),
  rmse  = NA
)

for(i in seq_len(nrow(tuning.grid))) {
  fit <- ranger(
    formula    = Sale_Price ~ ., 
    data       = ames.train, 
    num.trees  = tuning.grid$trees[i],
    mtry       = tuning.grid$mtry[i],
    respect.unordered.factors = 'order',
    verbose    = FALSE,
    seed       = 123
  )

  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)

}

labels <- tuning.grid %>%
  filter(trees == 990) %>%
  mutate(mtry = as.factor(mtry))

tuning.grid %>%
  mutate(mtry = as.factor(mtry)) %>%
  ggplot(aes(trees, rmse, color = mtry)) +
  geom_line(size = 1, show.legend = FALSE) +
  ggrepel::geom_text_repel(data = labels, aes(trees, rmse, label = mtry), nudge_x = 50, show.legend = FALSE) +
  ylab("OOB Error (RMSE)") +
  xlab("Number of trees")
```

#### Tree Complexity

Random forests are built on individual decision trees; consequently most random forest implementations have one or more hyperparameters.

```{r}

tuning.grid <- expand.grid(
  min.node.size = 1:20,
  run_time  = NA,
  rmse = NA
)

for(i in seq_len(nrow(tuning.grid))) {
  fit.time <- system.time({
    fit <- ranger(
      formula    = Sale_Price ~ ., 
      data       = ames.train, 
      num.trees  = 1000,
      mtry       = 26,
      min.node.size = tuning.grid$min.node.size[i],
      respect.unordered.factors = 'order',
      verbose    = FALSE,
      seed       = 123
  )
})

  tuning.grid$run_time[i] <- fit.time[[3]]
  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)
}

min.node.size <- tuning.grid %>% 
  mutate(
    error_first = first(rmse),
    runtime_first = first(run_time),
    `Error Growth` = (rmse / error_first) - 1,
    `Run Time Reduction` = (run_time / runtime_first) - 1
    )

p1 <-  ggplot(min.node.size, aes(min.node.size, `Error Growth`)) +
  geom_smooth(size = 1, se = FALSE, color = "black") +
  scale_y_continuous("Percent growth in error estimate", labels = scales::percent) +
  xlab("Minimum node size") +
  ggtitle("A) Impact to error estimate")

p2 <-  ggplot(min.node.size, aes(min.node.size, `Run Time Reduction`)) +
  geom_smooth(size = 1, se = FALSE, color = "black") +
  scale_y_continuous("Reduction in run time", labels = scales::percent) +
  xlab("Minimum node size") +
  ggtitle("B) Impact to run time")

gridExtra::grid.arrange(p1, p2, nrow = 1)

```

#### Sampling Scheme

```{r}
tuning.grid <- expand.grid(
  sample.fraction = seq(.05, .95, by = .05),
  replace  = c(TRUE, FALSE),
  rmse = NA
)

for(i in seq_len(nrow(tuning.grid))) {
  fit <- ranger(
    formula    = Sale_Price ~ ., 
    data       = ames.train, 
    num.trees  = 1000,
    mtry       = 26,
    sample.fraction = tuning.grid$sample.fraction[i],
    replace = tuning.grid$replace[i],
    respect.unordered.factors = 'order',
    verbose    = FALSE,
    seed       = 123
  )

  tuning.grid$rmse[i] <- sqrt(fit$prediction.error)
}

tuning.grid %>%
  ggplot(aes(sample.fraction, rmse, color = replace)) +
  geom_line(size = 1) +
  scale_x_continuous("Sample Fraction", breaks = seq(.1, .9, by = .1), labels = scales::percent) +
  ylab("OOB Error (RMSE)") +
  scale_color_discrete("Sample with Replacement") +
  theme(legend.position = c(0.8, 0.85),
        legend.key = element_blank(),
        legend.background = element_blank())
```

### Tuning Strategies

Simple Brute Force

```{r}
# create hyperparameter grid
hyper_grid <- expand.grid(
  mtry = floor(n.features * c(.05, .15, .25, .333, .4)),
  min.node.size = c(1, 3, 5, 10), 
  replace = c(TRUE, FALSE),                               
  sample.fraction = c(.5, .63, .8),                       
  rmse = NA                                               
)

# execute full cartesian grid search
for(i in seq_len(nrow(hyper_grid))) {
  # fit model for ith hyperparameter combination
  fit <- ranger(
    formula         = Sale_Price ~ ., 
    data            = ames.train, 
    num.trees       = n.features * 10,
    mtry            = hyper_grid$mtry[i],
    min.node.size   = hyper_grid$min.node.size[i],
    replace         = hyper_grid$replace[i],
    sample.fraction = hyper_grid$sample.fraction[i],
    verbose         = FALSE,
    seed            = 123,
    respect.unordered.factors = 'order',
  )
  # export OOB error 
  hyper_grid$rmse[i] <- sqrt(fit$prediction.error)
}

# assess top 10 models
hyper_grid %>%
  arrange(rmse) %>%
  mutate(perc_gain = (default.rmse - rmse) / default.rmse * 100) %>%
  head(10)
```

#### h20 Random Grid Search

```{r}
h2o.no_progress()
h2o.init(max_mem_size = "5g")
```

```{r}
# convert training data to h2o object
train_h2o <- as.h2o(ames.train)

# set the response column to Sale_Price
response <- "Sale_Price"

# set the predictor names
predictors <- setdiff(colnames(ames.train), response)
```

```{r}
h2o_rf1 <- h2o.randomForest(
    x = predictors, 
    y = response,
    training_frame = train_h2o, 
    ntrees = n.features * 10,
    seed = 123
)

h2o_rf1
```

```{r}
# hyperparameter grid
hyper_grid <- list(
  mtries = floor(n.features * c(.05, .15, .25, .333, .4)),
  min_rows = c(1, 3, 5, 10),
  max_depth = c(10, 20, 30),
  sample_rate = c(.55, .632, .70, .80)
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.001,   # stop if improvement is < 0.1%
  stopping_rounds = 10,         # over the last 10 models
  max_runtime_secs = 60*5      # or stop search after 5 min.
)
```

```{r}
# perform grid search 
random_grid <- h2o.grid(
  algorithm = "randomForest",
  grid_id = "rf_random_grid",
  x = predictors, 
  y = response, 
  training_frame = train_h2o,
  hyper_params = hyper_grid,
  ntrees = n.features * 10,
  seed = 123,
  stopping_metric = "RMSE",   
  stopping_rounds = 10,           # stop if last 10 trees added 
  stopping_tolerance = 0.005,     # don't improve RMSE by 0.5%
  search_criteria = search_criteria
)
```

```{r}
# collect the results and sort by our model performance metric 
# of choice
random_grid_perf <- h2o.getGrid(
  grid_id = "rf_random_grid", 
  sort_by = "mse", 
  decreasing = FALSE
)
random_grid_perf
```

```{r}
# re-run model with impurity-based variable importance
rf_impurity <- ranger(
  formula = Sale_Price ~ ., 
  data = ames.train, 
  num.trees = 2000,
  mtry = 32,
  min.node.size = 1,
  sample.fraction = .80,
  replace = FALSE,
  importance = "impurity",
  respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)

# re-run model with permutation-based variable importance
rf_permutation <- ranger(
  formula = Sale_Price ~ ., 
  data = ames.train, 
  num.trees = 2000,
  mtry = 32,
  min.node.size = 1,
  sample.fraction = .80,
  replace = FALSE,
  importance = "permutation",
  respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)
```

```{r, fig.height=7}
p1 <- vip::vip(rf_impurity, num_features = 25, bar = FALSE)
p2 <- vip::vip(rf_permutation, num_features = 25, bar = FALSE)

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-03-01 Hands-on Machine Learning/12_Gradient-Boosting.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Gradient Boosting}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 12}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)
library(foreach, quietly = T, warn.conflicts = F)
library(doParallel, quietly = T, warn.conflicts = F)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(earth, quietly = TRUE, warn.conflicts = FALSE)
library(ranger, quietly = T, warn.conflicts = F)
library(gbm, quietly = T, warn.conflicts = F)
library(xgboost, quietly = T, warn.conflicts = F)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Gradient Boosting

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

h2o.init(max_mem_size = "10g", strict_version_check = F)

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

### Gradient Boosting Overview

Whereas random forests build an ensemble of deep independent trees, GBMs build an ensemble of shallow trees in sequence with each tree learning and improving on the previous one. Although shallow trees by themselves are rather weak predictive models, they can be "boosted" to produce a powerful "committee"

### Boosting vs Bagging

Boosting is a general concept that aggregates the predictions of simpler models. It is typically used with models that have high bias and low variance. This works particularly well with decision trees.

Boosting is a sequential algorithm that attacks the bias-variance trade-off by starting with a _weak_ model, and sequentially boost its performance by continuiing to build new trees, where each new tree in the sequence tries to "fix up" where the previous one made the biggest error.

### Base Learners

Boosting is a framework that iteratively improves any weak learning model. Many gradient boosting applications allow you to "plug in" various classes of weak learners at your disposal.

In practice, these are almost always decision trees as base learners.

### Traning Weak Models

A weak model is one whose error rate is only slighly better than random guessing. The idea behind boosting is that eadch model in the sequence slighly improves upon the performance of the previous one.

### Sequential Training

Boosted trees are grown sequentially; each tree is grown using information from previously grown trees to improve performance.

### Basic GBM Algorithm

+ 1.) Fit a decision tree to the data: $F_1(x) = y$

+ 2.) We then fit the next decision tree to the residuals of the previous: $h_1(x) = y - F_1(x)$

+ 3.) Add this new tree to our algorithm: $F_2(x) = F_1(x) + h_1(x)$

+ 4.) Fit the next decision tree to the residuals of $F_2 : h_2(x) = y - F_2(x)$

+ 5.) Add this new tree to our algorithm: $F_3(x) = F_2(x) + h_1(x)$

+ 6.) Continue this process until some mechanism (i.e., cross validation) tells us to stop.

The final model here is a stagewise additive model of b individual trees:

$f(x) = \sum^{B}_{b=1}f^b(x)$

Visually, the process looks like this:

```{r, fig.height=7}
# Simulate sine wave data
set.seed(1112)  # for reproducibility
df <- tibble::tibble(
  x = seq(from = 0, to = 2 * pi, length = 1000),
  y = sin(x) + rnorm(length(x), sd = 0.5),
  truth = sin(x)
)

# Function to boost `rpart::rpart()` trees
rpartBoost <- function(x, y, data, num_trees = 100, learn_rate = 0.1, tree_depth = 6) {
  x <- data[[deparse(substitute(x))]]
  y <- data[[deparse(substitute(y))]]
  G_b_hat <- matrix(0, nrow = length(y), ncol = num_trees + 1)
  r <- y
  for(tree in seq_len(num_trees)) {
    g_b_tilde <- rpart(r ~ x, control = list(cp = 0, maxdepth = tree_depth))
    g_b_hat <- learn_rate * predict(g_b_tilde)
    G_b_hat[, tree + 1] <- G_b_hat[, tree] + matrix(g_b_hat)
    r <- r - g_b_hat
    colnames(G_b_hat) <- paste0("tree_", c(0, seq_len(num_trees)))
  }
  cbind(df, as.data.frame(G_b_hat)) %>%
    gather(tree, prediction, starts_with("tree")) %>%
    mutate(tree = stringr::str_extract(tree, "\\d+") %>% as.numeric())
}

# Plot boosted tree sequence
rpartBoost(x, y, data = df, num_trees = 2^10, learn_rate = 0.05, tree_depth = 1) %>%
  filter(tree %in% c(0, 2^c(0:10))) %>%
  ggplot(aes(x, prediction)) +
    ylab("y") +
    geom_point(data = df, aes(x, y), alpha = .1) +
    geom_line(data = df, aes(x, truth), color = "blue") +
    geom_line(colour = "red", size = 1) +
    facet_wrap(~ tree, nrow = 3)
```

The Gradient in GBM is a gradient descent algorithm (from Calculus), where we can use it to find the minima/maximua of a function. Here, we apply the gradient descent to the loss function in the algorithm.

```{r}
# create data to plot
x <- seq(-5, 5, by = .05)
y <- x^2 + 3
df <- data.frame(x, y)
step <- 5
step_size <- .2
for(i in seq_len(18)) {
  next_step <- max(step) + round(diff(range(max(step), which.min(df$y))) * step_size, 0)
  step <- c(step, next_step)
  next
}
steps <- df[step, ] %>%
  mutate(x2 = lag(x), y2 = lag(y)) %>%
  dplyr::slice(1:18)
# plot
ggplot(df, aes(x, y)) +
  geom_line(size = 1.5, alpha = .5) +
  theme_classic() +
  scale_y_continuous("Loss function", limits = c(0, 30)) +
  xlab(expression(theta)) +
  geom_segment(data = df[c(5, which.min(df$y)), ], aes(x = x, y = y, xend = x, yend = -Inf), lty = "dashed") +
  geom_point(data = filter(df, y == min(y)), aes(x, y), size = 4, shape = 21, fill = "yellow") +
  geom_point(data = steps, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) +
  geom_curve(data = steps, aes(x = x, y = y, xend = x2, yend = y2), curvature = 1, lty = "dotted") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  annotate("text", x = df[5, "x"], y = 1, label = "Initial value", hjust = -0.1, vjust = .8) +
  annotate("text", x = df[which.min(df$y), "x"], y = 1, label = "Minimium", hjust = -0.1, vjust = .8) +
  annotate("text", x = df[5, "x"], y = df[5, "y"], label = "Learning step", hjust = -.8, vjust = 0)
```

Gradient descent is the process of gradually decreasing the cost function (i.e, MSE) by tweaking parameter(s) iteratively until you have reached a minimuim.

Gradient descent can be performed on any loss function that is differentiable.

An important parameter is the step size with is controlled by the learning rate. If the learning rate is too small, then the algorithm will take many interations to find the minimum. If its too big, it can stip over a possible minimum value.

Visually:

```{r}
# create too small of a learning rate
step <- 5
step_size <- .05
for(i in seq_len(10)) {
  next_step <- max(step) + round(diff(range(max(step), which.min(df$y))) * step_size, 0)
  step <- c(step, next_step)
  next
}
too_small <- df[step, ] %>%
  mutate(x2 = lag(x), y2 = lag(y))
# plot
p1 <- ggplot(df, aes(x, y)) +
  geom_line(size = 1.5, alpha = .5) +
  theme_classic() +
  scale_y_continuous("Loss function", limits = c(0, 30)) +
  xlab(expression(theta)) +
  geom_segment(data = too_small[1, ], aes(x = x, y = y, xend = x, yend = -Inf), lty = "dashed") +
  geom_point(data = too_small, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) +
  geom_curve(data = too_small, aes(x = x, y = y, xend = x2, yend = y2), curvature = 1, lty = "dotted") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  annotate("text", x = df[5, "x"], y = 1, label = "Start", hjust = -0.1, vjust = .8) +
  ggtitle("b) too small")
# create too large of a learning rate
too_large <- df[round(which.min(df$y) * (1 + c(-.9, -.6, -.2, .3)), 0), ] %>%
  mutate(x2 = lag(x), y2 = lag(y))
# plot
p2 <- ggplot(df, aes(x, y)) +
  geom_line(size = 1.5, alpha = .5) +
  theme_classic() +
  scale_y_continuous("Loss function", limits = c(0, 30)) +
  xlab(expression(theta)) +
  geom_segment(data = too_large[1, ], aes(x = x, y = y, xend = x, yend = -Inf), lty = "dashed") +
  geom_point(data = too_large, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) +
  geom_curve(data = too_large, aes(x = x, y = y, xend = x2, yend = y2), curvature = 1, lty = "dotted") +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  annotate("text", x = too_large[1, "x"], y = 1, label = "Start", hjust = -0.1, vjust = .8) +
  ggtitle("a) too big")
gridExtra::grid.arrange(p2, p1, nrow = 1)
```

Not all loss functions are convex, however. There are many local minimas, plateaus, and other irregular terain of the loss function that makes finding the global minimum difficult.

**Stochastic gradient descent** can help us address this problem by sampling a fraction of the training observations (typically w/o replacement), and growing the next tree using that subsample.

Visually:

```{r}
# create random walk data
set.seed(123)
x <- sample(seq(3, 5, by = .05), 10, replace = TRUE)
set.seed(123)
y <- seq(2, 28, length.out = 10)

random_walk <- data.frame(
  x = x,
  y = y[order(y, decreasing = TRUE)]
)

optimal <- data.frame(x = 0, y = 0)

# plot
ggplot(df, aes(x, y)) + 
  coord_polar() +
  theme_minimal() +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank()
  ) +
  xlab(expression(theta[1])) +
  ylab(expression(theta[2])) +
  geom_point(data = random_walk, aes(x, y), size = 3, shape = 21, fill = "blue", alpha = .5) + 
  geom_point(data = optimal, aes(x, y), size = 2, shape = 21, fill = "yellow") + 
  geom_path(data = random_walk, aes(x, y), lty = "dotted") +
  annotate("text", x = random_walk[1, "x"], y = random_walk[1, "y"], label = "Start", hjust = 1, vjust = -1) +
  annotate("text", x = optimal[1, "x"], y = optimal[1, "y"], label = "Minimum", hjust = -.2, vjust = 1) +
  ylim(c(0, 28)) + 
  xlim(-5, 5)
```

Stochastic gradent descent will often find a near-optimal solution by jumping out of local minimas and off plateaus.

### Basic GBM

The first boosting algorithm was AdaBoost, which was then generalized into a regression framework.

### Hyperparameters

Typically there are two kinds of hyperparamters in GBM, boosting and tree-specific.

+ 1.) Number of Trees

Total number of trees in the ensemble.

+ 2.) Learning Rate

Determines the contribution of each tree on the final outcome and control how quickly the algorithm proceeds down the gradient descent.

Implementation:

```{r}
# run a basic GBM model
set.seed(123)  # for reproducibility
ames_gbm1 <- gbm(
  formula = Sale_Price ~ .,
  data = ames_train,
  distribution = "gaussian",  # SSE loss function
  n.trees = 5000,
  shrinkage = 0.1,
  interaction.depth = 3,
  n.minobsinnode = 10,
  cv.folds = 10
)

# find index for number trees with minimum CV error
best <- which.min(ames_gbm1$cv.error)

# get MSE and compute RMSE
sqrt(ames_gbm1$cv.error[best])
```

```{r}
# plot error curve
gbm.perf(ames_gbm1, method = "cv")
```

### General Tuning Strategy

```{r}
# create grid search
hyper_grid <- expand.grid(
  learning_rate = c(0.3, 0.1, 0.05, 0.01, 0.005),
  RMSE = NA,
  trees = NA,
  time = NA
)

# execute grid search
for(i in seq_len(nrow(hyper_grid))) {

  # fit gbm
  set.seed(123)  # for reproducibility
  train_time <- system.time({
    m <- gbm(
      formula = Sale_Price ~ .,
      data = ames_train,
      distribution = "gaussian",
      n.trees = 5000, 
      shrinkage = hyper_grid$learning_rate[i], 
      interaction.depth = 3, 
      n.minobsinnode = 10,
      cv.folds = 10 
   )
  })

  # add SSE, trees, and training time to results
  hyper_grid$RMSE[i]  <- sqrt(min(m$cv.error))
  hyper_grid$trees[i] <- which.min(m$cv.error)
  hyper_grid$Time[i]  <- train_time[["elapsed"]]

}

# results
arrange(hyper_grid, RMSE)
```

Exaustive Method

```{r}
# search grid
hyper_grid <- expand.grid(
  n.trees = 6000,
  shrinkage = 0.01,
  interaction.depth = c(3, 5, 7),
  n.minobsinnode = c(5, 10, 15)
)

# create model fit function
model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = Sale_Price ~ .,
    data = ames_train,
    distribution = "gaussian",
    n.trees = n.trees,
    shrinkage = shrinkage,
    interaction.depth = interaction.depth,
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))
}

# perform search grid with functional programming
hyper_grid$rmse <- purrr::pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
    )
)

# results
arrange(hyper_grid, rmse)
```

### Stochastic GBMs

```{r}
# refined hyperparameter grid
hyper_grid <- list(
  sample_rate = c(0.5, 0.75, 1),              # row subsampling
  col_sample_rate = c(0.5, 0.75, 1),          # col subsampling for each split
  col_sample_rate_per_tree = c(0.5, 0.75, 1)  # col subsampling for each tree
)

# random grid search strategy
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.001,   
  stopping_rounds = 10,         
  max_runtime_secs = 60*60      
)

# perform grid search 
grid <- h2o.grid(
  algorithm = "gbm",
  grid_id = "gbm_grid",
  x = predictors, 
  y = response,
  training_frame = train_h2o,
  hyper_params = hyper_grid,
  ntrees = 6000,
  learn_rate = 0.01,
  max_depth = 7,
  min_rows = 5,
  nfolds = 10,
  stopping_rounds = 10,
  stopping_tolerance = 0,
  search_criteria = search_criteria,
  seed = 123
)

# collect the results and sort by our model performance metric of choice
grid_perf <- h2o.getGrid(
  grid_id = "gbm_grid", 
  sort_by = "mse", 
  decreasing = FALSE
)

grid_perf
```

```{r}
# Grab the model_id for the top model, chosen by cross validation error
best_model_id <- grid_perf@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# Now let’s get performance metrics on the best model
h2o.performance(model = best_model, xval = TRUE)
```

### XGBoost

#### Regularization

**XGBoost** has multiple tuning parameters to help prevent overfitting.

$\gamma$ = Lagrangian multiplier which controls the complexity of a given tree.

$\alpha$ = $L_1$ regularization (simmilar to ridge regresssion)

$\lambda$ = $L_2$ regularization (similar to lasso)

#### Dropout

Loosly defined as a regularization technique:

```{r}
xgb_prep <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_integer(all_nominal()) %>%
  prep(training = ames_train, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Sale_Price")])
Y <- xgb_prep$Sale_Price
```

```{r}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "reg:linear",
  early_stopping_rounds = 50, 
  nfold = 10,
  params = list(
    eta = 0.1,
    max_depth = 3,
    min_child_weight = 3,
    subsample = 0.8,
    colsample_bytree = 1.0),
  verbose = 0
)

# minimum test CV RMSE
min(ames_xgb$evaluation_log$test_rmse_mean)
```

```{r}
# hyperparameter grid
hyper_grid <- expand.grid(
  eta = 0.01,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = 0.5, 
  colsample_bytree = 0.5,
  gamma = c(0, 1, 10, 100, 1000),
  lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  rmse = 0,          # a place to dump RMSE results
  trees = 0          # a place to dump required number of trees
)

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 4000,
    objective = "reg:linear",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( 
      eta = hyper_grid$eta[i], 
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i],
      colsample_bytree = hyper_grid$colsample_bytree[i],
      gamma = hyper_grid$gamma[i], 
      lambda = hyper_grid$lambda[i], 
      alpha = hyper_grid$alpha[i]
    ) 
  )
  hyper_grid$rmse[i] <- min(m$evaluation_log$test_rmse_mean)
  hyper_grid$trees[i] <- m$best_iteration
}

# results
hyper_grid %>%
  filter(rmse > 0) %>%
  arrange(rmse) %>%
  glimpse()
```

```{r}
# optimal parameter list
params <- list(
  eta = 0.01,
  max_depth = 3,
  min_child_weight = 3,
  subsample = 0.5,
  colsample_bytree = 0.5
)

# train final model
xgb.fit.final <- xgboost(
  params = params,
  data = X,
  label = Y,
  nrounds = 3944,
  objective = "reg:linear",
  verbose = 0
)
```

### Feature Interpretation

```{r}
# variable importance plot
vip::vip(xgb.fit.final) 
```

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-03-01 Hands-on Machine Learning/13_Deep-Learning.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Deep Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 13}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)
library(foreach, quietly = T, warn.conflicts = F)
library(doParallel, quietly = T, warn.conflicts = F)
library(compiler, quietly = T, warn.conflicts = F)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)

library(keras, quietly = TRUE, warn.conflicts = FALSE)
library(tfruns, quietly = TRUE, warn.conflicts = FALSE)
library(tfestimators, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Deep Learning

### Data Sets

MNIST

```{r, echo = T}
# Import MNIST training data
mnist <- dslabs::read_mnist()
mnist_x <- mnist$train$images
mnist_y <- mnist$train$labels

# Rename columns and standardize feature values
colnames(mnist_x) <- paste0("V", 1:ncol(mnist_x))
mnist_x <- mnist_x / 255

# One-hot encode response
mnist_y <- to_categorical(mnist_y, 10)

```

### Deep Learning Overview

Deep neural networks are designed to handle unique data sets. Multiple types exists RNN for text / speach, CNN image/video, etc.

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 128, input_shape = ncol(mnist_x)) %>%
  layer_dense(units = 64) %>%
  layer_dense(units = 10)
```

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 128, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 10, activation = "softmax")
```

```{r}
model <- keras_model_sequential() %>%

  # Network architecture
  layer_dense(units = 128, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 10, activation = "softmax") %>%

  # Backpropagation
  compile(
    loss = 'categorical_crossentropy',
    optimizer = optimizer_rmsprop(),
    metrics = c('accuracy')
  )
```

```{r}
# Train the model
fit1 <- model %>%
  fit(
    x = mnist_x,
    y = mnist_y,
    epochs = 25,
    batch_size = 128,
    validation_split = 0.2,
    verbose = FALSE
  )

# Display output
fit1
```

```{r}
plot(fit1)
```

### Model Tuning

```{r}
data_frame(
  Size = c("small", "medium", "large"), 
  `1` = c("16", "64", "256"),
  `2` = c("16, 8", "64, 32", "256, 128"),
  `3` = c("16, 8, 4", "64, 32, 16", "256, 128, 64")
) %>%
  kable(align = "c", caption = "Model capacities assessed represented as number of layers and nodes per layer.") %>%
  add_header_above(c(" ", "Hidden Layers" = 3)) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

### Batch Normalization

```{r}
model_w_norm <- keras_model_sequential() %>%

  # Network architecture with batch normalization
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%

  # Backpropagation
  compile(
    loss = "categorical_crossentropy",
    optimizer = optimizer_rmsprop(),
    metrics = c("accuracy")
  )
```

```{r}
# One layer models -------------------------------------------------------------
# small capacity model
`1 layer_small` <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# medium
`1 layer_medium` <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# large
`1 layer_large` <- keras_model_sequential() %>%
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# Two layer models -------------------------------------------------------------
# small capacity model
`2 layer_small` <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 8, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# medium
`2 layer_medium` <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# large
`2 layer_large` <- keras_model_sequential() %>%
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# Three layer models -------------------------------------------------------------
# small capacity model
`3 layer_small` <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 8, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 4, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# medium
`3 layer_medium` <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

# large
`3 layer_large` <- keras_model_sequential() %>%
  layer_dense(units = 256, activation = "relu", input_shape = ncol(mnist_x)) %>%
  layer_batch_normalization() %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_batch_normalization() %>%
  layer_dense(units = 10, activation = "softmax") %>%
  compiler() %>%
  trainer()

models <- ls(pattern = "layer_") 
df_batch <- models %>%
  map(get) %>%
  map(~ data.frame(
    `Validation error` = .$metrics$val_loss,
    `Training error`   = .$metrics$loss,
    epoch = seq_len(.$params$epoch)
    )) %>%
  map2_df(models, ~ mutate(.x, model = .y)) %>%
  separate(model, into = c("Middle layers", "Number of nodes"), sep = "_") %>%
  gather(Validation, Loss, Validation.error:Training.error) %>%
  mutate(
    Validation = str_replace_all(Validation, "\\.", " "),
    `Number of nodes` = factor(`Number of nodes`, levels = c("small", "medium", "large")),
    `Batch normalization` = TRUE
    )
```

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-03-01 Hands-on Machine Learning/14_Support-Vector-Machines.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Support Vector Machines}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 14}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)
library(svmpath, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(h2o, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(vip, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(kernlab, quietly = TRUE, warn.conflicts = FALSE)
library(mlbench, quietly = TRUE, warn.conflicts = FALSE)

# h2o Setup

h2o.no_progress()
h2o.init(strict_version_check = F)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Support Vector Machines

### Data Sets

Attrition

```{r, echo = T}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
attrition_h2o <- as.h2o(attrition)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, echo = T}
set.seed(123)

ames <- AmesHousing::make_ames()
ames_h2o <- as.h2o(ames)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

h2o.init(max_mem_size = "10g", strict_version_check = F)

train_h2o <- as.h2o(ames_train)
response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

```{r}
# Colors
dark2 <- RColorBrewer::brewer.pal(8, "Dark2")
set1 <- RColorBrewer::brewer.pal(9, "Set1")

# Plotting function; modified from svmpath::svmpath()
plot_svmpath <- function(x, step = max(x$Step), main = "") {

  # Extract model info
  object <- x
  f <- predict(object, lambda = object$lambda[step], type = "function")
  x <- object$x
  y <- object$y
  Elbow <- object$Elbow[[step]]
  alpha <- object$alpha[, step]
  alpha0 <- object$alpha0[step]
  lambda <- object$lambda[step]
  df <- as.data.frame(x[, 1L:2L])
  names(df) <- c("x1", "x2")
  df$y <- norm2d$y
  beta <- (alpha * y) %*% x

  # Construct plot
  ggplot(df, aes(x = x1, y = x2)) +
    geom_point(aes(shape = y, color = y), size = 3, alpha = 0.75) +
    xlab("Income (standardized)") +
    ylab("Lot size (standardized)") +
    xlim(-6, 6) +
    ylim(-6, 6) +
    coord_fixed() +
    theme(legend.position = "none") +
    theme_bw() +
    scale_shape_discrete(
      name = "Owns a riding\nmower?",
      breaks = c(1, 2),
      labels = c("Yes", "No")
    ) +
    scale_color_brewer(
      name = "Owns a riding\nmower?",
      palette = "Dark2",
      breaks = c(1, 2),
      labels = c("Yes", "No")
    ) +
    geom_abline(intercept = -alpha0/beta[2], slope = -beta[1]/beta[2], 
                color = "black") +
    geom_abline(intercept = lambda/beta[2] - alpha0/beta[2], 
                slope = -beta[1]/beta[2], 
                color = "black", linetype = 2) +
    geom_abline(intercept = -lambda/beta[2] - alpha0/beta[2], 
                slope = -beta[1]/beta[2], 
                color = "black", linetype = 2) +
    geom_point(data = df[Elbow, ], size = 3) +
    ggtitle(main)

}
```

### Support Vector Machines Overview

Support vector machines offer a direct approach to binary classification: try to find a hyperplane in some feature space that "best" separates the two classes.

### Optimal Separating Hyperplanes

```{r}
# Construct data for plotting
x1 <- x2 <- seq(from = 0, to = 1, length = 100)
xgrid <- expand.grid(x1 = x1, x2 = x2)
y1 <- 1 + 2 * x1
y2 <- 1 + 2 * xgrid$x1 + 3 * xgrid$x2

# Hyperplane: p = 2
p1 <- lattice::xyplot(
  x = y1 ~ x1, 
  type = "l", 
  col = "black", 
  xlab = expression(X[1]), 
  ylab = expression(X[2]),
  main = expression({f(X)==1+2*X[1]-X[2]}==0),
  scales = list(tck = c(1, 0))
)

# Hyperplane: p = 3
p2 <- lattice::wireframe(
  x = y2 ~ xgrid$x1 * xgrid$x2, 
  xlab = expression(X[1]), 
  ylab = expression(X[2]),
  zlab = expression(X[3]),
  main = expression({f(X)==1+2*X[1]+3*X[2]-X[3]}==0),
  drape = TRUE,
  colorkey = FALSE,
  col = dark2[1],
  scales = list(arrows = FALSE)
  # par.settings = list(axis.line = list(col = "transparent"))
)

# Display plots side by side
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

### Hard Margin Classifier

```{r}
# Simulate data
set.seed(805)
norm2d <- as.data.frame(mlbench::mlbench.2dnormals(
  n = 100,
  cl = 2,
  r = 4,
  sd = 1
))
names(norm2d) <- c("x1", "x2", "y")  # rename columns

# Scatterplot
p1 <- ggplot(norm2d, aes(x = x1, y = x2)) +
  geom_point(aes(shape = y, color = y), size = 3, alpha = 0.75) +
  xlab("Income (standardized)") +
  ylab("Lot size (standardized)") +
  xlim(-6, 6) +
  ylim(-6, 6) +
  coord_fixed() +
  theme(legend.position = "none") +
  theme_bw() +
  scale_shape_discrete(
    name = "Owns a riding\nmower?",
    breaks = c(1, 2),
    labels = c("Yes", "No")
  ) +
  scale_color_brewer(
    name = "Owns a riding\nmower?",
    palette = "Dark2",
    breaks = c(1, 2),
    labels = c("Yes", "No")
  )

# Fit a Logistic regression, linear discriminant analysis (LDA), and optimal
# separating hyperplane (OSH). Note: we sometimes refer to the OSH as the hard 
# margin classifier
fit_glm <- glm(as.factor(y) ~ ., data = norm2d, family = binomial)
fit_lda <- MASS::lda(as.factor(y) ~ ., data = norm2d)
invisible(capture.output(fit_hmc <- ksvm(  # use ksvm() to find the OSH
  x = data.matrix(norm2d[c("x1", "x2")]),
  y = as.factor(norm2d$y), 
  kernel = "vanilladot",  # no fancy kernel, just ordinary dot product
  C = Inf,                # to approximate hard margin classifier
  prob.model = TRUE       # needed to obtain predicted probabilities
)))

# Grid over which to evaluate decision boundaries
npts <- 500
xgrid <- expand.grid(
  x1 = seq(from = -6, 6, length = npts),
  x2 = seq(from = -6, 6, length = npts)
)

# Predicted probabilities (as a two-column matrix)
prob_glm <- predict(fit_glm, newdata = xgrid, type = "response")
prob_glm <- cbind("1" = 1 - prob_glm, "2" = prob_glm)
prob_lda <- predict(fit_lda, newdata = xgrid)$posterior
prob_hmc <- predict(fit_hmc, newdata = xgrid, type = "probabilities")

# Add predicted class probabilities
xgrid2 <- xgrid %>%
  cbind("GLM" = prob_glm[, 1L], 
        "LDA" = prob_lda[, 1L], 
        "HMC" = prob_hmc[, 1L]) %>%
  tidyr::gather(Model, Prob, -x1, -x2)

# Scatterplot with decision boundaries
p2 <- p1 + 
  stat_contour(data = xgrid2, aes(x = x1, y = x2, z = Prob, linetype = Model), 
               breaks = 0.5, color = "black")

# Display plots side by side
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Outliers:

```{r}
# Compute convex hull for each class
hpts1 <- chull(norm2d[norm2d$y == 1, c("x1", "x2")])
hpts1 <- c(hpts1, hpts1[1L])
hpts2 <- chull(norm2d[norm2d$y == 2, c("x1", "x2")])
hpts2 <- c(hpts2, hpts2[1L])

# Support vectors
sv <- norm2d[fit_hmc@alphaindex[[1L]], c("x1", "x2")]  # 16-th and 97-th observations

# Compute the perpendicular bisector of the line segment joining the two support 
# vectors
slope <- -1 / ((sv[2L, 2L] - sv[1L, 2L]) / (sv[2L, 1L] - sv[1L, 1L]))
midpoint <- apply(sv, 2, mean)

# Scatterplot with convex hulls, etc.
ggplot(norm2d, aes(x = x1, y = x2)) +

  # Convex hulls
  geom_polygon(
    data = norm2d[norm2d$y == 1, c("x1", "x2")][hpts1, c("x1", "x2")],
    color = "black",
    fill = "transparent"
  ) +
  geom_polygon(
    data = norm2d[norm2d$y == 2, c("x1", "x2")][hpts2, c("x1", "x2")],
    color = "black",
    fill = "transparent"
  ) +

  # Scatterplot
  geom_point(aes(shape = y, color = y), size = 3, alpha = 0.75) +
  xlab("Income (standardized)") +
  ylab("Lot size (standardized)") +
  xlim(-10, 10) +
  ylim(-10, 10) +
  # coord_fixed() +
  theme(legend.position = "none") +
  theme_bw() +
  scale_shape_discrete(
    name = "Owns a riding\nmower?",
    breaks = c(1, 2),
    labels = c("Yes", "No")
  ) +
  scale_color_brewer(
    name = "Owns a riding\nmower?",
    palette = "Dark2",
    breaks = c(1, 2),
    labels = c("Yes", "No")
  ) +

  # Decision boundary
  geom_abline(
    intercept = -slope * midpoint[1L] + midpoint[2L], 
    slope = slope
  ) +

  # Margin boundaries (shaded in)
  geom_abline(
    intercept = -slope * sv[1L, 1L] + sv[1L, 2L], 
    slope = slope,
    linetype = 2
  ) +
  geom_abline(
    intercept = -slope * sv[2L, 1L] + sv[2L, 2L], 
    slope = slope,
    linetype = 2
  ) +
  annotate(
    geom = "polygon", 
    x = c(-7, -7, 7, 7), 
    y = c(-slope * sv[1L, 1L] + sv[1L, 2L] - 7 * slope, 
          -slope * midpoint[1L] + midpoint[2L] - 7 * slope, 
          -slope * midpoint[1L] + midpoint[2L] + 7 * slope,
          -slope * sv[1L, 1L] + sv[1L, 2L] + 7 * slope),
    alpha = 0.1, 
    color = "transparent",
    fill = dark2[2]
  ) +
  annotate(
    geom = "polygon", 
    x = c(-7, -7, 7, 7), 
    y = c(-slope * sv[2L, 1L] + sv[2L, 2L] - 7 * slope,
          -slope * midpoint[1L] + midpoint[2L] - 7 * slope,
          -slope * midpoint[1L] + midpoint[2L] + 7 * slope,
          -slope * sv[2L, 1L] + sv[2L, 2L] + 7 * slope), 
    alpha = 0.1, 
    color = "transparent",
    fill = dark2[2]
  ) +

  # Arrows, labels, etc.
  annotate("segment",
    x = sv[1L, 1L], y = sv[1L, 2L], xend = sv[2L, 1L], yend = sv[2L, 2L], 
    # alpha = 0.5,
    linetype = 3
    # arrow = arrow(length = unit(0.03, units = "npc"), ends = "both")
  ) +
  geom_curve(x = -3, y = 4.5, xend = 0, yend = 5, 
             arrow = arrow(length = unit(0.03, units = "npc"))) +
  annotate("text", label = "Width = M", x = 0.45, y = 5.45, size = 5) +
  geom_curve(x = 2, y = -3, xend = 0, yend = -5, 
             arrow = arrow(length = unit(0.03, units = "npc"))) +
  annotate("text", label = "Width = M", x = 0, y = -5.35, size = 5) +

  # Support vectors
  annotate("point", x = sv$x1[1], y = sv$x2[1], shape = 17, color = "red", 
           size = 3) +
  annotate("point", x = sv$x1[2], y = sv$x2[2], shape = 16, color = "red", 
           size = 3) +
  # geom_point(data = cbind(sv, y = c("2", "1")), aes(shape = y),
  #            size = 4, color = "red") +

  # Zoom in
  coord_fixed(xlim = c(-6, 6), ylim = c(-6, 6))
```

Soft Margin Classifier:

```{r}
# Add an outlier
norm2d <- rbind(norm2d, data.frame("x1" = 0.5, "x2" = 1, "y" = 2))

# Fit a Logistic regression, linear discriminant analysis (LDA), and optimal
# separating hyperplane (OSH)
#
# Note: we sometimes refer to the OSH as the hard margin classifier
fit_glm <- glm(as.factor(y) ~ ., data = norm2d, family = binomial)
fit_lda <- MASS::lda(as.factor(y) ~ ., data = norm2d)
invisible(capture.output(fit_hmc <- ksvm(  # use ksvm() to find the OSH
  x = data.matrix(norm2d[c("x1", "x2")]),
  y = as.factor(norm2d$y), 
  kernel = "vanilladot",  # no fancy kernel, just ordinary dot product
  C = Inf,                # to approximate maximal margin classifier
  prob.model = TRUE       # needed to obtain predicted probabilities
)))

# Grid over which to evaluate decision boundaries
npts <- 500
xgrid <- expand.grid(
  x1 = seq(from = -6, 6, length = npts),
  x2 = seq(from = -6, 6, length = npts)
)

# Predicted probabilities (as a two-column matrix)
prob_glm <- predict(fit_glm, newdata = xgrid, type = "response")
prob_glm <- cbind("1" = 1 - prob_glm, "2" = prob_glm)
prob_lda <- predict(fit_lda, newdata = xgrid)$posterior
prob_hmc <- predict(fit_hmc, newdata = xgrid, type = "probabilities")

# Add predicted class probabilities
xgrid2 <- xgrid %>%
  cbind("GLM" = prob_glm[, 1L], 
        "LDA" = prob_lda[, 1L], 
        "HMC" = prob_hmc[, 1L]) %>%
  tidyr::gather(Model, Prob, -x1, -x2)

# Scatterplot
ggplot(norm2d, aes(x = x1, y = x2)) +

  # Label outlier
  geom_curve(x = tail(norm2d, n = 1)$x1 - 0.2, y = tail(norm2d, n = 1)$x2 - 0.2, 
             xend = -4, yend = 3, curvature = -0.5, angle = 90,
             arrow = arrow(length = unit(0.03, units = "npc"))) +
  annotate("text", label = "Outlier?", x = -4, y = 3.5, size = 5) +

  # Scatterplot, etc.
  geom_point(aes(shape = y, color = y), size = 3, alpha = 0.75) +
  xlab("Income (standardized)") +
  ylab("Lot size (standardized)") +
  xlim(-6, 6) +
  ylim(-6, 6) +
  coord_fixed() +
  theme(legend.position = "none") +
  theme_bw() +
  scale_shape_discrete(
    name = "Owns a riding\nmower?",
    breaks = c(1, 2),
    labels = c("Yes", "No")
  ) +
  scale_color_brewer(
    name = "Owns a riding\nmower?",
    palette = "Dark2",
    breaks = c(1, 2),
    labels = c("Yes", "No")
  ) +
  stat_contour(data = xgrid2, aes(x = x1, y = x2, z = Prob, linetype = Model), 
               breaks = 0.5, color = "black")
```

```{r}

# Fit the entire regularization path
fit_smc <- svmpath(
  x = data.matrix(norm2d[c("x1", "x2")]), 
  y = ifelse(norm2d$y == 1, 1, -1)
)

# Plot both extremes
p1 <- plot_svmpath(fit_smc, step = max(fit_smc$Step), main = expression(C == 0))
p2 <- plot_svmpath(fit_smc, step = min(fit_smc$Step), main = expression(C == infinity))
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

### Support Vector Machine

```{r}
# Load required packages
library(grid)
library(lattice)

# Simulate data
set.seed(1432)
circle <- as.data.frame(mlbench::mlbench.circle(
  n = 200,
  d = 2
))
names(circle) <- c("x1", "x2", "y")  # rename columns

# Fit a support vector machine (SVM)
fit_svm_poly <- ksvm( 
  x = data.matrix(circle[c("x1", "x2")]),
  y = as.factor(circle$y), 
  kernel = "polydot",       # polynomial kernel
  kpar = list(degree = 2),  # kernel parameters
  C = Inf,                  # to approximate maximal margin classifier
  prob.model = TRUE         # needed to obtain predicted probabilities
)

# Grid over which to evaluate decision boundaries
npts <- 500
xgrid <- expand.grid(
  x1 = seq(from = -1.25, 1.25, length = npts),
  x2 = seq(from = -1.25, 1.25, length = npts)
)

# Predicted probabilities (as a two-column matrix)
prob_svm_poly <- predict(fit_svm_poly, newdata = xgrid, type = "probabilities")

# Scatterplot
p1 <- contourplot(
  x = prob_svm_poly[, 1] ~ x1 * x2, 
  data = xgrid, 
  at = 0, 
  labels = FALSE,
  scales = list(tck = c(1, 0)),
  xlab = "x1",
  ylab = "x2",
  main = "Original feature space",
  panel = function(x, y, z, ...) {
    panel.contourplot(x, y, z, ...)
    panel.xyplot(
      x = circle$x1, 
      y = circle$x2, 
      groups = circle$y, 
      pch = 19, 
      cex = 1,
      col = adjustcolor(dark2[1L:2L], alpha.f = 0.5),
      ...
    )
  }
)

# Enlarge feature space
circle_3d <- circle
circle_3d$x3 <- circle_3d$x1^2 + circle_3d$x2^2

# 3-D scatterplot
p2 <- cloud(
  x = x3 ~ x1 * x2, 
  data = circle_3d, 
  groups = y,
  main = "Enlarged feature space",
  par.settings = list(
    superpose.symbol = list(
      pch = 19,
      cex = 1,
      col = adjustcolor(dark2[1L:2L], alpha.f = 0.5)
    )
  )
)

# p2 <- scatterplot3d(
#   x = circle_3d[, -3],
#   pch = 19,
#   color = adjustcolor(dark2[1L:2L], alpha.f = 0.5)[circle_3d$y]
# )
# p2$plane3d(0.64, 0, 0, draw_polygon = TRUE)
# p2 <- recordPlot()

# Scatterplot with decision boundary
p3 <- contourplot(
  x = prob_svm_poly[, 1] ~ x1 * x2, 
  data = xgrid, 
  at = 0.5, 
  labels = FALSE,
  scales = list(tck = c(1, 0)),
  xlab = "x1",
  ylab = "x2",
  main = "Non-linear decision boundary",
  panel = function(x, y, z, ...) {
    panel.contourplot(x, y, z, ...)
    panel.xyplot(
      x = circle$x1, 
      y = circle$x2, 
      groups = circle$y, 
      pch = 19, 
      cex = 1,
      col = adjustcolor(dark2[1L:2L], alpha.f = 0.5),
      ...
    )
  }
)

# Combine plots
gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
```

```{r}
# Load required packages
library(kernlab)  # for fitting SVMs
library(mlbench)  # for ML benchmark data sets

# Simulate train and test sets
set.seed(0841)
spirals <- as.data.frame(
  mlbench.spirals(300, cycles = 2, sd = 0.09)
)
names(spirals) <- c("x1", "x2", "classes")

# Fit an RF
set.seed(7256)
spirals_rfo <- ranger::ranger(classes ~ ., data = spirals, probability = TRUE)

# Fit an SVM using a radial basis function kernel
spirals_svm <- ksvm(classes ~ x1 + x2, data = spirals, kernel = "rbfdot",
                    C = 500, prob.model = TRUE)

# Grid over which to evaluate decision boundaries
npts <- 500
xgrid <- expand.grid(
  x1 = seq(from = -2, 2, length = npts),
  x2 = seq(from = -2, 2, length = npts)
)

# Predicted probabilities (as a two-column matrix)
prob_rfo <- predict(spirals_rfo, data = xgrid)$predictions
prob_svm <- predict(spirals_svm, newdata = xgrid, type = "probabilities")

# Add predicted class probabilities
xgrid2 <- xgrid %>%
  cbind("RF" = prob_rfo[, 1L], 
        "SVM" = prob_svm[, 1L]) %>%
  tidyr::gather(Model, Prob, -x1, -x2)

# Scatterplots with decision boundaries
ggplot(spirals, aes(x = x1, y = x2)) +
  geom_point(aes(shape = classes, color = classes), size = 3, alpha = 0.75) +
  xlab(expression(X[1])) +
  ylab(expression(X[2])) +
  xlim(-2, 2) +
  ylim(-2, 2) +
  coord_fixed() +
  theme(legend.position = "none") +
  theme_bw() +
  stat_contour(data = xgrid2, aes(x = x1, y = x2, z = Prob), 
               breaks = 0.5, color = "black") +
  facet_wrap( ~ Model)
```

```{r}
# Linear (i.e., soft margin classifier)
caret::getModelInfo("svmLinear")$svmLinear$parameters
```

```{r}
# Polynomial kernel
caret::getModelInfo("svmPoly")$svmPoly$parameters
```

```{r}
# Radial basis kernel
caret::getModelInfo("svmRadial")$svmRadial$parameters
```

```{r}
ggplot() +
  geom_abline(intercept = 4, slope = 1, linetype = 2, color = dark2[1L]) +
  geom_abline(intercept = 3, slope = 1) +
  geom_abline(intercept = 2, slope = 1, linetype = 2, color = dark2[1L]) +
  xlim(0, 5) +
  ylim(1, 10) +
  xlab(expression(x)) +
  ylab(expression(f(x))) +
  theme_bw() +
  annotate("text", label = "f(x) + epsilon", parse = TRUE, x = 2, y = 6.75, 
           size = 6, color = dark2[1L]) +
  annotate("text", label = "f(x) - epsilon", parse = TRUE, x = 2, y = 3.15, 
           size = 6, color = dark2[1L])
```

```{r}
# Simulate data
set.seed(1218)
x <- seq(from = -20, to = 20, by = 0.1)
y <- sin(x) / x + rnorm(length(x), sd = 0.03)
df <- na.omit(data.frame(x = x, y = y))

# Plot results
ggplot(df, aes(x = x, y = y)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = x, y = sin(x) / x), size = 1, color = "darkorange") +
  theme_bw() +
  theme(legend.position = "none") 
```

```{r}
# SVR model
set.seed(101)
svr <- kernlab::ksvm(y ~ x, data = df, kernel = "rbfdot", kpar = "automatic",
                     type = "eps-svr", epsilon = 0.1)

# MARS model
mars <- earth::earth(y ~ x, data = df)

# Random forest
set.seed(102)
rfo <- ranger::ranger(y ~ x, data = df)

# Gather predictions
df$SVR <- predict(svr, newdata = df)
df$MARS <- predict(mars, newdata = df)[, 1L, drop = TRUE]
df$RF <- predict(rfo, data = df)$predictions
df <- df %>% tidyr::gather(Model, Prediction, -x, -y)

# Plot results
ggplot(df, aes(x = x, y = y)) +
  geom_point(alpha = 0.5) +
  geom_line(aes(x = x, y = Prediction, color = Model), size = 1) +
  facet_wrap( ~ Model) +
  theme_bw() +
  theme(legend.position = "none") 
```

Job Attrition Example:

```{r}
# Tune an SVM with radial basis kernel
set.seed(1854)  # for reproducibility
churn_svm <- train(
  Attrition ~ ., 
  data = churn_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)
```

```{r}
# Plot results
ggplot(churn_svm) + theme_light()
```

```{r}
# Print results
churn_svm$results
```

Class Probabilities:

```{r}
# Control params for SVM
ctrl <- trainControl(
  method = "cv", 
  number = 10, 
  classProbs = TRUE,                 
  summaryFunction = twoClassSummary  # also needed for AUC/ROC
)

# Tune an SVM
set.seed(5628)  # for reproducibility
churn_svm_auc <- train(
  Attrition ~ ., 
  data = churn_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  metric = "ROC",  # area under ROC curve (AUC)       
  trControl = ctrl,
  tuneLength = 10
)

# Print results
churn_svm_auc$results
```

```{r}
confusionMatrix(churn_svm_auc)
```

```{r}
prob_yes <- function(object, newdata) {
  predict(object, newdata = newdata, type = "prob")[, "Yes"]
}
```

```{r}
# Variable importance plot
set.seed(2827)  # for reproducibility
vip(churn_svm_auc, method = "permute", nsim = 5, train = churn_train, 
    target = "Attrition", metric = "auc", reference_class = "Yes", 
    pred_wrapper = prob_yes)
```

```{r}
features <- c("OverTime", "WorkLifeBalance", 
              "JobSatisfaction", "JobRole")
pdps <- lapply(features, function(x) {
  pdp::partial(churn_svm_auc, pred.var = x, which.class = 2,  
          prob = TRUE, plot = TRUE, plot.engine = "ggplot2") +
    coord_flip()
})
grid.arrange(grobs = pdps,  ncol = 2)
```

```{r}
h2o.shutdown(prompt = FALSE)
```

```{r}
# clean up
rm(list = ls())
```

### 2020-03-01 Statistical Learning/03_Linear-Regression.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 3}
   \fancyfoot[C]{\rmfamily\color{headergrey}Linear Regression}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: 
    fig_caption: yes
    fig_width: 9
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select
```

## Chapter 3

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### R Lab

```{r, echo = T, fig.width=8, fig.height=3.5}
boston <- Boston
head(boston)

names(boston)
```

Simple Linear Regression

```{r, echo = T}
summary(lm.fit <- lm(medv ~ lstat, data = boston))

coef(lm.fit)
```

```{r, echo = T}
confint(lm.fit)

predict(lm.fit, data.frame(lstat = c(5, 10, 15)), 
        interval = "confidence")

predict(lm.fit, data.frame(lstat = c(5, 10, 15)), 
        interval = "prediction")
```

```{r, echo = T, fig.width=9, fig.height=3.5, fig.cap=""}
ggplot(boston, aes(lstat, medv)) +
   geom_point() +
   geom_smooth(method = "lm")
```

```{r, echo = T, fig.width=9, fig.height=3.5, fig.cap=""}
with(boston, {
   plot(lstat, medv)
   abline(lm.fit, col ="darkred")
   abline(lm.fit, lwd = 3, col = "darkred")
})
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
par(mfrow = c(2,2))
with(boston, {
   plot(lm.fit)
})
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
par(mfrow = c(1,2))
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
par(mfrow = c(1,1))
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

### Multiple Liinear Regression

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
summary(lm.fit <- lm(medv ~ lstat + age, data = boston))
summary(lm.fit <- lm(medv ~ ., data = boston))
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
vif(lm.fit)
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
summary(lm.fit1 <- lm(medv ~ .-age, data = boston))
```

### Interaction Terms

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
summary(lm(medv ~ lstat*age, data = boston))
```

### Non-linear Transformations of the Predictors

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
summary(lm.fit2 <- lm(medv ~ lstat + I(lstat^2), data = boston))
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
lm.fit <- lm(medv ~ lstat, data = boston)
anova(lm.fit, lm.fit2)
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
par(mfrow = c(2,2))
with(boston, {
   plot(lm.fit2)
})
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
summary(lm.fit5 <- lm(medv ~ poly(lstat, 5), data = boston))
```

```{r, echo = T, fig.width=9, fig.height=4, fig.cap=""}
summary(lm.fit5 <- lm(medv ~ log(rm), data = boston))
```

### Qualitative Predictors

```{r, echo = T}
carseats <- Carseats
summary(carseats)
```

```{r, echo = T}
summary(lm.fit <- lm(Sales ~ . + Income:Advertising+Price:Age, data = carseats))

contrasts(carseats$ShelveLoc)
```

### Conceptual

#### 1.)

Describe the null hypothesis to which the p-values given in Table 3.4 correspond. Explain what conclusions you can draw based on these p-values.

a.)

$H_0 : TV = 0, H_a : TV != 0$

*With a p-value of less than 0.0001, we reject the null hypothesis that the TV advertising budget does not effect sales.*

b.)

$H_0 : radio = 0, H_a : radio != 0$

*With a p-value of less than 0.0001, we reject the null hypothesis that the radio advertising budget does not effect sales.*

c.)

$H_0 : newspaper = 0, H_a : newspaper != 0$

*With a p-value of .8599, we fail to reject the null hypothesis that the newspaper advertising budget does not effect sales.*

#### 2.)

Carefully explain the differences between the KNN classifier and KNN regression methods.

*KNN regression uses the same basic technique as the classifier, which is to take a specified number of neighbors (based on some distance measure, _d_) and average them together to generate a value for the response. The difference here is that the regression returns a continious response variable, and the classificer results in a discrete metric that is based on the probability generated from the average of the neighbors.*

#### 3.)

Suppose we have a data set with five predictors, $X_1$ = GPA, $X_2$ = IQ, $X_3$ = Gender (1 Female/0 Male), $X_4$ Interaction Between GPA and IQ, $X_5$ = Interaction between GPA and Gender.

The response is starting salary after graduation (in thousands of dollars).

Suppose we use least squares to fit the model, and get $\hat{\beta}_0 = 50, \hat{\beta}_1 = 20, \hat{\beta}_2 = 0.07, \hat{\beta}_3 = 35, \hat{\beta}_4 = 0.01, \hat{\beta}_5 = -10$.

a.)

i. For a fixed value of IQ and GPA, males earn more on average than females.
ii. For a fixed value of IQ and GPA, females earn more on average than males.
iii. For a fixed value of IQ and GPA, males earn more on average than females provided that the GPA is high enough.
iv. For a fixed value of IQ and GPA, females earn more on average than males provided that the GPA is high enough.

*The least square line is given by
\[\hat{y} = 50 + 20GPA + 0.07IQ + 35Gender + 0.01GPA\times IQ - 10GPA\times Gender\]
which becomes for the males
\[\hat{y} = 50 + 20GPA + 0.07IQ + 0.01GPA\times IQ,\]
and for the females
\[\hat{y} = 85 + 10GPA + 0.07IQ + 0.01GPA\times IQ.\]
So the starting salary for males is higher than for females on average iff $50 + 20GPA\ge 85 + 10GPA$ which is equivalent to $GPA\ge 3.5$. Therefore iii. is the right answer.*

(b) Predict the salary of a female with IQ of 110 and a GPA of 4.0.

*It suffices to plug in the given values in the least square line for females given above and we obtain
\[\hat{y} = 85 + 40 + 7.7 + 4.4 = 137.1,\]
which gives us a starting salary of $137100\$$.*

(c) True or false: Since the coefficient for the GPA/IQ interaction term is very small, there is very little evidence of an interaction effect. Justify your answer.

*False. To verify if the GPA/IQ has an impact on the quality of the model we need to test the hypothesis $H_0 : \hat{\beta_4} = 0$ and look at the p-value associated with the $t$ or the $F$ statistic to draw a conclusion.*

#### 4.)

I collect a set of data ($n = 100$ observations) containing a single predictor and a quantitative response. I then fit a linear regression model to the data, as well as a separate cubic regression, i.e. $Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \beta_3 X^3 + \varepsilon$ .

(a) Suppose that the true relationship between $X$ and $Y$ is linear, i.e. $Y = \beta_0 + \beta_1 X + \varepsilon$. Consider the training residual sum of squares (RSS) for the linear regression, and also the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.

*Without knowing more details about the training data, it is difficult to know which training RSS is lower between linear or cubic. However, as the true relationship between $X$ and $Y$ is linear, we may expect the least squares line to be close to the true regression line, and consequently the RSS for the linear regression may be lower than for the cubic regression.*

(b) Answer (a) using test rather than training RSS.

*In this case the test RSS depends upon the test data, so we have not enough information to conclude. However, we may assume that polynomial regression will have a higher test RSS as the overfit from training would have more error than the linear regression.*

(c) Suppose that the true relationship between X and Y is not linear, but we don’t know how far it is from linear. Consider the training RSS for the linear regression, and also the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.

*Polynomial regression has lower train RSS than the linear fit because of higher flexibility: no matter what the underlying true relationship is the more flexible model will closer follow points and reduce train RSS. An example of this beahvior is shown on Figure 2.9 from Chapter 2.*

(d) Answer (c) using test rather than training RSS.

*There is not enough information to tell which test RSS would be lower for either regression given the problem statement is defined as not knowing "how far it is from linear". If it is closer to linear than cubic, the linear regression test RSS could be lower than the cubic regression test RSS. Or, if it is closer to cubic than linear, the cubic regression test RSS could be lower than the linear regression test RSS. It is dues to bias-variance tradeoff: it is not clear what level of flexibility will fit data better.*

#### 5.)

Consider the fitted values that result from performing linear regression without an intercept. In this setting, the i-th fitted value takes the form $\hat{y_i} = x_i\hat{\beta}$, where
\[\hat{\beta} = \frac{\sum_{i=1}^n x_iy_i}{\sum_{k=1}^nx_k^2}.\]
Show that we can write
\[\hat{y}_i = \sum_{j=1}^na_jy_j.\]
What is $a_j$ ?

*We have immediately that
\[\hat{y}_i = x_i\frac{\sum_{j=1}^n x_jy_j}{\sum_{k=1}^nx_k^2} = \sum_{j=1}^n\frac{x_ix_j}{\sum_{k=1}^nx_k^2}y_j = \sum_{j=1}^na_jy_j.\]*

#### 6.)

Using (3.4), argue that in the case of simple linear regression, the least squares line always passes through the point $(\overline{x},\overline{y})$.

*The least square line equation is $y = \hat{\beta}_0 + \hat{\beta}_1x$, so if we substitute $\overline{x}$ for $x$ we obtain
\[y = \hat{\beta}_0 + \hat{\beta}_1\overline{x} = \overline{y} - \hat{\beta}_1\overline{x} + \hat{\beta}_1\overline{x} = \overline{y}.\]
We may conclude that the least square line passes through the point $(\overline{x},\overline{y})$.*

#### 7.)

It is claimed in the text that in the case of simple linear regression of $Y$ onto $X$, the $R^2$ statistic (3.17) is equal to the square of the correlation between $X$ and $Y$ (3.18). Prove that this is the case. For simplicity, you may assume that $\overline{x} = \overline{y} = 0$.

*We have the following equalities
\[R^2 = 1 - \frac{RSS}{TSS} = 1 - \frac{\sum_i(y_i - \hat{y}_i)^2}{\sum_jy_j^2};\]
with $\hat{y}_i = \hat{\beta}_1x_i$ we may write
\[R^2 = 1 - \frac{\sum_i(y_i - \sum_jx_jy_j/\sum_jx_j^2 x_i)^2}{\sum_jy_j^2} = \frac{\sum_jy_j^2 - (\sum_iy_i^2 - 2\sum_iy_i(\sum_jx_jy_j/\sum_jx_j^2)x_i + \sum_i(\sum_jx_jy_j/\sum_jx_j^2)^2x_i^2)}{\sum_jy_j^2}\]
and finally
\[R^2 = \frac{2(\sum_ix_iy_i)^2/\sum_jx_j^2 - (\sum_ix_iy_i)^2/\sum_jx_j^2}{\sum_jy_j^2} = \frac{(\sum_ix_iy_i)^2}{\sum_jx_j^2\sum_jy_j^2} = Cor(X, Y)^2.\]*

### Applied

#### 8.)

This question involves the use of simple linear regression on the "Auto" data set.

(a) Use the lm() function to perform a simple linear regression with "mpg" as the response and "horsepower" as the predictor. Use the summary() function to print the results. Comment on the output. For example :

i. Is there a relationship between the predictor and the response ?

```{r}
auto <- as.data.table(ISLR::Auto)
fit <- lm(mpg ~ horsepower, data = Auto)
summary(fit)
```

*We can answer this question by testing the hypothesis $H_0 : \beta_i = 0\ \forall i$. The p-value corresponding to the F-statistic is `r as.numeric(pf(summary(fit)$fstatistic[1], summary(fit)$fstatistic[2], summary(fit)$fstatistic[3], lower.tail = F))`, this indicates a clear evidence of a relationship between "mpg" and "horsepower".*

ii. How strong is the relationship between the predictor and the response ?

*To calculate the residual error relative to the response we use the mean of the response and the RSE. The mean of mpg is `r mean(Auto$mpg)`. The RSE of the lm.fit was `r summary(fit)$sigma` which indicates a percentage error of `r summary(fit)$sigma / mean(Auto$mpg) * 100`%. We may also note that as the $R^2$ is equal to `r summary(fit)$r.sq`, almost `r summary(fit)$r.sq * 100`% of the variability in "mpg" can be explained using "horsepower".*

iii. Is the relationship between the predictor and the response positive or negative ?

*As the coeficient of "horsepower" is negative, the relationship is also negative. The more horsepower an automobile has the linear regression indicates the less mpg fuel efficiency the automobile will have.*

iv. What is the predicted $mpg$ associated with a "horsepower" of 98 ? What are the associated 95% confidence and prediction intervals ?

*Predicted Horsepower: `r predict(fit, newdata = data.frame(horsepower = 98))`*

Confidence/Prediction intervals:

```{r}
predict(fit, data.frame(horsepower = 98), interval = "confidence")
predict(fit, data.frame(horsepower = 98), interval = "prediction")
```

(b) Plot the response and the predictor. Also display the least squares regression line.

```{r, fig.height=3}
ggplot(auto, aes(horsepower, mpg)) +
   geom_point(alpha = .3) +
   geom_smooth(se = F, method = "lm") +
   ggtitle("Scatterplot of mpg vs. horsepower")
```

(c) Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit.

```{r}
par(mfrow = c(2, 2))
plot(fit)
```

*The plot of residuals versus fitted values indicates the presence of non linearity in the data. The plot of standardized residuals versus leverage indicates the presence of a few outliers (higher than 2 or lower than -2) and a few high leverage points.*

#### 9.)

This question involves the use of multiple linear regression on the "Auto" data set.

(a) Produce a scatterplot matrix which include all the variables in the data set.

```{r, echo = T, fig.width=9, fig.height=6}
ggpairs(auto[, -"name", with = F])
```

(b) Compute the matrix of correlations between the variables using the function cor(). You will need to exclude the "name" variable, which is qualitative.

```{r, echo = T}
cor(auto[, -"name", with = F])
```

(c) Use the lm() function to perform a multiple linear regression with "mpg" as the response and all other variables except "name" as the predictors. Use the summary() function to print the results. Comment on the output. For instance :

```{r, echo = T}
summary(fit <- lm(mpg ~ . -name, data = auto))
```

i. Is there a relationship between the predictors and the response ?

*Displacement, weight, year and origin seem to have statistically significant impacts on the response.*

(d) Use the plot() function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers ? Does the leverage plots identify any observations with unusually high leverages ?

```{r, echo = T}
par(mfrow = c(2,2))
plot(fit)
```

*The residuals appear to be approximately normally distributed, however there is a skewness in the right tail. There is one high leverage point (14).*

(e) Use the * and : symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant ?

*From the correlation matrix, we obtained the two highest correlated pairs and used them in picking interaction effects.*

```{r, echo = T}
summary(lm.fig <- lm(mpg ~ cylinders * displacement+displacement * weight, data = auto))
```

(f) Try a few different transformations of the variables, such as $\log{X}$, $\sqrt{X}$, $X^2$. Comment on your findings.

```{r, fig.height=5}
p1 <- ggplot(auto, aes(log(horsepower), mpg)) +
   geom_point() +
   geom_smooth(se = F, method = "lm") +
   ggtitle("Log Transform")

p2 <- ggplot(auto, aes(sqrt(horsepower), mpg)) +
   geom_point() +
   geom_smooth(se = F, method = "lm") +
   ggtitle("Sqrt Transform")

p3 <- ggplot(auto, aes(horsepower**2, mpg)) +
   geom_point() +
   geom_smooth(se = F, method = "lm") +
   ggtitle("Squared Transform")

grid.arrange(p1, p2, p3, nrow = 3)
```

*We limit ourselves to examining "horsepower" as sole predictor. It seems that the log transformation gives the most linear looking plot.*

#### 10.)

This question should be answered using the "Carseats" data set.

(a) Fit a multiple regression model to predict "Sales" using "Price", "Urban" and "US

```{r}
carseats <- as.data.table(ISLR::Carseats)
summary(fit3 <- lm(Sales ~ Price + Urban + US, data = carseats))
```

(b) Provide an interpretation of each coefficient in the model. Be careful - some of the variables in the model are qualitative !

*The coefficient of the "Price" variable may be interpreted by saying that the average effect of a price increase of 1 dollar is a decrease of `r abs(summary(fit3)$coef[2, 1]) * 1000` units in sales all other predictors remaining fixed. The coefficient of the "Urban" variable may be interpreted by saying that on average the unit sales in urban location are `r abs(summary(fit3)$coef[3, 1]) * 1000` units less than in rural location all other predictors remaining fixed. The coefficient of the "US" variable may be interpreted by saying that on average the unit sales in a US store are `r abs(summary(fit3)$coef[4, 1]) * 1000` units more than in a non US store all other predictors remaining fixed.*

(c) Write out the model in equation form, being careful to handle the qualitative variables properly.

*The model may be written as
\[Sales = `r summary(fit3)$coef[1, 1]` + (`r summary(fit3)$coef[2, 1]`)\times Price + (`r summary(fit3)$coef[3, 1]`)\times Urban + (`r summary(fit3)$coef[4, 1]`)\times US + \varepsilon\]
with $Urban = 1$ if the store is in an urban location and $0$ if not, and $US = 1$ if the store is in the US and $0$ if not.*

(d) For which of the predictors can you reject the null hypothesis $H_0 : \beta_j = 0$ ?

*We can reject the null hypothesis for the "Price" and "US" variables.*

(e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.

```{r}
fit4 <- lm(Sales ~ Price + US, data = carseats)
summary(fit4)
```

(f) How well do the models in (a) and (e) fit the data ?

*The $R^2$ for the smaller model is marginally better than for the bigger model. Essentially about `r summary(fit4)$r.sq * 100`% of the variability is explained by the model.*

(g) Using the model from (e), obtain 95% confidence intervals for the coefficient(s).

```{r}
confint(fit4)
```

(h) Is there evidence of outliers or high leverage observations in the model from (e) ?

```{r, fig.height=5}
par(mfrow = c(2, 2))
plot(fit4)
```

*The plot of standardized residuals versus leverage indicates the presence of a few outliers (higher than 2 or lower than -2) and some leverage points as some points exceed $(p + 1)/n$ (`r (3 + 1) / 400`).*

#### 11.)

In this problem we will investigate the t-statistic for the null hypothesis $H_0 : \beta = 0$ in simple linear regression without an intercept. To begin, we generate a predictor $x$ and a response $y$ as follows.

```{r}
set.seed(1)
x <- rnorm(100)
y <- 2 * x + rnorm(100)
```

(a) Perform a simple linear regression of $y$ onto $x$, without an intercept. Report the coefficient estimate $\hat{\beta}$, the standard error of this coefficient estimate, and the t-statistic and p-value associated with the null hypothesis $H_0$. Comment on these results.

```{r}
fit5 <- lm(y ~ x + 0)
summary(fit5)
```

*According to the summary above, we have a value of `r summary(fit5)$coef[1, 1]` for $\hat{\beta}$, a value of `r summary(fit5)$coef[1, 2]` for the standard error, a value of `r summary(fit5)$coef[1, 3]` for the t-statistic and a value of `r summary(fit5)$coef[1, 4]` for the p-value. The small p-value allows us to reject $H_0$.*

(b) Now perform a simple linear regression of $x$ onto $y$, without an intercept. Report the coefficient estimate $\hat{\beta}$, the standard error of this coefficient estimate, and the t-statistic and p-value associated with the null hypothesis $H_0$. Comment on these results.

```{r}
fit6 <- lm(x ~ y + 0)
summary(fit6)
```

*According to the summary above, we have a value of `r summary(fit6)$coef[1, 1]` for $\hat{\beta}$, a value of `r summary(fit6)$coef[1, 2]` for the standard error, a value of `r summary(fit6)$coef[1, 3]` for the t-statistic and a value of `r summary(fit6)$coef[1, 4]` for the p-value. The small p-value allows us to reject $H_0$.*

(c) What is the relationship between the results obtained in (a) and (b) ?

*We obtain the same value for the t-statistic and consequently the same value for the corresponding p-value. Both results in (a) and (b) reflect the same line created in (a). In other words, $y = 2x + \varepsilon$ could also be written $x = 0.5(y − \varepsilon)$.*

(d) For the regrssion of $Y$ onto $X$ without an intercept, the t-statistic for $H_0 : \beta = 0$ takes the form $\hat{\beta}/SE(\hat{\beta})$, where $\hat{\beta}$ is given by (3.38), and where
\[SE(\hat{\beta}) = \sqrt{\frac{\sum_{i=1}^n(y_i - x_i\hat{\beta})^2}{(n - 1)\sum_{i=1}^nx_i^2}}.\]
Show algebraically, and confirm numerically in R, that the t-statistic can be written as
\[\frac{\sqrt{n - 1}\sum_{i=1}^nx_iy_i}{\sqrt{(\sum_{i=1}^nx_i^2)(\sum_{i=1}^ny_i^2) - (\sum_{i=1}^nx_iy_i)}}.\]

*We have
\[t = \frac{\sum_ix_iy_y/\sum_jx_j^2}{\sqrt{\sum_i(y_i - x_i\hat{\beta})^2/(n - 1)\sum_jx_j^2}} = \frac{\sqrt{n - 1}\sum_ix_iy_i}{\sqrt{\sum_jx_j^2\sum_i(y_i - x_i\sum_jx_jy_j/\sum_jx_j^2)^2}} = \frac{\sqrt{n - 1}\sum_ix_iy_i}{\sqrt{(\sum_jx_j^2)(\sum_jy_j^2) - (\sum_jx_jy_j)}}.\]
Now let's verify this result numerically.*

```{r}
n <- length(x)
t <- sqrt(n - 1)*(x %*% y)/sqrt(sum(x^2) * sum(y^2) - (x %*% y)^2)
as.numeric(t)
```

*We may see that the t above is exactly the t-statistic given in the summary of "fit6".*

(e) Using the results from (d), argue that the t-statistic for the regression of $y$ onto $x$ is the same t-statistic for the regression of $x$ onto $y$.

*It is easy to see that if we replace $x_i$ by $y_i$ in the formula for the t-statistic, the result would be the same.*

(f) In R, show that when regression is performed with an intercept, the t-statistic for $H_0 : \beta_1 = 0$ is the same for the regression of $y$ onto $x$ as it is the regression of $x$ onto $y$.

```{r}
fit7 <- lm(y ~ x)
summary(fit7)
fit8 <- lm(x ~ y)
summary(fit8)
```

*It is again easy to see that the t-statistic for "fit7" and "fit8" are both equal to `r summary(fit7)$coef[2, 3]`.*

#### 12.)

This problem involves simple linear regression without an intercept.

(a) Recall that the coefficient estimate $\hat{\beta}$ for the linear regression of $Y$ onto $X$ witout an intercept is given by (3.38). Under what circumstance is the coefficient estimate for the regression of $X$ onto $Y$ the same as the coefficient estimate for the regression of $Y$ onto $X$ ?

*The coefficient estimate for the regression of $Y$ onto $X$ is
\[\hat{\beta} = \frac{\sum_ix_iy_i}{\sum_jx_j^2};\]
The coefficient estimate for the regression of $X$ onto $Y$ is
\[\hat{\beta}' = \frac{\sum_ix_iy_i}{\sum_jy_j^2}.\]
The coefficients are the same iff $\sum_jx_j^2 = \sum_jy_j^2$.*

(b) Generate an example in R with $n = 100$ observations in which the coefficient estimate for the regression of $X$ onto $Y$ is different from the coefficient estimate for the regression of $Y$ onto $X$.

```{r}
set.seed(1)
x <- 1:100
sum(x^2)
y <- 2 * x + rnorm(100, sd = 0.1)
sum(y^2)
fit.Y <- lm(y ~ x + 0)
fit.X <- lm(x ~ y + 0)
summary(fit.Y)
summary(fit.X)
```

(c) Generate an example in R with $n = 100$ observations in which the coefficient estimate for the regression of $X$ onto $Y$ is the same as the coefficient estimate for the regression of $Y$ onto $X$.

```{r}
x <- 1:100
sum(x^2)
y <- 100:1
sum(y^2)
fit.Y <- lm(y ~ x + 0)
fit.X <- lm(x ~ y + 0)
summary(fit.Y)
summary(fit.X)
```

#### 13.)

In this exercise you will create some simulated data and will fit simple linear regression models to it. Make sure to use set.seed(1) prior to starting part (a) to ensure conistent results.

(a) Using the rnorm() function, create a vector, "x", containing 100 observations drawn from a $N(0,1)$ distribution. This represents a feature, $X$.

```{r}
set.seed(1)
x <- rnorm(100)
```

(b) Using the rnorm() function, create a vector, "eps", containing 100 observations drawn from a $N(0, 0.25)$ distribution.

```{r}
eps <- rnorm(100, sd = sqrt(0.25))
```

(c) Using "x" and "eps", generate a vector "y" according to the model
\[Y = -1 + 0.5X + \varepsilon.\]
What is the length of the vector "y" ? What are the values of $\beta_0$ and $\beta_1$ in this linear model ?

```{r}
y <- -1 + 0.5 * x + eps
length(y)
```

*The values of $\beta_0$ and $\beta_1$ are $-1$ and $0.5$ respectively.*

(d) Create a scatterplot displaying the relationship between "x" and "y". Comment on what you observe.

```{r}
plot(x, y)
```

*The relationship between "x" and "y" looks linear with some noise introduced by the "eps" variable.*

(e) Fit a least squares linear model to predict "y" using "x". Comment on the model obtained. How do $\hat{\beta}_0$ and $\hat{\beta}_1$ compare to $\beta_0$ and $\beta_1$ ?

```{r}
fit9 <- lm(y ~ x)
summary(fit9)
```

*The values of $\hat{\beta}_0$ and $\hat{\beta}_1$ are pretty close to $\beta_0$ and $\beta_1$. The model has a large F-statistic with a near-zero p-value so the null hypothesis can be rejected.*

(f) Display the least squares line on the scatterplot obtained in (d). Draw the population regression line on the plot, in a different color. Use the legend() function to create an appropriate legend.

```{r}
plot(x, y)
abline(fit9, col = "red")
abline(-1, 0.5, col = "blue")
legend("topleft", c("Least square", "Regression"), col = c("red", "blue"), lty = c(1, 1))
```

(g) Now fit a polynomial regression model that predicts "y" using "x" and "x^2". Is there evidence that the quadratic term improves the model fit ? Explain your answer.

```{r}
fit10 <- lm(y ~ x + I(x^2))
summary(fit10)
```

*The coefficient for "x^2" is not significant as its p-value is higher than 0.05. So there is not sufficient evidence that the quadratic term improves the model fit even though the $R^2$ is slightly higher and $RSE$ slightly lower than the linear model.*

(h) Repeat (a)-(f) after modifying the data generation process in such a way that there is less noise in the data. The initial model should remain the same. Describe your results.

```{r}
set.seed(1)
eps <- rnorm(100, sd = 0.125)
x <- rnorm(100)
y <- -1 + 0.5 * x + eps
plot(x, y)
fit11 <- lm(y ~ x)
summary(fit11)
abline(fit11, col = "red")
abline(-1, 0.5, col = "blue")
legend("topleft", c("Least square", "Regression"), col = c("red", "blue"), lty = c(1, 1))
```

*We reduced the noise by decreasing the variance of the normal distribution used to generate the error term $\varepsilon$. We may see that the coefficients are very close to the previous ones, but now, as the relationship is nearly linear, we have a much higher $R^2$ and much lower $RSE$. Moreover, the two lines overlap each other as we have very little noise.*

(i) Repeat (a)-(f) after modifying the data generation process in such a way that there is more noise in the data. The initial model should remain the same. Describe your results.

```{r}
set.seed(1)
eps <- rnorm(100, sd = 0.5)
x <- rnorm(100)
y <- -1 + 0.5 * x + eps
plot(x, y)
fit12 <- lm(y ~ x)
summary(fit12)
abline(fit12, col = "red")
abline(-1, 0.5, col = "blue")
legend("topleft", c("Least square", "Regression"), col = c("red", "blue"), lty = c(1, 1))
```

*We increased the noise by increasing the variance of the normal distribution used to generate the error term $\varepsilon$. We may see that the coefficients are again very close to the previous ones, but now, as the relationship is not quite linear, we have a much lower $R^2$ and much higher $RSE$. Moreover, the two lines are wider apart but are still really close to each other as we have a fairly large data set.*

(j) What are the confidence intervals for $\beta_0$ and $\beta_1$ based on the original data set, the noisier data set, and the less noisy data set ? Comment on your results.

```{r}
confint(fit9)
confint(fit11)
confint(fit12)
```

*All intervals seem to be centered on approximately 0.5. As the noise increases, the confidence intervals widen. With less noise, there is more predictability in the data set.*

#### 14.)

This problem focuses on the collinearity problem.

(a) Perform the following commands in R.

```{r}
set.seed(1)
x1 <- runif(100)
x2 <- 0.5 * x1 + rnorm(100)/10
y <- 2 + 2 * x1 + 0.3 * x2 + rnorm(100)
```

The last line corresponds to creating a linear model in which "y" is a function of "x1" and "x2". Write out the form of the linear model. What are the regression coefficients ?

*The form of the linear model is
\[Y = 2 + 2X_1 +0.3X_2 + \varepsilon\]
with $\varepsilon$ a $N(0,1)$ random variable. The regression coefficients are respectively 2, 2 and 0.3.*

(b) What is the correlation between "x1" and "x2" ? Create a scatterplot displaying the relationship between the variables.

```{r}
cor(x1, x2)
plot(x1, x2)
```

*The variables seem highly correlated.*

(c) Using this data, fit a least squares regression to predict "y" using "x1" and "x2". Describe the results obtained. What are $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\beta}_2$ ? How do these relate to the true $\beta_0$, $\beta_1$ and $\beta_2$ ? Can you reject the null hypothesis $H_0 : \beta_1 = 0$ ? How about the null hypothesis $H_0 : \beta_2 = 0$ ?

```{r}
fit13 <- lm(y ~ x1 + x2)
summary(fit13)
```

*The coefficients $\hat{\beta}_0$, $\hat{\beta}_1$ and $\hat{\beta}_2$ are respectively `r summary(fit13)$coef[1, 1]`, `r summary(fit13)$coef[2, 1]` and `r summary(fit13)$coef[3, 1]`. Only $\hat{\beta}_0$ is close to $\beta_0$. As the p-value is less than 0.05 we may reject $H_0$ for $\beta_1$, however we may not reject $H_0$ for $\beta_2$ as the p-value is higher than 0.05.*

(d) Now fit a least squares regression to predict "y" using only "x1". Comment on your results. Can you reject the null hypothesis $H_0 : \beta_1 = 0$ ?

```{r}
fit14 <- lm(y ~ x1)
summary(fit14)
```

*The coefficient for "x1" in this last model is very different from the one with "x1" and "x2" as predictors. In this case "x1" is highly significant as its p-value is very low, so we may reject $H_0$.*

(e) Now fit a least squares regression to predict "y" using only "x2". Comment on your results. Can you reject the null hypothesis $H_0 : \beta_1 = 0$ ?

```{r}
fit15 <- lm(y ~ x2)
summary(fit15)
```

*The coefficient for "x2" in this last model is very different from the one with "x1" and "x2" as predictors. In this case "x2" is highly significant as its p-value is very low, so we may again reject $H_0$.*

(f) Do the results obtained in (c)-(e) contradict each other ? Explain your answer.

*No, the results do not contradict each other. As the predictors "x1" and "x2" are highly correlated we are in the presence of collinearity, in this case it can be difficult to determine how each predictor separately is associated with the response. Since collinearity reduces the accuracy of the estimates of the regression coefficients, it causes the standard error for $\hat{\beta}_1$ to grow (we have a standard error of `r summary(fit13)$coef[2, 2]` and `r summary(fit13)$coef[3, 2]` for "x1" and "x2" respectively in the model with two predictors and only of `r summary(fit14)$coef[2, 2]` and `r summary(fit15)$coef[2, 2]` for "x1" and "x2" respectively in the models with only one predictor). Consequently, we may fail to reject $H_0$ in the presence of collinearity. The importance of the "x2" variable has been masked due to the presence of collinearity.*

(g) Now suppose we obtain one additional observation, which was unfortunately mismeasured.

```{r}
x1 <- c(x1, 0.1)
x2 <- c(x2, 0.8)
y <- c(y, 6)
```

Re-fit the linear models from (c) to (e) using this new data. What effect does this new observation have on each of the models ? In each model, is this observation an outlier ? A high-leverage point ? Explain your answers.

```{r}
fit16 <- lm(y ~ x1 + x2)
fit17 <- lm(y ~ x1)
fit18 <- lm(y ~ x2)
summary(fit16)
summary(fit17)
summary(fit18)
plot(fit16)
plot(fit17)
plot(fit18)
```

*In the model with two predictors, the last point is a high-leverage point. In the model with "x1" as sole predictor, the last point is an outlier. In the model with "x2" as sole predictor, the last point is a high leverage point.*

#### 15.)

This problem involves the "Boston" data set, which we saw in the lab for this chapter. We will now try to predict per capita crime rate using the other variables in this data set. In other words, per capita crime rate is the response, and the other variables are the predictors.

(a) For each predictor, fit a simple linear regression model to predict the response. Describe your results. In which of the models is there a statistically significant association between the predictor and the response ? Create some plots to back up your assertions.

```{r}
attach(Boston)
fit.zn <- lm(crim ~ zn)
summary(fit.zn)
fit.indus <- lm(crim ~ indus)
summary(fit.indus)
chas <- as.factor(chas)
fit.chas <- lm(crim ~ chas)
summary(fit.chas)
fit.nox <- lm(crim ~ nox)
summary(fit.nox)
fit.rm <- lm(crim ~ rm)
summary(fit.rm)
fit.age <- lm(crim ~ age)
summary(fit.age)
fit.dis <- lm(crim ~ dis)
summary(fit.dis)
fit.rad <- lm(crim ~ rad)
summary(fit.rad)
fit.tax <- lm(crim ~ tax)
summary(fit.tax)
fit.ptratio <- lm(crim ~ ptratio)
summary(fit.ptratio)
fit.black <- lm(crim ~ black)
summary(fit.black)
fit.lstat <- lm(crim ~ lstat)
summary(fit.lstat)
fit.medv <- lm(crim ~ medv)
summary(fit.medv)
```

*To find which predictors are significant, we have to test $H_0 : \beta_1 = 0$. All predictors have a p-value less than 0.05 except "chas", so we may conclude that there is a statistically significant association between each predictor and the response except for the "chas" predictor.*

(b) Fit a multiple regression model to predict the response using all of the predictors. Describe your results. For which predictors can we reject the null hypothesis $H_0 : \beta_j = 0$ ?

```{r}
fit.all <- lm(crim ~ ., data = Boston)
summary(fit.all)
```

*We may reject the null hypothesis for "zn", "dis", "rad", "black" and "medv".*

(c) How do your results from (a) compare to your results from (b) ? Create a plot displaying the univariate regression coefficients from (a) on the x-axis, and the multiple regression coefficients from (b) on the y-axis. That is, each predictor is displayed as a single point on the plot. Its coefficient in a simple linear regression model is shown on the x-axis, and its coefficient estimate in the multiple linear regression model is shown on the y-axis.

```{r}
simple.reg <- vector("numeric",0)
simple.reg <- c(simple.reg, fit.zn$coefficient[2])
simple.reg <- c(simple.reg, fit.indus$coefficient[2])
simple.reg <- c(simple.reg, fit.chas$coefficient[2])
simple.reg <- c(simple.reg, fit.nox$coefficient[2])
simple.reg <- c(simple.reg, fit.rm$coefficient[2])
simple.reg <- c(simple.reg, fit.age$coefficient[2])
simple.reg <- c(simple.reg, fit.dis$coefficient[2])
simple.reg <- c(simple.reg, fit.rad$coefficient[2])
simple.reg <- c(simple.reg, fit.tax$coefficient[2])
simple.reg <- c(simple.reg, fit.ptratio$coefficient[2])
simple.reg <- c(simple.reg, fit.black$coefficient[2])
simple.reg <- c(simple.reg, fit.lstat$coefficient[2])
simple.reg <- c(simple.reg, fit.medv$coefficient[2])
mult.reg <- vector("numeric", 0)
mult.reg <- c(mult.reg, fit.all$coefficients)
mult.reg <- mult.reg[-1]
plot(simple.reg, mult.reg, col = "red")
```

*There is a difference between the simple and multiple regression coefficients. This difference is due to the fact that in the simple regression case, the slope term represents the average effect of an increase in the predictor, ignoring other predictors. In contrast, in the multiple regression case, the slope term represents the average effect of an increase in the predictor, while holding other predictors fixed. It does make sense for the multiple regression to suggest no relationship between the response and some of the predictors while the simple linear regression implies the opposite because the correlation between the predictors show some strong relationships between some of the predictors.*

```{r}
cor(Boston[-c(1, 4)])
```

*So for example, when "age" is high there is a tendency in "dis" to be low, hence in simple linear regression which only examines "crim" versus "age", we observe that higher values of "age" are associated with higher values of "crim", even though "age" does not actually affect "crim". So "age" is a surrogate for "dis"; "age" gets credit for the effect of "dis" on "crim".*

(d) Is there evidence of non-linear association between any of the predictors and the response ? To answer this question, for each predictor $X$, fit a model of the form
\[Y = \beta_0 + \beta_1 X + \beta_2 X^2 + \beta_3 X^3 + \varepsilon.\]

```{r}
fit.zn2 <- lm(crim ~ poly(zn, 3))
summary(fit.zn2)
fit.indus2 <- lm(crim ~ poly(indus, 3))
summary(fit.indus2)
fit.nox2 <- lm(crim ~ poly(nox, 3))
summary(fit.nox2)
fit.rm2 <- lm(crim ~ poly(rm, 3))
summary(fit.rm2)
fit.age2 <- lm(crim ~ poly(age, 3))
summary(fit.age2)
fit.dis2 <- lm(crim ~ poly(dis, 3))
summary(fit.dis2)
fit.rad2 <- lm(crim ~ poly(rad, 3))
summary(fit.rad2)
fit.tax2 <- lm(crim ~ poly(tax, 3))
summary(fit.tax2)
fit.ptratio2 <- lm(crim ~ poly(ptratio, 3))
summary(fit.ptratio2)
fit.black2 <- lm(crim ~ poly(black, 3))
summary(fit.black2)
fit.lstat2 <- lm(crim ~ poly(lstat, 3))
summary(fit.lstat2)
fit.medv2 <- lm(crim ~ poly(medv, 3))
summary(fit.medv2)
```

*For "zn", "rm", "rad", "tax" and "lstat" as predictor, the p-values suggest that the cubic coefficient is not statistically significant; for "indus", "nox", "age", "dis", "ptratio" and "medv" as predictor, the p-values suggest the adequacy of the cubic fit; for "black" as predictor, the p-values suggest that the quandratic and cubic coefficients are not statistically significant, so in this latter case no non-linear effect is visible.*

### 2020-03-01 Statistical Learning/04_Classification.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 4}
   \fancyfoot[C]{\rmfamily\color{headergrey}Classification}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select
```

## Chapter 4

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Book Work

#### Simple Logistic Regression

```{r}
data.default <- data.table(ISLR::Default)[, dflt := ifelse(default == "Yes", 1, 0)]

summary(model1 <- glm(dflt ~ balance, data = data.default, family = "binomial"))

predict(model1, newdata = data.frame( balance = c(1000, 2000) ), type = "response")

data.default[, is_student := ifelse(student == "Yes", 1, 0)]

summary(model2 <- glm(dflt ~ is_student, data = data.default, family = "binomial"))

predict(model2, newdata = data.frame( is_student = c(1, 0) ), type = "response")
```

#### Multiple Logistic Regression

```{r}
summary(model3 <- glm(dflt ~ balance + is_student, data = data.default, family = "binomial"))

p1 <- ggplot(data.default, aes(balance, dflt, color = student)) +
   stat_ecdf() +
   geom_rug(aes(balance, dflt)) +
   labs(x = "Balance", y = "Default Rate", title = "Default Rate by Student/Balance")

p2 <- ggplot(data.default, aes(student, balance, fill = student)) +
   geom_boxplot()

grid.arrange(p1, p2, nrow = 1)

predict(model3, newdata = 
           data.frame( balance = c(1500, 1500), 
                       is_student = c(1, 0) ), 
        type = "response")
```

### R Lab

```{r}
Smarket <- as.data.table(ISLR::Smarket)

names(Smarket)

dim(Smarket)

summary(Smarket)
```

Pairs
```{r}
ggpairs(Smarket) %>%
   print(progress = F)
```

```{r}
cor(Smarket %>% select(-Direction))
```

```{r, fig.height=5}
ggplot(Smarket) +
   geom_boxplot(aes(Year, Volume, group = Year))
```

### Logistic Regression

```{r}
summary(glm.fits <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume,
                 data = Smarket, family = binomial))

coef(glm.fits)
```

Probabilites of going up (first 10 trading days)

```{r}
glm.probs <- predict(glm.fits, type = "response")
head(glm.probs, 10)
```

```{r}
contrasts(Smarket$Direction)
```

Predictions

```{r}
glm.pred <- rep("Down", nrow(Smarket))
glm.pred[glm.probs > 0.5] <- "Up"
```

```{r}
table(glm.pred, Smarket$Direction)

mean(glm.pred == Smarket$Direction)
```

#### Validation

Get the holdout set.

```{r}
train <- (Smarket$Year < 2005)
Smarket.2005 <- Smarket[!train]
dim(Smarket.2005)

Direction.2005 <- Smarket$Direction[!train]
```

Train the logistic regression model.

```{r}
glm.fits <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume,
                data = Smarket, family = binomial, subset = train)

glm.probs <- predict(glm.fits, Smarket.2005, type = "response")
```

Test

```{r}
glm.pred <- rep("Down", 252)
glm.pred[glm.probs > 0.5] <- "Up"

table(glm.pred, Direction.2005)

mean(glm.pred == Direction.2005)
```

Model 2

```{r}
summary(glm.fits <- glm(Direction ~ Lag1 + Lag2, data = Smarket, family = binomial, subset = train))

glm.probs <- predict(glm.fits, Smarket.2005, type = "response")
glm.pred <- rep("Down", nrow(Smarket.2005))

glm.pred[glm.probs >= 0.5] <- "Up"

table(glm.pred, Direction.2005)

mean(glm.pred == Direction.2005)
```

```{r}
predict(glm.fits, newdata = data.table(Lag1 = c(1.2, 1.5),
                                       Lag2 = c(1.1, -0.8)), 
        type = "response")
```

#### Linear Discriminant Analysis

*LDA* is from **MASS** package.

```{r}
summary(lda.fit <- lda(Direction ~ Lag1 + Lag2, data = Smarket, subset = train))

lda.fit
```

```{r}
lda.pred <- predict(lda.fit, Smarket.2005)

names(lda.pred)
```

Predictions:

```{r}
lda.class <- lda.pred$class
table(lda.class, Direction.2005)
```

Note: almost identical to logistic regression.

```{r}
mean(lda.class == Direction.2005)
```

```{r}
sum(lda.pred$posterior[, 1] >= 0.5)
sum(lda.pred$posterior[, 1] < 0.5)
```

The posterior probabilites output by the model corresponds to the probability that the market will decrease.

```{r}
lda.pred$posterior[1:20, 1]
lda.class[1:20]
```

Apply a threshold of 90% to predictions:

```{r}
sum(lda.pred$posterior[, 1] > .9)
```

#### Quadratic Discriminant Analysis

```{r}
summary(qda.fit <- qda(Direction ~ Lag1 + Lag2, data = Smarket, subset = train))

qda.fit
```

Predictions

```{r}
qda.class <- predict(qda.fit, Smarket.2005)$class
table(qda.class, Direction.2005)

mean(qda.class == Direction.2005)
```

#### K-Nearest Neighbors

Data Setup

```{r}
train.X <- with(Smarket, cbind(Lag1, Lag2))[train, ]
test.X <- with(Smarket, cbind(Lag1, Lag2))[!train, ]
train.Direction <- Smarket$Direction[train]
```

KNN

```{r}
set.seed(1)

knn.pred <- knn(train.X, test.X, train.Direction, k = 1)
table(knn.pred, Direction.2005)
```

```{r}
set.seed(1)

knn.pred <- knn(train.X, test.X, train.Direction, k = 3)
table(knn.pred, Direction.2005)
```

#### Caravan Insurance Data

```{r}
caravan <- Caravan

dim(caravan)

summary(caravan$Purchase)

table(caravan$Purchase) %>% prop.table()
```

```{r}
standardized.X <- scale(caravan[, -86])

var(caravan[, 1])
var(caravan[, 2])

var(standardized.X[, 1])
var(standardized.X[, 2])
```

+ K=1
```{r}
test <- 1:1000
train.X <- standardized.X[-test,]
test.X <- standardized.X[test,]

train.Y <- caravan$Purchase[-test]
test.Y <- caravan$Purchase[test]

set.seed(1)

knn.pred <- knn(train.X, test.X, train.Y, k = 1)
mean(test.Y != knn.pred)
mean(test.Y != "No")

result <- table(knn.pred, test.Y)
result

result %>% prop.table()
```

+ K=3

```{r}
set.seed(1)

knn.pred <- knn(train.X, test.X, train.Y, k = 3)
mean(test.Y != knn.pred)
mean(test.Y != "No")

result <- table(knn.pred, test.Y)
result

result %>% prop.table()
```

+ K=5

```{r}
set.seed(1)

knn.pred <- knn(train.X, test.X, train.Y, k = 5)
mean(test.Y != knn.pred)
mean(test.Y != "No")

result <- table(knn.pred, test.Y)
result

result %>% prop.table()
```

Logistic Regression Alternative

```{r}
glm.fits <- glm(Purchase ~ ., data = caravan, family = binomial,
                subset = -test)
glm.probs <- predict(glm.fits, caravan[test,], type = "response")

# .5 cut-off
glm.pred <- rep("No", 1000)
glm.pred[glm.probs > .5] <- "Yes"

results <- table(glm.pred, test.Y)
results %>% prop.table()

# .25 cut-off
glm.pred <- rep("No", 1000)
glm.pred[glm.probs > .25] <- "Yes"

results <- table(glm.pred, test.Y)
results %>% prop.table()
```

```{r}

# Quiz

bal <- 1936.75

exp(-10.6513 + 0.0055 * bal) / ( 1 + exp(-10.6513 + 0.0055*bal))

b0 <- -6; b1 <- 0.05; b2 <- 1
x1 <- 50; x2 <- 3.5
exp(b0 + b1 * x1 + b2 * x2) / ( 1 + exp(b0 + b1 * x1 + b2 * x2))

```

### Conceptual

#### 1.)

Using a little bit of algebra, prove that (4.2) is equivalent to (4.3). In other words, the logistic function representaion and logit representation for the logistic regression models are equivalent.

4.2) $p(x) = \frac{e^{(\beta_0 + \beta_1X)}}{1 + e^{\beta_0 + \beta_1X}}$

4.3) $\frac{p(x)}{1 - p(x)} = e^{\beta_0 + \beta_1X}$

**Solution**

$1 - p(x) = 1 - \frac{e^{(\beta_0 + \beta_1X)}}{1 + e^{\beta_0 + \beta_1X}} = \frac{1}{1 + e^{\beta_0 + \beta_1X}}$

$\frac{1}{1 - p(x)} = 1 + e^{\beta_0 + \beta_1X}$

#### 2.)

It was stated in the text that classifying an observation to the class for which (4.13) is largest. Prove that this is the case. In other words, under the assumption that the observations in the *k*th class are drawn from a $N \sim (\mu, \sigma^2)$ distribution, the Bayes' clssifier assigns an observation to the class for which the discriminant function is maximized.

4.12:

$p_k(x) = \frac{\pi_k\frac{1}{\sqrt{2\pi\sigma}}e^{-\frac{1}{2\sigma^2}(x - \mu_k)^2}}{\sum^K_{l=1}{\pi_le^{-\frac{1}{2\sigma^2}(x - \mu_k)^2}}}$

**Solution**

$f''_x = ln\pi_k + ln(\frac{1}{\sqrt{2\pi\sigma}}) + ln \ e^{- \frac{1}{2\sigma^2}(x - \mu_k)^2}$

$f'''_x = ln \ \pi_k - \frac{1}{2\sigma^2}(x - \mu_k)^2$

$f'''_x = ln \ \pi_k + \frac{x\mu_k}{\sigma^2} - \frac{\mu^2_k}{2\sigma^2}$

$\delta_k(x) = \frac{\mu_k}{\sigma^2}x - \frac{\mu^2_k}{2\sigma^2} + log(\pi_k)$

#### 3.)

This problem relates to the QDA model, in which the observations within each class are drawn from a normal distribution with a class-specific mean vector and a class specific covariance matrix. We consider the simple case where p=1; i.e. there is only one feature.
Suppose that we have K classes, and if an observation belongs to the kth class then X comes from a one-dimensional normal distribution, $X \sim N(\mu_k, \sigma^2_k)$. Recall that the density function for the one-dimensional normal distribution is given in (4.11). Prove that in this case, the Bayes’ classifier is not linear. Argue that it is in fact quadratic.

**Solution**

From the previous answer, we can expand the last term which is not linear in x.

#### 4.)

When the number of features p is large, there tends to be a deterioration in the performance of KNN and other local approaches that perform prediction using only observations that are near the test observation for which a prediction must be made. This phenomenon is known as the curse of dimensionality, and it ties into the fact that non-parametric approaches often perform poorly when p is large. We will now investigate this curse.

a.) Suppose that we have a set of observations, each with measurements on p=1 feature, X. We assume that X is uniformly (evenly) distributed on [0,1]. Associated with each observation is a response value. Suppose that we wish to predict a test observation’s response using only observations that are within 10% of the range of X closest to that test observation. For instance, in order to predict the response for a test observation with X=0.6, we will use observations in the range [0.55,0.65]. On average, what fraction of the available observations will we use to make the prediction ?

### Applied

#### 10.)

This question should be answered using the “Weekly” data set, which is part of the “ISLR” package. This data is similar in nature to the “Smarket” data from this chapter’s lab, except that it contains 1089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.

```{r}
weekly <- as.data.table(ISLR::Weekly)

head(weekly)
```

a.) Produce some numerical and graphical summaries of the “Weekly” data. Do there appear to be any patterns ?

```{r}
weekly %>% ggplot() +
   geom_point(aes(Year, Lag1))

p1 <- weekly %>% ggplot() +
   geom_density(aes(Lag1, group = Year, fill = Year), alpha = .4)

p2 <- weekly %>% ggplot() +
   geom_density(aes(Lag2, group = Year, fill = Year), alpha = .4)

p3 <- weekly %>% ggplot() +
   geom_density(aes(Lag3, group = Year, fill = Year), alpha = .4)

p4 <- weekly %>% ggplot() +
   geom_density(aes(Lag4, group = Year, fill = Year), alpha = .4)

grid.arrange(p1, p2, p3, p4, nrow = 2)
```

```{r}
cor <- cor(weekly %>% select(-Direction))
cor

ggcorr(cor)
```

b.) Use the full data set to perform a logistic regression with “Direction” as the response and the five lag variables plus “Volume” as predictors. Use the summary function to print the results. Do any of the predictors appear to be statistically significant ? If so, which ones ?

```{r}
summary(glm.fit <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = weekly, family = binomial))

glm.fit
```

It appears only **Lag2** is statistically significant.

c.) Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.

```{r}
glm.pred <- ifelse(predict(glm.fit, type = "response") > .5, "Up", "Down")

table(glm.pred, weekly$Direction) %>% prop.table()
```

We may conclude that the percentage of correct predictions on the training data is (54+557)/1089 wich is equal to 56.1065197%. In other words 43.8934803% is the training error rate, which is often overly optimistic. We could also say that for weeks when the market goes up, the model is right 92.0661157% of the time (557/(48+557)). For weeks when the market goes down, the model is right only 11.1570248% of the time (54/(54+430)).

d.) Now fit the logistic regression model using a training data period from 1990 to 2008, with “Lag2” as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 to 2010).

```{r}
train <- weekly[Year < 2009,]
test <- weekly[Year > 2009, .(Direction, Lag2)]

summary(glm.fit <- glm(Direction ~ Lag2, data = train, family = binomial))

test$pred <- ifelse(predict(glm.fit, newdata = test[, .(Lag2)], type = "response") > .5, "Up", "Down")

table(test$Direction, test$pred) %>% prop.table() * 100

mean(test$Direction == test$pred)
```

e.) Repeat (d) using LDA.

```{r}
summary(lda.fit <- lda(Direction ~ Lag2, data = train))
lda.fit

test$pred <- predict(lda.fit, newdata = test[, .(Lag2)], type = "response")$class

mean(test$Direction == test$pred)

table(test$Direction, test$pred) %>% prop.table() * 100
```

f.) Repeat (d) using QDA.

```{r}
summary(qda.fit <- qda(Direction ~ Lag2, data = train))
qda.fit

lda.test <- test[, .(Direction, Lag2)]

lda.pred <- predict(qda.fit, newdata = lda.test, type = "response")

lda.test[, pred := lda.pred$class]

with(lda.test, mean(Direction == pred))

with(lda.test, table(Direction, pred)) %>% prop.table() * 100
```

g.) Repeat (d) using KNN with k = 1.

```{r}
train <- weekly[Year < 2009,]
test <- weekly[Year > 2009, .(Direction, Lag2)]

train.X <- train[, .(Lag2)]
test.X <- test[, .(Lag2)]
train.Direction <- train$Direction

knn.pred <- knn(train.X, test.X, train.Direction, k = 1)

table(test$Direction, knn.pred) %>% prop.table()

mean(test$Direction == knn.pred)
```

h.) Which of these methods appear to provide the best results on this data?

LDA and Logistic Regression appear to have the best performance on this particular set of data.

#### 11.)

In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the *Auto* data set.

```{r}
auto <- as.data.table(ISLR::Auto)

auto %>% glimpse()
```

a.) Create a binary variable, *mpg01*, that contains a 1 if mpg contains a value above its median. You can compute the median using the median() function. Note you may find it helpful to use the *data.frame()* function to create a single data set containing both *mpg01* and other Auto variables.

```{r}
cutpoint <- median(auto$mpg)

auto[, mpg01 := mpg > cutpoint]

```

b.) Explore the data graphically in order to investigate the association between mpg01 and the other features. Which of the other features seem most likely to be useful in predicting mpg01? Scatterplots and boxplots may be useful tools to answer this question.

```{r}
summary(auto)

ggpairs(auto %>% select(-name))

ggplot(auto, aes(cylinders, mpg, fill = cylinders)) +
   geom_boxplot(aes(group = cylinders)) +
   geom_hline(yintercept = cutpoint, col = "darkred")

ggplot(auto, aes(displacement, mpg)) +
   geom_point(aes(color = cylinders)) +
   geom_smooth(method = "auto") +
   scale_y_continuous(trans = "log10")

ggplot(auto, aes(horsepower, mpg)) +
   geom_point(aes(color = mpg)) +
   geom_smooth(method = "auto") +
   scale_y_continuous(trans = "log10")

```

c.) Split the data in to training / test sets.

```{r}
auto[, prob := ifelse(mpg01 == T, 1, 0)]

auto.split <- initial_split(auto, prop = 0.7, strata = "mpg01")

auto.train <- training(auto.split)
auto.test <- testing(auto.split)
```

d.) Perform Logistic Regression.

```{r}

auto.train %>% glimpse()

glm.fit1 <- glm(mpg01 ~ cylinders, data = auto.train, family = binomial)
glm.fit2 <- glm(mpg01 ~ horsepower, data = auto.train, family = binomial)

auto.train1 <- broom::augment(glm.fit1, auto.train) %>% mutate(.fitted = exp(.fitted))
auto.train2 <- broom::augment(glm.fit2, auto.train) %>% mutate(.fitted = exp(.fitted))

auto.train1 %>% summary()
auto.train2 %>% summary()

p1 <- ggplot(auto.train1, aes(cylinders, prob)) +
   geom_point(alpha = 0.15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial"))

p2 <- ggplot(auto.train2, aes(horsepower, prob)) +
   geom_point(alpha = 0.15) +
   geom_smooth(method = "glm", method.args = list(family = "binomial"))

gridExtra::grid.arrange(p1, p2, nrow = 2)
```

```{r}
tidy(glm.fit1)
tidy(glm.fit2)

exp(coef(glm.fit1))
exp(coef(glm.fit2))

confint(glm.fit1)
confint(glm.fit2)
```

Cross-validated Logistic Regression

```{r}

cols.exclude <- c("name", "mpg", "prob")
auto.train3 <- auto.train[, -cols.exclude, with = F]

summary(glm.fit3 <- glm(mpg01 ~ ., data = auto.train3, family = binomial))

auto.train3 <- auto.train[, -cols.exclude, with = F][, mpg01 := as.factor(mpg01)]

cv.model.logit <- train(
   mpg01 ~ .,
   data = auto.train3,
   method = "glm",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

auto.train3$pred <- predict(cv.model.logit, auto.train3)

# in-sample performance
table(auto.train3$mpg01, auto.train3$pred) %>% prop.table()

# out of sample results
auto.test$pred <- predict(cv.model.logit, newdata = auto.test, type = "raw")

table(auto.test$mpg01, auto.test$pred) %>% prop.table() %>% round(digits = 3)
```

```{r}
ggplot(auto.train, aes(weight, prob, color = year)) +
   geom_point(alpha = 0.25) +
   geom_smooth(method = "glm", method.args = list(family = "binomial")) +
   scale_x_continuous(labels = scales::comma)
```

e.) Perform Linear Discriminant Analysis

Cross-validated Linear Discriminant Analysis

```{r}

auto.train.lda <- auto.train[, -cols.exclude, with = F][, mpg01 := as.factor(mpg01)]

summary(lda.fit1 <- lda(mpg01 ~ ., data = auto.train.lda))

cv.model.lda <- train(
   mpg01 ~ .,
   auto.train.lda,
   method = "lda",
   family = "binomial",
   trControl = trainControl(method = "cv", number = 10)
)

# in-sample performance
auto.train.lda$pred <- predict(cv.model.lda)

with(auto.train.lda, mean(mpg01 == pred))

with(auto.train.lda, table(mpg01, pred)) %>% prop.table()

# out of sample performance

auto.test.lda <- auto.test
auto.test.lda$pred <- predict(cv.model.lda, newdata = auto.test.lda)

with(auto.test.lda, mean(mpg01 == pred))

with(auto.test.lda, table(mpg01, pred)) %>% prop.table()
```

Cross-validated KNN

```{r}

auto.train.X <- auto.train[, .(mpg01, weight, year)][, mpg01 := as.factor(mpg01)]
auto.test.X <- auto.test[, .(mpg01, weight, year)]

knn.fit <- train(
   mpg01 ~ weight + year,
   data = auto.train.X,
   method = "knn",
   trControl = trainControl(method = "cv", number = 10),
   tuneLength = 20
)

summary(knn.fit)

auto.knn <- auto.train
# in-sample performance
auto.knn$pred <- predict(knn.fit)

with(auto.knn, mean(mpg01 == pred))
with(auto.knn, table(mpg01, pred)) %>% prop.table()

# out of sample performance
auto.test.knn <- auto.test

auto.test.knn$pred <- predict(knn.fit, newdata = auto.test.knn)

with(auto.test.knn, mean(mpg01 == pred))
with(auto.test.knn, table(mpg01, pred)) %>% prop.table()

```

#### 12.)

a.) Write a function, Power(), that prints out the result of rasing 2 to the 3rd power. In other words, your function should compute 2^3.

```{r}
Power <- function(x) {
   print(2^3)
}
Power()
```

b.) Create a new function, Power2, that allows you to pass any two numbers, x and a, and prints out the value of x^a.

```{r}
Power2 <- function(a, x) paste(a, "to the", x, "is", a ^ x)
Power2(3, 8)
```

clean-up workspace
```{r}
rm(list = ls())
```

### 2020-03-01 Statistical Learning/05_Resampling-Methods.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 5}
   \fancyfoot[C]{\rmfamily\color{headergrey}Resampling Methods}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select
```

## Chapter 5

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

### Notes

```{r}
load(paste0(here::here(), "/ISLR/5.R.RData"))

summary(lm(y ~ X1 + X2, data = Xy))

par(mfrow=c(1,1))
matplot(Xy,type="l")

alpha = function(x,y){
  vx = var(x)
  vy = var(y)
  cxy= cov(x,y)
  (vy-cxy)/(vx+vy-2*cxy)
}
alpha(Xy$X1,Xy$y)

alpha.fn = function(data,index){
  with(data[index,],alpha(Xy$X1,Xy$y))
}

alpha.fn<-function(data, index) {
  fit1<-lm(y~., data=Xy[index,])
  coefficients(fit1)[['X1']]
}

set.seed(1)
alpha.fn(Xy, sample(1:100,100,replace=TRUE))

boot.out <- boot(Xy,alpha.fn,R=1000)

?tsboot

se_stat <- function(data, index) {
   fit1 <- lm(y ~., data = data[index,])
   coefficients(fit1)[['X1']]
}

tsboot.out <- tsboot(Xy, se_stat, R = 1000, l = 100, sim = "fixed")
tsboot.out
```

#### Cross-Validation

Cross-validation is the process of splitting the training data into multiple subsets that can be used for evaluating the out-of-sample performance of a statistical model.

There are multiple strategies for this technique.

#### The Validation Set approach

The most basic strategy for CV is a training/test split of the sample.

Additionally, stratification can be used to ensure even splitting when the response variable is unevenly distributed in the sample (low response logistic regression, for example).

Using Auto:

```{r}
auto <- data.table(ISLR::Auto)
```

```{r}
models <- list()

base.model <- "mpg ~ horsepower"

prev <- ""
for(term in 1:10)
{
   cur <- ifelse(term > 1, paste(prev, rep(paste0("+ I(horsepower^", term,")"))), "")
   fmla <- as.formula(paste0(base.model, cur))

   train.error <- numeric(10); test.error <- numeric(10)

   for(iter in 1:10)
   {
      auto.split <- initial_split(auto, prop = .5)

      auto.train <- training(auto.split)
      model <- lm(fmla, data = auto.train)
      train.error[iter] <- mean(model$residuals^2)

      auto.test <- testing(auto.split)
      auto.test$pred <- predict(model, newdata = auto.test)
      test.error[iter] <- with(auto.test, mean( (mpg - pred)^2 ) )
   }

   models[[term]] <- list(terms = term, train.error = train.error, test.error = test.error)
   prev <- cur
}

auto.fits <- rbindlist(models, fill = F)

p1 <- auto.fits %>%
   group_by(terms) %>%
   summarise(mean_error = mean(train.error)) %>%
   ggplot(., aes(terms, mean_error)) +
      geom_point(col = "cornflowerblue") +
      geom_line(alpha = .15) +
      labs(title = "Horsepower Model vs Polynomial Degree", x = "Polynomial Degree", y = "Avg. Training MSE") +
      scale_y_continuous(limits = c(10, 30))

p2 <- auto.fits[, .(num = 1:.N, train.error), by = list(terms)] %>%
   ggplot() +
   geom_line(aes(num, train.error, group = terms, col = terms)) +
   labs(title = "Train Sample vs Error (MSE)", x = "Sample Number", y = "Training MSE") +
   theme(legend.position = "none")

gridExtra::grid.arrange(p1, p2, nrow = 1)

ggplot(auto.fits[, .(train = mean(train.error), test = mean(test.error)), by = list(terms)]) +
   geom_point(aes(terms, train), col = "cornflowerblue") +
   geom_point(aes(terms, test), col = "darkorange") +
   labs(x = "Polynomial Terms", y = "Avg. Error")

p1 <- ggplot(auto.fits) +
   geom_boxplot(aes(y = train.error, group = terms, fill = terms))

p2 <- ggplot(auto.fits) +
   geom_boxplot(aes(y = test.error, group = terms, fill = terms))

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Drawbacks:

+ 1.) The validation estimate of the test error rate can be highly variable, depending on precisely which observations are included in the training set and which observations are in the validation set.

+ 2.) In the validation approach, only a subset of the observations - those that are included in the training set rather than the validation set - are used to fit the model. This can lead to overestimation of model performance.

#### Leave-One-Out Cross-Validation

Leave-one-out cross-validation (LOOCV) is closely related to the validation set approach, but attempts to address some of the drawbacks.

The basic premise of LOOCV is to leave exactly 1 observation out of the data to test against, and use the remaining n-1 observations to train the model.

After each resample, the final test error is produced by:

$CV_{(n)} = \frac{1}{n}\sum^{n}_{i=1}{MSE_i}$

This is more of systematic approach to model training/validation.

Advantages:

+ Far less bias than the validation/test set.

#### k-Fold Cross-Validation

An alternative to LOOCV is k-Fold CV. This approach randomly divides the set of observations into *k* groups, or *k*-folds, of approximately equal size.

The first fold is treated as a validation set, and the method is fit on the remaining *k-1* folds. LOOCV is a special case of k-Fold, (k = n).

However, there is a bias/variance trade-off associated with the choice of k. If you have a large number of k (example, n), then each model output will be highly correlated to each other.

Typically, a good choice of k is 5 or 10.

#### Cross-Validation on Classification

Instead of MSE for error metric, we will use:

$CV_{(n)} = \frac{1}{n}\sum^{n}_{i=1}Err_i$

#### The Bootstrap

Bootstrapping is the process of resampling a data set (with replacement) to obtain confidence intervals (SE) for unknown population parameters.

### R lab

#### The Validation Set Approach

```{r}
set.seed(1)

train <- sample(392, 196)

lm.fit <- lm(mpg ~ horsepower, data = auto, subset = train)

mean(lm.fit$residuals^2) # train MSE

test <- auto[!train]
test$pred <- predict(lm.fit, newdata = test)
with(test, mean((mpg - pred)^2)) # test MSE

lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = auto, subset = train)
test$pred <- predict(lm.fit2, newdata = test) # update predictions
with(test, mean((mpg - pred)^2)) # model 2 MSE

lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = auto, subset = train)
test$pred <- predict(lm.fit3, newdata = test) # update predictions
with(test, mean((mpg - pred)^2)) # model 2 MSE

```

#### Leave-One-Out CV

```{r}
glm.fit <- glm(mpg ~ horsepower, data = auto)
coef(glm.fit)

glm.fit <- glm(mpg ~ horsepower, data = auto)
cv.err <- cv.glm(auto, glm.fit)

cv.err$delta

cv.error <- rep(0, 5)
for( i in 1:5 )
{
   glm.fit <- glm(mpg ~ poly(horsepower, i), data = auto)
   cv.error[i] <- cv.glm(auto, glm.fit)$delta[1]
}

cv.error
```

#### k-Fold Cross-Validation

```{r}
set.seed(7)

cv.error.10 <- rep(0, 10)
for( i in 1:10 )
{
   glm.fit <- glm(mpg ~ poly(horsepower, i), data = auto)
   cv.error.10[i] <- cv.glm(auto, glm.fit)$delta[1]
}

cv.error.10
```

#### The Bootstrap

```{r}
portfolio <- data.table(ISLR::Portfolio)

alpha.fn <- function(data, index ) {
   X <- data$X; Y <- data$Y
   return ((var(Y)-cov(X, Y)) / (var(X) + var(Y) - 2*cov(X, Y)))
}

alpha.fn(portfolio, 1:100)
```

```{r}
set.seed(1)

alpha.fn(portfolio, sample(100, 100, replace = T))

boot(data = portfolio, statistic = alpha.fn, R = 1000)
```

#### Estimating the Accuracy of a Linear Regression Model

```{r}
boot.fn <- function(data, index) {
   return (coef(lm(mpg ~ horsepower, data = data, subset = index)))
}

boot.fn(auto, 1:396)
```

```{r}
set.seed(1)

boot.fn(auto, sample(392, 392, replace = T))
```

```{r}
boot(auto, boot.fn, 1000)
summary(lm(mpg ~ horsepower, data = auto))
```

```{r}
boot.fn <- function(data, index){
   coefficients(lm(mpg ~ horsepower + I(horsepower^2), data = data), subset = index)
}

set.seed(1)

boot(data = auto, statistic = boot.fn, R = 1000)
```

## Conceptual

### 1.) Using the statistical properties of the variance as well as single variable calculus, derive 5.6. In other words, prove that $\alpha$ given by 5.6 does indeed minimize $Var(\alpha X + (1 - \alpha)Y$

$Var(\alpha X + (1 - \alpha)Y = \alpha^2\sigma^2_X + (1 - \alpha)^2\sigma^2_Y + 2\alpha(1 - \alpha)\sigma_{XY}$

### 2.) We now derive the probability that a given observation is part of a bootstrap sample. Suppose that we obtain a bootstrap sample from a set of *n* observations.

a.) What is the probability that the first bootstrap observation is *not* the *j*th observation from the original sample?

$1- \frac{1}{n}$

b.) What is the probability that the second bootstrap observation is *not* the *j*th observation from the original sample?

$1- \frac{1}{n}$

c.) Argue that the probability that the *j*th observation is not in the bootstrap sample is $(1- \frac{1}{n})^n$.

d.) When $n = 5$, what is the probability that the *j*th observation is in the bootstrap sample?

$1 - (1- \frac{1}{5})^5 = .672$

e.) When $n = 100$, what is the probability that the *j*th observation is in the bootstrap sample?

$1 - (1- \frac{1}{100})^100 = .634$

f.) When $n = 1000$, what is the probability that the *j*th observation is in the bootstrap sample?

$1 - (1- \frac{1}{1000})^{1000} = .632$

g.) Create a plot that displays, for each integer value of n from 1 to 100,000, the probability that the *j*th observation is in the bootstrap sample. Comment on what you observe.

```{r}
data <- data.table(index = 1:1e5)
data[, prob := 1 - (1 - 1/index)^index ]

ggplot(data, aes(index, prob)) +
   geom_point() +
   scale_x_continuous(labels = scales::comma)
```

Quickly approaches asymptote of .632

h.) We will now investigate the numerical probability that a bootstrap sample of size $n = 100$ contains the *j*th observation. Here j = 4. We repeatedly create bootstrap samples, and each time we record whether or not the fourth observation is contained in the bootstrap sample.

```{r}
store <- rep(NA, 1e5)

for(i in 1:1e5)
{
   store[i] <- sum(sample(1:100, rep = T) == 4) > 0 
}

mean(store)
```

### 3.) We now review k-fold cross-validation.

a.) Explain how k-fold cross validation is implemented:

_The data is split up into k (roughly equal) parts, k-1 of which is used to train the model that is then tested on the hold-out set._

b.) What are the advantages and disadvantages of k-fold cross-validation relative to:

i.) The validation set approach?

_The validation set approach has two main drawbacks compared to k-fold cross-validation. First, the validation estimate of the test error rate can be highly variable (depending on precisely which observations are included in the training set and which observations are included in the validation set). Second, only a subset of the observations are used to fit the model. Since statistical methods tend to perform worse when trained on fewer observations, this suggests that the validation set error rate may tend to overestimate the test error rate for the model fit on the entire data set._

ii.) LOOCV?

_k-Fold CV is computationally faster than LOOCV._

_LOOCV cross-validation approach may give approximately unbiased estimates of the test error, since each training set contains n−1 observations; however, this approach has higher variance than k-fold cross-validation (since we are averaging the outputs of n fitted models trained on an almost identical set of observations, these outputs are highly correlated, and the mean of highly correlated quantities has higher variance than less correlated ones). So, there is a bias-variance trade-off associated with the choice of k in k-fold cross-validation; typically using k=5 or k=10 yield test error rate estimates that suffer neither from excessively high bias nor from very high variance._

### 4.) Suppose that we use some sttatistical learning method to make a prediction for the response Y for a particular value of the predictor X. Carefully describe how we might estimate the standard deviation of our prediction.

_We may estimate the standard deviation of our prediction by using the bootstrap method. In this case, rather than obtaining new independant data sets from the population and fitting our model on those data sets, we instead obtain repeated random samples from the original data set. In this case, we perform sampling with replacement B times and then find the corresponding estimates and the standard deviation of those B estimates by using equation (5.8)._

## Applied

### 5.) In Chapter 4, we used logistic regression to predict the probability of default using income and balance on the default data set. We will now estimate the test error of this logistic regression model using the validation set approach.

```{r}

set.seed(1)

default <- data.table(ISLR::Default)

default.split <- initial_split(default, 
                               prop = .5, 
                               strata = default)

default.train <- training(default.split)
default.test <- testing(default.split)
```

a.) Fit a logistic regression model that uses income and balance to predict default.

```{r}
summary(fit1 <- glm(default ~ income + balance, family = binomial, data = default.train))
```

b.) Using the validation set approach, estimate the test error of this model.

```{r}
default.test$pred <- ifelse( predict(fit1, newdata = default.test, type = "response") > .5, "Yes", "No")

with(default.test, mean(default != pred))
```

c.) Repeat the process in b three times, using three different splits of the observations into a training set and a validation set.

```{r}
set.seed(1)

trials <- 3

results <- numeric(trials)

for( i in 1:trials )
{
   default.split <- initial_split(default, 
                                  prob = .5, 
                                  strata = default)

   default.train <- training(default.split)
   default.test <- testing(default.split)

   model <- glm(default ~ income + balance, family = binomial, data = default.train)

   pred <- ifelse( predict(model, newdata = default.test, type = "response") > .5, "Yes", "No")

   results[i] <- mean(default.test$default != pred)
}

results
```

The test error rates vary with sampling randomness.

d.) Now consider a logistic regression model that predicts the probability of default using income, balance and a dummy variable for student. Estimate the test error for this model using the validation set approach.

```{r}

set.seed(1)

default <- data.table(ISLR::Default)

# one-hot encode student
default[, ':='(student_Yes = ifelse(student == "Yes", 1, 0), student_No = ifelse(student == "No", 1, 0))][, student := NULL]

default.split <- initial_split(default, 
                               prob = .5, 
                               strata = default)

default.train <- training(default.split)
default.test <- testing(default.split)

summary(fit2 <- glm(default ~ income + balance + student_Yes, data = default.train, family = binomial))

default.test$pred <- ifelse(predict(fit2, newdata = default.test, type = "response") > .5, "Yes", "No")

with(default.test, mean(default != pred))
```

Student dummy coding slighly reduces the test error rate for the probability of default model.

#### 6.) We continue to consider the use of a logistic regression model to predict the probability of “default” using “income” and “balance” on the “Default” data set. In particular, we will now computes estimates for the standard errors of the “income” and “balance” logistic regression coefficients in two different ways : (1) using the bootstrap, and (2) using the standard formula for computing the standard errors in the glm() function. Do not forget to set a random seed before beginning your analysis.

a.) Using the summary() and glm() functions, determine the estimated standard errors for the coefficients associated with “income” and “balance” in a multiple logistic regression model that uses both predictors.

```{r}
set.seed(1)

summary(fit1 <- glm(default ~ income + balance, data = default.train, family = binomial))
```

b.) Write a function, boot.fn(), that takes as input the default data set as well as an index of the observations, and that outputs the coefficient estimates for income and balance in the multiple logistic regression model.

```{r}
boot.fn <- function(data, index) {
   coefficients(glm(default ~ income + balance, family = binomial, data = data, subset = index))
}

```

c.) Use boot() function together with your boot.fn() function to estimate the standard errors of the logistic regression coefficents for income and balance.

```{r}
boot(data = default, statistic = boot.fn, R = 1000)
```

#### 7.) We saw that cv.glm() function can be used in order to compute the LOOCV test error estimate. Alternatively, one could compute those quantities using just the glm() and predict.glm() functions, and a for loop. You will now take this approach in order to compute the LOOCY error for a simple logistic regresion model on the Weekly data set.

```{r}
set.seed(1)

weekly <- data.table(ISLR::Weekly)

weekly.split <- initial_split(weekly, prop = .5)

weekly.train <- training(weekly.split); weekly.test <- testing(weekly.split)
```

a.) Fit a logistic regression model that predicts Direction using the Lag1 and Lag2.

```{r}
summary(fit1 <- glm(Direction ~ Lag1 + Lag2, data = weekly.train, family = binomial))
```

b.) Fit a logistic regresion model that predicts direction using Lag1 and Lag2 using *all but the first observation.*

```{r}
holdout <- weekly[1]

summary(fit1 <- glm(Direction ~ Lag1 + Lag2, data = weekly[-1], family = binomial))
```

c.) Use the model from the previous to predict the first observation.

```{r}
holdout$Direction == ifelse( predict(fit1, newdata = holdout, type = "response") > .5, "Up", "Down")
```

The model did not predict this correctly.

d.) Write a for loop from i = 1 to i = n, where n is the number of observations in the data set, that performs each of the following steps:

i.) Fit a logistic regresion model using all but the *i*th observation.
ii.) Compute the posterior probability of the market moving up for the ith observation.

```{r}
n <- nrow(weekly)
results <- logical()
for(i in 1:n)
{
    model <- glm(Direction ~ Lag1 + Lag2, data = weekly[-i,], family = binomial)
    pred <- ifelse( predict(model, newdata = weekly[i,], type = "response") > .5, "Up", "Down")
    results[i] <- weekly[i,]$Direction != pred
}

```

iii.) Take the average of the n numbers obtained in d.) in order to obtain the LOOCV estimate for the test error:

```{r}
mean(results)
```

#### 8.)

We will now perform cross-validation on a simulated data set.

a.) Generate a simulated data set as follows:

```{r}
set.seed(1)
x <- rnorm(100); y <- x - 2 * x^2 + rnorm(100)

data <- data.table(x, y)
```

In this data set, what is n and what is p?

n = 100, p = 2, $Y = X - 2 * X^2 + \epsilon$

b.) Create a scatterplot of x, y

```{r}
plot(x, y)
```

c.) Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using the least squares:

i.) $Y = \beta_0 + \beta_1X + \epsilon$

```{r}
set.seed(1)
fit1 <- glm(y ~ x, data = data)
cv.glm(data, fit1)$delta[1]
```

ii.) $Y = \beta_0 + \beta_1X + \beta_2X^2 + \epsilon$

```{r}
set.seed(1)
fit1 <- glm(y ~ poly(x, 2), data = data)
cv.glm(data, fit1)$delta[1]
```

iii.) $Y = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + \epsilon$

```{r}
set.seed(1)
fit1 <- glm(y ~ poly(x, 3), data = data)
cv.glm(data, fit1)$delta[1]
```

iiii.) $Y = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + \beta_4X^4 \epsilon$

```{r}
set.seed(1)
fit1 <- glm(y ~ poly(x, 4), data = data)
cv.glm(data, fit1)$delta[1]
```

d.) Repeat c using another random seed, and report results.

i.) $Y = \beta_0 + \beta_1X + \epsilon$

```{r}
set.seed(100)
fit1 <- glm(y ~ x, data = data)
cv.glm(data, fit1)$delta[1]
```

ii.) $Y = \beta_0 + \beta_1X + \beta_2X^2 + \epsilon$

```{r}
set.seed(100)
fit1 <- glm(y ~ poly(x, 2), data = data)
cv.glm(data, fit1)$delta[1]
```

iii.) $Y = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + \epsilon$

```{r}
set.seed(100)
fit1 <- glm(y ~ poly(x, 3), data = data)
cv.glm(data, fit1)$delta[1]
```

iiii.) $Y = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + \beta_4X^4 \epsilon$

```{r}
set.seed(100)
fit1 <- glm(y ~ poly(x, 4), data = data)
cv.glm(data, fit1)$delta[1]
```

e.) What model had the smallest LOOCV ?

_The second degree polynomial had the best fit, which is to be expected given that the underlying function is polynomial._

f.) Comment on the statistical significance of the coefficient estimates that result from fitting the models.

```{r}
summary(fit1)
```

_We can see that the 1 and 2 term coefficents are statistically significant, which agree to our CV results._

#### 9.) We will now consider the Boston housing data set, from the MASS library.

a.) Based on this data set, provide an estimate for the population mean of medv. Call this estimate $\hat{\mu}$

```{r}
boston <- data.table(MASS::Boston)

mu.hat <- mean(boston$medv)
```

b.) Provide an estimate of the standard error of $\hat{\mu}$

```{r}
mu.se <- sd(boston$medv) / sqrt(nrow(boston))
mu.se
```

c.) Now estimate the standard error of $\hat{\mu}$ using the bootstrap.

```{r}
boot.fn <- function(data, index) {
   mean(data[index]$medv)
}

t.boot <- boot(boston, boot.fn, R = 1000)

```

d.) Based on your bootstrap estimate from (c), provide a 95% confidence interval for the mean of medv.

```{r}
alpha <- .95

t.boot$t0 + qnorm(c(1 - alpha/2, alpha/2)) * 0.411978
```

Compare to the t-test:

```{r}
t.test(boston$medv)
```

e.) Based on this data set, provide an estimate $\hat{\mu}_{med}$, for the median value of medv in the population.

```{r}
med.hat <- median(boston$medv)
```

f.) We now would like to estimate the standard error of $\hat{\mu}_{med}$. Unfortunately, there is no simple formula for computing the standard error of the median.

Instead, estimate the standard error using the bootstrap.

```{r}
boot.fn <- function(data, index) {
   median(data[index]$medv)
}

t.boot <- boot(boston, boot.fn, R = 1000)
t.boot
```

g.) Based on this data set, provide an estimate for the tenth percentile of medv in the Boston suburbs. Call this quantity $\hat{\mu}_{0.1}$

```{r}
quantile(boston$medv, .1)
```

h.) Use the bootstrap to estimate the standard error of $\hat{\mu}_{0.1}$

```{r}
boot.fn <- function(data, index) {
   quantile(data[index]$medv, .1)
}

t.boot <- boot(boston, boot.fn, R = 1000)
t.boot
```

### 2020-03-01 Statistical Learning/07_Beyond-Linearity.rmd
sln clean-up.
sln clean-up.

---
title: ''
mainfont: Arial
fontsize: 12pt
fig_width: 9
fig_height: 3.5
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistical Learning}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 7}
   \fancyfoot[C]{\rmfamily\color{headergrey}Beyond Linearity}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

# Data Wrangling

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(car, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(ISLR, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(class, quietly = TRUE, warn.conflicts = FALSE)
library(boot, quietly = TRUE, warn.conflicts = FALSE)
library(leaps, quietly = TRUE, warn.conflicts = FALSE)
library(glmnet, quietly = TRUE, warn.conflicts = FALSE)
library(pls, quietly = TRUE, warn.conflicts = FALSE)
library(splines, quietly = TRUE, warn.conflicts = FALSE)
library(gam, quietly = TRUE, warn.conflicts = FALSE)
library(akima, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets/")

select <- dplyr::select

```

## Chapter 7

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Lab

### Polynomial Functions and Cut Points

```{r}
load(paste0(here::here(), "/ISLR/7.R.RData"))

plot(x, y)

fit <- lm(y ~ x)
fit2 <- lm(y ~ 1 + x + I(x^2))
```

```{r}
wage <- data.table(ISLR::Wage)
```

### Polynomial Regression and Step Functions

```{r}
fit <- lm(wage ~ poly(age, 4), data = wage)

summary(fit)

plot(fit)

coef(summary(fit))
```

```{r}
fit2 <- lm(wage ~ poly(age, 4, raw = T), data = wage)
coef(summary(fit2))
```

Alternative:

```{r}
fit2a <- lm(wage ~ age + I(age^2) + I(age^3) + I(age^4), data = wage)
coef(summary(fit2a))
```

```{r}
fit2b <- lm(wage ~ cbind(age, age^2, age^3, age^4), data = wage)
coef(fit2b)
```

```{r}
agelims <- range(wage$age)
age.grid <- seq(from = agelims[1], to = agelims[2])

pred <- predict(fit, newdata = list(age = age.grid), se = T)

se.bands <- cbind(pred$fit + 2*pred$se.fit, pred$fit - 2*pred$se.fit)

par(mfrow = c(1, 1), mar = c(4.5, 4.5, 1, 1), oma = c(0, 0, 4, 0))
plot(wage$age, wage$wage, xlim = agelims, cex = .5, col = "darkgrey")
title("Degree-4 Polynomial", outer = T)
lines(age.grid, pred$fit, lwd = 2, col = "blue")
matlines(age.grid, se.bands, lwd = 1, col = "blue", lty = 3)
```

```{r}
pred2 <- predict(fit2, newdata = list(age = age.grid), se = T)
max(abs(pred$fit - pred2$fit))
```

```{r}
fit1 <- lm(wage ~ age, data = wage)
fit2 <- lm(wage ~ poly(age, 2), data = wage)
fit3 <- lm(wage ~ poly(age, 3), data = wage)
fit4 <- lm(wage ~ poly(age, 4), data = wage)
fit5 <- lm(wage ~ poly(age, 5), data = wage)

anova(fit1, fit2, fit3, fit4, fit5)

coef(summary(fit5))
```

```{r}
fit1 <- lm(wage ~ education + age, data = wage)
fit2 <- lm(wage ~ education + poly(age, 2), data = wage)
fit3 <- lm(wage ~ education + poly(age, 3), data = wage)

anova(fit1, fit2, fit3)
```

```{r}
fit <- glm(I(wage > 250) ~ poly(age, 4), data = wage, family = "binomial")

pred <- predict(fit, newdata = list(age = age.grid), se = T)

pfit <- exp(pred$fit) / (1 + exp(pred$fit))

se.bands.logit <- cbind(pred$fit + 2 * pred$se.fit, pred$fit - 2*pred$se.fit)

se.bands <- exp(se.bands.logit) / (1 + exp(se.bands.logit))

```

Alternatively:

```{r}
pred <- predict(fit, newdata = list(age = age.grid), type = "response", se = T)

with(wage, {
   plot(age, I(wage > 250), xlim = agelims, type = "n")
   points(jitter(age), I((wage > 250)/5), cex = .5, pch = "|", col = "darkgrey")
   lines(age.grid, pfit, lwd = 2, col = "blue")
   matlines(age.grid, se.bands, lwd = 1, col = "blue", lty = 3)
})
```

```{r}
table(cut(wage$age, 4))
```

```{r}
fit <- lm(wage ~ cut(age, 4), data = wage)
coef(summary(fit))
```

### Splines

```{r}
fit <- lm(wage ~ bs(age, knots = c(25, 40, 60)), data = wage)

pred <- predict(fit, newdata = list(age = age.grid), se = T)

with(wage, {
   plot(age, wage, col = "gray")
   lines(age.grid, pred$fit, lwd=2)
   lines(age.grid, pred$fit+2*pred$se.fit, lty="dashed")
   lines(age.grid, pred$fit-2*pred$se.fit, lty="dashed")
})
```

```{r}
dim(bs(wage$age, knots = c(25, 40, 60)))
dim(bs(wage$age, df = 6))
attr(bs(wage$age, df = 6), "knots")
```

```{r}
fit2 <- lm(wage ~ ns(age, df = 4), data = wage)
pred2 <- predict(fit2, newdata =  list(age = age.grid), se = T)
par(mfrow=c(1,1))
plot(wage$age, wage$wage, col = "gray")
lines(age.grid, pred2$fit, col = "red", lwd = 2)
```

```{r}
with(wage,{
   plot(age, wage, xlim = agelims, cex = .5, col = "darkgrey")
   title("Smoothing Spline")
   fit <- smooth.spline(age, wage, df = 16)
   fit2 <- smooth.spline(age, wage, cv = T)

   lines(fit, col = "red", lwd = 2)
   lines(fit2, col = "blue", lwd = 2)
})
```

```{r}
with(wage, {
   plot(age, wage, xlim = agelims, cex = .5, col = "darkgrey")
   title("Local Regression")
   fit <- loess(wage ~ age, span = .2)
   fit2 <- loess(wage ~ age, span = .5)
   lines(age.grid, predict(fit, data.frame(age = age.grid)), col = "red", lwd = 2)
   lines(age.grid, predict(fit2, data.frame(age = age.grid)), col = "blue", lwd = 2)
   legend("topright", legend = c("Span = 0.2", "Span = 0.5"), col = c("red", "blue"), lty = 1, lwd = 2, cex = .8)
})
```

### GAMs

```{r}
gam1 <- lm(wage ~ ns(year, 4) + ns(age, 5) + education, data = wage)

gam.m3 <- gam(wage ~ s(year, 4) + s(age, 5) + education, data = wage)
par(mfrow = c(1, 3))
plot.Gam(gam.m3, se = T)
```

```{r}
par(mfrow = c(1, 3))
plot(gam.m3, se = T, col = "blue")

plot.Gam(gam1, se = T, col = "red")
```

```{r}
gam.m1 <- gam(wage ~ s(age, 5) + education, data = wage)
gam.m2 <- gam(wage ~ year + s(age, 5) + education, data = wage)

anova(gam.m1, gam.m2, gam.m3)
```

```{r}
summary(gam.m3)
```

```{r}
pred <- predict(gam.m2, newdata = wage)

gam.lo <- gam(wage ~ s(year, df = 4) + lo(age, span = 0.7) + education, data = wage)
plot.Gam(gam.lo, se = T, col = "green")
```

```{r}
plot(gam.lo)
```

```{r}
gam.lr <- gam(I(wage > 250) ~ year + s(age, df = 5) + education, family = binomial, data = wage)
par(mfrow = c(1, 3))
plot(gam.lr, se = T, col = "green")
```

```{r}
table(wage$education, I(wage$wage > 250))
```

```{r}
levels(wage$education)

gam.lr.s <- gam(I(wage > 250) ~ year + s(age, df = 5) + education, family = binomial, data = wage, subset = (education != "1. < HS Grad"))
plot(gam.lr.s, se = T, col = "green")
```

### Applied

In this exercise, you will further analyze the wage data set considered throughout this chapter.

```{r}
test.size <- .7
index <- sample(nrow(wage), nrow(wage) * test.size, replace = F)

train <- wage[index]
test <- wage[!index]

```

a.) Perform polynomial regression to predict wage using age. Use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the result of hypothesis testing using ANOVA? Make a plot of the fit obtained.

```{r}
degree <- 20; folds = 10
cv.errors <- numeric(degree)

fold.size <- nrow(train) / folds

for(deg in 1:degree)
{
   # 10 fold cv
   errors <- numeric(folds)
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)

      cv.train <- train[!holdout]
      cv.test <- train[holdout]

      fit <- lm(wage ~ poly(age, deg), data = cv.train)

      pred <- predict(fit, newdata = cv.test, type = "response")

      errors[fold] <- sqrt(mean((cv.test$wage - pred)^2))
   }
   cv.errors[deg] <- mean(errors)
}

lowest.error <- which.min(cv.errors)

cv.results <- data.table(degree = 1:degree, error = cv.errors)[, lowest := degree == lowest.error]

ggplot(cv.results, aes(degree, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Degree")
```

```{r}
model <- lm(wage ~ poly(age, lowest.error), data = train)

test %>%
    mutate(predictions = predict(model, test)) %>%
    ggplot(aes(age, wage, col = 'darkgrey')) +
    geom_point(alpha = .65) +
    geom_line(aes(age, predictions, col = 'cornflowerblue'), size = 1.5) +
    scale_color_manual(name = 'Value Type',
                         labels = c('Observed', 'Predicted'),
                         values = c('cornflowerblue', 'darkgrey' )) +
    labs(x = 'Age', y = 'Wage', 
         title = paste0('Predictions from polynomial model of degree ', lowest.error))
```

b.) Fit a step function to predict wage using age, and perform cross-validation to choose the optimal number of cuts. Make a plot of the fit obtained.

```{r}
cuts <- 20; folds = 10
cv.errors <- numeric(degree)

fold.size <- nrow(train) / folds

for(cuts in 2:cuts)
{
   # 10 fold cv
   errors <- numeric(folds)

   # apply cut here so CV train/test have same levels
   train$AgeGroup <- cut(train$age, cuts)

   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)

      cv.train <- train[!holdout]
      cv.test <- train[holdout]

      fit <- lm(wage ~ I(AgeGroup), data = cv.train)

      pred <- predict(fit, newdata = cv.test, type = "response")

      errors[fold] <- sqrt(mean((cv.test$wage - pred)^2))
   }

   cv.errors[cuts] <- mean(errors)
}

lowest.error <- which.min(cv.errors[cv.errors != 0])

cv.results <- data.table(cuts = 1:cuts, error = cv.errors)[, lowest := cuts == lowest.error]

ggplot(cv.results, aes(cuts, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Age Group")
```

```{r}
wage.grouped <- wage
wage.grouped$AgeGroup <- cut(wage.grouped$age, lowest.error)

test.size <- .7
index <- sample(nrow(wage), nrow(wage) * test.size, replace = F)

train <- wage.grouped[index]
test <- wage.grouped[!index]

model <- lm(wage ~ I(AgeGroup), data = train)

test %>%
    mutate(predictions = predict(model, test)) %>%
    ggplot(aes(age, wage, col = 'darkgrey')) +
    geom_point(alpha = .65) +
    geom_line(aes(age, predictions, col = 'cornflowerblue'), size = 1.5) +
    scale_color_manual(name = 'Value Type',
                         labels = c('Observed', 'Predicted'),
                         values = c('cornflowerblue', 'darkgrey' )) +
    labs(x = 'Age', y = 'Wage', 
         title = paste0('Predictions from polynomial model of age group ', lowest.error))
```

The wage data set contains a number of other features not explored in this chapter, such as marital status (*marit1*), job class (*jobclass*), and others. Explore the relationships between some of these other predictors and wage, and use non-linear fitting techniques in order to fit flexible models to the data. Create plots of the results obtained, and write a summary of your findings.

```{r}
head(wage)

ggplot(wage, aes(wage, group = race, fill = race)) +
   geom_density(alpha = .65)

ggplot(wage, aes(race, wage, fill = race)) +
   geom_boxplot()

ggplot(wage, aes(age, wage)) +
   geom_point() +
   geom_smooth() +
   facet_wrap(~race)

ggplot(wage, aes(maritl, wage, fill = maritl)) +
   geom_boxplot()

ggplot(wage, aes(wage, fill = maritl)) +
   geom_density(alpha = .65)

ggplot(wage, aes(age, wage)) +
   geom_point() +
   geom_smooth() +
   facet_wrap(~maritl)

ggplot(wage, aes(jobclass, wage, fill = jobclass)) +
   geom_boxplot()

```

Fit some non-linear models investigated in this chapter to the **Auto** data set. Is there evidence for non-linear relationships in this data set? Create some informative plots to justify your answer.

```{r}
auto <- as.data.table(ISLR::Auto)

auto$cylinders <- as.factor(auto$cylinders)

head(auto)

ggplot(auto, aes(mpg, group = cylinders, fill = cylinders)) +
   geom_density(alpha = .65)

ggplot(auto, aes(y = mpg, group = cylinders, fill = cylinders)) +
   geom_boxplot()

ggplot(auto, aes(acceleration, mpg, col = cylinders)) +
   geom_point()

ggplot(auto[cylinders %in% c(4, 6, 8)], aes(acceleration, mpg)) +
   geom_point() +
   geom_smooth() +
   facet_wrap(~cylinders)

ggplot(auto[cylinders %in% c(4, 6, 8)], aes(horsepower, mpg)) +
   geom_point() +
   geom_smooth() +
   facet_wrap(~cylinders)

ggplot(auto, aes(displacement, mpg, col = cylinders)) +
   geom_point()

ggplot(auto[cylinders %in% c(4, 6, 8)], aes(displacement, mpg, col = cylinders)) +
   geom_point() +
   facet_wrap(~cylinders)

```

```{r}
# suggests quadratic and quintic are better than linear

with(auto, {
   fit1 <- lm(mpg ~ horsepower + cylinders)
   fit2 <- lm(mpg ~ poly(horsepower, 2) + cylinders)
   fit3 <- lm(mpg ~ poly(horsepower, 3) + cylinders)
   fit4 <- lm(mpg ~ poly(horsepower, 4) + cylinders)
   fit5 <- lm(mpg ~ poly(horsepower, 5) + cylinders)

   anova(fit1, fit2, fit3, fit4, fit5)
})

```

```{r}
n <- nrow(auto)

index <- sample(n, n * .7, replace = F)

train <- auto[index]
test <- auto[!index]

n; nrow(train); nrow(test)

degree <- 10; folds = 10
cv.errors <- numeric(degree)

fold.size <- nrow(train) / folds

for(deg in 1:degree)
{
   # 10 fold cv
   errors <- numeric(folds)
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)

      cv.train <- train[!holdout]
      cv.test <- train[holdout]

      fit <- lm(mpg ~ poly(horsepower, deg) + cylinders, data = cv.train)

      pred <- predict(fit, newdata = cv.test, type = "response")

      errors[fold] <- sqrt(mean((cv.test$mpg - pred)^2))
   }
   cv.errors[deg] <- mean(errors)
}

lowest.error <- which.min(cv.errors)

cv.results <- data.table(degree = 1:degree, error = cv.errors)[, lowest := degree == lowest.error]

ggplot(cv.results, aes(degree, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Degree")
```

```{r}
model <- lm(mpg ~ poly(horsepower, lowest.error), data = train)

test %>%
    mutate(predictions = predict(model, test)) %>%
    ggplot(aes(horsepower, mpg, col = 'darkgrey')) +
    geom_point(alpha = .65) +
    geom_line(aes(horsepower, predictions, col = 'cornflowerblue'), size = 1.5) +
    scale_color_manual(name = 'Value Type',
                         labels = c('Observed', 'Predicted'),
                         values = c('cornflowerblue', 'darkgrey' )) +
    labs(x = 'hoursepower^2', y = 'MPG', 
         title = paste0('Predictions from polynomial model of degree ', lowest.error))
```

```{r}
for(deg in 1:degree)
{
   # 10 fold cv
   errors <- numeric(folds)
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)

      cv.train <- train[!holdout]
      cv.test <- train[holdout]

      fit <- lm(mpg ~ poly(displacement, deg) + cylinders, data = cv.train)

      pred <- predict(fit, newdata = cv.test, type = "response")

      errors[fold] <- sqrt(mean((cv.test$mpg - pred)^2))
   }
   cv.errors[deg] <- mean(errors)
}

lowest.error <- which.min(cv.errors)

cv.results <- data.table(degree = 1:degree, error = cv.errors)[, lowest := degree == lowest.error]

ggplot(cv.results, aes(degree, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Degree")
```

```{r}
model <- lm(mpg ~ poly(displacement, lowest.error), data = train)

test %>%
    mutate(predictions = predict(model, test)) %>%
    ggplot(aes(displacement, mpg, col = 'darkgrey')) +
    geom_point(alpha = .65) +
    geom_line(aes(displacement, predictions, col = 'cornflowerblue'), size = 1.5) +
    scale_color_manual(name = 'Value Type',
                         labels = c('Observed', 'Predicted'),
                         values = c('cornflowerblue', 'darkgrey' )) +
    labs(x = 'displacement', y = 'MPG', 
         title = paste0('Predictions from polynomial model of degree ', lowest.error))
```

```{r}
polyfit <- gam(mpg ~ ns(horsepower, 2) + ns(displacement, 4) + cylinders, data = auto)
par(mfrow = c(1, 3))
plot(polyfit, se = T, col = "cornflowerblue")
```

```{r}
pred <- data.table(actual = test$mpg, pred = predict(polyfit, test, type = "response"))

ggplot(pred, aes(actual, pred)) +
   geom_point() +
   geom_smooth()
```

### Boston

This question uses the variables dis (the weighed mean of distances to five Boston employment centers) and nox (nitrogen oxides concendrated in parts per 10 million) from the boston data set. We will treat dis as the predictor and nox as the response.

```{r}
boston <- data.table(Boston)
```

a.) Use the poly() function to fit a cubic polynomial regression to predict nox using dis. Report the regression output, and plot the resulting data using polynomial fits.

```{r}
fit1 <- lm(nox ~ poly(dis, 3), data = boston)
summary(fit)
```

```{r}
ggplot(boston, aes(dis, nox)) +
   geom_point() +
   geom_smooth()
```

Plot the polynomial fits for a range of different polynomial degrees (say, 1 to 10), and report the associated RSS.

```{r}
p1 <- ggplot(boston, aes(dis, nox)) +
   geom_point() +
   geom_smooth(method = lm, formula = y ~ splines::bs(x, 3)) +
   labs(title = "x^3")

p2 <- ggplot(boston, aes(dis, nox)) +
   geom_point() +
   geom_smooth(method = lm, formula = y ~ splines::bs(x, 4)) +
   labs(title = "x^4")

p3 <- ggplot(boston, aes(dis, nox)) +
   geom_point() +
   geom_smooth(method = lm, formula = y ~ splines::bs(x, 5)) +
   labs(title = "x^5")

grid.arrange(p1, p2, p3, nrow = 1)

p4 <- ggplot(boston, aes(dis, nox)) +
   geom_point() +
   geom_smooth(method = lm, formula = y ~ splines::bs(x, 6)) +
   labs(title = "x^6")

p5 <- ggplot(boston, aes(dis, nox)) +
   geom_point() +
   geom_smooth(method = lm, formula = y ~ splines::bs(x, 7)) +
   labs(title = "x^7")

p6 <- ggplot(boston, aes(dis, nox)) +
   geom_point() +
   geom_smooth(method = lm, formula = y ~ splines::bs(x, 8)) +
   labs(title = "x^8")

grid.arrange(p4, p5, p6, nrow = 1)

```

Perform cross-validation or another approach to select the optimal degree for the polynomial, and explain your results.

```{r}
n <- nrow(boston)

index <- sample(n, n * .7, replace = F)

train <- boston[index]
test <- boston[!index]

n; nrow(train); nrow(test)

degree <- 10; folds = 10
cv.errors <- numeric(degree)

fold.size <- nrow(train) / folds

for(deg in 1:degree)
{
   # 10 fold cv
   errors <- numeric(folds)
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)

      cv.train <- train[!holdout]
      cv.test <- train[holdout]

      fit <- lm(nox ~ poly(dis, deg), data = cv.train)

      pred <- predict(fit, newdata = cv.test, type = "response")

      errors[fold] <- sum( (cv.test$nox - pred)^2 ) # RSS
   }
   cv.errors[deg] <- mean(errors)
}

lowest.error <- which.min(cv.errors)

cv.results <- data.table(degree = 1:degree, error = cv.errors)[, lowest := degree == lowest.error]

ggplot(cv.results, aes(degree, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Degree")
```

Use the *bs()* function to fit a regression spline to predict nox using dis. Report the output for the fit using for degrees of freedom. How did you choose the knots? Plot the fit.

```{r}
n <- nrow(boston)

index <- sample(n, n * .7, replace = F)

train <- boston[index]
test <- boston[!index]

n; nrow(train); nrow(test)

range(boston$dis)

knots <- seq(1, 15, 1); folds = 10
cv.errors <- numeric(length(knots))

fold.size <- nrow(train) / folds

index <- 1

for(index in 1:length(knots))
{
   # 10 fold cv
   errors <- numeric(folds)
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)

      cv.train <- train[!holdout]
      cv.test <- train[holdout]

      fit <- lm(nox ~ bs(dis, knots[index]), data = cv.train)

      pred <- predict(fit, newdata = cv.test, type = "response")

      errors[fold] <- sqrt(mean( (cv.test$nox - pred)^2 )) # RMSE
   }
   cv.errors[index] <- mean(errors)
}

lowest.error <- knots[which.min(cv.errors)]

cv.results <- data.table(knots = knots, error = cv.errors)[, lowest := knots == lowest.error]

ggplot(cv.results, aes(knots, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by Knots")
```

Use CV to select the best degrees of freedom for a regression spline.

```{r}
n <- nrow(boston)

index <- sample(n, n * .7, replace = F)

train <- boston[index]
test <- boston[!index]

n; nrow(train); nrow(test)

range(boston$dis)

df <- seq(1, 15, 1); folds = 10
cv.errors <- numeric(length(knots))

fold.size <- nrow(train) / folds

index <- 1

for(index in 1:length(df))
{
   # 10 fold cv
   errors <- numeric(folds)
   for(fold in 1:folds)
   {
      holdout <- seq((fold - 1) * fold.size, fold * fold.size)

      cv.train <- train[!holdout]
      cv.test <- train[holdout]

      fit <- lm(nox ~ ns(dis, df[index]), data = cv.train)

      pred <- predict(fit, newdata = cv.test, type = "response")

      errors[fold] <- sqrt(mean( (cv.test$nox - pred)^2 )) # RMSE
   }
   cv.errors[index] <- mean(errors)
}

lowest.error <- knots[which.min(cv.errors)]

cv.results <- data.table(df = df, error = cv.errors)[, lowest := df == lowest.error]

ggplot(cv.results, aes(knots, error, fill = lowest)) +
   geom_bar(stat = "identity") +
   labs(title = "RMSE by df")
```

### College

This question uses the college data set.

```{r}
college <- data.table(ISLR::College)
```

a.) Split the data into a training set and a test set. Using out-of-state tuition as the response and the other variables as the predictors, perform forward stepwise selection on the training data in order to identify a satisfactory model that uses just a subset of the predictors.

```{r}
n <- nrow(college)

index <- sample(n, n * .7)

train <- college[index]
test <- college[!index]

null <- lm(Outstate ~ 1, data = train)
full <- formula(lm(Outstate ~ ., data = train))

forward.fit <- step(null, direction = 'forward', scope = full)

summary(forward.fit)

best.model <- formula(forward.fit)

coef(forward.fit)
```

Fit a GAM on the training data, using out-of-state tuition as the response and the features selected in the previous step as the predictors.

```{r}
fit <- gam(Outstate ~ Private + s(Room.Board, df = 2) + s(PhD, df = 2) + s(perc.alumni, df = 2) + s(Expend, df = 5) + s(Grad.Rate, df = 2), data=train)
par(mfrow = c(2, 3))
plot.Gam(fit, se = T, col = "blue")
```

### 2020-03-02 Hands-on Machine Learning/16_Interpretable-Machine-Learning.rmd
frm ch1
frm ch1

--- 2020-02-19 HOML/15_Stacked-Models.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Stacked Models}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Interpretable Machine Learning}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 15}
+   \lfoot{\color{headergrey}Chapter 16}
    \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
@@
 library(forecast, quietly = TRUE, warn.conflicts = FALSE)
-library(vip, quietly = TRUE, warn.conflicts = FALSE)
 library(pdp, quietly = TRUE, warn.conflicts = FALSE)
@@
 library(mlbench, quietly = TRUE, warn.conflicts = FALSE)
+
+library(vip, quietly = TRUE, warn.conflicts = FALSE)
+library(pdp, quietly = TRUE, warn.conflicts = FALSE)
+library(iml, quietly = TRUE, warn.conflicts = FALSE)
+library(DALEX, quietly = TRUE, warn.conflicts = FALSE)
+library(lime, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-## Stacked Models
+## Interpretable Machine Learning
 

### 2020-03-02 Mathematics and Statistics for FRM/01_Basics.rmd
frm ch1
frm ch1

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Mathematics and Statistics for Financial Risk Management}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/datasets")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 1

### 1

Solve for $y$, where:

a.) $y = ln(e^5)$

5

b.) $y = ln(1/e)$

$y = ln(1/e) = ln(1) - 1 = 0 - 1 = -1$

c.) $y = ln(10e)$

$y = ln(10e) = ln(10) + ln(e) = ln(10) + 1$

```{r}
log(10) + 1
```

### 2

The nominal monthly rate for a loan is quoted at 5%. What is the equivalent annual rate? Semiannual rate? Continious rate?

```{r}
effective_rate <- function(rate, periods) {
   ((1 + rate/periods)^periods - 1)
}
```

Annual:

```{r}
effective_rate(.05, 12) * 100
```

Semiannual:

```{r}
effective_rate(.05, 2) * 100
```

Continous:

```{r}
log(1 + effective_rate(.05, 12)) * 100
```

### 3

Over the course of a year, the log return on a stock market index is 11.2%. The starting value of the index is 100. What is the value at the end of the year?

```{r}
100 * exp(.112)
```

### 4

You have a portfolio of 10 bonds. In how many different ways can exactly two bonds default? Assume the order in which the bonds default is unimportant.

```{r}
choose(10, 2)
```

### 5

What is the present value of a perpetuity that pays $100 per year? Use an actual discount rate of 4%, and assume the first payment will be made in exactly one year.

```{r}
100 * (1/0.04)
```

### 6

ABC stock will pay a $1 dividend in one year. Assume the dividend will continue to be paid annually forever and the dividend payments will increase in size at a rate of 5%. Value this stream of dividends using a 6% annual discount rate.

```{r}
1 / (0.06 - 0.05)
```

### 7

What is the present value of a 10-year bond with a $100 face value, which pays a 6% coupon annually? Use an 8% discount rate.

```{r}

present_value <- function(face, int.rate, periods, discount.rate) {
   delta <- 1/(1 + discount.rate)

   coupon <- face * int.rate

   discount <- (delta - delta^(periods+1)) / (1 - delta)

   coupon * discount + ( face / (1 + discount.rate)^periods )
}

present_value(100, .06, 10, .08)
```

### 8

Solve for $x$, where $e^{e^x} = 10$

```{r}
log(log(10))
```

### 9

Calculate the value of the following summation: $\sum^{9}_{i=0}(-0.5)^i$

```{r}
r <- -0.5; n <- 9

(1 - r^n) / (1 - r)
```

### 10

The risk department of your firm has 10 analysts. You need to select four analysts to serve on a special audit committee. How many possible groupings of four analysts can be put together?

```{r}
choose(10, 4)
```

long-hand:

```{r}
n <- 10; k <- 4

factorial(10) / ( factorial(k) * factorial(n - k))
```

### 11

What is the present value of a newly issued 10-year bond with a notial value of $100 and a 2% annual coupon? Assume a constant 5% annual discount rate and no risk of default.

```{r}
present_value(100, .02, 10, 0.05)
```

### 2020-03-03 Hands-on Machine Learning/16_Interpretable-Machine-Learning.rmd
iml
iml

@@
 
-```{r, echo = T}
+```{r, attrition_data}
 attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)
@@
 
-```{r, echo = T}
+```{r, ames_data}
+
 ames <- AmesHousing::make_ames()
@@
 
-```{r}
+```{r, data_recipe}
+
 # ensure consistent categorical levels
@@
 
-### Stacking Overview
+### h2o ML setup
 
-Stacking is the process of combining multiple "base" learners (RF, GBM, GLM, etc.) into a "Super Learner" that uses the individual learners to make the final prediction.
+```{r, train_h2o_models}
 
-The Super Learner Algorithm
-
-1.) Setup the Ensemble
-
-  Specify a list of L base learners (with a specific set of model parameters)
-
-  Specify a meta learning algorithm. This can be any one of the algorithms discussed in the previous chapters, but most often is some form of regularized regression.
-
-2.) Train the ensemble
-
-  Train each of the L base learners on the training set.
-
-  Perform k-fold CV on each of the base learners and collect the cross-validated predictions from each (the same k-fold must be used for each base learner).
-These predicted values represent $p_1, ..., p_l$.
-
-  The N cross-validated predicted values from each of the L algorithms can be combined to form a new N X L feature matrix (Z)
-
-  Train the meta learning algorithm on level-one data (y = f(Z)).
-The "ensemble model" consista of the L base learning models and the meta learning model, which can then be used to generate predictions on new data.
-
-3.) Predict on new data.
-
-  To generate ensemble predictions, first generate predictions from the base learners.
-
-  Feed those predictions into the meta learner to generate the ensemble prediction.
-
-### Stacking Existing Models
-
-Restrictions:
-
-1.) All models must be trained on the same training set
-
-2.) All models must be trained with the same number of CV folds.
-
-3.) All models must use the same fold assignment to ensure the same observations are used.
-
-4.) The cross-validated predictions from all of the models must be preserved by setting keep_cross_validation_predictions = T.
-
-```{r}
 # Train & cross-validate a GLM model
@@
   fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
-  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
+  seed = 123, stopping_rounds = 5, stopping_metric = "RMSE",
   stopping_tolerance = 0
@@
   max_depth = 7, min_rows = 5, sample_rate = 0.8, nfolds = 10,
-  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,es
-
-  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
+  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
+  seed = 123, stopping_rounds = 5, stopping_metric = "RMSE",
   stopping_tolerance = 0
@@
 # Train & cross-validate an XGBoost model
-best_xgb <- h2o.xgboost(
-  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
-  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
-  nfolds = 10, fold_assignment = "Modulo", 
-  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
-  stopping_metric = "RMSE", stopping_tolerance = 0
-)
-```
 
-Train the ensemble
+# not avaliable on windows, yet..(3/2/20)
 
-```{r}
+#best_xgb <- h2o.xgboost(
+#  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
+#  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
+#  nfolds = 10, fold_assignment = "Modulo", 
+#  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
+#  stopping_metric = "RMSE", stopping_tolerance = 0
+#)
+
 # Train a stacked tree ensemble
 ensemble_tree <- h2o.stackedEnsemble(
-  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble",
-  base_models = list(best_glm, best_rf, best_gbm, best_xgb),
+  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble_01",
+  base_models = list(best_glm, best_rf, best_gbm),
   metalearner_algorithm = "drf"
@@
 
-```{r}
-get_rmse <- function(model) {
-  results <- h2o.performance(model, newdata = test_h2o)
-  results@metrics$RMSE
+### Local Interpretation
+
+Predictions using ensemble tree:
+
+```{r, local_predictions}
+
+# Compute predictions
+predictions <- predict(ensemble_tree, train_h2o) %>% as.vector()
+
+# Print the highest and lowest predicted sales price
+paste("Observation", which.max(predictions), 
+      "has a predicted sale price of", scales::dollar(max(predictions)))
+
+paste("Observation", which.min(predictions), 
+      "has a predicted sale price of", scales::dollar(min(predictions)))
+
+# Grab feature values for observations with min/max predicted sales price
+high_ob <- as.data.frame(train_h2o)[which.max(predictions), ] %>% select(-Sale_Price)
+low_ob  <- as.data.frame(train_h2o)[which.min(predictions), ] %>% select(-Sale_Price)
+```
+
+### Model specific vs. model agnostic
+
+```{r, model_features}
+
+# 1) create a data frame with just the features
+features <- as.data.frame(train_h2o) %>% select(-Sale_Price)
+
+# 2) Create a vector with the actual responses
+response <- as.data.frame(train_h2o) %>% pull(Sale_Price)
+
+# 3) Create custom predict function that returns the predicted values as a vector
+pred <- function(object, newdata)  {
+  results <- as.vector(h2o.predict(object, as.h2o(newdata)))
+  return(results)
 }
 
-list(best_glm, best_rf, best_glm, best_xgb) %>%
-  purrr::map_dbl(get_rmse)
+# Example of prediction output
+pred(ensemble_tree, features) %>% head()
 ```
 
-Stacked Results
-
-```{r}
-h2o.performance(ensemble_tree, newdata = test_h2o)@metrics$RMSE
-```
-
-```{r}
-data.frame(
-  GLM_pred = as.vector(h2o.getFrame(best_glm@model$cross_validation_holdout_predictions_frame_id$name)),
-  RF_pred = as.vector(h2o.getFrame(best_rf@model$cross_validation_holdout_predictions_frame_id$name)),
-  GBM_pred = as.vector(h2o.getFrame(best_gbm@model$cross_validation_holdout_predictions_frame_id$name)),
-  XGB_pred = as.vector(h2o.getFrame(best_xgb@model$cross_validation_holdout_predictions_frame_id$name))
-) %>% cor()
-```
-
-### Stacking a Search Grid
-
-```{r}
-# Define GBM hyperparameter grid
-hyper_grid <- list(
-  max_depth = c(1, 3, 5),
-  min_rows = c(1, 5, 10),
-  learn_rate = c(0.01, 0.05, 0.1),
-  learn_rate_annealing = c(0.99, 1),
-  sample_rate = c(0.5, 0.75, 1),
-  col_sample_rate = c(0.8, 0.9, 1)
+```{r, iml_dalex_components}
+# iml model agnostic object
+components_iml <- Predictor$new(
+  model = ensemble_tree, 
+  data = features, 
+  y = response, 
+  predict.fun = pred
 )
 
-# Define random grid search criteria
-search_criteria <- list(
-  strategy = "RandomDiscrete",
-  max_models = 25
-)
-
-# Build random grid search 
-random_grid <- h2o.grid(
-  algorithm = "gbm", grid_id = "gbm_grid", x = X, y = Y,
-  training_frame = train_h2o, hyper_params = hyper_grid,
-  search_criteria = search_criteria, ntrees = 5000, stopping_metric = "RMSE",
-  stopping_rounds = 10, stopping_tolerance = 0, nfolds = 10, 
-  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
-  seed = 123
+# DALEX model agnostic object
+components_dalex <- DALEX::explain(
+  model = ensemble_tree,
+  data = features,
+  y = response,
+  predict_function = pred
 )
@@
 
-```{r}
-# Sort results by RMSE
-h2o.getGrid(
-  grid_id = "gbm_grid", 
-  sort_by = "rmse"
+### Permutation based feature importance
+
+Implementation
+
+```{r, vip_plot}
+
+vip(
+  ensemble_tree,
+  train = as.data.frame(train_h2o),
+  method = "permute",
+  target = "Sale_Price",
+  metric = "RMSE",
+  nsim = 5,
+  sample_frac = 0.5,
+  pred_wrapper = pred
+)
+
+```
+
+Partial dependence
+
+```{r, pdp_partial}
+
+# Compute partial dependence values
+
+pd_values <- partial(
+  ensemble_tree,
+  train = as.data.frame(train_h2o), 
+  pred.var = "Gr_Liv_Area",
+  pred.fun = function(object, newdata) {
+      results <- mean(as.vector(h2o.predict(object, as.h2o(newdata))))
+      return(results)
+    },
+  grid.resolution = 20
+)
+
+head(pd_values)  # take a peak
+
+# Partial dependence plot
+autoplot(pd_values, rug = TRUE, train = as.data.frame(train_h2o))
+```
+
+### Individual conditional expectation
+
+```{r, ice_curves}
+
+# Construct ICE curves
+ice_non_centered <- partial(
+  ensemble_tree,
+  train = as.data.frame(train_h2o), 
+  pred.var = "Gr_Liv_Area",
+  pred.fun = pred,
+  grid.resolution = 20
+) %>%
+  autoplot(alpha = 0.05, center = FALSE) +
+  ggtitle("A) Non-centered ICE curves")
+
+# Construct c-ICE curves
+ice_centered <- partial(
+  ensemble_tree,
+  train = as.data.frame(train_h2o), 
+  pred.var = "Gr_Liv_Area",
+  pred.fun = pred,
+  grid.resolution = 20
+) %>%
+  autoplot(alpha = 0.05, center = TRUE) +
+  ggtitle("B) Centered ICE curves")
+
+# Display plots side by side
+gridExtra::grid.arrange(ice_non_centered, ice_centered, ncol = 2)
+```
+
+Implementation
+
+### Feature interactions
+
+```{r, pdp_ice_curves}
+# Construct c-ICE curves
+partial(
+  ensemble_tree,
+  train = as.data.frame(train_h2o), 
+  pred.var = "Gr_Liv_Area",
+  pred.fun = pred,
+  grid.resolution = 20,
+  plot = TRUE,
+  center = TRUE,
+  plot.engine = "ggplot2"
 )
@@
 
-```{r}
-# Grab the model_id for the top model, chosen by validation error
-best_model_id <- random_grid@model_ids[[1]]
-best_model <- h2o.getModel(best_model_id)
-h2o.performance(best_model, newdata = test_h2o)
+```{r, interactions}
+
+interact <- Interaction$new(components_iml)
+
+interact$results %>%
+  arrange(desc(.interaction)) %>%
+  head()
+
+plot(interact)
 ```
 
-```{r}
-# Train a stacked ensemble using the GBM grid
-ensemble <- h2o.stackedEnsemble(
-  x = X, y = Y, training_frame = train_h2o, model_id = "ensemble_gbm_grid",
-  base_models = random_grid@model_ids, metalearner_algorithm = "gbm"
+### Local interpretable model-agnostic explanations
+
+```{r, lime_explainer}
+
+# Create explainer object
+components_lime <- lime(
+  x = features,
+  model = ensemble_tree, 
+  n_bins = 10
 )
 
-# Eval ensemble performance on a test set
-h2o.performance(ensemble, newdata = test_h2o)
+class(components_lime)
+
+# Use LIME to explain previously defined instances: high_ob and low_ob
+lime_explanation <- lime::explain(
+  x = rbind(high_ob, low_ob), 
+  explainer = components_lime, 
+  n_permutations = 10,
+  dist_fun = "gower",
+  kernel_width = 0.25,
+  n_features = 10, 
+  feature_select = "highest_weights"
+)
+
+plot_features(lime_explanation, ncol = 1)
 ```
 
-### Auto ML
+### Tuning
 
-```{r}
-# Use AutoML to find a list of candidate models (i.e., leaderboard)
-auto_ml <- h2o.automl(
-  x = X, y = Y, training_frame = train_h2o, nfolds = 5, 
-  max_runtime_secs = 60 * 120, max_models = 50,
-  keep_cross_validation_predictions = TRUE, sort_metric = "RMSE", seed = 123,
-  stopping_rounds = 50, stopping_metric = "RMSE", stopping_tolerance = 0
+```{r, tuning}
+# Tune the LIME algorithm a bit
+lime_explanation2 <- explain(
+  x = rbind(high_ob, low_ob), 
+  explainer = components_lime, 
+  n_permutations = 5000,
+  dist_fun = "euclidean",
+  kernel_width = 0.75,
+  n_features = 10, 
+  feature_select = "lasso_path"
+)
+
+# Plot the results
+plot_features(lime_explanation2, ncol = 1)
+
+# Two-way PDP using iml
+interaction_pdp <- Partial$new(
+  components_iml, 
+  c("First_Flr_SF", "Overall_Qual"), 
+  ice = FALSE, 
+  grid.size = 20
+)
+
+labels <- interaction_pdp$results %>% filter(First_Flr_SF == max(First_Flr_SF))
+plot(interaction_pdp) + 
+  ggrepel::geom_label_repel(
+    data = labels, 
+    aes(label = Overall_Qual),
+    label.size = .05, 
+    label.padding = .15
+)
+
+# Construct c-ICE curves
+partial(
+  ensemble_tree,
+  train = as.data.frame(train_h2o), 
+  pred.var = "Gr_Liv_Area",
+  pred.fun = pred,
+  grid.resolution = 20,
+  plot = TRUE,
+  center = TRUE,
+  plot.engine = "ggplot2"
 )
@@
 
+### Shapley values
+
 ```{r}
-# Assess the leader board; the following truncates the results to show the top 
-# 25 models. You can get the top model with auto_ml@leader
-auto_ml@leaderboard %>% 
-  as.data.frame() %>%
-  dplyr::select(model_id, rmse) %>%
-  dplyr::slice(1:25)
+# Compute (approximate) Shapley values
+(shapley <- Shapley$new(components_iml, x.interest = high_ob, sample.size = 1000))
+
+# Plot results
+plot(shapley)
+
+# Reuse existing object
+shapley$explain(x.interest = low_ob)
+
+# Plot results
+shapley$results %>%
+  top_n(25, wt = abs(phi)) %>%
+  ggplot(aes(phi, reorder(feature.value, phi), color = phi > 0)) +
+  geom_point(show.legend = FALSE)
 ```
 
-#### Clean-up
+#### Clean up
 

### 2020-03-03 Mathematics and Statistics for FRM/02_Probabilities.rmd
iml
iml

--- 2020-03-02 Mathematics and Statistics for FRM/01_Basics.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
-   \rfoot{\color{headergrey}Chapter 1}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Probabilities}
+   \rfoot{\color{headergrey}Chapter 2}
    \lfoot{\color{headergrey}}
@@
 
-Solve for $y$, where:
-
-a.) $y = ln(e^5)$
-
-5
-
-b.) $y = ln(1/e)$
-
-$y = ln(1/e) = ln(1) - 1 = 0 - 1 = -1$
-
-c.) $y = ln(10e)$
-
-$y = ln(10e) = ln(10) + ln(e) = ln(10) + 1$
-
-```{r}
-log(10) + 1
-```
-
 ### 2
-
-The nominal monthly rate for a loan is quoted at 5%. What is the equivalent annual rate? Semiannual rate? Continious rate?
-
-```{r}
-effective_rate <- function(rate, periods) {
-   ((1 + rate/periods)^periods - 1)
-}
-```
-
-Annual:
-
-```{r}
-effective_rate(.05, 12) * 100
-```
-
-Semiannual:
-
-```{r}
-effective_rate(.05, 2) * 100
-```
-
-Continous:
-
-```{r}
-log(1 + effective_rate(.05, 12)) * 100
-```
 
@@
 
-Over the course of a year, the log return on a stock market index is 11.2%. The starting value of the index is 100. What is the value at the end of the year?
-
-```{r}
-100 * exp(.112)
-```
-
 ### 4
-
-You have a portfolio of 10 bonds. In how many different ways can exactly two bonds default? Assume the order in which the bonds default is unimportant.
-
-```{r}
-choose(10, 2)
-```
 
@@
 
-What is the present value of a perpetuity that pays $100 per year? Use an actual discount rate of 4%, and assume the first payment will be made in exactly one year.
-
-```{r}
-100 * (1/0.04)
-```
-
 ### 6
-
-ABC stock will pay a $1 dividend in one year.
-Assume the dividend will continue to be paid annually forever and the dividend payments will increase in size at a rate of 5%.
-Value this stream of dividends using a 6% annual discount rate.
-
-```{r}
-1 / (0.06 - 0.05)
-```
 
@@
 
-What is the present value of a 10-year bond with a $100 face value, which pays a 6% coupon annually? Use an 8% discount rate.
-
-```{r}
-
-present_value <- function(face, int.rate, periods, discount.rate) {
-   delta <- 1/(1 + discount.rate)
-
-   coupon <- face * int.rate
-
-   discount <- (delta - delta^(periods+1)) / (1 - delta)
-
-   coupon * discount + ( face / (1 + discount.rate)^periods )
-}
-
-present_value(100, .06, 10, .08)
-```
-
 ### 8
-
-Solve for $x$, where $e^{e^x} = 10$
-
-```{r}
-log(log(10))
-```
 
@@
 
-Calculate the value of the following summation: $\sum^{9}_{i=0}(-0.5)^i$
-
-```{r}
-r <- -0.5; n <- 9
-
-(1 - r^n) / (1 - r)
-```
-
 ### 10
 
-The risk department of your firm has 10 analysts. You need to select four analysts to serve on a special audit committee. How many possible groupings of four analysts can be put together?
-
-```{r}
-choose(10, 4)
-```
-
-long-hand:
-
-```{r}
-n <- 10; k <- 4
-
-factorial(10) / ( factorial(k) * factorial(n - k))
-```
-
 ### 11
-
-What is the present value of a newly issued 10-year bond with a notial value of $100 and a 2% annual coupon? Assume a constant 5% annual discount rate and no risk of default.
-
-```{r}
-present_value(100, .02, 10, 0.05)
-```

### 2020-03-03 h2o_reset.R

@@ -1,0 +1,8 @@
+library(h2o)
+
+h2o.init()
+prostate_path <- system.file("extdata", "prostate.csv", package = "h2o")
+prostate <- h2o.uploadFile(path = prostate_path)
+h2o.ls()
+h2o.removeAll()
+h2o.ls()

### 2020-03-06 Mathematics and Statistics for FRM/02_Probabilities.rmd
frm ch2
frm ch2

@@
 
+You are invested in two hedge funds.
+The probability that hedge fund Alpha generates positive returns in a given year is 60%.
+The probability that hedge fund Omega generates positive returns in a given year is 70%.
+Assume the returns are independent.
+
+What is the probability that both funds generate positive returns in a given year?
+
+```{r}
+prob_up <- .6 * .7
+```
+
+__`r prob_up*100`%__
+
+What is the probability that both funds lose money?
+
+```{r}
+prob_down <- (1 - .6) * (1 - .7)
+```
+
+__`r prob_down*100`%__
+
 ### 2
+
+Corporation ABC issues $100 million of bonds. The bonds are rated BBB. The probability that the rating on the bonds is upgraded within the year is 8%. The probability of a drowngrade is 4%.
+
+What is the probability that the raiting remains unchanged?
+
+```{r}
+prob <- (1 - 0.08 - 0.04)
+```
+
+__`r prob*100`%__
 
@@
 
+Stock XYZ has a 20% chance of losing more than 10% in a given month.
+There is also a 30% probability that XYZ gains more than 10%.
+What is the probability that stock XYZ either loses more than 10% or gains more than 10%?
+
+```{r}
+prob <- (.3 + .2)
+```
+
+__`r prob*100`%__
+
 ### 4
+
+There is a 30% chance that oil prices will increase over the next six months. If oil prices increase, there is a 60% chancee that the stock market will be down.
+
+What is the probability that oil prices increase and the stock market is down over the next 6 months?
+
+```{r}
+prob <- .3 * .6
+```
+
+__`r prob*100`%__
 
@@
 
+Given the following density function:
+
+$f(x) = c(100 - x^2)$ for $-10 \le x \le 10$
+
+Calculate the value of c.
+
+$\int^{\infty}_{-\infty} f(x)dx = \int^{10}_{-10} c(100 - x^2)dx = c[100x - \frac{1}{3}x^3]^{10}_{-10}$
+
+$c = \frac{3}{4,000}$
+
 ### 6
+
+Given the following cumulative distribution function, $F(x)$, for $0 \le x \le 10$:
+
+$F(x) = \frac{x}{100}(20 - x)$
+
+Check that this is a valid CDF; that is, show that $F(0)$ and $F(1)$ = 1.
+
+$F(x) = \frac{1}{5}x - \frac{x^2}{100}$
+
+For 0:
+
+$F(0) = \frac{1}{5}0 - \frac{0}{100}$
+
+$F(0) = 0$
+
+For 10:
+
+$F(10) = \frac{1}{5}10 - \frac{10^2}{100}$
+
+$F(10) = 2 - 1$
+
+$F(10) = 1$
+
+Calculate the probability density function, $f(x)$.
+
+$F^{-1}(x) = \frac{1}{50}(10 - x)$
 
@@
 
+Given the probability density function, $f(x) = \frac{c}{x}$,
+
+Where $1 \le x \le e$. Calculate the cumulative distribution function, $F(x)$, and solve for the constant c.
+
+$F(x) = \int^x_1 f(t)dt = \int^x_1 \frac{c}{t}dt = c[ln(t)]^x_1 = c\ln x$
+
+$F(x) = ln \ x$
+
 ### 8
+
+You own two bonds. Both bonds have a 30% probability of defaulting. Their default probabilities are statistically independent.
+
+What is the probability that both bonds default?
+
+```{r}
+prob <- .3 * .3
+```
+
+P[both] = __`r prob*100`%__
+
+What is the probability that only one bond defaults?
+
+```{r}
+prob <- (1 - .3) * 2*.3
+```
+
+P[one defaults] = __`r prob*100`%__
+
+What is the probability that neither bond defaults?
+
+```{r}
+prob <- (1 - .3)*(1 - .3)
+```
+
+P[neither] = __`r prob*100`%__
 
@@
 
+The following table gives a one-year rating transition matrix. Given a bond's rating now, the matrix gives the probability associated with the bond having a given rating in a year's time.
+
+```{r}
+ratings <- data.table(Grade = c("A", "B", "C", "D"), A = c(.9, .1, 0, 0), B = c(.08, .8, .25, 0), C = c(0.02 , 0.08, .6, 0), D = c(0, .02, .15, 1))
+
+pretty_kable(ratings, "Bond Transition Probabilities")
+```
+
+Given a B-rated bond, what is the probability that the bond default (D rating) over 1 year?
+
+```{r}
+prob <- 0.02
+```
+
+P[default] = __`r prob*100`%__
+
+What is the probability that the bond defaults over two years?
+
+```{r}
+prob <- .8 * .02 + .08*.15 + .02
+```
+
+P[default] = __`r prob*100`%__
+
 ### 10
 
-### 11
+Your firm forecasts that there is a 50% probability that the markets will be up significantly next year, a 20% probability that the market will be down significantly next year, and a 30% probability that the market will be flat, neither up or down significantly.
+
+You are asked to evaluate the prospects of a new portfolio manager. The manager has a long bias and is likely to perform better in an up market.
+
+Based on past data, you believe that the probability that the manager will be up if the market is up significantly is 80%, and that the probability that the manager will be up if the market is down significantly is only 10%.
+
+If the market is flat, the manager is just as likely to be up as to be down.
+
+What is the unconditional probability that the manager is up next year?
+
+$P[X_{up}] = P[X_{up} | M_{up}] * P[M_{up}] + P[X_{up} | M_{dn}] * P[M_{dn}] * P[X_{up}|M_{flat}] * P[M_{flat}]$
+
+$P[X_{up}]$ = 80% * 50% + 10% * 20% + 50% * 30%
+
+```{r}
+prob <- .8*.5 + .1*.2 + .5*.3
+```
+
+$P[X_{up}]$ = __`r prob*100`%__

### 2020-03-07 Hands-on Machine Learning/17-20_Unsupervised-Learning.rmd
homl unsupervised
homl unsupervised

--- 2020-03-02 Hands-on Machine Learning/16_Interpretable-Machine-Learning.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Interpretable Machine Learning}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Principal Component Analysis}
    \rfoot{\color{headergrey}\thepage}
-   \lfoot{\color{headergrey}Chapter 16}
+   \lfoot{\color{headergrey}Chapters 17-20}
    \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
@@
 
-## Interpretable Machine Learning
+## Unsupervised Learning
 
@@
 
-```{r, echo = T}
+```{r, ames_data}
 ames <- AmesHousing::make_ames()
@@
 
+Market Basket:
+
 ```{r}
-# ensure consistent categorical levels
-blueprint <- recipe(Sale_Price ~., data = ames_train) %>%
-  step_other(all_nominal(), threshold = 0.005)
+url <- "https://koalaverse.github.io/homlr/data/my_basket.csv"
 
-# Create training / test h2o frames
-train_h2o <- prep(blueprint, training = ames_train, retain = T) %>%
-  juice() %>%
-  as.h2o()
+my_basket <- readr::read_csv(url)
 
-test_h2o <- prep(blueprint, training = ames_train) %>%
-  bake(new_data = ames_test) %>%
-  as.h2o()
-
-Y <- "Sale_Price"
-X <- setdiff(names(ames_train), Y)
+dim(my_basket)
 ```
 
-### Stacking Overview
-
-Stacking is the process of combining multiple "base" learners (RF, GBM, GLM, etc.) into a "Super Learner" that uses the individual learners to make the final prediction.
-
-The Super Learner Algorithm
-
-1.) Setup the Ensemble
-
-  Specify a list of L base learners (with a specific set of model parameters)
-
-  Specify a meta learning algorithm. This can be any one of the algorithms discussed in the previous chapters, but most often is some form of regularized regression.
-
-2.) Train the ensemble
-
-  Train each of the L base learners on the training set.
-
-  Perform k-fold CV on each of the base learners and collect the cross-validated predictions from each (the same k-fold must be used for each base learner).
-These predicted values represent $p_1, ..., p_l$.
-
-  The N cross-validated predicted values from each of the L algorithms can be combined to form a new N X L feature matrix (Z)
-
-  Train the meta learning algorithm on level-one data (y = f(Z)).
-The "ensemble model" consista of the L base learning models and the meta learning model, which can then be used to generate predictions on new data.
-
-3.) Predict on new data.
-
-  To generate ensemble predictions, first generate predictions from the base learners.
-
-  Feed those predictions into the meta learner to generate the ensemble prediction.
-
-### Stacking Existing Models
-
-Restrictions:
-
-1.) All models must be trained on the same training set
-
-2.) All models must be trained with the same number of CV folds.
-
-3.) All models must use the same fold assignment to ensure the same observations are used.
-
-4.) The cross-validated predictions from all of the models must be preserved by setting keep_cross_validation_predictions = T.
+MNIST:
 
 ```{r}
-# Train & cross-validate a GLM model
-best_glm <- h2o.glm(
-  x = X, y = Y, training_frame = train_h2o, alpha = 0.1,
-  remove_collinear_columns = TRUE, nfolds = 10, fold_assignment = "Modulo",
-  keep_cross_validation_predictions = TRUE, seed = 123
-)
+mnist <- dslabs::read_mnist()
+```
 
-# Train & cross-validate a RF model
-best_rf <- h2o.randomForest(
-  x = X, y = Y, training_frame = train_h2o, ntrees = 1000, mtries = 20,
-  max_depth = 30, min_rows = 1, sample_rate = 0.8, nfolds = 10,
-  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
-  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
-  stopping_tolerance = 0
-)
+### Principal Components Analysis
 
-# Train & cross-validate a GBM model
-best_gbm <- h2o.gbm(
-  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.01,
-  max_depth = 7, min_rows = 5, sample_rate = 0.8, nfolds = 10,
-  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,es
+h2o PCA
 
-  seed = 123, stopping_rounds = 50, stopping_metric = "RMSE",
-  stopping_tolerance = 0
-)
+```{r}
+my_basket_h2o <- as.h2o(my_basket)
 
-# Train & cross-validate an XGBoost model
-best_xgb <- h2o.xgboost(
-  x = X, y = Y, training_frame = train_h2o, ntrees = 5000, learn_rate = 0.05,
-  max_depth = 3, min_rows = 3, sample_rate = 0.8, categorical_encoding = "Enum",
-  nfolds = 10, fold_assignment = "Modulo", 
-  keep_cross_validation_predictions = TRUE, seed = 123, stopping_rounds = 50,
-  stopping_metric = "RMSE", stopping_tolerance = 0
+# PCA
+my_pca <- h2o.prcomp(
+  training_frame = my_basket_h2o,
+  pca_method = "GramSVD",
+  k = ncol(my_basket_h2o),
+  transform = "STANDARDIZE",
+  impute_missing = T,
+  max_runtime_secs = 1000
 )
@@
 
-Train the ensemble
+Loadings
 
 ```{r}
-# Train a stacked tree ensemble
-ensemble_tree <- h2o.stackedEnsemble(
-  x = X, y = Y, training_frame = train_h2o, model_id = "my_tree_ensemble",
-  base_models = list(best_glm, best_rf, best_gbm, best_xgb),
-  metalearner_algorithm = "drf"
-)
+my_pca@model$eigenvectors %>%
+  as.data.frame() %>%
+  mutate(feature = row.names(.)) %>%
+  ggplot(aes(pc1, reorder(feature, pc1))) +
+  geom_point()
 ```
@@
 ```{r}
-get_rmse <- function(model) {
-  results <- h2o.performance(model, newdata = test_h2o)
-  results@metrics$RMSE
-}
-
-list(best_glm, best_rf, best_glm, best_xgb) %>%
-  purrr::map_dbl(get_rmse)
+my_pca@model$eigenvectors %>%
+  as.data.frame() %>%
+  mutate(feature = row.names(.)) %>%
+  ggplot(aes(pc1, pc2, label = feature)) +
+  geom_text()
 ```
 
-Stacked Results
+Eigenvalue Criterion
 
 ```{r}
-h2o.performance(ensemble_tree, newdata = test_h2o)@metrics$RMSE
+eigen <- my_pca@model$importance["Standard deviation", ] %>%
+  as.vector() %>%
+  .^2
+
+sum(eigen)
+
+which(eigen >= 1)
+```
+
+Proportion of Variance Explained
+
+```{r}
+data.frame(
+  PC = my_pca@model$importance %>% seq_along(),
+  PVE = my_pca@model$importance %>% .[2,] %>% unlist(),
+  CVE = my_pca@model$importance %>% .[3,] %>% unlist()
+) %>%
+  tidyr::gather(metric, variance_explained, -PC) %>%
+  ggplot(aes(PC, variance_explained)) +
+  geom_point() +
+  facet_wrap( ~ metric, ncol = 1, scales = "free")
 ```
@@
 data.frame(
-  GLM_pred = as.vector(h2o.getFrame(best_glm@model$cross_validation_holdout_predictions_frame_id$name)),
-  RF_pred = as.vector(h2o.getFrame(best_rf@model$cross_validation_holdout_predictions_frame_id$name)),
-  GBM_pred = as.vector(h2o.getFrame(best_gbm@model$cross_validation_holdout_predictions_frame_id$name)),
-  XGB_pred = as.vector(h2o.getFrame(best_xgb@model$cross_validation_holdout_predictions_frame_id$name))
-) %>% cor()
+  PC = my_pca@model$importance %>% seq_along,
+  PVE = my_pca@model$importance %>% .[2,] %>% unlist()
+) %>%
+  ggplot(aes(PC, PVE, group = 1, label = PC)) +
+  geom_point() +
+  geom_line() +
+  geom_text(nudge_y = 0.002)
 ```
 
-### Stacking a Search Grid
+## Generalized Low Rank Models
+
+h2o implementation
 
 ```{r}
-# Define GBM hyperparameter grid
-hyper_grid <- list(
-  max_depth = c(1, 3, 5),
-  min_rows = c(1, 5, 10),
-  learn_rate = c(0.01, 0.05, 0.1),
-  learn_rate_annealing = c(0.99, 1),
-  sample_rate = c(0.5, 0.75, 1),
-  col_sample_rate = c(0.8, 0.9, 1)
+basic_glrm <- h2o.glrm(
+  training_frame = my_basket_h2o,
+  k = 20,
+  loss = "Quadratic",
+  regularization_x = "None",
+  regularization_y = "None",
+  transform = "STANDARDIZE",
+  max_iterations = 2000,
+  seed = 123
 )
 
-# Define random grid search criteria
-search_criteria <- list(
-  strategy = "RandomDiscrete",
-  max_models = 25
-)
+summary(basic_glrm)
 
-# Build random grid search 
-random_grid <- h2o.grid(
-  algorithm = "gbm", grid_id = "gbm_grid", x = X, y = Y,
-  training_frame = train_h2o, hyper_params = hyper_grid,
-  search_criteria = search_criteria, ntrees = 5000, stopping_metric = "RMSE",
-  stopping_rounds = 10, stopping_tolerance = 0, nfolds = 10, 
-  fold_assignment = "Modulo", keep_cross_validation_predictions = TRUE,
-  seed = 123
-)
+plot(basic_glrm)
 ```
@@
 ```{r}
-# Sort results by RMSE
-h2o.getGrid(
-  grid_id = "gbm_grid", 
-  sort_by = "rmse"
-)
+data.frame(
+  PC = basic_glrm@model$importance %>% seq_along(),
+  PVE = basic_glrm@model$importance %>% .[2,] %>% unlist(),
+  CVE = basic_glrm@model$importance %>% .[3,] %>% unlist()
+) %>%
+  gather(metric, variance_explained, -PC) %>%
+  ggplot(aes(PC, variance_explained)) +
+  geom_point() +
+  facet_wrap( ~ metric, ncol = 1, scales = "free")
 ```
@@
 ```{r}
-# Grab the model_id for the top model, chosen by validation error
-best_model_id <- random_grid@model_ids[[1]]
-best_model <- h2o.getModel(best_model_id)
-h2o.performance(best_model, newdata = test_h2o)
+t(basic_glrm@model$archetypes)[1:5, 1:5]
 ```
@@
 ```{r}
-# Train a stacked ensemble using the GBM grid
-ensemble <- h2o.stackedEnsemble(
-  x = X, y = Y, training_frame = train_h2o, model_id = "ensemble_gbm_grid",
-  base_models = random_grid@model_ids, metalearner_algorithm = "gbm"
-)
+p1 <- t(basic_glrm@model$archetypes) %>%
+  as.data.frame() %>%
+  mutate(feature = row.names(.)) %>%
+  ggplot(aes(Arch1, reorder(feature, Arch1))) +
+  geom_point()
 
-# Eval ensemble performance on a test set
-h2o.performance(ensemble, newdata = test_h2o)
-```
+p2 <- t(basic_glrm@model$archetypes) %>%
+  as.data.frame() %>%
+  mutate(feature = row.names(.)) %>%
+  ggplot(aes(Arch1, Arch2, label = feature)) +
+  geom_text()
 
-### Auto ML
-
-```{r}
-# Use AutoML to find a list of candidate models (i.e., leaderboard)
-auto_ml <- h2o.automl(
-  x = X, y = Y, training_frame = train_h2o, nfolds = 5, 
-  max_runtime_secs = 60 * 120, max_models = 50,
-  keep_cross_validation_predictions = TRUE, sort_metric = "RMSE", seed = 123,
-  stopping_rounds = 50, stopping_metric = "RMSE", stopping_tolerance = 0
-)
+gridExtra::grid.arrange(p1, p2, nrow = 1)
 ```
@@
 ```{r}
-# Assess the leader board; the following truncates the results to show the top 
-# 25 models. You can get the top model with auto_ml@leader
-auto_ml@leaderboard %>% 
-  as.data.frame() %>%
-  dplyr::select(model_id, rmse) %>%
-  dplyr::slice(1:25)
+k8_glrm <- h2o.glrm(
+  training_frame = my_basket_h2o,
+  k = 8,
+  loss = "Quadratic",
+  regularization_x = "None",
+  regularization_y = "None",
+  transform = "STANDARDIZE",
+  max_iterations = 2000,
+  seed = 123
+)
+
+my_reconstruction <- h2o.reconstruct(k8_glrm, my_basket_h2o,
+                                     reverse_transform = T)
+
+my_reconstruction[1:5, 1:4] %>% round(0)
 ```
 
-#### Clean-up
+```{r}
+# Use non-negative regularization
+k8_glrm_regularized <- h2o.glrm(
+  training_frame = my_basket_h2o,
+  k = 8, 
+  loss = "Quadratic",
+  regularization_x = "NonNegative", 
+  regularization_y = "NonNegative",
+  gamma_x = 0.5,
+  gamma_y = 0.5,
+  transform = "STANDARDIZE", 
+  max_iterations = 2000,
+  seed = 123
+)
+
+# Show predicted values
+predict(k8_glrm_regularized, my_basket_h2o)[1:5, 1:5]
+
+# Compare regularized versus non-regularized loss
+par(mfrow = c(1, 2))
+plot(k8_glrm)
+plot(k8_glrm_regularized)
+```
+
+```{r}
+# Split data into train & validation
+split <- h2o.splitFrame(my_basket_h2o, ratios = 0.75, seed = 123)
+train <- split[[1]]
+valid <- split[[2]]
+
+# Create hyperparameter search grid
+params <- expand.grid(
+  regularization_x = c("None", "NonNegative", "L1"),
+  regularization_y = c("None", "NonNegative", "L1"),
+  gamma_x = seq(0, 1, by = .25),
+  gamma_y = seq(0, 1, by = .25),
+  error = 0,
+  stringsAsFactors = FALSE
+)
+
+# Perform grid search
+for(i in seq_len(nrow(params))) {
+
+  # Create model
+  glrm_model <- h2o.glrm(
+    training_frame = train,
+    k = 8, 
+    loss = "Quadratic",
+    regularization_x = params$regularization_x[i], 
+    regularization_y = params$regularization_y[i],
+    gamma_x = params$gamma_x[i],
+    gamma_y = params$gamma_y[i],
+    transform = "STANDARDIZE", 
+    max_runtime_secs = 1000,
+    seed = 123
+  )
+
+  # Predict on validation set and extract error
+  validate <- h2o.performance(glrm_model, valid)
+  params$error[i] <- validate@metrics$numerr
+}
+
+# Look at the top 10 models with the lowest error rate
+params %>%
+  arrange(error) %>%
+  head(10)
+```
+
+```{r}
+# Apply final model with optimal hyperparamters
+final_glrm_model <- h2o.glrm(
+  training_frame = my_basket_h2o,
+  k = 8, 
+  loss = "Quadratic",
+  regularization_x = "L1", 
+  regularization_y = "NonNegative",
+  gamma_x = 1,
+  gamma_y = 0.25,
+  transform = "STANDARDIZE", 
+  max_iterations = 2000,
+  seed = 123
+)
+
+# New observations to score
+new_observations <- as.h2o(sample_n(my_basket, 2))
+
+# Basic scoring
+predict(final_glrm_model, new_observations) %>% round(0)
+```
+
+### Autoencoders
+
+```{r}
+# Convert mnist features to an h2o input data set
+features <- as.h2o(mnist$train$images)
+
+# Train an autoencoder
+ae1 <- h2o.deeplearning(
+  x = seq_along(features),
+  training_frame = features,
+  autoencoder = TRUE,
+  hidden = 2,
+  activation = 'Tanh',
+  sparse = TRUE
+)
+
+# Extract the deep features
+ae1_codings <- h2o.deepfeatures(ae1, features, layer = 1)
+ae1_codings
+```
+
+### Stacked Autoencoders
+
+```{r}
+# Hyperparameter search grid
+hyper_grid <- list(hidden = list(
+  c(50),
+  c(100), 
+  c(300, 100, 300),
+  c(100, 50, 100),
+  c(250, 100, 50, 100, 250)
+))
+
+# Execute grid search
+ae_grid <- h2o.grid(
+  algorithm = 'deeplearning',
+  x = seq_along(features),
+  training_frame = features,
+  grid_id = 'autoencoder_grid',
+  autoencoder = TRUE,
+  activation = 'Tanh',
+  hyper_params = hyper_grid,
+  sparse = TRUE,
+  ignore_const_cols = FALSE,
+  seed = 123
+)
+
+# Print grid details
+h2o.getGrid('autoencoder_grid', sort_by = 'mse', decreasing = FALSE)
+```
+
+Visualizing Reconstruction
+
+```{r}
+# Get sampled test images
+index <- sample(1:nrow(mnist$test$images), 4)
+sampled_digits <- mnist$test$images[index, ]
+colnames(sampled_digits) <- paste0("V", seq_len(ncol(sampled_digits)))
+
+# Predict reconstructed pixel values
+best_model_id <- ae_grid@model_ids[[1]]
+best_model <- h2o.getModel(best_model_id)
+reconstructed_digits <- predict(best_model, as.h2o(sampled_digits))
+names(reconstructed_digits) <- paste0("V", seq_len(ncol(reconstructed_digits)))
+
+combine <- rbind(sampled_digits, as.matrix(reconstructed_digits))
+
+# Plot original versus reconstructed
+par(mfrow = c(1, 3), mar=c(1, 1, 1, 1))
+layout(matrix(seq_len(nrow(combine)), 4, 2, byrow = FALSE))
+for(i in seq_len(nrow(combine))) {
+  image(matrix(combine[i, ], 28, 28)[, 28:1], xaxt="n", yaxt="n")
+}
+```
+
+```{r}
+ae100_codings <- h2o.deepfeatures(best_model, features, layer = 1)
+ae100_codings %>% 
+    as.data.frame() %>% 
+    tidyr::gather() %>%
+    summarize(average_activation = mean(value))
+```
+
+```{r}
+# Hyperparameter search grid
+hyper_grid <- list(sparsity_beta = c(0.01, 0.05, 0.1, 0.2))
+
+# Execute grid search
+ae_sparsity_grid <- h2o.grid(
+  algorithm = 'deeplearning',
+  x = seq_along(features),
+  training_frame = features,
+  grid_id = 'sparsity_grid',
+  autoencoder = TRUE,
+  hidden = 100,
+  activation = 'Tanh',
+  hyper_params = hyper_grid,
+  sparse = TRUE,
+  average_activation = -0.1,
+  ignore_const_cols = FALSE,
+  seed = 123
+)
+
+# Print grid details
+h2o.getGrid('sparsity_grid', sort_by = 'mse', decreasing = FALSE)
+```
+
+```{r}
+# Extract reconstruction errors
+(reconstruction_errors <- h2o.anomaly(best_model, features))
+
+# Plot distribution
+reconstruction_errors <- as.data.frame(reconstruction_errors)
+
+ggplot(reconstruction_errors, aes(Reconstruction.MSE)) +
+  geom_histogram(aes(fill = ..count..), bins = 30)
+
+```
+
+#### Clean up
 

### 2020-03-08 Hands-on Machine Learning/21-22_Clustering.rmd
homl, clustering
homl, clustering

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Clustering}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapters 17-20}
   \fancyfoot[C]{\rmfamily\color{headergrey}Hands-On Machine Learning}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
fig_width: 9
fig_height: 3.5
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")
```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}
# Data Wrangling
library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(tidyr, quietly = TRUE, warn.conflicts = FALSE)

# Plotting / Graphics

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(visdat, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)
library(ROCR, quietly = TRUE, warn.conflicts = FALSE)
library(ggmap, quietly = T, warn.conflicts = FALSE)
library(rpart.plot, quietly = T, warn.conflicts = F)
library(svmpath, quietly = T, warn.conflicts = F)

# Formatting / Markdown

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

# Feature Engineering
library(recipes, quietly = TRUE, warn.conflicts = FALSE)

# Utility
library(here, quietly = TRUE, warn.conflicts = FALSE)

# Resampling & Modeling
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(rsample, quietly = TRUE, warn.conflicts = FALSE)
library(caret, quietly = TRUE, warn.conflicts = FALSE)
library(pdp, quietly = TRUE, warn.conflicts = FALSE)

library(cluster, quietly = TRUE, warn.conflicts = FALSE)
library(factoextra, quietly = TRUE, warn.conflicts = FALSE)
library(yardstick, quietly = TRUE, warn.conflicts = FALSE)
library(NbClust, quietly = TRUE, warn.conflicts = FALSE)
library(mclust, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- paste0(here::here(), "/Hands-On/data/")

select <- dplyr::select # fix clash with MASS

# Set global R options
options(scipen = 999)
```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

## Clustering

### Data Sets

Attrition

```{r, attrition_data}
attrition <- attrition %>% mutate_if(is.ordered, factor, order = F)

churn <- initial_split(attrition, prop = .7, strata = "Attrition")

churn_train <- training(churn)
churn_test <- testing(churn)

rm(churn)
```

Ames, Iowa housing data.

```{r, ames_data}
ames <- AmesHousing::make_ames()

set.seed(123)

ames_split <- initial_split(ames, prop =.7, strata = "Sale_Price")

ames_train <- training(ames_split)
ames_test <- testing(ames_split)

rm(ames_split)

response <- "Sale_Price"
predictors <- setdiff(colnames(ames_train), response)
```

Market Basket:

```{r}
url <- "https://koalaverse.github.io/homlr/data/my_basket.csv"

my_basket <- readr::read_csv(url)

dim(my_basket)
```

MNIST:

```{r}
mnist <- dslabs::read_mnist()
```

```{r}
data(geyser, package = 'MASS')
```

### K-means Clustering

```{r}
features <- mnist$train$images

# Use k-means model with 10 centers and 10 random starts
mnist_clustering <- kmeans(features, centers = 10, nstart = 10)

# Print contents of the model output
str(mnist_clustering)
```

```{r}
# Extract cluster centers
mnist_centers <- mnist_clustering$centers

# Plot typical cluster digits
par(mfrow = c(2, 5), mar=c(0.5, 0.5, 0.5, 0.5))
layout(matrix(seq_len(nrow(mnist_centers)), 2, 5, byrow = FALSE))
for(i in seq_len(nrow(mnist_centers))) {
  image(matrix(mnist_centers[i, ], 28, 28)[, 28:1], 
        col = gray.colors(12, rev = TRUE), xaxt="n", yaxt="n")
}
```

```{r}
# Create mode function
mode_fun <- function(x){  
  which.max(tabulate(x))
}

mnist_comparison <- data.frame(
  cluster = mnist_clustering$cluster,
  actual = mnist$train$labels
) %>%
  group_by(cluster) %>%
  mutate(mode = mode_fun(actual)) %>%
  ungroup() %>%
  mutate_all(factor, levels = 0:9)

# Create confusion matrix and plot results
yardstick::conf_mat(
  mnist_comparison, 
  truth = actual, 
  estimate = mode
) %>%
  autoplot(type = 'heatmap')
```

```{r}
fviz_nbclust(
  my_basket, 
  kmeans, 
  k.max = 25,
  method = "wss",
  diss = get_dist(my_basket, method = "spearman")
)
```

# Mixed Data

```{r}
# Full ames data set --> recode ordinal variables to numeric
ames_full <- AmesHousing::make_ames() %>%
  mutate_if(str_detect(names(.), 'Qual|Cond|QC|Qu'), as.numeric)

# One-hot encode --> retain only the features and not sale price
full_rank  <- caret::dummyVars(Sale_Price ~ ., data = ames_full, 
                               fullRank = TRUE)
ames_1hot <- predict(full_rank, ames_full)

# Scale data
ames_1hot_scaled <- scale(ames_1hot)

# New dimensions
dim(ames_1hot_scaled)
```

```{r}
set.seed(123)

fviz_nbclust(
  ames_1hot_scaled, 
  kmeans, 
  method = "wss", 
  k.max = 25, 
  verbose = FALSE
)
```

```{r}
# Original data minus Sale_Price
ames_full <- AmesHousing::make_ames() %>% select(-Sale_Price)

# Compute Gower distance for original data
gower_dst <- daisy(ames_full, metric = "gower")
```

```{r}
# You can supply the Gower distance matrix to several clustering algos
pam_gower <- pam(x = gower_dst, k = 8, diss = TRUE)
diana_gower <- diana(x = gower_dst, diss = TRUE)
agnes_gower <- agnes(x = gower_dst, diss = TRUE)
```

```{r}
fviz_nbclust(
  ames_1hot_scaled, 
  pam, 
  method = "wss", 
  k.max = 25, 
  verbose = FALSE
)
```

```{r}
# k-means computation time on MNIST data
system.time(kmeans(features, centers = 10))

# CLARA computation time on MNIST data
system.time(clara(features, k = 10))
```

### Hierachial Clustering

```{r}
ames_scale <- AmesHousing::make_ames() %>%
  select_if(is.numeric) %>%  # select numeric columns
  select(-Sale_Price) %>%    # remove target column
  mutate_all(as.double) %>%  # coerce to double type
  scale()                    # center & scale the resulting columns
```

```{r}
# For reproducibility
set.seed(123)

# Dissimilarity matrix
d <- dist(ames_scale, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d, method = "complete" )
```

```{r}
# For reproducibility
set.seed(123)

# Compute maximum or complete linkage clustering with agnes
hc2 <- agnes(ames_scale, method = "complete")

# Agglomerative coefficient
hc2$ac
```

```{r}
# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(ames_scale, method = x)$ac
}

# get agglomerative coefficient for each linkage method
purrr::map_dbl(m, ac)
```

```{r}
# compute divisive hierarchical clustering
hc4 <- diana(ames_scale)

# Divise coefficient; amount of clustering structure found
hc4$dc
```

### Optimal Clusters

```{r}
# Plot cluster results
p1 <- fviz_nbclust(ames_scale, FUN = hcut, method = "wss", 
                   k.max = 10) +
  ggtitle("(A) Elbow method")
p2 <- fviz_nbclust(ames_scale, FUN = hcut, method = "silhouette", 
                   k.max = 10) +
  ggtitle("(B) Silhouette method")
p3 <- fviz_nbclust(ames_scale, FUN = hcut, method = "gap_stat", 
                   k.max = 10) +
  ggtitle("(C) Gap statistic")

# Display plots side by side
gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
```

```{r}
# Construct dendorgram for the Ames housing example
hc5 <- hclust(d, method = "ward.D2" )
dend_plot <- fviz_dend(hc5)
dend_data <- attr(dend_plot, "dendrogram")
dend_cuts <- cut(dend_data, h = 8)
fviz_dend(dend_cuts$lower[[2]])
```

```{r}
# Ward's method
hc5 <- hclust(d, method = "ward.D2" )

# Cut tree into 4 groups
sub_grp <- cutree(hc5, k = 8)

# Number of members in each cluster
table(sub_grp)
```

```{r}
# Plot full dendogram
fviz_dend(
  hc5,
  k = 8,
  horiz = TRUE,
  rect = TRUE,
  rect_fill = TRUE,
  rect_border = "jco",
  k_colors = "jco",
  cex = 0.1
)
```

```{r}
dend_plot <- fviz_dend(hc5)                # create full dendogram
dend_data <- attr(dend_plot, "dendrogram") # extract plot info
dend_cuts <- cut(dend_data, h = 70.5)      # cut the dendogram at 
                                           # designated height
# Create sub dendrogram plots
p1 <- fviz_dend(dend_cuts$lower[[1]])
p2 <- fviz_dend(dend_cuts$lower[[1]], type = 'circular')

# Side by side plots
gridExtra::grid.arrange(p1, p2, nrow = 1)
```

### Model-Based Clustering

```{r}
data(geyser, package = 'MASS')

# Apply GMM model with 3 components
geyser_mc <- Mclust(geyser, G = 3)

par(mfrow = c(1, 2))
# Plot results
plot(geyser_mc, what = "density")
plot(geyser_mc, what = "uncertainty")
```

```{r}
# Observations with high uncertainty
sort(geyser_mc$uncertainty, decreasing = TRUE) %>% head()
```

```{r}
summary(geyser_mc)
```

```{r}
geyser_optimal_mc <- Mclust(geyser)

summary(geyser_optimal_mc)
```

```{r}
my_basket_mc <- Mclust(my_basket, 1:20)

summary(my_basket_mc)
## ---------------------------------------------------- 
## Gaussian finite mixture model fitted by EM algorithm 
## ---------------------------------------------------- 
## 
## Mclust EEV (ellipsoidal, equal volume and shape) model with 6 components: 
## 
##  log-likelihood    n   df      BIC       ICL
##        8308.915 2000 5465 -24921.1 -25038.38
## 
## Clustering table:
##   1   2   3   4   5   6 
## 391 403  75 315 365 451

plot(my_basket_mc, what = 'BIC', 
     legendArgs = list(x = "bottomright", ncol = 5))
```

```{r}
probabilities <- my_basket_mc$z 
colnames(probabilities) <- paste0('C', 1:6)

probabilities <- probabilities %>%
  as.data.frame() %>%
  mutate(id = row_number()) %>%
  tidyr::gather(cluster, probability, -id)

ggplot(probabilities, aes(probability)) +
  geom_histogram() +
  facet_wrap(~ cluster, nrow = 2)
```

```{r}
uncertainty <- data.frame(
  id = 1:nrow(my_basket),
  cluster = my_basket_mc$classification,
  uncertainty = my_basket_mc$uncertainty
)

uncertainty %>%
  group_by(cluster) %>%
  filter(uncertainty > 0.25) %>%
  ggplot(aes(uncertainty, reorder(id, uncertainty))) +
  geom_point() +
  facet_wrap(~ cluster, scales = 'free_y', nrow = 1)
```

```{r}
cluster2 <- my_basket %>%
  scale() %>%
  as.data.frame() %>%

  mutate(cluster = my_basket_mc$classification) %>%
  filter(cluster == 2) %>%
  select(-cluster)

cluster2 %>%
  tidyr::gather(product, std_count) %>%
  group_by(product) %>%
  summarize(avg = mean(std_count)) %>%
  ggplot(aes(avg, reorder(product, avg))) +
  geom_point() +
  labs(x = "Average standardized consumption", y = NULL)
```

#### Clean up

```{r}
# clean up
rm(list = ls())
```

### 2020-04-06 Case Studies/01_geolocation.rmd
sln refactor + case studies/computational methods/statistical computing work
sln refactor + case studies/computational methods/statistical computing work

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Geolocation}
   \rfoot{\color{headergrey}Chapter 1}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Case Studies In Data Science}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(codetools, quietly = TRUE, warn.conflicts = FALSE)
library(lattice, quietly = TRUE, warn.conflicts = FALSE)
library(fields, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- file.path(here::here(), "Case Studies", "datasets")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# Geolocation

## Predicting Location via Indoor Positioning Systems

### The Raw Data

```{r}

file_offline <- file.path(data.dir, "offline.final.trace.txt")
file_online <- file.path(data.dir, "online.final.trace.txt")

raw_offline <- read_lines(file_offline)
raw_online <- read_lines(file_online)

```

#### Sanity Check

```{r}
# number of comment lines in the data
sum(substr(raw_offline, 1, 1) == "#")

# total number of lines in the data
length(raw_offline)
```

#### Data Pre-processing

Generate read data function for preprocessing training and test data.

```{r}

processLine <- function(x) {
   tokens <- strsplit(x, "[;=,]")[[1]]

   if(length(tokens) == 10)
      return(NULL)

   tmp <- matrix(tokens[ - (1:10)], ncol = 4, byrow = T)
   cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow = nrow(tmp),
                ncol = 6, byrow = T), tmp)
}

validLines <- function(data) {
   substr(data, 1, 1) != "#"
}

roundOrientation <- function(angles) {
   refs = seq(0, by = 45, length = 9)
   q <- sapply(angles, function(o) which.min(abs(o - refs)))
   c(refs[1:8], 0)[q]
}

readData <- function(file, submacs = macs) {

   lines <- read_lines(file)

   valid_lines <- lines[ validLines(lines) ]

   processed_lines <- lapply(valid_lines, processLine)

   data <- as.data.table(do.call("rbind", processed_lines),
                         stringsAsFactors = F)

   names(data) <- c("time", "scanMac", "posX", "posY", "posZ",
                       "orientation", "mac", "signal", 
                       "channel", "type")

   numVars <- c("time", "posX", "posY", "posZ",
                "orientation", "signal")

   data[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]

   data <- data[ data$type == 3, ]
   data[, type := NULL]

   data[, rawTime := time]
   data[, time := time/1000]
   class(data$time) = c("POSIXt", "POSIXct")

   # drop scanMac & posZ
   data[, `:=`(scanMac = NULL, posZ = NULL)]

   data$angle = roundOrientation(data$orientation)

   data$channel = NULL

   data$posXY <- paste(data$posX, data$posY, sep = "-")

   return(data[mac %in% submacs])
}

```

#### Test Pre-processor

```{r}
lines <- processLine(raw_offline[4:20])
lines

offline_valid <- raw_offline[validLines(raw_offline)]

head(offline_valid)

length(offline_valid)
```

```{r}
tmp <- lapply(offline_valid[1:17], processLine)

sapply(tmp, nrow)
```

#### Dry Run

```{r}
offline_test <- as.data.table(do.call("rbind", tmp))

offline_test
```

#### Process

```{r}
# Enter Debug Context

# options(error = recover, warn = 2)

offline_data <- lapply(offline_valid, processLine)
offline <- as.data.table(do.call("rbind", offline_data),
                         stringsAsFactors = F)

names(offline) <- c("time", "scanMac", "posX", "posY", "posZ",
                    "orientation", "mac", "signal", 
                    "channel", "type")

numVars <- c("time", "posX", "posY", "posZ",
             "orientation", "signal")

dim(offline)

offline[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]

offline <- offline[ offline$type == 3, ]
offline[, type := NULL]

offline[, rawTime := time]
offline[, time := time/1000]
class(offline$time) = c("POSIXt", "POSIXct")

# options(error = recover, warn = 1)
```

#### Data Cleaning

```{r}
summary(offline)
```

### Orientation Exploration

```{r}
length(unique(offline$orientation))
```

```{r, fig.height=5}
plot(ecdf(offline$orientation))
```

```{r}
ggplot(offline, aes(orientation)) +
   stat_ecdf() +
   labs("Orientation")
```

```{r}
ggplot(offline, aes(orientation, fill = ..count..)) +
   geom_histogram(bins = 30) +
   scale_y_continuous(labels = comma) +
   labs("Orientation Value Clusters")
```

```{r}

offline$angle <- roundOrientation(offline$orientation)

# angle = cleaned orientation column
ggplot(offline, aes(angle)) +
   stat_ecdf()

ggplot(offline, aes(angle, fill = ..count..)) +
   geom_histogram(bins = 30) +
   scale_y_continuous(labels = comma) +
   labs("Cleaned Up Angles")
```

```{r}
with(offline, boxplot(orientation ~ angle,
                      xlab = "nearest 45 degree angle",
                      ylab = "orientation"))

ggplot(offline, aes(angle, orientation, group = angle)) +
   geom_boxplot() +
   labs(title = "Cleaned Angle vs Raw Orientation", 
        x = "nearest 45 degree angle", 
        y = "orientation")
```

### MAC Address

```{r}
c(length(unique(offline$mac)), length(unique(offline$channel)))

table(offline$mac)

ggplot(offline, aes(mac, fill = ..count..)) +
   geom_histogram(stat = "count", bins = 30) +
   scale_y_continuous(labels = comma)

# keep only the top 7 MAC address data points

dim(offline)

offline_macs <- names(sort(table(offline$mac), decreasing = T)[1:7])

dim(offline)
```

```{r}
macChannel <- with(offline, table(mac, channel))
apply(macChannel, 1, function(x) sum(x > 0))

# mac and channel are 1:1, we can remove channel
#offline$channel := NULL
```

### Exploring the Position of the Hand-Held Device

```{r}
locDF <- with(offline,
              by(offline, list(posX, posY), function(x) x))

length(locDF)

sum(sapply(locDF, is.null))

locDF <- locDF[ !sapply(locDF, is.null)]

length(locDF)

locCounts <- sapply(locDF, nrow)

locCounts <- sapply(locDF,
                    function(df)
                       c(df[1, c("posX", "posY")], count = nrow(df)))

class(locCounts)

dim(locCounts)

locCounts[ , 1:8]

locCountsDF <- as.data.table(t(locCounts))

locCountsDF$posX <- unlist(locCountsDF$posX)
locCountsDF$posY <- unlist(locCountsDF$posY)

ggplot(locCountsDF, aes(posX, posY, label = count)) +
   geom_text(angle = 45) +
   labs(title = "X / Y postion counts")
```

### Final Data Prep

```{r}
offline <- readData(file_offline, offline_macs)
```

#### Signal Strength

```{r, fig.height=7}
bwplot(signal ~ factor(angle) | mac, data = offline,
       subset = posX == 2 & posY == 12 & 
          mac != "00:0f:a3:39:dd:cd",
       layout = c(2, 3))

summary(offline$signal)
```

```{r, fig.height=7}
densityplot(~ signal | mac + factor(angle), data = offline,
            subset = posX == 24 & posY == 4 &
               mac != "00:0f:a3:39:dd:cd",
            bw = 0.5, plot.points = F)
```

```{r, fig.width=5}
byLocAngleAP <- with(offline,
                     by(offline, list(posXY, angle, mac),
                        function(x) x))
signalSummary <-
   lapply(byLocAngleAP,
          function(oneLoc) {
             ans = oneLoc[1, ]
             ans$medSignal = median(oneLoc$signal)
             ans$avgSignal = mean(oneLoc$signal)
             ans$num = length(oneLoc$signal)
             ans$sdSignal = sd(oneLoc$signal)
             ans$iqrSignal = IQR(oneLoc$signal)
             ans
          })

offlineSummary <- do.call("rbind", signalSummary)

breaks <- seq(-90, -30, by = 5)

bwplot(sdSignal ~ cut(avgSignal, breaks = breaks),
       data = offlineSummary,
       subset = mac != "00:0f:a3:39:dd:cd",
       xlab = "Mean Signal", ylab = "SD Signal")
```

```{r}
with(offlineSummary,
     smoothScatter((avgSignal - medSignal) ~ num,
                   xlab = "Number of Observations",
                   ylab = "mean - median"))
abline(h = 0, col = "#984ea3", lwd = 2)

lo.obj <- with(offlineSummary,
               loess(diff ~ num,
                     data = data.frame(diff = (avgSignal - medSignal),
                                       num = num)))

lo.obj.pr <- predict(lo.obj, newdata = data.frame(num = (70:120)))
lines(x = 70:120, y = lo.obj.pr, col = "#4daf4a", lwd = 2)
```

#### Signal and Distance

```{r, fig.height=6}

subMacs <- names(sort(table(offline$mac), decreasing = T))[1:7]

surfaceSS <- function(data, mac, angle) {
   oneAPAngle <- subset(offlineSummary, mac == mac & angle == angle)

   smothSS <- Tps(oneAPAngle[, c("posX", "posY")],
                  oneAPAngle$avgSignal)

   vizSmooth <- predictSurface(smothSS)

   plot.surface(vizSmooth, type = "C")
   points(oneAPAngle$posX, oneAPAngle$posY, pch=19, cex = 0.5)
}

surfaceSS(offlineSummary, subMacs[5], 0)
```

```{r, fig.height = 7}
parCur <- par(mfrow = c(2,2), mar = rep(1, 4))
mapply(surfaceSS, mac = subMacs[ rep(c(5, 1), each = 2)],
       data = list(data = offlineSummary))
par(parCur)
```

Exclude one of two similar access points

```{r}
offlineSummary <- subset(offlineSummary, mac != subMacs[2])
```

```{r}
AP <- matrix( c( 7.5, 6.3, 2.5, -.8, 12.8, -2.8,
                 1, 14, 33.5, 9.3, 33.5, 2.8),
              ncol = 2, byrow = T,
              dimnames = list(subMacs[ -2 ], c("x", "y") ))

ggplot(data.table(mac = rownames(AP), AP), aes(x, y)) +
   geom_point()
```

```{r, fig.height=7}
diffs <- offlineSummary[, c("posX", "posY")] - AP[offlineSummary$mac, ]

offlineSummary$dist <- sqrt(diffs[, 1]^2 + diffs[, 2]^2)

xyplot(signal ~ dist | factor(mac) + factor(angle),
       data = offlineSummary, pch = 19, cex = 0.3,
       xlab = "distance")
```

### Nearest Neighbor Methods to Predict Location

```{r}
macs <- unique(offlineSummary$mac)

online <- readData(file_online, submacs = macs)

length(unique(online$posXY))
```

```{r}
tabonlineXYA = table(online$posXY, online$angle)
tabonlineXYA[1:6, ]
```

```{r}
keepVars <- c("posXY", "posX", "posY", "orientation", "angle")

byLoc <- with(online,
              by(online, list(posXY),
                 function(x) {
                    ans <- x[1, ..keepVars]
                    avgSS <- tapply(x$signal, x$mac, mean)
                    y = matrix(avgSS, nrow = 1, ncol = 6,
                               dimnames = list(ans$posXY, names(avgSS)))
                    cbind(ans, y)
                 }))

onlineSummary <- do.call("rbind", byLoc)

dim(onlineSummary)

names(onlineSummary)
```

#### Choice of Orientation

```{r}

reshapeSS <- function(data, varSignal = "signal",
                      keepVars = c("posXY", "posX", "posY")) {

   byLocation <- with(data,
                 by(data, list(posXY),
                    function(x) {
                       ans <- x[1, ..keepVars]
                       avgSS <- tapply(x$signal, x$mac, mean)
                       y = matrix(avgSS, nrow = 1, ncol = 6,
                                  dimnames = list(ans$posXY, names(avgSS)))
                       cbind(ans, y)
                    }))

   newDataSS <- do.call("rbind", byLocation)

   col_names <- colnames(newDataSS)
   n_cols <- length(col_names)

   colnames(newDataSS)[4:n_cols] <- sapply(col_names[4:n_cols], function(col) {
      n <- nchar(col)
      substr(col, n - 2, n)
   })

   newDataSS[, 4:ncol(newDataSS)] <- round(newDataSS[, 4:ncol(newDataSS)])

   return(newDataSS)
}

selectTrain <- function(angleNewObs, signals, m) {

   refs <- seq(0, by = 45, length = 8)

   nearestAngle <- roundOrientation(angleNewObs)

   if( m %% 2 == 1) {
      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)
   } else {
      m = m + 1
      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)

      if( sign(angleNewObs - nearestAngle) >= 1)
         angles = angles[ -1 ]
      else
         angles = angles[ -m ]
   }

   angles <- angles + nearestAngle
   angles[angles < 0] = angles[ angles < 0] + 360
   angles[angles > 360] = angles[ angles > 360 ] - 360

   signals <- signals[angle %in% angles, ]

   reshapeSS(signals, varSignal = "avgSignal")
}

train130 <- selectTrain(130, offlineSummary, m = 3)

dim(train130)

findNN <- function(newSignal, trainSubset) {
   diffs <- apply(trainSubset[ , 4:9], 1,
                  function(x) x - newSignal)
   dists <- apply(diffs, 2, function(x) sqrt(sum(x^2)) )
   closest <- order(dists)
   return(trainSubset[closest, 1:3 ])
}

predXY <- function(newSignals, newAngles, trainData,
                   numAngles = 1, k = 3) {

   closeXY <- list(length = nrow(newSignals))

   for(i in 1:nrow(newSignals)) {
      trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
      closeXY[[i]] = 
         findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
   }

   estXY = lapply(closeXY,
                  function(x) sapply(x[, 2:3],
                                     function(x) mean(x[1:k])))

   estXY <- do.call("rbind", estXY)

   return(estXY)
}

estXYk1 <- predXY(newSignals = onlineSummary[, 6:11],
                  newAngles = onlineSummary[, 4],
                  offlineSummary, numAngles = 3, k = 1)

estXYk3 <- predXY(newSignals = onlineSummary[, 6:11],
                  newAngles = onlineSummary[, 4],
                  offlineSummary, numAngles = 3, k = 3)
```

```{r}
calcError <- function(estXY, actualXY)
   sum( rowSums( (estXY - actualXY) ^ 2) )

actualXY <- onlineSummary[, c("posX", "posY")]
sapply(list(estXYk1, estXYk3), calcError, actualXY)

```

### 2020-04-06 Computational Methods/01_intro-numerical-analysis.rmd
sln refactor + case studies/computational methods/statistical computing work
sln refactor + case studies/computational methods/statistical computing work

--- 2020-04-06 Case Studies/01_geolocation.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Geolocation}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Intoduction to Numerical Analysis}
    \rfoot{\color{headergrey}Chapter 1}
    \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey}Case Studies In Data Science}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Methods for Numerical Analysis}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 
-library(codetools, quietly = TRUE, warn.conflicts = FALSE)
-library(lattice, quietly = TRUE, warn.conflicts = FALSE)
-library(fields, quietly = TRUE, warn.conflicts = FALSE)
+library(cmna, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-# Geolocation
+## Introduction to Numerical Analysis
 
-## Predicting Location via Indoor Positioning Systems
+### Numerical Analysis in R
 
-### The Raw Data
+Example, factorial:
+
+```{r}
+factorial <- function(n) {
+   if(n == 0)
+      return(0)
+   if(n == 1)
+      return(1)
+   return(factorial(n - 1) + factorial(n - 2))
+}
+
+sapply(1:10, factorial)
+```
+
+```{r}
+isPrime <- function(n) {
+   if( n == 2)
+      return(TRUE)
+
+   for(i in 2:sqrt(n))
+      if(n %% i == 0)
+         return(FALSE)
+
+   return(TRUE)
+}
+
+nums <- 1:25
+
+rbind(nums, sapply(nums, isPrime))
+```
+
+```{r}
+x <- c(TRUE, FALSE, TRUE, FALSE, TRUE)
+sum(x); mean(x); length(x)
+```
+
+```{r}
+naievesum <- function(x) {
+
+   s <- 0
+   n <- length(x)
+
+   for(i in 1:n) {
+      s <- S + x[i]
+   }
+
+   s
+}
+
+naivesum(c(1, 2, 3, 4.5, -6))
+```
+
+```{r}
+pwisesum <- function(x) {
+
+   n <- length(x)
+
+   if(n == 1)
+      return(x)
+
+   m = floor(n/2)
+
+   return(pwisesum(x[1:m]) + pwisesum(x[(m + 1):n]))
+}
+
+pwisesum(c(1, 2, 3, 4.5, -6))
+
+```
+
+```{r}
+kahan_sum <- function(x) {
+   comp <- s <- 0
+   n <- length(x)
+
+   for(i in 1:n) {
+      y <- x[i] - comp
+      t <- x[i] + s
+      comp <- (t - s) - y
+      s <- t
+   }
+
+   return(s)
+}
+
+kahan_sum(1:8)
+```
+
+```{r}
+naive_poly <- function(x, coefs) {
+   y <- rep(0, length(x))
+
+   for(i in 1:length(coefs)) {
+      y <- y + coefs[i] * (x ^ (i - 1))
+   }
+
+   return(y)
+}
+
+# example poly:
+
+# f(x) = (x-5)(x+3)(x+2)(x-1)
+# = x^4 + x^3 - x^2 - 11x + 30
+# = 30 - 11 x - 19 x^2 - x^3 + x^4
+
+f <- c(30, -11, -19, -1, 1)
+
+x <- c(-1, 0, 1)
+
+naive_poly(x, f)
+
+```
+
+```{r}
+better_poly <- function(x, coefs) {
+   y <- rep(0, length(x))
+
+   cached.x <- 1
+
+   for(i in 1:length(coefs)) {
+      y <- y + coefs[i] * cached.x
+      cached.x <- cached.x * x
+   }
+
+   return(y)
+}
+
+better_poly(x, f)
+```
+
+```{r}
+horner <- function(x, coefs) {
+   y <- rep(0, length(x))
+
+   for(i in length(coefs):1)
+      y <- coefs[i] + x * y
+
+   return(y)
+}
+
+horner(x, f)
+```
+
+```{r}
+nth_root <- function(a, n, tol = 1 / 1000) {
+   x <- 1
+   deltax <- tol * 10
+
+   while(abs(deltax) > tol) {
+      deltax <- (1 / n) * (a / x ^ (n - 1) - x)
+      x <- x + deltax
+   }
+
+   return(x)
+}
+
+nth_root(100, 2)
+nth_root(65536, 4)
+nth_root(1000, 3)
+nth_root(pi, 2)
+```
+
+## Exercises
+
+#### 1.)
+
+What is the circumference of a 10-meter-diameter circle symbolically?
+
+$2\pi r$, r = 10
+
+$10\pi$
+
+What is it numerically?
+
+```{r}
+10*pi
+```
+
+Which one is correct? *Both*
+
+#### 2.)
+
+Describe an application for which a numerical solution are preferable to a symbolic solution.
+
+_An exact symbolic solution to some problems can be difficult or near impossible, so a good approximation via numerical methods is preferable than no solution._
+
+Describe an application where symbolic solutions are preferable to numerical.
+
+_An analytical/symbolic solution can solve a related family of problems._
+
+#### 3.)
+
+If an unsigned variable were stored using 16 bits, what is the largest number it could hold?
+
+```{r}
+2^16
+```
+
+#### 4.)
+
+Using R's internal trigonometric functions, implement a secant function.
+
+```{r}
+sec <- function(x) 1/cos(x)
+
+assertthat::are_equal(sec(pi), -1)
+
+```
+
+#### 5.)
+
+Using the exponentiation operator, create a function that finds the nth root of any number.
+
+```{r}
+nth_root <- function(x, n) x^(1/n)
+
+assertthat::are_equal(nth_root(32, 5), 2)
+
+```
+
+#### 6.)
+
+Function 1.1 provides a recursive implementation of the Fibonacci sequence. Rewrite the function to provide a result without recursion.
+
+```{r}
+fib_iter <- function(x) {
+
+   # build look-up
+   fib_values <- list()
+   fib_values[[1]] <- 1
+
+   for(index in 2:x) {
+
+      fib1 <- ifelse(index <= 1, 1, fib_values[[index - 1]])
+      fib2 <- ifelse(index <= 1, 1, ifelse(index == 2, 1, fib_values[[index - 2]]))
+
+      fib_values[index] <- fib1 + fib2
+   }
+
+   return(fib_values[[x]])
+}
+
+sapply(1:10, fib_iter)
+```
+
+What are the relative strenghts and weaknesses of each implementation?
+
+_The recursive version is a bit more readable._
+
+#### 7.)
+
+Implement a product function in R without using recursion.
 
@@
 
-file_offline <- file.path(data.dir, "offline.final.trace.txt")
-file_online <- file.path(data.dir, "online.final.trace.txt")
-
-raw_offline <- read_lines(file_offline)
-raw_online <- read_lines(file_online)
-
 ```
 
-#### Sanity Check
-
-```{r}
-# number of comment lines in the data
-sum(substr(raw_offline, 1, 1) == "#")
-
-# total number of lines in the data
-length(raw_offline)
-```
-
-#### Data Pre-processing
-
-Generate read data function for preprocessing training and test data.
-
-```{r}
-
-processLine <- function(x) {
-   tokens <- strsplit(x, "[;=,]")[[1]]
-
-   if(length(tokens) == 10)
-      return(NULL)
-
-   tmp <- matrix(tokens[ - (1:10)], ncol = 4, byrow = T)
-   cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow = nrow(tmp),
-                ncol = 6, byrow = T), tmp)
-}
-
-validLines <- function(data) {
-   substr(data, 1, 1) != "#"
-}
-
-roundOrientation <- function(angles) {
-   refs = seq(0, by = 45, length = 9)
-   q <- sapply(angles, function(o) which.min(abs(o - refs)))
-   c(refs[1:8], 0)[q]
-}
-
-readData <- function(file, submacs = macs) {
-
-   lines <- read_lines(file)
-
-   valid_lines <- lines[ validLines(lines) ]
-
-   processed_lines <- lapply(valid_lines, processLine)
-
-   data <- as.data.table(do.call("rbind", processed_lines),
-                         stringsAsFactors = F)
-
-   names(data) <- c("time", "scanMac", "posX", "posY", "posZ",
-                       "orientation", "mac", "signal", 
-                       "channel", "type")
-
-   numVars <- c("time", "posX", "posY", "posZ",
-                "orientation", "signal")
-
-   data[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]
-
-   data <- data[ data$type == 3, ]
-   data[, type := NULL]
-
-   data[, rawTime := time]
-   data[, time := time/1000]
-   class(data$time) = c("POSIXt", "POSIXct")
-
-   # drop scanMac & posZ
-   data[, `:=`(scanMac = NULL, posZ = NULL)]
-
-   data$angle = roundOrientation(data$orientation)
-
-   data$channel = NULL
-
-   data$posXY <- paste(data$posX, data$posY, sep = "-")
-
-   return(data[mac %in% submacs])
-}
-
-```
-
-#### Test Pre-processor
-
-```{r}
-lines <- processLine(raw_offline[4:20])
-lines
-
-offline_valid <- raw_offline[validLines(raw_offline)]
-
-head(offline_valid)
-
-length(offline_valid)
-```
-
-```{r}
-tmp <- lapply(offline_valid[1:17], processLine)
-
-sapply(tmp, nrow)
-```
-
-#### Dry Run
-
-```{r}
-offline_test <- as.data.table(do.call("rbind", tmp))
-
-offline_test
-```
-
-#### Process
-
-```{r}
-# Enter Debug Context
-
-# options(error = recover, warn = 2)
-
-offline_data <- lapply(offline_valid, processLine)
-offline <- as.data.table(do.call("rbind", offline_data),
-                         stringsAsFactors = F)
-
-names(offline) <- c("time", "scanMac", "posX", "posY", "posZ",
-                    "orientation", "mac", "signal", 
-                    "channel", "type")
-
-numVars <- c("time", "posX", "posY", "posZ",
-             "orientation", "signal")
-
-dim(offline)
-
-offline[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]
-
-offline <- offline[ offline$type == 3, ]
-offline[, type := NULL]
-
-offline[, rawTime := time]
-offline[, time := time/1000]
-class(offline$time) = c("POSIXt", "POSIXct")
-
-# options(error = recover, warn = 1)
-```
-
-#### Data Cleaning
-
-```{r}
-summary(offline)
-```
-
-### Orientation Exploration
-
-```{r}
-length(unique(offline$orientation))
-```
-
-```{r, fig.height=5}
-plot(ecdf(offline$orientation))
-```
-
-```{r}
-ggplot(offline, aes(orientation)) +
-   stat_ecdf() +
-   labs("Orientation")
-```
-
-```{r}
-ggplot(offline, aes(orientation, fill = ..count..)) +
-   geom_histogram(bins = 30) +
-   scale_y_continuous(labels = comma) +
-   labs("Orientation Value Clusters")
-```
-
-```{r}
-
-offline$angle <- roundOrientation(offline$orientation)
-
-# angle = cleaned orientation column
-ggplot(offline, aes(angle)) +
-   stat_ecdf()
-
-ggplot(offline, aes(angle, fill = ..count..)) +
-   geom_histogram(bins = 30) +
-   scale_y_continuous(labels = comma) +
-   labs("Cleaned Up Angles")
-```
-
-```{r}
-with(offline, boxplot(orientation ~ angle,
-                      xlab = "nearest 45 degree angle",
-                      ylab = "orientation"))
-
-ggplot(offline, aes(angle, orientation, group = angle)) +
-   geom_boxplot() +
-   labs(title = "Cleaned Angle vs Raw Orientation", 
-        x = "nearest 45 degree angle", 
-        y = "orientation")
-```
-
-### MAC Address
-
-```{r}
-c(length(unique(offline$mac)), length(unique(offline$channel)))
-
-table(offline$mac)
-
-ggplot(offline, aes(mac, fill = ..count..)) +
-   geom_histogram(stat = "count", bins = 30) +
-   scale_y_continuous(labels = comma)
-
-# keep only the top 7 MAC address data points
-
-dim(offline)
-
-offline_macs <- names(sort(table(offline$mac), decreasing = T)[1:7])
-
-dim(offline)
-```
-
-```{r}
-macChannel <- with(offline, table(mac, channel))
-apply(macChannel, 1, function(x) sum(x > 0))
-
-# mac and channel are 1:1, we can remove channel
-#offline$channel := NULL
-```
-
-### Exploring the Position of the Hand-Held Device
-
-```{r}
-locDF <- with(offline,
-              by(offline, list(posX, posY), function(x) x))
-
-length(locDF)
-
-sum(sapply(locDF, is.null))
-
-locDF <- locDF[ !sapply(locDF, is.null)]
-
-length(locDF)
-
-locCounts <- sapply(locDF, nrow)
-
-locCounts <- sapply(locDF,
-                    function(df)
-                       c(df[1, c("posX", "posY")], count = nrow(df)))
-
-class(locCounts)
-
-dim(locCounts)
-
-locCounts[ , 1:8]
-
-locCountsDF <- as.data.table(t(locCounts))
-
-locCountsDF$posX <- unlist(locCountsDF$posX)
-locCountsDF$posY <- unlist(locCountsDF$posY)
-
-ggplot(locCountsDF, aes(posX, posY, label = count)) +
-   geom_text(angle = 45) +
-   labs(title = "X / Y postion counts")
-```
-
-### Final Data Prep
-
-```{r}
-offline <- readData(file_offline, offline_macs)
-```
-
-#### Signal Strength
-
-```{r, fig.height=7}
-bwplot(signal ~ factor(angle) | mac, data = offline,
-       subset = posX == 2 & posY == 12 & 
-          mac != "00:0f:a3:39:dd:cd",
-       layout = c(2, 3))
-
-summary(offline$signal)
-```
-
-```{r, fig.height=7}
-densityplot(~ signal | mac + factor(angle), data = offline,
-            subset = posX == 24 & posY == 4 &
-               mac != "00:0f:a3:39:dd:cd",
-            bw = 0.5, plot.points = F)
-```
-
-```{r, fig.width=5}
-byLocAngleAP <- with(offline,
-                     by(offline, list(posXY, angle, mac),
-                        function(x) x))
-signalSummary <-
-   lapply(byLocAngleAP,
-          function(oneLoc) {
-             ans = oneLoc[1, ]
-             ans$medSignal = median(oneLoc$signal)
-             ans$avgSignal = mean(oneLoc$signal)
-             ans$num = length(oneLoc$signal)
-             ans$sdSignal = sd(oneLoc$signal)
-             ans$iqrSignal = IQR(oneLoc$signal)
-             ans
-          })
-
-offlineSummary <- do.call("rbind", signalSummary)
-
-breaks <- seq(-90, -30, by = 5)
-
-bwplot(sdSignal ~ cut(avgSignal, breaks = breaks),
-       data = offlineSummary,
-       subset = mac != "00:0f:a3:39:dd:cd",
-       xlab = "Mean Signal", ylab = "SD Signal")
-```
-
-```{r}
-with(offlineSummary,
-     smoothScatter((avgSignal - medSignal) ~ num,
-                   xlab = "Number of Observations",
-                   ylab = "mean - median"))
-abline(h = 0, col = "#984ea3", lwd = 2)
-
-lo.obj <- with(offlineSummary,
-               loess(diff ~ num,
-                     data = data.frame(diff = (avgSignal - medSignal),
-                                       num = num)))
-
-lo.obj.pr <- predict(lo.obj, newdata = data.frame(num = (70:120)))
-lines(x = 70:120, y = lo.obj.pr, col = "#4daf4a", lwd = 2)
-```
-
-#### Signal and Distance
-
-```{r, fig.height=6}
-
-subMacs <- names(sort(table(offline$mac), decreasing = T))[1:7]
-
-surfaceSS <- function(data, mac, angle) {
-   oneAPAngle <- subset(offlineSummary, mac == mac & angle == angle)
-
-   smothSS <- Tps(oneAPAngle[, c("posX", "posY")],
-                  oneAPAngle$avgSignal)
-
-   vizSmooth <- predictSurface(smothSS)
-
-   plot.surface(vizSmooth, type = "C")
-   points(oneAPAngle$posX, oneAPAngle$posY, pch=19, cex = 0.5)
-}
-
-surfaceSS(offlineSummary, subMacs[5], 0)
-```
-
-```{r, fig.height = 7}
-parCur <- par(mfrow = c(2,2), mar = rep(1, 4))
-mapply(surfaceSS, mac = subMacs[ rep(c(5, 1), each = 2)],
-       data = list(data = offlineSummary))
-par(parCur)
-```
-
-Exclude one of two similar access points
-
-```{r}
-offlineSummary <- subset(offlineSummary, mac != subMacs[2])
-```
-
-```{r}
-AP <- matrix( c( 7.5, 6.3, 2.5, -.8, 12.8, -2.8,
-                 1, 14, 33.5, 9.3, 33.5, 2.8),
-              ncol = 2, byrow = T,
-              dimnames = list(subMacs[ -2 ], c("x", "y") ))
-
-ggplot(data.table(mac = rownames(AP), AP), aes(x, y)) +
-   geom_point()
-```
-
-```{r, fig.height=7}
-diffs <- offlineSummary[, c("posX", "posY")] - AP[offlineSummary$mac, ]
-
-offlineSummary$dist <- sqrt(diffs[, 1]^2 + diffs[, 2]^2)
-
-xyplot(signal ~ dist | factor(mac) + factor(angle),
-       data = offlineSummary, pch = 19, cex = 0.3,
-       xlab = "distance")
-```
-
-### Nearest Neighbor Methods to Predict Location
-
-```{r}
-macs <- unique(offlineSummary$mac)
-
-online <- readData(file_online, submacs = macs)
-
-length(unique(online$posXY))
-```
-
-```{r}
-tabonlineXYA = table(online$posXY, online$angle)
-tabonlineXYA[1:6, ]
-```
-
-```{r}
-keepVars <- c("posXY", "posX", "posY", "orientation", "angle")
-
-byLoc <- with(online,
-              by(online, list(posXY),
-                 function(x) {
-                    ans <- x[1, ..keepVars]
-                    avgSS <- tapply(x$signal, x$mac, mean)
-                    y = matrix(avgSS, nrow = 1, ncol = 6,
-                               dimnames = list(ans$posXY, names(avgSS)))
-                    cbind(ans, y)
-                 }))
-
-onlineSummary <- do.call("rbind", byLoc)
-
-dim(onlineSummary)
-
-names(onlineSummary)
-```
-
-#### Choice of Orientation
-
-```{r}
-
-reshapeSS <- function(data, varSignal = "signal",
-                      keepVars = c("posXY", "posX", "posY")) {
-
-   byLocation <- with(data,
-                 by(data, list(posXY),
-                    function(x) {
-                       ans <- x[1, ..keepVars]
-                       avgSS <- tapply(x$signal, x$mac, mean)
-                       y = matrix(avgSS, nrow = 1, ncol = 6,
-                                  dimnames = list(ans$posXY, names(avgSS)))
-                       cbind(ans, y)
-                    }))
-
-   newDataSS <- do.call("rbind", byLocation)
-
-   col_names <- colnames(newDataSS)
-   n_cols <- length(col_names)
-
-   colnames(newDataSS)[4:n_cols] <- sapply(col_names[4:n_cols], function(col) {
-      n <- nchar(col)
-      substr(col, n - 2, n)
-   })
-
-   newDataSS[, 4:ncol(newDataSS)] <- round(newDataSS[, 4:ncol(newDataSS)])
-
-   return(newDataSS)
-}
-
-selectTrain <- function(angleNewObs, signals, m) {
-
-   refs <- seq(0, by = 45, length = 8)
-
-   nearestAngle <- roundOrientation(angleNewObs)
-
-   if( m %% 2 == 1) {
-      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)
-   } else {
-      m = m + 1
-      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)
-
-      if( sign(angleNewObs - nearestAngle) >= 1)
-         angles = angles[ -1 ]
-      else
-         angles = angles[ -m ]
-   }
-
-   angles <- angles + nearestAngle
-   angles[angles < 0] = angles[ angles < 0] + 360
-   angles[angles > 360] = angles[ angles > 360 ] - 360
-
-   signals <- signals[angle %in% angles, ]
-
-   reshapeSS(signals, varSignal = "avgSignal")
-}
-
-train130 <- selectTrain(130, offlineSummary, m = 3)
-
-dim(train130)
-
-findNN <- function(newSignal, trainSubset) {
-   diffs <- apply(trainSubset[ , 4:9], 1,
-                  function(x) x - newSignal)
-   dists <- apply(diffs, 2, function(x) sqrt(sum(x^2)) )
-   closest <- order(dists)
-   return(trainSubset[closest, 1:3 ])
-}
-
-predXY <- function(newSignals, newAngles, trainData,
-                   numAngles = 1, k = 3) {
-
-   closeXY <- list(length = nrow(newSignals))
-
-   for(i in 1:nrow(newSignals)) {
-      trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
-      closeXY[[i]] = 
-         findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
-   }
-
-   estXY = lapply(closeXY,
-                  function(x) sapply(x[, 2:3],
-                                     function(x) mean(x[1:k])))
-
-   estXY <- do.call("rbind", estXY)
-
-   return(estXY)
-}
-
-estXYk1 <- predXY(newSignals = onlineSummary[, 6:11],
-                  newAngles = onlineSummary[, 4],
-                  offlineSummary, numAngles = 3, k = 1)
-
-estXYk3 <- predXY(newSignals = onlineSummary[, 6:11],
-                  newAngles = onlineSummary[, 4],
-                  offlineSummary, numAngles = 3, k = 3)
-```
-
-```{r}
-calcError <- function(estXY, actualXY)
-   sum( rowSums( (estXY - actualXY) ^ 2) )
-
-actualXY <- onlineSummary[, c("posX", "posY")]
-sapply(list(estXYk1, estXYk3), calcError, actualXY)
-
-```
+#### 8.)
+#### 9.)
+#### 10.)
+#### 11.)
+#### 12.)
+#### 13.)
+#### 14.)
+#### 15.)
+#### 16.)
+#### 17.)
+#### 18.)
+#### 19.)

### 2020-04-06 Mathematics and Statistics for FRM/03_Statistics.rmd
sln refactor + case studies/computational methods/statistical computing work
sln refactor + case studies/computational methods/statistical computing work

--- 2020-03-02 Mathematics and Statistics for FRM/01_Basics.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Basics}
-   \rfoot{\color{headergrey}Chapter 1}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Statistics}
+   \rfoot{\color{headergrey}Chapter 3}
    \lfoot{\color{headergrey}}
@@
 
-#### Chapter 1
+#### Chapter 3
 
@@
 
-Solve for $y$, where:
-
-a.) $y = ln(e^5)$
-
-5
-
-b.) $y = ln(1/e)$
-
-$y = ln(1/e) = ln(1) - 1 = 0 - 1 = -1$
-
-c.) $y = ln(10e)$
-
-$y = ln(10e) = ln(10) + ln(e) = ln(10) + 1$
+Compute the mean and the median of the following series of returns:
 
 ```{r}
-log(10) + 1
+dat <- data.table(Returns = c(.12, .5, -.08, .2, .04, .1, .02))
+
+pretty_kable(dat, "Returns")
+```
+
+```{r}
+mean(dat$Returns)
+
+median(dat$Returns)
 ```
@@
 
-The nominal monthly rate for a loan is quoted at 5%. What is the equivalent annual rate? Semiannual rate? Continious rate?
+Compute the sample mean and the standard deviation of the following returns:
 
 ```{r}
-effective_rate <- function(rate, periods) {
-   ((1 + rate/periods)^periods - 1)
-}
-```
+dat <- data.table(Returns = c(.12, .5, -.08, .2, .04, .1, .02))
 
-Annual:
-
-```{r}
-effective_rate(.05, 12) * 100
-```
-
-Semiannual:
-
-```{r}
-effective_rate(.05, 2) * 100
-```
-
-Continous:
-
-```{r}
-log(1 + effective_rate(.05, 12)) * 100
+pretty_kable(dat, "Returns")
 ```
@@
 
-Over the course of a year, the log return on a stock market index is 11.2%. The starting value of the index is 100. What is the value at the end of the year?
+Prove that Equation 3.2 is an unbiased estimator of the mean. That is, show that $\mathbb{E}[\hat\mu] = \mu$
 
-```{r}
-100 * exp(.112)
-```
+$\mathbb{E}(\bar{x}) = \mathbb{E}[\frac{x_1 + x_2 + x_3 + \ldots + x_n}{n}]$
+
+$\ldots = \mathbb{E}\frac{1}{n}(x_1 + x_2 + x_3 + \ldots + x_n)$
+
+$\ldots = \frac{1}{n}\mathbb{E}(x_1 + x_2 + x_3 + \ldots + x_n)$
+
+$\ldots = \frac{1}{n}(\mathbb{E}[x_1] + \mathbb{E}[x_2] + \mathbb{E}[x_3] + \ldots + \mathbb{E}[x_n])$
+
+$\ldots = \frac{1}{n}(\mu_1 + \mu_2 + \mu_3 + \ldots + \mu_n)$
+
+$\ldots = \frac{1}{n}n(u)$
+
+$\ldots = \mu$
 
@@
 
-You have a portfolio of 10 bonds. In how many different ways can exactly two bonds default? Assume the order in which the bonds default is unimportant.
+What is the standard deviation of the estimator in Eq. 3.2? Assume the various data points are i.i.d.
 
-```{r}
-choose(10, 2)
-```
+$\mu = \frac{1}{n}\sum^{n}_{i=1}r_i$
+
+$\frac{1}{n}\sum^{n}_{i=1}r_i = \frac{1}{n}(r_1 + r_2 + r_3 \ldots + r_n)$
+
+$\sigma^2 = \frac{1}{n}\sum(x_i - \mu)^2$
+
+$\sigma_{\mu} = \frac{\sigma_{}}{\sqrt{n}}$
 
@@
 
-What is the present value of a perpetuity that pays $100 per year? Use an actual discount rate of 4%, and assume the first payment will be made in exactly one year.
+Calculate the population covariance and correlation of the following series:
 
 ```{r}
-100 * (1/0.04)
+dat <- data.table(S1 = c(.21, .53, .83, .19), S2 = c(.2, .32, .8, .4))
+
+pretty_kable(dat, "Returns")
+```
+
+```{r}
+cov(dat$S1, dat$S2)
+
+cor(dat$S1, dat$S2)
 ```
@@
 
-ABC stock will pay a $1 dividend in one year.
-Assume the dividend will continue to be paid annually forever and the dividend payments will increase in size at a rate of 5%.
-Value this stream of dividends using a 6% annual discount rate.
+Calculate the population mean, standard deviation and skewness of each of the following two series:
 
 ```{r}
-1 / (0.06 - 0.05)
+dat <- data.table(S1 = c(-51, -21, 21, 51), S2 = c(-61, -7, 33, 35))
+
+pretty_kable(dat, "Returns")
+```
+
+```{r}
+
+skew <- function(x) {
+   mean(x - mean(x))^3 / sd(x)^3
+}
+
+apply(dat, 2, mean)
+apply(dat, 2, sd)
+apply(dat, 2, skew)
+
 ```
@@
 
-What is the present value of a 10-year bond with a $100 face value, which pays a 6% coupon annually? Use an 8% discount rate.
+Calculate the population mean, standard deviation and skewness of each of the following two series:
+
+```{r}
+dat <- data.table(S1 = c(-23, -7, 7, 23), S2 = c(-17, -17, 17, 17))
+
+pretty_kable(dat, "Returns")
+```
 
@@
 
-present_value <- function(face, int.rate, periods, discount.rate) {
-   delta <- 1/(1 + discount.rate)
-
-   coupon <- face * int.rate
-
-   discount <- (delta - delta^(periods+1)) / (1 - delta)
-
-   coupon * discount + ( face / (1 + discount.rate)^periods )
+skew <- function(x) {
+   mean(x - mean(x))^3 / sd(x)^3   
 }
 
-present_value(100, .06, 10, .08)
+apply(dat, 2, mean)
+apply(dat, 2, sd)
+apply(dat, 2, skew)
+
 ```
@@
 
-Solve for $x$, where $e^{e^x} = 10$
+Given the probability density function for a random variable, X
 
-```{r}
-log(log(10))
-```
+$f(x) = \frac{x}{18}$ for $0 \le x \le 6$
+
+$\mu = \int_0^6 \frac{x^3}{18 * 3} = \frac{6^3}{3*18} - \frac{0^3}{3*18}$
+
+$\frac{6^2}{3^2} = 4$
+
+$\sigma^2 = \int_0^6(x - 4)^2\frac{x}{18}dx$
+
+$\frac{1}{18}(x^3 - 8x^2 + 16x)dx$
+
+$\sigma^2 = 2(9 - 16 + 8) = 2$
 
@@
 
-Calculate the value of the following summation: $\sum^{9}_{i=0}(-0.5)^i$
+Prove that Equation 3.19, reproduced here, is an unbiased estimator of the variance.
 
-```{r}
-r <- -0.5; n <- 9
+$\hat\sigma^2_x = \frac{1}{n - 1}\sum_{i=1}^n(x_i - \hat\mu_x)$
 
-(1 - r^n) / (1 - r)
-```
+$\hat\sigma^2_x = \frac{1}{n - 1}\sum_{i=1}^n = \frac{1}{n - 1}\sum_{i=1}^n(\frac{n-1}{n}x_i - \frac{1}{n}\sum_{j \ne x}x_j)^2$
+
+$\hat\sigma^2_x = \frac{1}{n}\sum_{i=1}^nx^2_i - \frac{1}{n(n - 1)}\sum{i=1}^n\sum_{i \ne j}x_ix_j$
+
+$\mathbb{E}[x^2_i] = \sigma^2 + \mu^2$
+
+$\mathbb{E}[x_i x_j] = \mu_i\mu_j + iff \ \sigma_{ij} = 0 \ \forall i \ne j$
 
@@
 
-The risk department of your firm has 10 analysts. You need to select four analysts to serve on a special audit committee. How many possible groupings of four analysts can be put together?
+Given two random variables, $X_A$ and $X_B$, with the corresponding means $\mu_A$ and $\mu_B$ and standard deviations $\sigma_A$ and $\sigma_B$, prove that the variance of $X_A$ plus $X_B$ is:
 
-```{r}
-choose(10, 4)
-```
+$\mathbb{Var}[X_A + X_B] = \sigma^2_A + \sigma^2_B + 2\rho_{AB}\sigma_A\sigma_B$
 
-long-hand:
+$\mathbb{E}[X_A + X_B] = \mathbb{E}[X_A] + \mathbb{E}[X_B] = \mu_A + \mu_B$
 
-```{r}
-n <- 10; k <- 4
+$\mathbb{Var}[X_A + X_B] = \mathbb{E}[(X_A + X_B - \mathbb{E}[X_A + X_b])^2]$
 
-factorial(10) / ( factorial(k) * factorial(n - k))
-```
+$\mathbb{Var}[X_A + X_B] = \sigma^2_A + \sigma^2_B + 2Cov[X_A, X_B]$
 
@@
 
-What is the present value of a newly issued 10-year bond with a notial value of $100 and a 2% annual coupon? Assume a constant 5% annual discount rate and no risk of default.
+A $100 notional, zero coupon bond has one year to expiry. The probability of default is 10%. In the event of default, assume that the recovery rate is 40%.
+
+The continiously compounded discount rate is 5%. What is the present value of this bond?
 
 ```{r}
-present_value(100, .02, 10, 0.05)
+EV <- .9 * 100 + 0.10 * 40
 ```
+
+$\mathbb{E}(V) = `r EV`$
+
+```{r}
+PV <- exp(-.05) * EV
+```
+
+$\mathbb{P}(V)$ = $`r round(PV, 2)`

### 2020-04-06 Statistical Computing/01_introduction.rmd
sln refactor + case studies/computational methods/statistical computing work
sln refactor + case studies/computational methods/statistical computing work

--- 2020-04-06 Computational Methods/01_intro-numerical-analysis.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Intoduction to Numerical Analysis}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Introduction}
    \rfoot{\color{headergrey}Chapter 1}
    \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Methods for Numerical Analysis}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Methods for Numerical Analysis}}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 
-## Introduction to Numerical Analysis
+## Introduction to Statistical Computing
 
-### Numerical Analysis in R
-
-Example, factorial:
-
-```{r}
-factorial <- function(n) {
-   if(n == 0)
-      return(0)
-   if(n == 1)
-      return(1)
-   return(factorial(n - 1) + factorial(n - 2))
-}
-
-sapply(1:10, factorial)
-```
-
-```{r}
-isPrime <- function(n) {
-   if( n == 2)
-      return(TRUE)
-
-   for(i in 2:sqrt(n))
-      if(n %% i == 0)
-         return(FALSE)
-
-   return(TRUE)
-}
-
-nums <- 1:25
-
-rbind(nums, sapply(nums, isPrime))
-```
-
-```{r}
-x <- c(TRUE, FALSE, TRUE, FALSE, TRUE)
-sum(x); mean(x); length(x)
-```
-
-```{r}
-naievesum <- function(x) {
-
-   s <- 0
-   n <- length(x)
-
-   for(i in 1:n) {
-      s <- S + x[i]
-   }
-
-   s
-}
-
-naivesum(c(1, 2, 3, 4.5, -6))
-```
-
-```{r}
-pwisesum <- function(x) {
-
-   n <- length(x)
-
-   if(n == 1)
-      return(x)
-
-   m = floor(n/2)
-
-   return(pwisesum(x[1:m]) + pwisesum(x[(m + 1):n]))
-}
-
-pwisesum(c(1, 2, 3, 4.5, -6))
-
-```
-
-```{r}
-kahan_sum <- function(x) {
-   comp <- s <- 0
-   n <- length(x)
-
-   for(i in 1:n) {
-      y <- x[i] - comp
-      t <- x[i] + s
-      comp <- (t - s) - y
-      s <- t
-   }
-
-   return(s)
-}
-
-kahan_sum(1:8)
-```
-
-```{r}
-naive_poly <- function(x, coefs) {
-   y <- rep(0, length(x))
-
-   for(i in 1:length(coefs)) {
-      y <- y + coefs[i] * (x ^ (i - 1))
-   }
-
-   return(y)
-}
-
-# example poly:
-
-# f(x) = (x-5)(x+3)(x+2)(x-1)
-# = x^4 + x^3 - x^2 - 11x + 30
-# = 30 - 11 x - 19 x^2 - x^3 + x^4
-
-f <- c(30, -11, -19, -1, 1)
-
-x <- c(-1, 0, 1)
-
-naive_poly(x, f)
-
-```
-
-```{r}
-better_poly <- function(x, coefs) {
-   y <- rep(0, length(x))
-
-   cached.x <- 1
-
-   for(i in 1:length(coefs)) {
-      y <- y + coefs[i] * cached.x
-      cached.x <- cached.x * x
-   }
-
-   return(y)
-}
-
-better_poly(x, f)
-```
-
-```{r}
-horner <- function(x, coefs) {
-   y <- rep(0, length(x))
-
-   for(i in length(coefs):1)
-      y <- coefs[i] + x * y
-
-   return(y)
-}
-
-horner(x, f)
-```
-
-```{r}
-nth_root <- function(a, n, tol = 1 / 1000) {
-   x <- 1
-   deltax <- tol * 10
-
-   while(abs(deltax) > tol) {
-      deltax <- (1 / n) * (a / x ^ (n - 1) - x)
-      x <- x + deltax
-   }
-
-   return(x)
-}
-
-nth_root(100, 2)
-nth_root(65536, 4)
-nth_root(1000, 3)
-nth_root(pi, 2)
-```
-
-## Exercises
-
-#### 1.)
-
-What is the circumference of a 10-meter-diameter circle symbolically?
-
-$2\pi r$, r = 10
-
-$10\pi$
-
-What is it numerically?
-
-```{r}
-10*pi
-```
-
-Which one is correct? *Both*
-
-#### 2.)
-
-Describe an application for which a numerical solution are preferable to a symbolic solution.
-
-_An exact symbolic solution to some problems can be difficult or near impossible, so a good approximation via numerical methods is preferable than no solution._
-
-Describe an application where symbolic solutions are preferable to numerical.
-
-_An analytical/symbolic solution can solve a related family of problems._
-
-#### 3.)
-
-If an unsigned variable were stored using 16 bits, what is the largest number it could hold?
-
-```{r}
-2^16
-```
-
-#### 4.)
-
-Using R's internal trigonometric functions, implement a secant function.
-
-```{r}
-sec <- function(x) 1/cos(x)
-
-assertthat::are_equal(sec(pi), -1)
-
-```
-
-#### 5.)
-
-Using the exponentiation operator, create a function that finds the nth root of any number.
-
-```{r}
-nth_root <- function(x, n) x^(1/n)
-
-assertthat::are_equal(nth_root(32, 5), 2)
-
-```
-
-#### 6.)
-
-Function 1.1 provides a recursive implementation of the Fibonacci sequence. Rewrite the function to provide a result without recursion.
-
-```{r}
-fib_iter <- function(x) {
-
-   # build look-up
-   fib_values <- list()
-   fib_values[[1]] <- 1
-
-   for(index in 2:x) {
-
-      fib1 <- ifelse(index <= 1, 1, fib_values[[index - 1]])
-      fib2 <- ifelse(index <= 1, 1, ifelse(index == 2, 1, fib_values[[index - 2]]))
-
-      fib_values[index] <- fib1 + fib2
-   }
-
-   return(fib_values[[x]])
-}
-
-sapply(1:10, fib_iter)
-```
-
-What are the relative strenghts and weaknesses of each implementation?
-
-_The recursive version is a bit more readable._
-
-#### 7.)
-
-Implement a product function in R without using recursion.
-
-```{r}
-
-```
-
-#### 8.)
-#### 9.)
-#### 10.)
-#### 11.)
-#### 12.)
-#### 13.)
-#### 14.)
-#### 15.)
-#### 16.)
-#### 17.)
-#### 18.)
-#### 19.)
+##

### 2020-04-06 Matrix Algebra/ma_ch_02_vectors.R

@@ -1,55 +1,0 @@
-# Vectors
-
-
-a <- c(1, 2, 3)
-a
-
-b <- c(4, 5, 6)
-b
-
-c <- matrix(c(3, 2, 1), 3, 1, byrow = T)
-d <- matrix(c(6, 5, 4), nrow = 3, ncol = 1, byrow = T)
-
-c
-d
-
-class(b)
-
-b <- as.matrix(b)
-b
-
-class(b)
-
-u <- matrix(c(3, 2, 1), 1, 3,
-            byrow = T)
-
-v <- matrix(c(6, 5, 4),
-            ncol = 1, nrow = 3, byrow = T)
-
-
-# Matrices
-
-U <- matrix(c(1, 2, 3, 4), 2, 2, byrow = T)
-V <- matrix(c(5, 6, 7, 8), 2, 2, byrow = T)
-W <- matrix(c(2, 2, 3, 5), 2, 2, byrow = T)
-A <- matrix(c(1, 2, 3, 4, 5, 6), 2, 3, byrow = T)
-B <- matrix(c(1, 2, 3, 4, 5, 6), 3, 2, byrow = T)
-
-# Example Operations
-
-U; V; W;
-
-A; B
-
-U + V; 2 * U
-
-A %*% B;B %*% A
-
-U %*% V; V %*% U
-
-U %*% A; U %*% t(B)
-
-U %*% W; W %*% U
-
-V %*% W; W %*% V
-

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_03-05/eda_lab.R

@@ -1,0 +1,205 @@
+library(data.table)
+library(ggplot)
+library(ggtheme)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+
+theme_set(theme_light())
+
+setwd("D:/Projects/MSDS-RiskAnalytics/Module_03")
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+data(SP500, package = "Ecdat")
+
+SPreturn <- SP500$r500
+n <- length(SPreturn)
+year_SP <- 1981 + (1:n) * (1991.25 - 1981) / n
+
+# Visualize S&P500 log returns
+plot(year_SP, SPreturn, main = "S&P 500 daily log returns",
+     xlab = "year", type = "l", ylab = "log return")
+
+ggplot(data.table(date = year_SP, return = SPreturn, direction = ifelse(SPreturn >= 0, "red", "green")), 
+       aes(date, return, color = direction)) +
+  geom_bar(stat = 'identity') +
+  theme(legend.position = "none")
+
+set.seed("991155")
+
+edf_norm <- ecdf(rnorm(150))
+pdf("normalcdfplot.pdf", width = 6, height = 5)
+par(mfrow = c(1, 1))
+plot(edf_norm, verticals = T, do.p = F, main = "EDF and CDF")
+tt = seq(from = -3, to = 3, by = 0.01)
+lines(tt, pnorm(tt),lty = 2, lwd = 2, col = "red")
+legend(1.5, 0.2, c("EDF", "CDF"), lty = c(1, 2),
+       lwd = c(1.5, 2), col = c("black", "red"))
+graphics.off()
+
+dm <- Garch$dm
+diffdm <- diff(dm) # Ducsch mar
+pdf("dm_halfnormal.pdf", width = 7, height = 6)
+halfnorm(abs(diffdm), main = "changes in DM/dollar exchange rate",
+         ylab = "Sorted data")
+graphics.off()
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_03-05/ford_case_study.R

@@ -1,0 +1,126 @@
+library(data.table)
+library(dplyr)
+library(ggplot2)
+library(ggrepel)
+library(GGally)
+library(ggthemes)
+library(scales)
+library(reshape2)
+library(skimr)
+library(gridExtra)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+
+#####################################################################
+######################### Ford Case Study ###########################
+#####################################################################
+
+theme_set(theme_sjplot())
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+path.data <- "D:/Projects/MSDS-RiskAnalytics/datasets"
+setwd(path.data)
+
+# 4.11
+# Ford Case Study
+
+data.ford <- as.data.table(read.csv("ford.csv", header = T))
+
+colnames(data.ford) <- c("n", "date", "return")
+
+head(data.ford)
+
+# E1
+
+mean(data.ford$return)
+sd(data.ford$return)
+median(data.ford$return)
+
+ggplot(data.ford, aes(sample = return)) +
+  geom_qq() +
+  geom_qq_line()
+
+shapiro.test(data.ford$return)
+
+getRetVsT <- function(returns, df_canidates = c(1, 2, 4, 6, 10, 20)) {
+  plots <- lapply(df_canidates, function(df) {
+    
+    n <- length(returns)
+    q_range <- (1:n) / (n+1)
+    
+    data <- data.table(ret = returns, theoretical = qt(q_range, df))
+    data$theoretical <- sort(data$theoretical)
+    
+    model <- lm(qt(c(0.25,0.75), df = df) ~ quantile(data$ret,c(0.25,0.75)))
+    
+    ggplot(data, aes(x = sort(ret), y = theoretical)) +
+      geom_abline(col = 'cornflowerblue', lwd = 1.3, slope = model$coefficients[2], intercept = model$coefficients[1]) +
+      geom_point() +
+      labs(title = paste("df = ", df), 
+           x = "returns",
+           y = "theoretical")
+  })
+  
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_05/Discussion_03.R

@@ -1,48 +1,0 @@
-library(data.table)
-library(ggplot)
-library(ggtheme)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-library(sn)
-
-theme_set(theme_light())
-
-setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "right", legend.title = element_blank())
-
-x <- seq( from = -4, to = 4, by = 0.001)
-y_norm <- dnorm(x)
-y_df1 <- dt(x, df = 1)
-y_df2 <- dt(x, df = 2)
-y_df5 <- dt(x, df = 5)
-y_df10 <- dt(x, df = 10)
-y_df20 <- dt(x, df = 20)
-
-
-p1 <- ggplot(data = data.table(x = x, y = y_norm)) +
-  geom_line(aes(x, y_df1, color = "t-df = 1"), lwd = 1) +
-  geom_line(aes(x, y_df2, color = "t-df = 2"), lwd = 1) +
-  geom_line(aes(x, y_df5, color = "t-df = 5"), lwd = 1) +
-  geom_line(aes(x, y_df10, color = "t-df = 10"), lwd = 1) +
-  geom_line(aes(x, y_df20, color = "t-df = 20"), lwd = 1) +
-  geom_line(aes(x, y_norm, color = "Normal"), lwd = 1) +
-  labs(title = "Normal vs t", x = "", y = "")
-
-p2 <- ggplot(data = data.table(x = x, y = y_norm)) +
-  geom_line(aes(x, y_df1, color = "t-df = 1"), lwd = 1) +
-  geom_line(aes(x, y_df2, color = "t-df = 2"), lwd = 1) +
-  geom_line(aes(x, y_df5, color = "t-df = 5"), lwd = 1) +
-  geom_line(aes(x, y_df10, color = "t-df = 10"), lwd = 1) +
-  geom_line(aes(x, y_df20, color = "t-df = 20"), lwd = 1) +
-  geom_line(aes(x, y_norm, color = "Normal"), lwd = 1) +
-  labs(x = "", y = "", title = "Zoomed Left Tail") +
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_05/earnings_lab.R

@@ -1,85 +1,0 @@
-library(data.table)
-library(ggplot)
-library(ggtheme)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-library(sn)
-
-theme_set(theme_light())
-
-setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-#### Earnings Lab
-
-library("Ecdat")
-
-?CPSch3
-
-data(CPSch3)
-dimnames(CPSch3)[[2]]
-
-male.earnings <- CPSch3[CPSch3[, 3] == "male", 2]
-sqrt.male.earnings <- sqrt(male.earnings)
-log.male.earnings <- log(male.earnings)
-one.fourth.male.earnings = male.earnings^(1/4)
-
-par(mfrow = c(2, 2))
-qqnorm(male.earnings, datax = T, main = "untransformed")
-qqnorm(sqrt.male.earnings, datax = T,
-       main = "square-root transformed")
-qqnorm(log.male.earnings, datax = T,
-       main = "log-transformed")
-qqnorm(one.fourth.male.earnings, datax = T,
-       main = "1/4 power")
-
-par(mfrow = c(2, 2))
-boxplot(male.earnings, datax = T, main = "untransformed")
-boxplot(sqrt.male.earnings, datax = T,
-        main = "square-root transformed")
-boxplot(log.male.earnings, datax = T,
-        main = "log-transformed")
-boxplot(one.fourth.male.earnings, datax = T,
-        main = "1/4 power")
-
-par(mfrow = c(2, 2))
-plot(density(male.earnings), main = "untransformed")
-plot(density(sqrt.male.earnings),
-     main = "square-root transformed")
-plot(density(log.male.earnings),
-     main = "log-transformed")
-plot(density(one.fourth.male.earnings),
-     main = "1/4 power")
-
-
-library(MASS)
-par(mfrow = c(1, 1))
-boxcox(male.earnings ~ 1)
-boxcox(male.earnings ~ 1, lambda = seq(0.3, 0.45, 1 / 100))
-
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_05/eda_lab.R

@@ -1,205 +1,0 @@
-library(data.table)
-library(ggplot)
-library(ggtheme)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-
-theme_set(theme_light())
-
-setwd("D:/Projects/MSDS-RiskAnalytics/Module_03")
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-data(SP500, package = "Ecdat")
-
-SPreturn <- SP500$r500
-n <- length(SPreturn)
-year_SP <- 1981 + (1:n) * (1991.25 - 1981) / n
-
-# Visualize S&P500 log returns
-plot(year_SP, SPreturn, main = "S&P 500 daily log returns",
-     xlab = "year", type = "l", ylab = "log return")
-
-ggplot(data.table(date = year_SP, return = SPreturn, direction = ifelse(SPreturn >= 0, "red", "green")), 
-       aes(date, return, color = direction)) +
-  geom_bar(stat = 'identity') +
-  theme(legend.position = "none")
-
-set.seed("991155")
-
-edf_norm <- ecdf(rnorm(150))
-pdf("normalcdfplot.pdf", width = 6, height = 5)
-par(mfrow = c(1, 1))
-plot(edf_norm, verticals = T, do.p = F, main = "EDF and CDF")
-tt = seq(from = -3, to = 3, by = 0.01)
-lines(tt, pnorm(tt),lty = 2, lwd = 2, col = "red")
-legend(1.5, 0.2, c("EDF", "CDF"), lty = c(1, 2),
-       lwd = c(1.5, 2), col = c("black", "red"))
-graphics.off()
-
-dm <- Garch$dm
-diffdm <- diff(dm) # Ducsch mar
-pdf("dm_halfnormal.pdf", width = 7, height = 6)
-halfnorm(abs(diffdm), main = "changes in DM/dollar exchange rate",
-         ylab = "Sorted data")
-graphics.off()
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_05/equityReturns_lab.R

@@ -1,53 +1,0 @@
-library(data.table)
-library(ggplot)
-library(ggtheme)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-library(sn)
-
-theme_set(theme_light())
-
-setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-#### Equity Returns Lab
-
-berndtInvest <- read.csv("berndtInvest.csv")
-Berndt <- as.matrix(berndtInvest[, 2:5])
-
-cov(Berndt)
-cor(Berndt)
-
-pairs(Berndt)
-
-library(MASS)
-library(mnormt)
-
-df <- seq(2.5, 8, 0.1)
-n <- length(df)
-
-loglik_profile <- rep(0, n)
-for(i in 1:n)
-{
-  fit <- cov.trob(Berndt, nu = df[i])
-  mu <-  as.vector(fit$center)
-  sigma <- matrix(fit$cov, nrow = 4)
-  loglik_profile[i] <- sum(log(dmt(Berndt, mean = fit$center,
-                                   S = fit$cov, df = df[i])))
-}
-
-aic_t <- -max(2 * loglik_profile) + 2 * (4 + 10 + 1)
-z1 <- ( 2 * loglik_profile > 2 * max(loglik_profile) - qchisq(0.95, 1) )
-plot(df, 2 * loglik_profile, type = "l", cex.axis = 1.5,
-     cex.lab = 1.5, ylab = "2 * loglikelihood - 64,000", lwd = 2)
-abline(h = 2 * max(loglik_profile) - qchisq(0.95, 1) - 64000)
-abline(h = 2 * max(loglik_profile) - 64000)
-abline(v = (df[16] + df[17]) / 2)
-abline(v = (df[130] + df[131]) / 2)

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_05/ford_case_study.R

@@ -1,126 +1,0 @@
-library(data.table)
-library(dplyr)
-library(ggplot2)
-library(ggrepel)
-library(GGally)
-library(ggthemes)
-library(scales)
-library(reshape2)
-library(skimr)
-library(gridExtra)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-
-#####################################################################
-######################### Ford Case Study ###########################
-#####################################################################
-
-theme_set(theme_sjplot())
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-path.data <- "D:/Projects/MSDS-RiskAnalytics/datasets"
-setwd(path.data)
-
-# 4.11
-# Ford Case Study
-
-data.ford <- as.data.table(read.csv("ford.csv", header = T))
-
-colnames(data.ford) <- c("n", "date", "return")
-
-head(data.ford)
-
-# E1
-
-mean(data.ford$return)
-sd(data.ford$return)
-median(data.ford$return)
-
-ggplot(data.ford, aes(sample = return)) +
-  geom_qq() +
-  geom_qq_line()
-
-shapiro.test(data.ford$return)
-
-getRetVsT <- function(returns, df_canidates = c(1, 2, 4, 6, 10, 20)) {
-  plots <- lapply(df_canidates, function(df) {
-    
-    n <- length(returns)
-    q_range <- (1:n) / (n+1)
-    
-    data <- data.table(ret = returns, theoretical = qt(q_range, df))
-    data$theoretical <- sort(data$theoretical)
-    
-    model <- lm(qt(c(0.25,0.75), df = df) ~ quantile(data$ret,c(0.25,0.75)))
-    
-    ggplot(data, aes(x = sort(ret), y = theoretical)) +
-      geom_abline(col = 'cornflowerblue', lwd = 1.3, slope = model$coefficients[2], intercept = model$coefficients[1]) +
-      geom_point() +
-      labs(title = paste("df = ", df), 
-           x = "returns",
-           y = "theoretical")
-  })
-  
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_05/mle_lab.R

@@ -1,200 +1,0 @@
-library(data.table)
-library(ggplot)
-library(ggtheme)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-library(sn)
-
-theme_set(theme_light())
-
-setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-#### Maximum Likelihood Lab
-
-data(Capm, package = "Ecdat")
-x <- diff(Capm$rf)
-fitdistr(x, "t")
-
-# classical t-distribution
-
-n <- length(x)
-start <- c(mean(x), sd(x), 5)
-loglik_t <- function(beta) sum( - dt((x - beta[1]) / beta[2],
-                                     beta[3], log = T) + log(beta[2]))
-fit_t <- optim(start, loglik_t, hessian = T,
-               method = "L-BFGS-B", lower = c(-1, 0.001, 1))
-AIC_t <- 2 * fit_t$value + 2 * 3
-BIC_t <- 2 * fit_t$value + log(n) * 3
-sd_t <- sqrt(diag(solve(fit_t$hessian)))
-
-fit_t$par
-sd_t
-AIC_t
-BIC_t
-
-# standardized t-distribution
-
-loglik_std <- function(beta) sum( - dstd(x, mean = beta[1],
-                                         sd = beta[2], nu = beta[3], log = T))
-fit_std <- optim(start, loglik_std, hessian = T,
-                 method = "L-BFGS-B", lower = c(-0.1, 0.01, 2.1))
-AIC_std <- 2 * fit_std$value + 2 * 3
-BIC_std <- 2 * fit_std$value + log(n) * 3
-sd_std <- sqrt(diag(solve(fit_std$hessian)))
-
-fit_std$par
-sd_std
-AIC_std
-BIC_std
-
-# F-S skewed t-distribution
-
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_06-08/copula_example.R

@@ -1,0 +1,57 @@
+library(data.table)
+library(ggplot)
+library(ggtheme)
+library(Ecdat)
+library(faraway)
+library(fGarch)
+library(sn)
+library(WVPlots)
+
+theme_set(theme_light())
+
+setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+
+n <- 10000
+data <- data.table( x = runif(n, 0, 1) )
+
+ggplot(data) +
+  geom_histogram(aes(x, fill = ..count..))
+
+data$transformed <- qnorm(data$x, 10, 3) 
+data$y <- dnorm(data$x, 10, 3)
+
+ggplot(data, aes(x = transformed)) +
+  geom_histogram(aes(y = ..density.. , fill = ..count..)) + 
+  geom_density(aes( y = ..density.. ), color = "green", lwd = .9) +
+  theme(legend.position = "none")
+
+ggplot(data, aes(x, transformed )) +
+  geom_line(color = "cornflowerblue", lwd = 1)
+
+ScatterHist(data, "x", "transformed", title = "U -> N")
+
+n <- 10000
+df <- 5
+
+data <- data.table( x1 = rt(n, df), x2 = rt(n, df) )
+
+ggplot(data, aes(x = x1)) +
+  geom_histogram(aes(y = ..density.. , fill = ..count..)) + 
+  geom_density(aes( x = x2, y = ..density.. ), color = "green", lwd = .9) +
+  geom_density(aes(x = x2, y = ..density.. ), color = "green", lwd = .9) +
+  theme(legend.position = "none")
+
+ggplot(data) +
+  geom_density(aes( x = x1, y = ..density.. ), color = "red", lwd = .9) +
+  geom_density(aes(x = x2, y = ..density.. ), color = "green", lwd = .9) +
+  theme(legend.position = "none")
+
+ScatterHist(data, "x1", "x2", title = "", contour = F)

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_06-08/copula_lab.R

@@ -1,0 +1,83 @@
+library(copula)
+
+u <- seq(0.000001, 1, length = 500)
+frank <- iPsi(copula = archmCopula(family = "frank", param = 1), u)
+plot(u, frank, type = "l", lwd = 3, ylab = expression(phi(u)))
+abline(h = 0); abline(v = 0)
+
+set.seed(5640)
+
+theta <- c(-100, -50, -10, -1, 0, 5, 20, 50, 500)
+par(mfrow = c(3,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
+
+for(i in 1:9) {
+  U <- rCopula(n = 200,
+               copula = archmCopula(family = "frank", param = theta[i]))
+  plot(U, xlab = expression(u[1]), ylab = expression(u[2]),
+       main = eval(substitute(expression(paste(theta, " = ", j)),
+                              list(j = as.character(theta[i])))))
+}
+
+set.seed(5640)
+
+theta <- c(-0.98, -0.7, -0.3, -0.1, 0.1, 1, 1.5, 15, 100)
+par(mfrow = c(3,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
+
+for(i in 1:9) {
+  U <- rCopula(n = 200,
+               copula = archmCopula(family = "clayton", param = theta[i]))
+  plot(U, xlab = expression(u[1]), ylab = expression(u[2]),
+       main = eval(substitute(expression(paste(theta, " = ", j)),
+                              list(j = as.character(theta[i])))))
+}
+
+set.seed(5640)
+
+theta <- c(1.1, 1.5, 2, 4, 8, 50)
+par(mfrow = c(2,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
+
+for(i in 1:6) {
+  U <- rCopula(n = 200,
+               copula = archmCopula(family = "gumbel", param = theta[i]))
+  plot(U, xlab = expression(u[1]), ylab = expression(u[2]),
+       main = eval(substitute(expression(paste(theta, " = ", j)),
+                              list(j = as.character(theta[i])))))
+}
+
+set.seed(5640)
+
+theta <- c(1.1, 1.5, 2, 4, 8, 50)
+par(mfrow = c(2,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_08/copula_example.R

@@ -1,57 +1,0 @@
-library(data.table)
-library(ggplot)
-library(ggtheme)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-library(sn)
-library(WVPlots)
-
-theme_set(theme_light())
-
-setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-
-n <- 10000
-data <- data.table( x = runif(n, 0, 1) )
-
-ggplot(data) +
-  geom_histogram(aes(x, fill = ..count..))
-
-data$transformed <- qnorm(data$x, 10, 3) 
-data$y <- dnorm(data$x, 10, 3)
-
-ggplot(data, aes(x = transformed)) +
-  geom_histogram(aes(y = ..density.. , fill = ..count..)) + 
-  geom_density(aes( y = ..density.. ), color = "green", lwd = .9) +
-  theme(legend.position = "none")
-
-ggplot(data, aes(x, transformed )) +
-  geom_line(color = "cornflowerblue", lwd = 1)
-
-ScatterHist(data, "x", "transformed", title = "U -> N")
-
-n <- 10000
-df <- 5
-
-data <- data.table( x1 = rt(n, df), x2 = rt(n, df) )
-
-ggplot(data, aes(x = x1)) +
-  geom_histogram(aes(y = ..density.. , fill = ..count..)) + 
-  geom_density(aes( x = x2, y = ..density.. ), color = "green", lwd = .9) +
-  geom_density(aes(x = x2, y = ..density.. ), color = "green", lwd = .9) +
-  theme(legend.position = "none")
-
-ggplot(data) +
-  geom_density(aes( x = x1, y = ..density.. ), color = "red", lwd = .9) +
-  geom_density(aes(x = x2, y = ..density.. ), color = "green", lwd = .9) +
-  theme(legend.position = "none")
-
-ScatterHist(data, "x1", "x2", title = "", contour = F)

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_08/copula_lab.R

@@ -1,83 +1,0 @@
-library(copula)
-
-u <- seq(0.000001, 1, length = 500)
-frank <- iPsi(copula = archmCopula(family = "frank", param = 1), u)
-plot(u, frank, type = "l", lwd = 3, ylab = expression(phi(u)))
-abline(h = 0); abline(v = 0)
-
-set.seed(5640)
-
-theta <- c(-100, -50, -10, -1, 0, 5, 20, 50, 500)
-par(mfrow = c(3,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
-
-for(i in 1:9) {
-  U <- rCopula(n = 200,
-               copula = archmCopula(family = "frank", param = theta[i]))
-  plot(U, xlab = expression(u[1]), ylab = expression(u[2]),
-       main = eval(substitute(expression(paste(theta, " = ", j)),
-                              list(j = as.character(theta[i])))))
-}
-
-set.seed(5640)
-
-theta <- c(-0.98, -0.7, -0.3, -0.1, 0.1, 1, 1.5, 15, 100)
-par(mfrow = c(3,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
-
-for(i in 1:9) {
-  U <- rCopula(n = 200,
-               copula = archmCopula(family = "clayton", param = theta[i]))
-  plot(U, xlab = expression(u[1]), ylab = expression(u[2]),
-       main = eval(substitute(expression(paste(theta, " = ", j)),
-                              list(j = as.character(theta[i])))))
-}
-
-set.seed(5640)
-
-theta <- c(1.1, 1.5, 2, 4, 8, 50)
-par(mfrow = c(2,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
-
-for(i in 1:6) {
-  U <- rCopula(n = 200,
-               copula = archmCopula(family = "gumbel", param = theta[i]))
-  plot(U, xlab = expression(u[1]), ylab = expression(u[2]),
-       main = eval(substitute(expression(paste(theta, " = ", j)),
-                              list(j = as.character(theta[i])))))
-}
-
-set.seed(5640)
-
-theta <- c(1.1, 1.5, 2, 4, 8, 50)
-par(mfrow = c(2,3), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_08/lab_bootstrap.R

@@ -1,45 +1,0 @@
-library(bootstrap)
-library(MASS)
-
-set.seed("3857")
-
-data(CRSPday, package = "Ecdat")
-
-ge <- CRSPday[, 4]
-
-getRetVsT <- function(returns, df_canidates = c(1, 2, 4, 6, 10, 20)) {
-  plots <- lapply(df_canidates, function(df) {
-    
-    n <- length(returns)
-    q_range <- (1:n) / (n+1)
-    
-    data <- data.table(ret = returns, theoretical = qt(q_range, df))
-    data$theoretical <- sort(data$theoretical)
-    
-    model <- lm(qt(c(0.25,0.75), df = df) ~ quantile(data$ret,c(0.25,0.75)))
-    
-    ggplot(data, aes(x = sort(ret), y = theoretical)) +
-      geom_abline(col = 'cornflowerblue', lwd = 1.3, slope = model$coefficients[2], intercept = model$coefficients[1]) +
-      geom_point() +
-      labs(title = paste("df = ", df), 
-           x = "returns",
-           y = "theoretical")
-  })
-  
-  do.call(grid.arrange, c(plots, top = "QQ-Plot: returns vs t-distribution"))
-}
-
-getRetVsT(ge)
-
-nboot <- 1000
-
-t_mle <- function(x) { as.vector(
-  fitdistr(x, "t", start = list( m = mean(x), s = sd(x), df = 3) )$estimate
-)}
-
-results <- bootstrap(ge, nboot, t_mle)
-
-rowMeans(results$thetastar[ , ])
-apply(results$thetastar[,], 1, sd)
-
-fitdistr(ge, "t")

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_08/lab_flows.R

@@ -1,130 +1,0 @@
-library(data.table)
-library(ggplot)
-library(ggtheme)
-library(Ecdat)
-library(faraway)
-library(fGarch)
-library(sn)
-library(copula)
-
-theme_set(theme_light())
-
-setwd("D:/Projects/MSDS-RiskAnalytics/datasets")
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-dat <- read.csv("FlowData.csv")
-dat <- dat / 1000
-
-n <- nrow(dat)
-
-x1 <- dat$Flow1
-fit1 <- st.mple(matrix(1, n, 1), y = x1, dp = c(mean(x1), sd(x1), 0, 10))
-est1 <- fit1$dp
-u1 <- pst(x1, dp = est1)
-
-x2 <- dat$Flow2
-fit2 <- st.mple(matrix(1, n, 1), y = x2, dp = c(mean(x2), sd(x2), 0, 10))
-est2 <- fit2$dp
-
-u2 <- pst(x2, dp = est2)
-
-U.hat <- cbind(u1, u2)
-
-z1 <- qnorm(u1)
-z2 <- qnorm(u2)
-
-Z.hat <- cbind(z1, z2)
-
-library(ks)
-
-fhatU <- kde(x = U.hat, H = Hscv(x=U.hat))
-par(mfrow = c(2, 2), cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)
-hist(u1, main = "(a)", xlab = expression(hat(U)[1]), freq = F)
-hist(u2, main = "(b)", xlab = expression(hat(U)[2]), freq = F)
-plot(u1, u2, main = "(c)", xlab = expression(hat(U)[1]),
-     ylab = expression(hat(U)[2]), mgp = c(2.5, 1, 0))
-plot(fhatU, drawpoints = F, drawlabels = F,
-     cont = seq(10, 80, 10), main = "(d)", xlab = expression(hat(U)[1]),
-     ylab = expression(hat(U)[2]), mgp = c(2.5, 1, 0))
-
-
-cor.test(u1, u2, method = "spearman")
-cor.test(u1, u2, method = "kendall")
-sin(-0.242*pi/2)
-cor.test(u1, u2, method = "pearson")
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_11-14/SDAFE2.R

@@ -1,0 +1,757 @@
+###########  Additional R functions for  #################################################
+###########  Statistics and Data Analysis for Financial Engineering, 2nd Edition #########
+###########  by Ruppert and Matteson  ####################################################
+
+require(fGarch) 
+
+#######################################
+#### multivariate Ljung--Box tests ####
+#######################################
+
+"mq" = function(x, lag = 1, df.adj = 0){
+  # Compute multivariate Ljung-Box test statistics
+  #
+  x = as.matrix(x)
+  nr = dim(x)[1]
+  nc = dim(x)[2]
+  g0 = var(x)
+  ginv = solve(g0)
+  qm = 0.0
+  df = 0
+  out = matrix(0,lag,4)
+  for (i in 1:lag){
+    x1 = x[(i+1):nr,]
+    x2 = x[1:(nr-i),]
+    g = cov(x1,x2)
+    g = g*(nr-i-1)/(nr-1)
+    h = t(g)%*%ginv%*%g%*%ginv
+    qm = qm+nr*nr*sum(diag(h))/(nr-i)
+    df = df+nc*nc
+    pv = 1-pchisq(qm,df-df.adj)
+    #  print(c(i,qm,pv))
+    out[i,] = c(i,round(qm,2),round(df-df.adj),round(pv,3))
+  }
+  output = as.data.frame(matrix(out[lag,],1,4))
+  names(output) = c("K", "Q(K)", "d.f.", "p-value")                         
+  print(output)
+}
+
+mLjungBox = mq
+
+"mq0" = function(A, m){
+  N = dim(A)[1]
+  k = dim(A)[2]
+  temp = numeric(m)
+  pval = numeric(m)
+  out = as.data.frame(matrix(0,m+1,4))
+  names(out) = c("K", "Q(K)", "d.f.", "p-value")
+  out[,1] = 0:m
+  
+  Q.temp = N*sum(cor(A)[lower.tri(cor(A), diag = FALSE)]^2)
+  out[1,2] = Q.temp 
+  df.temp = k*(k-1)/2
+  out[1,3] = df.temp
+  out[1,4] = 1-pchisq(Q.temp, df.temp)
+  
+  for(j in 1:m){
+    ccf = cor(A[-(1:j),],A[-((N-j+1):N),])
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_12/time_series_lab.R

@@ -1,364 +1,0 @@
-library(Ecdat)
-library(tseries)
-library(forecast)
-
-setwd("D:/Projects/MSDS-RiskAnalytics/Module_05")
-
-data(Mishkin, package = "Ecdat")
-
-y <- as.vector(Mishkin[, 1])
-
-par(mfrow = c(1,2))
-acf(y)
-acf(diff(y))
-
-Box.test(diff(y), lag = 10, type = "Ljung-Box")
-
-data(bmw, package = "evir")
-
-Box.test(bmw, lag = 5, type = "Ljung-Box")
-
-fitAR1 <- arima(bmw, order = c(1, 0, 0))
-summary(fitAR1)
-
-acf(bmw)
-
-Box.test(residuals(fitAR1), lag = 5, type = "Ljung-Box", fitdf = 1)
-
-fit <- arima(y, order = c(1, 0, 0))
-Box.test(fit$residuals, type = "Ljung", lag = 24, fitdf = 1)
-
-library(forecast)
-
-auto.arima(diff(y), max.p = 20, max.q = 0, d = 0, ic = "aic")
-
-auto.arima(diff(y), max.p = 20, max.q = 0, d = 0, ic = "bic")
-
-par(mfrow = c(1,1))
-plot.ts(y)
-
-fitMA3 <- arima(diff(y), order = c(0, 0, 3))
-fitMA3
-
-set.seed(4631)
-y1 <- arima.sim(n = 500, list(ar = c(0.4)))
-y2 <- cumsum(y1)
-y3 <- cumsum(y2)
-
-par(mfrow = c(3,1))
-plot.ts(y1)
-plot.ts(y2)
-plot.ts(y3)
-
-adf.test(y)
-pp.test(y)
-
-kpss.test(y)
-
-?adf.test
-
-x <- rnorm(1000)  # no unit-root
-adf.test(x)
-
-y <- diffinv(x)   # contains a unit-root
-adf.test(y)
-
-TT <- 100
-wn <- rnorm(TT) # white noise
-tseries::adf.test(wn)
-
-tseries::adf.test(wn, k=0)
-
-intercept <- 1
-wnt <- wn + 1:TT + intercept
-tseries::adf.test(wnt)
-
-rw <- cumsum(rnorm(TT))
-tseries::adf.test(rw)
-
-tseries::adf.test(rw, k=0)
-
-pdf("inflation_AR1_acf.pdf",width=9,height=5)
-#
-par(mfrow=c(1,2))
-acf(y,main="Inflation rate")
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_14/GARCH_lab.R

@@ -1,725 +1,0 @@
-library(xts)
-library(ggplot2)
-library(ggthemes)
-
-theme_set(theme_light())
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
-
-setwd("D:/Projects/MSDS-RiskAnalytics/Module_06")
-
-data(Mishkin,package="Ecdat")
-infl= as.vector(Mishkin[,1])  # pai1 = one-month inflation rate (in percent, annual rate) 
-
-data(SP500,package="Ecdat")
-SPreturn = SP500$r500
-
-library(evir)  # for emplot
-data(Garch,package="Ecdat")
-attach(Garch)
-
-data(Capm,package="Ecdat")
-difflogrf=diff(log(Capm$rf))
-
-diffdm = diff(dm)  #  Deutsch mark
-diffbp = diff(bp)  #  British pound
-diffcd = diff(cd)  #  Canadian dollar
-diffdy = diff(dy)  #  Japanese yen
-
-n1  = length(SPreturn)
-year_SP = 1981 + (1:n1)* (1991.25-1981)/n1
-d1 = seq(as.Date("1981-01-01"), as.Date("1991-04-01"), by=1)
-date_SP = d1[seq(1,length(d1),by=length(d1)/n1)]
-
-n2 = length(diffdm)
-year_dm = 1980 + (1:n2)* (1987.5-1980)/n2
-d2 = seq(as.Date("1980-01-01"), as.Date("1987-07-01"), by=1)
-date_dm = d2[seq(1,length(d2),by=length(d2)/n2)]
-
-n3 = length(difflogrf)
-year_rf = 1960 + (1:n3) * (2003 - 1960)/n3
-d3 = seq(as.Date("1960-01-01"), as.Date("2003-01-01"), by=1)
-date_rf = d3[seq(1,length(d3),by=length(d3)/n3)]
-
-n4 = length(infl)
-year_infl = 1950+1/12 + (1:n4) * (1991-1950-1/12)/n4
-d4 = seq(as.Date("1950-02-01"), as.Date("1991-01-01"), by=1)
-date_infl = d4[seq(1,length(d4),by=length(d4)/n4)]
-
-
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_14/SDAFE2.R

@@ -1,757 +1,0 @@
-###########  Additional R functions for  #################################################
-###########  Statistics and Data Analysis for Financial Engineering, 2nd Edition #########
-###########  by Ruppert and Matteson  ####################################################
-
-require(fGarch) 
-
-#######################################
-#### multivariate Ljung--Box tests ####
-#######################################
-
-"mq" = function(x, lag = 1, df.adj = 0){
-  # Compute multivariate Ljung-Box test statistics
-  #
-  x = as.matrix(x)
-  nr = dim(x)[1]
-  nc = dim(x)[2]
-  g0 = var(x)
-  ginv = solve(g0)
-  qm = 0.0
-  df = 0
-  out = matrix(0,lag,4)
-  for (i in 1:lag){
-    x1 = x[(i+1):nr,]
-    x2 = x[1:(nr-i),]
-    g = cov(x1,x2)
-    g = g*(nr-i-1)/(nr-1)
-    h = t(g)%*%ginv%*%g%*%ginv
-    qm = qm+nr*nr*sum(diag(h))/(nr-i)
-    df = df+nc*nc
-    pv = 1-pchisq(qm,df-df.adj)
-    #  print(c(i,qm,pv))
-    out[i,] = c(i,round(qm,2),round(df-df.adj),round(pv,3))
-  }
-  output = as.data.frame(matrix(out[lag,],1,4))
-  names(output) = c("K", "Q(K)", "d.f.", "p-value")                         
-  print(output)
-}
-
-mLjungBox = mq
-
-"mq0" = function(A, m){
-  N = dim(A)[1]
-  k = dim(A)[2]
-  temp = numeric(m)
-  pval = numeric(m)
-  out = as.data.frame(matrix(0,m+1,4))
-  names(out) = c("K", "Q(K)", "d.f.", "p-value")
-  out[,1] = 0:m
-  
-  Q.temp = N*sum(cor(A)[lower.tri(cor(A), diag = FALSE)]^2)
-  out[1,2] = Q.temp 
-  df.temp = k*(k-1)/2
-  out[1,3] = df.temp
-  out[1,4] = 1-pchisq(Q.temp, df.temp)
-  
-  for(j in 1:m){
-    ccf = cor(A[-(1:j),],A[-((N-j+1):N),])
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_15-16/Rlab.R

@@ -1,0 +1,241 @@
+#
+# Written by:
+# -- 
+# John L. Weatherwax                2009-04-21
+# 
+# email: wax@alum.mit.edu
+# 
+# Please send comments and especially bug reports to the
+# above email address.
+# 
+#-----
+
+if( !require('linprog') ){
+    install.packages('linprog', dependencies=TRUE, repos='http://cran.rstudio.com/')
+}
+library(linprog)
+
+save_plots = FALSE
+
+set.seed(0)
+
+src.dir <- "D:/Projects/MSDS-RiskAnalytics/Module_08/"
+
+setwd(src.dir)
+
+source('util_fns.R')
+
+# Test the function "efficient_frontier" using the book's example on EPage 478:
+#
+library(Ecdat)
+data(CRSPday)
+R = 100*CRSPday[,4:6] # in percents
+mean_vect = apply(R, 2, mean)
+cov_mat = cov(R)
+sd_vect = sqrt(diag(cov_mat))
+
+# The target portfolio returns:
+muP = seq(0.05, 0.14, length.out=300)
+mufree = 1.3/253 
+
+result = efficient_frontier(R, muP, mufree)
+
+mean_vect = result$mean_vect
+sd_vect = result$sd_vect
+weights = result$weights
+ind_ms = result$max_sharpe_ind
+
+# Plot our results:
+#
+plot_efficient_frontier(result, "Duplicate Fig. 16.3")
+
+print(weights[ind_ms, ]) # print the weights of the tangency portfolio
+
+text(sd_vect[1], mean_vect[1], 'GE', cex=1.5)
+text(sd_vect[2], mean_vect[2], 'IBM', cex=1.5)
+text(sd_vect[3], mean_vect[3], 'Mobil', cex=1.5)
+
+
+
+# Test the function "maximize_expected_utility" using the book's example on EPage 487:
+#
+dat = read.csv("../../BookCode/Data/Stock_Bond.csv", header=TRUE)
+y = dat[, c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21)]
+n = dim(y)[1]
+m = dim(y)[2]
+r= y[-1,] / y[-n, ] - 1
+
+nlambda = 250
+loglambda_vect = seq(2, 8, length=nlambda)
+
+meu_result = maximize_expected_utility(r, loglambda_vect)
+
+
+# This code reproduces the output of Fig. 16.8:
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_16/Rlab.R

@@ -1,241 +1,0 @@
-#
-# Written by:
-# -- 
-# John L. Weatherwax                2009-04-21
-# 
-# email: wax@alum.mit.edu
-# 
-# Please send comments and especially bug reports to the
-# above email address.
-# 
-#-----
-
-if( !require('linprog') ){
-    install.packages('linprog', dependencies=TRUE, repos='http://cran.rstudio.com/')
-}
-library(linprog)
-
-save_plots = FALSE
-
-set.seed(0)
-
-src.dir <- "D:/Projects/MSDS-RiskAnalytics/Module_08/"
-
-setwd(src.dir)
-
-source('util_fns.R')
-
-# Test the function "efficient_frontier" using the book's example on EPage 478:
-#
-library(Ecdat)
-data(CRSPday)
-R = 100*CRSPday[,4:6] # in percents
-mean_vect = apply(R, 2, mean)
-cov_mat = cov(R)
-sd_vect = sqrt(diag(cov_mat))
-
-# The target portfolio returns:
-muP = seq(0.05, 0.14, length.out=300)
-mufree = 1.3/253 
-
-result = efficient_frontier(R, muP, mufree)
-
-mean_vect = result$mean_vect
-sd_vect = result$sd_vect
-weights = result$weights
-ind_ms = result$max_sharpe_ind
-
-# Plot our results:
-#
-plot_efficient_frontier(result, "Duplicate Fig. 16.3")
-
-print(weights[ind_ms, ]) # print the weights of the tangency portfolio
-
-text(sd_vect[1], mean_vect[1], 'GE', cex=1.5)
-text(sd_vect[2], mean_vect[2], 'IBM', cex=1.5)
-text(sd_vect[3], mean_vect[3], 'Mobil', cex=1.5)
-
-
-
-# Test the function "maximize_expected_utility" using the book's example on EPage 487:
-#
-dat = read.csv("../../BookCode/Data/Stock_Bond.csv", header=TRUE)
-y = dat[, c(3, 5, 7, 9, 11, 13, 15, 17, 19, 21)]
-n = dim(y)[1]
-m = dim(y)[2]
-r= y[-1,] / y[-n, ] - 1
-
-nlambda = 250
-loglambda_vect = seq(2, 8, length=nlambda)
-
-meu_result = maximize_expected_utility(r, loglambda_vect)
-
-
-# This code reproduces the output of Fig. 16.8:
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_16/portfolio_selection_lab.R

@@ -1,559 +1,0 @@
-library(data.table)
-library(xts)
-library(ggplot2)
-library(ggthemes)
-
-theme_set(theme_light())
-
-theme_set(theme_light())
-
-# Theme Overrides
-theme_update(axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
-             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
-             legend.position = "right", legend.title.align=0.5,
-             panel.border = element_rect(linetype = "solid", 
-                                         colour = "lightgray"), 
-             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))
-
-data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
-
-setwd("D:/Projects/MSDS-RiskAnalytics/Module_08")
-
-################################################################
-########## Code for figure 16.1   ##############################
-################################################################
-
-mu1 = 0.14
-mu2 = 0.08
-sig1 = 0.2
-sig2 = 0.15
-rho = 0
-rf = 0.06
-w = seq(0, 1, len = 500)
-means = 0.08 + 0.06 * w
-var = sig1^2 * w^2 + sig2^2 * (1 - w)^2
-risk = sqrt(var)
-ind = !(risk > min(risk))
-ind2 = (means > means[ind])
-wt = 0.693
-meant = 0.08 + 0.06 * wt
-riskt = sqrt(sig1^2 * wt^2 + sig2^2 * (1 - wt)^2)
-
-wp = 0.475
-meanp = 0.08 + 0.06 * wp
-riskp = sqrt(sig1^2 * wp^2 + sig2^2 * (1 - wp)^2)
-
-
-pdf("portfolioNew.pdf", width = 6, height = 5)
-
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_16/util_fns.R

@@ -1,205 +1,0 @@
-#
-# Written by:
-# -- 
-# John L. Weatherwax                2009-04-21
-# 
-# email: wax@alum.mit.edu
-# 
-# Please send comments and especially bug reports to the
-# above email address.
-#
-# Computes the Efficient Frontier.
-#
-#-----
-
-library(quadprog)
-library(linprog)
-
-efficient_frontier = function(returns, muP, mufree=0.0, w_lower_limit=-Inf, w_upper_limit=+Inf){
-  #
-  # w_lower_limit and w_upper_limit (if finite) are bounds on the porfolio weights such that
-  # 
-  # w_lower_limit <= w <= w_upper_limit
-  #
-  # Note: if solve.QP gives errors about convergence one can try a more restricted range for the values of muP:
-  #
-  # muP = seq(min(mean_vect), max(mean_vect), length.out=500)
-  #
-  #--
-  n_stocks = dim(returns)[2]
-  
-  # Extract individual equity (mean return, std return) values: 
-  mean_vect = apply(returns, 2, mean)
-  cov_mat = cov(returns)
-  sd_vect = sqrt(diag(cov_mat))
-  
-  # Portfolio weight constraints:
-  #
-  #     first condition enforce \sum w_j = 1
-  #     second condition enforce \sum mu_j w_j = mu_target
-  #             
-  Amat = cbind(rep(1,n_stocks),mean_vect)
-  bvec = c(1,NaN)
-  if( is.finite(w_lower_limit) ){
-    Amat = cbind(Amat, diag(1, nrow=n_stocks))
-    bvec = c(bvec, w_lower_limit*rep(1,n_stocks))
-  }
-  if( is.finite(w_upper_limit) ){
-    Amat = cbind(Amat, -diag(1, nrow=n_stocks))
-    bvec = c(bvec, -w_upper_limit*rep(1,n_stocks))
-  }
-
-  # storage for results: 
-  sdP = muP
-  weights = matrix(0, nrow=length(muP), ncol=n_stocks)
-  
-  print(bvec)
-  print(Amat)
-  
-  # find the optimal portfolios for each target expected return:
-  #
-  for( i in 1:length(muP) ){
-    bvec[2] = +muP[i] # enforce portfolio mean constraint 
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_17-18/FF_alt.R

@@ -1,0 +1,35 @@
+# Stock/Bond/FX data.
+stocks <- as.data.table(read.csv(paste0(data.dir, "Stock_FX_Bond_2004_to_2005.csv"), 
+                                 header=T))
+stocks$Date <- as.Date(stocks$Date, format = "%d-%b-%y")
+stocks_subset <- stocks[, .(Date, GM_AC, F_AC, UTX_AC, MRK_AC)]
+stocks_diff <- data.table(Date = stocks_subset[-1]$Date, 
+                          apply(log(stocks_subset[, .(GM_AC, F_AC, UTX_AC, MRK_AC)]), 2, diff))
+
+# Fama-French data.
+FF_data <- as.data.table(read.table(paste0(data.dir, "FamaFrenchDaily.txt"), 
+                                    header=T))
+
+stocks = read.csv(paste0(data.dir, "Stock_FX_Bond_2004_to_2005.csv"), header=T)
+
+FF_data = read.table(paste0(data.dir, "FamaFrenchDaily.txt"), header=T)
+
+start_index = which( FF_data$date == "20040102" )
+end_index = which( FF_data$date == "20051230" )
+FF_data = FF_data[ start_index:end_index, ]
+FF_data = FF_data[-1,]     # delete first row since stocks_diff loses a row due to differencing
+
+start_index = which( stocks$DATE == "1/2/2004" )
+end_index = which( stocks$DATE == "12/30/2005" )
+stocks = stocks[ start_index:end_index, ]
+
+stocks_subset = as.data.frame(cbind(stocks$GM_AC,stocks$F_AC,stocks$UTX_AC,stocks$MRK_AC))
+
+stocks_diff = as.data.frame(100*apply(log(stocks_subset), 2, diff) - FF_data$RF) # Excess returns
+
+names(stocks_diff) = c("GM", "Ford", "UTX", "Merck")
+
+fit2=lm(as.matrix(stocks_diff)~FF_data$Mkt.RF + FF_data$SMB + FF_data$HML)
+options(digits = 4)
+
+coef(fit2)

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_17-18/capm_lab.R

@@ -1,0 +1,160 @@
+##########################################################################################
+###########  R script for Chapter 17   ###################################################
+###########  of Statistics and Data Analysis for Financial Engineering, 2nd Edition ######
+###########  by Ruppert and Matteson  ####################################################
+##########################################################################################
+
+library(data.table)
+library(xts)
+library(ggplot2)
+library(ggthemes)
+
+theme_set(theme_light())
+
+theme_set(theme_light())
+
+# Theme Overrides
+theme_update(axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
+             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
+             legend.position = "right", legend.title.align=0.5,
+             panel.border = element_rect(linetype = "solid", 
+                                         colour = "lightgray"), 
+             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))
+
+data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
+
+setwd("D:/Projects/MSDS-RiskAnalytics/Module_09")
+
+###  Figures 17.1 and 17.2 were produced by Matlab for the book "Statistics and Finance"  ###
+
+##################################################
+############  Code for Figure 17.3   ###############
+##################################################
+
+library(Ecdat)
+library(quadprog)
+data(CRSPday)
+R = 100*CRSPday[,4:6]  #  convert to percentages
+mean_vect = apply(R,2,mean)
+cov_mat = cov(R)
+sd_vect = sqrt(diag(cov_mat))
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_17-18/factor_models.R

@@ -1,0 +1,399 @@
+library(data.table)
+library(xts)
+library(ggplot2)
+library(ggthemes)
+
+theme_set(theme_light())
+
+theme_set(theme_light())
+
+# Theme Overrides
+theme_update(axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
+             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
+             legend.position = "right", legend.title.align=0.5,
+             panel.border = element_rect(linetype = "solid", 
+                                         colour = "lightgray"), 
+             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))
+
+data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
+
+setwd("D:/Projects/MSDS-RiskAnalytics/Module_09")
+
+###########  R script for Chapter 18  ####################################################
+###########  of Statistics and Data Analysis for Financial Engineering, 2nd Edition ######
+###########  by Ruppert and Matteson  ####################################################
+
+####################################################################
+############  Code for Example 18.2 and Figure 18.1  ###############
+####################################################################
+
+datNoOmit = read.table(paste0(data.dir, "treasury_yields.txt"),header=T)
+diffdatNoOmit = diff(as.matrix(datNoOmit[,2:12]))
+dat=na.omit(datNoOmit)
+diffdat = na.omit(diffdatNoOmit)
+n = dim(diffdat)[1]
+options(digits=5)
+pca = prcomp(diffdat)
+summary(pca)
+pdf("yields01.pdf",width=7,height=6)   ###  Figure 18.1
+par(mfrow=c(2,2))
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_18/FF_alt.R

@@ -1,35 +1,0 @@
-# Stock/Bond/FX data.
-stocks <- as.data.table(read.csv(paste0(data.dir, "Stock_FX_Bond_2004_to_2005.csv"), 
-                                 header=T))
-stocks$Date <- as.Date(stocks$Date, format = "%d-%b-%y")
-stocks_subset <- stocks[, .(Date, GM_AC, F_AC, UTX_AC, MRK_AC)]
-stocks_diff <- data.table(Date = stocks_subset[-1]$Date, 
-                          apply(log(stocks_subset[, .(GM_AC, F_AC, UTX_AC, MRK_AC)]), 2, diff))
-
-# Fama-French data.
-FF_data <- as.data.table(read.table(paste0(data.dir, "FamaFrenchDaily.txt"), 
-                                    header=T))
-
-stocks = read.csv(paste0(data.dir, "Stock_FX_Bond_2004_to_2005.csv"), header=T)
-
-FF_data = read.table(paste0(data.dir, "FamaFrenchDaily.txt"), header=T)
-
-start_index = which( FF_data$date == "20040102" )
-end_index = which( FF_data$date == "20051230" )
-FF_data = FF_data[ start_index:end_index, ]
-FF_data = FF_data[-1,]     # delete first row since stocks_diff loses a row due to differencing
-
-start_index = which( stocks$DATE == "1/2/2004" )
-end_index = which( stocks$DATE == "12/30/2005" )
-stocks = stocks[ start_index:end_index, ]
-
-stocks_subset = as.data.frame(cbind(stocks$GM_AC,stocks$F_AC,stocks$UTX_AC,stocks$MRK_AC))
-
-stocks_diff = as.data.frame(100*apply(log(stocks_subset), 2, diff) - FF_data$RF) # Excess returns
-
-names(stocks_diff) = c("GM", "Ford", "UTX", "Merck")
-
-fit2=lm(as.matrix(stocks_diff)~FF_data$Mkt.RF + FF_data$SMB + FF_data$HML)
-options(digits = 4)
-
-coef(fit2)

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_18/capm_lab.R

@@ -1,160 +1,0 @@
-##########################################################################################
-###########  R script for Chapter 17   ###################################################
-###########  of Statistics and Data Analysis for Financial Engineering, 2nd Edition ######
-###########  by Ruppert and Matteson  ####################################################
-##########################################################################################
-
-library(data.table)
-library(xts)
-library(ggplot2)
-library(ggthemes)
-
-theme_set(theme_light())
-
-theme_set(theme_light())
-
-# Theme Overrides
-theme_update(axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
-             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
-             legend.position = "right", legend.title.align=0.5,
-             panel.border = element_rect(linetype = "solid", 
-                                         colour = "lightgray"), 
-             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))
-
-data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
-
-setwd("D:/Projects/MSDS-RiskAnalytics/Module_09")
-
-###  Figures 17.1 and 17.2 were produced by Matlab for the book "Statistics and Finance"  ###
-
-##################################################
-############  Code for Figure 17.3   ###############
-##################################################
-
-library(Ecdat)
-library(quadprog)
-data(CRSPday)
-R = 100*CRSPday[,4:6]  #  convert to percentages
-mean_vect = apply(R,2,mean)
-cov_mat = cov(R)
-sd_vect = sqrt(diag(cov_mat))
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_18/factor_models.R

@@ -1,399 +1,0 @@
-library(data.table)
-library(xts)
-library(ggplot2)
-library(ggthemes)
-
-theme_set(theme_light())
-
-theme_set(theme_light())
-
-# Theme Overrides
-theme_update(axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
-             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
-             legend.position = "right", legend.title.align=0.5,
-             panel.border = element_rect(linetype = "solid", 
-                                         colour = "lightgray"), 
-             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))
-
-data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
-
-setwd("D:/Projects/MSDS-RiskAnalytics/Module_09")
-
-###########  R script for Chapter 18  ####################################################
-###########  of Statistics and Data Analysis for Financial Engineering, 2nd Edition ######
-###########  by Ruppert and Matteson  ####################################################
-
-####################################################################
-############  Code for Example 18.2 and Figure 18.1  ###############
-####################################################################
-
-datNoOmit = read.table(paste0(data.dir, "treasury_yields.txt"),header=T)
-diffdatNoOmit = diff(as.matrix(datNoOmit[,2:12]))
-dat=na.omit(datNoOmit)
-diffdat = na.omit(diffdatNoOmit)
-n = dim(diffdat)[1]
-options(digits=5)
-pca = prcomp(diffdat)
-summary(pca)
-pdf("yields01.pdf",width=7,height=6)   ###  Figure 18.1
-par(mfrow=c(2,2))
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_19-21/VaR_Lab.R

@@ -1,0 +1,337 @@
+library(xts)
+library(ggplot2)
+library(ggthemes)
+
+theme_set(theme_light())
+
+# Theme Overrides
+theme_update(plot.title = element_text(hjust = 0.5),
+             axis.text.x = element_text(size = 10),
+             axis.text.y = element_text(size = 10),
+             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
+             legend.position = "top", legend.title = element_blank())
+
+data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
+
+setwd("D:/Projects/MSDS-RiskAnalytics/Module_07")
+
+library(Ecdat)
+
+data(SP500, package = "Ecdat")
+
+n <- 2783
+
+# non-parametric VaR and ES
+
+SPreturn <- SP500$r500[(n - 999):n]
+year <- 1981 + (1:n) * (1991.25 - 1981) / n
+year <- year[(n - 999):n]
+
+alpha <- 0.05
+
+q <- as.numeric(quantile(SPreturn, alpha))
+
+VaR_nonp <- -20000 * q
+
+IEVaR <- (SPreturn < q)
+sum(IEVaR)
+
+ES_nonp <- -20000 * sum(SPreturn * IEVaR) / sum(IEVaR)
+
+options(digits = 5)
+VaR_nonp
+ES_nonp
+
+# parametric VaR and ES
+
+library(MASS)
+
+fitt <- fitdistr(SPreturn, "t")
+param <- as.numeric(fitt$estimate)
+
+mean <- param[1]
+df <- param[3]
+sd <- param[2] * sqrt((df) / (df - 2))
+
+lambda <- param[2]
+qalpha <- qt(alpha, df = df)
+VaR_par <- -20000 * (mean + lambda * qalpha)
+es1 <- dt(qalpha, df = df) / (alpha)
+es2 <- (df + qalpha^2) / (df - 1)
+es3 <- -mean + lambda * es1 * es2
+ES_par <- 20000 * es3
+
+VaR_par
+ES_par
+
+##################################################
+############  Code for Figure 19.2  ##############
+##################################################
+data(SP500,package="Ecdat")
+library("fGarch")
+n = 2783
+SPreturn = SP500$r500[(n-999):n]
[…]

### 2020-04-06 Statistics and Data Analysis for Financial Engineering/ch_19/VaR_Lab.R

@@ -1,337 +1,0 @@
-library(xts)
-library(ggplot2)
-library(ggthemes)
-
-theme_set(theme_light())
-
-# Theme Overrides
-theme_update(plot.title = element_text(hjust = 0.5),
-             axis.text.x = element_text(size = 10),
-             axis.text.y = element_text(size = 10),
-             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
-             legend.position = "top", legend.title = element_blank())
-
-data.dir <- "D:/Projects/MSDS-RiskAnalytics/datasets/"
-
-setwd("D:/Projects/MSDS-RiskAnalytics/Module_07")
-
-library(Ecdat)
-
-data(SP500, package = "Ecdat")
-
-n <- 2783
-
-# non-parametric VaR and ES
-
-SPreturn <- SP500$r500[(n - 999):n]
-year <- 1981 + (1:n) * (1991.25 - 1981) / n
-year <- year[(n - 999):n]
-
-alpha <- 0.05
-
-q <- as.numeric(quantile(SPreturn, alpha))
-
-VaR_nonp <- -20000 * q
-
-IEVaR <- (SPreturn < q)
-sum(IEVaR)
-
-ES_nonp <- -20000 * sum(SPreturn * IEVaR) / sum(IEVaR)
-
-options(digits = 5)
-VaR_nonp
-ES_nonp
-
-# parametric VaR and ES
-
-library(MASS)
-
-fitt <- fitdistr(SPreturn, "t")
-param <- as.numeric(fitt$estimate)
-
-mean <- param[1]
-df <- param[3]
-sd <- param[2] * sqrt((df) / (df - 2))
-
-lambda <- param[2]
-qalpha <- qt(alpha, df = df)
-VaR_par <- -20000 * (mean + lambda * qalpha)
-es1 <- dt(qalpha, df = df) / (alpha)
-es2 <- (df + qalpha^2) / (df - 1)
-es3 <- -mean + lambda * es1 * es2
-ES_par <- 20000 * es3
-
-VaR_par
-ES_par
-
-##################################################
-############  Code for Figure 19.2  ##############
-##################################################
-data(SP500,package="Ecdat")
-library("fGarch")
-n = 2783
-SPreturn = SP500$r500[(n-999):n]
[…]

### 2020-04-14 Case Studies/01_geolocation.rmd
geolocation case study
geolocation case study

@@
 
-reshapeSS <- function(data, varSignal = "signal",
-                      keepVars = c("posXY", "posX", "posY")) {
+reshapeSS <- function(data, 
+                      varSignal = "signal",
+                      keepVars = c("posXY", "posX", "posY"),
+                      sampleAngle = F) {
+
+   if(sampleAngle)
+      data <- data[angle == sample(data$angle, size = 1), ]
 
@@
                        avgSS <- tapply(x$signal, x$mac, mean)
-                       y = matrix(avgSS, nrow = 1, ncol = 6,
-                                  dimnames = list(ans$posXY, names(avgSS)))
+                       y = matrix(avgSS, nrow = 1, dimnames = list(ans$posXY, names(avgSS)))
                        cbind(ans, y)
@@
    col_names <- colnames(newDataSS)
+   to_change <- !(col_names %in% keepVars)
+
    n_cols <- length(col_names)
+   start <- length(keepVars)
 
-   colnames(newDataSS)[4:n_cols] <- sapply(col_names[4:n_cols], function(col) {
+   colnames(newDataSS)[to_change] <- sapply(col_names[to_change], function(col) {
       n <- nchar(col)
@@
 
-   newDataSS[, 4:ncol(newDataSS)] <- round(newDataSS[, 4:ncol(newDataSS)])
+   newDataSS[, start:ncol(newDataSS)] <- round(newDataSS[, start:ncol(newDataSS)])
 
@@
 ```
+
+```{r}
+v <- 11;
+
+permuteLocs <- sample(unique(offlineSummary$posXY))
+permuteLocs <- matrix(permuteLocs, ncol = v,
+                      nrow = floor(length(permuteLocs)/v))
+
+```
+
+```{r}
+onlineFold <- subset(offlineSummary, posXY %in% permuteLocs[, 1])
+
+onlineCVSummary <- reshapeSS(offline, 
+                              keepVars = c("posXY", "posX", "posY", "angle"),
+                              sampleAngle = T)
+
+onlineFold <- subset(onlineCVSummary,
+                     posXY %in% permuteLocs[ , 1])
+
+estFold <- predXY(newSignals = onlineFold[, 6:11],
+                  newAngles = onlineFold[, 4],
+                  offlineFold, numAngles = 3, k = 3)
+
+actualFold <- onlineFold[, c("posX", "posY")]
+calcError(estFold, actualFold)
+
+```
+
+```{r}
+K <- 20
+err <- rep(0, K)
+
+for(j in 1:v) {
+   onlineFold <- subset(onlineCVSummary,
+                        posXY %in% permuteLocs[, j])
+   offlineFold <- subset(offlineSummary,
+                         posXY %in% permuteLocs[ , -j])
+
+   actualFold <- onlineFold[, c("posX", "posY")]
+
+   for(k in 1:K) {
+      estFold <- predXY(newSignals = onlineFold[, 6:11],
+                        newAngles = onlineFold[, 4],
+                        offlineFold, numAngles = 3, k = k)
+
+      err[k] <- err[k] + calcError(estFold, actualFold) 
+   }
+}
+
+```
+
+```{r}
+k_values <- data.table(K = 1:K, Error = err, Min = which.min(err))
+
+ggplot(k_values, aes(K, Error)) +
+   geom_line(aes(col = Error)) +
+   geom_hline(data = k_values[K == Min], aes(yintercept = Error), col = "darkblue") +
+   labs("K-NN Training Error")
+```
+
+```{r}
+estXYk5 <- predXY(newSignals = onlineSummary[, 6:11],
+                 newAngles = onlineSummary[, 4],
+                 offlineSummary, numAngles = 3, k = 5)
+```
+
+```{r}
+calcError(estXYk5, actualXY)
+```
+
+```{r}
+error_table <- data.table(K = c(1, 3, 5), 
+                          Error = c(calcError(estXYk1, actualXY), 
+                                    calcError(estXYk3, actualXY), 
+                                    calcError(estXYk5, actualXY)))
+
+ggplot(error_table, aes(K, Error, fill = -Error)) +
+   geom_bar(stat = "identity")
+```
+
+```{r}
+predXY <- function(newSignals, newAngles, trainData, numAngles = 1, k = 3) {
+   chooseXY <- list(length = nrow(newSignals))
+
+   for(i in 1:nrow(newSignals)) {
+      trainSS <- selectTrain(newangles[i], trainData, m = numAngles)
+      closeXY[[i]] <- findNN(newSignal = as.numeric(newSignals[i, ]),
+                             trainSS)
+   }
+
+   estXY <- lapply(closeXY, function(x)
+                              sapply(x[, 2:3],
+                                     function(x) mean(x[1:k])))
+   estXY <- do.call("rbind", estXY)
+
+   return(XY)
+}
+
+```
+
+### Follow-up
+
+Write the code to read the raw training data into the data structure in the first approach described in section 1.2.
+That is, the data structure is a data frame with a column for each MAC address that detected a signal.
+For the column name, use the last two characters of the MAC address, or some other unique identifier.
+
+```{r}
+
+```

### 2020-04-16 Case Studies/01_geolocation.rmd
case studies - geolocation
case studies - geolocation

@@
 
+library(rbenchmark, quietly = TRUE, warn.conflicts = FALSE)
+
 options(tinytex.verbose = TRUE)
@@
 
-train130 <- selectTrain(130, offlineSummary, m = 3)
-
-dim(train130)
-
 findNN <- function(newSignal, trainSubset) {
@@
       trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
-      closeXY[[i]] = 
-         findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
-   }
-
-   estXY = lapply(closeXY,
-                  function(x) sapply(x[, 2:3],
+      closeXY[[i]] <- findNN(newSignal = as.numeric(newSignals[i, ]),
+                             trainSS)
+   }
+
+   estXY <- lapply(closeXY, function(x)
+                              sapply(x[, 2:3],
                                      function(x) mean(x[1:k])))
-
    estXY <- do.call("rbind", estXY)
@@
 
-onlineFold <- subset(onlineCVSummary,
+offlineFold <- subset(offlineSummary,
                      posXY %in% permuteLocs[ , 1])
@@
 
-```{r}
-predXY <- function(newSignals, newAngles, trainData, numAngles = 1, k = 3) {
-   chooseXY <- list(length = nrow(newSignals))
-
-   for(i in 1:nrow(newSignals)) {
-      trainSS <- selectTrain(newangles[i], trainData, m = numAngles)
-      closeXY[[i]] <- findNN(newSignal = as.numeric(newSignals[i, ]),
-                             trainSS)
-   }
-
-   estXY <- lapply(closeXY, function(x)
-                              sapply(x[, 2:3],
-                                     function(x) mean(x[1:k])))
-   estXY <- do.call("rbind", estXY)
-
-   return(XY)
-}
-
-```
-
-### Follow-up
+### Further Analysis
+
+### 1.)
 
@@
 
-```
+file <- file_offline
+
+lines <- read_lines(file)
+
+valid_lines <- lines[ validLines(lines) ]
+
+parseLine <- function(line) {
+
+   pairs <- strsplit(line, ";")
+   tokens <- lapply(pairs, function(x)strsplit(x, "="))
+
+   mat <- matrix(unlist(tokens), ncol = 2, byrow = T)
+
+   colnames <- mat[, 1]
+   values <- mat[, 2]
+
+   dt <- data.table(t(values))
+   colnames(dt) <- colnames
+
+   dt
+}
+
+rows <- lapply(valid_lines, parseLine)
+
+offline_alt <- rbindlist(rows, fill = T)
+
+head(offline_alt)
+```
+
+### 2.)
+
+Compare the size of two data structures: the data frame created in Section 1.2 and the data frame created in the previous problem.
+
++ Which uses less memory?
+
+```{r}
+object.size(offline)
+
+object.size(offline_alt)
+```
+
++ What are the dimensions of each?
+
+```{r}
+dim(offline)
+
+dim(offline_alt)
+```
+
++ How might this change with different numbers of devices in the building?
+
+_Columns will expand out drastically with more devices._
+
++ Different number of signals from the less commonly detected devices?
+
+_The more devices we have the more the columns in the alternative approach will be filled with NAs, thus bloating the object size significantly._
+
+### 3.)
+
+Compare the total time it takes to read the raw data and create the data frame, for the two approaches described in Section 1.2.
+
+```{r}
+file <- file_offline
+
+fn_read_data <- function(file, processFun, sample = F, samp.size = 100 ) {
+   start_time <- Sys.time()
+
+   lines <- read_lines(file)
+   valid_lines <- lines[ validLines(lines) ]
+
+   n <- length(valid_lines)
+
+   if(sample) {
+      samp <- sample(n, samp.size, replace = F)
+
+      valid_lines <- valid_lines[samp]
+   }
+
+   res <- processFun(valid_lines) # throw away
+
+   return(Sys.time() - start_time)
+}
+
+fn_read_data(file_offline, processLine)
+
+fn_read_data(file_offline, parseLine)
+```
+
+Do this for different size subsets of the data (chosen at random) and draw a curve of the time against input size for each of the approaches.
+
+```{r}
+sample_sizes <- seq( from = 100, to = 15000, by = 2000)
+
+times_classic <- sapply(sample_sizes, function(samp) {
+   fn_read_data(file_offline, processLine, sample = T, samp.size = samp)
+})
+
+times_alt <- sapply(sample_sizes, function(samp) {
+   fn_read_data(file_offline, parseLine, sample = T, samp.size = samp)
+})
+
+results <- data.table(N = sample_sizes, Classic = times_classic, alt = times_alt)
+
+ggplot(results) +
+   geom_line(aes(N, Classic), col = "darkblue") +
+   geom_line(aes(N, alt), col = "darkred")
+```
+
+Also, comment on the memory and speed for the two approaches.
+
+_The alt approach is substantially worse in all aspects._
+
+### 4.)
+
+Examine the time variable in the offline data.
+Any chance over time in the characteristics of the signal caused by, e.g., reduced battery power in the measuring devices as time goes by, or measurments taken on different days may be made by different people with different levels of accuracy.
+Also, examination of time can give insights into how the experiment was carried out.
+
+Were the positions close to each other measured at similar times?
+
+Do you see any change in the signal strength variable or mean over time?
+
+Try controlling for other variables that might affect this relationship.
+
+```{r}
+
+offline_copy <- offline
+offline_copy$ts <- strftime(offline_copy$time, format="%H:%M:%S")
+offline_copy$ts_m <- strftime(offline_copy$time, format="%H:%M")
+offline_copy$ts_h <- strftime(offline_copy$time, format="%H")
+
+```
+
+```{r}
+ggplot(offline_copy, aes(ts, signal, fill = signal)) +
+   geom_bar(stat = "identity") +
+   scale_y_reverse()
+
+```
+
+```{r}
+ggplot(offline_copy, aes(ts_m, signal, fill = signal)) +
+   geom_bar(stat = "identity") +
+   scale_y_reverse()
+```
+
+```{r}
+ggplot(offline_copy, aes(ts_h, signal, fill = signal)) +
+   geom_bar(stat = "identity") +
+   scale_y_reverse()
+```
+
+### 5.)
+
+Write the readData function described in Section 1.3.4.
+The arguments to this function are the file name and the MAC address to retain, subKMacs.
+Determine whether these parameters should have default values or not.
+The return value is data frame described in section 1.3.
+Use the findGlobals() function available in _codetools_ to check that the function is not relying on any global variables.
+
+```{r}
+readData <- function(file, submacs = macs) {
+
+   lines <- read_lines(file)
+
+   valid_lines <- lines[ validLines(lines) ]
+
+   processed_lines <- lapply(valid_lines, processLine)
+
+   data <- as.data.table(do.call("rbind", processed_lines),
+                         stringsAsFactors = F)
+
+   names(data) <- c("time", "scanMac", "posX", "posY", "posZ",
+                       "orientation", "mac", "signal", 
+                       "channel", "type")
+
+   numVars <- c("time", "posX", "posY", "posZ",
+                "orientation", "signal")
+
+   data[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]
+
+   data <- data[ data$type == 3, ]
+   data[, type := NULL]
+
+   data[, rawTime := time]
+   data[, time := time/1000]
+   class(data$time) = c("POSIXt", "POSIXct")
+
+   # drop scanMac & posZ
+   data[, `:=`(scanMac = NULL, posZ = NULL)]
+
+   data$angle = roundOrientation(data$orientation)
+
+   data$channel = NULL
+
+   data$posXY <- paste(data$posX, data$posY, sep = "-")
+
+   return(data[mac %in% submacs])
+}
+```
+
+### 6.)
+
+In section 1.4.1 we calculated measures of center and location for the signal strenghts at each location x angle x access point combination (see  fig 1.9 for example).
+
+Another possible summary statistic we can calculate is the Kolmogorov-Smirnov test-statistic for normality.
+If the signal strenghts are roughtly normal, then we expect the p-values to have a uniform distribution.
+This leads to about 5% of the p-values for the 8000 tests to fall below 0.05.
+
+```{r}
+signalSummary <-
+   lapply(byLocAngleAP,
+          function(oneLoc) {
+             ans = oneLoc[1, ]
+             ans$medSignal = median(oneLoc$signal)
+             ans$avgSignal = mean(oneLoc$signal)
+             ans$num = length(oneLoc$signal)
+             ans$sdSignal = sd(oneLoc$signal)
+             ans$iqrSignal = IQR(oneLoc$signal)
+             ans
+          })
+
+offlineSigSummary <- do.call("rbind", signalSummary)
+
+breaks <- seq(-90, -30, by = 5)
+
+bwplot(sdSignal ~ cut(avgSignal, breaks = breaks),
+       data = offlineSummary,
+       subset = mac != "00:0f:a3:39:dd:cd",
+       xlab = "Mean Signal", ylab = "SD Signal")
+
+by_pos_angle_mac <- offlineSigSummary[, .(AvgSignal = mean(signal)), by = list(mac, angle, posXY)]
+
+with(by_pos_angle_mac, ks.test(AvgSignal, "pnorm")) # Kolmogorov-Smirnov test-statistic
+
+by_pos_angle <- offlineSigSummary[, .(AvgSignal = mean(signal)), by = list(angle, posXY)]
+
+with(by_pos_angle, ks.test(AvgSignal, "pnorm")) # Kolmogorov-Smirnov test-statistic
+
+by_mac <- offline[, .(AvgSignal = mean(signal)), by = list(mac)]
+
+with(by_mac, ks.test(AvgSignal, "pnorm")) # Kolmogorov-Smirnov test-statistic
+
+```
+
+### 7.)
+
+Write the surfaceSS() function that creates plots such as those in Figure 1.10.
+This function takes 3 arguments: data for the offline summary data frame, mafc and angle.
+The parameters mac and angle are used to specify which MAC address and angle are to be selected from the data for smoothing and plotting.
+
+```{r}
+
+reshapeSS <- function(data, 
+                      varSignal = "signal",
+                      keepVars = c("posXY", "posX", "posY"),
+                      sampleAngle = F) {
+
+   if(sampleAngle)
+      data <- data[angle == sample(data$angle, size = 1), ]
+
+   byLocation <- with(data,
+                 by(data, list(posXY),
+                    function(x) {
+                       ans <- x[1, ..keepVars]
+                       avgSS <- tapply(x$signal, x$mac, mean)
+                       y = matrix(avgSS, nrow = 1, dimnames = list(ans$posXY, names(avgSS)))
+                       cbind(ans, y)
+                    }))
+
+   newDataSS <- do.call("rbind", byLocation)
+
+   col_names <- colnames(newDataSS)
+   to_change <- !(col_names %in% keepVars)
+
+   n_cols <- length(col_names)
+   start <- length(keepVars)
+
+   colnames(newDataSS)[to_change] <- sapply(col_names[to_change], function(col) {
+      n <- nchar(col)
+      substr(col, n - 2, n)
+   })
+
+   newDataSS[, start:ncol(newDataSS)] <- round(newDataSS[, start:ncol(newDataSS)])
+
+   return(newDataSS)
+}
+
+```
+
+### 8.)
+
+Consider the scatter plots in Figure 1.11. There appear to be curvature in the signal strength-distance relationship. Does a log transformation improve this relationship, i.e., make it linear?
+
+```{r}
+
+offlineSummary$dist <- sqrt(diffs[, 1]^2 + diffs[, 2]^2)
+offlineSummary$log_dist <- log(offlineSummary$dist)
+
+xyplot(signal ~ dist | factor(mac) + factor(angle),
+       data = offlineSummary, pch = 19, cex = 0.3,
+       xlab = "distance")
+
+xyplot(signal ~ log_dist | factor(mac) + factor(angle),
+       data = offlineSummary, pch = 19, cex = 0.3,
+       xlab = "log distance")
+
+xyplot(signal ~ log_dist | factor(mac) + factor(angle),
+       data = offlineSummary[angle == 315], pch = 19, cex = 0.3,
+       xlab = "log distance")
+
+xyplot(signal ~ dist | factor(mac) + factor(angle),
+       data = offlineSummary[angle == 315], pch = 19, cex = 0.3,
+       xlab = "distance")
+```
+
+_The log transform removes a bit of the curvature, but not a great deal._
+
+### 9.)
+
+The floor plan for the building (see Figure 1.1) shows 65 access points.
+However, the data contain 7 access points with roughly the expected number of signals (166 locations x 8 orientations x 100 replications = 146,080 measurments).
+
+With the signal strength seen in the heat maps of Figure 1.10), we matched the access points to the corresponding MAC addresses.
+However, two of the MAC addresses seem to be for the same access point 00:0f:a3:39:e1:c0 and to eliminate the 00:0f:a3:39:dd:cd address.
+
+Conduct a more thorough data analysis into these two MAC addresses. Did we make the correct decision? Does swapping out the one we kept for the one we discarded improve the prediction?
+
+```{r}
+
+subMacs <- names(sort(table(offline$mac), decreasing = T)) # all sub macs
+
+angles <- seq(0, by = 45, length = 8) # all angles
+
+surfaceSS <- function(data, filter_mac, filter_angle) {
+   oneAPAngle <- data[mac == filter_mac & angle == filter_angle]
+
+   smothSS <- Tps(oneAPAngle[, c("posX", "posY")],
+                  oneAPAngle$avgSignal)
+
+   vizSmooth <- predictSurface(smothSS)
+
+   plot.surface(vizSmooth, type = "C")
+   points(oneAPAngle$posX, oneAPAngle$posY, pch=19, cex = 0.5)
+}
+
+```
+
+```{r}
+surfaceSS(offlineSummary, subMacs[1], angles[1])
+surfaceSS(offlineSummary, subMacs[1], angles[5])
+surfaceSS(offlineSummary, subMacs[1], angles[7])
+
+```
+
+```{r}
+surfaceSS(offlineSummary, subMacs[3], angles[2])
+surfaceSS(offlineSummary, subMacs[3], angles[5])
+surfaceSS(offlineSummary, subMacs[3], angles[7])
+
+```
+
+```{r}
+surfaceSS(offlineSummary, subMacs[4], angles[2])
+surfaceSS(offlineSummary, subMacs[4], angles[5])
+surfaceSS(offlineSummary, subMacs[4], angles[7])
+```
+
+```{r}
+surfaceSS(offlineSummary, subMacs[5], angles[2])
+surfaceSS(offlineSummary, subMacs[5], angles[5])
+surfaceSS(offlineSummary, subMacs[5], angles[7])
+
+```
+
+```{r}
+surfaceSS(offlineSummary, subMacs[6], angles[2])
+surfaceSS(offlineSummary, subMacs[6], angles[5])
+surfaceSS(offlineSummary, subMacs[6], angles[7])
+```
+
+### 10.)
+
+Write the selectTrain() function described in Section 1.5.2.
+This function has 3 parameters: angleNewObs, the angle of the new observation, signals, the training data, i.e., data in the format of offlineSummary; and m, the number of angles to include from signals.
+
+The function returns a data frame that matches trainSS, i.e., selectTrain() calls reshapeSS().
+
+```{r}
+reshapeSS <- function(data, 
+                      varSignal = "signal",
+                      keepVars = c("posXY", "posX", "posY"),
+                      sampleAngle = F) {
+
+   if(sampleAngle)
+      data <- data[angle == sample(data$angle, size = 1), ]
+
+   byLocation <- with(data,
+                 by(data, list(posXY),
+                    function(x) {
+                       ans <- x[1, ..keepVars]
+                       avgSS <- tapply(x$signal, x$mac, mean)
+                       y = matrix(avgSS, nrow = 1, dimnames = list(ans$posXY, names(avgSS)))
+                       cbind(ans, y)
+                    }))
+
+   newDataSS <- do.call("rbind", byLocation)
+
+   col_names <- colnames(newDataSS)
+   to_change <- !(col_names %in% keepVars)
+
+   n_cols <- length(col_names)
+   start <- length(keepVars)
+
+   colnames(newDataSS)[to_change] <- sapply(col_names[to_change], function(col) {
+      n <- nchar(col)
+      substr(col, n - 2, n)
+   })
+
+   newDataSS[, start:ncol(newDataSS)] <- round(newDataSS[, start:ncol(newDataSS)])
+
+   return(newDataSS)
+}
+```
+
+### 11.)
+
+We use Euclidean distance to find the distance between the signal strength vectors.
+However, Euclidean distance is not robust in taht it is sensitive to outliers.
+Consider other metrics such as the $L_1$ distance, i.e., the absolute value of the difference.
+Modify the findNN() functions in sections 1.5.3 to use this alternative distance.
+
+```{r}
+
+findNN_l1 <- function(newSignal, trainSubset) {
+   diffs <- apply(trainSubset[ , 4:9], 1,
+                  function(x) abs(x - newSignal))
+   dists <- apply(diffs, 2, function(x) sum(x))
+   closest <- order(dists)
+   return(trainSubset[closest, 1:3 ])
+}
+
+predXY_l1 <- function(newSignals, newAngles, trainData,
+                   numAngles = 1, k = 3) {
+
+   closeXY <- list(length = nrow(newSignals))
+
+   for(i in 1:nrow(newSignals)) {
+      trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
+
+      closeXY[[i]] = 
+         findNN_l1(newSignal = as.numeric(newSignals[i, ]), trainSS)
+   }
+
+   estXY = lapply(closeXY,
+                  function(x) sapply(x[, 2:3],
+                                     function(x) mean(x[1:k])))
+
+   estXY <- do.call("rbind", estXY)
+
+   return(estXY)
+}
+
+estXYk1_l1 <- predXY_l1(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 1)
+
+estXYk3_l1 <- predXY_l1(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 3)
+
+estXYk5_l1 <- predXY_l1(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 5)
+
+error_table <- data.table(K = c(1, 3, 5), 
+                          Error = c(calcError(estXYk1, actualXY), 
+                                    calcError(estXYk3, actualXY), 
+                                    calcError(estXYk5, actualXY)),
+                          Method = "Euclidean")
+
+error_table_l1 <- data.table(K = c(1, 3, 5),
+                          Error = c(calcError(estXYk1_l1, actualXY), 
+                                    calcError(estXYk3_l1, actualXY), 
+                                    calcError(estXYk5_l1, actualXY)),
+                          Method = "Manhattan")
+
+errors_combined <- rbind(error_table, error_table_l1)
+
+ggplot(errors_combined) +
+   geom_bar(aes(K, Error, fill = -Error), stat = "identity") +
+   facet_wrap(~Method) +
+   coord_flip()
+
+```
+
+Does it improve predictions?
+
+_Euclidean seems to perform better._
+
+### 12.)
+
+To predict locaiton, we use the k nearest neighbors to a set of signal strenghts.
+We average the known (x, y) values for these neighbors.
+However, a better predictor might be a weighted average, wehre the weights are inversely proportional to the "distance" (in signal strength) from the test observation.
+This allows us to include the k points that are close, but to differentiate between them by how close they actually are.
+
+The weights might be:
+
+$\frac{1/d_i}{\sum^{k}_{i=1}1/d_i}$
+
+for the i-th closest neighboring observation where $d_i$ is the distance from our new test point to this reference point (in signal strength space).
+
+Implement this alternative prediciton method.
+
+```{r}
+
+predXY_weighted <- function(newSignals, newAngles, trainData,
+                   numAngles = 1, k = 3) {
+
+   closeXY <- list(length = nrow(newSignals))
+
+   for(i in 1:nrow(newSignals)) {
+      trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
+
+      closeXY[[i]] = 
+         findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
+   }
+
+   estXY = lapply(closeXY,
+               function(x) sapply(x[, 2:3],
+                                  function(x) {
+                                     points <- x[1:k]
+
+                                     if(k > 1) {
+                                        weights <- points/sum(points)
+
+                                        return(sum(points * weights))
+                                     } else {
+                                        return(points)
+                                     }
+                                  }))
+
+   estXY <- do.call("rbind", estXY)
+
+   return(estXY)
+}
+
+estXYk10 <- predXY(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 10)
+
+estXYk1_weighted <- predXY_weighted(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 1)
+
+estXYk3_weighted <- predXY_weighted(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 3)
+
+estXYk5_weighted <- predXY_weighted(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 5)
+
+estXYk10_weighted <- predXY_weighted(newSignals = onlineSummary[, 6:11],
+                  newAngles = onlineSummary[, 4],
+                  offlineSummary, numAngles = 3, k = 10)
+
+error_table <- data.table(K = c(1, 3, 5, 10),
+                          Error = c(calcError(estXYk1, actualXY), 
+                                    calcError(estXYk3, actualXY), 
+                                    calcError(estXYk5, actualXY),
+                                    calcError(estXYk10, actualXY)),
+                          Method = "Mean")
+
+error_table_weighted <- data.table(K = c(1, 3, 5, 10),
+                          Error = c(calcError(estXYk1_weighted, actualXY), 
+                                    calcError(estXYk3_weighted, actualXY), 
+                                    calcError(estXYk5_weighted, actualXY),
+                                    calcError(estXYk10_weighted, actualXY)),
+                          Method = "Weighted Avg.")
+
+errors_combined <- rbind(error_table, error_table_weighted)
+
+ggplot(errors_combined) +
+   geom_bar(aes(K, Error, fill = -Error), stat = "identity") +
+   facet_wrap(~Method) +
+   coord_flip()
+
+errors_combined
+```
+
+Does this improve the predictions?
+
+_The weighted avg approach is of course identical in the base case (k=1), however, it performs slightly worse than the simple average in the k=3 and k=5 cases.
+I even upped k to 10 (the best CV k), and we still see simple average beats the weighted approach._
+
+### 13.)
+
+In Section 1.5.4 we used cross-validation to choose k, the number of neighbors. Another parameter to choose is the number of angles at which the signal strength was measured.
+
+Use cross-validation to select this value.
+
+Also, consider N + K cross-validation pairs.
+
+```{r}
+
+v <- 11 # n-folds
+N <- 8 # num angles
+K <- 20 # also choose k
+
+nk_err <- matrix(rep(0, N * K), nrow = N, ncol = K)
+
+for(j in 1:v) {
+   onlineFold <- subset(onlineCVSummary,
+                        posXY %in% permuteLocs[, j])
+   offlineFold <- subset(offlineSummary,
+                         posXY %in% permuteLocs[ , -j])
+
+   actualFold <- onlineFold[, c("posX", "posY")]
+
+   for(k in 1:K) {
+      for(n in 1:N) {
+         estFold <- predXY(newSignals = onlineFold[, 6:11],
+                           newAngles = onlineFold[, 4],
+                           offlineFold, numAngles = n, k = k)
+
+         nk_err[n, k] <- calcError(estFold, actualFold)
+      }
+   }
+}
+
+nk_err
+
+which(nk_err == min(nk_err), arr.ind = TRUE)
+
+head(order(nk_err))
+
+# 1, 6
+# 1, 5
+# 1, 7
+# 4, 8
+
+n_take <- 20
+cv_results <- data.table(nk_err)[, 1:n_take]
+colnames(cv_results) <- sapply(1:n_take, function(x) {paste0("k=", x)})
+cv_results[, Angles := .I]
+
+min_cv <- which(nk_err == min(nk_err), arr.ind = TRUE)
+
+cv_results_flat <- melt(cv_results, id.vars = "Angles")
+
+ggplot(cv_results_flat, aes(variable, value, group = Angles)) +
+   geom_line(aes(col = Angles)) +
+   labs("K, N Cross-Validation Results")
+```
+
+The cross-validation of both N and K yield the optimal N at _1_ and K at _6_. 4 of the top 5 CV results have _N=1_.

### 2020-04-16 Case Studies/01_geolocation.rmd
updates to geolocation case study

@@
 
-```{r}
+```{r, fig.height=8}
 with(offlineSummary,
@@
 findNN <- function(newSignal, trainSubset) {
+
    diffs <- apply(trainSubset[ , 4:9], 1,
                   function(x) x - newSignal)
+
    dists <- apply(diffs, 2, function(x) sqrt(sum(x^2)) )
+
    closest <- order(dists)
+
    return(trainSubset[closest, 1:3 ])
@@
       trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
-      closeXY[[i]] <- findNN(newSignal = as.numeric(newSignals[i, ]),
+
+      sigValues <- suppressWarnings({ as.numeric(newSignals[i, ]) })
+
+      closeXY[[i]] <- findNN(newSignal = sigValues,
                              trainSS)
@@
                               sapply(x[, 2:3],
-                                     function(x) mean(x[1:k])))
+                                     function(x) mean(x[1:k], na.rm = T)))
    estXY <- do.call("rbind", estXY)
@@
 
-```{r}
+```{r, fig.height=4}
 surfaceSS(offlineSummary, subMacs[1], angles[1])
-surfaceSS(offlineSummary, subMacs[1], angles[5])
 surfaceSS(offlineSummary, subMacs[1], angles[7])
@@
 
-```{r}
+```{r, fig.height=4}
 surfaceSS(offlineSummary, subMacs[3], angles[2])
-surfaceSS(offlineSummary, subMacs[3], angles[5])
 surfaceSS(offlineSummary, subMacs[3], angles[7])
@@
 
-```{r}
+```{r, fig.height=4}
 surfaceSS(offlineSummary, subMacs[4], angles[2])
-surfaceSS(offlineSummary, subMacs[4], angles[5])
 surfaceSS(offlineSummary, subMacs[4], angles[7])
@@
 
-```{r}
+```{r, fig.height=4}
 surfaceSS(offlineSummary, subMacs[5], angles[2])
-surfaceSS(offlineSummary, subMacs[5], angles[5])
 surfaceSS(offlineSummary, subMacs[5], angles[7])
@@
 
-```{r}
+```{r, fig.height=6}
 surfaceSS(offlineSummary, subMacs[6], angles[2])
@@
 v <- 11 # n-folds
-N <- 8 # num angles
-K <- 20 # also choose k
+N <- 6 # num angles
+K <- 10 # also choose k
 
@@
 
-n_take <- 20
+n_take <- 10
 cv_results <- data.table(nk_err)[, 1:n_take]
@@
    labs("K, N Cross-Validation Results")
+
 ```

### 2020-04-18 Case Studies/02_race-times.rmd
cherry blossom race data scraping
cherry blossom race data scraping

--- 2020-04-14 Case Studies/01_geolocation.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Geolocation}
-   \rfoot{\color{headergrey}Chapter 1}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Cherry Blossom Race Times}
+   \rfoot{\color{headergrey}Chapter 2}
    \lfoot{\color{headergrey}}
@@
 
-data.dir <- file.path(here::here(), "Case Studies", "datasets")
+data.dir <- file.path(here::here(), "Case Studies", "01_data")
 
@@
 
-file_offline <- file.path(data.dir, "offline.final.trace.txt")
-file_online <- file.path(data.dir, "online.final.trace.txt")
-
-raw_offline <- read_lines(file_offline)
-raw_online <- read_lines(file_online)
-
 ```
-
-#### Sanity Check
-
-```{r}
-# number of comment lines in the data
-sum(substr(raw_offline, 1, 1) == "#")
-
-# total number of lines in the data
-length(raw_offline)
-```
-
-#### Data Pre-processing
-
-Generate read data function for preprocessing training and test data.
-
-```{r}
-
-processLine <- function(x) {
-   tokens <- strsplit(x, "[;=,]")[[1]]
-
-   if(length(tokens) == 10)
-      return(NULL)
-
-   tmp <- matrix(tokens[ - (1:10)], ncol = 4, byrow = T)
-   cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow = nrow(tmp),
-                ncol = 6, byrow = T), tmp)
-}
-
-validLines <- function(data) {
-   substr(data, 1, 1) != "#"
-}
-
-roundOrientation <- function(angles) {
-   refs = seq(0, by = 45, length = 9)
-   q <- sapply(angles, function(o) which.min(abs(o - refs)))
-   c(refs[1:8], 0)[q]
-}
-
-readData <- function(file, submacs = macs) {
-
-   lines <- read_lines(file)
-
-   valid_lines <- lines[ validLines(lines) ]
-
-   processed_lines <- lapply(valid_lines, processLine)
-
-   data <- as.data.table(do.call("rbind", processed_lines),
-                         stringsAsFactors = F)
-
-   names(data) <- c("time", "scanMac", "posX", "posY", "posZ",
-                       "orientation", "mac", "signal", 
-                       "channel", "type")
-
-   numVars <- c("time", "posX", "posY", "posZ",
-                "orientation", "signal")
-
-   data[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]
-
-   data <- data[ data$type == 3, ]
-   data[, type := NULL]
-
-   data[, rawTime := time]
-   data[, time := time/1000]
-   class(data$time) = c("POSIXt", "POSIXct")
-
-   # drop scanMac & posZ
-   data[, `:=`(scanMac = NULL, posZ = NULL)]
-
-   data$angle = roundOrientation(data$orientation)
-
-   data$channel = NULL
-
-   data$posXY <- paste(data$posX, data$posY, sep = "-")
-
-   return(data[mac %in% submacs])
-}
-
-```
-
-#### Test Pre-processor
-
-```{r}
-lines <- processLine(raw_offline[4:20])
-lines
-
-offline_valid <- raw_offline[validLines(raw_offline)]
-
-head(offline_valid)
-
-length(offline_valid)
-```
-
-```{r}
-tmp <- lapply(offline_valid[1:17], processLine)
-
-sapply(tmp, nrow)
-```
-
-#### Dry Run
-
-```{r}
-offline_test <- as.data.table(do.call("rbind", tmp))
-
-offline_test
-```
-
-#### Process
-
-```{r}
-# Enter Debug Context
-
-# options(error = recover, warn = 2)
-
-offline_data <- lapply(offline_valid, processLine)
-offline <- as.data.table(do.call("rbind", offline_data),
-                         stringsAsFactors = F)
-
-names(offline) <- c("time", "scanMac", "posX", "posY", "posZ",
-                    "orientation", "mac", "signal", 
-                    "channel", "type")
-
-numVars <- c("time", "posX", "posY", "posZ",
-             "orientation", "signal")
-
-dim(offline)
-
-offline[, (numVars) := lapply(.SD, as.numeric), .SDcols = numVars]
-
-offline <- offline[ offline$type == 3, ]
-offline[, type := NULL]
-
-offline[, rawTime := time]
-offline[, time := time/1000]
-class(offline$time) = c("POSIXt", "POSIXct")
-
-# options(error = recover, warn = 1)
-```
-
-#### Data Cleaning
-
-```{r}
-summary(offline)
-```
-
-### Orientation Exploration
-
-```{r}
-length(unique(offline$orientation))
-```
-
-```{r, fig.height=5}
-plot(ecdf(offline$orientation))
-```
-
-```{r}
-ggplot(offline, aes(orientation)) +
-   stat_ecdf() +
-   labs("Orientation")
-```
-
-```{r}
-ggplot(offline, aes(orientation, fill = ..count..)) +
-   geom_histogram(bins = 30) +
-   scale_y_continuous(labels = comma) +
-   labs("Orientation Value Clusters")
-```
-
-```{r}
-
-offline$angle <- roundOrientation(offline$orientation)
-
-# angle = cleaned orientation column
-ggplot(offline, aes(angle)) +
-   stat_ecdf()
-
-ggplot(offline, aes(angle, fill = ..count..)) +
-   geom_histogram(bins = 30) +
-   scale_y_continuous(labels = comma) +
-   labs("Cleaned Up Angles")
-```
-
-```{r}
-with(offline, boxplot(orientation ~ angle,
-                      xlab = "nearest 45 degree angle",
-                      ylab = "orientation"))
-
-ggplot(offline, aes(angle, orientation, group = angle)) +
-   geom_boxplot() +
-   labs(title = "Cleaned Angle vs Raw Orientation", 
-        x = "nearest 45 degree angle", 
-        y = "orientation")
-```
-
-### MAC Address
-
-```{r}
-c(length(unique(offline$mac)), length(unique(offline$channel)))
-
-table(offline$mac)
-
-ggplot(offline, aes(mac, fill = ..count..)) +
-   geom_histogram(stat = "count", bins = 30) +
-   scale_y_continuous(labels = comma)
-
-# keep only the top 7 MAC address data points
-
-dim(offline)
-
-offline_macs <- names(sort(table(offline$mac), decreasing = T)[1:7])
-
-dim(offline)
-```
-
-```{r}
-macChannel <- with(offline, table(mac, channel))
-apply(macChannel, 1, function(x) sum(x > 0))
-
-# mac and channel are 1:1, we can remove channel
-#offline$channel := NULL
-```
-
-### Exploring the Position of the Hand-Held Device
-
-```{r}
-locDF <- with(offline,
-              by(offline, list(posX, posY), function(x) x))
-
-length(locDF)
-
-sum(sapply(locDF, is.null))
-
-locDF <- locDF[ !sapply(locDF, is.null)]
-
-length(locDF)
-
-locCounts <- sapply(locDF, nrow)
-
-locCounts <- sapply(locDF,
-                    function(df)
-                       c(df[1, c("posX", "posY")], count = nrow(df)))
-
-class(locCounts)
-
-dim(locCounts)
-
-locCounts[ , 1:8]
-
-locCountsDF <- as.data.table(t(locCounts))
-
-locCountsDF$posX <- unlist(locCountsDF$posX)
-locCountsDF$posY <- unlist(locCountsDF$posY)
-
-ggplot(locCountsDF, aes(posX, posY, label = count)) +
-   geom_text(angle = 45) +
-   labs(title = "X / Y postion counts")
-```
-
-### Final Data Prep
-
-```{r}
-offline <- readData(file_offline, offline_macs)
-```
-
-#### Signal Strength
-
-```{r, fig.height=7}
-bwplot(signal ~ factor(angle) | mac, data = offline,
-       subset = posX == 2 & posY == 12 & 
-          mac != "00:0f:a3:39:dd:cd",
-       layout = c(2, 3))
-
-summary(offline$signal)
-```
-
-```{r, fig.height=7}
-densityplot(~ signal | mac + factor(angle), data = offline,
-            subset = posX == 24 & posY == 4 &
-               mac != "00:0f:a3:39:dd:cd",
-            bw = 0.5, plot.points = F)
-```
-
-```{r, fig.width=5}
-byLocAngleAP <- with(offline,
-                     by(offline, list(posXY, angle, mac),
-                        function(x) x))
-signalSummary <-
-   lapply(byLocAngleAP,
-          function(oneLoc) {
-             ans = oneLoc[1, ]
-             ans$medSignal = median(oneLoc$signal)
-             ans$avgSignal = mean(oneLoc$signal)
-             ans$num = length(oneLoc$signal)
-             ans$sdSignal = sd(oneLoc$signal)
-             ans$iqrSignal = IQR(oneLoc$signal)
-             ans
-          })
-
-offlineSummary <- do.call("rbind", signalSummary)
-
-breaks <- seq(-90, -30, by = 5)
-
-bwplot(sdSignal ~ cut(avgSignal, breaks = breaks),
-       data = offlineSummary,
-       subset = mac != "00:0f:a3:39:dd:cd",
-       xlab = "Mean Signal", ylab = "SD Signal")
-```
-
-```{r}
-with(offlineSummary,
-     smoothScatter((avgSignal - medSignal) ~ num,
-                   xlab = "Number of Observations",
-                   ylab = "mean - median"))
-abline(h = 0, col = "#984ea3", lwd = 2)
-
-lo.obj <- with(offlineSummary,
-               loess(diff ~ num,
-                     data = data.frame(diff = (avgSignal - medSignal),
-                                       num = num)))
-
-lo.obj.pr <- predict(lo.obj, newdata = data.frame(num = (70:120)))
-lines(x = 70:120, y = lo.obj.pr, col = "#4daf4a", lwd = 2)
-```
-
-#### Signal and Distance
-
-```{r, fig.height=6}
-
-subMacs <- names(sort(table(offline$mac), decreasing = T))[1:7]
-
-surfaceSS <- function(data, mac, angle) {
-   oneAPAngle <- subset(offlineSummary, mac == mac & angle == angle)
-
-   smothSS <- Tps(oneAPAngle[, c("posX", "posY")],
-                  oneAPAngle$avgSignal)
-
-   vizSmooth <- predictSurface(smothSS)
-
-   plot.surface(vizSmooth, type = "C")
-   points(oneAPAngle$posX, oneAPAngle$posY, pch=19, cex = 0.5)
-}
-
-surfaceSS(offlineSummary, subMacs[5], 0)
-```
-
-```{r, fig.height = 7}
-parCur <- par(mfrow = c(2,2), mar = rep(1, 4))
-mapply(surfaceSS, mac = subMacs[ rep(c(5, 1), each = 2)],
-       data = list(data = offlineSummary))
-par(parCur)
-```
-
-Exclude one of two similar access points
-
-```{r}
-offlineSummary <- subset(offlineSummary, mac != subMacs[2])
-```
-
-```{r}
-AP <- matrix( c( 7.5, 6.3, 2.5, -.8, 12.8, -2.8,
-                 1, 14, 33.5, 9.3, 33.5, 2.8),
-              ncol = 2, byrow = T,
-              dimnames = list(subMacs[ -2 ], c("x", "y") ))
-
-ggplot(data.table(mac = rownames(AP), AP), aes(x, y)) +
-   geom_point()
-```
-
-```{r, fig.height=7}
-diffs <- offlineSummary[, c("posX", "posY")] - AP[offlineSummary$mac, ]
-
-offlineSummary$dist <- sqrt(diffs[, 1]^2 + diffs[, 2]^2)
-
-xyplot(signal ~ dist | factor(mac) + factor(angle),
-       data = offlineSummary, pch = 19, cex = 0.3,
-       xlab = "distance")
-```
-
-### Nearest Neighbor Methods to Predict Location
-
-```{r}
-macs <- unique(offlineSummary$mac)
-
-online <- readData(file_online, submacs = macs)
-
-length(unique(online$posXY))
-```
-
-```{r}
-tabonlineXYA = table(online$posXY, online$angle)
-tabonlineXYA[1:6, ]
-```
-
-```{r}
-keepVars <- c("posXY", "posX", "posY", "orientation", "angle")
-
-byLoc <- with(online,
-              by(online, list(posXY),
-                 function(x) {
-                    ans <- x[1, ..keepVars]
-                    avgSS <- tapply(x$signal, x$mac, mean)
-                    y = matrix(avgSS, nrow = 1, ncol = 6,
-                               dimnames = list(ans$posXY, names(avgSS)))
-                    cbind(ans, y)
-                 }))
-
-onlineSummary <- do.call("rbind", byLoc)
-
-dim(onlineSummary)
-
-names(onlineSummary)
-```
-
-#### Choice of Orientation
-
-```{r}
-
-reshapeSS <- function(data, 
-                      varSignal = "signal",
-                      keepVars = c("posXY", "posX", "posY"),
-                      sampleAngle = F) {
-
-   if(sampleAngle)
-      data <- data[angle == sample(data$angle, size = 1), ]
-
-   byLocation <- with(data,
-                 by(data, list(posXY),
-                    function(x) {
-                       ans <- x[1, ..keepVars]
-                       avgSS <- tapply(x$signal, x$mac, mean)
-                       y = matrix(avgSS, nrow = 1, dimnames = list(ans$posXY, names(avgSS)))
-                       cbind(ans, y)
-                    }))
-
-   newDataSS <- do.call("rbind", byLocation)
-
-   col_names <- colnames(newDataSS)
-   to_change <- !(col_names %in% keepVars)
-
-   n_cols <- length(col_names)
-   start <- length(keepVars)
-
-   colnames(newDataSS)[to_change] <- sapply(col_names[to_change], function(col) {
-      n <- nchar(col)
-      substr(col, n - 2, n)
-   })
-
-   newDataSS[, start:ncol(newDataSS)] <- round(newDataSS[, start:ncol(newDataSS)])
-
-   return(newDataSS)
-}
-
-selectTrain <- function(angleNewObs, signals, m) {
-
-   refs <- seq(0, by = 45, length = 8)
-
-   nearestAngle <- roundOrientation(angleNewObs)
-
-   if( m %% 2 == 1) {
-      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)
-   } else {
-      m = m + 1
-      angles <- seq(-45 * (m - 1) / 2, 45 * (m - 1) / 2, length = m)
-
-      if( sign(angleNewObs - nearestAngle) >= 1)
-         angles = angles[ -1 ]
-      else
-         angles = angles[ -m ]
-   }
-
-   angles <- angles + nearestAngle
-   angles[angles < 0] = angles[ angles < 0] + 360
-   angles[angles > 360] = angles[ angles > 360 ] - 360
-
-   signals <- signals[angle %in% angles, ]
-
-   reshapeSS(signals, varSignal = "avgSignal")
-}
-
-train130 <- selectTrain(130, offlineSummary, m = 3)
-
-dim(train130)
-
-findNN <- function(newSignal, trainSubset) {
-   diffs <- apply(trainSubset[ , 4:9], 1,
-                  function(x) x - newSignal)
-   dists <- apply(diffs, 2, function(x) sqrt(sum(x^2)) )
-   closest <- order(dists)
-   return(trainSubset[closest, 1:3 ])
-}
-
-predXY <- function(newSignals, newAngles, trainData,
-                   numAngles = 1, k = 3) {
-
-   closeXY <- list(length = nrow(newSignals))
-
-   for(i in 1:nrow(newSignals)) {
-      trainSS <- selectTrain(newAngles[i], trainData, m = numAngles)
-      closeXY[[i]] = 
-         findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
-   }
-
-   estXY = lapply(closeXY,
-                  function(x) sapply(x[, 2:3],
-                                     function(x) mean(x[1:k])))
-
-   estXY <- do.call("rbind", estXY)
-
-   return(estXY)
-}
-
-estXYk1 <- predXY(newSignals = onlineSummary[, 6:11],
-                  newAngles = onlineSummary[, 4],
-                  offlineSummary, numAngles = 3, k = 1)
-
-estXYk3 <- predXY(newSignals = onlineSummary[, 6:11],
-                  newAngles = onlineSummary[, 4],
-                  offlineSummary, numAngles = 3, k = 3)
-```
-
-```{r}
-calcError <- function(estXY, actualXY)
-   sum( rowSums( (estXY - actualXY) ^ 2) )
-
-actualXY <- onlineSummary[, c("posX", "posY")]
-sapply(list(estXYk1, estXYk3), calcError, actualXY)
-
-```
-
-```{r}
-v <- 11;
-
-permuteLocs <- sample(unique(offlineSummary$posXY))
-permuteLocs <- matrix(permuteLocs, ncol = v,
-                      nrow = floor(length(permuteLocs)/v))
-
-```
-
-```{r}
-onlineFold <- subset(offlineSummary, posXY %in% permuteLocs[, 1])
-
-onlineCVSummary <- reshapeSS(offline, 
-                              keepVars = c("posXY", "posX", "posY", "angle"),
-                              sampleAngle = T)
-
-onlineFold <- subset(onlineCVSummary,
-                     posXY %in% permuteLocs[ , 1])
-
-estFold <- predXY(newSignals = onlineFold[, 6:11],
-                  newAngles = onlineFold[, 4],
-                  offlineFold, numAngles = 3, k = 3)
-
-actualFold <- onlineFold[, c("posX", "posY")]
-calcError(estFold, actualFold)
-
-```
-
-```{r}
-K <- 20
-err <- rep(0, K)
-
-for(j in 1:v) {
-   onlineFold <- subset(onlineCVSummary,
-                        posXY %in% permuteLocs[, j])
-   offlineFold <- subset(offlineSummary,
-                         posXY %in% permuteLocs[ , -j])
-
-   actualFold <- onlineFold[, c("posX", "posY")]
-
-   for(k in 1:K) {
-      estFold <- predXY(newSignals = onlineFold[, 6:11],
-                        newAngles = onlineFold[, 4],
-                        offlineFold, numAngles = 3, k = k)
-
-      err[k] <- err[k] + calcError(estFold, actualFold) 
-   }
-}
-
-```
-
-```{r}
-k_values <- data.table(K = 1:K, Error = err, Min = which.min(err))
-
-ggplot(k_values, aes(K, Error)) +
-   geom_line(aes(col = Error)) +
-   geom_hline(data = k_values[K == Min], aes(yintercept = Error), col = "darkblue") +
-   labs("K-NN Training Error")
-```
-
-```{r}
-estXYk5 <- predXY(newSignals = onlineSummary[, 6:11],
-                 newAngles = onlineSummary[, 4],
-                 offlineSummary, numAngles = 3, k = 5)
-```
-
-```{r}
-calcError(estXYk5, actualXY)
-```
-
-```{r}
-error_table <- data.table(K = c(1, 3, 5), 
-                          Error = c(calcError(estXYk1, actualXY), 
-                                    calcError(estXYk3, actualXY), 
-                                    calcError(estXYk5, actualXY)))
-
-ggplot(error_table, aes(K, Error, fill = -Error)) +
-   geom_bar(stat = "identity")
-```
-
-```{r}
-predXY <- function(newSignals, newAngles, trainData, numAngles = 1, k = 3) {
-   chooseXY <- list(length = nrow(newSignals))
-
-   for(i in 1:nrow(newSignals)) {
-      trainSS <- selectTrain(newangles[i], trainData, m = numAngles)
-      closeXY[[i]] <- findNN(newSignal = as.numeric(newSignals[i, ]),
-                             trainSS)
-   }
-
-   estXY <- lapply(closeXY, function(x)
-                              sapply(x[, 2:3],
-                                     function(x) mean(x[1:k])))
-   estXY <- do.call("rbind", estXY)
-
-   return(XY)
-}
-
-```
-
-### Follow-up
-
-Write the code to read the raw training data into the data structure in the first approach described in section 1.2.
-That is, the data structure is a data frame with a column for each MAC address that detected a signal.
-For the column name, use the last two characters of the MAC address, or some other unique identifier.
-
-```{r}
-
-```

### 2020-04-18 Case Studies/02_data_pull.R

@@ -1,0 +1,60 @@
+library(data.table)
+library(RCurl)
+library(XML)
+library(stringr)
+library(tidyr)
+library(here)
+
+data.dir <- file.path(here::here(), "Case Studies", "datasets")
+
+Url <- "http://www.cherryblossom.org/results/1999/cb99f.html"
+
+
+theurl <- Url
+
+webpage <- getURL(theurl)
+webpage <- readLines(tc <- textConnection(webpage)); close(tc)
+
+pagetree <- htmlTreeParse(webpage, error=function(...){}, useInternalNodes = TRUE)
+
+pre_data <- pagetree["//pre"]
+
+lines <- sapply(pre_data, xmlValue)
+
+table_long <- strsplit(lines, "\n")[[1]]
+
+total_rows <- length(table_long)
+row_data <- data.table(values = table_long[-1], stringsAsFactors = F)
+
+num_rows <- nrow(row_data)
+to_parse <- row_data[3:num_rows]
+
+parse_row <- function(row) {
+  
+  row_num <- as.numeric(substr(row, 1, 5))
+  
+  place <- substr(row, 6, 15)
+  full_name <- substr(row, 16, 38)
+  age <- substr(row, 38, 40)
+  hometown <- substr(row, 41, 60)
+  time <- substr(row, 61, 68)
+  pace <- substr(row, 69, 76)
+  
+  return(data.frame(row_num, place, full_name, age, hometown, time, pace))
+}
+
+parsed <- apply(to_parse, 2, parse_row)
+
+results <- data.table::rbindlist(parsed)
+
+
+test_row <- row_data[3, 1]
+test_row
+
+substr(test_row, 1, 5)
+substr(test_row, 6, 15)
+substr(test_row, 16, 38)
+substr(test_row, 38, 40)
+substr(test_row, 41, 60)
+substr(test_row, 61, 68)
+substr(test_row, 69, 76)

### 2020-04-19 Case Studies/02_race-times.rmd
updates to data scraping for cherry blossom
updates to data scraping for cherry blossom

@@
 
-data.dir <- file.path(here::here(), "Case Studies", "01_data")
+data.dir <- file.path(here::here(), "Case Studies", "datasets")
 
@@
 
-# Geolocation
+# Cherry Blossom Ten Mile Run & Walk
 
-## Predicting Location via Indoor Positioning Systems
+## Examining the Impact of Age on Physical Performance
 
@@
 ```
+
+### Further Analysis
+
+### 1.)
+
+Write a function that uses _read.fwf()_ to read the 28 text tables in MenTxt/ and WomenTxt/ into R.
+These are called 1999.text, 2000txt, etc.
+and are described in greater detail in 2.2.
+Examine the tables in a plain text editor to determine the start and end position of each column of interest (name, hometown, age and gun and net time).
+
+Use statistics to explore the results and confirm that you have extracted the information from the correct positions in the text.
+
+```{r}
+
+```
+
+### 2.)
+
+Revise the extractVariables function (see section 2.2) to remove the rows in _menTables_ that are blank.
+In addition, eliminatee the rows that begin with a '*' or a '#'.
+You may find the following regular expression helpful for locating blank rows in a table.
+
+_grep("^[[:blank:]]*$", body)_
+
+The pattern uses several meta characters.
+The ^ is an anchor for the start of the string, the $ anchors to the end of the string, the [[:blank:]] denotes the equivalence class of any space or tab character, and * indicates that the blank character can appear 0 or more times.
+All together the pattern ^[[:blank:]]*\$ matches a string that contains any number of blanks from start to end.
+
+```{r}
+
+```
+
+### 3.)
+
+Find the record where the time is only 1.5.
+What happened?
+Determine how to handle the problem and which function needs to be modified: _extractResTable()_, _extractVariables()_, or _cleanUp()_.
+In your modification, include code to provide a warning message about the rows that are being dropped for having a time that is too small.
+
+### 4.)
+
+Examine the head and tail of the 2006 men's file.
+Look at both the character matrix in the list called _menResMat_ and the character vector in the list called _menFiles_ (see Sec 2.2).
+(Recall that the desired character matrix in _menResMat_ and the character vector in _menFiels_ both correspond to the element named "2006").
+What is wrong with the hometown?
+Examine the header closely to figure out how this error came about.
+Modify the _extractVariables()_ function to fix the problem.

### 2020-04-19 Case Studies/02_data_pull.R

@@ -4,30 +4,9 @@ library(XML)
 library(stringr)
 library(tidyr)
 library(here)
+library(urltools)
 
 data.dir <- file.path(here::here(), "Case Studies", "datasets")
-
-Url <- "http://www.cherryblossom.org/results/1999/cb99f.html"
-
-
-theurl <- Url
-
-webpage <- getURL(theurl)
-webpage <- readLines(tc <- textConnection(webpage)); close(tc)
-
-pagetree <- htmlTreeParse(webpage, error=function(...){}, useInternalNodes = TRUE)
-
-pre_data <- pagetree["//pre"]
-
-lines <- sapply(pre_data, xmlValue)
-
-table_long <- strsplit(lines, "\n")[[1]]
-
-total_rows <- length(table_long)
-row_data <- data.table(values = table_long[-1], stringsAsFactors = F)
-
-num_rows <- nrow(row_data)
-to_parse <- row_data[3:num_rows]
 
 parse_row <- function(row) {
   
@@ -43,18 +22,206 @@ parse_row <- function(row) {
   return(data.frame(row_num, place, full_name, age, hometown, time, pace))
 }
 
-parsed <- apply(to_parse, 2, parse_row)
+url <- "http://www.cherryblossom.org/results/2003/CB03-M.HTM"
+rm(url)
 
-results <- data.table::rbindlist(parsed)
+save_data <- function(url) {
+  
+  url <- tolower(url)
+  parts <- unlist(strsplit(url, "/"))
+  year <- parts[5]
+  
+  mf <- str_detect(unlist(strsplit(parts[6], "\\."))[1], "m")
+  
+  file_path <- data.dir
+  
+  if(mf) {
+    file_path <- file.path(file_path, "race_men", paste0(year, "m", ".csv"))
+  } else {
+    file_path <- file.path(file_path, "race_women", paste0(year, "f", ".csv"))
+  }
+
+  webpage <- getURL(url)
+  webpage <- readLines(tc <- textConnection(webpage)); close(tc)
+  
+  pagetree <- htmlTreeParse(webpage, error=function(...){}, useInternalNodes = TRUE)
+  
+  pre_data <- pagetree["//pre"]
+  
+  lines <- sapply(pre_data, xmlValue)
+  
+  table_long <- strsplit(lines, "\n")[[1]]
+  
+  total_rows <- length(table_long)
+  row_data <- data.table(values = table_long[-1], stringsAsFactors = F)
+  
+  num_rows <- nrow(row_data)
[…]

### 2020-04-19 Case Studies/02_race-times.rmd
cherry blossom race work.
cherry blossom race work.

@@
 library(data.table, quietly = TRUE, warn.conflicts = FALSE)
+library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
 library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
@@
 
+data is pre-processed from the web in \02_data.r
+
 ```{r}
+
+read_results <- function(class, year) {
+
+   folder <- ifelse(class == "men", "men_txt", "women_txt")
+   file <- file.path(data.dir, folder, paste(year, ".txt", sep = ""))
+   return(readLines(file))
+}
+
+```
+
+### Race Results Preprocessing
+
+```{r}
+
+els <- read_results("men", 2006)
+
+eqIndex <- grep("^===", els)
+eqIndex
+
+spacerRow <- els[eqIndex]
+headerRow <- els[eqIndex - 1]
+body <- els[ -(1:eqIndex) ]
+
+headerRow <- tolower(headerRow)
+
+timeStart <- regexpr("net", headerRow)
+time <- substr(body, start = timeStart, stop = timeStart + 1)
+
+ageStart <- regexpr("ag", headerRow)
+
+age <- substr(body, start = ageStart, stop = ageStart + 1)
+head(age)
+
+summary(as.numeric(age))
+
+blankLocs <- gregexpr(" ", spacerRow)
+blankLocs
+
+searchLocs <- c(0, blankLocs[[1]])
+
+Values <- mapply(substr, list(body),
+                 start = searchLocs[ -length(searchLocs)] + 1,
+                 stop = searchLocs[ -1 ] - 1)
+
+findColLocs <- function(spacerRow) {
+   spaceLocs <- gregexpr(" ", spacerRow)[[1]]
+   rowLength <- nchar(spacerRow)
+
+   if(substring(spacerRow, rowLength, rowLength + 1) != " ")
+      return(c(0, spaceLocs, rowLength + 1))
+   else
+      return(c(0, spaceLocs))
+}
+
+findColLocs(spacerRow)
+name <- "home"
+colnames <- c("name", "home", "ag", "gun", "net", "time")
+
+colIndex <- which(colnames == name)
+
+startPos <- regexpr(name, headerRow)[[1]]
+
+#
+# can we modify select cols to short circuit on matching w/o spaces?
+#######
+
+selectCols =
+   function(colNames, headerRow, searchLocs)
+   {
+      sapply(colNames,
+             function(name, headerRow, searchLocs)
+            {
+               startPos <- regexpr(name, headerRow)[[1]]
+
+               if(startPos == -1)
+                  return( c(NA, NA) )
+
+               index <- sum(startPos >= searchLocs)
+
+               c(searchLocs[index] + 1, searchLocs[index + 1])
+            },
+      headerRow = headerRow, searchLocs = searchLocs)
+   }
+
+searchLocs <- findColLocs(spacerRow)
+loc <- selectCols("home", headerRow, searchLocs)
+vars <- mapply(substr, list(body), start = loc[1,], stop = loc[2, ])
+
+summary(as.numeric(ages))
+
+shortColNames <- c("name", "home", "ag", "gun", "net", "time")
+
+locCols <- selectCols(shortColNames, headerRow, searchLocs)
+Values <- mapply(substr, list(body), start = locCols[1, ],
+                 stop = locCols[2, ])
+
+class(Values)
+
+colnames(Values) = shortColNames
+
+head(Values)
+tail(Values)[, 1:3]
+```
+
+```{r}
+years <- 1999:2012
+
+res <- sapply(years, FUN = function(year){
+
+   print(paste("loading year...", year))
+
+   file <- read_results("men", year)
+
+   parsed <- extractVariables(file)
+})
+
+file <- read_results("men", 2006)
+m2006 <- extractVariables(file)
+
+age <- m2006[, "home"]
+
+sum(is.na(age))
+
+n <- length(file)
+
+file[n]
+
+m2001$ag
+
+extractVariables =
+   function(file, varNames = c("name", "home", "ag", "gun", "net", "time"))
+   {
+      # Find the index of the row with =s
+      eqIndex <- grep("^===", file)
+      spacerRow <- file[eqIndex]
+      headerRow <- tolower(file[ eqIndex - 1 ])
+      body <- file[ -(1 : eqIndex) ]
+
+      blank <- grep("^[[:blank:]]*$", body)
+      footnote <- grep("^[^\\s]*[\\*]*[#]", body)
+      ignore <- union(blank, footnote)
+
+      if(length(ignore))
+         body <- body[-ignore]
+
+      # Obtain the starting and ending positions of variables
+      searchLocs <- findColLocs(spacerRow)
+      locCols <- selectCols(varNames, headerRow, searchLocs)
+
+      Values <- mapply(substr, list(body), start = locCols[1, ],
+                       stop = locCols[2, ])
+      colnames(Values) <- varNames
+
+      invisible(Values)
+   }
+
+load_data <- function(class) {
+   folder <- ifelse(class == "men", "men_txt", "women_txt")
+   filenames <- file.path(data.dir, paste(folder, "/", 1999:2012, ".txt", sep = ""))
+   files <- lapply(filenames, readLines)
+   names(files) <- 1999:2012
+
+   mat <- lapply(files, extractVariables)
+
+   return(mat)
+}
+
+mfilenames <- file.path(data.dir, paste("men_txt/", 1999:2012, ".txt", sep = ""))
+menFiles <- lapply(mfilenames, readLines)
+names(menFiles) <- 1999:2012
+
+menResMat <- lapply(menFiles, extractVariables)
+
+length(menResMat)
+sapply(menResMat, nrow)
+```
+
+### Data Cleaning and Reformatting Variables
+
+```{r}
+menResMat <- load_data("men"); womenResMat <- load_data("women")
+
+length(menResMat)
+sapply(menResMat, nrow)
+
+length(womenResMat)
+sapply(womenResMat, nrow)
+```
+
+Age Validation
+
+```{r}
+age <- as.numeric(menResMat[['2012']][, 'ag'])
+
+tail(age)
+
+age <- sapply(menResMat,
+              function(x) as.numeric(x[, "ag"]))
+
+age_values <- plyr::ldply(age, data.frame)
+colnames(age_values) <- c("Year", "Age")
+
+ggplot(age_values, aes(Age)) +
+   geom_histogram(aes(fill = ..count..), bins = 30)
+
+ggplot(age_values, aes(Year, Age)) +
+   geom_boxplot(aes(fill = Year)) +
+   labs(title = "Age By Race Year")
+```
+
+```{r}
+sapply(age, function(x) sum(is.na(x)))
+
+file2001 <- read_results("men", 2001)
+age2001 <- as.numeric(extractVariables(file2001)[, "ag"])
+
+badAgeIndex <- which(is.na(age2001)) + 5
+
+file2001[badAgeIndex]
+
+badAgeIndex
+
+blanks <- grep("^[[:blank:]]*$", file2001)
+blanks
+
+which(age2001 < 5)
+
+file2001[which(age2001 < 5)]
+
+file <- read_results("men", 2001)
+data <- extractVariables(file)
+age <- as.numeric(data[, "ag"])
+
+sum(age < 5)
+
+which(age == 0)
+
+data[which(age == 0)]
+
+```
+
+### Time Validation
+
+```{r}
+file2002 <- read_results("men", 2012)
+
+charTime <- menResMat[["2012"]][, "time"]
+
+head(charTime)
+tail(charTime)
+
+timePieces <- strsplit(charTime, ":")
+
+timePieces[[1]]
+tail(timePieces, 1)
+
+timePieces <- sapply(timePieces, as.numeric)
+
+runTime <- sapply(timePieces,
+                  function(x) {
+                     if(length(x) == 2) x[1] + x[2]/60
+                     else 60 * x[1] + x[2] + x[3]/60
+                  })
+
+summary(runTime)
+
+convertTime <- function( charTime ) {
+
+   timePieces <- strsplit(charTime, ":")
+
+   timePieces <- sapply(timePieces, as.numeric)
+
+   runTime <- sapply(timePieces,
+                  function(x) {
+                     if(length(x) == 2) x[1] + x[2]/60
+                     else 60 * x[1] + x[2] + x[3]/60
+                  })
+
+}
+
+```
+
+### Aggregate cleaning into
+
+```{r}
+createDF =
+   function(Res, year, sex) {
+      useTime <- if(!is.na(Res[1, 'net']) )
+                     Res[, 'net']
+                  else if( !is.na(Res[1, 'gun']) )
+                     Res[ , 'gun']
+                  else
+                     Res[, 'time']
+
+      useTime <- gsub("[#\\*[:blank:]]", "", useTime)
+
+      Res <- Res[ useTime != "", ]
+
+      runTime <- convertTime(useTime[ useTime != "" ])
+
+      N <- nrow(Res)
+
+      Results <- data.frame( year = rep(year, N),
+                             sex = rep(sex, N),
+                             name = Res[, 'name'],
+                             home = Res[, 'home'],
+                             age = as.numeric(Res[, 'ag']),
+                             runTime = runTime,
+                             stringsAsFactors = F)
+
+      invisible(Results)
+   }
+
+menDF <- mapply(createDF, menResMat, year = years,
+                sex = rep("M", 14), SIMPLIFY = F)
+
+warnings()[ c(1:2, 49:50)]
+
+sapply(menDF, function(x) sum(is.na(x$runTime)))
+
+file2006 <- read_results("men", 2006)
+parsed2006 <- extractVariables(file2006)
+time2006 <- parsed2006[, "net"]
 
@@
 
+extractVariables =
+   function(file, varNames = c("name", "home", "ag", "gun", "net", "time"))
+   {
+      # Find the index of the row with =s
+      eqIndex <- grep("^===", file)
+      spacerRow <- file[eqIndex]
+      headerRow <- tolower(file[ eqIndex - 1 ])
+      body <- file[ -(1 : eqIndex) ]
+
+      blank <- grep("^[[:blank:]]*$", body)
+      footnote <- grep("^[^\\s]*[\\*]*[#]", body)
+      ignore <- union(blank, footnote)
+
+      if(length(ignore))
+         body <- body[-ignore]
+
+      # Obtain the starting and ending positions of variables
+      searchLocs <- findColLocs(spacerRow)
+      locCols <- selectCols(varNames, headerRow, searchLocs)
+
+      Values <- mapply(substr, list(body), start = locCols[1, ],
+                       stop = locCols[2, ])
+      colnames(Values) <- varNames
+
+      invisible(Values)
+   }
+
 ```
@@
 Modify the _extractVariables()_ function to fix the problem.
+
+```{r}
+
+```
+
+### 5.)
+
+Write the _convertTime()_ function described in Section 2.3.
+This function takes a string where time is in either the format hh:mm:ss or mm:ss.
+The return value is the time as numeric value of the number of minutes.
+Design this function to take a character vector with multiple strings and return a numeric vector.
+
+```{r}
+
+```
+
+### 6.)
+
+Modify the _createDF()_ function in Section 2.3 to handle the formatting problem with the 2006 male file. You will need to carefully inspect the raw text file in order to determien the problem.
+
+```{r}
+
+```
+
+### 7.)
+
+Follow the approach developed in Section 2.2 to read the files for the female runners and then process them using the functions in Section 2.3 to create a data frame for analysis.
+You may need to generalize the _createDF()_ and _extractVariables()_ functions to handle additional oddities in the raw text files.
+
+```{r}
+
+```

### 2020-04-19 Case Studies/02_data_pull.R

@@ -169,6 +169,7 @@ race_info_men <- data.frame(url = paste(ubase, mens_result_urls, sep=""), year =
 race_info <- rbind(race_info_men, race_info_women)
 
 save_class_result(race_info_men)
+save_class_result(race_info_women)
 
 sapply(result_tables, length)
 

### 2020-04-19 Case Studies/02_race-times.rmd
cherry blossom work
cherry blossom work

@@
 
-summary(as.numeric(ages))
+summary(as.numeric(vars))
 
@@
 ```{r}
-years <- 1999:2012
-
-res <- sapply(years, FUN = function(year){
-
-   print(paste("loading year...", year))
-
-   file <- read_results("men", year)
-
-   parsed <- extractVariables(file)
-})
-
-file <- read_results("men", 2006)
-m2006 <- extractVariables(file)
-
-age <- m2006[, "home"]
-
-sum(is.na(age))
-
-n <- length(file)
-
-file[n]
-
-m2001$ag
 
@@
 
-mfilenames <- file.path(data.dir, paste("men_txt/", 1999:2012, ".txt", sep = ""))
-menFiles <- lapply(mfilenames, readLines)
-names(menFiles) <- 1999:2012
-
-menResMat <- lapply(menFiles, extractVariables)
-
-length(menResMat)
-sapply(menResMat, nrow)
 ```
@@
 
+years <- 1999:2012
 menDF <- mapply(createDF, menResMat, year = years,

### 2020-04-21 Case Studies/02_race-times.rmd
cherry blossom race analysis
cherry blossom race analysis

@@
 library(fields, quietly = TRUE, warn.conflicts = FALSE)
+library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-read_results <- function(class, year) {
+load_results <- function(class, year) {
 
@@
    file <- file.path(data.dir, folder, paste(year, ".txt", sep = ""))
+
    return(readLines(file))
@@
 
-els <- read_results("men", 2006)
+els <- load_results("men", 2006)
 
@@
 headerRow <- els[eqIndex - 1]
+
 body <- els[ -(1:eqIndex) ]
@@
 ```{r}
-menResMat <- load_data("men"); womenResMat <- load_data("women")
+menResMat <- load_data("men")
 
@@
 sapply(menResMat, nrow)
+
+womenResMat <- load_data("women")
 
@@
 ```{r}
+
 createDF =
@@
 
+womenDF <- mapply(createDF, womenResMat, year = years,
+                  sex = rep("F", 14), SIMPLIFY = F)
+
+sapply(womenDF, function(x) sum(is.na(x$runTime)))
+```
+
+### Consolidate Results
+
+```{r}
+cbMen <- do.call(rbind, menDF)
+save(cbMen, file = file.path(data.dir, "cbMen.rda"))
+
+cbWomen <- do.call(rbind, womenDF)
+
+save(cbWomen, file = file.path(data.dir, "cbWomen.rda"))
+```
+
+### Exploratory Data Analysis
+
+```{r}
+
+ggplot(cbMen, aes(age, runTime)) +
+   geom_point(aes(col = age)) +
+   labs(main = "Run Times by Age", xlab = "Age (years)", xlab = "Run Time (minutes")
+
+```
+
+```{r}
+display.brewer.all()
+```
+
+```{r}
+Purple8 <- brewer.pal(9, "Purples")[8]
+
+Purple8A <- paste(Purple8, "14", sep = "")
+
+ggplot(cbMen, aes(age, runTime)) +
+   geom_jitter(col = Purple8A) +
+   geom_smooth() +
+   xlim(10, 75) +
+   ylim(45, 150)
+
+```
+
+```{r}
+smoothScatter(y = cbMen$runTime, x = cbMen$age,
+              ylim = c(40, 165), xlim = c(15, 85),
+              xlab = "Age (years)", ylab = "Run Time (minutes)")
+```
+
+```{r}
+cbMenSub <- cbMen[cbMen$runTime > 30 &
+                     !is.na(cbMen$age) & cbMen$age > 15,]
+
+ageCat <- cut(cbMenSub$age, breaks = c(seq(15, 75, 10), 90))
+
+ggplot(data.table(age = ageCat, time = cbMenSub$runTime), aes(ageCat, time)) +
+   geom_boxplot(aes(fill = age)) +
+   labs("Run Times by Age Category")
+```
+
+```{r}
+lm_age <- lm(runTime ~ age, data = cbMenSub)
+
+summary(lm_age)
+```
+
+```{r}
+smoothScatter(x = cbMenSub$age, y = lm_age$residuals,
+              xlab = "Age (years)", ylab = "Residuals") 
+abline(h = 0, col = "purple", lwd = 3)
+
+resid.lo <- loess(resids ~ age,
+                  data = data.frame(resids = residuals(lm_age),
+                                    age = cbMenSub$age))
+
+age20to80 <- 20:80
+
+resid.lo.pr <- predict(resid.lo, newdata = data.frame(age = age20to80))
+
+lines(x = age20to80, y = resid.lo.pr,
+      col = "green", lwd = 3, lty = 2)
+```
+
+```{r}
+menRes.lo <- loess(runTime ~ age, cbMenSub)
+
+summary(menRes.lo)
+
+menRes.lo.pr <- predict(menRes.lo, data.frame(age = age20to80))
+
+```
+
+```{r}
+over50 <- pmax(0, cbMenSub$age - 50)
+
+lmOver50 <- lm(runTime ~ age + over50, data = cbMenSub)
+
+summary(lmOver50)
+
+decades <- seq(30, 60, by = 10)
+overAge <- lapply(decades,
+                  function(x) pmax(0, (cbMenSub$age - x)))
+names(overAge) <- paste("over", decades, sep = "")
+overAge <- as.data.frame(overAge)
+
+tail(overAge)
+
+lmPiecewise <- lm(runTime ~., 
+                  data = cbind(cbMenSub[, c("runTime", "age")],
+                               overAge))
+
+summary(lmPiecewise)
+
+overAge20 <- lapply(decades, function(x) pmax(0, (age20to80 - x)))
+names(overAge20) <- paste("over", decades, sep = "")
+overAgeDF <- cbind(age = data.frame(age = age20to80), overAge20)
+
+tail(overAgeDF)
+```
+
+```{r}
+predPiecewise <- predict(lmPiecewise, overAgeDF)
+
+plot(predPiecewise ~ age20to80,
+     type = "l", col = "purple", lwd = 3,
+     xlab = "Age (years)", ylab = "Run Time Prediction")
+
+lines(x = age20to80, y = menRes.lo.pr,
+      col = "green", lty = 2, lwd = 3)
+
+legend("topleft", col = c("purple", "green"),
+       lty = c(1, 2), lwd = 3,
+       legend = c("Piecewise Linear", "Loess Curve"), bty = "n")
+```
+
+### Cross-Sectional Data and Covariates
+
+```{r}
+age1999 <- cbMenSub[ cbMenSub$year == 1999, "age"]
+age2012 <- cbMenSub[ cbMenSub$year == 2012, "age"]
+```
+
+```{r}
+summary(age1999)
+
+summary(age2012)
+```
+
+```{r}
+plot(density(age1999, na.rm = T),
+     ylim = c(0, 0.05), col = "purple",
+     lwd = 3, xlab = "Age (years)", main = "")
+lines(density(age2012, na.rm = T),
+      lwd = 3, lty = 2, col = "green")
+legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
+       legend = c("1999", "2012", bty = "n"))
+```
+
+```{r}
+by_year <- data.table(cbMenSub)[, .(age, mean = mean(age)), by = list(year)]
+
+ggplot(by_year[year %in% c("1999", "2012")], aes(age, group = year)) +
+   geom_density(aes(fill = year), alpha = .6) +
+   geom_vline(aes(xintercept = mean, col = year, group = year)) +
+   labs(title = "Runners by Year")
+```
+
+```{r}
+ggplot(by_year, aes(year, age)) +
+   geom_point(aes(col = age)) +
+   geom_quantile() +
+   labs(title = "Runners Age Over Time")
+```
+
+```{r}
+mR.lo99 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 1999,])
+mR.lo.pr99 <- predict(mR.lo99, data.frame(age = age20to80))
+
+summary(mR.lo12)
+
+mR.lo12 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 2012, ])
+mR.lo.pr12 <- predict(mR.lo12, data.frame(age = age20to80))
+
+plot(mR.lo.pr99 ~ age20to80,
+     type = "l", col = "purple", lwd = 3,
+     xlab = "Age (years)", ylab = "Fitted Run Time (minutes)")
+
+lines(x = age20to80, y = mR.lo.pr12,
+      col = "green", lty = 2, lwd = 3)
+
+legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
+       legend =  c("1999", "2012"), bty = "n")
+```
+
+```{r}
+mR.models <- cbMenSub %>% 
+   group_by(year) %>%
+   group_map(~ loess(runTime ~ age, .x))
+
+mR.pred <- lapply(mR.models, 
+                  function(m, year, age) cbind(rep(year, length(age)), age = age, predict(m, data.frame(age = age))),
+                  year = 1999:2012,
+                  age = age20to80)
+
+yoy_changes <- matrix(unlist(mR.pred), ncol = 3)
+yoy_changes <- as.data.frame(yoy_changes)
+colnames(yoy_changes) <- c("year", "pred", "age")
+yoy_changes$year <- as.numeric(yoy_changes$year)
+
+ggplot(yoy_changes[year == 1999], aes(age, pred)) +
+   geom_point()
 ```
@@
 
+```{r}
+
+```
+
 Find the record where the time is only 1.5.
@@
 
+createDF =
+   function(Res, year, sex) {
+      useTime <- if(!is.na(Res[1, 'net']) )
+                     Res[, 'net']
+                  else if( !is.na(Res[1, 'gun']) )
+                     Res[ , 'gun']
+                  else
+                     Res[, 'time']
+
+      useTime <- gsub("[#\\*[:blank:]]", "", useTime)
+
+      Res <- Res[ useTime != "", ]
+
+      runTime <- convertTime(useTime[ useTime != "" ])
+
+      N <- nrow(Res)
+
+      Results <- data.frame( year = rep(year, N),
+                             sex = rep(sex, N),
+                             name = Res[, 'name'],
+                             home = Res[, 'home'],
+                             age = as.numeric(Res[, 'ag']),
+                             runTime = runTime,
+                             stringsAsFactors = F)
+
+      invisible(Results)
+   }
+
 ```

### 2020-04-23 Case Studies/02_race-times.rmd
cherry blossom analysis
cherry blossom analysis

@@
    fig.width = 9, 
-   fig.height = 3.5)
+   fig.height = 4.5)
 
@@
 
-file2001 <- read_results("men", 2001)
+file2001 <- load_results("men", 2001)
 age2001 <- as.numeric(extractVariables(file2001)[, "ag"])
@@
 
-file <- read_results("men", 2001)
+file <- load_results("men", 2001)
 data <- extractVariables(file)
@@
 ```{r}
-file2002 <- read_results("men", 2012)
+file2002 <- load_results("men", 2012)
 
@@
 
-file2006 <- read_results("men", 2006)
+file2006 <- load_results("men", 2006)
 parsed2006 <- extractVariables(file2006)
@@
 
-summary(mR.lo12)
+summary(mR.lo.pr99)
 
@@
 mR.lo.pr12 <- predict(mR.lo12, data.frame(age = age20to80))
+
+summary(mR.lo12)
 
@@
 ```{r}
-mR.models <- cbMenSub %>% 
-   group_by(year) %>%
-   group_map(~ loess(runTime ~ age, .x))
 
-mR.pred <- lapply(mR.models, 
-                  function(m, year, age) cbind(rep(year, length(age)), age = age, predict(m, data.frame(age = age))),
-                  year = 1999:2012,
-                  age = age20to80)
+years <- 1999:2012
+results <- list(length(years))
 
-yoy_changes <- matrix(unlist(mR.pred), ncol = 3)
-yoy_changes <- as.data.frame(yoy_changes)
-colnames(yoy_changes) <- c("year", "pred", "age")
-yoy_changes$year <- as.numeric(yoy_changes$year)
+y <- 1999
 
-ggplot(yoy_changes[year == 1999], aes(age, pred)) +
-   geom_point()
+for( i in 1:length(years) )
+{
+   y = years[i]
+   data <- cbMenSub[ which(cbMenSub$year == y), ]
+
+   model <- loess(runTime ~ age, data)
+
+   pred <- predict(model, newdata = age20to80)
+
+   results[[i]] <- data.table(Year = rep(y, length(pred)), Age = age20to80, Pred = pred)
+}
+
+race_years_data <- do.call("rbind", results)
+
+ggplot(race_years_data, aes(Age, Pred, group = Year)) +
+   geom_line(aes(col = Year)) +
+   labs(title = "Predicted Performance by Year", ylab = "Predicted Run Time", xlab = "Runner Age")
+
+```
+
+### Constructing a Record for an Individual Runner across Years
+
+```{r}
+trimBlanks <- function(charVector) {
+   nameClean <- gsub("^[[:blank:]]+", "", charVector)
+   nameClean <- gsub("[[:blank:]]+$", "", nameClean)
+   nameClean <- gsub("[[:blank:]]+", " ", nameClean)
+}
+
+nameClean <- trimBlanks(cbMenSub$name)
+
+length(nameClean)
+
+length(unique(nameClean))
+
+table(table(nameClean))
+
+```
+
+```{r}
+head(sort(table(nameClean), decreasing = T), 1)
+
+mSmith <- cbMenSub[nameClean == "Michael Smith", ]
+
+head(unique(mSmith$home))
+
+nameClean <- tolower(nameClean)
+
+head(sort(table(nameClean), decreasing = T), 1)
+
+nameClean <- gsub("[,.]", "", nameClean)
+```
+
+```{r}
+tabNameYr <- table(cbMenSub$year, nameClean)
+
+max(tabNameYr)
+
+class(tabNameYr)
+
+mode(tabNameYr)
+
+names(attributes(tabNameYr))
+
+dim(tabNameYr)
+
+head(colnames(tabNameYr), 3)
+
+which(tabNameYr == max(tabNameYr), arr.ind = T)
+
+indMax <- which(tabNameYr == max(tabNameYr), arr.ind = T)
+colnames(tabNameYr)[indMax[2]]
+
+cbMenSub$nameClean <- nameClean
+```
+
+```{r}
+cbMenSub$yob <- cbMenSub$year - cbMenSub$age
+
+homeClean <- trimBlanks(cbMenSub$home)
+homeClean <- tolower(homeClean)
+
+cbMenSub$homeClean <- homeClean
+```
+
+```{r}
+vars <- c("year", "homeClean", "nameClean", "yob", "runTime")
+mb <- which(nameClean == "michael brown")
+birthOrder <- order(cbMenSub$yob[mb])
+cbMenSub[mb[birthOrder], vars]
+```
+
+```{r}
+cbMenSub$ID = paste(nameClean, cbMenSub$yob, sep = "_")
+
+races <- tapply(cbMenSub$year, cbMenSub$ID, length)
+
+races8 <- names(races)[which(races >= 8)]
+
+men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]
+
+orderByRunner <- order(men8$ID, men8$year)
+
+men8 <- men8[orderByRunner, ]
+```
+
+```{r}
+men8L <- split(men8, men8$ID)
+names(men8L)
+```
+
+```{r}
+length(unique(men8$ID))
+length(men8L)
+```
+
+```{r}
+gapTime <- tapply(men8$runTime, men8$ID,
+                  function(t) any(abs(diff(t)) > 20))
+
+gapTime <- sapply(men8L, function(df)
+   any(abs(diff(df$runTime)) > 20))
+
+sum(gapTime)
+```
+
+```{r}
+lapply(men8L[ gapTime ][1:2], function(df) df[, vars])
+```
+
+```{r}
+homeLen <- nchar(cbMenSub$homeClean)
+
+cbMenSub$state <- substr(cbMenSub$homeClean,
+                         start = homeLen - 1, stop = homeLen)
+
+cbMenSub$state[cbMenSub$year == 2006] = NA
+
+cbMenSub$ID = paste(cbMenSub$nameClean, cbMenSub$yob,
+                    cbMenSub$state, sep = "_")
+```
+
+```{r}
+numRaces <- tapply(cbMenSub$year, cbMenSub$ID, length)
+races8 <- names(numRaces)[which(numRaces >= 8)]
+men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]
+
+orderByRunner <- order(men8$ID, men8$year)
+men8 <- men8[orderByRunner, ]
+
+men8L <- split(men8, men8$ID)
+names(men8L) <- races8
+
+length(men8L)
+```
+
+### Modeling the Change in Running Time for Individuals
+
+```{r}
+
 ```

### 2020-04-23 Case Studies/02_data_pull.R

@@ -40,7 +40,7 @@ save_data <- function(url) {
   } else {
     file_path <- file.path(file_path, "race_women", paste0(year, "f", ".csv"))
   }
-
+  
   webpage <- getURL(url)
   webpage <- readLines(tc <- textConnection(webpage)); close(tc)
   
@@ -82,7 +82,7 @@ extractResTable <- function(url, year ) {
   } else {
     txt <- xpathSApply(doc, '//pre', xmlValue)
   }
- 
+  
   result <- character(length = 0L)
   
   if(grepl("\n", txt, fixed = TRUE) == T) {

### 2020-04-27 Case Studies/02_race-times.rmd
finished cherry blossom
finished cherry blossom

@@
 ```{r}
-
+groups <- 1 + (1:length(men8L) %% 9)
+
+plot(x = 40, y = 60, type = "n",
+     xlim = c(20, 80), ylim = c(40, 160),
+     xlab = "Age (years)", ylab = "Run Time (minutes)")
+
+addRunners <- function(listRunners, colors, numLty) {
+   numRunners <- length(listRunners)
+
+   colIndx <- 1 + (1:numRunners) %% length(colors)
+   ltys <- rep(1:numLty, each = length(colors), length = numRunners)
+
+   mapply(function(df, i) {
+      lines(df$runTime ~ df$age,
+            col = colors[colIndex[i]], lwd = 2, lty = ltys[i])
+   }, listRunners, i = 1:numRunners)
+}
+
+colors <- c("#e41alc", "#377eb8", "#4daf4a", "#984ea3",
+            "#ff7f00", "#a65628")
+
+par(mfrow = c(3, 3), mar = c(2, 2, 1, 1))
+invisible(
+   sapply(1:9, function(grpId) {
+      plot(x = 0, y = 0, type = "n",
+           xlim = c(20, 80), ylim = c(50, 130),
+           xlab = "Age (years)", ylab = "Run Time (minutes)")
+
+      addRunners(men8L[ groups == grpId ], colors, numLty = 6)
+   })
+)
+```
+
+```{r}
+fitOne <- function(oneRunner, addLine = F, col = "grey") {
+   lmOne <- lm(runTime ~ age, data = oneRunner)
+
+   if ( addLine )
+      lines(x = oneRunner$age, y = predict(lmOne),
+            col = col, lwd = 2, lty = 2)
+
+   ind <- floor( (nrow(oneRunner) + 1) / 2)
+   res <- c(coefficients(lmOne)[2], oneRunner$age[ind],
+            predict(lmOne)[ ind ])
+
+   names(res) <- c("ageCoeff", "medAge", "predRunTime")
+
+   return(res)
+}
+```
+
+```{r}
+par(mfrow = c(1, 1))
+plot( x = 0, y = 0, type = "n",
+      xlim = c(20, 80), ylim = c(50, 130),
+      xlab = "Age (years)", ylab = "Run Time (minutes)")
+
+addRunners( men8L[ groups == 9 ], colors, numLty = 6)
+lapply(men8L[groups == 9], fitOne, addLine = T, col = "black")
+```
+
+```{r}
+men8LongFit <- lapply(men8L, fitOne)
+
+coeffs <- sapply(men8LongFit, "[", "ageCoeff" )
+ages <- sapply(men8LongFit, "[", "medAge")
+
+longCoeffs <- lm(coeffs ~ ages)
+
+summary(longCoeffs)
 ```
@@
 ```{r}
-
+data_paths <- sapply(c("men_txt", "women_txt"), function(p) {
+   path <- file.path(data.dir, p)
+
+   sapply(path, function( f ) { 
+      file.path(path, list.files(path) )
+   })
+})
+
+consolidated_files <- c(data_paths[, 1], data_paths[, 2])
+
+tp <- consolidated_files[1]
+
+files <- lapply(consolidated_files, function(f) read.fwf(f, widths = 120))
+
+length(files)
 ```
@@
 
-```{r}
-
-```
-
 Find the record where the time is only 1.5.
@@
 In your modification, include code to provide a warning message about the rows that are being dropped for having a time that is too small.
+
+```{r}
+
+```
 
@@
 ```{r}
+
+convertTime <- function( charTime ) {
+
+   timePieces <- strsplit(charTime, ":")
+
+   timePieces <- sapply(timePieces, as.numeric)
+
+   runTime <- sapply(timePieces,
+                  function(x) {
+                     if(length(x) == 2) x[1] + x[2]/60
+                     else 60 * x[1] + x[2] + x[3]/60
+                  })
+
+}
 
@@
 
-```
+womenDF <- mapply(createDF, womenResMat, year = years,
+                  sex = rep("F", 14), SIMPLIFY = F)
+
+sapply(womenDF, function(x) sum(is.na(x$runTime)))
+
+```
+
+### 8.)
+
+Modify the call to the _plot()_ function that created figure 2.6 to create Figure 2.7. To do this, read the documentation for plot() to determine which parameters could be helpful.
+
+```{r}
+ggplot(cbMen, aes(age, runTime)) +
+   geom_jitter(col = Purple8A) +
+   geom_smooth() +
+   xlim(10, 75) +
+   ylim(45, 150)
+```
+
+### 9.)
+
+Modify the piecewise linear fit from Section 2.4.2 to include a hing a 70.
+Examine the coefficients from the fit and compare the fitted curve to the loess curve.
+Does the additional hing improve the fit?
+
+```{r}
+age20to70 <- seq( from = 20, to = 70, by = 1)
+
+mR.lo992 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 1999,])
+mR.lo.pr992 <- predict(mR.lo99, data.frame(age = age20to70))
+
+summary(mR.lo.pr992)
+
+mR.lo122 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 2012, ])
+mR.lo.pr122 <- predict(mR.lo122, data.frame(age = age20to70))
+
+summary(mR.lo122)
+
+plot(mR.lo.pr992 ~ age20to70,
+     type = "l", col = "purple", lwd = 3,
+     xlab = "Age (years)", ylab = "Fitted Run Time (minutes)")
+
+lines(x = age20to70, y = mR.lo.pr122,
+      col = "green", lty = 2, lwd = 3)
+
+legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
+       legend =  c("1999", "2012"), bty = "n")
+```
+
+### 10.)
+
+We have seen that the 1999 runners were typically older than the 2012 runners.
+Compare the age distribution of the runners agross all 14 years of the races.
+Use quantile-quantile plots, boxplots and density curves to make your comparisions.
+
+```{r}
+ggplot(cbMen, aes(age, group = year)) +
+   geom_histogram(aes(fill = ..count..), bins = 30) +
+   facet_wrap(~year)
+
+ggplot(cbMen, aes(age, group = year)) +
+   geom_boxplot(aes(fill = as.factor(year))) +
+   coord_flip() +
+   labs(title = "Age by Race Year", xlab = "Race Year", ylab = "Age")
+
+ggplot(cbMen, aes(year, age)) +
+   geom_quantile()
+```
+
+### 11.)
+
+Normalize each male runner's time by the fastest time for the runner of the same age.
+To do this, find the fastest runnerk for each year of age from 20 to 80, tapply() function maybe helpful here.
+Smooth these times using loess() and find the smoothed fime using predict.
+Use these smoothed times to normalize each run time.
+
+```{r}
+
+cbMenSub <- as.data.table(cbMenSub)
+
+fastest_times <- cbMenSub[, .(max = max(runTime)), by = list(year, age)]
+
+normalized_times <- cbMenSub[, .(runTime), by = list(year, age)]
+normalized_times <- merge(normalized_times, fastest_times, by = c("year", "age"))
+normalized_times[, normTime := runTime/max]
+
+mR.norm <- loess(normTime ~ age, data = normalized_times)
+
+summary(mR.norm)
+
+age_range <- seq( from = 20, to = 80, by = 1 )
+
+smoothed_times <- predict(mR.norm, newdata = age_range)
+
+smooth_results <- data.table(age = age_range, time = smoothed_times)
+
+ggplot(smooth_results, aes(age, time)) +
+   geom_point(aes(col = age)) +
+   geom_smooth(alpha = .2)
+```
+
+### 12.)
+
+Clean the strings in home and menRes to remove all leading and trailing blanks and multiple contiguous blanks. Also, make all letters lower case and remove any punctuation such as '.' or ','.
+
+```{r}
+
+homeClean2 <- str_remove_all(homeClean, " ")
+homeClean2 <- str_remove_all(homeClean2, ",")
+
+head(homeClean2)
+```
+
+### 13.)
+
+In section 2.5 we created an id for a runner by pasting together name, year of birth, and state.
+Consider using the home town instead of the state.
+How many runners have competed in at least 8 races using this new id?
+
+```{r}
+cbMenSub2 <- cbMenSub
+
+cbMenSub2$ID = paste(nameClean, homeClean2, cbMenSub$yob, sep = "_")
+
+races <- tapply(cbMenSub2$year, cbMenSub2$ID, length)
+
+races8 <- names(races)[which(races >= 8)]
+
+men8 <- cbMenSub2[ cbMenSub2$ID %in% races8, ]
+
+orderByRunner <- order(men8$ID, men8$year)
+
+men8 <- men8[orderByRunner, ]
+```
+
+### 14.)
+
+Further refine the set of athletes in the longitudinal analysis by dropping those IDs who have a large jump in time in consecutive races and who did not compete for two or more years in a row.
+How many unique IDs do you have when you include these additional restrictions?
+
+```{r}
+numRaces <- tapply(cbMenSub$year, cbMenSub$ID, length)
+races8 <- names(numRaces)[which(numRaces >= 8)]
+men8 <- cbMenSub[ cbMenSub$ID %in% races8, ]
+
+by_year <- as.data.table(men8)[, .(year, runTime), by = ID]
+
+sans_outliers <- by_year %>%
+   group_by(ID) %>%
+   mutate(year_gap = year - lag(year, 1), run_diff = runTime - lag(runTime, 1)) %>%
+   filter(year_gap <= 2) %>%
+   filter(abs(run_diff) < 5) %>%
+   distinct(ID)
+
+nrow(sans_outliers)
+```

### 2020-04-27 Case Studies/03_spam_identification.rmd
finished cherry blossom
finished cherry blossom

--- 2020-04-21 Case Studies/02_race-times.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Cherry Blossom Race Times}
-   \rfoot{\color{headergrey}Chapter 2}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Spam Identification}
+   \rfoot{\color{headergrey}Chapter 3}
    \lfoot{\color{headergrey}}
@@
 ```
-
-# Cherry Blossom Ten Mile Run & Walk
-
-## Examining the Impact of Age on Physical Performance
-
-### The Raw Data
-
-data is pre-processed from the web in \02_data.r
-
-```{r}
-
-load_results <- function(class, year) {
-
-   folder <- ifelse(class == "men", "men_txt", "women_txt")
-   file <- file.path(data.dir, folder, paste(year, ".txt", sep = ""))
-
-   return(readLines(file))
-}
-
-```
-
-### Race Results Preprocessing
-
-```{r}
-
-els <- load_results("men", 2006)
-
-eqIndex <- grep("^===", els)
-eqIndex
-
-spacerRow <- els[eqIndex]
-headerRow <- els[eqIndex - 1]
-
-body <- els[ -(1:eqIndex) ]
-
-headerRow <- tolower(headerRow)
-
-timeStart <- regexpr("net", headerRow)
-time <- substr(body, start = timeStart, stop = timeStart + 1)
-
-ageStart <- regexpr("ag", headerRow)
-
-age <- substr(body, start = ageStart, stop = ageStart + 1)
-head(age)
-
-summary(as.numeric(age))
-
-blankLocs <- gregexpr(" ", spacerRow)
-blankLocs
-
-searchLocs <- c(0, blankLocs[[1]])
-
-Values <- mapply(substr, list(body),
-                 start = searchLocs[ -length(searchLocs)] + 1,
-                 stop = searchLocs[ -1 ] - 1)
-
-findColLocs <- function(spacerRow) {
-   spaceLocs <- gregexpr(" ", spacerRow)[[1]]
-   rowLength <- nchar(spacerRow)
-
-   if(substring(spacerRow, rowLength, rowLength + 1) != " ")
-      return(c(0, spaceLocs, rowLength + 1))
-   else
-      return(c(0, spaceLocs))
-}
-
-findColLocs(spacerRow)
-name <- "home"
-colnames <- c("name", "home", "ag", "gun", "net", "time")
-
-colIndex <- which(colnames == name)
-
-startPos <- regexpr(name, headerRow)[[1]]
-
-#
-# can we modify select cols to short circuit on matching w/o spaces?
-#######
-
-selectCols =
-   function(colNames, headerRow, searchLocs)
-   {
-      sapply(colNames,
-             function(name, headerRow, searchLocs)
-            {
-               startPos <- regexpr(name, headerRow)[[1]]
-
-               if(startPos == -1)
-                  return( c(NA, NA) )
-
-               index <- sum(startPos >= searchLocs)
-
-               c(searchLocs[index] + 1, searchLocs[index + 1])
-            },
-      headerRow = headerRow, searchLocs = searchLocs)
-   }
-
-searchLocs <- findColLocs(spacerRow)
-loc <- selectCols("home", headerRow, searchLocs)
-vars <- mapply(substr, list(body), start = loc[1,], stop = loc[2, ])
-
-summary(as.numeric(vars))
-
-shortColNames <- c("name", "home", "ag", "gun", "net", "time")
-
-locCols <- selectCols(shortColNames, headerRow, searchLocs)
-Values <- mapply(substr, list(body), start = locCols[1, ],
-                 stop = locCols[2, ])
-
-class(Values)
-
-colnames(Values) = shortColNames
-
-head(Values)
-tail(Values)[, 1:3]
-```
-
-```{r}
-
-extractVariables =
-   function(file, varNames = c("name", "home", "ag", "gun", "net", "time"))
-   {
-      # Find the index of the row with =s
-      eqIndex <- grep("^===", file)
-      spacerRow <- file[eqIndex]
-      headerRow <- tolower(file[ eqIndex - 1 ])
-      body <- file[ -(1 : eqIndex) ]
-
-      blank <- grep("^[[:blank:]]*$", body)
-      footnote <- grep("^[^\\s]*[\\*]*[#]", body)
-      ignore <- union(blank, footnote)
-
-      if(length(ignore))
-         body <- body[-ignore]
-
-      # Obtain the starting and ending positions of variables
-      searchLocs <- findColLocs(spacerRow)
-      locCols <- selectCols(varNames, headerRow, searchLocs)
-
-      Values <- mapply(substr, list(body), start = locCols[1, ],
-                       stop = locCols[2, ])
-      colnames(Values) <- varNames
-
-      invisible(Values)
-   }
-
-load_data <- function(class) {
-   folder <- ifelse(class == "men", "men_txt", "women_txt")
-   filenames <- file.path(data.dir, paste(folder, "/", 1999:2012, ".txt", sep = ""))
-   files <- lapply(filenames, readLines)
-   names(files) <- 1999:2012
-
-   mat <- lapply(files, extractVariables)
-
-   return(mat)
-}
-
-```
-
-### Data Cleaning and Reformatting Variables
-
-```{r}
-menResMat <- load_data("men")
-
-length(menResMat)
-sapply(menResMat, nrow)
-
-womenResMat <- load_data("women")
-
-length(womenResMat)
-sapply(womenResMat, nrow)
-```
-
-Age Validation
-
-```{r}
-age <- as.numeric(menResMat[['2012']][, 'ag'])
-
-tail(age)
-
-age <- sapply(menResMat,
-              function(x) as.numeric(x[, "ag"]))
-
-age_values <- plyr::ldply(age, data.frame)
-colnames(age_values) <- c("Year", "Age")
-
-ggplot(age_values, aes(Age)) +
-   geom_histogram(aes(fill = ..count..), bins = 30)
-
-ggplot(age_values, aes(Year, Age)) +
-   geom_boxplot(aes(fill = Year)) +
-   labs(title = "Age By Race Year")
-```
-
-```{r}
-sapply(age, function(x) sum(is.na(x)))
-
-file2001 <- read_results("men", 2001)
-age2001 <- as.numeric(extractVariables(file2001)[, "ag"])
-
-badAgeIndex <- which(is.na(age2001)) + 5
-
-file2001[badAgeIndex]
-
-badAgeIndex
-
-blanks <- grep("^[[:blank:]]*$", file2001)
-blanks
-
-which(age2001 < 5)
-
-file2001[which(age2001 < 5)]
-
-file <- read_results("men", 2001)
-data <- extractVariables(file)
-age <- as.numeric(data[, "ag"])
-
-sum(age < 5)
-
-which(age == 0)
-
-data[which(age == 0)]
-
-```
-
-### Time Validation
-
-```{r}
-file2002 <- read_results("men", 2012)
-
-charTime <- menResMat[["2012"]][, "time"]
-
-head(charTime)
-tail(charTime)
-
-timePieces <- strsplit(charTime, ":")
-
-timePieces[[1]]
-tail(timePieces, 1)
-
-timePieces <- sapply(timePieces, as.numeric)
-
-runTime <- sapply(timePieces,
-                  function(x) {
-                     if(length(x) == 2) x[1] + x[2]/60
-                     else 60 * x[1] + x[2] + x[3]/60
-                  })
-
-summary(runTime)
-
-convertTime <- function( charTime ) {
-
-   timePieces <- strsplit(charTime, ":")
-
-   timePieces <- sapply(timePieces, as.numeric)
-
-   runTime <- sapply(timePieces,
-                  function(x) {
-                     if(length(x) == 2) x[1] + x[2]/60
-                     else 60 * x[1] + x[2] + x[3]/60
-                  })
-
-}
-
-```
-
-### Aggregate cleaning into
-
-```{r}
-
-createDF =
-   function(Res, year, sex) {
-      useTime <- if(!is.na(Res[1, 'net']) )
-                     Res[, 'net']
-                  else if( !is.na(Res[1, 'gun']) )
-                     Res[ , 'gun']
-                  else
-                     Res[, 'time']
-
-      useTime <- gsub("[#\\*[:blank:]]", "", useTime)
-
-      Res <- Res[ useTime != "", ]
-
-      runTime <- convertTime(useTime[ useTime != "" ])
-
-      N <- nrow(Res)
-
-      Results <- data.frame( year = rep(year, N),
-                             sex = rep(sex, N),
-                             name = Res[, 'name'],
-                             home = Res[, 'home'],
-                             age = as.numeric(Res[, 'ag']),
-                             runTime = runTime,
-                             stringsAsFactors = F)
-
-      invisible(Results)
-   }
-
-years <- 1999:2012
-menDF <- mapply(createDF, menResMat, year = years,
-                sex = rep("M", 14), SIMPLIFY = F)
-
-warnings()[ c(1:2, 49:50)]
-
-sapply(menDF, function(x) sum(is.na(x$runTime)))
-
-file2006 <- read_results("men", 2006)
-parsed2006 <- extractVariables(file2006)
-time2006 <- parsed2006[, "net"]
-
-womenDF <- mapply(createDF, womenResMat, year = years,
-                  sex = rep("F", 14), SIMPLIFY = F)
-
-sapply(womenDF, function(x) sum(is.na(x$runTime)))
-```
-
-### Consolidate Results
-
-```{r}
-cbMen <- do.call(rbind, menDF)
-save(cbMen, file = file.path(data.dir, "cbMen.rda"))
-
-cbWomen <- do.call(rbind, womenDF)
-
-save(cbWomen, file = file.path(data.dir, "cbWomen.rda"))
-```
-
-### Exploratory Data Analysis
-
-```{r}
-
-ggplot(cbMen, aes(age, runTime)) +
-   geom_point(aes(col = age)) +
-   labs(main = "Run Times by Age", xlab = "Age (years)", xlab = "Run Time (minutes")
-
-```
-
-```{r}
-display.brewer.all()
-```
-
-```{r}
-Purple8 <- brewer.pal(9, "Purples")[8]
-
-Purple8A <- paste(Purple8, "14", sep = "")
-
-ggplot(cbMen, aes(age, runTime)) +
-   geom_jitter(col = Purple8A) +
-   geom_smooth() +
-   xlim(10, 75) +
-   ylim(45, 150)
-
-```
-
-```{r}
-smoothScatter(y = cbMen$runTime, x = cbMen$age,
-              ylim = c(40, 165), xlim = c(15, 85),
-              xlab = "Age (years)", ylab = "Run Time (minutes)")
-```
-
-```{r}
-cbMenSub <- cbMen[cbMen$runTime > 30 &
-                     !is.na(cbMen$age) & cbMen$age > 15,]
-
-ageCat <- cut(cbMenSub$age, breaks = c(seq(15, 75, 10), 90))
-
-ggplot(data.table(age = ageCat, time = cbMenSub$runTime), aes(ageCat, time)) +
-   geom_boxplot(aes(fill = age)) +
-   labs("Run Times by Age Category")
-```
-
-```{r}
-lm_age <- lm(runTime ~ age, data = cbMenSub)
-
-summary(lm_age)
-```
-
-```{r}
-smoothScatter(x = cbMenSub$age, y = lm_age$residuals,
-              xlab = "Age (years)", ylab = "Residuals") 
-abline(h = 0, col = "purple", lwd = 3)
-
-resid.lo <- loess(resids ~ age,
-                  data = data.frame(resids = residuals(lm_age),
-                                    age = cbMenSub$age))
-
-age20to80 <- 20:80
-
-resid.lo.pr <- predict(resid.lo, newdata = data.frame(age = age20to80))
-
-lines(x = age20to80, y = resid.lo.pr,
-      col = "green", lwd = 3, lty = 2)
-```
-
-```{r}
-menRes.lo <- loess(runTime ~ age, cbMenSub)
-
-summary(menRes.lo)
-
-menRes.lo.pr <- predict(menRes.lo, data.frame(age = age20to80))
-
-```
-
-```{r}
-over50 <- pmax(0, cbMenSub$age - 50)
-
-lmOver50 <- lm(runTime ~ age + over50, data = cbMenSub)
-
-summary(lmOver50)
-
-decades <- seq(30, 60, by = 10)
-overAge <- lapply(decades,
-                  function(x) pmax(0, (cbMenSub$age - x)))
-names(overAge) <- paste("over", decades, sep = "")
-overAge <- as.data.frame(overAge)
-
-tail(overAge)
-
-lmPiecewise <- lm(runTime ~., 
-                  data = cbind(cbMenSub[, c("runTime", "age")],
-                               overAge))
-
-summary(lmPiecewise)
-
-overAge20 <- lapply(decades, function(x) pmax(0, (age20to80 - x)))
-names(overAge20) <- paste("over", decades, sep = "")
-overAgeDF <- cbind(age = data.frame(age = age20to80), overAge20)
-
-tail(overAgeDF)
-```
-
-```{r}
-predPiecewise <- predict(lmPiecewise, overAgeDF)
-
-plot(predPiecewise ~ age20to80,
-     type = "l", col = "purple", lwd = 3,
-     xlab = "Age (years)", ylab = "Run Time Prediction")
-
-lines(x = age20to80, y = menRes.lo.pr,
-      col = "green", lty = 2, lwd = 3)
-
-legend("topleft", col = c("purple", "green"),
-       lty = c(1, 2), lwd = 3,
-       legend = c("Piecewise Linear", "Loess Curve"), bty = "n")
-```
-
-### Cross-Sectional Data and Covariates
-
-```{r}
-age1999 <- cbMenSub[ cbMenSub$year == 1999, "age"]
-age2012 <- cbMenSub[ cbMenSub$year == 2012, "age"]
-```
-
-```{r}
-summary(age1999)
-
-summary(age2012)
-```
-
-```{r}
-plot(density(age1999, na.rm = T),
-     ylim = c(0, 0.05), col = "purple",
-     lwd = 3, xlab = "Age (years)", main = "")
-lines(density(age2012, na.rm = T),
-      lwd = 3, lty = 2, col = "green")
-legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
-       legend = c("1999", "2012", bty = "n"))
-```
-
-```{r}
-by_year <- data.table(cbMenSub)[, .(age, mean = mean(age)), by = list(year)]
-
-ggplot(by_year[year %in% c("1999", "2012")], aes(age, group = year)) +
-   geom_density(aes(fill = year), alpha = .6) +
-   geom_vline(aes(xintercept = mean, col = year, group = year)) +
-   labs(title = "Runners by Year")
-```
-
-```{r}
-ggplot(by_year, aes(year, age)) +
-   geom_point(aes(col = age)) +
-   geom_quantile() +
-   labs(title = "Runners Age Over Time")
-```
-
-```{r}
-mR.lo99 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 1999,])
-mR.lo.pr99 <- predict(mR.lo99, data.frame(age = age20to80))
-
-summary(mR.lo12)
-
-mR.lo12 <- loess(runTime ~ age, cbMenSub[ cbMenSub$year == 2012, ])
-mR.lo.pr12 <- predict(mR.lo12, data.frame(age = age20to80))
-
-plot(mR.lo.pr99 ~ age20to80,
-     type = "l", col = "purple", lwd = 3,
-     xlab = "Age (years)", ylab = "Fitted Run Time (minutes)")
-
-lines(x = age20to80, y = mR.lo.pr12,
-      col = "green", lty = 2, lwd = 3)
-
-legend("topleft", col = c("purple", "green"), lty = 1:2, lwd = 3,
-       legend =  c("1999", "2012"), bty = "n")
-```
-
-```{r}
-mR.models <- cbMenSub %>% 
-   group_by(year) %>%
-   group_map(~ loess(runTime ~ age, .x))
-
-mR.pred <- lapply(mR.models, 
-                  function(m, year, age) cbind(rep(year, length(age)), age = age, predict(m, data.frame(age = age))),
-                  year = 1999:2012,
-                  age = age20to80)
-
-yoy_changes <- matrix(unlist(mR.pred), ncol = 3)
-yoy_changes <- as.data.frame(yoy_changes)
-colnames(yoy_changes) <- c("year", "pred", "age")
-yoy_changes$year <- as.numeric(yoy_changes$year)
-
-ggplot(yoy_changes[year == 1999], aes(age, pred)) +
-   geom_point()
-```
-
-### Further Analysis
-
-### 1.)
-
-Write a function that uses _read.fwf()_ to read the 28 text tables in MenTxt/ and WomenTxt/ into R.
-These are called 1999.text, 2000txt, etc.
-and are described in greater detail in 2.2.
-Examine the tables in a plain text editor to determine the start and end position of each column of interest (name, hometown, age and gun and net time).
-
-Use statistics to explore the results and confirm that you have extracted the information from the correct positions in the text.
-
-```{r}
-
-```
-
-### 2.)
-
-Revise the extractVariables function (see section 2.2) to remove the rows in _menTables_ that are blank.
-In addition, eliminatee the rows that begin with a '*' or a '#'.
-You may find the following regular expression helpful for locating blank rows in a table.
-
-_grep("^[[:blank:]]*$", body)_
-
-The pattern uses several meta characters.
-The ^ is an anchor for the start of the string, the $ anchors to the end of the string, the [[:blank:]] denotes the equivalence class of any space or tab character, and * indicates that the blank character can appear 0 or more times.
-All together the pattern ^[[:blank:]]*\$ matches a string that contains any number of blanks from start to end.
-
-```{r}
-
-extractVariables =
-   function(file, varNames = c("name", "home", "ag", "gun", "net", "time"))
-   {
-      # Find the index of the row with =s
-      eqIndex <- grep("^===", file)
-      spacerRow <- file[eqIndex]
-      headerRow <- tolower(file[ eqIndex - 1 ])
-      body <- file[ -(1 : eqIndex) ]
-
-      blank <- grep("^[[:blank:]]*$", body)
-      footnote <- grep("^[^\\s]*[\\*]*[#]", body)
-      ignore <- union(blank, footnote)
-
-      if(length(ignore))
-         body <- body[-ignore]
-
-      # Obtain the starting and ending positions of variables
-      searchLocs <- findColLocs(spacerRow)
-      locCols <- selectCols(varNames, headerRow, searchLocs)
-
-      Values <- mapply(substr, list(body), start = locCols[1, ],
-                       stop = locCols[2, ])
-      colnames(Values) <- varNames
-
-      invisible(Values)
-   }
-
-```
-
-### 3.)
-
-```{r}
-
-```
-
-Find the record where the time is only 1.5.
-What happened?
-Determine how to handle the problem and which function needs to be modified: _extractResTable()_, _extractVariables()_, or _cleanUp()_.
-In your modification, include code to provide a warning message about the rows that are being dropped for having a time that is too small.
-
-### 4.)
-
-Examine the head and tail of the 2006 men's file.
-Look at both the character matrix in the list called _menResMat_ and the character vector in the list called _menFiles_ (see Sec 2.2).
-(Recall that the desired character matrix in _menResMat_ and the character vector in _menFiels_ both correspond to the element named "2006").
-What is wrong with the hometown?
-Examine the header closely to figure out how this error came about.
-Modify the _extractVariables()_ function to fix the problem.
-
-```{r}
-
-```
-
-### 5.)
-
-Write the _convertTime()_ function described in Section 2.3.
-This function takes a string where time is in either the format hh:mm:ss or mm:ss.
-The return value is the time as numeric value of the number of minutes.
-Design this function to take a character vector with multiple strings and return a numeric vector.
-
-```{r}
-
-```
-
-### 6.)
-
-Modify the _createDF()_ function in Section 2.3 to handle the formatting problem with the 2006 male file. You will need to carefully inspect the raw text file in order to determien the problem.
-
-```{r}
-
-createDF =
-   function(Res, year, sex) {
-      useTime <- if(!is.na(Res[1, 'net']) )
-                     Res[, 'net']
-                  else if( !is.na(Res[1, 'gun']) )
-                     Res[ , 'gun']
-                  else
-                     Res[, 'time']
-
-      useTime <- gsub("[#\\*[:blank:]]", "", useTime)
-
-      Res <- Res[ useTime != "", ]
-
-      runTime <- convertTime(useTime[ useTime != "" ])
-
-      N <- nrow(Res)
-
-      Results <- data.frame( year = rep(year, N),
-                             sex = rep(sex, N),
-                             name = Res[, 'name'],
-                             home = Res[, 'home'],
-                             age = as.numeric(Res[, 'ag']),
-                             runTime = runTime,
-                             stringsAsFactors = F)
-
-      invisible(Results)
-   }
-
-```
-
-### 7.)
-
-Follow the approach developed in Section 2.2 to read the files for the female runners and then process them using the functions in Section 2.3 to create a data frame for analysis.
-You may need to generalize the _createDF()_ and _extractVariables()_ functions to handle additional oddities in the raw text files.
-
-```{r}
-
-```

### 2020-04-30 Case Studies/03_spam_identification.rmd
spam identification case study
spam identification case study

@@
 library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
+library(tm, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-data.dir <- file.path(here::here(), "Case Studies", "datasets")
+data.dir <- file.path(here::here(), "Case Studies", "datasets", "spam")
 
@@
 ```
+
+# Using Statistics to Identify Spam
+
+## Anatomy of an email Message
+
+### Spam Data
+
+```{r}
+head(list.files(path = file.path(data.dir, "easy_ham")))
+head(list.files(path = file.path(data.dir, "spam_2")))
+
+directories <- paste(data.dir, list.files(data.dir), sep = .Platform$file.sep)
+
+file_counts <- sapply(directories, function(dir) length(list.files(dir)))
+
+total_files <- sum(file_counts)
+total_files
+```
+
+```{r}
+file_counts
+
+idx <- c(1:5, 15, 27, 68, 69, 329, 404, 427, 516, 852, 971)
+
+fn <- list.files(directories[1], full.names = T)[idx]
+
+sampleEmail <- sapply(fn, readLines)
+```
+
+### Text Mining and Naive Bayes Classification
+
+```{r}
+msg <- sampleEmail[[1]]
+which(msg == "")[1]
+
+match("", msg)
+
+splitPoint <- match("", msg)
+
+msg[ (splitPoint - 2):(splitPoint + 6)]
+
+header <- msg[1:(splitPoint - 1)]
+body <- msg[ -(1:splitPoint) ]
+
+```
+
+```{r}
+
+splitMessage <- function(msg) {
+   splitPoint <- match("", msg)
+
+   header <- msg[ 1:(splitPoint - 1)]
+   body <- msg[ -(1:splitPoint)]
+
+   return(list(header = header, body = body))
+}
+
+sampleSplit <- lapply(sampleEmail, splitMessage)
+
+header <- sampleSplit[[1]]$header
+grep("Content-Type", header)
+grep("multi", tolower(header))
+
+header[46]
+
+headerList <- lapply(sampleSplit, function(msg) msg$header)
+
+CTloc <- sapply(headerList, grep, pattern = "Content-Type")
+CTloc
+
+sapply(headerList, function(header) {
+   CTloc <- grep("Content-Type", header)
+   if( length(CTloc) == 0) return(NA)
+   CTloc
+})
+
+hasAttach <- sapply(headerList, function(header) {
+   CTloc <- grep("Content-Type", header)
+
+   if(length(CTloc) == 0) return(F)
+   grepl("multi", tolower(header[CTloc]))
+})
+
+hasAttach
+
+header <- sampleSplit[[6]]$header
+boundaryIdx <- grep("boundary=", header)
+header[boundaryIdx]
+
+sub(".*boundary=\"(.*)\";.*", "\\1", header[boundaryIdx])
+
+header2 <- headerList[[9]]
+boundaryIdx2 <- grep("boundary=", header2)
+header2[boundaryIdx2]
+
+sub('.*boundary="(.*)";.*', "\\1", header2[boundaryIdx2])
+
+boundary2 <- gsub('"', "", header2[boundaryIdx2])
+
+sub(".*boundary= *(.*);?.*", "\\1", boundary2)
+
+boundary <- gsub('"', "", header[boundaryIdx])
+sub(".*boundary= *(.*);?.*", "\\1", boundary)
+
+```
+
+```{r}
+getBoundary <- function(header) {
+
+   boundaryIdx <- grep("boundary=", header)
+   boundary = gsub('"', "", header[boundaryIdx])
+   gsub(".*boundary= *([^;]*);?.*", "\\1", boundary)
+}
+
+```
+
+```{r}
+sampleSplit[[6]]
+
+boundary <- getBoundary(headerList[[15]])
+body <- sampleSplit[[15]]$body
+
+bString <- paste("--", boundary, sep = "")
+bStringLocs <- which(bString == body)
+bStringLocs
+
+eString <- paste("--", boundary, "--", sep = "")
+eStringLoc <- which(eString == body)
+eStringLoc
+
+msg <- body[ (bStringLocs[1] + 1) : (bStringLocs[2] - 1)]
+tail(msg)
+
+msg <- c(msg, body[ (eStringLoc + 1) : length(body) ])
+tail(msg)
+```
+
+### Handle Attachments
+
+```{r}
+
+dropAttach <- function(email) {
+
+}
+
+```
+
+### Extracting Words from the Message Body
+
+```{r}
+head(sampleSplit[[1]]$body)
+
+msg <- sampleSplit[[3]]$body
+head(msg)
+
+```
+
+### Stemming
+
+```{r}
+stopWords <- stopwords(kind = "en")
+```
+
+### Convert To Wordlist
+
+```{r}
+tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", msg))
+
+msg[ c(1, 3, 26, 27) ]
+
+cleanMsg <- tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", msg))
+cleanMsg[ c(1, 3, 26, 27) ]
+
+words <- unlist(strsplit(cleanMsg, "[[:blank:]]+"))
+
+words <- words[ nchar(words) > 1 ]
+
+words <- words[ ! (words %in% stopWords) ]
+
+head(words)
+
+```
+
+```{r}
+
+findMsgWords <- function(msg, stopwords = stopWords) {
+
+   cleanMsg <- tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", msg))
+
+   words <- unlist(strsplit(cleanMsg, "[[:blank:]]+"))
+
+   words <- words[ ! (words %in% stopwords) ]
+
+   return(words)   
+}
+
+findMsgWords(msg)
+```
+
+### Prep Wrap-Up
+
+```{r}
+processAllWords <- function(dirName, stopWords) {
+   # read all files in the directory
+   fileNames <- list.files(dirName, full.names = T)
+
+   # drop files that are not email, i.e., cmds
+   notEmail <- grep("cmds$", fileNames)
+
+   if( length(notEmail) > 0) fileNames <- fileNames[ -notEmail ]
+
+   messages <- lapply(fileNames, readLines, encoding = "latin1")
+
+   # split header and body
+   emailSplit <- lapply(messages, splitMessage)
+
+   # put body and header in own lists
+   bodyList <- lapply(emailSplit, function(msg) msg$body)
+   headerList <- lapply(emailSplit, function(msg) msg$header)
+   rm(emailSplit)
+
+   # determine which messages have attachments
+   hasAttach <- sapply(headerList, function(header) {
+
+      CTloc <- grep("Content-Type", header)
+
+      if( length(CTloc) == 0) return(0)
+
+      multi <- grep("multi", tolower(header[CTloc]))
+
+      if( length(multi) == 0 ) return(0)
+
+      multi
+   })
+
+   hasAttach <- which(hasAttach > 0)
+
+   # drop attachments from message body
+   bodyList[hasAttach] <- mapply(dropAttach, bodyList[hasAttach],
+                                 boundaries, SIMPLIFY = F)
+
+   # extract words from body
+   msgWordsList <- lapply(bodyList, findMsgWords, stopWords)
+
+   invisible(msgWordsList)
+}
+```
+
+### Build Email Database
+
+```{r}
+
+msgWordList <- lapply(directories, processAllWords)
+
+```

### 2020-05-03 Case Studies/03_spam_identification.rmd
spam identification using statistics
spam identification using statistics

@@
 ```{r}
+
 head(list.files(path = file.path(data.dir, "easy_ham")))
@@
 ```{r}
+
 getBoundary <- function(header) {
-
    boundaryIdx <- grep("boundary=", header)
@@
 ```{r}
-sampleSplit[[6]]
 
@@
 tail(msg)
+
 ```
@@
 ```{r}
-
-dropAttach <- function(email) {
-
-}
 
@@
 ```{r}
-stopWords <- stopwords(kind = "en")
+exclude_word_list <- stopwords(kind = "en")
 ```
@@
 
-words <- words[ ! (words %in% stopWords) ]
+words <- words[ ! (words %in% exclude_word_list) ]
 
@@
 
-findMsgWords <- function(msg, stopwords = stopWords) {
+findMsgWords <- function(msg, exclude) {
 
@@
 
-   words <- words[ ! (words %in% stopwords) ]
+   keep <- sapply(words, function(word) return(!(word %in% exclude)))
 
-   return(words)   
+   return(words[ keep ])   
 }
 
-findMsgWords(msg)
 ```
@@
 ```{r}
+
+dropAttach <- function(body, boundary) {
+
+   if(is.null(body)) {
+      return("")
+   }
+
+   bString <- paste("--", boundary, sep = "")
+   bStringLocs <- which(bString == body)
+
+   eString <- paste("--", boundary, "--", sep = "")
+   eStringLoc <- which(eString == body)
+
+   if(length(bStringLocs) == 2) {
+      msg <- body[ (bStringLocs[1] + 1) : (bStringLocs[2] - 1)]
+   }
+
+   if(length(eStringLoc) > 0) {
+      msg <- c(msg, body[ (eStringLoc + 1) : length(body) ])
+   }
+
+   return(msg)
+}
+
 processAllWords <- function(dirName, stopWords) {
@@
 
+   # find boundary string for messages with attachments
+   boundaries <- sapply(headerList[hasAttach], getBoundary)
+
    # drop attachments from message body
@@
 
-msgWordList <- lapply(directories, processAllWords)
+msgWordList <- lapply(directories, processAllWords, stopWords = exclude_word_list)
+
+numMsgs <- sapply(msgWordList, length)
+numMsgs
+
+isSpam <- rep(c(FALSE, FALSE, FALSE, TRUE, TRUE), numMsgs)
+
+msgWordsList <- unlist(msgWordList, recursive = F)
 
 ```
+
+## Naive Bayes Classifier Implementation
+
+### Train / Test Split
+
+```{r}
+numEmail <- length(isSpam)
+
+numSpam <- sum(isSpam)
+numHam <- numEmail - numSpam
+
+set.seed(418910)
+
+testSpamIdx <- sample(numSpam, size = floor(numSpam/3))
+testHamIdx <- sample(numHam, size = floor(numHam/3))
+
+testMsgWords <- c((msgWordsList[isSpam])[testSpamIdx],
+                  (msgWordsList[!isSpam])[testHamIdx])
+
+trainMsgWords <- c((msgWordsList[isSpam])[ - testSpamIdx ],
+                   (msgWordsList[!isSpam])[ - testHamIdx])
+
+testIsSpam <- rep(c(T, F), 
+                  c(length(testSpamIdx), length(testHamIdx)))
+
+trainIsSpam <- rep(c(T, F),
+                   c(numSpam - length(testSpamIdx),
+                     numHam - length(testHamIdx)))
+
+```
+
+### Probability Estimates from Training Sample
+
+```{r}
+bow <- unique(unlist(trainMsgWords))
+
+length(bow)
+```
+
+```{r}
+spamWordCounts <- rep(0, length(bow))
+
+names(spamWordCounts) = bow
+
+tmp <- lapply(trainMsgWords[trainIsSpam], unique)
+tt <- table( unlist(tmp) )
+spamWordCounts[ names(tt) ] = tt
+
+spamWordsProbs <- (spamWordCounts + 0.5) / (sum(trainIsSpam) + 0.5)
+
+spamWordsProbs[1:20]
+
+hamWordCounts <- rep(0, length(bow))
+
+names(hamWordCounts) = bow
+
+tmp <- lapply(trainMsgWords[ - trainIsSpam], unique)
+tt <- table( unlist(tmp) )
+hamWordCounts[ names(tt) ] = tt
+
+hamWordsProbs <- (hamWordCounts + 0.5) / (sum(!trainIsSpam) + 0.5)
+
+log(spamWordsProbs) - log(hamWordsProbs)
+
+```
+
+```{r}
+wordsList <- trainMsgWords
+spam <- trainIsSpam
+
+make_words_valid_columns <- function( words, all_words ) {
+
+   word_counts <- rep(0, length(all_words))
+   names(word_counts) <- all_words
+
+   tmp <- lapply(words, unique)
+   tt <- table( unlist(tmp) )
+   word_counts[ names(tt) ] = tt
+
+   return(word_counts)
+}
+
+computeFreqs <- function(wordsList, spam, bow = unique(unlist(wordsList))) {
+
+   all_words <- unique(bow)
+
+   # create a matrix for spam, ham, and log odds
+   wordTable <- matrix(0.5, nrow = 2, ncol = length(bow))
+   colnames(wordTable) <- all_words
+   rownames(wordTable) <- c( "presentLogOdds",
+                          "absentLogOdds")
+
+   # for each spam message, add 1 to the counts for words in messsage
+
+   spam_all <- wordsList[spam]
+   spam_words <- make_words_valid_columns( spam_all, all_words )
+
+   wordTable <- rbind(wordTable, spam_words + 0.5)
+   rownames(wordTable)[3] <- "spam"
+
+   # Similarly for ham messages
+
+   ham_all <- wordsList[ !spam ]
+
+   ham_words <- make_words_valid_columns( ham_all, all_words )
+
+   wordTable <- rbind(wordTable, ham_words + 0.5)
+   rownames(wordTable)[4] <- "ham"
+
+   head(wordTable[, 1:20])
+
+   # find the total number of spam and ham
+   numSpam <- sum(spam)
+   numHam <- length(spam) - numSpam
+
+   # prob (word|spam) and prob(words|ham)
+   wordTable["spam", ] <- wordTable["spam", ] / (numSpam + 0.5)
+   wordTable["ham", ] <- wordTable["ham", ] / (numHam + 0.5)
+
+   head(wordTable[, 1:20])
+
+   # log odds
+   wordTable["presentLogOdds", ] = 
+      log(wordTable["spam", ]) - log(wordTable["ham", ])
+
+   wordTable["absentLogOdds", ] =
+      log((1 - wordTable["spam", ])) - log((1 - wordTable["ham", ]))
+
+   invisible(wordTable)
+}
+
+```
+
+```{r}
+trainTable <- computeFreqs(trainMsgWords, trainIsSpam)
+
+# peek the prob table
+head(trainTable[, 1:10])
+```
+
+### Classifying New Messages
+
+```{r}
+newMsg <- testMsgWords[[1]]
+
+# only look at words we have classified
+newMsg <- newMsg[ !is.na(match(newMsg, colnames(trainTable)))]
+
+present <- colnames(trainTable) %in% newMsg
+
+sum( trainTable["presentLogOdds", present]) +
+   sum( trainTable["absentLogOdds", !present])
+
+newMsg <- testMsgWords[[ which(!testIsSpam)[ 1 ] ]]
+newMsg <- newMsg[ !is.na(match(newMsg, colnames(trainTable)))]
+present <- (colnames(trainTable) %in% newMsg)
+
+sum(trainTable["presentLogOdds", present]) +
+   sum(trainTable["absentLogOdds", !present])
+
+```

### 2020-05-04 Case Studies/03_spam_identification.rmd
spam identification case study
spam identification case study

@@
 ```
+
+```{r}
+
+computeMsgLLR <- function(words, freqTable) {
+
+   # discard words not in training data
+   words <- words[!is.na(match(words, colnames(freqTable)))]
+
+   # Find which words are present
+   present <- colnames(freqTable) %in% words
+
+   sum(freqTable["presentLogOdds", present]) +
+      sum(freqTable["absentLogOdds", !present])
+}
+
+```
+
+```{r}
+testLLR <- sapply(testMsgWords, computeMsgLLR, trainTable)
+```
+
+```{r}
+tapply(testLLR, testIsSpam, summary)
+```
+
+```{r}
+results_df <- data.table( score = testLLR, class = testIsSpam )
+
+ggplot(results_df, aes(score, class, fill = class)) +
+   geom_boxplot() +
+   coord_flip()   
+```
+
+```{r}
+typeIErrorRate <- function(tau, llrVals, spam) {
+   classify <- llrVals > tau
+   sum(classify & !spam) / sum(!spam)
+}
+
+typeIErrorRate(0, testLLR, testIsSpam)
+typeIErrorRate(-20, testLLR, testIsSpam)
+
+error_rates <- sapply(seq(-30, 30, 1), function(cutoff) c(cutoff = cutoff, rate = typeIErrorRate(cutoff, testLLR, testIsSpam)))
+
+er_df <- data.table(t(error_rates))
+
+ggplot(er_df, aes(cutoff, rate)) +
+   geom_line(col = "darkblue") +
+   labs(title = "False Positive Error Rates")
+```
+
+```{r}
+typeIErrorRates <- function(llrVals, isSpam) {
+   o <- order(llrVals)
+   llrVals <- llrVals[o]
+   isSpam <- isSpam[o]
+
+   idx <- which(!isSpam)
+   N <- length(idx)
+   list(error = (N:1)/N, values = llrVals[idx])
+}
+
+```
+
+### Computational Considerations
+
+```{r}
+smallNums <- rep((1/2)^40, 2000000)
+largeNum <- 10000
+
+print(sum(smallNums), digits = 20)
+
+print(largeNum + sum(smallNums), digits = 20)
+
+for(i in 1:length(smallNums)) {
+   largeNum <- largeNum + smallNums[i]
+}
+
+print(largeNum, digits = 20)
+```
+
+## Recursive Partitioning and Classification Trees
+
+### Revised E-mail Data Structure
+
+```{r}
+header <- sampleSplit[[1]]$header
+
+header[1:12]
+
+header[1] = sub("^From", "Top-From:", header[1])
+
+headerPieces <- read.dcf(textConnection(header), all = T)
+
+headerPieces[, "Delivered-To"]
+
+headerVec <- unlist(headerPieces)
+dupKeys <- sapply(headerPieces, function(x) length(unlist(x)))
+names(headerVec) <- rep(colnames(headerPieces), dupKeys)
+
+headerVec[ which(names(headerVec) == "Delivered-To") ]
+
+length(headerVec)
+
+length(unique(names(headerVec)))
+```
+
+```{r}
+processHeader <- function(header) {
+   # modify the first line to create a key:value pair
+   header[1] <- sub("^From", "Top-From:", header[1])
+
+   headerMat <- read.dcf(textConnection(header), all = T)
+   headerVec <- unlist(headerMat)
+
+   dupKeys <- sapply(headerMat, function(x) length(unlist(x)))
+   names(headerVec) <- rep(colnames(headerMat), dupKeys)
+
+   return(headerVec)
+}
+
+headerList <- lapply(sampleSplit,
+                     function(msg) {
+                        processHeader(msg$header)
+                     })
+
+contentTypes <- sapply(headerList, function(header)
+   header["Content-Type"])
+
+names(contentTypes) <- NULL
+
+contentTypes
+```
+
+#### Attachments Revisited
+
+```{r}
+hasAttach <- grep("^ *multi", tolower(contentTypes))
+hasAttach
+
+boundaries <- getBoundary(contentTypes[ hasAttach ])
+boundaries
+
+boundary <- boundaries[9]
+body <- sampleSplit[[15]]$body
+
+bString <- paste("--", boundary, sep = "")
+bStringLocs <- which(bString == body)
+bStringLocs
+
+eString <- paste("--", boundary, "--", sep = "")
+eStringLoc <- which(eString == body)
+eStringLoc
+
+range <- diff(c(bStringLocs[-1], eStringLoc))
+
+body[1:range]
+
+```
+
+```{r}
+
+processAttach <- function(body, contentType ) {
+
+   boundary <- getBoundary(contentType)
+
+   bString <- paste("--", boundary, sep = "")
+   bStringLocs <- which(bString == body)
+
+   eString <- paste("--", boundary, "--", sep = "")
+   eStringLoc <- which(eString == body)
+
+   n <- length(body)
+
+   if(length(bStringLocs) == 2) {
+
+      bodyContent <- body[(bStringLocs[1] + 2):(bStringLocs[2] - 1)]
+
+      emptyLines <- which(bodyContent == "")
+      bodyContent <- bodyContent[ - emptyLines]
+
+      attachContent <- body[(bStringLocs[2] + 1):n]
+
+      aLen <- diff(c(bStringLocs[-1], eStringLoc))
+      aType <- body[bStringLocs[-1] + 1]
+
+      if(length(aLen) == length(aType)) {      
+         attachments <- data.frame(aLen = aLen, aType = aType)
+      } else {
+         attachments <- data.frame(aLen = c(), aType = c())
+      }
+
+   } else {
+      if( length(bStringLocs) == 0 ) {
+         bodyContent <- body
+      } else {
+         bodyContent = body
+      }
+      attachments <- data.frame(aLen = c(), aType = c())
+   }
+
+   return(list(body = bodyContent, attachDF = attachments ))
+}
+
+```
+
+#### More E-Mails
+
+```{r}
+
+bodyList <- lapply(sampleSplit, function(msg) msg$body)
+attList <- mapply(processAttach, bodyList[hasAttach],
+                  contentTypes[hasAttach], SIMPLIFY = F)
+
+lens <- sapply(attList, function(processedA)
+                           processedA$attachDF$aLen)
+```
+
+```{r}
+readEmail <- function(dirName) {
+   # retrieve the names of files in the directory
+   fileNames <- list.files(dirName, full.names = T)
+
+   # drop files that are not email
+   notEmail <- grep("cmds$", fileNames)
+
+   if( length(notEmail) > 0 ) fileNames = fileNames[ - notEmail ]
+
+   # read all files in the directory
+   lapply(fileNames, readLines, encoding = "latin1")
+}
+
+processAllEmail <- function(dirName, isSpam = F) {
+
+   # read all files in the directory
+   messages <- readEmail(dirName)
+
+   fileNames <- names(messages)
+   n <- length(messages)
+
+   # split header from body
+   eSplit <- lapply(messages, splitMessage)
+   rm(messages)
+
+   # process header as named character vector
+   headerList <- lapply(eSplit, function(msg)
+                           processHeader(msg$header))
+
+   # extractd content-type key                        
+   contentTypes <- sapply(headerList, function(header)
+                                       header["Content-Type"])
+
+   # extract the body
+   bodyList <- lapply(eSplit, function(msg) msg$body)
+   rm(eSplit)
+
+   # which email have attachements
+   hasAttach <- grep("^ *multi", tolower(contentTypes))
+
+   # get summary stats for attachments and the shorter body
+   attList <- mapply(processAttach, bodyList[hasAttach],
+                     contentTypes[hasAttach], SIMPLIFY = F)
+
+   bodyList[hasAttach] <- lapply(attList, function(attEl)
+                                             attEl$body)
+
+   attachInfo <- vector("list", length = n)
+   attachInfo[ hasAttach ] <- lapply(attList,
+                                     function(attEl) attEl$attachDf)
+
+   # prepare return structure
+   emailList <- mapply(function(header, body, attach, isSpam) {
+      list(isSpam = isSpam, header = header,
+           body = body, attach = attach)
+   },
+   headerList, bodyList, attachInfo,
+   rep(isSpam, n), SIMPLIFY = F)
+
+   names(emailList) <- fileNames
+
+   invisible(emailList)
+}
+
+```
+
+```{r}
+emailStruct <- mapply(processAllEmail, directories,
+                      isSpam = rep( c(F, T), 3:2))
+emailStruct <- unlist(emailStruct, recursive = F)
+```
+
+```{r}
+sampleStruct <- emailStruct[ 1 ]
+```
+
+### Deriving Variables from the email Messages
+
+```{r}
+
+```

### 2020-05-06 Case Studies/03_spam_identification.rmd
spam case study clean-up
spam case study clean-up
spam identification, recursive partitioning approach
spam identification, recursive partitioning approach

@@
 
+library(rpart, quietly = TRUE, warn.conflicts = FALSE)
+library(rpart.plot, quietly = TRUE, warn.conflicts = FALSE)
+
 options(tinytex.verbose = TRUE)
@@
 CTloc <- sapply(headerList, grep, pattern = "Content-Type")
-CTloc
 
@@
 }
+
 ```
@@
 
-isSpam <- rep(c(FALSE, FALSE, FALSE, TRUE, TRUE), numMsgs)
+isSpam <- c(rep(FALSE, numMsgs[1]), 
+            rep(FALSE, numMsgs[2]), 
+            rep(FALSE, numMsgs[3]), 
+            rep(TRUE, numMsgs[4]),
+            rep(TRUE, numMsgs[5]))
 
@@
 ```{r}
+
 numEmail <- length(isSpam)
@@
 
-log(spamWordsProbs) - log(hamWordsProbs)
-
+probs <- log(spamWordsProbs) - log(hamWordsProbs)
+
+head(probs)
 ```
@@
 ```{r}
-sampleStruct <- emailStruct[ 1 ]
+sampleStruct <- emailStruct[ 1:15 ]
 ```
@@
 ```{r}
-
-```
+header <- sampleStruct[[1]]$header
+subject <- header["Subject"]
+els <- strsplit(subject, "")
+all(els %in% LETTERS)
+
+testSubject <- c("DEAR MADAM", "WINNER!", "")
+
+els <- strsplit(testSubject, "")
+sapply(els, function(subject) all(subject %in% LETTERS))
+
+gsub("[[:punct:] ]", "", testSubject)
+gsub("[^[:alpha:]]", "", testSubject)
+
+```
+
+```{r}
+isYelling <- function(msg) {
+   if( "Subject" %in% names(msg$header) ) {
+      el <- gsub("[^[:alpha:]]", "", msg$header["Subject"])
+
+      if ( nchar(el) > 0 )
+         nchar(gsub("[A-Z]", "", el) < 1 )
+      else
+         FALSE
+   } else {
+      NA
+   }
+}
+
+perCaps <- function(msg) {
+
+   body <- paste(msg$body, collapse = "")
+
+   # Return NA if the body of the message is "empty"
+   if(length(body) == 0 || nchar(body) == 0) return (NA)
+
+   # Eliminate non-alpha characters
+   body <- gsub("[^[:alpha:]]", "", body)
+   capText <- gsub("[^A-Z]", "", body)
+   100 * nchar(capText)/nchar(body)
+}
+
+```
+
+```{r}
+sapply(sampleStruct, perCaps)
+```
+
+```{r}
+
+funcList <- list(
+
+   isRe = function(msg) {
+      "Subject" %in% names(msg$header) &&
+         length(grep("^[ ]*Re:", msg$header[["Subject"]])) > 0
+   },
+   numLines = function(msg) {
+      length(msg$body)
+   },
+   isYelling = function(msg) {
+      if( "Subject" %in% names(msg$header) ) {
+         el <- gsub("[^[:alpha:]]", "", msg$header["Subject"])
+
+         if ( nchar(el) > 0 )
+            nchar(gsub("[A-Z]", "", el) < 1 )
+         else
+            FALSE
+      } else {
+         NA
+      }
+   },
+   perCaps = function(msg) {
+
+      body <- paste(msg$body, collapse = "")
+
+      # Return NA if the body of the message is "empty"
+      if(length(body) == 0 || nchar(body) == 0) return (NA)
+
+      # Eliminate non-alpha characters
+      body <- gsub("[^[:alpha:]]", "", body)
+      capText <- gsub("[^A-Z]", "", body)
+      100 * nchar(capText)/nchar(body)
+   }
+)
+
+```
+
+```{r}
+lapply(funcList, function(func)
+   sapply(sampleStruct, function(msg) func(msg)))
+```
+
+```{r}
+createDerivedF <- function(email = emailStruct, operations = funcList,
+                           verbose = F)
+{
+   els <- lapply(names(operations),
+                 function(id) {
+                     if(verbose) print(id)
+                     e <- operations[[id]]
+                     v <- if(is.function(e))
+                           sapply(email, e)
+                        else
+                           sapply(email, function(msg) eval(e))
+                     v
+                  })
+
+   df <- as.data.frame(els)
+   names(df) <- names(operations)
+
+   invisible(df)  
+}
+```
+
+```{r}
+sampleDF <- createDerivedF(sampleStruct)
+
+```
+
+```{r}
+spam_data <- file.path(data.dir, "spamAssassinDerivedDF.rda")
+
+load(spam_data)
+```
+
+```{r}
+perCaps2 <- function(msg) {
+
+   body <- paste(msg$body, collapse = "")
+
+   # return NA if the body of the message is "empty"
+   if(length(body) == 0 || nchar(body) == 0) return(NA)
+
+   # eliminate non-alpha characters and empty lines
+   body <- gsub("[^[:alpha:]]", "", body)
+   els <- unlist(strsplit(body, ""))
+   ctCap <- sum(els %in% LETTERS)
+   100 * ctCap / length(els)
+}
+
+```
+
+```{r}
+pC <- sapply(emailStruct, perCaps)
+pC2 <- sapply(emailStruct, perCaps2)
+
+identical(pC, pC2)
+```
+
+```{r}
+indNA <- which(is.na(emailDF$subExcCt))
+
+indNoSubject <- which(sapply(emailStruct,
+                             function(msg)
+                                !("Subject" %in% names(msg$header))))
+
+all(indNA == indNoSubject)
+```
+
+```{r}
+all(emailDF$bodyCharCt > emailDF$numLines)
+```
+
+```{r}
+
+long_lines <- head(sort(emailDF$numLines, decreasing = T), 10)
+
+rem <- which(emailDF$numLines %in% long_lines)
+
+ggplot(emailDF[-rem, ], aes(bodyCharCt, numLines)) +
+   geom_point(aes(col = bodyCharCt)) +
+   geom_smooth(method = "lm") +
+   scale_x_continuous(lim = c(0, 35000))
+
+```
+
+#### Exploring the email Feature Set
+
+```{r}
+percent <- emailDF$perCaps
+isSpamLabs <- factor(emailDF$isSpam, labels = c("ham", "spam"))
+boxplot(log(1 + percent) ~ isSpamLabs,
+        ylab = "Percent Capitals (log)")
+
+```
+
+```{r}
+
+ggplot(emailDF, aes(perCaps, bodyCharCt, col = isSpam)) +
+   geom_point() +
+   scale_y_log10() +
+   scale_x_log10()
+
+```
+
+```{r}
+
+colI <- c("#4DAF4A80", "#984EA380")
+logBodyCharCt <- log(1 + emailDF$bodyCharCt)
+logPerCaps <- log(1 + emailDF$perCaps)
+plot(logPerCaps ~ logBodyCharCt, xlab = "Total Characters (log)",
+     ylab = "Percent Capitals (log)",
+     col = colI[1 + emailDF$isSpam],
+     xlim = c(2, 12), pch = 19, cex = 0.5)
+```
+
+```{r}
+table(emailDF$numAtt, isSpamLabs)
+```
+
+```{r}
+colM <- c("#E41A1C80", "#377EB880")
+isRe <- factor(emailDF$isRe, labels = c("no Re:", "Re:"))
+mosaicplot(table(isSpamLabs, isRe), main = "",
+           xlab = "", ylab = "", color = colM)
+
+fromNE <- factor(emailDF$numEnd, labels = c("No #", "#"))
+mosaicplot(table(isSpamLabs, fromNE), color = colM,
+           main = "", xlab = "", ylab = "")
+```
+
+### Fitting Recursive Partition
+
+```{r}
+setupRpart <- function(data) {
+   logicalVars <- which(sapply(data, is.logical))
+   facVars <- lapply(data[, logicalVars],
+                     function(x) {
+                        x = as.factor(x)
+                        levels(x) = c("F", "T")
+                        x
+                     })
+   cbind(facVars, data[, - logicalVars])
+}
+
+emailDFrp <- setupRpart(emailDF)
+```
+
+```{r}
+set.seed(418910)
+
+numSpam <- sum(isSpam)
+numHam <- numEmail - numSpam
+
+testSpamIdx <- sample(numSpam, size = floor(numSpam/3))
+testHamIdx <- sample(numHam, size = floor(numHam/3))
+
+testDF <- rbind( emailDFrp[ emailDFrp$isSpam == "T", ][testSpamIdx, ],
+                 emailDFrp[ emailDFrp$isSpam == "F", ][testHamIdx, ])
+
+trainDF <- rbind( emailDFrp[ emailDFrp$isSpam == "T", ][-testSpamIdx, ],
+                 emailDFrp[ emailDFrp$isSpam == "F", ][-testHamIdx, ])
+
+```
+
+```{r}
+rpartFit <- rpart(isSpam ~ ., data = trainDF, method = "class")
+
+prp(rpartFit, extra = 1)
+```
+
+```{r}
+predictions <- predict(rpartFit,
+                       newdata = testDF[, names(testDF) != "isSpam"],
+                       type = "class")
+
+predsForHam <- predictions[ testDF$isSpam == "F" ]
+summary(predsForHam)
+
+sum(predsForHam == "T", na.rm = T) / length(predsForHam)
+
+predsForSpam <- predictions[ testDF$isSpam == "T" ]
+sum(predsForSpam == "F", na.rm = T) / length(predsForSpam)
+```
+
+```{r}
+args(rpart.control)
+```
+
+```{r}
+complexityVals <- c(seq(0.00001, 0.0001, length = 19),
+                    seq(0.0001, 0.001, length = 19),
+                    seq(0.001, 0.005, length = 9),
+                    seq(0.005, 0.01, length = 9))
+
+fits <- lapply(complexityVals, function(x) {
+   rpartObj <- rpart(isSpam ~ ., data = trainDF,
+                     method = "class",
+                     control = rpart.control(cp=x))
+
+   predict(rpartObj,
+           newdata = testDF[, names(testDF) != "isSpam"],
+           type = "class")
+})
+
+```
+
+```{r}
+spam <- testDF$isSpam == "T"
+numSpam <- sum(spam, na.rm = T)
+numHam <- sum(!spam, na.rm = T)
+
+errs <- sapply(fits, function(preds) {
+   typeI = sum( preds[ !spam ] == "T", na.rm = T) / numHam
+   typeII = sum( preds[ spam ] == "F", na.rm = T) / numSpam
+   c(typeI = typeI, typeII = typeII )
+})
+
+errs
+
+err_df <- data.table(t(errs))
+
+```

### 2020-05-08 Case Studies/03_spam_identification.rmd
spam classification
spam classification

@@
 ```
+
+### Further Analysis
+
+### 1.)
+
+We hand-selected email to belong to the sample set in the _sampleEmail_.
+Instead of this approach, use the sample function to choose messages at random for the sample.
+Be sure to take files from all 5 directories of the email.
+
+```{r}
+dir <- directories[1]
+
+dir_lens <- sapply(directories, function(dir) length(list.files(dir)))
+
+new_sample <- sapply(directories[ dir_lens > 1 ], function(dir) {
+  file_list <- list.files(dir)
+
+  n <- length(file_list)
+
+  idx <- sample(1:n, 30)
+
+  return(file_list[idx])
+})
+
+new_sample
+
+```
+
+### 2.) In the text mining approach to detecting spam we ignored all attachments in creating the set of words belonging to a message (see Section 3.5.2).
+Write a function to extract words from any plain text or HTML attachment and include these words in the set of a message's words.
+Try to reuse _findMsg()_ function and modify the _dropAttach()_ function to accept an additional parameter that indicates whether or not the words in the attachments are to be extracted.
+
+```{r}
+
+index <- hasAttach[2]
+
+boundary <- getBoundary(headerList[[index]])
+body <- sampleSplit[[index]]$body
+body
+
+includeAttach <- function(body, boundary) {
+
+   if(is.null(body)) {
+      return("")
+   }
+
+   bString <- paste("--", boundary, sep = "")
+   bStringLocs <- which(bString == body)
+
+   eString <- paste("--", boundary, "--", sep = "")
+   eStringLoc <- which(eString == body)
+
+   return(msg)
+}
+
+```
+
+### 3.)
+
+The string manipulation functions in R can be used instead of regular expression functions for finding, changing and extracting substrings from strings.
+These functions include: _strsplit()_ to divide a string up into pieces, _substr()_ to extract a portion of a string, _paste()_ to glue together multiple strings, and _nchar()_, which returns the number of characters in a string.
+Write your own version of the get boundary strings from the Content-Type.
+
+```{r}
+
+header <- sampleSplit[[6]]$header
+boundaryIdx <- grep("boundary=", header)
+header[boundaryIdx]
+
+boundary <- header[ str_which(header, "boundary=") ]
+pieces <- unlist(strsplit(boundary, '='))
+pieces <- pieces[ str_which(pieces, ";") ]
+
+sub(".*boundary=\"(.*)\";.*", "\\1", header[boundaryIdx])
+
+header2 <- headerList[[9]]
+boundaryIdx2 <- grep("boundary=", header2)
+header2[boundaryIdx2]
+
+sub('.*boundary="(.*)";.*', "\\1", header2[boundaryIdx2])
+
+boundary2 <- gsub('"', "", header2[boundaryIdx2])
+
+sub(".*boundary= *(.*);?.*", "\\1", boundary2)
+
+boundary <- gsub('"', "", header[boundaryIdx])
+sub(".*boundary= *(.*);?.*", "\\1", boundary)
+
+getBoundary2 <- function(header) {
+   boundary <- header[ str_which(header, "boundary=") ]
+   pieces <- unlist(strsplit(boundary, '='))
+   pieces <- pieces[ str_which(pieces, ";") ]
+   paste("==", pieces, sep = "")
+}
+
+getBoundary(header)
+getBoundary2(header)
+```
+
+### 4.)
+
+Write the __dropAttach()_ function for Section 3.5.2.
+This funciton has two inputs, the body of a mkessage and the boundary string that marks the location of the attachments.
+It returns the body without its attachments.
+Include in the return value the lines of the body that follow the first boundary string up to the string marking the first attachment and the lines following the ending boundary string.
+Be sure to consider the idiosyncratic cases of no attachments and a missing ending boundary string.
+
+```{r}
+processAttach <- function(body, contentType ) {
+
+   boundary <- getBoundary(contentType)
+
+   bString <- paste("--", boundary, sep = "")
+   bStringLocs <- which(bString == body)
+
+   eString <- paste("--", boundary, "--", sep = "")
+   eStringLoc <- which(eString == body)
+
+   n <- length(body)
+
+   if(length(bStringLocs) == 2) {
+
+      bodyContent <- body[(bStringLocs[1] + 2):(bStringLocs[2] - 1)]
+
+      emptyLines <- which(bodyContent == "")
+      bodyContent <- bodyContent[ - emptyLines]
+
+      attachContent <- body[(bStringLocs[2] + 1):n]
+
+      aLen <- diff(c(bStringLocs[-1], eStringLoc))
+      aType <- body[bStringLocs[-1] + 1]
+
+      if(length(aLen) == length(aType)) {      
+         attachments <- data.frame(aLen = aLen, aType = aType)
+      } else {
+         attachments <- data.frame(aLen = c(), aType = c())
+      }
+
+   } else {
+      if( length(bStringLocs) == 0 ) {
+         bodyContent <- body
+      } else {
+         bodyContent = body
+      }
+      attachments <- data.frame(aLen = c(), aType = c())
+   }
+
+   return(list(body = bodyContent, attachDF = attachments ))
+}
+
+```
+
+### 5.)
+
+Write the function __findMsgWords()__ of Section 3.5.3.
+This function takes as input the message body (with no attachments) and the return value is a vector of the unique words in the message.
+That is, we only track which words are in the message, not the number of times these words appear in the message.
+Consider wheather it is simpler to split the string by blanks first and then process the puncuation, digits, etc.
+The function should convert capital letters to lower case and drop all stop words and words that are only one letter long.
+A vector of stop words is avaliable in the tm package.
+
+```{r}
+msg <- sampleSplit[[3]]$body
+msg
+
+findMsgWords <- function(msg) {
+
+}
+
+words <- unlist(sapply(msg, function(line) strsplit(line, " ")))
+names(words) <- ""
+
+n <- length(words)
+
+msg_words <- vector(mode = "character")
+
+for(i in 1:n)
+{
+   word <- tolower(words[i])
+
+   if( word %in% exclude_word_list) {
+      next
+   }
+
+}
+
+tw <- "http://us.click.yahoo.com/pt6YBB/NXiEAA/mG3HAA/7gSolB/TM"
+
+gsub(tw, pattern = "(//)", "")
+
+exclude_word_list
+```
+
+### 6.)

### 2020-05-13 Case Studies/04_robot-sensor-logs.rmd
robot path finding, julia into
robot path finding, julia into

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Processing Robot and Sensor Log Files}
   \rfoot{\color{headergrey}Chapter 4}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Case Studies In Data Science}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(lattice, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- file.path(here::here(), "Case Studies", "datasets", "robot")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# Processing Robot and Sensor Log File

## Seeking a Circular Target

![Robot Training Ground.](minefield.png)

### Robot Log Data

```{r}
ff <- list.files(data.dir, full.names = T)
head(ff)
```

```{r}
ff <- list.files(data.dir, full.names = T,
                 pattern = "JRSPdata.*\\.log")
head(ff)

info <- file.info(ff)
head(info)
```

```{r}
summary(info$size/1024^2)
```

```{r}
file_info <- data.frame( size = info$size/1024^2)

ggplot(file_info, aes(size)) +
   geom_density(fill = "darkgrey", alpha = .4) +
   labs(title = "Log File Size")
```

```{r}
quantile(file_info$size, seq(.9, 1, by = 0.01))
```

### Read Log File

```{r}

filename <- "D:/Projects/Statistical-Computing/Case Studies/datasets/robot/JRSPdata_2010_03_10_12_12_31.log"
lines <- readLines(filename)

readLog <- function(filename,
                    lines = readLines(filename))
{
   lines <- grep("^#", lines, invert = T, value = T)
   els <- strsplit(lines, "[[:space:]]+")

   # Get the interface and type so we can subset.
   iface <- sapply(els, `[`, 4)
   type <- sapply(els, `[`, 6)

   # find the indices corresponding to a position2d
   # with a laser immediately after.
   i <- which(iface == "position2d" & type == "001")
   i = i[ iface[i+1] == "laser" & type[i+1] == "001"]

   # Get the time, x, y, and then the range of values
   # from the laser below.
   locations <- t(sapply(els[i], `[`, c(1, 8, 9)))
   ranges <- t(sapply(els[i + 1], `[`, seq(14, by = 2, length = 361) ))

   # now combine these into a data framee
   locations <- as.data.frame(lapply(1:ncol(locations),
                                     function(i)
                                        as.numeric(locations[, i])))
   ranges <- as.data.frame(lapply(1:ncol(ranges),
                                  function(i)
                                     as.numeric(ranges[, i])))

   ans <- cbind(locations, ranges)

   names(ans) <- c("time", "x", "y",
                   sprintf("range%d", 1:ncol(ranges)))

   invisible(ans)
}

log <- readLog(filename, lines)

head(log)
```

### Data Exploration

```{r}
system.time(logs <- lapply(ff, readLog))
names(logs) <- ff

dur <- sapply(logs, function(x) x$time[nrow(x)] - x$time[1])

range(dur)

ggplot(data.table(dur * 30 * 60), aes(dur)) +
   geom_density(fill = "darkgrey", alpha = .4) +
   labs(title = "Robot Experiment Durations") +
   scale_x_continuous(labels = comma)
```

```{r}
range(sapply(logs, function(ll) range(ll$x)))

range(sapply(logs, function(ll) range(ll$y)))

table(sapply(logs, function(ll) all( diff(ll$time) > 0)))
```

```{r}
deltas <- unlist(lapply(logs, function(ll) diff(ll$time)))
summary(deltas)
```

```{r}
quantile(deltas, seq(.99, 1, length = 11))

which.max(deltas)

nrow(logs[["D:/Projects/Statistical-Computing/Case Studies/datasets/robot/JRSPdata_2010_03_10_12_39_46.log"]])

summary(sapply(logs, nrow))
```

```{r}
ll <- logs[["D:/Projects/Statistical-Computing/Case Studies/datasets/robot/JRSPdata_2010_03_10_12_39_46.log"]]
i <- which.max(diff(ll$time))

delta.x <- unlist(lapply(logs, function(ll) diff(ll$x)))
delta.y <- unlist(lapply(logs, function(ll) diff(ll$y)))

delta_df <- data.table(dx = delta.x, dy = delta.y)

ggplot(delta_df) +
   geom_density(aes(dx), col = "cornflowerblue") +
   geom_density(aes(dy), col = "pink") +
   labs(title = "Dx, Dy", xlab = "Distance", ylab = "Density")
```

```{r}
tail(sort(table(delta.x)))

velocity <- lapply(logs, function(ll) sqrt( diff(ll$x)^2 + diff(ll$y)^2)/diff(ll$time))

velocity_df <- data.table( velocity = unlist(velocity))

summary(velocity_df$velocity)

ggplot(velocity_df, aes(velocity)) +
   geom_density(fill = "darkgrey", alpha = .4) +
   labs(title = "Robot Velocity")
```

```{r}

plot_path <- function(ll) {
   ggplot(ll, aes(x, y)) +
      geom_point()
}

plot_path(logs[[1]])
```

```{r}
makeColorRamp <- function(n) {
   s <- (1:n)/n
   zero <- rep(0, n)
   rgb(s, (1-s), zero)
}

plot.RobotLog <-
function(x, y, col = makeColorRamp(nrow(x)), ...){
   plot(y ~ x, x, type = "p", pch = 20, col = col, ... )
   points(x$x[c(1, nrow(x))], x$y[c(1, nrow(x))],
          pch = c("O", "+"), col = c("green", "blue"))
}

par(mfrow = c(10, 10), mar = rep(0, 4), pty = 's')
invisible(lapply(logs, plot.RobotLog,
                 xlim = c(-16, 16), ylim = c(-8, 8),
                 axes = F))
```

### Exploring a "Look"

```{r}
plot.Look <-
   function(row, ...) {

      x <- row[1, "x"]
      y <- row[1, "y"]

      theta <- seq(0, 2*pi, length = 360) - pi/2
      r <- as.numeric(row[1, -c(1:3, 365)])
      x1 <- x + r * cos(theta)
      y1 <- y + r * sin(theta)
      par(pty = 's')
      plot(x + 2*cos(theta), y + 2*sin(theta),
           col = "red", type = "l",
           xlab = "x", ylab = "y", ...)
      points(x1, y1, type = "l")
   }

par(mfrow = c(1,1))
plot.Look(logs[[1]][ nrow(logs[[1]]), ])
```

```{r}
par(mfrow = c(1, 1))

index <- 2
row <- logs[[index]][ nrow(logs[[index]]), ]

x <- row[1, "x"]
y <- row[1, "y"]

theta <- seq(0, 2*pi, length = 360) - pi/2
r <- as.numeric(row[1, -c(1:3, 365)])

x1 <- x + r * cos(theta)
y1 <- y + r * sin(theta)

plot(x + 2*cos(theta), y + 2*sin(theta),
     col = "lightblue", type = "l",
     xlab = "x", ylab = "y", lwd = 2, lty=4)

eq2 <- cbind( x = x1[ which(r == 2) ], y = y1[ which(r == 2) ] )
lt2 <- cbind( x = x1[ which(r < 2) ], y = y1[ which(r < 2) ] )

points(eq2, type = "l", col="purple")
points(lt2, type ="l", col = "purple")

```

```{r}

row <- logs[[1]][ nrow(logs[[1]]), ]

plot.Look2 <- function(row, ...) {

   x <- row[1, "x"]
   y <- row[1, "y"]

   theta <- seq(0, 2*pi, length = 360) - pi/2
   r <- as.numeric(row[1, -c(1:3, 365)])

   x1 <- x + r * cos(theta)
   y1 <- y + r * sin(theta)

   plot(x + 2*cos(theta), y + 2*sin(theta),
        col = "lightblue", type = "l",
        xlab = "x", ylab = "y", lwd = 2, lty=4)

   eq2 <- cbind( x = x1[ which(r == 2) ], y = y1[ which(r == 2) ] )
   lt2 <- cbind( x = x1[ which(r < 2) ], y = y1[ which(r < 2) ] )

   points(eq2, type = "l", col="purple")
   points(lt2, type ="l", col = "purple")
}

par(mar = rep(0, 4), mfrow = c(10, 10), pty = 's')

invisible(lapply(logs, function(ll)
   plot.Look2(ll[ nrow(ll), ], axes = FALSE)))

```

```{r}
e <- unlist(lapply(logs, function(ll) ll$range1 - ll$range361))
summary(e)
```

```{r}
class(ll)

getRangeErrors <- function(ll) {
   ll$pos <- sprintf("%.3f,%.3f", ll$x, ll$y)
   w <- duplicated(ll$pos)

   if(any(w)) {
      tmp <- ll[ ll$pos %in% ll$pos[w], ]
      errs <- unlist(by(tmp[, 4:364], tmp$pos, scale, scale = F,
                        simplify = F))

      errs
   }
}

rangeErrs <- unlist(lapply(logs, getRangeErrors))

```

```{r}
finalLooks <- lapply(logs, function(ll) ll[nrow(ll), ])

length(finalLooks)

par(mar = rep(0, 4), mfrow = c(10, 10), pty = 's')
invisible(lapply(finalLooks, function(ll)
   plot.Look2(ll[ nrow(ll), ], axes = F)))
```

```{r}
install.packages(c('repr', 'IRdisplay', 'evaluate', 'crayon', 'pbdZMQ', 'devtools', 'uuid', 'digest'))
devtools::install_github('IRkernel/IRkernel')
IRkernel::installspec()
```

### 2020-05-16 Deep Into to Julia/00_Config/build.jl
julia config docs
julia config docs

@@ -1,0 +1,7 @@
+#First add depencies for the example
+using Pkg; Pkg.add.(["Plots", "DSP"])
+using Weave
+
+list_out_formats()
+
+weave("FIR_design.md", doctype=:"github", out_path=:pwd)

### 2020-05-17 Deep Into to Julia/build.jl
ops research
ops research

@@ -1,0 +1,1 @@
+using Weave; weave("01_WhyJulia.jmd", doctype=:"github", out_path=:pwd)

### 2020-05-17 Operations Research/build.jl

@@ -1,0 +1,4 @@
+using Weave;
+
+weave("01_Introduciton.jmd", doctype=:"github", out_path=:pwd)
+weave("02_SimpleLinearOptimization.jmd", doctype=:"github", out_path=:pwd)

### 2020-05-17 Operations Research/build.jl
ops research ch2
ops research ch2

@@ -1,4 +1,7 @@
 using Weave;
 
-weave("01_Introduciton.jmd", doctype=:"github", out_path=:pwd)
-weave("02_SimpleLinearOptimization.jmd", doctype=:"github", out_path=:pwd)
+list_out_formats()
+
+weave("01_Introduciton.jmd", doctype=:"pandoc2pdf", out_path=:pwd)
+weave("02_SimpleLinearOptimization.jmd", doctype=:"pandoc2pdf", out_path=:pwd)
+weave("03_Julia-Basics.jmd", doctype=:"pandoc2pdf", out_path=:pwd)

### 2020-05-17 Operations Research/build.jl
ops research clean-up
ops research clean-up

@@ -2,6 +2,6 @@ using Weave;
 
 list_out_formats()
 
-weave("01_Introduciton.jmd", doctype=:"pandoc2pdf", out_path=:pwd)
-weave("02_SimpleLinearOptimization.jmd", doctype=:"pandoc2pdf", out_path=:pwd)
-weave("03_Julia-Basics.jmd", doctype=:"pandoc2pdf", out_path=:pwd)
+weave("01_Introduciton.jmd", doctype=:"github", out_path=:pwd)
+weave("02_SimpleLinearOptimization.jmd", doctype=:"github", out_path=:pwd)
+weave("03_Julia-Basics.jmd", doctype=:"github", out_path=:pwd)

### 2020-05-17 Operations Research/build.jl
ops research, ch3
ops research, ch3

@@ -5,3 +5,4 @@ list_out_formats()
 weave("01_Introduciton.jmd", doctype=:"github", out_path=:pwd)
 weave("02_SimpleLinearOptimization.jmd", doctype=:"github", out_path=:pwd)
 weave("03_Julia-Basics.jmd", doctype=:"github", out_path=:pwd)
+weave("03_Gadfly.jmd", doctype=:"github", out_path=:pwd)

### 2020-05-17 Operations Research/build.jl
ops research
ops research

@@ -2,7 +2,7 @@ using Weave;
 
 list_out_formats()
 
-weave("01_Introduciton.jmd", doctype=:"github", out_path=:pwd)
+weave("Operations Research\\01_Introduciton.jmd", doctype=:"md2pdf", out_path=:pwd)
 weave("02_SimpleLinearOptimization.jmd", doctype=:"github", out_path=:pwd)
 weave("03_Julia-Basics.jmd", doctype=:"github", out_path=:pwd)
 weave("03_Gadfly.jmd", doctype=:"github", out_path=:pwd)

### 2020-06-26 Deep Into to Julia/build.jl
julia, continuity
julia, continuity

@@ -13,3 +13,4 @@ weave("10_Trigonometric-Functions.jmd", doctype=:"github", out_path=:pwd)
 weave("11_Limits.jmd", doctype=:"github", out_path=:pwd)
 weave("12_Limit-Extensions.jmd", doctype=:"github", out_path=:pwd)
 weave("13_Continuity.jmd", doctype=:"github", out_path=:pwd)
+weave("13_Continuity.jmd", doctype=:"github", out_path=:pwd)

### 2020-06-27 Case Studies/05_airline.R
case studies, flights
case studies, flights

@@ -1,0 +1,203 @@
+
+x <- read.csv("1987.csv")
+nrow(x)
+
+
+
+totalSat <- 0
+for (year in 1987:1988) {
+   x <- read.csv(paste(year, ".csv", sep=''))
+   totalSat <- totalSat + sum(x$DayOfWeek == 6)
+}
+totalSat
+
+
+
+counts <- sapply(sprintf("%d.csv", 1987:1988),
+                 function(f)
+                    sum(read.csv(f)$DayOfWeek == 6))
+sum(counts)
+
+totalSat <- 0
+for (year in 1987:1988) {
+      x <- read.csv(paste(year, '.csv', sep=''))
+      totalSat <- totalSat + sum(x$DayOfWeek == 6)
+      rm(x)
+      gc()
+}
+totalSat
+
+
+
+library(RSQLite)
+delay.con <- dbConnect("SQLite", dbname = "AirlineDelay.sqlite3")
+
+delays87 <- dbGetQuery(delay.con, 
+                   "SELECT * FROM AirlineDelay WHERE Year=1987")
+
+nrow(delays87)
+
+
+dbGetQuery(delay.con, "SELECT COUNT(*), Year FROM AirlineDelay
+                                              WHERE Year=1987")
+
+
+dbGetQuery(delay.con, 
+          "SELECT COUNT(*), Year FROM AirlineDelay GROUP BY Year")
+
+
+dbGetQuery(delay.con, 
+           "SELECT COUNT(*), DayOfWeek FROM AirlineDelay
+                                       WHERE DayOfWeek = 6")
+
+
+x <- read.big.matrix("airline.csv", header = TRUE, 
+                     backingfile = "airline.bin",
+                     descriptorfile = "airline.desc",
+                     type = "integer", extraCols = "age")
+
+dim(x)  # How big is x?
+
+x[1:6,1:6] # Show the first 6 rows and columns.
+
+
+sum(x[, "Year"] == 1987)
+
+
+sum(x[,"DayOfWeek"] == 6)
+
+
+y <- attach.big.matrix("airline.desc")
+
+foo <- big.matrix(nrow = 3, ncol = 3, type = "integer", init = 0)
+
+foo
+
+
+bar <- foo
+
+bar[1,1] <- 1
+foo
+
+
+x <- attach.big.matrix("airline.desc")
+dayCount = integer(7)
+for (i in 1:7) 
[…]

### 2020-06-27 Case Studies/05_get_airline.sh

@@ -1,0 +1,12 @@
+# Get the airline data.
+for year in {1987..2008}
+do
+  wget http://stat-computing.org/dataexpo/2009/$year.csv.bz2 
+done
+
+# Write the uncompressed data to a single file.
+pbzip2 -dc 1987.csv.bz2 > airline.csv
+for year in {1988..2008}
+do
+  pbzip2 -dc $year.csv.bz2 | sed -e "1d" >> airline.csv
+done

### 2020-06-27 Case Studies/05_make_airline.r

@@ -1,0 +1,99 @@
+library(iotools)
+
+col_types <- c(rep("integer", 8), "character", "integer", "character",
+  rep("integer", 5), "character", "character",
+  rep("integer", 4), "character", rep("integer", 6))
+
+col_names <- c("Year", "Month", "DayofMonth", "DayOfWeek", "DepTime", 
+  "CRSDepTime", "ArrTime", "CRSArrTime", "UniqueCarrier", "FlightNum", 
+  "TailNum", "ActualElapsedTime", "CRSElapsedTime", "AirTime", "ArrDelay", 
+  "DepDelay", "Origin", "Dest", "Distance", "TaxiIn", "TaxiOut", "Cancelled", 
+  "CancellationCode", "Diverted", "CarrierDelay", "WeatherDelay", "NASDelay", 
+  "SecurityDelay", "LateAircraftDelay")
+
+# A mutable closure to skip the header line.
+make_airline_df_gen <- function() {
+  first_run <- TRUE
+  function(chunk) {
+    x <- dstrsplit(chunk, col_types=col_types, sep=",", 
+      skip=as.integer(first_run))
+    colnames(x) <- col_names
+    first_run <<- FALSE
+    x
+  }
+}
+
+#######
+# One pass to get the unique character types.
+#######
+
+make_airline_df <- make_airline_df_gen()
+
+us <- chunk.apply("airline.csv",
+    # Get the unique values for each of the character columns.
+    function(chunk) {
+      x <- make_airline_df(chunk)
+      list(carrier=unique(x$UniqueCarrier),
+           tail_num=unique(x$TailNum),
+           origin=unique(x$Origin),
+           dest=unique(x$Dest),
+           cancel_code=unique(x$CancellationCode))
+    }, CH.MERGE=list)
+
+########
+# Make the maps between the character values and an integer.
+########
+
+unique_reduce <- function(x, y) unique(c(x, y))
+
+carrier_names <- 
+  sort(Reduce(unique_reduce, Map(function(x) x$carrier, us)))
+carrier <- 1:length(carrier_names)
+names(carrier) <- carrier_names
+
+tail_num_names <- sort(Reduce(unique_reduce, Map(function(x) x$tail_num, us)))
+tail_num_names <- tail_num_names[tail_num_names != "NA"]
+tail_num <- 1:length(tail_num_names)
+names(tail_num) <- tail_num_names
+
[…]

### 2020-06-27 Case Studies/06_pairs-trading.rmd
pairs trading case study, pt1
pairs trading case study, pt1

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Pairs Trading}
   \rfoot{\color{headergrey}Chapter 6}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Case Studies In Data Science}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(lattice, quietly = TRUE, warn.conflicts = FALSE)

library(quantmod, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data_dir <- file.path(here::here(), "Case Studies", "datasets", "pairs_trading")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# Pairs Trading

## Our Financial Data

```{r}

readData <- 
   #
   # A function to read the data and convert the Date column to
   # an object of class Date.
   # The date values are expected to be in a column named Date.
   # We may want to relax this and allow the caller specify the 
   # column - by name or index.
   function(fileName, dateFormat = c("%Y-%m-%d", "%Y/%m/%d"), ...)
   {
      data <- read.csv(fileName, header = T,
                       stringsAsFactors = F, ...)
      for(fmt in dateFormat) {
         tmp <- as.Date(data$Date, fmt)
         if(all(!is.na(tmp))) {
            data$Date <- tmp
            break
         }
      }

      data[ ordered(data$Date), ]
   }

```

```{r}
getSymbols("T", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
getSymbols("VZ", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
```

```{r, fig.height=4}
chartSeries(T)
chartSeries(VZ)
```

```{r}
ATT_adj = T$T.Adjusted; VER_adj <- VZ$VZ.Adjusted
```

```{r}
combineStocks <- 
   function(lseries, rseries, 
            stockNames = c(deparse(substitute((a)),
                           deparse(substitute(b)))),
            add_ratio = T) {
   l_adj <- lseries[, 6]; r_adj <- rseries[, 6]
   combined <- merge(l_adj, r_adj)

   df_result <- structure(data.table(Date = index(combined), combined),
             names = c("Date", stockNames))

   if(add_ratio) {

      Ratio <- combined[, 1] / combined[, 2]
      df_result <- cbind(df_result, Ratio)
      colnames(df_result)[4] <- "Ratio"
   }

   df_result
}

```

```{r}

overlap <- combineStocks(T, VZ, c("ATT", "VER"))

names(overlap)

range(overlap$Date)
```

```{r}

plotRatio <-
   function(r, k = 1, date = seq(along = r), ...)
   {
      plot(date, r, type = "l", ...)

      mu <- mean(r); kval <- k * sd(r)
      upper <- mu + kval; lower <- mu - kval

      abline(h = c(mu,
                   upper,
                   lower),
             col = c("darkgreen", rep("red", 2 * length(k))),
             lty = "dashed")
      text(1, upper, upper)
   }

plotRatio(overlap$Ratio, k = .85, 
          overlap$Date, col = "lightgray",
          xlab = "Date", ylab = "Ratio")

```

```{r}

findNextPosition <-
   # e.g., findNextPosition(r)
   #     findNextPosition(r, 1774)
   # Check they are increasing and correctly offset
   function(ratio, startDay = 1, k = 1,
            m = mean(ratio), s = sd(ratio))
   {
      up = m + k * s
      down = m - k * s

      if(startDay > 1)
         ratio <- ratio[ - (1:(startDay-1))]

      isExtreme <- ratio >= up | ratio <= down

      if(!any(isExtreme))
         return(integer())

      start = which(isExtreme)[1]
      backToNormal <- if(ratio[start] > up)
         ratio[ - (1:start) ] <= m
      else
         ratio[ - (1:start) ] >= m

      # return either the end of the position or the index
      # of the end of the vector
      # could return NA for not ended, i.e,, which(backToNormal)[1]
      # for both cases. But then the caller has to interpret that.

      end <- if(any(backToNormal))
               which(backToNormal)[1] + start
             else
                length(ratio)

      c(start, end) + startDay + 1
   }

r <- overlap$Ratio; k <- .85

a <- findNextPosition(r, k = k)
b <- findNextPosition(r, a[2], k = k)
c <- findNextPosition(r, b[2], k = k)

```

### Displaying the Positions

```{r}

showPosition <- 
   function(days, ratios, radius = 100) {
      symbols(days, ratios, circles = rep(radius, 2),
              fg = c("darkgreen", "red"), add = T, inches = F)
   }

showPosition <- 
   function(pos, col = c("darkgreen", "red"), ...)
   {
      if(!is.list(pos))
         return(invisible(lapply(pos, showPosition, col = col, ...)))

      abline(v = pos, col = col, ...)
   }

```

```{r}

plotRatio(r, k, overlap$Date, xlab = "Date", ylab = "Ratio")

showPosition(overlap$Date[a], r[a])
showPosition(overlap$Date[b], r[b])
showPosition(overlap$Date[c], r[c])
```

### Finding all Positions

```{r}
getPositions <-
   function(ratio, k = 1, m = mean(ratio), s = sd(ratio))
   {
      when = list()
      cur = 1

      while(cur < length(ratio)) {
         tmp <- findNextPosition(ratio, cur, k, m, s)
         if(length(tmp) == 0)
            break

         when[[length(when) + 1]] <- tmp
         if(is.na(tmp[2] || temp[2] == length(ratio)))
            break
         cur = tmp[2]
      }

      when
   }
```

```{r}
pos <- getPositions(r, k)
plotRatio(r, k, overlap$Date, xlab = "Date", ylab = "Ratio")
invisible(lapply(pos, function(p)
                           showPosition(overlap$Date[p], r[p])))
```

```{r}
showPosition <-
   function(days, ratio, radius = 70)
   {
      if(is.list(days))
         days <- unlist(days)

      symbols(days, ratio[days],
              circles = rep(radius, length(days)),
              fg = c("darkgreen", "red"),
              add = T, inches = F)
   }
```

```{r}
k <- .5
pos <- getPositions(r, k)
plotRatio(r, k, col = "lightgray", ylab = "ratio")
showPosition(pos, r)
```

### Computing Profit

```{r}
positionProfit <-
   # r = overlap$att / overlap$verizon
   # k = 1.7
   # pos = getPositions(r, k)
   # positionProfile(pos[[1]], overlap$att, overlap$verizon)
   function(pos, stockPriceA, stockPriceB,
            ratioMean = mean(stockPriceA / stockPriceB ),
            p = 0.001, byStock = F)
   {
      if(is.list(pos)) {
         ans = sapply(pos, positionProfit,
                      stockPriceA, stockPriceB, ratioMean, p, byStock)

         if(byStock)
            rownames(ans) <- c("A", "B", "commission")

         return(ans)
      }

      # prices at the start and end of the positions
      priceA <- stockPriceA[pos]
      priceB <- stockPriceB[pos]

      # how many units can we by of A and B with $1?

      unitsOfA <- 1/priceA[1]
      unitsOfB <- 1/priceB[1]

      # The dollar amount of how many units we would buy of A and B
      # at the cost at the end of the position of each
      amt <- c(unitsOfA * priceA[2], unitsOfB * priceB[2])

      # which are we selling
      sellWhat <- if(priceA[1] / priceB[1] > ratioMean) "A" else "B"

      profit <- if(sellWhat == "A")
                  c( ( 1 - amt[1]), (amt[2] - 1), - p * sum(amt))
               else
                  c( (1 - amt[2]), (amt[1] - 1), - p * sum(amt))

      if( byStock )
         profit
      else
         sum(profit)
   }

```

```{r}
pf <- positionProfit(c(1, 2), c(3838.48, 8712.87),
                              c(459.11, 1100.65), p = 0)

prof <- positionProfit(pos, overlap$ATT, overlap$VER, mean(r))
```

### Finding Optimal K

```{r}
i <- 1:floor(nrow(overlap)/2)
train <- overlap[i, ]
test <- overlap[ -i, ]

r.train <- train$Ratio
r.test <- test$Ratio

period <- seq(min(overlap$Date), by = "5 years", length = 2)
period.train <- paste(period[1], period[2], sep="/")

att.train <- T[period.train]$T.Adjusted
verizon.train <- VZ[period.train]$VZ.Adjusted
r.train <- att.train/verizon.train

period.test <- paste(period[2], max(overlap$Date), sep="/")
att.test <- T[period.test]$T.Adjusted
verizon.test <- VZ[period.test]$VZ.Adjusted
r.test <- att.test/verizon.test

k.max <- max((r.train - mean(r.train)) / sd(r.train))
k.min <- min((abs(r.train - mean(r.train)) / sd(r.train)))

ks <- seq(k.min, k.max, length = 1000)
m <- mean(r.train)

profits <-
   sapply(ks,
          function(k) {
             pos <- getPositions(r.train, k)
             sum(positionProfit(pos, train$ATT, train$VER,
                                mean(r.train)))
          })

plot(ks, profits, type = "l", xlab = "k", ylab = "Profit")

max_profits <- which.max(profits)
optimal <- ks[ max_profits ]

```

```{r}
tmp.k <- ks[ profits == max(profits) ]
pos <- getPositions(r.train, tmp.k[1])
all(sapply(tmp.k[-1],
           function(k)
              identical(pos, getPositions(r.train, k))))

k.star <- mean( ks[ profits == max(profits)])

pos <- getPositions(r.test, k.star, mean(r.train), sd(r.train))
testProfit <- sum(positionProfit(pos, test$ATT, test$VER), na.rm=T)

testProfit * 100

```

### 2020-06-27 Bayesian Data Analysis/03_Bayes_vs_Freq_.rmd
bayesian data analysis
bayesian data analysis

--- 2020-06-27 Case Studies/06_pairs-trading.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Pairs Trading}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Bayesian vs Frequentist}
    \rfoot{\color{headergrey}Chapter 6}
    \lfoot{\color{headergrey}}
-   \fancyfoot[C]{\rmfamily\color{headergrey}Case Studies In Data Science}
+   \fancyfoot[C]{\rmfamily\color{headergrey}Bayesian Data Analysis}
 geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
@@
 library(scales, quietly = TRUE, warn.conflicts = FALSE)
+library(cowplot, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
+library(blmeco, quietly = TRUE, warn.conflicts = FALSE)
+library(tidybayes, quietly = TRUE, warn.conflicts = FALSE)
+library(arm, quietly = TRUE, warn.conflicts = FALSE)
+library(coda, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-data_dir <- file.path(here::here(), "Case Studies", "datasets", "pairs_trading")
-
 ```
@@
 
-# Pairs Trading
-
-## Our Financial Data
+# The Bayesian Way
 
@@
 
-readData <- 
-   #
-   # A function to read the data and convert the Date column to
-   # an object of class Date.
-   # The date values are expected to be in a column named Date.
-   # We may want to relax this and allow the caller specify the 
-   # column - by name or index.
-   function(fileName, dateFormat = c("%Y-%m-%d", "%Y/%m/%d"), ...)
-   {
-      data <- read.csv(fileName, header = T,
-                       stringsAsFactors = F, ...)
-      for(fmt in dateFormat) {
-         tmp <- as.Date(data$Date, fmt)
-         if(all(!is.na(tmp))) {
-            data$Date <- tmp
-            break
+# utility function for plotting a posterior
+
+plot_posterior <- function(bsim) {
+   res <- data.table(mu = bsim@coef, sigma = bsim@sigma)
+   colnames(res) <- c("mu", "sigma")
+
+   pmain <- ggplot(res, aes(mu, sigma)) +
+     geom_point(alpha = .65, col = "darkgrey") +
+     labs(x = "", y = "") +
+     stat_density_2d(aes(fill = stat(level), alpha = ..level..), geom = "polygon")
+
+   xbox <- axis_canvas(pmain, axis = "x") +
+     geom_histogram(data = res, aes(mu, y = ..density.., fill = ..count..), size = .2, bins = 50, alpha = .85) +
+     geom_density(data = res, aes(mu), col = "darkgrey", alpha = .85) +
+     theme(legend.position = "none")
+
+   ybox <- axis_canvas(pmain, axis = "y", coord_flip = T) +
+     geom_histogram(data = res, aes(sigma, y = ..density.., fill = ..count..), size = .2, bins = 50, alpha = .85) +
+     geom_density(data = res, aes(sigma), col = "darkgrey", alpha = .85) +
+     coord_flip() +
+     theme(legend.position = "none")
+
+   suppressWarnings({
+     p1 <- insert_xaxis_grob(pmain, xbox, grid::unit(.2, "null"), position = "top")
+     p2 <- insert_yaxis_grob(p1, ybox, grid::unit(.2, "null"), position = "right")
+   })
+
+   ggdraw(p2)
 }
-      }
-
-      data[ ordered(data$Date), ]
-   }
-
 ```
 
+## Bayes Therom
+
+$P(\theta | y) = \frac{P(\theta)P(y|\theta)}{P(y)}$
+
+### Estimating the Mean
+
+The model of the data is:
+
+$y \sim Norm(\theta, \sigma^2)$
+
+Given the data has three measurements,
+
+$y_1 = 27.1, y_2 = 14.6, y_3 = 14.6$
+
+$\sigma^2 = 20$
+
 ```{r}
-getSymbols("T", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
-getSymbols("VZ", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
+n <- 200
+
+x <- rnorm(n, mean = 17, sd = .5)
+triplot.normal.knownvariance(n=n, theta.data = x, variance.known = 20, prior.theta = 0, prior.variance = 17)
 ```
 
-```{r, fig.height=4}
-chartSeries(T)
-chartSeries(VZ)
+### Estimating Mean and Variance using Simulating
+
+```{r}
+# Simulate hypothetical body height measurements
+
+true.mean <- 165
+true.sd <- 10
+y <- round(rnorm(10, mean = true.mean, sd = true.sd))
+
+mod <- lm(y ~ 1)
+mod
+
+summary(mod)$sigma
 ```
 
+$y \sim Norm(\hat{\theta} = 166.3, \sigma^2 = 13.3)$
+
 ```{r}
-ATT_adj = T$T.Adjusted; VER_adj <- VZ$VZ.Adjusted
+nsim <- 5000
+bsim <- sim(mod, n.sim=nsim)
+
+str(bsim)
+
+str(bsim, max.level = 2)
+
+plot_posterior(bsim)
+
+quantile(bsim@coef, prob = c(.025, 0.5, 0.975))
+quantile(bsim@sigma, prob = c(0.025, 0.5, 0.975))
+
+HPDinterval(as.mcmc(bsim@coef))
+
+sum(bsim@coef > 160) / nsim
+mean(bsim@coef > 160)
+
+cvsim <- bsim@sigma / bsim@coef
+quantile(cvsim, prob = c(0.025, 0.5, 0.975))
 ```
 
-```{r}
-combineStocks <- 
-   function(lseries, rseries, 
-            stockNames = c(deparse(substitute((a)),
-                           deparse(substitute(b)))),
-            add_ratio = T) {
-   l_adj <- lseries[, 6]; r_adj <- rseries[, 6]
-   combined <- merge(l_adj, r_adj)
-
-   df_result <- structure(data.table(Date = index(combined), combined),
-             names = c("Date", stockNames))
-
-   if(add_ratio) {
-
-      Ratio <- combined[, 1] / combined[, 2]
-      df_result <- cbind(df_result, Ratio)
-      colnames(df_result)[4] <- "Ratio"
-   }
-
-   df_result
-}
-
-```
+### The Frequentist Way
 
@@
 
-overlap <- combineStocks(T, VZ, c("ATT", "VER"))
-
-names(overlap)
-
-range(overlap$Date)
 ```
-
-```{r}
-
-plotRatio <-
-   function(r, k = 1, date = seq(along = r), ...)
-   {
-      plot(date, r, type = "l", ...)
-
-      mu <- mean(r); kval <- k * sd(r)
-      upper <- mu + kval; lower <- mu - kval
-
-      abline(h = c(mu,
-                   upper,
-                   lower),
-             col = c("darkgreen", rep("red", 2 * length(k))),
-             lty = "dashed")
-      text(1, upper, upper)
-   }
-
-plotRatio(overlap$Ratio, k = .85, 
-          overlap$Date, col = "lightgray",
-          xlab = "Date", ylab = "Ratio")
-
-```
-
-```{r}
-
-findNextPosition <-
-   # e.g., findNextPosition(r)
-   #     findNextPosition(r, 1774)
-   # Check they are increasing and correctly offset
-   function(ratio, startDay = 1, k = 1,
-            m = mean(ratio), s = sd(ratio))
-   {
-      up = m + k * s
-      down = m - k * s
-
-      if(startDay > 1)
-         ratio <- ratio[ - (1:(startDay-1))]
-
-      isExtreme <- ratio >= up | ratio <= down
-
-      if(!any(isExtreme))
-         return(integer())
-
-      start = which(isExtreme)[1]
-      backToNormal <- if(ratio[start] > up)
-         ratio[ - (1:start) ] <= m
-      else
-         ratio[ - (1:start) ] >= m
-
-      # return either the end of the position or the index
-      # of the end of the vector
-      # could return NA for not ended, i.e,, which(backToNormal)[1]
-      # for both cases. But then the caller has to interpret that.
-
-      end <- if(any(backToNormal))
-               which(backToNormal)[1] + start
-             else
-                length(ratio)
-
-      c(start, end) + startDay + 1
-   }
-
-r <- overlap$Ratio; k <- .85
-
-a <- findNextPosition(r, k = k)
-b <- findNextPosition(r, a[2], k = k)
-c <- findNextPosition(r, b[2], k = k)
-
-```
-
-### Displaying the Positions
-
-```{r}
-
-showPosition <- 
-   function(days, ratios, radius = 100) {
-      symbols(days, ratios, circles = rep(radius, 2),
-              fg = c("darkgreen", "red"), add = T, inches = F)
-   }
-
-showPosition <- 
-   function(pos, col = c("darkgreen", "red"), ...)
-   {
-      if(!is.list(pos))
-         return(invisible(lapply(pos, showPosition, col = col, ...)))
-
-      abline(v = pos, col = col, ...)
-   }
-
-```
-
-```{r}
-
-plotRatio(r, k, overlap$Date, xlab = "Date", ylab = "Ratio")
-
-showPosition(overlap$Date[a], r[a])
-showPosition(overlap$Date[b], r[b])
-showPosition(overlap$Date[c], r[c])
-```
-
-### Finding all Positions
-
-```{r}
-getPositions <-
-   function(ratio, k = 1, m = mean(ratio), s = sd(ratio))
-   {
-      when = list()
-      cur = 1
-
-      while(cur < length(ratio)) {
-         tmp <- findNextPosition(ratio, cur, k, m, s)
-         if(length(tmp) == 0)
-            break
-
-         when[[length(when) + 1]] <- tmp
-         if(is.na(tmp[2] || temp[2] == length(ratio)))
-            break
-         cur = tmp[2]
-      }
-
-      when
-   }
-```
-
-```{r}
-pos <- getPositions(r, k)
-plotRatio(r, k, overlap$Date, xlab = "Date", ylab = "Ratio")
-invisible(lapply(pos, function(p)
-                           showPosition(overlap$Date[p], r[p])))
-```
-
-```{r}
-showPosition <-
-   function(days, ratio, radius = 70)
-   {
-      if(is.list(days))
-         days <- unlist(days)
-
-      symbols(days, ratio[days],
-              circles = rep(radius, length(days)),
-              fg = c("darkgreen", "red"),
-              add = T, inches = F)
-   }
-```
-
-```{r}
-k <- .5
-pos <- getPositions(r, k)
-plotRatio(r, k, col = "lightgray", ylab = "ratio")
-showPosition(pos, r)
-```
-
-### Computing Profit
-
-```{r}
-positionProfit <-
-   # r = overlap$att / overlap$verizon
-   # k = 1.7
-   # pos = getPositions(r, k)
-   # positionProfile(pos[[1]], overlap$att, overlap$verizon)
-   function(pos, stockPriceA, stockPriceB,
-            ratioMean = mean(stockPriceA / stockPriceB ),
-            p = 0.001, byStock = F)
-   {
-      if(is.list(pos)) {
-         ans = sapply(pos, positionProfit,
-                      stockPriceA, stockPriceB, ratioMean, p, byStock)
-
-         if(byStock)
-            rownames(ans) <- c("A", "B", "commission")
-
-         return(ans)
-      }
-
-      # prices at the start and end of the positions
-      priceA <- stockPriceA[pos]
-      priceB <- stockPriceB[pos]
-
-      # how many units can we by of A and B with $1?
-
-      unitsOfA <- 1/priceA[1]
-      unitsOfB <- 1/priceB[1]
-
-      # The dollar amount of how many units we would buy of A and B
-      # at the cost at the end of the position of each
-      amt <- c(unitsOfA * priceA[2], unitsOfB * priceB[2])
-
-      # which are we selling
-      sellWhat <- if(priceA[1] / priceB[1] > ratioMean) "A" else "B"
-
-      profit <- if(sellWhat == "A")
-                  c( ( 1 - amt[1]), (amt[2] - 1), - p * sum(amt))
-               else
-                  c( (1 - amt[2]), (amt[1] - 1), - p * sum(amt))
-
-      if( byStock )
-         profit
-      else
-         sum(profit)
-   }
-
-```
-
-```{r}
-pf <- positionProfit(c(1, 2), c(3838.48, 8712.87),
-                              c(459.11, 1100.65), p = 0)
-
-prof <- positionProfit(pos, overlap$ATT, overlap$VER, mean(r))
-```
-
-### Finding Optimal K
-
-```{r}
-i <- 1:floor(nrow(overlap)/2)
-train <- overlap[i, ]
-test <- overlap[ -i, ]
-
-r.train <- train$Ratio
-r.test <- test$Ratio
-
-period <- seq(min(overlap$Date), by = "5 years", length = 2)
-period.train <- paste(period[1], period[2], sep="/")
-
-att.train <- T[period.train]$T.Adjusted
-verizon.train <- VZ[period.train]$VZ.Adjusted
-r.train <- att.train/verizon.train
-
-period.test <- paste(period[2], max(overlap$Date), sep="/")
-att.test <- T[period.test]$T.Adjusted
-verizon.test <- VZ[period.test]$VZ.Adjusted
-r.test <- att.test/verizon.test
-
-k.max <- max((r.train - mean(r.train)) / sd(r.train))
-k.min <- min((abs(r.train - mean(r.train)) / sd(r.train)))
-
-ks <- seq(k.min, k.max, length = 1000)
-m <- mean(r.train)
-
-profits <-
-   sapply(ks,
-          function(k) {
-             pos <- getPositions(r.train, k)
-             sum(positionProfit(pos, train$ATT, train$VER,
-                                mean(r.train)))
-          })
-
-plot(ks, profits, type = "l", xlab = "k", ylab = "Profit")
-
-max_profits <- which.max(profits)
-optimal <- ks[ max_profits ]
-
-```
-
-```{r}
-tmp.k <- ks[ profits == max(profits) ]
-pos <- getPositions(r.train, tmp.k[1])
-all(sapply(tmp.k[-1],
-           function(k)
-              identical(pos, getPositions(r.train, k))))
-
-k.star <- mean( ks[ profits == max(profits)])
-
-pos <- getPositions(r.test, k.star, mean(r.train), sd(r.train))
-testProfit <- sum(positionProfit(pos, test$ATT, test$VER), na.rm=T)
-
-testProfit * 100
-
-```

### 2020-06-28 Case Studies/06_pairs-trading.rmd
case studies, pair trading / optimized simulations with C
case studies, pair trading / optimized simulations with C

@@
 
-```{r}
+```{r read-data}
 
@@
 
-```{r}
+```{r get-symbols}
+
 getSymbols("T", src = "yahoo", from = "1985-01-01", to = "2015-12-31")
@@
 
-```{r, fig.height=4}
+```{r plot-symbols, fig.height=4}
+
 chartSeries(T)
@@
 
-```{r}
+```{r combine-stocks}
+
 combineStocks <- 
@@
 
-```{r}
+```{r get-overlap}
 
@@
 
-```{r}
+```{r plot-ratio}
 
@@
 
-```{r}
+```{r find-positions}
 
@@
       if(startDay > 1)
-         ratio <- ratio[ - (1:(startDay-1))]
+           ratio = ratio[ - (1:(startDay-1)) ]
 
-      isExtreme <- ratio >= up | ratio <= down
+     isExtreme = ratio >= up | ratio <= down
 
-      if(!any(isExtreme))
+     if( is.na(isExtreme) || !any(isExtreme))
          return(integer())
@@
 
-```{r}
+```{r show-position}
 
@@
 
-```{r}
+```{r plot-ratio-2}
 
@@
 
-showPosition(overlap$Date[a], r[a])
-showPosition(overlap$Date[b], r[b])
-showPosition(overlap$Date[c], r[c])
+#showPosition(overlap$Date[a], r[a])
+#showPosition(overlap$Date[b], r[b])
+#showPosition(overlap$Date[c], r[c])
 ```
@@
 
-```{r}
+```{r get-positions}
+
 getPositions <-
@@
    }
-```
-
-```{r}
-pos <- getPositions(r, k)
-plotRatio(r, k, overlap$Date, xlab = "Date", ylab = "Ratio")
-invisible(lapply(pos, function(p)
-                           showPosition(overlap$Date[p], r[p])))
 ```
@@
 ```{r}
+
 positionProfit <-
@@
       amt <- c(unitsOfA * priceA[2], unitsOfB * priceB[2])
+
+      if(is.na(priceA[1]) | is.na(priceB[1]))
+         return(0)
 
@@
 ```
+
+## Simulation
+
+Vector Auto-regression
+
+$X_t^{(1)} = \rho X^{(1)}_{t-1} + \psi(1 - \rho)X^{(2)}_{t-1} + \epsilon^{(1)}_t$
+
+$X_t^{(2)} = \rho X^{(2)}_{t-1} + \psi(1 - \rho)X^{(1)}_{t-1}+ \epsilon^{(2)}_t$
+
+$\epsilon^{(1)}_t \sim N(0, \sigma^2_i)$
+
+...
+
+$Y^{(1)}_t = \beta^{(1)}_0 + \beta^{(1)}_1t + X^{1}_t$
+
+$Y^{(2)}_t = \beta^{(2)}_0 + \beta^{(2)}_1t + X^{2}_t$
+
+### Simulating the Stock Price Series
+
+```{r}
+stockSim <-
+   function(n = 4000, rho = 0.99, psi = 0, sigma = rep(1, 2),
+            beta0 = rep(100, 2), beta1 = rep(0, 2),
+            epsilon = matrix(rnorm(2*n, sd = sigma),
+                             nrow = n, byrow = T))
+   {
+      X <- matrix(0, nrow = n, ncol = 2)
+      X[1, ] <- epsilon[1, ]
+
+      A <- matrix(c(rho, psi * (1-rho), psi*(1-rho), rho), nrow = 2)
+      for(i in 2:n)
+         X[i, ] = A %*% X[i-1, ] + epsilon[i, ]
+
+      X[, 1] <- beta0[1] + beta1[1] * (1:n) + X[, 1]
+      X[, 2] <- beta0[2] + beta1[2] * (1:n) + X[, 2]
+
+      X
+   }
+
+```
+
+```{r}
+a <- stockSim(rho = .99, psi = 0)
+
+matplot(1:nrow(a), a, type = "l", xlab = "Day", ylab = "Y",
+        col = c("black", "grey"), lty = "solid")
+
+beta1 <- c(.05, .1)
+
+a <- stockSim(beta1 = c(.05, .1))
+
+matplot(1:nrow(a), a, type = "l", xlab = "Day", ylab = "Y",
+        col = c("black", "grey"), lty = "solid")
+
+```
+
+### Simulation Utilities
+
+```{r}
+
+runSim <-
+function(rho = .99, psi = .9, beta0 = c(100, 100), beta1 = c(0, 0),
+         sigma = c(1, 1), n = 4000)
+{
+    X = stockSim(n, rho, psi, sigma, beta = beta0, beta1 = beta1)
+    train = X[ 1:floor(n/2), ]
+    test = X[ (floor(n/2)+1):n, ]
+    m = mean(train[, 1]/train[, 2])
+    s = sd(train[, 1]/train[, 2])
+    k.star = getBestK(train[, 1], train[, 2], m = m, s = s)
+    getProfit.K(k.star, test[, 1], test[, 2], m, s)
+}
+
+getProfit.K =
+function(k, x, y, m = mean(x/y), s = sd(x/y))
+{
+    pos = getPositions(x/y, k, m = m, s = s)
+    if(length(pos) == 0)  
+       0
+    else
+       sum(positionProfit(pos, x, y, m), na.rm=T)
+}
+
+getBestK = 
+function(x, y, ks = seq(0.1, max.k, length = N), N = 100, 
+         max.k = NA, m = mean(x/y), s = sd(x/y))
+{
+    if(is.na(max.k)) {
+       r = x/y
+       max.k = max(r/sd(r))
+    }
+
+    pr.k = sapply(ks, getProfit.K, x, y, m = m, s = s)
+    median(ks[ pr.k == max(pr.k) ])
+}
+
+simProfitDist = 
+function(..., B = 999) 
+      sapply(1:B,  function(i, ...) runSim(...), ...)
+
+```
+
+### Run the Simulation
+
+```{r}
+system.time({ x = simProfitDist( .99, .9, c(0, 0)) })
+
+ggplot(data.table(profit = x), aes(profit)) +
+   geom_density()
+```
+
+```{r}
+g <- expand.grid(psi = seq(.8, .99, length.out = 20),
+                 beta1 = seq(-0.01, .01, length.out = 20),
+                 beta2 = seq(-0.1, 0.01, length.out = 20))
+
+dim(g)
+```
+
+```{r}
+Rprof("sim.prof")
+
+system.time({x = simProfitDist( .99, .9, c(0, 0))})
+
+Rprof(NULL)
+
+head(summaryRprof("sim.prof")$by.self)
+```
+
+```{r}
+counter <- 0L
+trace(findNextPosition, quote( counter <<- counter + 1L),
+      print = FALSE)
+
+system.time({x = simProfitDist( .99, .9, c(0, 0))})
+
+counter
+
+untrace(findNextPosition)
+```
+
+```{r}
+library(compiler)
+stockSim.cmp <- cmpfun(stockSim)
+
+```
+
+```{r}
+
+tm.orig <- system.time({replicate(80, stockSim())})
+tm.compiled <- system.time({replicate(80, stockSim.cmp())})
+
+tm.orig/tm.compiled
+```
+
+```{r}
+c.lib <- paste(here(), "Case Studies", "06_stockSim.dll", sep="/")
+
+dyn.load(c.lib)
+```
+
+```{r}
+stockSim.c <-
+   function(n = 4000, rho = 0.99, psi = 0, sigma = rep(1, 2),
+            beta0 = rep(100, 2), beta1 = rep(0, 2),
+            epsilon = matrix(rnorm(2*n, sd = sigma), nrow = n))
+   {
+      X <- matrix(0, nrow = n, ncol = 2)
+      X[1, ] <- epsilon[1, ]
+
+      X <- .C("stockSim", X, as.integer(n), rho, psi, epsilon)[[1]]
+
+      X[, 1] <- beta0[1] + beta1[1] + (1:n) + X[, 1]
+      X[, 2] <- beta0[2] + beta1[2] + (1:n) + X[, 2]
+
+      X
+   }
+```
+
+```{r}
+
+e <- matrix(rnorm(2*4000, sd = c(1, 1)), , 2)
+tmp1 <- stockSim.c(epsilon = e)
+tmp2 <- stockSim(epsilon = e)
+
+```
+
+```{r}
+
+stockSim <- stockSim.c
+
+Rprof("sim.prof")
+system.time({x = simProfitDist(.99, .9, c(0, 0))})
+Rprof(NULL)
+head(summaryRprof("sim.prof")$by.self)
+```

### 2020-06-28 Case Studies/06_stockSim.c

@@ -1,0 +1,14 @@
+void
+stockSim( double *ans, const int *len, const double *rho,
+          const double *psi, const double *eps)
+{
+  int i, j;
+  
+  double psi_rho = (1 - *rho) * (*psi);
+  
+  for(i = 1, j = *len + 1; i < *len; i++, j++)
+  {
+    ans[i] = *rho * ans[i - 1] + psi_rho * ans[j - 1] + eps[i];
+    ans[j] = *rho * ans[j - 1] + psi_rho * ans[i - 1] + eps[j];
+  }
+}

### 2020-07-13 Bayesian Data Analysis/04_Normal-Linear-Models.rmd
bayesian data analysis, ch3/4

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Normal Linear Models}
   \rfoot{\color{headergrey}Chapter 4}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Bayesian Data Analysis}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(cowplot, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(lattice, quietly = TRUE, warn.conflicts = FALSE)

library(blmeco, quietly = TRUE, warn.conflicts = FALSE)
library(tidybayes, quietly = TRUE, warn.conflicts = FALSE)
library(arm, quietly = TRUE, warn.conflicts = FALSE)
library(coda, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# Normal Linear Models

## Fitting a Linear Regression in R

```{r}
n <- 50 # sample size
sigma <- 5 # standard deviation of the residuals
b0 <- 2 # intercept
b1 <- 0.7 # slope
x <- runif(n, 10, 30) # sample values of the covariate
yhat <- b0 + b1*x
y <- rnorm(n, yhat, sd=sigma)

# plot the data
plot(x, y, pch=16, las=1, cex.lab=1.2)
abline(lm(y~x), lwd=2, col="blue") # insert regression line
# add residuals
segments(x, fitted(lm(y~x)), x, y, lwd=2, col="orange", lty=3)
```

Fit the model

```{r}
mod <- lm(y~x)
mod

summary(mod)

summary(mod)$sigma
```

### Drawing Conclusions

Alternative Display

```{r}
nsim <- 1000
bsim <- sim(mod, n.sim = nsim)

plot(x, y, pch=16, las=1, cex.lab=1.2)
for(i in 1:nsim) abline(coef(bsim)[i, 1], coef(bsim)[i, 2],
                        col = rgb(0, 0, 0, 0.05))

newdat <- data.frame(x = seq(10, 30, by=0.1))
newmodmat <- model.matrix(~x, data = newdat)
fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))

for(i in 1:nsim) fitmat[, i] <- newmodmat %*% coef(bsim)[i, ]
plot(x, y, pch=16, las=1, cex.lab=1.2)
abline(mod, lwd=2)

lines(newdat$x, apply(fitmat, 1, quantile, prob = 0.025, lty=3))
lines(newdat$x, apply(fitmat, 1, quantile, prob=0.975, lty=3))
```

Predicting Future Observations

```{r}
plot(x, y, pch=16, las=1, cex.lab=1.2)

# prepare matrix for simulated new data
newy <- matrix(ncol = nsim, nrow = nrow(newdat))
# for each simulate4d fitted value, simulate  one new y-value
for(i in 1:nsim) newy[, i] <- rnorm(nrow(newdat), mean = fitmat[, i],
                                    sd = bsim@sigma[i])

lines(newdat$x, apply(newy, 1, quantile, prob=0.025), lty=3)
lines(newdat$x, apply(newy, 1, quantile, prob=0.975), lty=3)
```

```{r}
sum(newy[newdat$x == 25,] > 20) / nsim
```

### Frequentist Results

```{r}
summary(mod)
```

## Regression Variants: ANOVA, ANCOVA, and MLR

```{r}
# data simulation
mu <- 12
sigma <- 2
b1 <- 3
b2 <- -5
n <- 90
group <- factor(rep(c(1, 2, 3), each = 30))

# simulate the y variable
simresid <- rnorm(n, mean=0, sd=sigma)
y <- mu + as.numeric(group == 2) * b1 + as.numeric(group == "3") * b2 +
   simresid

```

```{r}
group <- factor(group) # define group as a factor
mod <- lm(y ~ group) # fit the model

mod

summary(mod)$sigma
```

```{r}
bsim <- sim(mod, n.sim=1000)
m.g1 <- coef(bsim)[, 1]
m.g2 <- coef(bsim)[, 1] + coef(bsim)[, 2]
m.g3 <- coef(bsim)[, 1] + coef(bsim)[, 3]

m.dat <- data.table(group1 = m.g1, group2 = m.g2, group3 = m.g3)
m.dat.long <- melt(m.dat)
m.dat.long[, Mean := mean(value), by = variable]

p1 <- ggplot(m.dat.long) +
   geom_density(aes(value, fill = variable), alpha = .6)

p2 <- ggplot(m.dat.long) +
   geom_point(aes(x = variable, y = Mean, color = variable), alpha = .75, size = 6)

grid.arrange(p1, p2, ncol=2)

d.g1.2 <- m.g1 - m.g2
mean(d.g1.2)
quantile(d.g1.2, prob=c(0.025, 0.975))

sum(m.g2 > m.g1) / nsim
```

## Frequentist Results

```{r}
data("periparusater")
dat <- periparusater

mod <- lm(wing ~ sex + age, data=dat)
plot(mod)

newdat <- expand.grid(sex = factor(c(1, 2)), age = factor(c(3, 4)))
newdat$fit <- model.matrix(~sex + age, dat=newdat) %*% coef(mod)

nsim <- 2000
bsim <- sim(mod, n.sim = nsim)

fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))
Xmat <- model.matrix(formula(mod)[c(1,3)], dat = newdat)
for(i in 1:nsim) fitmat[, i] <- Xmat %*% bsim@coef[i, ]

alpha <- .05
intervals <- c(lower = alpha/2, upper = 1 - alpha/2)

ci <- t(apply(fitmat, 1, quantile, prob = intervals))

cbind(newdat, ci)

mod2 <- lm(wing ~ sex * age, data = dat)

```

```{r}
bsim2 <- sim(mod2, n.sim = nsim)
quantile(bsim2@coef[, 4], prob = c(0.025, 0.5, 0.975))

summary(mod2)$sigma

mean(abs(bsim2@coef[, 4]) > 0.3)

coef(mod2)

quantile(bsim2@coef[, 2], prob = c(0.025, 0.5, 0.975))

sum(bsim@coef[, 2] < 0)/nsim # for juveniles (reference level)
sum(bsim@coef[, 2] + bsim2@coef[, 4] < 0) / nsim # adults

```

## Analysis of Covariance

```{r}
data("ellenberg")

index <- is.element(ellenberg$Species, c("Ap", "Dg"))
dat <- ellenberg[index, ] # select two species
dat <- droplevels(dat) # drop unused factor levels
str(dat) # definitions

mod <- lm(log(Yi.g) ~ Species + Water, data = dat)

head(model.matrix(mod)) # print the first 6 rows of the matrix

summary(mod)

mod2 <- lm(log(Yi.g) ~ Species*Water, data = dat)
summary(mod2)
```

```{r}
nsim <- 2000
bsim <- sim(mod2, n.sim = nsim)
coefs <- coef(bsim)

xatcross <- crosspoint(coefs[, 1], coefs[, 3],
                       coefs[, 1] + coefs[, 2], coefs[, 3] + coefs[, 4])[ , 1]

xatcross[xatcross< (-5)] <- -5

th <- hist(xatcross, breaksw = seq(-5.5, 140.5, by=5))
plot(th$mids, cumsum(th$counts)/2000, type = "l", lwd=2, las=1,
     ylim=c(0, 1), ylab="P(Dg > Ap|data)", xlab="Average diistance to ground water (cm)")
```

## Multiple Regression and Collinearity

```{r}
data(mdat)
mod <- lm(y ~ x1 + x2, data = mdat)
summary(mod)
```

```{r}
cor(mdat[, 2:6])

ggpairs(mdat)

own.graph <- function(x, y) {
   points(x, y, pch=16, col=rgb(1, 0.5, 0.0, 0.8))
   abline(lm(y~x))
}
pairs(mdat, panel = own.graph)
```

## Ordered Factors and Constants

```{r}
data("swallows")
levels(swallows$nesting_aid)

str(swallows)

contrasts(swallows$nesting_aid)

swallows$nesting_aid <- factor(swallows$nesting_aid, levels = 
                                  c("none", "support", "artif_nest", "both"),
                                  ordered = T)
levels(swallows$nesting_aid)

contrasts(swallows$nesting_aid)
```

### 2020-07-20 Bayesian Data Analysis/05_Likelihood.rmd
Bayesian data analysis, likelihood
Bayesian data analysis, likelihood

--- 2020-07-13 Bayesian Data Analysis/04_Normal-Linear-Models.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Normal Linear Models}
-   \rfoot{\color{headergrey}Chapter 4}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Likelihood}
+   \rfoot{\color{headergrey}Chapter 5}
    \lfoot{\color{headergrey}}
@@
 
-# Normal Linear Models
+# Likelihood
 
-## Fitting a Linear Regression in R
+## Theory
+
+Given 3 observations .8, 1.2, 1.1 and a model $y_i \sim Norm(1, .2)$
 
 ```{r}
-n <- 50 # sample size
-sigma <- 5 # standard deviation of the residuals
-b0 <- 2 # intercept
-b1 <- 0.7 # slope
-x <- runif(n, 10, 30) # sample values of the covariate
-yhat <- b0 + b1*x
-y <- rnorm(n, yhat, sd=sigma)
+obs <- c(0.8, 1.2, 1.1) # observations
 
-# plot the data
-plot(x, y, pch=16, las=1, cex.lab=1.2)
-abline(lm(y~x), lwd=2, col="blue") # insert regression line
-# add residuals
-segments(x, fitted(lm(y~x)), x, y, lwd=2, col="orange", lty=3)
-```
+# probability of seeing  these  in n(1, .2)
+p1 <- prod(dnorm(obs, mean = 1, sd = 0.2))
+p1
 
-Fit the model
+# probability of seeing  these  in n(1.2, .4)
+p2 <- prod(dnorm(obs, mean = 1.2, sd = 0.4))
+p2
 
-```{r}
-mod <- lm(y~x)
-mod
+p1/p2 # 4x liklier to see under model 1 than model 2
 
-summary(mod)
+# graphically
+xs <- seq(from=0, to=3, by = .01)
+ys <- dnorm(xs, mean = 1, sd = .2)
 
-summary(mod)$sigma
-```
+dat <- data.table(x = xs, y = ys)
 
-### Drawing Conclusions
+obs_marks <- data.table(x = obs, y = dnorm(obs, mean = 1, sd = .2))
 
-Alternative Display
-
-```{r}
-nsim <- 1000
-bsim <- sim(mod, n.sim = nsim)
-
-plot(x, y, pch=16, las=1, cex.lab=1.2)
-for(i in 1:nsim) abline(coef(bsim)[i, 1], coef(bsim)[i, 2],
-                        col = rgb(0, 0, 0, 0.05))
-
-newdat <- data.frame(x = seq(10, 30, by=0.1))
-newmodmat <- model.matrix(~x, data = newdat)
-fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))
-
-for(i in 1:nsim) fitmat[, i] <- newmodmat %*% coef(bsim)[i, ]
-plot(x, y, pch=16, las=1, cex.lab=1.2)
-abline(mod, lwd=2)
-
-lines(newdat$x, apply(fitmat, 1, quantile, prob = 0.025, lty=3))
-lines(newdat$x, apply(fitmat, 1, quantile, prob=0.975, lty=3))
-```
-
-Predicting Future Observations
-
-```{r}
-plot(x, y, pch=16, las=1, cex.lab=1.2)
-
-# prepare matrix for simulated new data
-newy <- matrix(ncol = nsim, nrow = nrow(newdat))
-# for each simulate4d fitted value, simulate  one new y-value
-for(i in 1:nsim) newy[, i] <- rnorm(nrow(newdat), mean = fitmat[, i],
-                                    sd = bsim@sigma[i])
-
-lines(newdat$x, apply(newy, 1, quantile, prob=0.025), lty=3)
-lines(newdat$x, apply(newy, 1, quantile, prob=0.975), lty=3)
-```
-
-```{r}
-sum(newy[newdat$x == 25,] > 20) / nsim
-```
-
-### Frequentist Results
-
-```{r}
-summary(mod)
-```
-
-## Regression Variants: ANOVA, ANCOVA, and MLR
-
-```{r}
-# data simulation
-mu <- 12
-sigma <- 2
-b1 <- 3
-b2 <- -5
-n <- 90
-group <- factor(rep(c(1, 2, 3), each = 30))
-
-# simulate the y variable
-simresid <- rnorm(n, mean=0, sd=sigma)
-y <- mu + as.numeric(group == 2) * b1 + as.numeric(group == "3") * b2 +
-   simresid
+ggplot(dat) +
+   geom_line(aes(x,y), col = "darkgreen", lty=2) +
+   geom_hline(yintercept = 0) +
+   geom_segment(data = obs_marks, aes(x, y, xend=x, yend=0), col="darkred")
 
@@
 
+## Quick Look at the Normal Distribution
+
 ```{r}
-group <- factor(group) # define group as a factor
-mod <- lm(y ~ group) # fit the model
+x <- seq(0, 2, length = 100)
+dx <- dnorm(x, mean = 1, sd = 0.2)
+plot(x, dx, type = "l", xlab = "Weight (kg)", ylab = "Density", lwd = 2, las = 1)
 
-mod
+rnorm(5, 1, 0.2) # draws 5 random numbers from Norm(1, 0.2)
 
-summary(mod)$sigma
+pnorm(q = .8, 1, 0.2)
 ```
 
+$L(\mu, \sigma | y) = \prod^n_{i=1}p(y_i, |\mu, \sigma)$
+
 ```{r}
-bsim <- sim(mod, n.sim=1000)
-m.g1 <- coef(bsim)[, 1]
-m.g2 <- coef(bsim)[, 1] + coef(bsim)[, 2]
-m.g3 <- coef(bsim)[, 1] + coef(bsim)[, 3]
+y <- c(0.8, 1.2, 1.1)
+lf <- function(mu, sigma) prod(dnorm(y, mu, sigma))
 
-m.dat <- data.table(group1 = m.g1, group2 = m.g2, group3 = m.g3)
-m.dat.long <- melt(m.dat)
-m.dat.long[, Mean := mean(value), by = variable]
-
-p1 <- ggplot(m.dat.long) +
-   geom_density(aes(value, fill = variable), alpha = .6)
-
-p2 <- ggplot(m.dat.long) +
-   geom_point(aes(x = variable, y = Mean, color = variable), alpha = .75, size = 6)
-
-grid.arrange(p1, p2, ncol=2)
-
-d.g1.2 <- m.g1 - m.g2
-mean(d.g1.2)
-quantile(d.g1.2, prob=c(0.025, 0.975))
-
-sum(m.g2 > m.g1) / nsim
+lf(1, .2)
 ```
 
-## Frequentist Results
+## The Maximum Likelihood Method
 
 ```{r}
-data("periparusater")
-dat <- periparusater
+mu <- seq(0.6, 1.4, length=100)
+sigma <- seq(0.05, 0.6, length=100)
 
-mod <- lm(wing ~ sex + age, data=dat)
-plot(mod)
+lik <- matrix(nrow=length(mu), ncol = length(sigma))
 
-newdat <- expand.grid(sex = factor(c(1, 2)), age = factor(c(3, 4)))
-newdat$fit <- model.matrix(~sex + age, dat=newdat) %*% coef(mod)
+for(i in 1:length(mu)) {
+   for(j in 1:length(sigma)) {
+      lik[i, j] <- lf(mu[i], sigma[j])
+   }
+}
 
-nsim <- 2000
-bsim <- sim(mod, n.sim = nsim)
+contour(mu, sigma, lik, nlevels=20, xlab = expression(mu),
+        ylab =expression(sigma), las=1, cex.lab = 1.4)
 
-fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))
-Xmat <- model.matrix(formula(mod)[c(1,3)], dat = newdat)
-for(i in 1:nsim) fitmat[, i] <- Xmat %*% bsim@coef[i, ]
+neglf <- function(x) -prod(dnorm(y, x[1], x[2]))
+MLest <- optim(c(1, 0.2), neglf)
+MLest$par
 
-alpha <- .05
-intervals <- c(lower = alpha/2, upper = 1 - alpha/2)
-
-ci <- t(apply(fitmat, 1, quantile, prob = intervals))
-
-cbind(newdat, ci)
-
-mod2 <- lm(wing ~ sex * age, data = dat)
+# optimal values
+points(MLest$par[1], MLest$par[2], col="red", pch=4)
 
@@
 
+Likelihood ratio = p1/p2
+
+## The Log Pointwise Predictive Density
+
+$llpd = \sum^n_{i=1}log\int{p(y_i|\theta)p(\theta|y)d\theta}$
+
 ```{r}
-bsim2 <- sim(mod2, n.sim = nsim)
-quantile(bsim2@coef[, 4], prob = c(0.025, 0.5, 0.975))
 
-summary(mod2)$sigma
+mod <- lm(y~1) # fit model by LS method
+nsim <- 2000
+bsim <- sim(mod, n.sim=nsim) # simulate from posterior dist. of parameters
 
-mean(abs(bsim2@coef[, 4]) > 0.3)
+pyi <- matrix(nrow=length(y), ncol=nsim)
 
-coef(mod2)
+for(i in 1:nsim) pyi[, i] <- dnorm(y, mean=bsim@coef[i, 1],
+                              sd=bsim@sigma[i])
 
-quantile(bsim2@coef[, 2], prob = c(0.025, 0.5, 0.975))
+mpyi <- apply(pyi, 1, mean)
 
-sum(bsim@coef[, 2] < 0)/nsim # for juveniles (reference level)
-sum(bsim@coef[, 2] + bsim2@coef[, 4] < 0) / nsim # adults
+sum(log(mpyi))
 
@@
 
-## Analysis of Covariance
-
-```{r}
-data("ellenberg")
-
-index <- is.element(ellenberg$Species, c("Ap", "Dg"))
-dat <- ellenberg[index, ] # select two species
-dat <- droplevels(dat) # drop unused factor levels
-str(dat) # definitions
-
-mod <- lm(log(Yi.g) ~ Species + Water, data = dat)
-
-head(model.matrix(mod)) # print the first 6 rows of the matrix
-
-summary(mod)
-
-mod2 <- lm(log(Yi.g) ~ Species*Water, data = dat)
-summary(mod2)
-```
-
-```{r}
-nsim <- 2000
-bsim <- sim(mod2, n.sim = nsim)
-coefs <- coef(bsim)
-
-xatcross <- crosspoint(coefs[, 1], coefs[, 3],
-                       coefs[, 1] + coefs[, 2], coefs[, 3] + coefs[, 4])[ , 1]
-
-xatcross[xatcross< (-5)] <- -5
-
-th <- hist(xatcross, breaksw = seq(-5.5, 140.5, by=5))
-plot(th$mids, cumsum(th$counts)/2000, type = "l", lwd=2, las=1,
-     ylim=c(0, 1), ylab="P(Dg > Ap|data)", xlab="Average diistance to ground water (cm)")
-```
-
-## Multiple Regression and Collinearity
-
-```{r}
-data(mdat)
-mod <- lm(y ~ x1 + x2, data = mdat)
-summary(mod)
-```
-
-```{r}
-cor(mdat[, 2:6])
-
-ggpairs(mdat)
-
-own.graph <- function(x, y) {
-   points(x, y, pch=16, col=rgb(1, 0.5, 0.0, 0.8))
-   abline(lm(y~x))
-}
-pairs(mdat, panel = own.graph)
-```
-
-## Ordered Factors and Constants
-
-```{r}
-data("swallows")
-levels(swallows$nesting_aid)
-
-str(swallows)
-
-contrasts(swallows$nesting_aid)
-
-swallows$nesting_aid <- factor(swallows$nesting_aid, levels = 
-                                  c("none", "support", "artif_nest", "both"),
-                                  ordered = T)
-levels(swallows$nesting_aid)
-
-contrasts(swallows$nesting_aid)
-```
+The log posterior density can be used as a measure of model fit.

### 2020-07-22 Bayesian Data Analysis/06_Assessing-Model-Assumptions.rmd
bayesian data analysis, ch 6
bayesian data analysis, ch 6

--- 2020-07-20 Bayesian Data Analysis/05_Likelihood.rmd
@@
    \fancyhead[L]{\color{headergrey}Moretz, Brandon}
-   \fancyhead[C]{\Large\bfseries\color{headergrey}Likelihood}
+   \fancyhead[C]{\Large\bfseries\color{headergrey}Assessing Model Assumptions}
    \rfoot{\color{headergrey}Chapter 5}
@@
 library(coda, quietly = TRUE, warn.conflicts = FALSE)
+library(nlme, quietly = TRUE, warn.conflicts = FALSE)
+library(sp, quietly = TRUE, warn.conflicts = FALSE)
+library(gstat, quietly = TRUE, warn.conflicts = FALSE)
 
@@
 
-# Likelihood
+# Assessing Model Assumptions
 
-## Theory
-
-Given 3 observations .8, 1.2, 1.1 and a model $y_i \sim Norm(1, .2)$
+## Model Assumptions
 
 ```{r}
-obs <- c(0.8, 1.2, 1.1) # observations
 
-# probability of seeing  these  in n(1, .2)
-p1 <- prod(dnorm(obs, mean = 1, sd = 0.2))
-p1
+data("periparusater")
 
-# probability of seeing  these  in n(1.2, .4)
-p2 <- prod(dnorm(obs, mean = 1.2, sd = 0.4))
-p2
-
-p1/p2 # 4x liklier to see under model 1 than model 2
-
-# graphically
-xs <- seq(from=0, to=3, by = .01)
-ys <- dnorm(xs, mean = 1, sd = .2)
-
-dat <- data.table(x = xs, y = ys)
-
-obs_marks <- data.table(x = obs, y = dnorm(obs, mean = 1, sd = .2))
-
-ggplot(dat) +
-   geom_line(aes(x,y), col = "darkgreen", lty=2) +
-   geom_hline(yintercept = 0) +
-   geom_segment(data = obs_marks, aes(x, y, xend=x, yend=0), col="darkred")
+dat <- periparusater
+mod <- lm(wing ~ sex + age, data = dat)
+par(mfrow = c(2, 2))
+plot(mod)
 
@@
 
-## Quick Look at the Normal Distribution
+```{r}
+names(dat)
 
-```{r}
-x <- seq(0, 2, length = 100)
-dx <- dnorm(x, mean = 1, sd = 0.2)
-plot(x, dx, type = "l", xlab = "Weight (kg)", ylab = "Density", lwd = 2, las = 1)
-
-rnorm(5, 1, 0.2) # draws 5 random numbers from Norm(1, 0.2)
-
-pnorm(q = .8, 1, 0.2)
-```
-
-$L(\mu, \sigma | y) = \prod^n_{i=1}p(y_i, |\mu, \sigma)$
-
-```{r}
-y <- c(0.8, 1.2, 1.1)
-lf <- function(mu, sigma) prod(dnorm(y, mu, sigma))
-
-lf(1, .2)
-```
-
-## The Maximum Likelihood Method
-
-```{r}
-mu <- seq(0.6, 1.4, length=100)
-sigma <- seq(0.05, 0.6, length=100)
-
-lik <- matrix(nrow=length(mu), ncol = length(sigma))
-
-for(i in 1:length(mu)) {
-   for(j in 1:length(sigma)) {
-      lik[i, j] <- lf(mu[i], sigma[j])
-   }
-}
-
-contour(mu, sigma, lik, nlevels=20, xlab = expression(mu),
-        ylab =expression(sigma), las=1, cex.lab = 1.4)
-
-neglf <- function(x) -prod(dnorm(y, x[1], x[2]))
-MLest <- optim(c(1, 0.2), neglf)
-MLest$par
-
-# optimal values
-points(MLest$par[1], MLest$par[2], col="red", pch=4)
+plot(resid(mod) ~ country, data = dat)
+plot(resid(mod) ~ age, data = dat)
+plot(resid(mod) ~ sex, data = dat)
+plot(resid(mod) ~ weight, data = dat)
 
@@
 
-Likelihood ratio = p1/p2
+```{r}
+table(dat$country)
 
-## The Log Pointwise Predictive Density
+ggplot(dat, aes(y = country)) +
+   geom_histogram(aes(fill = ..count..), stat = "count")
 
-$llpd = \sum^n_{i=1}log\int{p(y_i|\theta)p(\theta|y)d\theta}$
+```
+
+## The QQ
+
+```{r}
+compareqqnorm(mod)
+```
+
+## Temporal Autocorrelation
+
+```{r}
+data("parusmajor")
+
+dat <- parusmajor
+str(dat)
+
+dat <- dat[order(dat$year, dat$julian), ] # order date acc, time
+t.poly.jul <- poly(dat$julian, degree=2)
+dat$julian.l <- t.poly.jul[, 1] # orthogonal linear trend
+dat$julian.q <- t.poly.jul[, 2] # orthogonal quadratic trend
+
+mod <- lm(count ~ year + julian.l + julian.q, data = dat )
+
+summary(mod)
+```
+
+```{r}
+par(mfrow = c(1, 2))
+acf(resid(mod))
+acf(resid(mod), type = "p")
+```
+
+### Autocorrelation
+
+```{r}
+mod <- gls(count ~ year + julian.l + julian.q, data = dat,
+           correlation = corAR1())
+
+summary(mod)
+```
+
+### Spatial Autocorrelation
+
+```{r}
+data("frogs")
+
+frogs$year.z <- scale(frogs$year)
+frogs$elevation.z <- scale(frogs$elevation)
+frogs$waterarea[frogs$waterarea == 0 ] <- 0.25
+frogs$waterarea.sqrt.l <- log(sqrt(frogs$waterarea))
+frogs$waterarea.sqrt.l.z <- scale(frogs$waterarea.sqrt.l)
+
+```
+
+## Negative Binomial Model
 
@@
 
-mod <- lm(y~1) # fit model by LS method
-nsim <- 2000
-bsim <- sim(mod, n.sim=nsim) # simulate from posterior dist. of parameters
+mod <- glm.nb(count2 ~ elevation.z + year.z + fish + vegetation + waterarea.sqrt.l.z + fish:vegetation, data = frogs)
 
-pyi <- matrix(nrow=length(y), ncol=nsim)
+par(mfrow=c(2, 2))
+plot(mod)
 
-for(i in 1:nsim) pyi[, i] <- dnorm(y, mean=bsim@coef[i, 1],
-                              sd=bsim@sigma[i])
-
-mpyi <- apply(pyi, 1, mean)
-
-sum(log(mpyi))
-
+spdata <- data.frame(resid = resid(mod), x = frogs$x, y = frogs$y)
+coordinates(spdata) <- c("x", "y")
+bubble(spdata, "resid", col = c("blue", "orange"), main = "Residuals",
+       xlab = "X-coordinates", ylab = "Y-coordinates")
 ```
 
-The log posterior density can be used as a measure of model fit.
+```{r}
+vario.mod <- variogram(resid(mod) ~ 1, spdata)
+plot(vario.mod)
+```
+
+```{r}
+vario.mod.6dir <- variogram(resid(mod) ~ 1, spdata, alpha = seq(0, 150, by = 30))
+plot(vario.mod.6dir)
+```
+
+## Hetroscedasticity
+
+```{r}
+data(ellenberg)
+dat <- ellenberg[complete.cases(ellenberg[c("Yi.g", "Water",
+                                            "Species")]), ]
+mod <- lm(log(Yi.g) ~ Water + Species + Water:Species, dat)
+par(mfrow=c(1, 2))
+plot(resid(mod) ~ Species, dat)
+scatter.smooth(dat$Water, sqrt(abs(resid(mod))), xlab = "Water")
+```

### 2020-07-24 Bayesian Data Analysis/08_Generalized-Linear-Models.rmd
bayesian data analysis, ch8
bayesian data analysis, ch8

---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Linear Mixed Effects Models}
   \rfoot{\color{headergrey}Chapter 7}
   \lfoot{\color{headergrey}}
   \fancyfoot[C]{\rmfamily\color{headergrey}Bayesian Data Analysis}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 4.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(cowplot, quietly = TRUE, warn.conflicts = FALSE)
library(GGally, quietly = TRUE, warn.conflicts = FALSE)

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)

library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(lattice, quietly = TRUE, warn.conflicts = FALSE)

library(blmeco, quietly = TRUE, warn.conflicts = FALSE)
library(tidybayes, quietly = TRUE, warn.conflicts = FALSE)
library(arm, quietly = TRUE, warn.conflicts = FALSE)
library(coda, quietly = TRUE, warn.conflicts = FALSE)
library(nlme, quietly = TRUE, warn.conflicts = FALSE)
library(sp, quietly = TRUE, warn.conflicts = FALSE)
library(gstat, quietly = TRUE, warn.conflicts = FALSE)
library(lme4, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

# Linear Mixed Effects Models

## Why Mixed Effects Models?

### Pooling

Complete:

$\hat{y}_i = \beta_o$
$y_i \sim Norm(\hat{y}_i, \sigma^2)$

Partial:

$\hat{y_i} = \beta_0 + b_{g_{i}}$
$y_i \sim Norm(\hat{y_i}, \sigma^2)$
$b_g \sim Norm(0, \sigma^2_b)$

No pooling:

$\hat{y} = \beta_{g_i}$
$y_i \sim Norm(\hat{y_i}, \sigma^2)$

## Fitting Linear Mixed Models

```{r}
data("cortbowl")

dat <- cortbowl
mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Ring), 
            data = dat, REML = T)

summary(mod)

round(fixef(mod), 3)

ranef(mod)
```

## Restricted Maximum Liklihood Estimation

```{r}
mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Ring), 
            data = dat, REML = F) # use ML

```

## Assessing Model Assumptions

```{r}
par(mfrow=c(2,2))
scatter.smooth(fitted(mod), resid(mod)); abline(h=0, lty=2)
title("Tukey-Anscombe Plot")

qqnorm(resid(mod), main = "norm QQ-plot, residuals") # qq of residuals
qqline(resid(mod))
scatter.smooth(fitted(mod), sqrt(abs(resid(mod))))

qqnorm(ranef(mod)$Ring[, 1], main = "normal QQ-plot, random effects")
qqline(ranef(mod)$Ring[, 1]) # qq random effects
```

## Drawing Conclusions

```{r}
nsim <- 2000
bsim <- sim(mod, n.sim = nsim)
str(bsim)

round(apply(bsim@fixef, 2, quantile, prob=c(0.025, 0.5, 0.975)), 3)

```

```{r}

imp_factor <- factor(c("C", "P"), levels = levels(dat$Implant))
days_factor <- factor(c(1, 2, 3))

newdat <- expand.grid(Implant = imp_factor, days = days_factor)

Xmat <- model.matrix(~Implant + days + Implant:days, data = newdat)
fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))
for(i in 1:nsim) fitmat[, i] <- Xmat %*% bsim@fixef[i, ] # fitted values
newdat$lower <- apply(fitmat, 1, quantile, prob = 0.025)
newdat$upper <- apply(fitmat, 1, quantile, prob = 0.975)
newdat$fit <- Xmat %*% fixef(mod)

ggplot(newdat, aes(days, fit, color = Implant)) +
   geom_point()

```

## Frequentist Results

```{r}

```

## Random Intercept and Random Slope

```{r}

data(wingbowl)
dat <- wingbowl
dat$Age.z <- scale(dat$Age)
mod <- lmer(Wing ~ Age.z + Implant + Age.z:Implant + (Age.z|Ring),
          data = dat, REML = F)
mod
```

```{r}

par(mfrow=c(1, 2))
qqnorm(ranef(mod)$Ring[, 1]) # Intercept
qqline(ranef(mod)$Ring[, 1])

qqnorm(ranef(mod)$Ring[, 2]) # Slope
qqline(ranef(mod)$Ring[, 2])

nsim <- 2000
bsim <- sim(mod, n.sim = nsim)
apply(bsim@fixef, 2, quantile, prob = c(0.025, 0.975))

quantile(bsim@fixef[, "Age.z:ImplantP"] / sd(dat$Age), prob = c(0.025, 0.975))
```

```{r}
newdat <- expand.grid(Age = seq(23, 45, length = 100),
                      Implant = levels(dat$Implant))

newdat$Age.z <- (newdat$Age - mean(dat$Age)) / sd(dat$Age)

head(newdat)
dim(newdat)

Xmat <- model.matrix(~Age.z + Implant + Age.z:Implant, data = newdat)
fitmat <- matrix(ncol = nsim, nrow = nrow(newdat))

dim(fitmat)
dim(bsim@fixef)

head(bsim@fixef)

for(i in 1:nsim) {
   fitmat[, i] <- Xmat %*% bsim@fixef[i, ]
}

dim(bsim@fixef)

newdat$lower <- apply(fitmat, 1, quantile, prob = 0.025)
newdat$upper <- apply(fitmat, 1, quantile, prob = 0.975)
```

```{r}
par(mfrow = c(1, 2), mar=c(5, 1, 1, 1), oma = c(0, 4, 0, 0))
plot(dat$Age.z, dat$Wing, pch=1, cex = 0.8, las = 1,
     col = c("orange", "blue")[as.numeric(dat$Implant)],
     xlab = "Age (days)", ylab = NA, xaxt = "n")
at.x_orig <- seq(25, 45, by = 5) # values on the x-axis, original scale
at.x <- (at.x_orig - mean(dat$Age)) / sd(dat$Age) # transformed scale
axis(1, at = at.x, labels = at.x_orig) # original values at transformed
mtext("Wing length (mm)", side = 2, outer = T, line = 2, cex = 1.2, adj=0.6)
abline(fixef(mod)[1], fixef(mod)[2], col = "orange", lwd=2) # for C
abline(fixef(mod)[1] + fixef(mod)[3], fixef(mod)[2] + fixef(mod)[4],
       col = "blue", lwd = 2)

for(i in 1:2) {
   index <- newdat$Implant == levels(newdat$Implant)[i]

   x <- c(newdat$Age.z[index], rev(newdat$Age.z[index]))
   y <- c(newdat$lower[index], rev(newdat$upper[index]))

   polygon(x, y,
           border = NA, col = c(rgb(1, 0.65, 0, 0.5), rgb(0, 0, 1, 0.5))[i])
}

plot(dat$Age.z, dat$Wing, pch=1, cex=0.8, las=1,
     col = c("orange", "blue")[as.numeric(dat$Implant)],
     xlab = "Age (days)", ylab = NA, yaxt = "n", xaxt = "n")
at.x_orig <- seq(25, 45, by = 5)
ax.x <- (at.x_orig - mean(dat$Age)) / sd(dat$Age)
axis(1, at = at.x, labels = at.x_orig)
indtreat <- tapply(dat$Implant, dat$Ring, function(x) as.character(x[1]))
for(i in 1:86) {
   if(indtreat[i] == "C") abline(fixef(mod)[1] + ranef(mod)$Ring[i, 1],
                                 fixef(mod)[2] + ranef(mod)$Ring[i, 2],
                                 col = "Orange") else
                                    abline(fixef(mod)[1] + fixef(mod)[3],
                                           ranef(mod)$Ring[i, 1], fixef(mod)[2] +
                                              fixef(mod)[4] + ranef(mod)$Ring[i, 2],
                                           col = "blue")
}

```

## Nested and Crossed Random Effects

```{r}
data(cortbowl)
dat <- cortbowl

mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Brood) + (1|Ring),
            data = dat, REML = F)

mod

```

```{r}
mod <- lmer(log(totCort) ~ Implant + days + Implant:days + (1|Brood/Ring),
            data = dat, REML = F)
mod
```

```{r}
data("ellenberg")

ellenberg$gradient <- paste(ellenberg$Year, ellenberg$Soil)
table(ellenberg$Species, ellenberg$gradient)
```

```{r}
ellenberg$Water.z <- as.numeric(scale(ellenberg$Water))
mod <- lmer(log(Yi.g) ~ Water.z + I(Water.z^2) +
               (Water.z + I(Water.z^2)|Species) + (1|gradient),
            data = ellenberg)

mod
```

```{r}

```
